### 2017-09-06 paper_template/main.tex
Add paper template

%
%  $Author: ienne $
%  $Date: 1995/09/15 15:20:59 $
%  $Revision: 1.4 $
%

% \documentclass[10pt,journal,cspaper,compsoc]{IEEEtran}   %%%tc version
\documentclass[10pt, conference]{IEEEtran}
%\documentclass[conference,compsoc]{IEEEtran}
%\documentclass[10pt, conference]{IEEEtran}
%\documentclass[times, 10pt,onecolumn]{article}
\usepackage{amsmath, amssymb, enumerate}

%%%%%%%%%%%%%%%% page control%%%%%%%%%%%%%%%%%
%\usepackage[margin=0.75in]{geometry}

%\linespread{0.991}  %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% this is really useful
%\usepackage{cite}
\usepackage{fancybox}
\usepackage{amsfonts}
%\usepackage{algorithm}
%\usepackage[noend]{algorithmic}
\usepackage[usenames]{color}
%\usepackage{colortbl}
%\usepackage[ figure, boxed, vlined]{algorithm2e}
%\usepackage[linesnumbered,vlined]{algorithm2e}
%\usepackage[lined,boxed]{algorithm2e}
\usepackage{listings}

\usepackage[linesnumbered,vlined]{algorithm2e}
\usepackage{graphicx}
\usepackage{times}
\usepackage{psfrag}
\usepackage{subfigure}
\usepackage{caption}
%\usepackage{subcaption}
\usepackage{multirow}
%\usepackage{setspace}
%\usepackage{listings}
\usepackage{epsfig}
%\usepackage{epstopdf}
%\usepackage[font=small,labelfont=bf]{caption}
\usepackage{url}

\usepackage{color}
\def\fixme#1{\typeout{FIXED in page \thepage : {#1}}
%\bgroup \color{red}{} \egroup}
\bgroup \color{red}{[FIXME: {#1}]} \egroup}

%\usepackage[pdftex]{hyperref}
\usepackage{rotating,tabularx}

\interfootnotelinepenalty=10000

%% Define a new 'leo' style for the package that will use a smaller font.
\makeatletter
\def\url@leostyle{%
  \@ifundefined{selectfont}{\def\UrlFont{\sf}}{\def\UrlFont{\small\ttfamily}}}
\makeatother

%\documentstyle[times,art10,twocolumn,latex8]{article}

%-------------------------------------------------------------------------
% take the % away on next line to produce the final camera-ready version
\pagestyle{plain}
%\thispagestyle{empty}
%\pagestyle{empty}

\newtheorem{theorem}{Theorem}
\newtheorem{lemma}[theorem]{Lemma}

%% remaining budget share, used in task stall section.
\newcommand{\bottomrule}{\hline}
\newcommand{\toprule}{\hline}
\newcommand{\midrule}{\hline}
%-------------------------------------------------------------------------
\begin{document}

\title{Title of Your Paper}
\author{Author One, Author Two\\
\{author1,author2\}@ku.edu\\
University of Kansas, USA\\ 
}

\maketitle
\thispagestyle{empty}
\begin{abstract}

Abstract goes here.

\end{abstract}

%-------------------------------------------------------------------------

\section{Introduction}
Introduction goes here.

\section{Background}
Cite a paper \cite{barroso2009datacenter}.

Cite multiple papers \cite{banga99resourcecontainers,barroso2009datacenter}

\section{Your System}
...
\section{Evaluation}
...
\section{Conclusion}
...
%-------------------------------------------------------------------------

\bibliographystyle{plain}
\bibliography{reference}
\end{document}

### 2017-09-06 paper_template/main.tex
Add section headers and preliminary introduction

@@
 
-\title{Title of Your Paper}
+\title{DeepPicar:​ ​A​ ​Low-cost​ ​Deep​ ​Neural​ ​Network-based​ ​Autonomous​ ​Car}
 \author{Author One, Author Two\\
@@
 \section{Introduction}
-Introduction goes here.
+Autonomous vehicles have been a topic of increasing interest in many disciplines in recent years.
+Due to the nature of autonomous vehicles, it is necessary that all real-time operations are successfully performed prior to their deadlines.
+This requires that the AV platform be capable of completing all necessary computations in a timely manner, while maintaining a high level of accuracy.
+Consequently, the cost of creating a platform capable of self-driving can be relatively high, especially in the current cost-sensitive climate of the automotive industry \cite{}.
+The overall price for an autonomous vehicle platform currently acts as a bottleneck, especially from an academic standpoint, where cost is an important factor.
+For many researchers and students, the charge associated with autonomous vehicle research is a significant barrier.
+These same parties may also be deterred from participating in autonomous vehicle research due to the potential for their developed platforms to break in the experimental process.
+The idea of building one relatively expensive platform may not be too daunting for some, but the notion of potentially making multiple identical platforms in case of accidents may be overwhelming.
+Toward this end, we explore the possibility of developing a low-cost system that still employs state-of-the-art AI technologies and is capable of executing real-time autonomous vehicle operations.
+
+In this paper, we present DeepPicar, or Picar for short, a low-cost autonomous car platform that can be utilized in research and education.
+We follow the same methodology found in other platforms, such as NVIDIA’s end-to-end vision based control \cite{}, but implement it on a smaller scale.
+The Picar is comprised of a Raspberry Pi 3 Model B, a camera and a low cost RC car.
+By creating and using the Picar platform, we seek to accomplish the following three goals.
+First, the significant reduction of the overall cost required would make autonomous driving research more accessible to interested individuals/parties.
+Furthermore, it would remove the concerns of creating replacement platforms as each part, or the entire system, could be replaced relatively inexpensively.
+Second, the utilization of a more cost-efficient platform would allow for a better understanding of the performance requirements necessary for efficiently operating autonomous vehicles.
+Finally, we strive to achieve the ability to compare the real-time performance of multiple embedded computing platforms.
+
+We display the efficacy of the Picar by training and testing the DeepTesla DNN \cite{} in a custom-made environment.
+We find that the Picar is capable of consistently accomplishing all real-time operations required for autonomous driving in a 50 millisecond frame.
+
+The remaining sections of the paper are as follows: Section II gives an overview of the platform, including the high-level system and the methods used for training and inference.
+Section III discusses the ways/methodologies in which training data was collected.
+Section IV outlines the network architecture.
+This is followed by the real-time DNN inferencing in Section V.
+Section VI reviews the online/offline training done with the platform, with an evaluation given in Section VII.
+Section VIII gives a discussion of related works, and the paper finishes with conclusions in Section IX.
 
@@
 
-\section{Your System}
+\section{Platform Overview}
+...
+\section{Data Collection}
+...
+\section{Network Architecture}
+...
+\section{Real-Time DNN Inferencing}
+...
+\section{Online/Offline Training}
 ...
@@
 ...
+\section{Related Work}
+...
 \section{Conclusion}
-...
 %-------------------------------------------------------------------------

### 2017-09-07 picar-mini-rc-adafruit.py
swb,swa switch (record, dnn)

@@ -76,21 +76,17 @@ throttle = 0
 rec_start_time = 0
 null_frame = np.zeros((160,120,3), np.uint8)
 
-def deg2rad(deg):
-    return deg * math.pi / 180.0
-def rad2deg(rad):
-    return 180.0 * rad / math.pi
-
 def rc_to_throttle_angle(str_rc, thr_rc):
     if str_rc < 800 or thr_rc < 800:
         return 0, 0
-    th = (thr_rc - 1400) * 256 / 450
-    an = (str_rc - 1400) * 180 / 450 # degree
+    th = float(thr_rc - 1350) * 256 / 450
+    an = float(str_rc - 1350) * 180 / 450 # degree
     return th, an
     
 def control_motor_differential(th, an):    
-    l_speed = th - an / 5
-    r_speed = th + an / 5
+    an_adj = an - 41.6
+    l_speed = th - an_adj / 5
+    r_speed = th + an_adj / 5
 
     if l_speed > 0:
         left_motor.run(Adafruit_MotorHAT.BACKWARD)
@@ -133,30 +129,34 @@ while True:
         
     # 1. machine input
     img = preprocess.preprocess(frame)
-    angle_rad = model.y.eval(feed_dict={model.x: [img], model.keep_prob: 1.0})[0][0]
-    angle_dnn = rad2deg(angle_rad)
+    angle_dnn = model.y.eval(feed_dict={model.x: [img], model.keep_prob: 1.0})[0][0]
 
     # 2. human input (RC)
     line = ser.readline()    
-    rc1, rc2 = string.split(line[:-1])
-    rc1 = int(rc1) # str
+    rc1, rc2, rc3, rc4 = string.split(line[:-1])
+    rc1 = int(rc1) # steering
     rc2 = int(rc2) # throttle
+    rc3 = int(rc3) # swa: >1000 = dnn, <1000 = human
+    rc4 = int(rc4) # swb: >1000 = start record, <1000 = stop record
+
+    # print "DBG:", rc1, rc2, rc3, rc4
+    
     throttle, angle_rc = rc_to_throttle_angle(rc1, rc2)
     
     # 3. make a decision
-    if use_dnn:
+    if rc3 > 1000:
+        print "DNN:",
         angle = angle_dnn
     else:
+        print "MAN:",
         angle = angle_rc
         
     # 4. control actuators
     l_speed, r_speed = control_motor_differential(throttle, angle)
 
+    print ("throttle: %.1f angle: %.1f" % (throttle, angle))
[…]

### 2017-09-07 params.py
differential model

@@ -7,12 +7,12 @@ from collections import OrderedDict
 batch_size = 255
 # save_dir = os.path.abspath('models')
 save_dir = os.path.abspath('../picar/models')
-training_steps = 20000
+training_steps = 4000
 img_height = 66
 img_width = 200
 img_channels = 3
 write_summary = True
-use_category_normal = True # if ture, center/curve images are equally selected.
+use_category_normal = False # if ture, center/curve images are equally selected.
 
 # change this to the directory that contains the source videos
 data_dir = os.path.abspath('../picar/epochs')
@@ -30,8 +30,8 @@ if not os.path.isdir(out_dir):
 epochs = OrderedDict()
 # 8/23 picar-v2.0, 50ms fix
 
-epochs['train'] = range(400,420)
-epochs['val'] = range(420,425)
+epochs['train'] = [501, 502, 503, 504] 
+epochs['val'] = [500, 505]
 
 #epochs['train'] += [414, 415, 416, 417, 419, 420, 421, 422]
 #epochs['val'] += [418, 423]

### 2017-09-19 test-model.py
multi-model and multi-threading experiments
1. multi-threading experiment:

  method:

    1) change #of cores to use
       modify "NCPU" variable in test-model.py

    2) run
       $ python test-model.py

  files:

    logs/log-1c.txt   1 core result
    logs/log-2c.txt   2 core result
    logs/log-3c.txt   3 core result
    logs/log-4c.txt   4 core result

2. multi-model experiment

  method:

    1) set NCPU=1 in test-model.py

    2) run
      > 4 models:
      $ for c in 0 1 2 3; do python test-model.py & done

      > 2 models:
      $ for c in 0 1; do python test-model.py & done

  fi

@@ -13,7 +13,7 @@ import preprocess
 # import visualize
 import time
 import math
-
+import numpy as np
 import local_common as cm
 
 def deg2rad(deg):
@@ -21,7 +21,12 @@ def deg2rad(deg):
 def rad2deg(rad):
         return 180.0 * rad / math.pi
 
-sess = tf.InteractiveSession()
+NCPU=1
+config = tf.ConfigProto(intra_op_parallelism_threads=NCPU, inter_op_parallelism_threads=NCPU, \
+                        allow_soft_placement=True, device_count = {'CPU': 1})
+# sess = tf.Session(config=config)
+
+sess = tf.InteractiveSession(config=config)
 saver = tf.train.Saver()
 model_name = 'model.ckpt'
 model_path = cm.jn(params.save_dir, model_name)
@@ -29,7 +34,9 @@ saver.restore(sess, model_path)
 
 epoch_ids = sorted(list(set(itertools.chain(*params.epochs.values()))))
 
-# epoch_ids = [6] # DBG - heechul
+epoch_ids = [6] # DBG - heechul
+
+tot_time_list = []
 
 for epoch_id in epoch_ids:
     print '---------- processing video for epoch {} ----------'.format(epoch_id)
@@ -45,6 +52,7 @@ for epoch_id in epoch_ids:
     print 'performing inference...'
     time_start = time.time()
     for frame_id in xrange(frame_count):
+        cam_start = time.time()
         ret, img = cap.read()
         assert ret
 
@@ -56,12 +64,15 @@ for epoch_id in epoch_ids:
         deg = rad2deg(rad)
         pred_end   = time.time()
 
-        prep_time = pred_start - prep_start
-        pred_time = pred_end - pred_start
+        cam_time  = (prep_start - cam_start)*1000
+        prep_time = (pred_start - prep_start)*1000
+        pred_time = (pred_end - pred_start)*1000
+        tot_time  = (pred_end - cam_start)*1000
 
-        print 'pred: {} deg. took {} ms'.format(deg, pred_time * 1000)
+        print 'pred: {:0.2f} deg. took: {:0.2f} ms | cam={:0.2f} prep={:0.2f} pred={:0.2f}'.format(deg, tot_time, cam_time, prep_time, pred_time)
         # print 'pred: {} deg (rad={})'.format(deg, rad)
-
+        if frame_id > 0:
+	    tot_time_list.append(tot_time)
         machine_steering.append(deg)
 
     cap.release()
[…]

### 2017-09-19 test-model2.py
Add second model test file and multicore timing data
test-model2.py: change numCore to equal the number of cores being used.

Timing files:
Timing/time_1c.csv : times using one core
Timing/time_2c.csv : times using two cores
Timing/time_3c.csv : times using three cores
Timing/time_4c.csv : times using four cores
Timing/time_data.csv: statistical data of each multicore run

@@ -1,0 +1,172 @@
+#!/usr/bin/env python 
+from __future__ import division
+
+import model
+import cv2
+import subprocess as sp
+import itertools
+import params
+import sys
+import os
+import preprocess
+# import visualize
+import time
+import math
+import csv
+import numpy as np
+import resource
+import tensorflow as tf
+
+import local_common as cm
+
+def deg2rad(deg):
+        return deg * math.pi / 180.0
+def rad2deg(rad):
+        return 180.0 * rad / math.pi
+        
+numCore = 1
+sess = tf.InteractiveSession(config=tf.ConfigProto(intra_op_parallelism_threads=numCore, inter_op_parallelism_threads=numCore, allow_soft_placement=True))
+saver = tf.train.Saver()
+model_name = 'model.ckpt'
+model_path = cm.jn(params.save_dir, model_name)
+saver.restore(sess, model_path)
+
+epoch_ids = sorted(list(set(itertools.chain(*params.epochs.values()))))
+
+# epoch_ids = [6] # DBG - heechul
+
+cap_time_list = []
+process_time_list = []
+angle_time_list = []
+frame_time_list = []
+
+frameNum = 0
+
+cap = cv2.VideoCapture(0)
+
+for epoch_id in epoch_ids:
+    print '---------- processing video for epoch {} ----------'.format(epoch_id)
+    # vid_path = cm.jn(params.data_dir, 'epoch{:0>2}_front.mkv'.format(epoch_id))
+
+    vid_path = cm.jn(params.data_dir, 'out-video-{}.avi'.format(epoch_id))
+    assert os.path.isfile(vid_path)
+    frame_count = cm.frame_count(vid_path)
+    #cap = cv2.VideoCapture(vid_path)
+
+    machine_steering = []
+
+    print 'performing inference...'
+    time_start = time.time()
+    for frame_id in xrange(frame_count):
+        if frameNum < 1001:
+            cap_start = time.time()
+            ret, img = cap.read()
+            assert ret
+            
+            prep_start = time.time()
+
+            img = preprocess.preprocess(img)
+
+            pred_start = time.time()
+            #with sess.as_default():
+            rad = model.y.eval(feed_dict={model.x: [img], model.keep_prob: 1.0})[0][0]
+            deg = rad2deg(rad)
+            #with sess2.as_default():
+                #rad2 = model.y.eval(feed_dict={model.x: [img], model.keep_prob: 1.0})[0][0]
[…]

### 2017-09-20 paper/main.tex
intro revision

--- 2017-09-06 paper_template/main.tex
@@
 
+%% UPenn's f1/10 BOM: $3,628.37	
+%% http://f1tenth.org/
+%% http://selfdrivingcars.mit.edu/
+%% http://fast.scripts.mit.edu/racecar/
+%% https://github.com/mit-racecar
+
 \section{Introduction}
-Autonomous vehicles have been a topic of increasing interest in many disciplines in recent years.
-Due to the nature of autonomous vehicles, it is necessary that all real-time operations are successfully performed prior to their deadlines.
-This requires that the AV platform be capable of completing all necessary computations in a timely manner, while maintaining a high level of accuracy.
-Consequently, the cost of creating a platform capable of self-driving can be relatively high, especially in the current cost-sensitive climate of the automotive industry \cite{}.
-The overall price for an autonomous vehicle platform currently acts as a bottleneck, especially from an academic standpoint, where cost is an important factor.
-For many researchers and students, the charge associated with autonomous vehicle research is a significant barrier.
-These same parties may also be deterred from participating in autonomous vehicle research due to the potential for their developed platforms to break in the experimental process.
-The idea of building one relatively expensive platform may not be too daunting for some, but the notion of potentially making multiple identical platforms in case of accidents may be overwhelming.
-Toward this end, we explore the possibility of developing a low-cost system that still employs state-of-the-art AI technologies and is capable of executing real-time autonomous vehicle operations.
+Autonomous vehicles have been a topic of increasing interest in many
+disciplines in recent years. Due to the nature of autonomous vehicles,
+it is necessary that all real-time operations are successfully
+performed prior to their deadlines. This requires that the AV platform
+be capable of completing all necessary computations in a timely
+manner, while maintaining a high level of accuracy. Consequently, the
+cost of creating a platform capable of self-driving can be relatively
+high, especially in the current cost-sensitive climate of the
+automotive industry \cite{}. The overall price for an autonomous
+vehicle platform currently acts as a bottleneck, especially from an
+academic standpoint, where cost is an important factor. For many
+researchers and students, the charge associated with autonomous
+vehicle research is a significant barrier. These same parties may also
+be deterred from participating in autonomous vehicle research due to
+the potential for their developed platforms to break in the
+experimental process. The idea of building one relatively expensive
+platform may not be too daunting for some, but the notion of
+potentially making multiple identical platforms in case of accidents
+may be overwhelming. Toward this end, we explore the possibility of
+developing a low-cost system that still employs state-of-the-art AI
+technologies and is capable of executing real-time autonomous vehicle
+operations.
 
-In this paper, we present DeepPicar, or Picar for short, a low-cost autonomous car platform that can be utilized in research and education.
-We follow the same methodology found in other platforms, such as NVIDIA’s end-to-end vision based control \cite{}, but implement it on a smaller scale.
-The Picar is comprised of a Raspberry Pi 3 Model B, a camera and a low cost RC car.
-By creating and using the Picar platform, we seek to accomplish the following three goals.
-First, the significant reduction of the overall cost required would make autonomous driving research more accessible to interested individuals/parties.
-Furthermore, it would remove the concerns of creating replacement platforms as each part, or the entire system, could be replaced relatively inexpensively.
-Second, the utilization of a more cost-efficient platform would allow for a better understanding of the performance requirements necessary for efficiently operating autonomous vehicles.
-Finally, we strive to achieve the ability to compare the real-time performance of multiple embedded computing platforms.
+In this paper, we present DeepPicar, a low-cost autonomous car
+platform for research and education. From hardware perspective, 
+DeepPicar is comprised of a Raspberry Pi 3 Model B quad-core
+computer, a web camera and a RC car, all of which are very
+affordable. The DeepPicar, however, employs state-of-the-art AI
+techonologies, including a vision-based end-to-end control system that
+utilizes a deep convolutional nueral network (CNN), which has 9
+layers, about 27 million connections and 250 thousand parameters
+(weights). The network architecture is in fact identifical to the one
+used in Nvidia's DAVE-2 self-driving car~\cite{Bojarski2016}. Other
+than the obvious difference in scale (RC car vs. real car), from the 
+perspective of computing systems, the only differences between
+our system and Nvidia's DAVE-2 are that our system is implemented in
+TensorFlow~\cite{abadi2016tensorflow} and runs on a 
+Raspberry Pi 3 whereas Nvidia's DAVE-2 systems is implemented in Torch
+7~\cite{collobert2011torch7} and runs on a Drive PX self-driving car
+computer~\cite{drivepx}, which is much more powerful but also a lot more
+expensive.
 
-We display the efficacy of the Picar by training and testing the DeepTesla DNN \cite{} in a custom-made environment.
-We find that the Picar is capable of consistently accomplishing all real-time operations required for autonomous driving in a 50 millisecond frame.
+%% We chose to use a Raspberry Pi 3 not only because of its
+%% affordability and availability but also because it is representative
+%% of today's mainstream low-end multicore processors, found in
+%% smartphones, tables and other embedded devices.
+%% The DeepPicar is comprised of a Raspberry
+%% Pi 3 Model B, a camera and a low cost RC car.
 
-The remaining sections of the paper are as follows: Section II gives an overview of the platform, including the high-level system and the methods used for training and inference.
-Section III discusses the ways/methodologies in which training data was collected.
-Section IV outlines the network architecture.
-This is followed by the real-time DNN inferencing in Section V.
-Section VI reviews the online/offline training done with the platform, with an evaluation given in Section VII.
-Section VIII gives a discussion of related works, and the paper finishes with conclusions in Section IX.
+By creating and using the Picar platform, we seek to accomplish the
+following three goals. First, the significant reduction of the overall cost required
+would make autonomous driving research more accessible to interested
+individuals/parties. Furthermore, it would remove the concerns of
+creating replacement platforms as each part, or the entire system,
+could be replaced relatively inexpensively. Second, the utilization of
+a more cost-efficient platform would allow for a better understanding
+of the performance requirements necessary for efficiently operating
+autonomous vehicles. Finally, we strive to achieve the ability to
+compare the real-time performance of multiple embedded computing
+platforms.
+
+We display the efficacy of the Picar by training and testing the
+DeepTesla DNN \cite{} in a custom-made environment. We find that the
+Picar is capable of consistently accomplishing all real-time
+operations required for autonomous driving in a 50 millisecond frame.
+
+The remaining sections of the paper are as follows: Section II gives
+an overview of the platform, including the high-level system and the
+methods used for training and inference. Section III discusses the
+ways/methodologies in which training data was collected. Section IV
+outlines the network architecture. This is followed by the real-time
+DNN inferencing in Section V. Section VI reviews the online/offline
+training done with the platform, with an evaluation given in Section
+VII. Section VIII gives a discussion of related works, and the paper
+finishes with conclusions in Section IX.
 
@@
 \section{Related Work}
-...
+\cite{Bojarski2016}
+
 \section{Conclusion}

### 2017-09-25 multicore_timings.sh
multicore_timings test script and UP testing

@@ -1,0 +1,51 @@
+timestamp=$(date +%y-%m-%d_%H:%M:%S)
+
+mkdir -p logs/$timestamp
+
+echo single-core
+perf stat taskset -c 0 python test-model.py 1 > logs/$timestamp/single-core.txt &
+wait
+echo end single-core
+
+echo 2-parallel-single-core
+perf stat taskset -c 0 python test-model.py 1 > logs/$timestamp/2-single-core_cpu0.txt &
+perf stat taskset -c 1 python test-model.py 1 > logs/$timestamp/2-single-core_cpu1.txt &
+wait
+echo end 2-parallel-single-core
+
+echo 3-parallel-single-core
+perf stat taskset -c 0 python test-model.py 1 > logs/$timestamp/3-single-core_cpu0.txt &
+perf stat taskset -c 1 python test-model.py 1 > logs/$timestamp/3-single-core_cpu1.txt &
+perf stat taskset -c 2 python test-model.py 1 > logs/$timestamp/3-single-core_cpu2.txt &
+wait
+echo end 3-parallel-single-core
+
+echo 4-parallel-single-core
+perf stat taskset -c 0 python test-model.py 1 > logs/$timestamp/4-single-core_cpu0.txt &
+perf stat taskset -c 1 python test-model.py 1 > logs/$timestamp/4-single-core_cpu1.txt &
+perf stat taskset -c 2 python test-model.py 1 > logs/$timestamp/4-single-core_cpu2.txt &
+perf stat taskset -c 3 python test-model.py 1 > logs/$timestamp/4-single-core_cpu3.txt &
+wait
+echo end 4-parallel-single-core
+
+echo dual-core
+perf stat taskset -c 0,1 python test-model.py 2 > logs/$timestamp/dual-core.txt &
+wait
+echo end dual-core
+
+echo 2-parallel-dual-core
+perf stat taskset -c 0,1 python test-model.py 2 > logs/$timestamp/2-dual-core_cpu01.txt &
+perf stat taskset -c 2,3 python test-model.py 2 > logs/$timestamp/2-dual-core_cpu23.txt &
+wait
+echo end 2-parallel-dual-core
+
+echo tri-core
+perf stat taskset -c 0,1,2 python test-model.py 3 > logs/$timestamp/tri-core.txt &
+wait
+echo end tri-core
+
+echo quad-core
+perf stat taskset -c 0,1,2,3 python test-model.py 4 > logs/$timestamp/quad-core.txt &
+wait
+echo end quad-core
+

### 2017-09-25 paper/main.tex
intro revision

@@
 DeepPicar is comprised of a Raspberry Pi 3 Model B quad-core
-computer, a web camera and a RC car, all of which are very
-affordable. The DeepPicar, however, employs state-of-the-art AI
+computer, a web camera and a RC car, all of which are affordable
+components (less than \$200 in total).
+The DeepPicar, however, employs state-of-the-art AI
 techonologies, including a vision-based end-to-end control system that
-utilizes a deep convolutional nueral network (CNN), which has 9
-layers, about 27 million connections and 250 thousand parameters
-(weights). The network architecture is in fact identifical to the one
+utilizes a deep convolutional nueral network (CNN).
+The network receives an image frame from a single forward
+looking camera, as input and generates a predicted steering angle
+value as output at each control period in \emph{real-time}. 
+The network is deep as it has 9 layers, about 27 million connections
+and 250 thousand parameters (weights).
+The network architecture is in fact identifical to the one
 used in Nvidia's DAVE-2 self-driving car~\cite{Bojarski2016}. Other
 than the obvious difference in scale (RC car vs. real car), from the
-perspective of computing systems, the only differences between
-our system and Nvidia's DAVE-2 are that our system is implemented in
+computing perspective, the only differences between
+the two systems are that our system is implemented in
 TensorFlow~\cite{abadi2016tensorflow} and runs on a 
 Raspberry Pi 3 whereas Nvidia's DAVE-2 systems is implemented in Torch
-7~\cite{collobert2011torch7} and runs on a Drive PX self-driving car
-computer~\cite{drivepx}, which is much more powerful but also a lot more
-expensive.
+7~\cite{collobert2011torch7} and runs on a Drive PX computer (NVIDIA's
+automotive specialized computing system~\cite{drivepx}), which is more
+powerful but also more expensive.
+
+We follow the standard supervised learning methodology to train the
+system with a goal of following lanes on the ground, mimicking real
+roads. We first collected data for taining and validation by manually 
+controlling the RC car and recording the vision (from the webcam
+mounted on the RC-car) and the human control inputs. We then train the
+network offline using the collected data on a desktop computer, which
+equips a Nvidia GTX 1060 GPU. Finally, the trained network is copied
+back to the Raspberry Pi, which is then used to perform interference
+operations---locally on the Pi---in the car's main control loop in
+real-time. For real-time control, each \emph{inference} operation must
+be completed within the desired control period. (e.g., 50ms period for
+20Hz control frequency.)
+
+The primary {\bf contributions} of this paper are as follows:
+\begin{itemize}
+  \item We present the design, implementation, and case-study of a low-cost
+    autonomous vehicle that utilizes the state-of-the-art aritificial 
+    intelligence techniques.
+  \item We systematically analyze real-time characteristics of
+    multiple embedded computing platform, including the Raspberry Pi 3
+    Model B, in the contex of a vision-based autnomous driving.
+\end{itemize}
 
@@
 
-By creating and using the Picar platform, we seek to accomplish the
-following three goals. First, the significant reduction of the overall cost required
-would make autonomous driving research more accessible to interested
-individuals/parties. Furthermore, it would remove the concerns of
-creating replacement platforms as each part, or the entire system,
-could be replaced relatively inexpensively. Second, the utilization of
-a more cost-efficient platform would allow for a better understanding
-of the performance requirements necessary for efficiently operating
-autonomous vehicles. Finally, we strive to achieve the ability to
-compare the real-time performance of multiple embedded computing
-platforms.
+%% By creating and using the Picar platform, we seek to accomplish the 
+%% following three goals. First, the significant reduction of the overall
+%% cost required 
+%% would make autonomous driving research more accessible to interested
+%% individuals/parties. Furthermore, it would remove the concerns of
+%% creating replacement platforms as each part, or the entire system,
+%% could be replaced relatively inexpensively. Second, the utilization of
+%% a more cost-efficient platform would allow for a better understanding
+%% of the performance requirements necessary for efficiently operating
+%% autonomous vehicles. Finally, we strive to achieve the ability to
+%% compare the real-time performance of multiple embedded computing
+%% platforms.
 
-We display the efficacy of the Picar by training and testing the
-DeepTesla DNN \cite{} in a custom-made environment. We find that the
-Picar is capable of consistently accomplishing all real-time
-operations required for autonomous driving in a 50 millisecond frame.
+%% We display the efficacy of the Picar by training and testing the
+%% DeepTesla DNN \cite{} in a custom-made environment. We find that the
+%% Picar is capable of consistently accomplishing all real-time
+%% operations required for autonomous driving in a 50 millisecond frame.
 

### 2017-09-26 test-model.py
test script

@@ -21,7 +21,7 @@ def deg2rad(deg):
 def rad2deg(rad):
         return 180.0 * rad / math.pi
 
-NCPU=1
+NCPU=int(sys.argv[1])
 config = tf.ConfigProto(intra_op_parallelism_threads=NCPU, inter_op_parallelism_threads=NCPU, \
                         allow_soft_placement=True, device_count = {'CPU': 1})
 # sess = tf.Session(config=config)

### 2017-09-27 paper/background.tex
restructure

\section{Background} \label{sec:background}

(1) AI and Robotics
- explosion of AI
- in particular, application of DNN in perception and control of robotics systems.
- end-to-end control is a promising technique.
  levine's publications?
- examples: nvidia's DAVE-II prototype, forest navigating drone
challenge problem: computing at low cost?

%% UPenn's f1/10 BOM: $3,628.37	
%% http://f1tenth.org/
%% http://selfdrivingcars.mit.edu/
%% http://fast.scripts.mit.edu/racecar/
%% https://github.com/mit-racecar

(2) Multicore SBCs
- computing has been a obstacle.
- performance, but also constraints size, weight, and power as well as
cost senstive nature of industries including automtive.
- many new embeded computing platforms emerged: affordable and
powerful. raspberry pi, nvidia's embedded platforms tout their
superiority in GPU based acceleration of the AI tasks.

Our primary objective of this study are
- to understand the necessary computing performance for applying AI
technology based robotics systems, and 
- what kind of computing architecture and runtime supports
are most appropriate for such workload.

Toward to achieve the two goals, we implement a low-cost autonomous
car platform as a case study, as we will explain in the following
section.

### 2017-09-27 paper/intro.tex
restructure

--- 2017-09-25 paper/main.tex
@@
-%
-%  $Author: ienne $
-%  $Date: 1995/09/15 15:20:59 $
-%  $Revision: 1.4 $
-%
+\section{Introduction} \label{sec:intro}
 
-% \documentclass[10pt,journal,cspaper,compsoc]{IEEEtran}   %%%tc version
-\documentclass[10pt, conference]{IEEEtran}
-%\documentclass[conference,compsoc]{IEEEtran}
-%\documentclass[10pt, conference]{IEEEtran}
-%\documentclass[times, 10pt,onecolumn]{article}
-\usepackage{amsmath, amssymb, enumerate}
-
-%%%%%%%%%%%%%%%% page control%%%%%%%%%%%%%%%%%
-%\usepackage[margin=0.75in]{geometry}
-
-%\linespread{0.991}  %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% this is really useful
-%\usepackage{cite}
-\usepackage{fancybox}
-\usepackage{amsfonts}
-%\usepackage{algorithm}
-%\usepackage[noend]{algorithmic}
-\usepackage[usenames]{color}
-%\usepackage{colortbl}
-%\usepackage[ figure, boxed, vlined]{algorithm2e}
-%\usepackage[linesnumbered,vlined]{algorithm2e}
-%\usepackage[lined,boxed]{algorithm2e}
-\usepackage{listings}
-
-\usepackage[linesnumbered,vlined]{algorithm2e}
-\usepackage{graphicx}
-\usepackage{times}
-\usepackage{psfrag}
-\usepackage{subfigure}
-\usepackage{caption}
-%\usepackage{subcaption}
-\usepackage{multirow}
-%\usepackage{setspace}
-%\usepackage{listings}
-\usepackage{epsfig}
-%\usepackage{epstopdf}
-%\usepackage[font=small,labelfont=bf]{caption}
-\usepackage{url}
-
-\usepackage{color}
-\def\fixme#1{\typeout{FIXED in page \thepage : {#1}}
-%\bgroup \color{red}{} \egroup}
-\bgroup \color{red}{[FIXME: {#1}]} \egroup}
-
-%\usepackage[pdftex]{hyperref}
-\usepackage{rotating,tabularx}
-
-\interfootnotelinepenalty=10000
-
-%% Define a new 'leo' style for the package that will use a smaller font.
-\makeatletter
-\def\url@leostyle{%
-  \@ifundefined{selectfont}{\def\UrlFont{\sf}}{\def\UrlFont{\small\ttfamily}}}
-\makeatother
-
-%\documentstyle[times,art10,twocolumn,latex8]{article}
-
-%-------------------------------------------------------------------------
-% take the % away on next line to produce the final camera-ready version
-\pagestyle{plain}
-%\thispagestyle{empty}
-%\pagestyle{empty}
-
-\newtheorem{theorem}{Theorem}
-\newtheorem{lemma}[theorem]{Lemma}
-
-%% remaining budget share, used in task stall section.
-\newcommand{\bottomrule}{\hline}
-\newcommand{\toprule}{\hline}
-\newcommand{\midrule}{\hline}
-%-------------------------------------------------------------------------
-\begin{document}
-
-\title{DeepPicar:​ ​A​ ​Low-cost​ ​Deep​ ​Neural​ ​Network-based​ ​Autonomous​ ​Car}
-\author{Author One, Author Two\\
-\{author1,author2\}@ku.edu\\
-University of Kansas, USA\\ 
-}
-
-\maketitle
-\thispagestyle{empty}
-\begin{abstract}
-
-Abstract goes here.
-
-\end{abstract}
-
-%-------------------------------------------------------------------------
-
-%% UPenn's f1/10 BOM: $3,628.37	
-%% http://f1tenth.org/
-%% http://selfdrivingcars.mit.edu/
-%% http://fast.scripts.mit.edu/racecar/
-%% https://github.com/mit-racecar
-
-\section{Introduction}
 Autonomous vehicles have been a topic of increasing interest in many
@@
 
-The primary {\bf contributions} of this paper are as follows:
+The {\bf contributions} of this paper are as follows:
 \begin{itemize}
-  \item We present the design, implementation, and case-study of a low-cost
-    autonomous vehicle that utilizes the state-of-the-art aritificial 
-    intelligence techniques.
-  \item We systematically analyze real-time characteristics of
-    multiple embedded computing platform, including the Raspberry Pi 3
-    Model B, in the contex of a vision-based autnomous driving.
+  \item We present the design and implementation of a
+    low-cost autonomous vehicle testbed DeepPicar, which utilizes the
+    state-of-the-art aritificial intelligence techniques.
+  \item We provide an analysis and case-study of real-time issues in the
+    DeepPicar.
+  \item We systematically compare real-time computing capabilities of
+    multiple embedded computing platforms in the contex of a
+    vision-based autnomous driving.
 \end{itemize}
@@
 finishes with conclusions in Section IX.
-
-\section{Background}
-Cite a paper \cite{barroso2009datacenter}.
-
-Cite multiple papers \cite{banga99resourcecontainers,barroso2009datacenter}
-
-\section{Platform Overview}
-...
-\section{Data Collection}
-...
-\section{Network Architecture}
-...
-\section{Real-Time DNN Inferencing}
-...
-\section{Online/Offline Training}
-...
-\section{Evaluation}
-...
-\section{Related Work}
-\cite{Bojarski2016}
-
-\section{Conclusion}
-%-------------------------------------------------------------------------
-
-\bibliographystyle{plain}
-\bibliography{reference}
-\end{document}

### 2017-09-27 paper/main.tex
restructure

--- 2017-09-06 paper_template/main.tex
@@
 
-\section{Introduction}
-Autonomous vehicles have been a topic of increasing interest in many disciplines in recent years.
-Due to the nature of autonomous vehicles, it is necessary that all real-time operations are successfully performed prior to their deadlines.
-This requires that the AV platform be capable of completing all necessary computations in a timely manner, while maintaining a high level of accuracy.
-Consequently, the cost of creating a platform capable of self-driving can be relatively high, especially in the current cost-sensitive climate of the automotive industry \cite{}.
-The overall price for an autonomous vehicle platform currently acts as a bottleneck, especially from an academic standpoint, where cost is an important factor.
-For many researchers and students, the charge associated with autonomous vehicle research is a significant barrier.
-These same parties may also be deterred from participating in autonomous vehicle research due to the potential for their developed platforms to break in the experimental process.
-The idea of building one relatively expensive platform may not be too daunting for some, but the notion of potentially making multiple identical platforms in case of accidents may be overwhelming.
-Toward this end, we explore the possibility of developing a low-cost system that still employs state-of-the-art AI technologies and is capable of executing real-time autonomous vehicle operations.
-
-In this paper, we present DeepPicar, or Picar for short, a low-cost autonomous car platform that can be utilized in research and education.
-We follow the same methodology found in other platforms, such as NVIDIA’s end-to-end vision based control \cite{}, but implement it on a smaller scale.
-The Picar is comprised of a Raspberry Pi 3 Model B, a camera and a low cost RC car.
-By creating and using the Picar platform, we seek to accomplish the following three goals.
-First, the significant reduction of the overall cost required would make autonomous driving research more accessible to interested individuals/parties.
-Furthermore, it would remove the concerns of creating replacement platforms as each part, or the entire system, could be replaced relatively inexpensively.
-Second, the utilization of a more cost-efficient platform would allow for a better understanding of the performance requirements necessary for efficiently operating autonomous vehicles.
-Finally, we strive to achieve the ability to compare the real-time performance of multiple embedded computing platforms.
-
-We display the efficacy of the Picar by training and testing the DeepTesla DNN \cite{} in a custom-made environment.
-We find that the Picar is capable of consistently accomplishing all real-time operations required for autonomous driving in a 50 millisecond frame.
-
-The remaining sections of the paper are as follows: Section II gives an overview of the platform, including the high-level system and the methods used for training and inference.
-Section III discusses the ways/methodologies in which training data was collected.
-Section IV outlines the network architecture.
-This is followed by the real-time DNN inferencing in Section V.
-Section VI reviews the online/offline training done with the platform, with an evaluation given in Section VII.
-Section VIII gives a discussion of related works, and the paper finishes with conclusions in Section IX.
-
-\section{Background}
-Cite a paper \cite{barroso2009datacenter}.
-
-Cite multiple papers \cite{banga99resourcecontainers,barroso2009datacenter}
+\input{intro}
+\input{background}
 

### 2017-09-27 paper/main.tex

@@
 \section{Evaluation}
-...
+In evaluating the real-time efficacy of the Raspberry Pi 3, the same methodology is utilized in all 
+experiments conducted. The performance of the Pi is measured over a set of 1001 video frames that are 
+each individually fed to the model. The processing time for the first frame is ignored as, due to 
+cache warmup, it is uncharacteristically high and doesn't accurately represent the Pi's capabilities. A 
+deadline of 50 ms, or 20 Hz, is used as a baseline to assess the Pi's ability to complete all 
+necessary real-time operations in a timely manner. Also, please note that frame processing times 
+would be approximately the same if the input stream was a 
+camera instead of a video.
+
+\subsection{Real-Time Operations}
+In our platform, three main real-time tasks are performed during autonomous operation. 
+In order, those operations are: (1) capturing and reading the input frame from the designated camera 
+or video stream, (2) preprocessing the acquired frame so that it is compatible with the DNN, and (3) 
+feeding the frame to and getting the angle prediction from the model. However, the time it takes to 
+complete each of the operations will differ, with at least one of them being the dominating step in 
+processing each frame.
+
+\begin{figure}[h]
+  \centering
+  \includegraphics[width=.5\textwidth]{Operation_Time}
+  \caption{ Real-time operation completion time in processing frames.}
+\end{figure}
+
+In order to determine which operation(s) take the longest to execute, we measured the time it 
+took for each step to be completed. For this experiment, all four of the Pi's cpu cores were utilized, 
+and only one model was run. As is shown in Fig. 1, the angle prediction operation (3) consumes the 
+majority of the processing for each frame. Furthermore, the time it takes for the operation to 
+complete is volatile, and can range anywhere between 30 ms and 50 ms for any particular frame. On the 
+other hand, both the frame capture (1) and preprocessing (2) operations take substantially less time 
+and are relatively more consistent in their times, at 2 ms and 3 ms, respectively.
+
+\subsection{Multicore Performance}
+It may not always be the case that all four cores of the Raspberry Pi 3's Cortex A-53 CPU can be used 
+solely for the purpose of operating an autonomous vehicle. Thus, we test how the number of cores 
+utilized for real-time operations affects the Pi's overall ability to function as an autonomous 
+vehicle platform.
+
+\begin{table*}
+  \begin{tabular} {| l | l | l | l | l | l | l | l | l | l |}
+    \hline
+    \textbf{num cores} & \textbf{mean} & \textbf{max} & \textbf{99.999pct} & \textbf{99.99pct} & 
+      \textbf{99.9pct} & \textbf{99pct} & \textbf{min} & \textbf{median} & \textbf{stdev} \\ \hline 
+    1 & 61.95813584 & 65.99712 & 65.99277693 & 65.9536558 & 65.562445 & 63.30799 & 59.304 & 61.975479 & 
+      0.506466 \\ \hline
+    2 & 50.49184 895 & 71.54608 & 71.5404019 & 71.4893194 & 70.978494 & 70.02651 & 46.93723 & 50.011516 & 
+      3.162015 \\ \hline
+    3 & 48.11392045 & 72.21985 & 72.21789317 & 72.200294 & 72.024303 & 58.44643 & 41.15391 & 48.444033 & 
+      4.178325 \\ \hline
+    4 & 42.66992688 & 56.36907 & 56.33994633 & 56.0778672 & 53.457076 & 50.70305 & 38.26094 & 42.286992 & 
+      2.800991 \\
+    \hline
+  \end{tabular}
+  \caption{Real-time performance of the Raspberry Pi 3 depending on the number of cores used.}
+\end{table*}
+
+As is depicted in Table 1, the Raspberry Pi performed better, on average, when it utilized more 
+cores. With 4 cores, the Pi was able to meet the vast majority of its 50 ms deadlines, doing so in 
+almost 99\% of the processed frames. The Pi had the worst average performance when using only 1 core, 
+as it was unable to meet any of the 50 ms deadlines (its fastest processing time was just under 60 
+ms). Another observation is that the difference between using 2 cores and 3 cores is relatively 
+small. On average, using 3 cores only performed better by 2 ms, so the addition of one core in that 
+specific case offers relatively little improvement. However, please note that the average time when 
+using 3 cores was below the 50 ms deadline, while the average time using 2 cores was slightly above it. 
+One important observation that can be made is that of consistency, as using only 1 core had a much 
+steadier performance. As a result, the use of multiple cores is very beneficial in terms of reducing 
+the time it takes to complete real-time operations, but may ultimately result in processing times 
+that are more volatile.
+
+\subsection{Multimodel Performance}
+In a real world scenario, it is highly probable that multiple models will need to be used 
+simultaneously in order to perform various important tasks (angle prediction, object detection, 
+etc.)\cite{}. As such, we also tested the capability of the Raspberry Pi to run multple models at the 
+same time, and measure whether all models are able to meet their respective deadlines on a consistent 
+basis. Specifically, the Pi is tested in the cases of running 2 and 4 models simultaneously. For each 
+case, all models are allocated an equal number of cores, with 2 models having 2 cores each, and 4 
+models having 1 core each.
+
+\begin{table*}[t]
+  \begin{tabular} {| l | l | l | l | l | l | l | l | l | l |}
+  \hline
+  \textbf{num models} & \textbf{cores} & \textbf{mean} & \textbf{max} & \textbf{L1 refs} & \textbf{L1 
+    misses} & \textbf{L1 miss \%} & textbf{L2 refs} & \textbf{L2 misses} & \textbf{L2 miss \%} \\ \hline
+  1 & 0 &  62.48010421 & 66.83707237 & 27,809,824,750 & 435,932,903	& 1.568 & 2,830,285,204	& 
+    358,764,318 & 12.68 \\ \hline
+  1 & 0,1 & 51.35843444 & 72.92294502 & 30,354,597,625 &  477,953,407 & 1.575 & 3,308,819,327 & 
+    367,823,926 & 11.12 \\ \hline
+  2 & 0,1 & 58.03400564 & 83.67586136 & 30,426,714,898 & 490,709,126 & 1.613 & 3,912,431,599 & 
+    425,658,305 & 10.88 \\ \hline
+  2 & 2,3 & 56.39569068 & 82.88908005 & 30,350,459,058 & 480,166,192 & 1.582 & 3,877,553,207 & 
+    421,339,847 & 10.87 \\ \hline
+  4 & 0 & 77.90289187 & 104.2420864 & 27,853,570,186 & 452,700,705 & 1.625 & 3,360,219,598 & 
+    443,088,325 & 13.19 \\ \hline
+  4 & 1 & 78.88618684 & 108.7918282 & 27,868,355,666 & 464,332,042 & 1.666 & 3,415,627,978 & 
+    437,962,230 & 12.82 \\ \hline
+  4 & 2 & 77.80583382 & 111.9449139 & 27,799,436,982 & 444,828,792 & 1.6 & 3,449,341,398 & 440,580,899 
+    & 12.77 \\ \hline
+  4 & 3 & 77.86954451 & 99.28488731 & 27,875,785,919 & 445,229,272 & 1.597 & 3,408,895,945 & 
+    439,091,391 & 12.88 \\
+  \hline
+  \end{tabular}
+  \caption{Real-time performance of the Raspberry Pi 3 depending on the number of models being run 
+    simultaneously.}
+\end{table*}
+
+Ideally, each model run would replicate a single model run. However, Table 2 shows that such was not 
+the case. In the tests where 2 models ran simultaneously, both of the models showed average time 
+increases of around 5-7 ms, around 10\%, when compared to a baseline of 1 model running on two cores. 
+This change, however, did not seem to be caused by any form of cache interference as both L1 and L2 
+cache misses remained constant, regardless of the number of models running. The number of L1 misses 
+being close to 1.6\% of all references and the number of L2 misses being about 11\% of all 
+references.
+
+The difference was even greater in the case of 4 models running concurrently, as each one displayed 
+an average time increase of approximately 15 ms, around 30\%, when compared to a single model 
+running on 1 core. Once again, though, cache interference was not a factor as the number of L1 cache 
+misses remained 1.6\% of all references, and the number of L2 cache misses stayed at 13\% of all 
+references.
+
+\subsection{Performance Requirements}
+In the utilization of the Raspberry Pi 3 in our platform, there are a few factors that need to be 
+considered and/or enforced in order to guarantee that the Pi is able to consistently perform at a 
+desired level. Specifically, these issues all have the potential to negatively affect the cpu clock 
+speed/frequency, which would result in decreased performance. While, in the above experiments, the cpu 
+operated at a preferred clock speed of 1.2 GHz, it is entirely possible for the cpu to operate at a 
+lower frequency if the following problems are not taken into account.
+
+The most notable issue that can affect the cpu clock speed is that of the power supplied to the 
+Raspberry Pi. In essence, it is necessary that the Pi be supplied with 2 Amps, as any less could 
+hinder the Pi's ability to maintain a 1.2 GHz frequency. In experiments done with a power supply that 
+only provided 1 Amp, the Pi was unable to sustain a 1.2 GHz clock speed and, instead, fluctuated 
+between operating at 600 MHz and 1.2 GHz. As a result, it is necessary, or at least highly 
+recommended, that the power supply used for the Raspberry Pi 3 be capable of outputting 2 Amps, 
+otherwise optimal performance isn't guaranteed.
+
+Another factor that can affect clock speed is that of the cpu's temperature. Some model operations can 
+be computationally intensive, thus it is possible for the temperature of the cpu to become relatively 
+high. This can be especially problematic in situations where multiple models are running 
+simultaneously on the Pi. As a consequence, thermal throttling may be used to decrease the clock 
+speed so that the cpu temperature stays at a safe level. As such, the Raspberry Pi may not be suited 
+for prolonged use, especially in cases where the workload is relatively larger, like running multiple 
+models. Rather, the Pi seems to be better suited for running in set periods, after which it is turned 
+off or made idle so that the cpu is given time to cool down.
 \section{Related Work}

### 2017-09-27 paper/main.tex

@@
 
-\input{intro}
-\input{background}
+%% UPenn's f1/10 BOM: $3,628.37	
+%% http://f1tenth.org/
+%% http://selfdrivingcars.mit.edu/
+%% http://fast.scripts.mit.edu/racecar/
+%% https://github.com/mit-racecar
+
+\section{Introduction}
+Autonomous vehicles have been a topic of increasing interest in many
+disciplines in recent years. Due to the nature of autonomous vehicles,
+it is necessary that all real-time operations are successfully
+performed prior to their deadlines. This requires that the AV platform
+be capable of completing all necessary computations in a timely
+manner, while maintaining a high level of accuracy. Consequently, the
+cost of creating a platform capable of self-driving can be relatively
+high, especially in the current cost-sensitive climate of the
+automotive industry \cite{}. The overall price for an autonomous
+vehicle platform currently acts as a bottleneck, especially from an
+academic standpoint, where cost is an important factor. For many
+researchers and students, the charge associated with autonomous
+vehicle research is a significant barrier. These same parties may also
+be deterred from participating in autonomous vehicle research due to
+the potential for their developed platforms to break in the
+experimental process. The idea of building one relatively expensive
+platform may not be too daunting for some, but the notion of
+potentially making multiple identical platforms in case of accidents
+may be overwhelming. Toward this end, we explore the possibility of
+developing a low-cost system that still employs state-of-the-art AI
+technologies and is capable of executing real-time autonomous vehicle
+operations.
+
+In this paper, we present DeepPicar, a low-cost autonomous car
+platform for research and education. From hardware perspective, 
+DeepPicar is comprised of a Raspberry Pi 3 Model B quad-core
+computer, a web camera and a RC car, all of which are affordable
+components (less than \$200 in total).
+The DeepPicar, however, employs state-of-the-art AI
+techonologies, including a vision-based end-to-end control system that
+utilizes a deep convolutional nueral network (CNN).
+The network receives an image frame from a single forward
+looking camera, as input and generates a predicted steering angle
+value as output at each control period in \emph{real-time}. 
+The network is deep as it has 9 layers, about 27 million connections
+and 250 thousand parameters (weights).
+The network architecture is in fact identifical to the one
+used in Nvidia's DAVE-2 self-driving car~\cite{Bojarski2016}. Other
+than the obvious difference in scale (RC car vs. real car), from the
+computing perspective, the only differences between
+the two systems are that our system is implemented in
+TensorFlow~\cite{abadi2016tensorflow} and runs on a 
+Raspberry Pi 3 whereas Nvidia's DAVE-2 systems is implemented in Torch
+7~\cite{collobert2011torch7} and runs on a Drive PX computer (NVIDIA's
+automotive specialized computing system~\cite{drivepx}), which is more
+powerful but also more expensive.
+
+We follow the standard supervised learning methodology to train the
+system with a goal of following lanes on the ground, mimicking real
+roads. We first collected data for taining and validation by manually 
+controlling the RC car and recording the vision (from the webcam
+mounted on the RC-car) and the human control inputs. We then train the
+network offline using the collected data on a desktop computer, which
+equips a Nvidia GTX 1060 GPU. Finally, the trained network is copied
+back to the Raspberry Pi, which is then used to perform interference
+operations---locally on the Pi---in the car's main control loop in
+real-time. For real-time control, each \emph{inference} operation must
+be completed within the desired control period. (e.g., 50ms period for
+20Hz control frequency.)
+
+The primary {\bf contributions} of this paper are as follows:
+\begin{itemize}
+  \item We present the design, implementation, and case-study of a low-cost
+    autonomous vehicle that utilizes the state-of-the-art aritificial 
+    intelligence techniques.
+  \item We systematically analyze real-time characteristics of
+    multiple embedded computing platform, including the Raspberry Pi 3
+    Model B, in the contex of a vision-based autnomous driving.
+\end{itemize}
+
+%% We chose to use a Raspberry Pi 3 not only because of its
+%% affordability and availability but also because it is representative
+%% of today's mainstream low-end multicore processors, found in
+%% smartphones, tables and other embedded devices.
+%% The DeepPicar is comprised of a Raspberry
+%% Pi 3 Model B, a camera and a low cost RC car.
+
+%% By creating and using the Picar platform, we seek to accomplish the 
+%% following three goals. First, the significant reduction of the overall
+%% cost required 
+%% would make autonomous driving research more accessible to interested
+%% individuals/parties. Furthermore, it would remove the concerns of
+%% creating replacement platforms as each part, or the entire system,
+%% could be replaced relatively inexpensively. Second, the utilization of
+%% a more cost-efficient platform would allow for a better understanding
+%% of the performance requirements necessary for efficiently operating
+%% autonomous vehicles. Finally, we strive to achieve the ability to
+%% compare the real-time performance of multiple embedded computing
+%% platforms.
+
+%% We display the efficacy of the Picar by training and testing the
+%% DeepTesla DNN \cite{} in a custom-made environment. We find that the
+%% Picar is capable of consistently accomplishing all real-time
+%% operations required for autonomous driving in a 50 millisecond frame.
+
+The remaining sections of the paper are as follows: Section II gives
+an overview of the platform, including the high-level system and the
+methods used for training and inference. Section III discusses the
+ways/methodologies in which training data was collected. Section IV
+outlines the network architecture. This is followed by the real-time
+DNN inferencing in Section V. Section VI reviews the online/offline
+training done with the platform, with an evaluation given in Section
+VII. Section VIII gives a discussion of related works, and the paper
+finishes with conclusions in Section IX.
+
+\section{Background}
+Cite a paper \cite{barroso2009datacenter}.
+
+Cite multiple papers \cite{banga99resourcecontainers,barroso2009datacenter}
 

### 2017-09-27 paper/intro.tex
intro revision

@@
 
-Autonomous vehicles have been a topic of increasing interest in many
-disciplines in recent years. Due to the nature of autonomous vehicles,
+% broad context:
+% - advance in ai sparked interests in the robotics application, such as
+%   self-driving cars.
+% - in particular, deep neural network models are increasingly used
+%   for perception and control of a vehicle. say. AI workloads.
+%
+% 
+Self-driving cars have been a topic of increasing interest in recent
+years as multiple companies are actively developing related
+technologies. One of the key technologies is the use of deep neural
+network (DNNs) models for preception and control of the vehicle.
+
+% requirements:
+%   (1) real-time processing.
+%   (2) cost
+%   (3) safety
+%   (4) SWaP
+%
+% our goals:
+%   (1) build a safe and low cost testbed but utilizing the
+%       state-of-the-art AI technologies: DNN based end-to-end control
+%   (2) analyze real-time issues in DNN based end-to-end control.
+%   (3) evalute popular embedded platforms for the AI workloads
+%       (end-to-end control, perception).
+%       - Can apsberry pi handle AI workloads?
+%       - Can embedded GPU do better? How much?
+\fixme{TBD: requirements and goals reivison}
 it is necessary that all real-time operations are successfully
@@
 
+% DeepPicar introduction
 In this paper, we present DeepPicar, a low-cost autonomous car
@@
 
+% how we trained and tested (in terms of network performance)
 We follow the standard supervised learning methodology to train the
@@
 
+% how we evaluated (in terms real-time performance)
+% what are our findings?
+
+% contributions
 The {\bf contributions} of this paper are as follows:

### 2017-09-28 paper/intro.tex
still revising

@@
 %
-Self-driving cars have been a topic of increasing interest in recent
-years as multiple companies are actively developing related
-technologies. One of the key technologies is the use of deep neural
-network (DNNs) models for preception and control of the vehicle.
+Autonomous cars have been a topic of increasing interest in recent
+years as many companies are actively developing related hardware
+and software technologies for ever more autonomous decision making
+capabilities---ultimately fully autonomous driving capability without
+any human intervention. One of the key technologies for increased
+autonomy of a vehicle is deep neural network (DNN) models, which have
+been sucessfully applied in various perception and control tasks in
+recent years. Already, DNN models are used in commercial
+semi-autonomous vehicles such as Tesla Model S, which includes a poweful
+computer and accelerators (e.g., MobileEye EyeQ) to execute various
+DNN models, espeically for (vision-based) real-time perception
+tasks. DNN based end-to-end control of a vehicle is also actively
+explored in recent years~\cite{Bojarski2016}.
 
@@
 %       - Can embedded GPU do better? How much?
-\fixme{TBD: requirements and goals reivison}
+There are several challenges in studying autonomous vehicles.
+
 it is necessary that all real-time operations are successfully
@@
 The network receives an image frame from a single forward
-looking camera, as input and generates a predicted steering angle
+looking camera as input and generates a predicted steering angle
 value as output at each control period in \emph{real-time}. 
-The network is deep as it has 9 layers, about 27 million connections
+The network is \emph{deep} as it has 9 layers, about 27 million connections
 and 250 thousand parameters (weights).
 The network architecture is in fact identifical to the one
-used in Nvidia's DAVE-2 self-driving car~\cite{Bojarski2016}. Other
-than the obvious difference in scale (RC car vs. real car), from the
-computing perspective, the only differences between
-the two systems are that our system is implemented in
+used in NVIDIA's DAVE-2 self-driving car~\cite{Bojarski2016}. Other
+than the obvious difference in scale (RC car vs. real car) the only
+differences between the two systems---from the computing
+perspective---are that our system is implemented in 
 TensorFlow~\cite{abadi2016tensorflow} and runs on a 
-Raspberry Pi 3 whereas Nvidia's DAVE-2 systems is implemented in Torch
+Raspberry Pi 3 whereas NVIDIA's DAVE-2 systems is implemented in Torch
 7~\cite{collobert2011torch7} and runs on a Drive PX computer (NVIDIA's
@@
 network offline using the collected data on a desktop computer, which
-equips a Nvidia GTX 1060 GPU. Finally, the trained network is copied
+equips a NVIDIA GTX 1060 GPU. Finally, the trained network is copied
 back to the Raspberry Pi, which is then used to perform interference

### 2017-09-29 paper/intro.tex
still revising (unsatisfactory.)

@@
 capabilities---ultimately fully autonomous driving capability without
-any human intervention. One of the key technologies for increased
-autonomy of a vehicle is deep neural network (DNN) models, which have
-been sucessfully applied in various perception and control tasks in
-recent years. Already, DNN models are used in commercial
-semi-autonomous vehicles such as Tesla Model S, which includes a poweful
-computer and accelerators (e.g., MobileEye EyeQ) to execute various
-DNN models, espeically for (vision-based) real-time perception
-tasks. DNN based end-to-end control of a vehicle is also actively
-explored in recent years~\cite{Bojarski2016}.
+any human intervention. Deep neural networks (DNNs) have been
+sucessfully applied in various perception and control tasks in 
+recent years, and they are important workloads for autonomous vehicles
+as well. For example, Tesla Model S was known to use a specialized SoC
+(MobileEye EyeQ), which uses a deep neural network for vision-based
+real-time obstacle detection and avoidance. More recently, researchers
+are investigating DNN based end-to-end control of
+cars~\cite{Bojarski2016} and other robots. It is expected that more
+DNNs based Artifical Intelligence workloads may be used in future
+autonomous vehicles.
 
-% requirements:
-%   (1) real-time processing.
-%   (2) cost
-%   (3) safety
-%   (4) SWaP
-%
+% need testbeds.
+% but challenges 
+%   (1) real-time processing (2) cost (3) safety (4) SWaP
+Developing and testing DNN models for autonomous vehicles is
+however, challenging for a variety of reasons. The first challegen is
+finding a suitable computing platform. It is generally known that DNN
+models are computationally demainding. Furtheremore, because many of
+the AI workloads, such as models for end-to-end control or perception,
+have strict real-time requirements as cars have to interact with the
+physical world in real-time. On the other, the computing platform must
+also satisfy cost, size, weight, and power constraints. 
+The second challenge is preparing a
+physical car platform to collect data and test the software. While
+instrumenting a real-car would be most realistic, as many
+companies such as google and uber are doing, but doing it is costly
+and would pose safety concerns. 
 % our goals:
@@
 %       - Can embedded GPU do better? How much?
-There are several challenges in studying autonomous vehicles.
+Therefore, there is a strong need for low-cost testbeds that still
+employs state-of-the-art AI technologies and are capable of executing
+real-time autonomous vehicle operations.
 
-it is necessary that all real-time operations are successfully
-performed prior to their deadlines. This requires that the AV platform
-be capable of completing all necessary computations in a timely
-manner, while maintaining a high level of accuracy. Consequently, the
-cost of creating a platform capable of self-driving can be relatively
-high, especially in the current cost-sensitive climate of the
-automotive industry \cite{}. The overall price for an autonomous
-vehicle platform currently acts as a bottleneck, especially from an
-academic standpoint, where cost is an important factor. For many
-researchers and students, the charge associated with autonomous
-vehicle research is a significant barrier. These same parties may also
-be deterred from participating in autonomous vehicle research due to
-the potential for their developed platforms to break in the
-experimental process. The idea of building one relatively expensive
-platform may not be too daunting for some, but the notion of
-potentially making multiple identical platforms in case of accidents
-may be overwhelming. Toward this end, we explore the possibility of
-developing a low-cost system that still employs state-of-the-art AI
-technologies and is capable of executing real-time autonomous vehicle
-operations.
+%% it is necessary that all real-time operations are successfully
+%% performed prior to their deadlines. This requires that the AV platform
+%% be capable of completing all necessary computations in a timely
+%% manner, while maintaining a high level of accuracy. Consequently, the
+%% cost of creating a platform capable of self-driving can be relatively
+%% high, especially in the current cost-sensitive climate of the
+%% automotive industry \cite{}. The overall price for an autonomous
+%% vehicle platform currently acts as a bottleneck, especially from an
+%% academic standpoint, where cost is an important factor. For many
+%% researchers and students, the charge associated with autonomous
+%% vehicle research is a significant barrier. These same parties may also
+%% be deterred from participating in autonomous vehicle research due to
+%% the potential for their developed platforms to break in the
+%% experimental process. The idea of building one relatively expensive
+%% platform may not be too daunting for some, but the notion of
+%% potentially making multiple identical platforms in case of accidents
+%% may be overwhelming.
+%% Toward this end, we explore the possibility of
+%% developing a low-cost system that still employs state-of-the-art AI
+%% technologies and is capable of executing real-time autonomous vehicle
+%% operations.
 

### 2017-09-29 multicore_timings.sh
removed colons from folder name and updated script

@@ -1,51 +1,0 @@
-timestamp=$(date +%y-%m-%d_%H:%M:%S)
-
-mkdir -p logs/$timestamp
-
-echo single-core
-perf stat taskset -c 0 python test-model.py 1 > logs/$timestamp/single-core.txt &
-wait
-echo end single-core
-
-echo 2-parallel-single-core
-perf stat taskset -c 0 python test-model.py 1 > logs/$timestamp/2-single-core_cpu0.txt &
-perf stat taskset -c 1 python test-model.py 1 > logs/$timestamp/2-single-core_cpu1.txt &
-wait
-echo end 2-parallel-single-core
-
-echo 3-parallel-single-core
-perf stat taskset -c 0 python test-model.py 1 > logs/$timestamp/3-single-core_cpu0.txt &
-perf stat taskset -c 1 python test-model.py 1 > logs/$timestamp/3-single-core_cpu1.txt &
-perf stat taskset -c 2 python test-model.py 1 > logs/$timestamp/3-single-core_cpu2.txt &
-wait
-echo end 3-parallel-single-core
-
-echo 4-parallel-single-core
-perf stat taskset -c 0 python test-model.py 1 > logs/$timestamp/4-single-core_cpu0.txt &
-perf stat taskset -c 1 python test-model.py 1 > logs/$timestamp/4-single-core_cpu1.txt &
-perf stat taskset -c 2 python test-model.py 1 > logs/$timestamp/4-single-core_cpu2.txt &
-perf stat taskset -c 3 python test-model.py 1 > logs/$timestamp/4-single-core_cpu3.txt &
-wait
-echo end 4-parallel-single-core
-
-echo dual-core
-perf stat taskset -c 0,1 python test-model.py 2 > logs/$timestamp/dual-core.txt &
-wait
-echo end dual-core
-
-echo 2-parallel-dual-core
-perf stat taskset -c 0,1 python test-model.py 2 > logs/$timestamp/2-dual-core_cpu01.txt &
-perf stat taskset -c 2,3 python test-model.py 2 > logs/$timestamp/2-dual-core_cpu23.txt &
-wait
-echo end 2-parallel-dual-core
-
-echo tri-core
-perf stat taskset -c 0,1,2 python test-model.py 3 > logs/$timestamp/tri-core.txt &
-wait
-echo end tri-core
-
-echo quad-core
-perf stat taskset -c 0,1,2,3 python test-model.py 4 > logs/$timestamp/quad-core.txt &
-wait
-echo end quad-core
-

### 2017-09-29 test-model_core_timings.sh

@@ -1,0 +1,68 @@
+timestamp=$(date +%y-%m-%d_%H-%M-%S)
+
+echo "Which board is this test being run on?  Input a string and press [ENTER]"
+read board
+
+identification=$board
+identification+="_"
+identification+=$timestamp
+
+mkdir -p logs/test-model/test-model/$identification
+echo "This test will take approximately 45 minutes to complete"
+echo "The output files can be found in: logs/test-model/$identification"
+
+echo single-core
+perf stat -o logs/test-model/$identification/single-core_perf.txt taskset -c 0 python test-model.py 1 > logs/test-model/$identification/single-core.txt &
+wait
+echo end single-core
+sleep 300
+
+echo 2-parallel-single-core
+perf stat -o logs/test-model/$identification/2-single-core_cpu0_perf.txt taskset -c 0 python test-model.py 1 > logs/test-model/$identification/2-single-core_cpu0.txt &
+perf stat -o logs/test-model/$identification/2-single-core_cpu1_perf.txt taskset -c 1 python test-model.py 1 > logs/test-model/$identification/2-single-core_cpu1.txt &
+wait
+echo end 2-parallel-single-core
+sleep 300
+
+echo 3-parallel-single-core
+perf stat -o logs/test-model/$identification/3-single-core_cpu0_perf.txt taskset -c 0 python test-model.py 1 > logs/test-model/$identification/3-single-core_cpu0.txt &
+perf stat -o logs/test-model/$identification/3-single-core_cpu1_perf.txt taskset -c 1 python test-model.py 1 > logs/test-model/$identification/3-single-core_cpu1.txt &
+perf stat -o logs/test-model/$identification/3-single-core_cpu2_perf.txt taskset -c 2 python test-model.py 1 > logs/test-model/$identification/3-single-core_cpu2.txt &
+wait
+echo end 3-parallel-single-core
+sleep 300
+
+echo 4-parallel-single-core
+perf stat -o logs/test-model/$identification/4-single-core_cpu0_perf.txt taskset -c 0 python test-model.py 1 > logs/test-model/$identification/4-single-core_cpu0.txt &
+perf stat -o logs/test-model/$identification/4-single-core_cpu1_perf.txt taskset -c 1 python test-model.py 1 > logs/test-model/$identification/4-single-core_cpu1.txt &
[…]

### 2017-09-29 paper/intro.tex

@@
 years as many companies are actively developing related hardware
-and software technologies for ever more autonomous decision making
-capabilities---ultimately fully autonomous driving capability without
+and software technologies toward fully autonomous driving capability without
 any human intervention. Deep neural networks (DNNs) have been
@@
 recent years, and they are important workloads for autonomous vehicles
-as well. For example, Tesla Model S was known to use a specialized SoC
-(MobileEye EyeQ), which uses a deep neural network for vision-based
+as well. For example, Tesla Model S was known to use a specialized
+chip (MobileEye EyeQ), which uses a deep neural network for vision-based
 real-time obstacle detection and avoidance. More recently, researchers
@@
 cars~\cite{Bojarski2016} and other robots. It is expected that more
-DNNs based Artifical Intelligence workloads may be used in future
+DNN based Artifical Intelligence workloads may be used in future
 autonomous vehicles.
 
-% need testbeds.
-% but challenges 
-%   (1) real-time processing (2) cost (3) safety (4) SWaP
-Developing and testing DNN models for autonomous vehicles is
-however, challenging for a variety of reasons. The first challegen is
-finding a suitable computing platform. It is generally known that DNN
-models are computationally demainding. Furtheremore, because many of
-the AI workloads, such as models for end-to-end control or perception,
-have strict real-time requirements as cars have to interact with the
-physical world in real-time. On the other, the computing platform must
-also satisfy cost, size, weight, and power constraints. 
-The second challenge is preparing a
-physical car platform to collect data and test the software. While
-instrumenting a real-car would be most realistic, as many
-companies such as google and uber are doing, but doing it is costly
-and would pose safety concerns. 
-% our goals:
-%   (1) build a safe and low cost testbed but utilizing the
-%       state-of-the-art AI technologies: DNN based end-to-end control
-%   (2) analyze real-time issues in DNN based end-to-end control.
-%   (3) evalute popular embedded platforms for the AI workloads
-%       (end-to-end control, perception).
-%       - Can apsberry pi handle AI workloads?
-%       - Can embedded GPU do better? How much?
-Therefore, there is a strong need for low-cost testbeds that still
-employs state-of-the-art AI technologies and are capable of executing
-real-time autonomous vehicle operations.
+% big problem
+Executing these AI workloads on a embedded computing platform that can
+be used in a car poses several additional challenges. First, many AI
+workloads in vehicles are computationaly demanding and have strict
+real-time requirements. For example, latency in a vision-based object
+detection task may directly linked to safety of the vehicle. This
+requires a high computing capacity as well as means to guaranteeing
+the timings. On the other hand, the computing hardware platform must
+also satisfy cost, size, weight, and power constraints, which require
+highly efficient computing platform. These two conflicting
+requirements  complicate the platform selection process as observed in
+~\cite{Otterness2017}, which evaluated vision workloads on
+NVIDIA's Tegra TX1 platform.
+%% For example, while today's self-driving car
+%% prototype equip more \$100,000 
+%% of computers and sensors~\cite{juliussen2014emerging}, a study
+%% found that aveage consumers are willing to pay much less amount of
+%% extra cost for a self-driving capability~\cite{Daziano2017}.
+%% https://qz.com/924212/what-it-really-costs-to-turn-a-car-into-a-self-driving-vehicle/
 
-%% it is necessary that all real-time operations are successfully
-%% performed prior to their deadlines. This requires that the AV platform
-%% be capable of completing all necessary computations in a timely
-%% manner, while maintaining a high level of accuracy. Consequently, the
-%% cost of creating a platform capable of self-driving can be relatively
-%% high, especially in the current cost-sensitive climate of the
-%% automotive industry \cite{}. The overall price for an autonomous
-%% vehicle platform currently acts as a bottleneck, especially from an
-%% academic standpoint, where cost is an important factor. For many
-%% researchers and students, the charge associated with autonomous
-%% vehicle research is a significant barrier. These same parties may also
-%% be deterred from participating in autonomous vehicle research due to
-%% the potential for their developed platforms to break in the
-%% experimental process. The idea of building one relatively expensive
-%% platform may not be too daunting for some, but the notion of
-%% potentially making multiple identical platforms in case of accidents
-%% may be overwhelming.
-%% Toward this end, we explore the possibility of
-%% developing a low-cost system that still employs state-of-the-art AI
-%% technologies and is capable of executing real-time autonomous vehicle
-%% operations.
+% related work and remaing problems
+To understand what kind of computing hardware is needed for AI
+workloads, we need a testbed and realistic workloads. While a real car
+testbed would be most ideal, it is not only highly expensive but also
+poses serious safety concerns that hinder development and exploration.
+Therefore, there is a strong need for safer and less costly testbeds
+that still employ state-of-the-art AI technologies. There are already
+several relatively inexpensive RC-car based testbeds, such as MIT
+RaceCar~\cite{shin2017project} and UPenn's F1$/$10 racecar~\cite{upennf1tenth}.
+However, these RC-car testbeds still cost more than \$3,000.
+
+% our goals
+Instead, we want to build a low cost testbed that still employs the
+state-of-the art AI technologies. Specifically, we focus on a deep
+convolutional neural network (CNN) based end-to-end control system,
+which was developed for a real self-driving car, NVIDIA
+DAVE-2~\cite{Bojarski2016}, and use the same moethodology on a
+smaller, safer and less costly setup. In developing the testbed, our
+goals are (1) analyze real-time issues in DNN based end-to-end
+control; (2) evalute real-time performance of contemporarly embedded
+multicore platforms for the AI workloads.
 
@@
 value as output at each control period in \emph{real-time}. 
-The network is \emph{deep} as it has 9 layers, about 27 million connections
+The network has 9 layers, about 27 million connections
 and 250 thousand parameters (weights).
-The network architecture is in fact identifical to the one
-used in NVIDIA's DAVE-2 self-driving car~\cite{Bojarski2016}. Other
-than the obvious difference in scale (RC car vs. real car) the only
-differences between the two systems---from the computing
-perspective---are that our system is implemented in 
-TensorFlow~\cite{abadi2016tensorflow} and runs on a 
-Raspberry Pi 3 whereas NVIDIA's DAVE-2 systems is implemented in Torch
-7~\cite{collobert2011torch7} and runs on a Drive PX computer (NVIDIA's
-automotive specialized computing system~\cite{drivepx}), which is more
-powerful but also more expensive.
+The network architecture is identifical to that of NVIDIA's DAVE-2
+self-driving car~\cite{Bojarski2016}, which uses much more powerful
+computer (NVIDIA's automotive specialized DRIV-PX
+computer~\cite{drivepx}) than the Raspberry Pi~3.
+We chose to use the Pi 3 not only because of its
+affordability but also because it is representative
+of today's mainstream low-end multicore platforms, found in
+smartphones and other embedded devices.
 
-% how we trained and tested (in terms of network performance)
+%% Other than the difference in scale (RC car vs. real car), the only other
+%% differences between the two systems---from the computing
+%% perspective---are that our system is implemented in
+%% TensorFlow~\cite{abadi2016tensorflow} and runs on a Raspberry Pi 3
+%% whereas NVIDIA's DAVE-2 systems is implemented in Torch
+%% 7~\cite{collobert2011torch7} and runs on a Drive PX computer (NVIDIA's
+%% automotive specialized computing system~\cite{drivepx}), which is more
+%% powerful but also more expensive.
+
+% how we trained (maybe moved to a later section)
 We follow the standard supervised learning methodology to train the
-system with a goal of following lanes on the ground, mimicking real
-roads. We first collected data for taining and validation by manually 
+system to follow tracks on the ground. We collect data for
+taining and validation by manually  
 controlling the RC car and recording the vision (from the webcam
@@
 operations---locally on the Pi---in the car's main control loop in
-real-time. For real-time control, each \emph{inference} operation must
+real-time. For real-time control, each inference operation must
 be completed within the desired control period. (e.g., 50ms period for
 20Hz control frequency.)
+% how we evaluated (in terms real-time performance)
 
-% how we evaluated (in terms real-time performance)
 % what are our findings?
+\fixme{TBD: Summary of our findings}
 
@@
 
-%% We chose to use a Raspberry Pi 3 not only because of its
-%% affordability and availability but also because it is representative
-%% of today's mainstream low-end multicore processors, found in
-%% smartphones, tables and other embedded devices.
 %% The DeepPicar is comprised of a Raspberry
@@
 finishes with conclusions in Section IX.
+
+%% it is necessary that all real-time operations are successfully
+%% performed prior to their deadlines. This requires that the AV platform
+%% be capable of completing all necessary computations in a timely
+%% manner, while maintaining a high level of accuracy. Consequently, the
+%% cost of creating a platform capable of self-driving can be relatively
+%% high, especially in the current cost-sensitive climate of the
+%% automotive industry \cite{}. The overall price for an autonomous
+%% vehicle platform currently acts as a bottleneck, especially from an
+%% academic standpoint, where cost is an important factor. For many
+%% researchers and students, the charge associated with autonomous
+%% vehicle research is a significant barrier. These same parties may also
+%% be deterred from participating in autonomous vehicle research due to
+%% the potential for their developed platforms to break in the
+%% experimental process. The idea of building one relatively expensive
+%% platform may not be too daunting for some, but the notion of
+%% potentially making multiple identical platforms in case of accidents
+%% may be overwhelming.
+%% Toward this end, we explore the possibility of
+%% developing a low-cost system that still employs state-of-the-art AI
+%% technologies and is capable of executing real-time autonomous vehicle
+%% operations.

### 2017-09-29 paper/main.tex
Add parts of platform overview and modify figures

--- 2017-09-27 paper/main.tex
@@
 \section{Platform Overview}
-...
+
+\begin{figure}[h]
+  \centering
+  \includegraphics[width=.4\textwidth]{Picar_Picture}
+  \caption{ The Picar Platform. }
+\end{figure}
+
+\subsection{Platform Components}
+We employ a small and relatively inexpensive RC car that is capable of performing basic automotive 
+operations. However, the car we use doesn't replicate the capabilities of other autonomous vehicle 
+platforms, as it is more simplistic in nature. Notably, the RC car we use only has three possible 
+options in terms of turning: center, left, and right. As such, control of the car may be less precise at 
+times, and may negatively affect performance. The car is capable of other operations that aren't used 
+within the scope of this platform. Specifically, the car is able to drive in reverse and the driving 
+speed can be changed. In our experiments, we have the RC car drive forward, and at a constant speed.
+
+Our platform also comes equipped with a camera that is used for both recording training videos and 
+providing input frames to the model whenever the Picar is self-driving. We chose the Playstation Eye 
+Camera due to its ability to reach and maintain higher fps levels while also remaining relatively 
+low-cost. One concern, however, is the affect of camera latency on the self-driving performance of our 
+platform. While humans are able to see environmental changes in real-time, the same can't be said for 
+cameras since there is a delay between when frame capture by the camera, and when it is read by the 
+Raspberry Pi 3. As a result, it is possible for the model to be given an input frame that is different 
+from the real world, thus impacting the performance of the Picar.
+
+\subsection{Model Training}
+In order to operate the Picar as an autonomous vehicle, we utilize the DeepTesla library, which is 
+capable of training a deep neural network (DNN) with end-to-end learning, that could then be used by the 
+Picar for autonomous driving. Please note that it is more efficient to execute the actual model training 
+process on a different system with an NVIDIA GPU, rather than on the Raspberry Pi 3 itself. This is 
+because DeepTesla trains the DNN with the Tensorflow library, which greatly benefits from the use of 
+gpu-enabled operations. For comparison, training a model on the Pi took approximately 6 hours, whereas 
+the time it took on a computer with a NVIDIA GPU was only around 5 minutes!
+
+We train our model on a custom made track/lane composed of black and white duct tape, where the black 
+tape represents the bounds of the track, and the white tape marks the center of the lane. The model is 
+taught to stay close to the center of the lane for as long as possible, and to turn whenever it 
+reaches/crosses the outer lane bounds so that it remains on the track. For data, we navigate and record 
+the car going both ways across the track, and use the collected videos to train the model(s) that will 
+be used later for angle prediction while the car is self-driving.
+
 \section{Data Collection}
@@
 \subsection{Real-Time Operations}
+In a real-time operations, the system has to be capable of consistently executing all necessary 
+functions before their given deadlines. In the case of our platform, it has to process every given frame 
+and get the predicted angles from the model. In order to determine if the Raspberry Pi 3 is capable of 
+this, we tested the platform and recorded the time it took for the Pi to complete all real-time 
+functions for every given frame.For this experiment, all four of the Pi's cpu cores were utilized, and 
+only one model was run. As can be seen in Figure 1, the Pi was able to meet almost all of its deadlines, 
+while only missing it in a few instances.
+
+\begin{figure}[h]
+  \centering
+  \includegraphics[width=.5\textwidth]{Total_Processing_Time}
+  \caption{ Real-time performance of the Raspberry Pi 3 in processing 1000 frames.}
+\end{figure}
+
 In our platform, three main real-time tasks are performed during autonomous operation. 
@@
 or video stream, (2) preprocessing the acquired frame so that it is compatible with the DNN, and (3) 
-feeding the frame to and getting the angle prediction from the model. However, the time it takes to 
+feeding the frame to, and getting the angle prediction from, the model. However, the time it takes to 
 complete each of the operations will differ, with at least one of them being the dominating step in 
@@
 
-\begin{figure}[h]
+\begin{table}[h]
   \centering
-  \includegraphics[width=.5\textwidth]{Operation_Time}
-  \caption{ Real-time operation completion time in processing frames.}
-\end{figure}
+  \begin{tabular} {| l | l | l | l | l |}
+    \hline
+    \textbf{operation} & \textbf{mean} & \textbf{max} & \textbf{99pct} & \textbf{stdev} \\ \hline 
+    Frame Capture & 2.28 & 4.94 & 63.31 &  .51\\ \hline
+    Preprocessing & 3.09 & 4.6 & 3.31 & .1 \\ \hline
+    Angle Prediction & 37.3 & 51.03 & 45.48 & 2.75 \\ 
+    \hline
+  \end{tabular}
+  \caption{Real-time performance of the Raspberry Pi 3 depending on the number of cores used.}
+\end{table}
 
@@
 
-\begin{table*}
+\begin{table}[h]
+  \centering
   \begin{tabular} {| l | l | l | l | l | l | l | l | l | l |}
     \hline
-    \textbf{num cores} & \textbf{mean} & \textbf{max} & \textbf{99.999pct} & \textbf{99.99pct} & 
-      \textbf{99.9pct} & \textbf{99pct} & \textbf{min} & \textbf{median} & \textbf{stdev} \\ \hline 
-    1 & 61.95813584 & 65.99712 & 65.99277693 & 65.9536558 & 65.562445 & 63.30799 & 59.304 & 61.975479 & 
-      0.506466 \\ \hline
-    2 & 50.49184 895 & 71.54608 & 71.5404019 & 71.4893194 & 70.978494 & 70.02651 & 46.93723 & 50.011516 & 
-      3.162015 \\ \hline
-    3 & 48.11392045 & 72.21985 & 72.21789317 & 72.200294 & 72.024303 & 58.44643 & 41.15391 & 48.444033 & 
-      4.178325 \\ \hline
-    4 & 42.66992688 & 56.36907 & 56.33994633 & 56.0778672 & 53.457076 & 50.70305 & 38.26094 & 42.286992 & 
-      2.800991 \\
+    \textbf{num cores} & \textbf{mean} & \textbf{max} & \textbf{99pct} & \textbf{stdev} \\ \hline 
+    1 & 61.96 & 66 & 63.31 &  .51\\ \hline
+    2 & 50.49 & 71.55 & 70.03 & 4.18 \\ \hline
+    3 & 48.11 & 72.22 & 58.45 & 2.8 \\ \hline
+    4 & 42.67 & 56.37 & 50.70 & 2.8 \\
     \hline
@@
   \caption{Real-time performance of the Raspberry Pi 3 depending on the number of cores used.}
-\end{table*}
+\end{table}
 
@@
 
-\begin{table*}[t]
-  \begin{tabular} {| l | l | l | l | l | l | l | l | l | l |}
+\begin{table*}
+  \begin{tabular} {| l | l | l | l | l | l | l | l | l |}
   \hline
-  \textbf{num models} & \textbf{cores} & \textbf{mean} & \textbf{max} & \textbf{L1 refs} & \textbf{L1 
+  \textbf{num models} & \textbf{cores} & \textbf{mean} & \textbf{L1 refs} & \textbf{L1 
     misses} & \textbf{L1 miss \%} & textbf{L2 refs} & \textbf{L2 misses} & \textbf{L2 miss \%} \\ \hline
-  1 & 0 &  62.48010421 & 66.83707237 & 27,809,824,750 & 435,932,903	& 1.568 & 2,830,285,204	& 
-    358,764,318 & 12.68 \\ \hline
-  1 & 0,1 & 51.35843444 & 72.92294502 & 30,354,597,625 &  477,953,407 & 1.575 & 3,308,819,327 & 
-    367,823,926 & 11.12 \\ \hline
-  2 & 0,1 & 58.03400564 & 83.67586136 & 30,426,714,898 & 490,709,126 & 1.613 & 3,912,431,599 & 
-    425,658,305 & 10.88 \\ \hline
-  2 & 2,3 & 56.39569068 & 82.88908005 & 30,350,459,058 & 480,166,192 & 1.582 & 3,877,553,207 & 
-    421,339,847 & 10.87 \\ \hline
-  4 & 0 & 77.90289187 & 104.2420864 & 27,853,570,186 & 452,700,705 & 1.625 & 3,360,219,598 & 
-    443,088,325 & 13.19 \\ \hline
-  4 & 1 & 78.88618684 & 108.7918282 & 27,868,355,666 & 464,332,042 & 1.666 & 3,415,627,978 & 
-    437,962,230 & 12.82 \\ \hline
-  4 & 2 & 77.80583382 & 111.9449139 & 27,799,436,982 & 444,828,792 & 1.6 & 3,449,341,398 & 440,580,899 
-    & 12.77 \\ \hline
-  4 & 3 & 77.86954451 & 99.28488731 & 27,875,785,919 & 445,229,272 & 1.597 & 3,408,895,945 & 
-    439,091,391 & 12.88 \\
-  \hline
+  1 & 0,1 & 51.35 & 3.04E+10 & 4.78E+08 & 1.58 & 3.31E+09 & 3.68E+08 & 11.12\\ \hline
+  2 & 0,1 & 58.03 & 3.04E+10 & 4.91E+08 & 1.61 & 3.91E+09 & 4.26E+08 & 10.88 \\ \hline
+  2 & 2,3 & 56.4 & 3.04E+10 & 4.80E+08 & 1.58 & 3.88E+09 & 4.21E+08 & 10.87 \\ \hline
   \end{tabular}
-  \caption{Real-time performance of the Raspberry Pi 3 depending on the number of models being run 
-    simultaneously.}
+  \caption{Real-time performance of the Raspberry Pi 3 when 2 models are running simultaneously.}
 \end{table*}
@@
 references.
+
+\begin{table*}
+  \begin{tabular} {| l | l | l | l | l | l | l | l | l |}
+  \hline
+  \textbf{num models} & \textbf{core} & \textbf{mean} & \textbf{L1 refs} & \textbf{L1 
+    misses} & \textbf{L1 miss \%} & textbf{L2 refs} & \textbf{L2 misses} & \textbf{L2 miss \%} \\ \hline
+    1 & 0 &  62.48 & 2.78E+10 & 4.36E+08 & 1.57 & 2.83E+09 & 3.59E+08 & 12.68  \\ \hline
+    4 & 0 & 77.9 & 2.79E+10 & 4.53E+08 & 1.63 & 3.36E+09 & 4.43E+08 & 13.19 \\ \hline
+    4 & 1 & 78.89 & 2.79E+10 & 4.64E+08 & 1.67 & 3.42E+09 & 4.38E+08 & 12.82 \\ \hline
+    4 & 2 & 77.81 & 2.78E+10 & 4.45E+08 & 1.6 & 3.45E+09 & 4.41E+08 & 12.77 \\ \hline
+    4 & 3 & 77.87 & 2.79E+10 & 4.45E+08 & 1.60 & 3.41E+09 & 4.39E+08 & 12.88 \\ \hline
+  \end{tabular}
+  \caption{Real-time performance of the Raspberry Pi when 4 models are running simultaneously.}
+\end{table*}
 

### 2017-09-29 paper/intro.tex
minor revision

@@
 affordability but also because it is representative
-of today's mainstream low-end multicore platforms, found in
+of today's mainstream low-end embedded multicore platforms, found in
 smartphones and other embedded devices.

### 2017-09-29 paper/background.tex
still revising (unsatisfactory.)
background revision (WIP)

@@
 
-(1) AI and Robotics
+\subsection{Deep End-to-End Control}
+End-to-end training of deep neural network models is an active
+research area in the fields of AI~\cite{Levine2016}.
+
+\begin{figure}[h]
+  \centering
+  \includegraphics[width=.5\textwidth]{figs/endtoend}
+  \caption{Standard robotics control vs. DNN based end-to-end
+    control. \fixme{figure must be redrawn}}
+\end{figure}
+
 - explosion of AI
@@
 %% https://github.com/mit-racecar
+%% https://mit-racecar.github.io/
 
-(2) Multicore SBCs
+\subsection{Embedded Multicore Single-Board-Computers}
+
 - computing has been a obstacle.

### 2017-09-29 paper/intro.tex
background revision (WIP)

--- 2017-09-29 paper/intro.tex
@@
 smaller, safer and less costly setup. In developing the testbed, our
-goals are (1) analyze real-time issues in DNN based end-to-end
-control; (2) evalute real-time performance of contemporarly embedded
+goals are (1) to analyze real-time issues in DNN based end-to-end
+control; (2) to evalute real-time performance of contemporarly embedded
 multicore platforms for the AI workloads.
@@
 computer, a web camera and a RC car, all of which are affordable
-components (less than \$200 in total).
+components (less than \$100 in total).
 The DeepPicar, however, employs state-of-the-art AI
@@
 self-driving car~\cite{Bojarski2016}, which uses much more powerful
-computer (NVIDIA's automotive specialized DRIV-PX
-computer~\cite{drivepx}) than the Raspberry Pi~3.
-We chose to use the Pi 3 not only because of its
+computer (Drive PX computer~\cite{drivepx}) than Raspberry Pi~3.
+We chose to use Pi 3 not only because of its
 affordability but also because it is representative
-of today's mainstream low-end multicore platforms, found in
+of today's mainstream low-end embedded multicore platforms, found in
 smartphones and other embedded devices.
@@
 % how we trained (maybe moved to a later section)
-We follow the standard supervised learning methodology to train the
-system to follow tracks on the ground. We collect data for
+We apply a standard imitation learning methodology to train the cat to
+follow tracks on the ground. We collect data for 
 taining and validation by manually  

### 2017-09-30 paper/background.tex
background revision (WIP)

@@
 
-\subsection{Deep End-to-End Control}
-End-to-end training of deep neural network models is an active
-research area in the fields of AI~\cite{Levine2016}.
+In this section, we provide background on the application of deep
+learning in robotics, particulary autonomous vehicles.
+
+\subsection{Deep End-to-End Learning in Autonomous Vehicles}
+
+%% - explosion of AI
+%% - in particular, application of DNN in perception and control of robotics systems.
+%% - end-to-end control is a promising technique.
+%%   levine's publications?
+%% - examples: nvidia's DAVE-II prototype, forest navigating drone
+%% challenge problem: computing at low cost?
+
+To solve the problem of autonomous driving, a standard approach has
+been decomposing the problem into multiple sub-problems,
+such as lane marking detection, path planning, and low-level
+control, which together form a processing pipeline~\cite{Bojarski2016}.
+Recently, researchers are exploring another approach that dramatically
+simplifies the standard control pipeline by applying deep neural
+networks to directly produce control outputs from senor
+inputs~\cite{Levine2016}}. Figure~\ref{fig:end-to-end-control}
+shows the differences between two two approaches.
 
@@
     control. \fixme{figure must be redrawn}}
+  \label{fig:end-to-end-control}
 \end{figure}
 
-- explosion of AI
-- in particular, application of DNN in perception and control of robotics systems.
-- end-to-end control is a promising technique.
-  levine's publications?
-- examples: nvidia's DAVE-II prototype, forest navigating drone
-challenge problem: computing at low cost?
+The use of neural networks for end-to-end control of autonomous
+driving was first demonstrated in late 1980s~\cite{Pomerleau1989}, but
+its success has been limited at the time as its performance was not
+sufficient enough.
+
+%% \cite{Levine2016}: ``In this paper, we aim to answer
+%% the following question: does training the perception and control systems jointly end-toend
+%% provide better performance than training each component separately?''
+
+%% \cite{Bojarski2016} nvidia paper
+%% ``We trained a convolutional neural network (CNN) to map raw pixels from
+%% a sin- gle front-facing camera directly to steering commands.''
+
+%% ``Compared to explicit decomposition of the problem, such as lane
+%% marking detec- tion, path planning, and control, our end-to-end system
+%% optimizes all processing steps simultaneously. We''
 

### 2017-10-01 paper/background.tex
..

@@
 networks to directly produce control outputs from senor
-inputs~\cite{Levine2016}}. Figure~\ref{fig:end-to-end-control}
+inputs~\cite{Levine2016}. Figure~\ref{fig:end-to-end-control}
 shows the differences between two two approaches.

### 2017-10-01 paper/background.tex
background revision (WIP)

@@
 The use of neural networks for end-to-end control of autonomous
-driving was first demonstrated in late 1980s~\cite{Pomerleau1989}, but
-its success has been limited at the time as its performance was not
-sufficient enough.
+driving was first demonstrated in late 1980s~\cite{Pomerleau1989},
+using a small fully connected neural network; and subsequently in a
+DARPA Autonomous Vehicle (DAVE) project in early
+2000s~\cite{LeCun:04}, using a 6 layer convolutional neural network
+(CNN); and most recently in NVIDIA's DAVE-2
+project~\cite{Bojarski2016}, using 9 layer CNN. In all these projects,
+the neural network models take raw image pixels as input and directly
+produce steering control commands bypassing all intermediary steps and
+hand-written rules in conventional robotics control approach. The
+NVIDIA's latest effort reports that their trained CNN successfully 
+self-drive their modified cars in public roads without human
+intervention~\cite{Bojarski2016}.
 
@@
 
+Recent successes in AI, including the NVIDIA's DAVE-2 showing, are due
+in large part to the increased computing capability of computers,
+which affored researchers to train ever deeper networks with high
+accuracy. In automotive systems,
+
 - computing has been a obstacle.

### 2017-10-01 paper/background.tex
revision

@@
 In this section, we provide background on the application of deep
-learning in robotics, particulary autonomous vehicles.
+learning in robotics, particularly autonomous vehicles.
 
@@
 inputs~\cite{Levine2016}. Figure~\ref{fig:end-to-end-control}
-shows the differences between two two approaches.
+shows the differences between two approaches.
 
@@
 
-Recent successes in AI, including the NVIDIA's DAVE-2 showing, are due
-in large part to the increased computing capability of computers,
-which affored researchers to train ever deeper networks with high
-accuracy. In automotive systems,
+Recent successes in AI, including NVIDIA's DAVE-2 showing, are due
+in large part to the increased computing performance,
+which afforded researchers to train and use ever deeper neural networks with
+high accuracy. For practical, a computer platform in an automotive vehicle must
+satisfy two often conflicting goals. One the one hand, the platform must provide
+enough computing capacity for real-time processing of computationally expensive
+AI workloads. On the other hand, the platform must also satisfy various
+constrains such as cost, size, weight, and power consumption limits.
+Because of these reasons, finding a right computing platform is non-trivial
+task, one that requires deep understanding of the workloads and their
+characteristics as well as the hardware platform that executes the
+workloads. It is further complicated by the heterogeneity of available
+computing hardware resources---multicore, GPU, DSP and other specialized
+hardware accelerators.
 
-- computing has been a obstacle.
-- performance, but also constraints size, weight, and power as well as
-cost senstive nature of industries including automtive.
-- many new embeded computing platforms emerged: affordable and
-powerful. raspberry pi, nvidia's embedded platforms tout their
-superiority in GPU based acceleration of the AI tasks.
-
-Our primary objective of this study are
-- to understand the necessary computing performance for applying AI
-technology based robotics systems, and 
-- what kind of computing architecture and runtime supports
+The primary objective of this study are (1) to understand the necessary
+computing performance for applying AI technology based robotics systems, and
+(2) what kind of computing architecture and runtime supports
 are most appropriate for such workload.
-
 Toward to achieve the two goals, we implement a low-cost autonomous

### 2017-10-01 paper/main.tex
Add initial platform comparison table/paragraph

@@
 \subsection{Platform Components}
+
+\begin{table*}[h]
+  \centering
+  \begin{tabular} {| l | l | l | l | l |}
+    \hline
+    \textbf{Platform} & \textbf{Picar} & \textbf{F1/10} & \textbf{NVIDIA DAVE-2}\\ \hline 
+    Car & Mini RC Car (\$10) & Traxxas 1/10th car platform (\$299.97) & TBD\\ \hline
+    Embedded system & Raspberry Pi 3 Model B (\$35.00) & NVIDIA Jetson TK1 (\$192.99) & NVIDIA DRIVE PX \\ \hline
+    CPU & Cortex A-53 quad core & Cortex A-15 quad core & 8x Cortex A-57 quad core, 8x Cortex A-53 quad core \\ \hline
+    Camera & Playstation Eye camera (\$6.99) & ZED camera (\$449.00) & 3x cameras\\ \hline
+    Power Supply & Mobile battery (\$5.99) & Energizer battery pack (\$159.00) & TBD\\
+    \hline
+  \end{tabular}
+  \caption{Comparison of components used in autonomous vehicle platforms.}
+\end{table*}
+
 We employ a small and relatively inexpensive RC car that is capable of performing basic automotive 
@@
 from the real world, thus impacting the performance of the Picar.
+
+Compared to other existing autonomous vehicle platforms, such as the F1/10 and NVIDIA DAVE-2, the Picar 
+platform is capable of performing the same operations while using more cost inexpensive components. As 
+shown by Table 1, the components selected for the Picar all cost considerably less than the other 
+platforms when focusing on the common components (camera, power supply, etc.). Including the other 
+components used in the F1/10 and DAVE-2, such as the additional sensors employed by both, the difference 
+in cost would be even greater.
 

### 2017-10-01 paper/background.tex
..

@@
 
-\subsection{Deep End-to-End Learning in Autonomous Vehicles}
+\subsection{End-to-End Deep Learning for Autonomous Vehicles}
 
@@
 intervention~\cite{Bojarski2016}. 
+\fixme{Discuss the differences between Inferencing and
+  Training. Real-time is most needed for inferencing operations.}
 
 %% \cite{Levine2016}: ``In this paper, we aim to answer
-%% the following question: does training the perception and control systems jointly end-toend
+%% the following question: does training the perception and control
+%% systems jointly end-toend 
 %% provide better performance than training each component separately?''
@@
 
-\subsection{Embedded Multicore Single-Board-Computers}
+\subsection{Embedded Computing Platforms for Real-Time Inferencing}
 
@@
 workloads. It is further complicated by the heterogeneity of available
-computing hardware resources---multicore, GPU, DSP and other specialized
-hardware accelerators.
+computing hardware resources---multicore, GPU, DSP, specialized
+hardware accelerators, etc.
 

### 2017-10-01 paper/evaluation.tex
reorg

--- 2017-09-29 paper/main.tex
@@
-%
-%  $Author: ienne $
-%  $Date: 1995/09/15 15:20:59 $
-%  $Revision: 1.4 $
-%
+\section{Evaluation}\label{sec:evaluation}
 
-% \documentclass[10pt,journal,cspaper,compsoc]{IEEEtran}   %%%tc version
-\documentclass[10pt, conference]{IEEEtran}
-%\documentclass[conference,compsoc]{IEEEtran}
-%\documentclass[10pt, conference]{IEEEtran}
-%\documentclass[times, 10pt,onecolumn]{article}
-\usepackage{amsmath, amssymb, enumerate}
-
-%%%%%%%%%%%%%%%% page control%%%%%%%%%%%%%%%%%
-%\usepackage[margin=0.75in]{geometry}
-
-%\linespread{0.991}  %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% this is really useful
-%\usepackage{cite}
-\usepackage{fancybox}
-\usepackage{amsfonts}
-%\usepackage{algorithm}
-%\usepackage[noend]{algorithmic}
-\usepackage[usenames]{color}
-%\usepackage{colortbl}
-%\usepackage[ figure, boxed, vlined]{algorithm2e}
-%\usepackage[linesnumbered,vlined]{algorithm2e}
-%\usepackage[lined,boxed]{algorithm2e}
-\usepackage{listings}
-
-\usepackage[linesnumbered,vlined]{algorithm2e}
-\usepackage{graphicx}
-\usepackage{times}
-\usepackage{psfrag}
-\usepackage{subfigure}
-\usepackage{caption}
-%\usepackage{subcaption}
-\usepackage{multirow}
-%\usepackage{setspace}
-%\usepackage{listings}
-\usepackage{epsfig}
-%\usepackage{epstopdf}
-%\usepackage[font=small,labelfont=bf]{caption}
-\usepackage{url}
-
-\usepackage{color}
-\def\fixme#1{\typeout{FIXED in page \thepage : {#1}}
-%\bgroup \color{red}{} \egroup}
-\bgroup \color{red}{[FIXME: {#1}]} \egroup}
-
-%\usepackage[pdftex]{hyperref}
-\usepackage{rotating,tabularx}
-
-\interfootnotelinepenalty=10000
-
-%% Define a new 'leo' style for the package that will use a smaller font.
-\makeatletter
-\def\url@leostyle{%
-  \@ifundefined{selectfont}{\def\UrlFont{\sf}}{\def\UrlFont{\small\ttfamily}}}
-\makeatother
-
-%\documentstyle[times,art10,twocolumn,latex8]{article}
-
-%-------------------------------------------------------------------------
-% take the % away on next line to produce the final camera-ready version
-\pagestyle{plain}
-%\thispagestyle{empty}
-%\pagestyle{empty}
-
-\newtheorem{theorem}{Theorem}
-\newtheorem{lemma}[theorem]{Lemma}
-
-%% remaining budget share, used in task stall section.
-\newcommand{\bottomrule}{\hline}
-\newcommand{\toprule}{\hline}
-\newcommand{\midrule}{\hline}
-%-------------------------------------------------------------------------
-\begin{document}
-
-\title{DeepPicar:​ ​A​ ​Low-cost​ ​Deep​ ​Neural​ ​Network-based​ ​Autonomous​ ​Car}
-\author{Author One, Author Two\\
-\{author1,author2\}@ku.edu\\
-University of Kansas, USA\\ 
-}
-
-\maketitle
-\thispagestyle{empty}
-\begin{abstract}
-
-Abstract goes here.
-
-\end{abstract}
-
-%-------------------------------------------------------------------------
-
-\input{intro}
-\input{background}
-
-\section{Platform Overview}
-
-\begin{figure}[h]
-  \centering
-  \includegraphics[width=.4\textwidth]{Picar_Picture}
-  \caption{ The Picar Platform. }
-\end{figure}
-
-\subsection{Platform Components}
-We employ a small and relatively inexpensive RC car that is capable of performing basic automotive 
-operations. However, the car we use doesn't replicate the capabilities of other autonomous vehicle 
-platforms, as it is more simplistic in nature. Notably, the RC car we use only has three possible 
-options in terms of turning: center, left, and right. As such, control of the car may be less precise at 
-times, and may negatively affect performance. The car is capable of other operations that aren't used 
-within the scope of this platform. Specifically, the car is able to drive in reverse and the driving 
-speed can be changed. In our experiments, we have the RC car drive forward, and at a constant speed.
-
-Our platform also comes equipped with a camera that is used for both recording training videos and 
-providing input frames to the model whenever the Picar is self-driving. We chose the Playstation Eye 
-Camera due to its ability to reach and maintain higher fps levels while also remaining relatively 
-low-cost. One concern, however, is the affect of camera latency on the self-driving performance of our 
-platform. While humans are able to see environmental changes in real-time, the same can't be said for 
-cameras since there is a delay between when frame capture by the camera, and when it is read by the 
-Raspberry Pi 3. As a result, it is possible for the model to be given an input frame that is different 
-from the real world, thus impacting the performance of the Picar.
-
-\subsection{Model Training}
-In order to operate the Picar as an autonomous vehicle, we utilize the DeepTesla library, which is 
-capable of training a deep neural network (DNN) with end-to-end learning, that could then be used by the 
-Picar for autonomous driving. Please note that it is more efficient to execute the actual model training 
-process on a different system with an NVIDIA GPU, rather than on the Raspberry Pi 3 itself. This is 
-because DeepTesla trains the DNN with the Tensorflow library, which greatly benefits from the use of 
-gpu-enabled operations. For comparison, training a model on the Pi took approximately 6 hours, whereas 
-the time it took on a computer with a NVIDIA GPU was only around 5 minutes!
-
-We train our model on a custom made track/lane composed of black and white duct tape, where the black 
-tape represents the bounds of the track, and the white tape marks the center of the lane. The model is 
-taught to stay close to the center of the lane for as long as possible, and to turn whenever it 
-reaches/crosses the outer lane bounds so that it remains on the track. For data, we navigate and record 
-the car going both ways across the track, and use the collected videos to train the model(s) that will 
-be used later for angle prediction while the car is self-driving.
-
-\section{Data Collection}
-...
-\section{Network Architecture}
-...
-\section{Real-Time DNN Inferencing}
-...
-\section{Online/Offline Training}
-...
-\section{Evaluation}
 In evaluating the real-time efficacy of the Raspberry Pi 3, the same methodology is utilized in all 
@@
 off or made idle so that the cpu is given time to cool down.
-\section{Related Work}
-\cite{Bojarski2016}
-
-\section{Conclusion}
-%-------------------------------------------------------------------------
-
-\bibliographystyle{plain}
-\bibliography{reference}
-\end{document}

### 2017-10-01 paper/overview-michael.tex
reorg

--- 2017-10-01 paper/main.tex
@@
-%
-%  $Author: ienne $
-%  $Date: 1995/09/15 15:20:59 $
-%  $Revision: 1.4 $
-%
-
-% \documentclass[10pt,journal,cspaper,compsoc]{IEEEtran}   %%%tc version
-\documentclass[10pt, conference]{IEEEtran}
-%\documentclass[conference,compsoc]{IEEEtran}
-%\documentclass[10pt, conference]{IEEEtran}
-%\documentclass[times, 10pt,onecolumn]{article}
-\usepackage{amsmath, amssymb, enumerate}
-
-%%%%%%%%%%%%%%%% page control%%%%%%%%%%%%%%%%%
-%\usepackage[margin=0.75in]{geometry}
-
-%\linespread{0.991}  %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% this is really useful
-%\usepackage{cite}
-\usepackage{fancybox}
-\usepackage{amsfonts}
-%\usepackage{algorithm}
-%\usepackage[noend]{algorithmic}
-\usepackage[usenames]{color}
-%\usepackage{colortbl}
-%\usepackage[ figure, boxed, vlined]{algorithm2e}
-%\usepackage[linesnumbered,vlined]{algorithm2e}
-%\usepackage[lined,boxed]{algorithm2e}
-\usepackage{listings}
-
-\usepackage[linesnumbered,vlined]{algorithm2e}
-\usepackage{graphicx}
-\usepackage{times}
-\usepackage{psfrag}
-\usepackage{subfigure}
-\usepackage{caption}
-%\usepackage{subcaption}
-\usepackage{multirow}
-%\usepackage{setspace}
-%\usepackage{listings}
-\usepackage{epsfig}
-%\usepackage{epstopdf}
-%\usepackage[font=small,labelfont=bf]{caption}
-\usepackage{url}
-
-\usepackage{color}
-\def\fixme#1{\typeout{FIXED in page \thepage : {#1}}
-%\bgroup \color{red}{} \egroup}
-\bgroup \color{red}{[FIXME: {#1}]} \egroup}
-
-%\usepackage[pdftex]{hyperref}
-\usepackage{rotating,tabularx}
-
-\interfootnotelinepenalty=10000
-
-%% Define a new 'leo' style for the package that will use a smaller font.
-\makeatletter
-\def\url@leostyle{%
-  \@ifundefined{selectfont}{\def\UrlFont{\sf}}{\def\UrlFont{\small\ttfamily}}}
-\makeatother
-
-%\documentstyle[times,art10,twocolumn,latex8]{article}
-
-%-------------------------------------------------------------------------
-% take the % away on next line to produce the final camera-ready version
-\pagestyle{plain}
-%\thispagestyle{empty}
-%\pagestyle{empty}
-
-\newtheorem{theorem}{Theorem}
-\newtheorem{lemma}[theorem]{Lemma}
-
-%% remaining budget share, used in task stall section.
-\newcommand{\bottomrule}{\hline}
-\newcommand{\toprule}{\hline}
-\newcommand{\midrule}{\hline}
-%-------------------------------------------------------------------------
-\begin{document}
-
-\title{DeepPicar:​ ​A​ ​Low-cost​ ​Deep​ ​Neural​ ​Network-based​ ​Autonomous​ ​Car}
-\author{Author One, Author Two\\
-\{author1,author2\}@ku.edu\\
-University of Kansas, USA\\ 
-}
-
-\maketitle
-\thispagestyle{empty}
-\begin{abstract}
-
-Abstract goes here.
-
-\end{abstract}
-
-%-------------------------------------------------------------------------
-
-\input{intro}
-\input{background}
-
 \section{Platform Overview}
@@
 be used later for angle prediction while the car is self-driving.
-
-\section{Data Collection}
-...
-\section{Network Architecture}
-...
-\section{Real-Time DNN Inferencing}
-...
-\section{Online/Offline Training}
-...
-\section{Evaluation}
-In evaluating the real-time efficacy of the Raspberry Pi 3, the same methodology is utilized in all 
-experiments conducted. The performance of the Pi is measured over a set of 1001 video frames that are 
-each individually fed to the model. The processing time for the first frame is ignored as, due to 
-cache warmup, it is uncharacteristically high and doesn't accurately represent the Pi's capabilities. A 
-deadline of 50 ms, or 20 Hz, is used as a baseline to assess the Pi's ability to complete all 
-necessary real-time operations in a timely manner. Also, please note that frame processing times 
-would be approximately the same if the input stream was a 
-camera instead of a video.
-
-\subsection{Real-Time Operations}
-In a real-time operations, the system has to be capable of consistently executing all necessary 
-functions before their given deadlines. In the case of our platform, it has to process every given frame 
-and get the predicted angles from the model. In order to determine if the Raspberry Pi 3 is capable of 
-this, we tested the platform and recorded the time it took for the Pi to complete all real-time 
-functions for every given frame.For this experiment, all four of the Pi's cpu cores were utilized, and 
-only one model was run. As can be seen in Figure 1, the Pi was able to meet almost all of its deadlines, 
-while only missing it in a few instances.
-
-\begin{figure}[h]
-  \centering
-  \includegraphics[width=.5\textwidth]{Total_Processing_Time}
-  \caption{ Real-time performance of the Raspberry Pi 3 in processing 1000 frames.}
-\end{figure}
-
-In our platform, three main real-time tasks are performed during autonomous operation. 
-In order, those operations are: (1) capturing and reading the input frame from the designated camera 
-or video stream, (2) preprocessing the acquired frame so that it is compatible with the DNN, and (3) 
-feeding the frame to, and getting the angle prediction from, the model. However, the time it takes to 
-complete each of the operations will differ, with at least one of them being the dominating step in 
-processing each frame.
-
-\begin{table}[h]
-  \centering
-  \begin{tabular} {| l | l | l | l | l |}
-    \hline
-    \textbf{operation} & \textbf{mean} & \textbf{max} & \textbf{99pct} & \textbf{stdev} \\ \hline 
-    Frame Capture & 2.28 & 4.94 & 63.31 &  .51\\ \hline
-    Preprocessing & 3.09 & 4.6 & 3.31 & .1 \\ \hline
-    Angle Prediction & 37.3 & 51.03 & 45.48 & 2.75 \\ 
-    \hline
-  \end{tabular}
-  \caption{Real-time performance of the Raspberry Pi 3 depending on the number of cores used.}
-\end{table}
-
-In order to determine which operation(s) take the longest to execute, we measured the time it 
-took for each step to be completed. For this experiment, all four of the Pi's cpu cores were utilized, 
-and only one model was run. As is shown in Fig. 1, the angle prediction operation (3) consumes the 
-majority of the processing for each frame. Furthermore, the time it takes for the operation to 
-complete is volatile, and can range anywhere between 30 ms and 50 ms for any particular frame. On the 
-other hand, both the frame capture (1) and preprocessing (2) operations take substantially less time 
-and are relatively more consistent in their times, at 2 ms and 3 ms, respectively.
-
-\subsection{Multicore Performance}
-It may not always be the case that all four cores of the Raspberry Pi 3's Cortex A-53 CPU can be used 
-solely for the purpose of operating an autonomous vehicle. Thus, we test how the number of cores 
-utilized for real-time operations affects the Pi's overall ability to function as an autonomous 
-vehicle platform.
-
-\begin{table}[h]
-  \centering
-  \begin{tabular} {| l | l | l | l | l | l | l | l | l | l |}
-    \hline
-    \textbf{num cores} & \textbf{mean} & \textbf{max} & \textbf{99pct} & \textbf{stdev} \\ \hline 
-    1 & 61.96 & 66 & 63.31 &  .51\\ \hline
-    2 & 50.49 & 71.55 & 70.03 & 4.18 \\ \hline
-    3 & 48.11 & 72.22 & 58.45 & 2.8 \\ \hline
-    4 & 42.67 & 56.37 & 50.70 & 2.8 \\
-    \hline
-  \end{tabular}
-  \caption{Real-time performance of the Raspberry Pi 3 depending on the number of cores used.}
-\end{table}
-
-As is depicted in Table 1, the Raspberry Pi performed better, on average, when it utilized more 
-cores. With 4 cores, the Pi was able to meet the vast majority of its 50 ms deadlines, doing so in 
-almost 99\% of the processed frames. The Pi had the worst average performance when using only 1 core, 
-as it was unable to meet any of the 50 ms deadlines (its fastest processing time was just under 60 
-ms). Another observation is that the difference between using 2 cores and 3 cores is relatively 
-small. On average, using 3 cores only performed better by 2 ms, so the addition of one core in that 
-specific case offers relatively little improvement. However, please note that the average time when 
-using 3 cores was below the 50 ms deadline, while the average time using 2 cores was slightly above it. 
-One important observation that can be made is that of consistency, as using only 1 core had a much 
-steadier performance. As a result, the use of multiple cores is very beneficial in terms of reducing 
-the time it takes to complete real-time operations, but may ultimately result in processing times 
-that are more volatile.
-
-\subsection{Multimodel Performance}
-In a real world scenario, it is highly probable that multiple models will need to be used 
-simultaneously in order to perform various important tasks (angle prediction, object detection, 
-etc.)\cite{}. As such, we also tested the capability of the Raspberry Pi to run multple models at the 
-same time, and measure whether all models are able to meet their respective deadlines on a consistent 
-basis. Specifically, the Pi is tested in the cases of running 2 and 4 models simultaneously. For each 
-case, all models are allocated an equal number of cores, with 2 models having 2 cores each, and 4 
-models having 1 core each.
-
-\begin{table*}
-  \begin{tabular} {| l | l | l | l | l | l | l | l | l |}
-  \hline
-  \textbf{num models} & \textbf{cores} & \textbf{mean} & \textbf{L1 refs} & \textbf{L1 
-    misses} & \textbf{L1 miss \%} & textbf{L2 refs} & \textbf{L2 misses} & \textbf{L2 miss \%} \\ \hline
-  1 & 0,1 & 51.35 & 3.04E+10 & 4.78E+08 & 1.58 & 3.31E+09 & 3.68E+08 & 11.12\\ \hline
-  2 & 0,1 & 58.03 & 3.04E+10 & 4.91E+08 & 1.61 & 3.91E+09 & 4.26E+08 & 10.88 \\ \hline
-  2 & 2,3 & 56.4 & 3.04E+10 & 4.80E+08 & 1.58 & 3.88E+09 & 4.21E+08 & 10.87 \\ \hline
-  \end{tabular}
-  \caption{Real-time performance of the Raspberry Pi 3 when 2 models are running simultaneously.}
-\end{table*}
-
-Ideally, each model run would replicate a single model run. However, Table 2 shows that such was not 
-the case. In the tests where 2 models ran simultaneously, both of the models showed average time 
-increases of around 5-7 ms, around 10\%, when compared to a baseline of 1 model running on two cores. 
-This change, however, did not seem to be caused by any form of cache interference as both L1 and L2 
-cache misses remained constant, regardless of the number of models running. The number of L1 misses 
-being close to 1.6\% of all references and the number of L2 misses being about 11\% of all 
-references.
-
-\begin{table*}
-  \begin{tabular} {| l | l | l | l | l | l | l | l | l |}
-  \hline
-  \textbf{num models} & \textbf{core} & \textbf{mean} & \textbf{L1 refs} & \textbf{L1 
-    misses} & \textbf{L1 miss \%} & textbf{L2 refs} & \textbf{L2 misses} & \textbf{L2 miss \%} \\ \hline
-    1 & 0 &  62.48 & 2.78E+10 & 4.36E+08 & 1.57 & 2.83E+09 & 3.59E+08 & 12.68  \\ \hline
-    4 & 0 & 77.9 & 2.79E+10 & 4.53E+08 & 1.63 & 3.36E+09 & 4.43E+08 & 13.19 \\ \hline
-    4 & 1 & 78.89 & 2.79E+10 & 4.64E+08 & 1.67 & 3.42E+09 & 4.38E+08 & 12.82 \\ \hline
-    4 & 2 & 77.81 & 2.78E+10 & 4.45E+08 & 1.6 & 3.45E+09 & 4.41E+08 & 12.77 \\ \hline
-    4 & 3 & 77.87 & 2.79E+10 & 4.45E+08 & 1.60 & 3.41E+09 & 4.39E+08 & 12.88 \\ \hline
-  \end{tabular}
-  \caption{Real-time performance of the Raspberry Pi when 4 models are running simultaneously.}
-\end{table*}
-
-The difference was even greater in the case of 4 models running concurrently, as each one displayed 
-an average time increase of approximately 15 ms, around 30\%, when compared to a single model 
-running on 1 core. Once again, though, cache interference was not a factor as the number of L1 cache 
-misses remained 1.6\% of all references, and the number of L2 cache misses stayed at 13\% of all 
-references.
-
-\subsection{Performance Requirements}
-In the utilization of the Raspberry Pi 3 in our platform, there are a few factors that need to be 
-considered and/or enforced in order to guarantee that the Pi is able to consistently perform at a 
-desired level. Specifically, these issues all have the potential to negatively affect the cpu clock 
-speed/frequency, which would result in decreased performance. While, in the above experiments, the cpu 
-operated at a preferred clock speed of 1.2 GHz, it is entirely possible for the cpu to operate at a 
-lower frequency if the following problems are not taken into account.
-
-The most notable issue that can affect the cpu clock speed is that of the power supplied to the 
-Raspberry Pi. In essence, it is necessary that the Pi be supplied with 2 Amps, as any less could 
-hinder the Pi's ability to maintain a 1.2 GHz frequency. In experiments done with a power supply that 
-only provided 1 Amp, the Pi was unable to sustain a 1.2 GHz clock speed and, instead, fluctuated 
-between operating at 600 MHz and 1.2 GHz. As a result, it is necessary, or at least highly 
-recommended, that the power supply used for the Raspberry Pi 3 be capable of outputting 2 Amps, 
-otherwise optimal performance isn't guaranteed.
-
-Another factor that can affect clock speed is that of the cpu's temperature. Some model operations can 
-be computationally intensive, thus it is possible for the temperature of the cpu to become relatively 
-high. This can be especially problematic in situations where multiple models are running 
-simultaneously on the Pi. As a consequence, thermal throttling may be used to decrease the clock 
-speed so that the cpu temperature stays at a safe level. As such, the Raspberry Pi may not be suited 
-for prolonged use, especially in cases where the workload is relatively larger, like running multiple 
-models. Rather, the Pi seems to be better suited for running in set periods, after which it is turned 
-off or made idle so that the cpu is given time to cool down.
-\section{Related Work}
-\cite{Bojarski2016}
-
-\section{Conclusion}
-%-------------------------------------------------------------------------
-
-\bibliographystyle{plain}
-\bibliography{reference}
-\end{document}

### 2017-10-01 paper/overview.tex
reorg

\section{DeepPicar Overview}

\begin{figure}[h]
  \centering
  \includegraphics[width=.4\textwidth]{Picar_Picture}
  \caption{DeepPicar platform.}
  \label{fig:overview}
\end{figure}

In this section, we provide an overview of our DeepPicar platform. In
developing DeepPicar, one of our primary goals is to replicate the
NVIDIA's DAVE-2 system on a smaller scale with using a low cost
multicore platform, Raspberry Pi 3. Because Pi 3's computing
performance is much lower than that of the DRIVE PX platform, used in
DAVE-2, we are interested in if and how we can process
computationally expensive neural network operations in
real-time. Specifically, inferencing (forward pass processing)
operation must be completed within each control period
duration---e.g., WCET of 50ms for 20Hz control frequency---locally on
the Pi 3 platform, although training of the network (back-propagation
for weight updates) can be done offline and remotely using a desktop
computer.

Figure~\ref{fig:overview} shows the DeepPicar, which is comprised of a
set of inexpensive components: a Raspberry Pi 3 Single Board Computer
(SBC), a Pololu DRV8835 motor driver, a Playstation Eye Webcam, a
battery, and a 1:24 scale RC car. Table~\ref{tbl:carbom} shows
estimated cost of the system.

\begin{table}[h]
  \centering
  \begin{tabular}{|l|l|}
    \hline
    Item                    & cost (\$) \\
    \hline
    Raspberry Pi 3 Model B  & 35 \\
    New Bright 1:24 scale RC car       & 10 \\
    Playstation Eye camera  &  7 \\
    Pololu DRV8835 motor hat&  8 \\
    External battery pack   & 10 \\
    \hline
    Total                   & 70 \\
    \hline
  \end{tabular}
  \caption{DeepPicar's bill of materials (BOM)}
  \label{tbl:carbom}
\end{table}

For the neural network architecture, we adopt a tensorflow version of
NVIDIA's DAVE-2 convolutional neural network (CNN), published by
Fridman at  MIT~\footnote{https://github.com/lexfridman/deeptesla}. As
in NVIDIA's DAVE-2, the CNN takes raw color image (200x66 RGB pixels)
as input and produce a single steering angle value as
output. Figure~\ref{fig:architecture} shows the network architecture, which
is comprised of 9 layers, 250K parameters, and about 27 million
connections.

\begin{figure}[h]
  \centering
  \includegraphics[width=.4\textwidth]{figs/architecture}
  \caption{DeepPicar's neural network architecture: 9 layers (5
    convolutional, 4 fully-connected layers), 250K
    parameters. Adopted from ~\cite{Bojarski2016}.}
  \label{fig:architecture}
\end{figure}

% data collection and training.
To collect the training data, a human pilot manually drives the RC car
on a small track we created (Figure~\ref{fig:track}) to record
timestamped video and contol commands. The stored data is then copied 
to a desktop computer, which equips a NVIDIA GTX 1060 GPU, where we
train the network to accerate training speed. 
For comparison, training the network on the Raspbeery Pi 3 takes
approximately 6 hours, whereas it takes only about 5 minutes on the
desktop computer using the GTX 1060 GPU.

\begin{figure}[h]
  \centering
  \includegraphics[width=.4\textwidth]{figs/track}
  \caption{A track for training/testing.\fixme{replace this}}
  \label{fig:track}
\end{figure}

% inferencing on pi3
Once the network is trained on the desktop computer, the trained model
is copied back to the Raspberry Pi computer. The network is then used
by the car's main controller, which feeds an image frame the web
camera as input to the network. The produced steering angle output is
then converted as the PWM value of the steering motor of the
car. Figure~\ref{fig:controlloop} shows simplified psuedo code of the
controller's main loop.

\begin{figure}[t]
   \lstset{language=python,
           basicstyle=\ttfamily\small,
           keywordstyle=\color{blue}\ttfamily,
           stringstyle=\color{red}\ttfamily,
           commentstyle=\color{green}\ttfamily
          }  
  \lstinputlisting[language=python]{control.py}
  \caption{Control loop}
  \label{fig:controlloop}
\end{figure}

%% Although the steering angle output of the network is a continuous
%% real value,

%% We employ a small and relatively inexpensive RC car that is capable of
%% performing basic automotive operations. However, the car we use
%% doesn't replicate the capabilities of other autonomous vehicle
%% platforms, as it is more simplistic in nature. Notably, the RC car we
%% use only has three possible options in terms of turning: center, left,
%% and right. As such, control of the car may be less precise at times,
%% and may negatively affect performance. The car is capable of other
%% operations that aren't used within the scope of this
%% platform. Specifically, the car is able to drive in reverse and the
%% driving speed can be changed. In our experiments, we have the RC car
%% drive forward, and at a constant speed.

%% Our platform also comes equipped with a camera that is used for both
%% recording training videos and providing input frames to the model
%% whenever the Picar is self-driving. We chose the Playstation Eye
%% Camera due to its ability to reach and maintain higher fps levels
%% while also remaining relatively low-cost. One concern, however, is the
%% affect of camera latency on the self-driving performance of our
%% platform. While humans are able to see environmental changes in
%% real-time, the same can't be said for cameras since there is a delay
%% between when frame capture by the camera, and when it is read by the
%% Raspberry Pi 3. As a result, it is possible for the model to be given
%% an input frame that is different from the real world, thus impacting
%% the performance of the Picar.

%% Compared to other existing autonomous vehicle platforms, such as the
%% F1/10 and NVIDIA DAVE-2, the Picar platform is capable of performing
%% the same operations while using more cost inexpensive components. As
%% shown by Table 1, the components selected for the Picar all cost
%% considerably less than the other  platforms when focusing on the
%% common components (camera, power supply, etc.). Including the other
%% components used in the F1/10 and DAVE-2, such as the additional
%% sensors employed by both, the difference  in cost would be even
%% greater.

%% \subsection{Model Training}
%% In order to operate the Picar as an autonomous vehicle, we utilize the
%% DeepTesla library, which is capable of training a deep neural network
%% (DNN) with end-to-end learning, that could then be used by the  Picar
%% for autonomous driving. Please note that it is more efficient to
%% execute the actual model training  process on a different system with
%% an NVIDIA GPU, rather than on the Raspberry Pi 3 itself. This is
%% because DeepTesla trains the DNN with the Tensorflow library, which
%% greatly benefits from the use of  gpu-enabled operations. For
%% comparison, training a model on the Pi took approximately 6 hours,
%% whereas  the time it took on a computer with a NVIDIA GPU was only
%% around 5 minutes!

%% We train our model on a custom made track/lane composed of black and
%% white duct tape, where the black  tape represents the bounds of the
%% track, and the white tape marks the center of the lane. The model is
%% taught to stay close to the center of the lane for as long as
%% possible, and to turn whenever it  reaches/crosses the outer lane
%% bounds so that it remains on the track. For data, we navigate and
%% record  the car going both ways across the track, and use the
%% collected videos to train the model(s) that will  be used later for
%% angle prediction while the car is self-driving.

%% \begin{table*}[h]
%%   \centering
%%   \begin{tabular} {| l | l | l | l | l |}
%%     \hline
%%     \textbf{Platform} & \textbf{Picar} & \textbf{F1/10} & \textbf{NVIDIA DAVE-2}\\ \hline  
%%     Car & Mini RC Car (\$10) & Traxxas 1/10th car platform (\$299.97) & TBD\\ \hline
%%     Embedded system & Raspberry Pi 3 Model B (\$35.00) & NVIDIA Jetson TK1 (\$192.99) & NVIDIA DRIVE PX \\ \hline
%%     CPU & Cortex A-53 quad core & Cortex A-15 quad core & 8x Cortex A-57 quad core, 8x Cortex A-53 quad core \\ \hline
%%     Camera & Playstation Eye camera (\$6.99) & ZED camera (\$449.00) & 3x cameras\\ \hline
%%     Power Supply & Mobile battery (\$5.99) & Energizer battery pack (\$159.00) & TBD\\
%%     \hline
%%   \end{tabular}
%%   \caption{Comparison of components used in autonomous vehicle platforms.}
%% \end{table*}

### 2017-10-01 paper/control.py

@@ -1,0 +1,13 @@
+
+while True:
+  # 1. read from the forward camera
+  frame = camera.read()
+  # 2. convert to 200x66 rgb pixels
+  frame = preprocess(frame)
+  # 3. perform inferencing operation
+  angle = DNN_inferencing(frame)
+  # 4. motor control                       
+  steering_motor_control(angle)
+  # 5. wait till next period begins                       
+  wait_till_next_period() 
+  

### 2017-10-02 paper/evaluation.tex
Add preliminary conclusion and revised evaluation
Also added redrawn Figure 1 powerpoint and image.

@@
 
-In evaluating the real-time efficacy of the Raspberry Pi 3, the same methodology is utilized in all 
-experiments conducted. The performance of the Pi is measured over a set of 1001 video frames that are 
-each individually fed to the model. The processing time for the first frame is ignored as, due to 
-cache warmup, it is uncharacteristically high and doesn't accurately represent the Pi's capabilities. A 
-deadline of 50 ms, or 20 Hz, is used as a baseline to assess the Pi's ability to complete all 
-necessary real-time operations in a timely manner. Also, please note that frame processing times 
-would be approximately the same if the input stream was a 
-camera instead of a video.
+In evaluating the real-time efficacy of the DeepPicar, the methodology is consistent across all 
+experiments. The performance of the platform is measured over a set of 1001 video frames that are 
+each individually fed to the model. The processing time for the first frame is omitted as, due to 
+cache warmup, it is uncharacteristically high and doesn't accurately represent the Picar's capabilities. 
+A WCET of 50 ms, or 20 Hz, is used as a baseline to assess the platform's ability to complete all 
+necessary real-time operations in a timely manner.
 
 \subsection{Real-Time Operations}
-In a real-time operations, the system has to be capable of consistently executing all necessary 
-functions before their given deadlines. In the case of our platform, it has to process every given frame 
-and get the predicted angles from the model. In order to determine if the Raspberry Pi 3 is capable of 
-this, we tested the platform and recorded the time it took for the Pi to complete all real-time 
-functions for every given frame.For this experiment, all four of the Pi's cpu cores were utilized, and 
-only one model was run. As can be seen in Figure 1, the Pi was able to meet almost all of its deadlines, 
-while only missing it in a few instances.
+We seek to determine if the DeepPicar is capable of consistently executing all necessary functions 
+before their given deadlines. In the case of our platform, it has to process every given frame and get 
+the predicted angles from the model within a WCET of 50 ms. In order to determine if this was achievable, 
+we tested the platform by running a single model that utilized all 4 cpu cores and measured the time it 
+took for each frame to be processed. The performance of the DeepPicar can be seen in Figure 6. We found 
+that DeepPicar was able to completely process the vast majority of the provided frames within 50 ms, and 
+was unable to do so for very few frames.
 
@@
 
-In our platform, three main real-time tasks are performed during autonomous operation. 
-In order, those operations are: (1) capturing and reading the input frame from the designated camera 
-or video stream, (2) preprocessing the acquired frame so that it is compatible with the DNN, and (3) 
-feeding the frame to, and getting the angle prediction from, the model. However, the time it takes to 
-complete each of the operations will differ, with at least one of them being the dominating step in 
-processing each frame.
+In our platform, three main real-time tasks are performed during autonomous operation. In order, these 
+operations are: (1) capturing and reading the input frame from the designated camera or video stream, 
+(2) preprocessing the acquired frame so that it is compatible with the DNN, and (3) feeding the frame 
+to, and getting the angle prediction from, the model. We aim to determine which operation(s), if any, 
+require the most time to be performed.
 
@@
     \textbf{operation} & \textbf{mean} & \textbf{max} & \textbf{99pct} & \textbf{stdev} \\ \hline 
-    Frame Capture & 2.28 & 4.94 & 63.31 &  .51\\ \hline
-    Preprocessing & 3.09 & 4.6 & 3.31 & .1 \\ \hline
-    Angle Prediction & 37.3 & 51.03 & 45.48 & 2.75 \\ 
+    Frame Capture & 2.28 & 4.94 & 4.54  & 0.52\\ \hline
+    Preprocessing & 3.09 & 4.60 & 3.31 & 0.10 \\ \hline
+    Angle Prediction & 37.30 & 51.03 & 45.48 & 2.75 \\  \hline
+    Total Time & 42.67 & 56.37 & 50.70 & 2.80 \\
     \hline
   \end{tabular}
-  \caption{Real-time performance of the Raspberry Pi 3 depending on the number of cores used.}
+  \caption{Real-time performance of DeepPicar for each autonomous vehicle operation performed.}
 \end{table}
@@
 took for each step to be completed. For this experiment, all four of the Pi's cpu cores were utilized, 
-and only one model was run. As is shown in Fig. 1, the angle prediction operation (3) consumes the 
+and only one model was run. As is shown in Table II, the angle prediction operation consumes the 
 majority of the processing for each frame. Furthermore, the time it takes for the operation to 
 complete is volatile, and can range anywhere between 30 ms and 50 ms for any particular frame. On the 
-other hand, both the frame capture (1) and preprocessing (2) operations take substantially less time 
+other hand, both the frame capture and preprocessing operations take substantially less time 
 and are relatively more consistent in their times, at 2 ms and 3 ms, respectively.
@@
     \textbf{num cores} & \textbf{mean} & \textbf{max} & \textbf{99pct} & \textbf{stdev} \\ \hline 
-    1 & 61.96 & 66 & 63.31 &  .51\\ \hline
+    1 & 61.96 & 66.00 & 63.31 & 0.51\\ \hline
     2 & 50.49 & 71.55 & 70.03 & 4.18 \\ \hline
-    3 & 48.11 & 72.22 & 58.45 & 2.8 \\ \hline
-    4 & 42.67 & 56.37 & 50.70 & 2.8 \\
+    3 & 48.11 & 72.22 & 58.45 & 2.80 \\ \hline
+    4 & 42.67 & 56.37 & 50.70 & 2.80 \\
     \hline
   \end{tabular}
-  \caption{Real-time performance of the Raspberry Pi 3 depending on the number of cores used.}
+  \caption{Real-time performance of DeepPicar depending on the number of cores used.}
 \end{table}
 
-As is depicted in Table 1, the Raspberry Pi performed better, on average, when it utilized more 
-cores. With 4 cores, the Pi was able to meet the vast majority of its 50 ms deadlines, doing so in 
-almost 99\% of the processed frames. The Pi had the worst average performance when using only 1 core, 
-as it was unable to meet any of the 50 ms deadlines (its fastest processing time was just under 60 
-ms). Another observation is that the difference between using 2 cores and 3 cores is relatively 
-small. On average, using 3 cores only performed better by 2 ms, so the addition of one core in that 
-specific case offers relatively little improvement. However, please note that the average time when 
-using 3 cores was below the 50 ms deadline, while the average time using 2 cores was slightly above it. 
-One important observation that can be made is that of consistency, as using only 1 core had a much 
-steadier performance. As a result, the use of multiple cores is very beneficial in terms of reducing 
-the time it takes to complete real-time operations, but may ultimately result in processing times 
-that are more volatile.
+The performance of the DeepPicar is summarized in Table III. The DeepPicar performed better, on average, 
+when it utilized more cores. With 4 cores, it was able to meet the vast majority of its deadlines, doing 
+so in almost 99\% of the input frames. The platform performed the worst when using only 1 core, 
+as it was unable to meet any of its deadlines.  On average, using 3 cores instead of 2 only had a 
+performance increase of approximately 2 ms, so the addition of one core in that specific case offers 
+relatively little improvement. One important observation is that the performance was more consistent 
+when only 1 core was used. As a result, the use of multiple cores is very beneficial in terms of 
+reducing the time it takes to complete inference operations, but may result in times that are more 
+volatile.
 
 \subsection{Multimodel Performance}
-In a real world scenario, it is highly probable that multiple models will need to be used 
-simultaneously in order to perform various important tasks (angle prediction, object detection, 
-etc.)\cite{}. As such, we also tested the capability of the Raspberry Pi to run multple models at the 
-same time, and measure whether all models are able to meet their respective deadlines on a consistent 
-basis. Specifically, the Pi is tested in the cases of running 2 and 4 models simultaneously. For each 
-case, all models are allocated an equal number of cores, with 2 models having 2 cores each, and 4 
-models having 1 core each.
+We also test the capability of the DeepPicar to run multiple models at the same time, and whether each 
+model can successfully perform within the given deadline. Specifically, the platform is tested in the 
+cases of running 2 and 4 models simultaneously. For each case, all models are allocated an equal number 
+of cores. That is, 2 models are given 2 cores each and 4 models are given 1 core each.
 
 \begin{table*}
+  \centering
   \begin{tabular} {| l | l | l | l | l | l | l | l | l |}
@@
   \textbf{num models} & \textbf{cores} & \textbf{mean} & \textbf{L1 refs} & \textbf{L1 
-    misses} & \textbf{L1 miss \%} & textbf{L2 refs} & \textbf{L2 misses} & \textbf{L2 miss \%} \\ \hline
+    misses} & \textbf{L1 miss \%} & \textbf{L2 refs} & \textbf{L2 misses} & \textbf{L2 miss \%} \\ \hline
   1 & 0,1 & 51.35 & 3.04E+10 & 4.78E+08 & 1.58 & 3.31E+09 & 3.68E+08 & 11.12\\ \hline
   2 & 0,1 & 58.03 & 3.04E+10 & 4.91E+08 & 1.61 & 3.91E+09 & 4.26E+08 & 10.88 \\ \hline
-  2 & 2,3 & 56.4 & 3.04E+10 & 4.80E+08 & 1.58 & 3.88E+09 & 4.21E+08 & 10.87 \\ \hline
+  2 & 2,3 & 56.40 & 3.04E+10 & 4.80E+08 & 1.58 & 3.88E+09 & 4.21E+08 & 10.87 \\ \hline
   \end{tabular}
-  \caption{Real-time performance of the Raspberry Pi 3 when 2 models are running simultaneously.}
+  \caption{Real-time performance of the DeepPicar when 2 models are running simultaneously.}
 \end{table*}
 
-Ideally, each model run would replicate a single model run. However, Table 2 shows that such was not 
-the case. In the tests where 2 models ran simultaneously, both of the models showed average time 
-increases of around 5-7 ms, around 10\%, when compared to a baseline of 1 model running on two cores. 
-This change, however, did not seem to be caused by any form of cache interference as both L1 and L2 
-cache misses remained constant, regardless of the number of models running. The number of L1 misses 
-being close to 1.6\% of all references and the number of L2 misses being about 11\% of all 
-references.
+The results for the 2 model test are outlined in Table IV, and the 4 model test is summarized in Table 
+V. In the the 2 model test, both of the models showed average inference time increases of around 5-7 
+ms, $\sim$10\%, when compared to a baseline of 1 model running on two cores. The difference was even 
+greater in the 4 model test, as each model displayed an average inference time increase of 
+approximately 15 ms, $\sim$30\%, when compared to a single model running on 1 core.
+
+The increase in inference times, however, was not the result of increased cache misses due to 
+additional accesses. For all models, the number of L1 misses was always $\sim$1.6\% of all 
+references. In the 2 model test, L2 cache misses remained at $\sim$11\% of all references, and each 
+model in the 4 model test missed $\sim$13\% of all L2 references.
 
 \begin{table*}
+  \centering
   \begin{tabular} {| l | l | l | l | l | l | l | l | l |}
@@
   \textbf{num models} & \textbf{core} & \textbf{mean} & \textbf{L1 refs} & \textbf{L1 
-    misses} & \textbf{L1 miss \%} & textbf{L2 refs} & \textbf{L2 misses} & \textbf{L2 miss \%} \\ \hline
+    misses} & \textbf{L1 miss \%} & \textbf{L2 refs} & \textbf{L2 misses} & \textbf{L2 miss \%} \\ \hline
     1 & 0 & 62.48 & 2.78E+10 & 4.36E+08 & 1.57 & 2.83E+09 & 3.59E+08 & 12.68 \\ \hline
-    4 & 0 & 77.9 & 2.79E+10 & 4.53E+08 & 1.63 & 3.36E+09 & 4.43E+08 & 13.19 \\ \hline
+    4 & 0 & 77.90 & 2.79E+10 & 4.53E+08 & 1.63 & 3.36E+09 & 4.43E+08 & 13.19 \\ \hline
     4 & 1 & 78.89 & 2.79E+10 & 4.64E+08 & 1.67 & 3.42E+09 & 4.38E+08 & 12.82 \\ \hline
-    4 & 2 & 77.81 & 2.78E+10 & 4.45E+08 & 1.6 & 3.45E+09 & 4.41E+08 & 12.77 \\ \hline
+    4 & 2 & 77.81 & 2.78E+10 & 4.45E+08 & 1.60 & 3.45E+09 & 4.41E+08 & 12.77 \\ \hline
     4 & 3 & 77.87 & 2.79E+10 & 4.45E+08 & 1.60 & 3.41E+09 & 4.39E+08 & 12.88 \\ \hline
   \end{tabular}
-  \caption{Real-time performance of the Raspberry Pi when 4 models are running simultaneously.}
+  \caption{Real-time performance of the DeepPicar when 4 models are running simultaneously.}
 \end{table*}
 
-The difference was even greater in the case of 4 models running concurrently, as each one displayed 
-an average time increase of approximately 15 ms, around 30\%, when compared to a single model 
-running on 1 core. Once again, though, cache interference was not a factor as the number of L1 cache 
-misses remained 1.6\% of all references, and the number of L2 cache misses stayed at 13\% of all 
-references.
+In all multimodel tests, it was found that a greater number of models run simultaneously resulted in 
+interference that led to a noticable increase in the average inferencing time that wasn't 
+attributable to a change in cache misses. Instead, the overall time increases can most likely be 
+traced to an increase in cache latency. Even if the models had a consistent number of cache hits, it 
+is highly probable that the contention for the cache increased the access time for each model, 
+consequently increasing the time it took for the models to execute their operations.
+
+In terms of real-time performance under a WCET of 50 ms, the DeepPicar was unsatisfactory in all 
+multimodel tests. On average, the 2 model test saw models miss their deadline by 6-8 ms, and the 4 
+model test was worse as each model went over their deadlines by 27-29 ms. As a result, it can be 
+ascertained that, if a WCET of 50 ms is required, DeepPicar would not be able to perform as necessary.
+
+\subsection{Summary}
+We found that the DeepPicar is capable of successfully completing all necessary inferencing 
+operations within a given deadline of 50 ms. Between these operations, it was discovered that the 
+angle prediction time of the model was the dominating step in the processing time of a single frame. 
+The other operations were found to execute in very little time, with either one taking 5 ms at most.
+
+When running a single model, DeepPicar performed the best when it used all 4 cpu cores, as it was 
+able to process a frame in 43 ms on average. This feat can still be accomplished when running a 
+single model on 3 cpu cores, and is also possible when using 2 cores. The only time in which the Pi 
+can't run a single model under 50 ms is when only 1 core is used. This was also the case when 
+multiple models were run concurrently as no models were able to consistently complete inference 
+operations in under 50 ms.
+
+However, the DeepPicar was shown to be capable of handling other potential WCETs. If given a larger 
+deadline, the capabilities of DeepPicar as an autonomous vehicle platform would be greatly increased. 
+For example, if tested with a deadline of 66.67 ms, or 15 fps, the platform would have passed several 
+of the experiments performed. In actuality, the DeepPicar would have only been unsatisfactory in the 
+4 model test where it would have missed the deadlines by $\sim$10 ms.
 
@@
 Raspberry Pi. In essence, it is necessary that the Pi be supplied with 2 Amps, as any less could 
-hinder the Pi's ability to maintain a 1.2 GHz frequency. In experiments done with a power supply that 
+hinder the Pi's ability to maintain a 1.2 GHz frequency. In experiments conducted with a power supply that 
 only provided 1 Amp, the Pi was unable to sustain a 1.2 GHz clock speed and, instead, fluctuated 
@@
 high. This can be especially problematic in situations where multiple models are running 
-simultaneously on the Pi. As a consequence, thermal throttling may be used to decrease the clock 
-speed so that the cpu temperature stays at a safe level. As such, the Raspberry Pi may not be suited 
-for prolonged use, especially in cases where the workload is relatively larger, like running multiple 
+simultaneously on the Pi. Consequently, thermal throttling may be used to decrease the clock 
+speed so that the cpu temperature stays at a safe level. Thus the Raspberry Pi may not be suited 
+for prolonged use, especially in cases where the workload is relatively larger, such as running multiple 
 models. Rather, the Pi seems to be better suited for running in set periods, after which it is turned 
-off or made idle so that the cpu is given time to cool down.
+off or made idle so that the cpu is allowed time to cool down.

### 2017-10-02 paper/overview.tex
tbd
Change images for platform and track
Add evaluation subsection for benchmark tests
Also add log for training a model on Raspberry Pi 3 that took over 4
hours to finish.

@@
   \centering
-  \includegraphics[width=.4\textwidth]{Picar_Picture}
+  \includegraphics[width=.4\textwidth]{figs/DeepPicar_platform}
   \caption{DeepPicar platform.}
@@
 For comparison, training the network on the Raspbeery Pi 3 takes
-approximately 6 hours, whereas it takes only about 5 minutes on the
+approximately 4 hours, whereas it takes only about 5 minutes on the
 desktop computer using the GTX 1060 GPU.
@@
   \centering
-  \includegraphics[width=.4\textwidth]{figs/track}
+  \includegraphics[width=.4\textwidth, angle=270]{figs/track_new}
   \caption{A track for training/testing.\fixme{replace this}}
@@
 \end{figure}
+
+\fixme{Discuss the limitation of the RC car platform: discrete
+  control, web camera latency}
 

### 2017-10-02 benchmark_timings.sh
Time test script update

@@ -1,0 +1,57 @@
+timestamp=$(date +%y-%m-%d_%H:%M:%S)
+
+echo "Which board is this test being run on?  Input a string and press [ENTER]"
+read board
+
+identification=$board
+identification+="_"
+identification+=$timestamp
+
+mkdir -p logs/test-model-and-benchmark/$identification
+echo "This test will take approximately 45 minutes to complete"
+echo "The output files can be found in: logs/test-model/$identification"
+
+echo 2-parallel-single-core
+./bandwidth -a write -m 16384 -t 10000 -c 1 &
+perf stat -o logs/test-model-and-benchmark/$identification/1-sc-bench_1-sc-model_cpu0_perf.txt taskset -c 0 python test-model.py 1 > logs/test-model-and-benchmark/$identification/1-sc-bench_1-sc-model_cpu0.txt
+jobs -p | xargs kill
+echo end 2-parallel-single-core
+sleep 300
+
+echo 3-parallel-single-core
+./bandwidth -a write -m 16384 -t 10000 -c 1 &
+./bandwidth -a write -m 16384 -t 10000 -c 2 &
+perf stat -o logs/test-model-and-benchmark/$identification/2-sc-bench_1-sc-model_cpu0_perf.txt taskset -c 0 python test-model.py 1 > logs/test-model-and-benchmark/$identification/2-sc-bench_1-sc-model_cpu0.txt
+jobs -p | xargs kill
+echo end 3-parallel-single-core
+sleep 300
+
+echo 4-parallel-single-core
+./bandwidth -a write -m 16384 -t 10000 -c 1 &
+./bandwidth -a write -m 16384 -t 10000 -c 2 &
+./bandwidth -a write -m 16384 -t 10000 -c 3 &
+perf stat -o logs/test-model-and-benchmark/$identification/3-sc-bench_1-sc-model_cpu0_perf.txt taskset -c 0 python test-model.py 1 > logs/test-model-and-benchmark/$identification/3-sc-bench_1-sc-model_cpu0.txt
+echo end 4-parallel-single-core
+sleep 300
+
+echo 1-single-core_1-dual-core
+./bandwidth -a write -m 16384 -t 10000 -c 2 &
[…]

### 2017-10-02 multicore_timings.sh

@@ -1,0 +1,68 @@
+timestamp=$(date +%y-%m-%d_%H:%M:%S)
+
+echo "Which board is this test being run on?  Input a string and press [ENTER]"
+read board
+
+identification=$board
+identification+="_"
+identification+=$timestamp
+
+mkdir -p logs/test-model/test-model/$identification
+echo "This test will take approximately 45 minutes to complete"
+echo "The output files can be found in: logs/test-model/$identification"
+
+echo single-core
+perf stat -o logs/test-model/$identification/single-core_perf.txt taskset -c 0 python test-model.py 1 > logs/test-model/$identification/single-core.txt &
+wait
+echo end single-core
+sleep 300
+
+echo 2-parallel-single-core
+perf stat -o logs/test-model/$identification/2-single-core_cpu0_perf.txt taskset -c 0 python test-model.py 1 > logs/test-model/$identification/2-single-core_cpu0.txt &
+perf stat -o logs/test-model/$identification/2-single-core_cpu1_perf.txt taskset -c 1 python test-model.py 1 > logs/test-model/$identification/2-single-core_cpu1.txt &
+wait
+echo end 2-parallel-single-core
+sleep 300
+
+echo 3-parallel-single-core
+perf stat -o logs/test-model/$identification/3-single-core_cpu0_perf.txt taskset -c 0 python test-model.py 1 > logs/test-model/$identification/3-single-core_cpu0.txt &
+perf stat -o logs/test-model/$identification/3-single-core_cpu1_perf.txt taskset -c 1 python test-model.py 1 > logs/test-model/$identification/3-single-core_cpu1.txt &
+perf stat -o logs/test-model/$identification/3-single-core_cpu2_perf.txt taskset -c 2 python test-model.py 1 > logs/test-model/$identification/3-single-core_cpu2.txt &
+wait
+echo end 3-parallel-single-core
+sleep 300
+
+echo 4-parallel-single-core
+perf stat -o logs/test-model/$identification/4-single-core_cpu0_perf.txt taskset -c 0 python test-model.py 1 > logs/test-model/$identification/4-single-core_cpu0.txt &
+perf stat -o logs/test-model/$identification/4-single-core_cpu1_perf.txt taskset -c 1 python test-model.py 1 > logs/test-model/$identification/4-single-core_cpu1.txt &
[…]

### 2017-10-02 paper/evaluation.tex

@@
 
+\subsection{Benchmark Performance}
+In order to determine how cache misses affects the performance of the DeepPicar, we test its ability by 
+running synthetic benchmarks concurrently with the model. In each experiment, we run a single model on a 
+number of cores, and a synthetic benchmark on the remaining idle cores. The results of each test can be 
+seen in Table VII.
+
+\begin{table}[h]
+\centering
+  \begin{tabular} {| l | l | l | l | l | l |}
+  \hline
+  \textbf{num cores} & 1 & 2 & 3 \\ \hline
+  \textbf{mean} & 562.33 & 391.98 & 98.05 \\ \hline
+  \textbf{L1 refs} & 3.16E+10 & 3.17E+10 & 3.06E+10 \\ \hline
+  \textbf{L1 misses} & 5.41E+08 & 5.35E+08 & 4.98E+08 \\ \hline
+  \textbf{L1 miss \%} & 1.71 & 1.69 & 1.62 \\ \hline
+  \textbf{L2 refs} & 3.49E+09 & 3.23E+09 & 3.21E+09 \\ \hline
+  \textbf{L2 misses} & 8.05E+08 & 7.54E+08 & 4.98E+08 \\ \hline
+  \textbf{L2 miss \%} & 23.05 & 23.37 & 15.51 \\ 
+  \hline
+  \end{tabular}
+  \caption{Real-time performance of the DeepPicar when benchmarks are run concurrently with the model.}
+\end{table}
+
+The presence of the benchmark(s) had a noticable effect on the DeepPicar, with all tests showing 
+increases in inferencing operations. The least change happened when the model was run on 3 cores, and a 
+single benchmark was run on the final core. Even with a slight increase in the number of L2 cache 
+misses, inferencing was able to be completed in under 100 ms. The other tests, however, showed dramatic 
+changes when the model was run on 1 and 2 cores. In both instances, L2 misses rise to $\sim$23\% of all 
+references. The additional benchmark in the 1 core test was even more detrimental as the average 
+inferencing time was at least 150 ms longer than when the model used 2 cores.
+
+With the introduction of synthetic benchmarks, the DeepPicar failed to comply with the 50 ms WCET across 
+all tests run. Even in the best case, inferencing operations take twice as long as the deadline to 
+complete. As a result, if benchmarks were required to run during autonomous operation, the DeepPicar 
+wouldn't be capable of meeting its deadlines.
+
 \subsection{Summary}
@@
 of the experiments performed. In actuality, the DeepPicar would have only been unsatisfactory in the 
-4 model test where it would have missed the deadlines by $\sim$10 ms.
+4 model test where it would have missed the deadlines by $\sim$10 ms, and the benchmark tests, where 
+it would still miss by $\sim$33 ms at best.
 

### 2017-10-02 paper/intro.tex

@@
 be used in a car poses several additional challenges. First, many AI
-workloads in vehicles are computationaly demanding and have strict
+workloads in vehicles are computationally demanding and have strict
 real-time requirements. For example, latency in a vision-based object
@@
 NVIDIA's Tegra TX1 platform.
+
 %% For example, while today's self-driving car
@@
 
-% related work and remaing problems
+% related work and remaining problems
 To understand what kind of computing hardware is needed for AI
@@
 which was developed for a real self-driving car, NVIDIA
-DAVE-2~\cite{Bojarski2016}, and use the same moethodology on a
+DAVE-2~\cite{Bojarski2016}, and use the same methodology on a
 smaller, safer and less costly setup. In developing the testbed, our
 goals are (1) to analyze real-time issues in DNN based end-to-end
-control; (2) to evalute real-time performance of contemporarly embedded
+control; (2) to evaluate real-time performance of contemporary embedded
 multicore platforms for the AI workloads.
@@
 The DeepPicar, however, employs state-of-the-art AI
-techonologies, including a vision-based end-to-end control system that
-utilizes a deep convolutional nueral network (CNN).
+technologies, including a vision-based end-to-end control system that
+utilizes a deep convolutional neural network (CNN).
 The network receives an image frame from a single forward
@@
 and 250 thousand parameters (weights).
-The network architecture is identifical to that of NVIDIA's DAVE-2
+The network architecture is identical to that of NVIDIA's DAVE-2
 self-driving car~\cite{Bojarski2016}, which uses much more powerful
@@
 follow tracks on the ground. We collect data for
-taining and validation by manually  
+training and validation by manually
 controlling the RC car and recording the vision (from the webcam
@@
     low-cost autonomous vehicle testbed DeepPicar, which utilizes the
-    state-of-the-art aritificial intelligence techniques.
+    state-of-the-art artificial intelligence techniques.
   \item We provide an analysis and case-study of real-time issues in the
@@
   \item We systematically compare real-time computing capabilities of
-    multiple embedded computing platforms in the contex of a
-    vision-based autnomous driving.
+    multiple embedded computing platforms in the context of a
+    vision-based autonomous driving.
 \end{itemize}

### 2017-10-02 paper/conclusion.tex

\section{Conclusion}\label{sec:conclusion}

We present a low cost autonomous vehicle platform that uses state-of-the art AI techniques and 
technologies. It employs a deep convolutional neural network (CNN) that is trained to effectively 
navigate the car by predicting the steering angle given an input frame. We have tested the real time 
efficacy of our platform and found that it is capable of performing inferencing operations within a WCET 
of 50 ms, or 20 fps, when a single model is run on more than one core. Further studies should be 
conducted on the interference that occurs between cores to allow for improved performance in utilizing 
multiple models for autonomous vehicles. With minimal further research, the DeepPicar can 
become a valudable, reliable and inexpensive tool in the continued research of autonomous vehicles.

### 2017-10-02 paper/main.tex
..

--- 2017-09-29 paper/main.tex
@@
 \begin{abstract}
-
 Abstract goes here.
-
 \end{abstract}
@@
 %-------------------------------------------------------------------------
-
 \input{intro}
 \input{background}
-
-\section{Platform Overview}
-
-\begin{figure}[h]
-  \centering
-  \includegraphics[width=.4\textwidth]{Picar_Picture}
-  \caption{ The Picar Platform. }
-\end{figure}
-
-\subsection{Platform Components}
-We employ a small and relatively inexpensive RC car that is capable of performing basic automotive 
-operations. However, the car we use doesn't replicate the capabilities of other autonomous vehicle 
-platforms, as it is more simplistic in nature. Notably, the RC car we use only has three possible 
-options in terms of turning: center, left, and right. As such, control of the car may be less precise at 
-times, and may negatively affect performance. The car is capable of other operations that aren't used 
-within the scope of this platform. Specifically, the car is able to drive in reverse and the driving 
-speed can be changed. In our experiments, we have the RC car drive forward, and at a constant speed.
-
-Our platform also comes equipped with a camera that is used for both recording training videos and 
-providing input frames to the model whenever the Picar is self-driving. We chose the Playstation Eye 
-Camera due to its ability to reach and maintain higher fps levels while also remaining relatively 
-low-cost. One concern, however, is the affect of camera latency on the self-driving performance of our 
-platform. While humans are able to see environmental changes in real-time, the same can't be said for 
-cameras since there is a delay between when frame capture by the camera, and when it is read by the 
-Raspberry Pi 3. As a result, it is possible for the model to be given an input frame that is different 
-from the real world, thus impacting the performance of the Picar.
-
-\subsection{Model Training}
-In order to operate the Picar as an autonomous vehicle, we utilize the DeepTesla library, which is 
-capable of training a deep neural network (DNN) with end-to-end learning, that could then be used by the 
-Picar for autonomous driving. Please note that it is more efficient to execute the actual model training 
-process on a different system with an NVIDIA GPU, rather than on the Raspberry Pi 3 itself. This is 
-because DeepTesla trains the DNN with the Tensorflow library, which greatly benefits from the use of 
-gpu-enabled operations. For comparison, training a model on the Pi took approximately 6 hours, whereas 
-the time it took on a computer with a NVIDIA GPU was only around 5 minutes!
-
-We train our model on a custom made track/lane composed of black and white duct tape, where the black 
-tape represents the bounds of the track, and the white tape marks the center of the lane. The model is 
-taught to stay close to the center of the lane for as long as possible, and to turn whenever it 
-reaches/crosses the outer lane bounds so that it remains on the track. For data, we navigate and record 
-the car going both ways across the track, and use the collected videos to train the model(s) that will 
-be used later for angle prediction while the car is self-driving.
-
-\section{Data Collection}
-...
-\section{Network Architecture}
-...
-\section{Real-Time DNN Inferencing}
-...
-\section{Online/Offline Training}
-...
-\section{Evaluation}
-In evaluating the real-time efficacy of the Raspberry Pi 3, the same methodology is utilized in all 
-experiments conducted. The performance of the Pi is measured over a set of 1001 video frames that are 
-each individually fed to the model. The processing time for the first frame is ignored as, due to 
-cache warmup, it is uncharacteristically high and doesn't accurately represent the Pi's capabilities. A 
-deadline of 50 ms, or 20 Hz, is used as a baseline to assess the Pi's ability to complete all 
-necessary real-time operations in a timely manner. Also, please note that frame processing times 
-would be approximately the same if the input stream was a 
-camera instead of a video.
-
-\subsection{Real-Time Operations}
-In a real-time operations, the system has to be capable of consistently executing all necessary 
-functions before their given deadlines. In the case of our platform, it has to process every given frame 
-and get the predicted angles from the model. In order to determine if the Raspberry Pi 3 is capable of 
-this, we tested the platform and recorded the time it took for the Pi to complete all real-time 
-functions for every given frame.For this experiment, all four of the Pi's cpu cores were utilized, and 
-only one model was run. As can be seen in Figure 1, the Pi was able to meet almost all of its deadlines, 
-while only missing it in a few instances.
-
-\begin{figure}[h]
-  \centering
-  \includegraphics[width=.5\textwidth]{Total_Processing_Time}
-  \caption{ Real-time performance of the Raspberry Pi 3 in processing 1000 frames.}
-\end{figure}
-
-In our platform, three main real-time tasks are performed during autonomous operation. 
-In order, those operations are: (1) capturing and reading the input frame from the designated camera 
-or video stream, (2) preprocessing the acquired frame so that it is compatible with the DNN, and (3) 
-feeding the frame to, and getting the angle prediction from, the model. However, the time it takes to 
-complete each of the operations will differ, with at least one of them being the dominating step in 
-processing each frame.
-
-\begin{table}[h]
-  \centering
-  \begin{tabular} {| l | l | l | l | l |}
-    \hline
-    \textbf{operation} & \textbf{mean} & \textbf{max} & \textbf{99pct} & \textbf{stdev} \\ \hline 
-    Frame Capture & 2.28 & 4.94 & 63.31 &  .51\\ \hline
-    Preprocessing & 3.09 & 4.6 & 3.31 & .1 \\ \hline
-    Angle Prediction & 37.3 & 51.03 & 45.48 & 2.75 \\ 
-    \hline
-  \end{tabular}
-  \caption{Real-time performance of the Raspberry Pi 3 depending on the number of cores used.}
-\end{table}
-
-In order to determine which operation(s) take the longest to execute, we measured the time it 
-took for each step to be completed. For this experiment, all four of the Pi's cpu cores were utilized, 
-and only one model was run. As is shown in Fig. 1, the angle prediction operation (3) consumes the 
-majority of the processing for each frame. Furthermore, the time it takes for the operation to 
-complete is volatile, and can range anywhere between 30 ms and 50 ms for any particular frame. On the 
-other hand, both the frame capture (1) and preprocessing (2) operations take substantially less time 
-and are relatively more consistent in their times, at 2 ms and 3 ms, respectively.
-
-\subsection{Multicore Performance}
-It may not always be the case that all four cores of the Raspberry Pi 3's Cortex A-53 CPU can be used 
-solely for the purpose of operating an autonomous vehicle. Thus, we test how the number of cores 
-utilized for real-time operations affects the Pi's overall ability to function as an autonomous 
-vehicle platform.
-
-\begin{table}[h]
-  \centering
-  \begin{tabular} {| l | l | l | l | l | l | l | l | l | l |}
-    \hline
-    \textbf{num cores} & \textbf{mean} & \textbf{max} & \textbf{99pct} & \textbf{stdev} \\ \hline 
-    1 & 61.96 & 66 & 63.31 &  .51\\ \hline
-    2 & 50.49 & 71.55 & 70.03 & 4.18 \\ \hline
-    3 & 48.11 & 72.22 & 58.45 & 2.8 \\ \hline
-    4 & 42.67 & 56.37 & 50.70 & 2.8 \\
-    \hline
-  \end{tabular}
-  \caption{Real-time performance of the Raspberry Pi 3 depending on the number of cores used.}
-\end{table}
-
-As is depicted in Table 1, the Raspberry Pi performed better, on average, when it utilized more 
-cores. With 4 cores, the Pi was able to meet the vast majority of its 50 ms deadlines, doing so in 
-almost 99\% of the processed frames. The Pi had the worst average performance when using only 1 core, 
-as it was unable to meet any of the 50 ms deadlines (its fastest processing time was just under 60 
-ms). Another observation is that the difference between using 2 cores and 3 cores is relatively 
-small. On average, using 3 cores only performed better by 2 ms, so the addition of one core in that 
-specific case offers relatively little improvement. However, please note that the average time when 
-using 3 cores was below the 50 ms deadline, while the average time using 2 cores was slightly above it. 
-One important observation that can be made is that of consistency, as using only 1 core had a much 
-steadier performance. As a result, the use of multiple cores is very beneficial in terms of reducing 
-the time it takes to complete real-time operations, but may ultimately result in processing times 
-that are more volatile.
-
-\subsection{Multimodel Performance}
-In a real world scenario, it is highly probable that multiple models will need to be used 
-simultaneously in order to perform various important tasks (angle prediction, object detection, 
-etc.)\cite{}. As such, we also tested the capability of the Raspberry Pi to run multple models at the 
-same time, and measure whether all models are able to meet their respective deadlines on a consistent 
-basis. Specifically, the Pi is tested in the cases of running 2 and 4 models simultaneously. For each 
-case, all models are allocated an equal number of cores, with 2 models having 2 cores each, and 4 
-models having 1 core each.
-
-\begin{table*}
-  \begin{tabular} {| l | l | l | l | l | l | l | l | l |}
-  \hline
-  \textbf{num models} & \textbf{cores} & \textbf{mean} & \textbf{L1 refs} & \textbf{L1 
-    misses} & \textbf{L1 miss \%} & textbf{L2 refs} & \textbf{L2 misses} & \textbf{L2 miss \%} \\ \hline
-  1 & 0,1 & 51.35 & 3.04E+10 & 4.78E+08 & 1.58 & 3.31E+09 & 3.68E+08 & 11.12\\ \hline
-  2 & 0,1 & 58.03 & 3.04E+10 & 4.91E+08 & 1.61 & 3.91E+09 & 4.26E+08 & 10.88 \\ \hline
-  2 & 2,3 & 56.4 & 3.04E+10 & 4.80E+08 & 1.58 & 3.88E+09 & 4.21E+08 & 10.87 \\ \hline
-  \end{tabular}
-  \caption{Real-time performance of the Raspberry Pi 3 when 2 models are running simultaneously.}
-\end{table*}
-
-Ideally, each model run would replicate a single model run. However, Table 2 shows that such was not 
-the case. In the tests where 2 models ran simultaneously, both of the models showed average time 
-increases of around 5-7 ms, around 10\%, when compared to a baseline of 1 model running on two cores. 
-This change, however, did not seem to be caused by any form of cache interference as both L1 and L2 
-cache misses remained constant, regardless of the number of models running. The number of L1 misses 
-being close to 1.6\% of all references and the number of L2 misses being about 11\% of all 
-references.
-
-\begin{table*}
-  \begin{tabular} {| l | l | l | l | l | l | l | l | l |}
-  \hline
-  \textbf{num models} & \textbf{core} & \textbf{mean} & \textbf{L1 refs} & \textbf{L1 
-    misses} & \textbf{L1 miss \%} & textbf{L2 refs} & \textbf{L2 misses} & \textbf{L2 miss \%} \\ \hline
-    1 & 0 &  62.48 & 2.78E+10 & 4.36E+08 & 1.57 & 2.83E+09 & 3.59E+08 & 12.68  \\ \hline
-    4 & 0 & 77.9 & 2.79E+10 & 4.53E+08 & 1.63 & 3.36E+09 & 4.43E+08 & 13.19 \\ \hline
-    4 & 1 & 78.89 & 2.79E+10 & 4.64E+08 & 1.67 & 3.42E+09 & 4.38E+08 & 12.82 \\ \hline
-    4 & 2 & 77.81 & 2.78E+10 & 4.45E+08 & 1.6 & 3.45E+09 & 4.41E+08 & 12.77 \\ \hline
-    4 & 3 & 77.87 & 2.79E+10 & 4.45E+08 & 1.60 & 3.41E+09 & 4.39E+08 & 12.88 \\ \hline
-  \end{tabular}
-  \caption{Real-time performance of the Raspberry Pi when 4 models are running simultaneously.}
-\end{table*}
-
-The difference was even greater in the case of 4 models running concurrently, as each one displayed 
-an average time increase of approximately 15 ms, around 30\%, when compared to a single model 
-running on 1 core. Once again, though, cache interference was not a factor as the number of L1 cache 
-misses remained 1.6\% of all references, and the number of L2 cache misses stayed at 13\% of all 
-references.
-
-\subsection{Performance Requirements}
-In the utilization of the Raspberry Pi 3 in our platform, there are a few factors that need to be 
-considered and/or enforced in order to guarantee that the Pi is able to consistently perform at a 
-desired level. Specifically, these issues all have the potential to negatively affect the cpu clock 
-speed/frequency, which would result in decreased performance. While, in the above experiments, the cpu 
-operated at a preferred clock speed of 1.2 GHz, it is entirely possible for the cpu to operate at a 
-lower frequency if the following problems are not taken into account.
-
-The most notable issue that can affect the cpu clock speed is that of the power supplied to the 
-Raspberry Pi. In essence, it is necessary that the Pi be supplied with 2 Amps, as any less could 
-hinder the Pi's ability to maintain a 1.2 GHz frequency. In experiments done with a power supply that 
-only provided 1 Amp, the Pi was unable to sustain a 1.2 GHz clock speed and, instead, fluctuated 
-between operating at 600 MHz and 1.2 GHz. As a result, it is necessary, or at least highly 
-recommended, that the power supply used for the Raspberry Pi 3 be capable of outputting 2 Amps, 
-otherwise optimal performance isn't guaranteed.
-
-Another factor that can affect clock speed is that of the cpu's temperature. Some model operations can 
-be computationally intensive, thus it is possible for the temperature of the cpu to become relatively 
-high. This can be especially problematic in situations where multiple models are running 
-simultaneously on the Pi. As a consequence, thermal throttling may be used to decrease the clock 
-speed so that the cpu temperature stays at a safe level. As such, the Raspberry Pi may not be suited 
-for prolonged use, especially in cases where the workload is relatively larger, like running multiple 
-models. Rather, the Pi seems to be better suited for running in set periods, after which it is turned 
-off or made idle so that the cpu is given time to cool down.
-\section{Related Work}
-\cite{Bojarski2016}
-
-\section{Conclusion}
+\input{overview}
+\input{evaluation}
+\input{related}
+\input{conclusion}
 %-------------------------------------------------------------------------
-
 \bibliographystyle{plain}

### 2017-10-02 paper/related.tex

\section{Related Work}\label{sec:related}

\cite{Bojarski2016}

### 2017-10-02 benchmark_timings.sh
updated scripts that broke

@@ -36,7 +36,7 @@ sleep 300
 
 echo 1-single-core_1-dual-core
 ./bandwidth -a write -m 16384 -t 10000 -c 2 &
-perf stat -o logs/test-model-and-benchmark/$timestamp/1-sc-bench_1-dc-model_cpu01_perf.txt taskset -c 0,1 python test-model.py 2 > logs/test-model-and-benchmark/$timestamp/1-sc-bench_1-dc-model_cpu01.txt
+perf stat -o logs/test-model-and-benchmark/$identification/1-sc-bench_1-dc-model_cpu01_perf.txt taskset -c 0,1 python test-model.py 2 > logs/test-model-and-benchmark/$identification/1-sc-bench_1-dc-model_cpu01.txt
 jobs -p | xargs kill
 echo end 2-parallel-dual-core
 sleep 300
@@ -44,14 +44,14 @@ sleep 300
 echo 2-parallel-dual-core
 ./bandwidth -a write -m 16384 -t 10000 -c 2 &
 ./bandwidth -a write -m 16384 -t 10000 -c 3 &
-perf stat -o logs/test-model-and-benchmark/$timestamp/2-sc-bench_1-dc-model_cpu01_perf.txt taskset -c 0,1 python test-model.py 2 > logs/test-model-and-benchmark/$timestamp/2-sc-bench_1-dc-model_cpu01.txt
+perf stat -o logs/test-model-and-benchmark/$identification/2-sc-bench_1-dc-model_cpu01_perf.txt taskset -c 0,1 python test-model.py 2 > logs/test-model-and-benchmark/$identification/2-sc-bench_1-dc-model_cpu01.txt
 jobs -p | xargs kill
 echo end 2-parallel-dual-core
 sleep 300
 
 echo tri-core
 ./bandwidth -a write -m 16384 -t 10000 -c 3 &
-perf stat -o logs/test-model-and-benchmark/$timestamp/1-sc-bench_1-tc-model_cpu012_perf.txt taskset -c 0,1,2 python test-model.py 3 > logs/test-model-and-benchmark/$timestamp/1-sc-bench_1-dc-model_cpu012.txt
+perf stat -o logs/test-model-and-benchmark/$identifcation/1-sc-bench_1-tc-model_cpu012_perf.txt taskset -c 0,1,2 python test-model.py 3 > logs/test-model-and-benchmark/$identification/1-sc-bench_1-dc-model_cpu012.txt
 jobs -p | xargs kill
 echo end tri-core
 

### 2017-10-02 test-model_timings.sh

@@ -7,7 +7,7 @@ identification=$board
 identification+="_"
 identification+=$timestamp
 
-mkdir -p logs/test-model/test-model/$identification
+mkdir -p logs/test-model/$identification
 echo "This test will take approximately 45 minutes to complete"
 echo "The output files can be found in: logs/test-model/$identification"
 
@@ -48,20 +48,20 @@ echo end dual-core
 sleep 300
 
 echo 2-parallel-dual-core
-perf stat -o logs/test-model/$timestamp/2-dual-core_cpu01_perf.txt taskset -c 0,1 python test-model.py 2 > logs/test-model/$timestamp/2-dual-core_cpu01.txt &
-perf stat -o logs/test-model/$timestamp/2-dual-core_cpu23_perf.txt taskset -c 2,3 python test-model.py 2 > logs/test-model/$timestamp/2-dual-core_cpu23.txt &
+perf stat -o logs/test-model/$identification/2-dual-core_cpu01_perf.txt taskset -c 0,1 python test-model.py 2 > logs/test-model/$identification/2-dual-core_cpu01.txt &
+perf stat -o logs/test-model/$identification/2-dual-core_cpu23_perf.txt taskset -c 2,3 python test-model.py 2 > logs/test-model/$identification/2-dual-core_cpu23.txt &
 wait
 echo end 2-parallel-dual-core
 sleep 300
 
 echo tri-core
-perf stat -o logs/test-model/$timestamp/tri-core_perf.txt taskset -c 0,1,2 python test-model.py 3 > logs/test-model/$timestamp/tri-core.txt &
+perf stat -o logs/test-model/$identification/tri-core_perf.txt taskset -c 0,1,2 python test-model.py 3 > logs/test-model/$identification/tri-core.txt &
 wait
 echo end tri-core
 sleep 300
 
 echo quad-core
-perf stat -o logs/test-model/$timestamp/quad-core_perf.txt taskset -c 0,1,2,3 python test-model.py 4 > logs/test-model/$timestamp/quad-core.txt &
+perf stat -o logs/test-model/$identification/quad-core_perf.txt taskset -c 0,1,2,3 python test-model.py 4 > logs/test-model/$identification/quad-core.txt &
 wait
 echo end quad-core
 sleep 300

### 2017-10-02 paper/overview.tex
Change track image
Added new image, track_new2.png
..

@@
   \centering
-  \begin{tabular}{|l|l|}
+  \begin{tabular}{|r|r|}
     \hline
@@
     Pololu DRV8835 motor hat&  8 \\
-    External battery pack   & 10 \\
+    External battery pack \& misc.   & 10 \\
     \hline
@@
   \centering
-  \includegraphics[width=.4\textwidth, angle=270]{figs/track_new}
+  \includegraphics[width=.4\textwidth]{figs/track_new2}
   \caption{A track for training/testing.\fixme{replace this}}

### 2017-10-02 paper/overview.tex
background revision

@@
 set of inexpensive components: a Raspberry Pi 3 Single Board Computer
-(SBC), a Pololu DRV8835 motor driver, a Playstation Eye Webcam, a
+(SBC), a Pololu DRV8835 motor driver, a Playstation Eye webcam, a
 battery, and a 1:24 scale RC car. Table~\ref{tbl:carbom} shows

### 2017-10-02 paper/overview.tex
caption in Figure 3.

@@
   \caption{DeepPicar's neural network architecture: 9 layers (5
-    convolutional, 4 fully-connected layers), 250K
-    parameters. Adopted from ~\cite{Bojarski2016}.}
+    convolutional, 4 fully-connected layers), 27 million connections,
+    250K parameters. The architecture is identical to the one used
+    in~\cite{Bojarski2016}.}
   \label{fig:architecture}

### 2017-10-02 paper/background.tex
figure 1 revision

@@
   \centering
-  \includegraphics[width=.5\textwidth]{figs/endtoend}
+  \includegraphics[width=.5\textwidth]{figs/endtoend_redrawn}
   \caption{Standard robotics control vs. DNN based end-to-end
-    control. \fixme{figure must be redrawn}}
+    control. Adopted from ~\cite{Levine2017cs294}.}
   \label{fig:end-to-end-control}
@@
 intervention~\cite{Bojarski2016}.
-\fixme{Discuss the differences between Inferencing and
-  Training. Real-time is most needed for inferencing operations.}
+
+Using deep nueral networks involves two distinct
+phases~\cite{NVIDIA2015}. The first
+phase is \emph{training} during which the weights of the network is
+incrementally updated by backpropating errors it sees from the
+training examples. Once the network is trained---i.e., the weights of
+the network minimize errors in the training examples---the next phase
+is \emph{inferencing} during which unseen data is fed to the network
+as input to produce predicted output (e.g., predicted image
+classification). In general, the training phase is most compute
+intensive and requires high throughput, which is generally not
+available on embedded platforms. The inferencing phase, on the
+other hand, is relatively less compute intensive and latency becomes
+as important, if not more, as compute throughput because many
+use-cases have strict real-time requirements (e.g., search query
+latency).
 
@@
 \subsection{Embedded Computing Platforms for Real-Time Inferencing}
+Real-time embedded systems, such as an autonomous vehicle, present
+unique challenges for deep learning as the computing platforms of such
+systems must satisfy two often conflicting goals~\cite{Otterness2017}:
+%% Recent successes in AI, including NVIDIA's DAVE-2 showing, are due
+%% in large part to the increased computing performance,
+%% which afforded researchers to train and use ever deeper neural networks with
+%% high accuracy.
+%% For practical applications, the computer platform in an
+%% autonomous vehicle must satisfy two often conflicting goals:
+On the one hand, the platform must provide 
+enough computing capacity for real-time processing of computationally
+expensive AI workloads (deep neural networks);
+On the other hand, the platform must also satisfy various
+constrains such as cost, size, weight, and power consumption limits.
 
-Recent successes in AI, including NVIDIA's DAVE-2 showing, are due
-in large part to the increased computing performance,
-which afforded researchers to train and use ever deeper neural networks with
-high accuracy. For practical, a computer platform in an automotive vehicle must
-satisfy two often conflicting goals. One the one hand, the platform must provide
-enough computing capacity for real-time processing of computationally expensive
-AI workloads. On the other hand, the platform must also satisfy various
-constrains such as cost, size, weight, and power consumption limits.
-Because of these reasons, finding a right computing platform is non-trivial
-task, one that requires deep understanding of the workloads and their
-characteristics as well as the hardware platform that executes the
-workloads. It is further complicated by the heterogeneity of available
-computing hardware resources---multicore, GPU, DSP, specialized
-hardware accelerators, etc.
+Accelerating AI workloads, especially inferencing
+operations, has gotten a lot of attentions from academia and industry
+in recent years as applications of deep learning are broadening to
+areas real-time embedded systems such as autonomous vehicles. These
+efforts include development of various heterogeneous architecture
+based system-on-a-chips (SOCs) that may include multiple cores, GPU,
+DSP, FPGA and neural network optimized ASIC hardware~\cite{Jouppi2017}.
+Consolidating multiple tasks on SoCs with lots of shared hardware
+resources while guaranteeing real-time performance is also an active
+research area, which is orthogonal to improving raw
+performance. Consolidation is necessary for efficiency but unmanaged 
+interference can nullify the benefits of consolidation~\cite{Kim2016}.
+Because of these reasons, finding a right computing platform is
+non-trivial task, one that requires deep understanding of the
+workloads and the hardware platform that executes the workloads.
 
-The primary objective of this study are (1) to understand the necessary
-computing performance for applying AI technology based robotics systems, and
-(2) what kind of computing architecture and runtime supports
-are most appropriate for such workload.
-Toward to achieve the two goals, we implement a low-cost autonomous
-car platform as a case study, as we will explain in the following
-section.
+The primary objectives of this study are (1) to understand the
+necessary computing performance for applying AI technology based
+robotics systems, and (2) what kind of computing architecture and
+runtime supports are most appropriate for such workload. Toward to
+achieve these goals, we implement a low-cost autonomous car platform
+as a case-study and systematically conduct experiments, as we will 
+describe in the subsequent sections.

### 2017-10-03 paper/overview.tex
overview revision

@@
   \includegraphics[width=.4\textwidth]{figs/track_new2}
-  \caption{A track for training/testing.\fixme{replace this}}
+  \caption{One of the custom tracks used for training/testing.}
   \label{fig:track}
@@
 is copied back to the Raspberry Pi computer. The network is then used
-by the car's main controller, which feeds an image frame the web
+by the car's main controller, which feeds a image frame from the web
 camera as input to the network. The produced steering angle output is
@@
 car. Figure~\ref{fig:controlloop} shows simplified psuedo code of the
-controller's main loop.
+controller's main loop. Among the five steps, step 3---network
+inferencing---is most computationally intensive and dominates the
+execution time.
 
@@
 
-\fixme{Discuss the limitation of the RC car platform: discrete
-  control, web camera latency}
+Note that although the steering angle output of the network $angle$ is
+a continuous real value, the RC car platform we used unfortunately
+only supports three discrete angles---left (-30$^{\circ}$), center
+(0$^{\circ}$), and right (+30$^{\circ}$)---as control
+inputs. Currently, we approximate the network generated real-valued
+angle to the closest one of the three angles. Although this may
+introduce inaccuracy in control.
+In the future, we plan to use different (more expensive) RC car
+platforms that can precisely control the car's steering angle. We
+would like to stress, however, that the use of different RC car
+platforms has no impact in terms of computational 
+aspects of the system, and that our main focus of this study is
+not in improving network accuracy but in closely replicating the
+DAVE-2's network architecture and its real-time characteristics.
 
-%% Although the steering angle output of the network is a continuous
-%% real value,
+Nevertheless, the trained model was still able to achieve reasonable a
+degree of accuracy, successfully navigating several different tracks 
+we trained. A collection of self-driving videos of the DeepPicar can
+be found at: \url{https://github.com/heechul/picar/videos}.
 

### 2017-10-03 paper/evaluation.tex
Revise evaluation figures

@@
   \includegraphics[width=.5\textwidth]{Total_Processing_Time}
-  \caption{ Real-time performance of the Raspberry Pi 3 in processing 1000 frames.}
+  \caption{ Real-time performance of the DeepPicar in processing 1000 frames.}
 \end{figure}
@@
     1 & 61.96 & 66.00 & 63.31 & 0.51\\ \hline
-    2 & 50.49 & 71.55 & 70.03 & 4.18 \\ \hline
-    3 & 48.11 & 72.22 & 58.45 & 2.80 \\ \hline
+    2 & 50.49 & 71.55 & 70.03 & 3.16 \\ \hline
+    3 & 48.11 & 72.22 & 58.45 & 4.18 \\ \hline
     4 & 42.67 & 56.37 & 50.70 & 2.80 \\
@@
 
-\begin{table*}
+\begin{figure}[h]
   \centering
-  \begin{tabular} {| l | l | l | l | l | l | l | l | l |}
+  \includegraphics[width=.5\textwidth]{figs/2ModelChart}
+  \caption{ Change in inferencing time when 2 models are run concurrently. }
+\end{figure}
+
+\begin{table}[h]
+\centering
+  \begin{tabular} {| l | l | l |}
   \hline
-  \textbf{num models} & \textbf{cores} & \textbf{mean} & \textbf{L1 refs} & \textbf{L1 
-    misses} & \textbf{L1 miss \%} & \textbf{L2 refs} & \textbf{L2 misses} & \textbf{L2 miss \%} \\ \hline
-  1 & 0,1 & 51.35 & 3.04E+10 & 4.78E+08 & 1.58 & 3.31E+09 & 3.68E+08 & 11.12\\ \hline
-  2 & 0,1 & 58.03 & 3.04E+10 & 4.91E+08 & 1.61 & 3.91E+09 & 4.26E+08 & 10.88 \\ \hline
-  2 & 2,3 & 56.40 & 3.04E+10 & 4.80E+08 & 1.58 & 3.88E+09 & 4.21E+08 & 10.87 \\ \hline
+  \textbf{num models} & 1 & 2 \\ \hline
+  \textbf{L1 refs} &  3.04E+10 &  3.04E+10 \\ \hline
+  \textbf{L1 misses} & 4.78E+08 & 4.91E+08 \\ \hline
+  \textbf{L1 miss \%} & 1.58 & 1.61 \\ \hline
+  \textbf{L2 refs} & 3.31E+09 &  3.91E+09 \\ \hline
+  \textbf{L2 misses} & 3.68E+08 & 4.62E+08 \\ \hline
+  \textbf{L2 miss \%} & 11.12 & 10.88 \\ \hline
   \end{tabular}
-  \caption{Real-time performance of the DeepPicar when 2 models are running simultaneously.}
-\end{table*}
+  \caption{ Effect of 2 simulatneous models on cache accesses. }
+\end{table}
 
-The results for the 2 model test are outlined in Table IV, and the 4 model test is summarized in Table 
-V. In the the 2 model test, both of the models showed average inference time increases of around 5-7 
-ms, $\sim$10\%, when compared to a baseline of 1 model running on two cores. The difference was even 
+The results for the 2 model test are outlined in Figure 3 and Table IV, and the 4 model test is summarized in 
+Figure 4 Table V. In the the 2 model test, both of the models showed average inference time increases of around 
+5-7 ms, $\sim$10\%, when compared to a baseline of 1 model running on two cores. The difference was even 
 greater in the 4 model test, as each model displayed an average inference time increase of 
@@
 
-\begin{table*}
+\begin{figure}[h]
   \centering
-  \begin{tabular} {| l | l | l | l | l | l | l | l | l |}
+  \includegraphics[width=.5\textwidth]{figs/4ModelChart}
+  \caption{Change in inferencing time when 4 models are run simultaneously. }
+\end{figure}
+
+\begin{table}[h]
+\centering
+  \begin{tabular} {| l | l | l |}
   \hline
-  \textbf{num models} & \textbf{core} & \textbf{mean} & \textbf{L1 refs} & \textbf{L1 
-    misses} & \textbf{L1 miss \%} & \textbf{L2 refs} & \textbf{L2 misses} & \textbf{L2 miss \%} \\ \hline
-    1 & 0 & 62.48 & 2.78E+10 & 4.36E+08 & 1.57 & 2.83E+09 & 3.59E+08 & 12.68 \\ \hline
-    4 & 0 & 77.90 & 2.79E+10 & 4.53E+08 & 1.63 & 3.36E+09 & 4.43E+08 & 13.19 \\ \hline
-    4 & 1 & 78.89 & 2.79E+10 & 4.64E+08 & 1.67 & 3.42E+09 & 4.38E+08 & 12.82 \\ \hline
-    4 & 2 & 77.81 & 2.78E+10 & 4.45E+08 & 1.60 & 3.45E+09 & 4.41E+08 & 12.77 \\ \hline
-    4 & 3 & 77.87 & 2.79E+10 & 4.45E+08 & 1.60 & 3.41E+09 & 4.39E+08 & 12.88 \\ \hline
+  \textbf{num models} & 1 & 4 \\  \hline
+  \textbf{L1 refs} & 2.78E+10 & 2.79E+10 \\ \hline
+  \textbf{L1 misses} & 4.36E+08 & 4.53E+08 \\ \hline
+  \textbf{L1 miss \%} & 1.57 & 1.63 \\ \hline
+  \textbf{L2 refs} & 2.83E+09 & 3.36E+09 \\ \hline 
+  \textbf{L2 misses} & 3.59E+08 & 4.43E+08 \\ \hline
+  \textbf{L2 miss \%} & 12.68 & 13.19 \\ 
+\hline
   \end{tabular}
-  \caption{Real-time performance of the DeepPicar when 4 models are running simultaneously.}
-\end{table*}
+  \caption{ Effect of 4 concurrent models on cache accesses. }
+\end{table}
 
@@
 running synthetic benchmarks concurrently with the model. In each experiment, we run a single model on a 
-number of cores, and a synthetic benchmark on the remaining idle cores. The results of each test can be 
-seen in Table VII.
+number of cores, and a synthetic benchmark on the remaining idle cores. The effects of the benchmarks can
+seen in Figure 9 and Table VI.
+
+\begin{figure}[h]
+  \centering
+  \includegraphics[width=.5\textwidth]{figs/BenchmarkChart}
+  \caption{ Effect of benchmarks on inferencing time. }
+\end{figure}
 
@@
   \textbf{num cores} & 1 & 2 & 3 \\ \hline
-  \textbf{mean} & 562.33 & 391.98 & 98.05 \\ \hline
   \textbf{L1 refs} & 3.16E+10 & 3.17E+10 & 3.06E+10 \\ \hline
@@
   \end{tabular}
-  \caption{Real-time performance of the DeepPicar when benchmarks are run concurrently with the model.}
+  \caption{ Effect of benchmarks on cache accesses. }
 \end{table}

### 2017-10-03 paper/overview.tex
use hyperref package

@@
 inputs. Currently, we approximate the network generated real-valued
-angle to the closest one of the three angles. Although this may
+angle to the closest one of the three angles, which may
 introduce inaccuracy in control.
@@
 
-Nevertheless, the trained model was still able to achieve reasonable a
-degree of accuracy, successfully navigating several different tracks 
-we trained. A collection of self-driving videos of the DeepPicar can
-be found at: \url{https://github.com/heechul/picar/videos}.
+Nevertheless, the trained models were still able to achieve reasonable
+a degree of accuracy, successfully navigating several different tracks 
+we trained. The source code, build instruction, and a collection of
+self-driving videos of the DeepPicar can be found at:
+\url{https://github.com/heechul/picar}.
+%% \begin{center}
+%% \end{center}
 

### 2017-10-03 paper/evaluation.tex
Fix evaluation typos

@@
 solely for the purpose of operating an autonomous vehicle. Thus, we test how the number of cores 
-utilized for real-time operations affects the Pi's overall ability to function as an autonomous 
+utilized for real-time operations affects the DeepPicar's overall ability to function as an autonomous 
 vehicle platform.
@@
 misses, inferencing was able to be completed in under 100 ms. The other tests, however, showed dramatic 
-changes when the model was run on 1 and 2 cores. In both instances, L2 misses rise to $\sim$23\% of all 
+changes when the model was run on 1 and 2 cores. In both instances, L2 misses rose to $\sim$23\% of all 
 references. The additional benchmark in the 1 core test was even more detrimental as the average 

### 2017-10-03 paper/related.tex
working on related works -- probably a little short

@@
 
-\cite{Bojarski2016}
+Autonomous vehicles are becoming an increasingly interesting research topic as computing platforms capable of safely processing sensor input into safe vehicle controls become smaller and more affordable.
+The model proposed in End-To-End Learning for Self-Driving Cars \cite{Bojarski2016} is viable on a platform as simple as a Raspberry Pi.
+
+Given smaller, affordable platforms with multiple cores and even GPU based parallel processing such as the Tegra X1 \cite{NVIDIA2015} and the recently released Tegra X2 \cite{Amert2017}, allocation and management of available shared CPU and GPU resources \cite{Kim2016} becomes increasingly important.
+Research on improving the performance of neural networks \cite{Jouppi2017} also attributes to the viability of using these models in small systems.
+
+Together with hardware and platform improvements, there have been significant improvements in the algorithms such as a neural network algorithm capable of surpassing human level control at atari games \cite{DBLP}.
+A popular technique had been to take sensor input and determine a depth map, such as the model used in \cite{Michels:2005}.
+End-to-end models such as the one used in \cite{Bojarski2016} can improve the efficiency of the model.
+Improving the flexibility of the model and it abilty to adapt to different situations and avoiding overfitting to training data continues to be an important topic in autonomous driving \cite{Pomerleau1989}.
+
+Producing large platforms such as DAVE \cite{LeCun:04} has often been the approach to developing the technology for self-driving vehicles.
+This approach, while certainly effective, presents a higher barrier to entry for research groups.
+Smaller, more focused platforms, such as the one found in \cite{Michels:2005} present a more accesible way forward.

### 2017-10-03 paper/background.tex

@@
 The use of neural networks for end-to-end control of autonomous
-driving was first demonstrated in late 1980s~\cite{Pomerleau1989},
-using a small fully connected neural network; and subsequently in a
-DARPA Autonomous Vehicle (DAVE) project in early
+car was first demonstrated in late 1980s~\cite{Pomerleau1989},
+using a small 3-layer fully connected neural network; and subsequently
+in a DARPA Autonomous Vehicle (DAVE) project in early
 2000s~\cite{LeCun:04}, using a 6 layer convolutional neural network

### 2017-10-03 paper/evaluation.tex
evaluation revision (WIP)

@@
 
-In evaluating the real-time efficacy of the DeepPicar, the methodology is consistent across all 
-experiments. The performance of the platform is measured over a set of 1001 video frames that are 
-each individually fed to the model. The processing time for the first frame is omitted as, due to 
-cache warmup, it is uncharacteristically high and doesn't accurately represent the Picar's capabilities. 
-A WCET of 50 ms, or 20 Hz, is used as a baseline to assess the platform's ability to complete all 
-necessary real-time operations in a timely manner.
+In this section, we experimentally analyze various real-time aspects
+of the DeepPicar. This includes
+(1) measurement based worst-case execution time (WCET) analysis of
+deep learning inferencing,
+(2) the effect of using multiple cores in accelerating inferencing,
+(3) the effect of co-scheduling multiple deep neural network models,
+and 
+(4) the effect of co-scheduling memory bandwidth intensive co-runners.
 
-\subsection{Real-Time Operations}
+\subsection{Setup}
+
+The Raspberry Pi 3 Model B platform used in DeepPicar equip a Boardcom
+BCM2837 SoC, which has a quad-core 64bit ARM Cortex-A53 cluster,
+running at up to 1.2GHz. The chip also includes Broadcom's Videocore IV
+GPU, although we did not use the GPU in our evaluation due to the lack
+of sofware support (Tensorflow is not compatible with the GPU).
+For software, we use Ubuntu MATE 16.04, Tensorflow 1.1 and python
+2.7. We disabled DVFS (dynamic voltage frequency scaling) and
+configure the clock speed of each core at the maximum 1.2GHz.
+
+\subsection{Inference Timing for Real-Time Control}
+
+For real-time control of a car (or any robots), the control loop
+frequency must be sufficiently high so that the car can quickly
+react to the changing environment and its internal states. In general,
+control performance improves when the frequency is higher, though
+computation time and the type of particular physical system are
+factors in determining a proper control loop latency. While a standard
+control system may be comprized of multiple control loops with
+differning control frequencies---e.g., a inner control loop for lower-level
+PD control, a outer loop for motion planning, etc.---DeepPicar's
+control loop is a single layer, as shown earlier in
+Figure~\ref{fig:controlloop}, as a single deep neural network
+replaces the traditional multi-layer control pipline. (Recall
+Figure~\ref{fig:end-to-end-control} on the differences between the
+standard robotics control vs. end-to-end deep learning approach).
+This means that the DNN interference operation must be completed
+within the inner-most control loop update frequency. To understand
+achievable control-loop update frequencies, we experimentally measured
+the execution times of DeepPicar's DNN inference operations.
+
+% 50-200Hz for quadcopters:
+% https://robotics.stackexchange.com/questions/231/what-frequency-does-my-quadcopter-output-sense-calculate-output-update-loop-need
+
+\begin{figure}[t]
+  \centering
+  \includegraphics[width=.5\textwidth]{Total_Processing_Time}
+  \caption{DeepPicar's control loop processing times over 1000 input image frames.}
+  \label{fig:control-loop-timing}
+\end{figure}
+
+Figure~\ref{fig:control-loop-timing} shows the measured control loop
+processing times of the DeepPicar in which 1001 image frames, from the
+camera, are fed to the DNN, one frame per each control loop. The
+measured time is total execution time of a control loop.
+
+%% In evaluating the real-time efficacy of the DeepPicar, the methodology is consistent across all 
+%% experiments. The performance of the platform is measured over a set of 1001 video frames that are 
+%% each individually fed to the model. The processing time for the first frame is omitted as, due to 
+%% cache warmup, it is uncharacteristically high and doesn't accurately represent the Picar's capabilities. 
+%% A WCET of 50 ms, or 20 Hz, is used as a baseline to assess the platform's ability to complete all 
+%% necessary real-time operations in a timely manner.
+
 We seek to determine if the DeepPicar is capable of consistently executing all necessary functions 
@@
 was unable to do so for very few frames.
-
-\begin{figure}[h]
-  \centering
-  \includegraphics[width=.5\textwidth]{Total_Processing_Time}
-  \caption{ Real-time performance of the DeepPicar in processing 1000 frames.}
-\end{figure}
 

### 2017-10-03 paper/evaluation.tex
Add draft evaluation of Pi3 vs UP Board
Does not have results from Tegra TX2 and will need to be updated to
include those. Also, the figures will need to be updated to include the
Tegra TX2 findings.

@@
 off or made idle so that the cpu is allowed time to cool down.
+
+\subsection{System Comparisons}
+We compare the real time performance of the Raspberry Pi 3 to two other embedded computing platforms, the Intel UP Board the NVIDIA Tegra TX2, to evaluate their efficacy in vision-based autonomous vehicles.
+We conduct the same experiments done on the DeepPicar with each platform, and analyze the results.
+
+\begin{figure}[h]
+  \centering
+  \includegraphics[width=.5\textwidth]{figs/system_multicore}
+  \caption{Inferencing time across platforms based on number of cores used.}
+  \label{fig:sys_core}
+\end{figure}
+
+The results of the multicore tests can be seen in Figure~\ref{fig:sys_core}.
+It was found that the Intel UP Board performed better in all experiments.
+On average, it was able to perform inferencing operations, at least, twice as fast as the Raspberry Pi 3.
+As a result, the UP Board was able to satisfy the 50 ms WCET by a very clear margin (in the worst case, it would still meet it by $\sim$20 ms), and, thus, demonstrates that it is more capable of autonomous driving than the Raspberry Pi 3.
+
+The Intel UP Board also performed better in the multimodel tests, as is shown in Figure~\ref{fig:sys_model}.
+Once again, the UP was able to complete inferecing operations in half the time when compared to the Pi.
+Furthermore, the Intel UP Board was capable of satisfying the 50 ms WCET, while the Raspberry Pi 3 was unable to do so.
+As such, the Intel UP Board displays greater potential for the simultaneous execution of multiple models during self-driving operation, while the Raspberry Pi 3 would struggle to do so.
+
+\begin{figure}[h]
+  \centering
+  \includegraphics[width=.5\textwidth]{figs/system_multimodel}
+  \caption{Inferencing time across platforms based on number of models run concurrently.}
+  \label{fig:sys_model}
+\end{figure}
+
+Finally, we examined if the performance of the Intel UP Board and Tegra TX2 would be affected by the addition of the synthetic benchmarks.
+As summarized in Figure~\ref{fig;sys_bench}, the Intel UP Board did experience a change in its inferencing times, but the increase was not as drastic as is seen by the Raspberry Pi 3.
+in the worst case, the Intel UP Board produced times that were over two times as large, whereas the Raspberry Pi 3 output times that were around 10 times greater.
+The Intel UP Board was successful in completing inferencing operations in under 50 ms when benchmarks were run on a maximum of 2 cores, while the Raspberry Pi 3 failed to do so when any benchmark was introduced.
+As such, it can be concluded that the Intel UP Board would be more capable of running computationally heavy processes during autonomous driving, and that the Pi wouldn't be able to do the same.
+
+\begin{figure}[h]
+  \centering
+  \includegraphics[width=.5\textwidth]{figs/system_benchmark}
+  \caption{Inferencing time across platforms based on number of cores used.}
+  \label{fig:sys_bench}
+\end{figure}
+
+In the comparison of the real-time capabilities of three embedded computing platforms, we found that the Raspberry Pi 3 performed the worst.
+When compared to the Intel UP Board, inferencing on the Pi took twice as long across all multicore and multimodel experiments.
+The difference was even more noticeable in the the addition of computationally heavy synthetic benchmarks that had a much more dire effect on the Pi.

### 2017-10-03 paper/evaluation.tex
Add line breaks
evaluation revision subsection A.

--- 2017-10-03 paper/evaluation.tex
@@
 % https://robotics.stackexchange.com/questions/231/what-frequency-does-my-quadcopter-output-sense-calculate-output-update-loop-need 
+% https://quadmeup.com/pid-looptime-why-it-is-not-only-about-frequency/
 
@@
 
+\begin{table}[t]
+  \centering
+  \begin{tabular} {| c | r | r | r | r |}
+    \hline
+    \textbf{Operation} & \textbf{Mean} & \textbf{Max} &   \textbf{99pct.} & \textbf{Stdev.} \\ \hline
+    Image capture        & 2.28  &  4.94 &  4.54  & 0.52 \\ \hline
+    Image pre-processing & 3.09  &  4.60 &  3.31  & 0.10 \\ \hline
+    DNN inferencing      & 37.30 & 51.03 & 45.48  & 2.75 \\ \hline
+    Total Time           & 42.67 & 56.37 & 50.70  & 2.80 \\ \hline
+  \end{tabular}
+  \caption{Control loop timing breakdown.}
+  \label{tbl:control-loop-breakdown}
+\end{table}
+
 Figure~\ref{fig:control-loop-timing} shows the measured control loop 
-processing times of the DeepPicar in which 1001 image frames, from the
-camera, are fed to the DNN, one frame per each control loop. The
-measured time is total execution time of a control loop.
+processing times of the DeepPicar over 1000 image frames (one per each
+control loop). We omit the first frame processing time for cache
+warmup. Table~\ref{tbl:control-loop-breakdown} shows the time
+breakdown of each control loop. Note that all four CPU cores of the
+Raspberry Pi 3 were used by the Tensorflow library when performing the
+DNN inference operations.
+
+First, as expected, we find that the inference operation
+dominates the control loop execution time, accounting about 85\% of
+the execution time.
+Second, more importantly, we also find that the measured average
+execution time of a single control loop is 42.67 ms, or 23.4 Hz and
+the 99 percentail time is 50.70ms.
+This means that the DeepPicar can operates
+at about 20 Hz control frequency in real-time using only the on-board
+Raspberry Pi 3 computing platform, without needing any remote computing
+resources. We consider these results respectable given the complexity
+of the deep neural network  and the fact that the inference operation
+performed by Tensorflow only utilizes the CPU cores of the
+Raspberry Pi 3 (its GPU is not supported by Tensorflow.)
+
+In comparison, NVIDIA's DAVE-2 system, which has the exact same neural
+network architecture, is reportedly run at 30 Hz~\cite{Bojarski2016},
+just a bit faster than the DeepPicar. Although we believe it was not
+limited by their computing platform (we will experimentally compare
+performance differences among multiple embedded computing platforms,
+including NVIDIA's Jetson TK2, later in
+Section~\ref{sec:eval-platforms}), the fact that the low-cost
+Raspberry Pi 3 can achieve similar real-time control performance is
+surprising.
+
+%% In our platform, three main real-time tasks are performed during autonomous operation. In order, these 
+%% operations are: (1) capturing and reading the input frame from the designated camera or video stream, 
+%% (2) preprocessing the acquired frame so that it is compatible with the DNN, and (3) feeding the frame 
+%% to, and getting the angle prediction from, the model. We aim to determine which operation(s), if any, 
+%% require the most time to be performed.
+
+%% In order to determine which operation(s) take the longest to execute, we measured the time it 
+%% took for each step to be completed. For this experiment, all four of the Pi's cpu cores were utilized, 
+%% and only one model was run. As is shown in Table II, the angle prediction operation consumes the 
+%% majority of the processing for each frame. Furthermore, the time it takes for the operation to 
+%% complete is volatile, and can range anywhere between 30 ms and 50 ms for any particular frame. On the 
+%% other hand, both the frame capture and preprocessing operations take substantially less time 
+%% and are relatively more consistent in their times, at 2 ms and 3 ms, respectively.
 
@@
 %% each individually fed to the model. The processing time for the first frame is omitted as, due to 
-%% cache warmup, it is uncharacteristically high and doesn't accurately represent the Picar's capabilities. 
+%% cache warmup, it is uncharacteristically high and doesn't
+%% accurately represent the Picar's capabilities.  
 %% A WCET of 50 ms, or 20 Hz, is used as a baseline to assess the platform's ability to complete all 
@@
 
-We seek to determine if the DeepPicar is capable of consistently executing all necessary functions 
-before their given deadlines. In the case of our platform, it has to process every given frame and get 
-the predicted angles from the model within a WCET of 50 ms. In order to determine if this was achievable, 
-we tested the platform by running a single model that utilized all 4 cpu cores and measured the time it 
-took for each frame to be processed. The performance of the DeepPicar can be seen in Figure 6. We found 
-that DeepPicar was able to completely process the vast majority of the provided frames within 50 ms, and 
-was unable to do so for very few frames.
+%% We seek to determine if the DeepPicar is capable of consistently executing all necessary functions 
+%% before their given deadlines. In the case of our platform, it has to process every given frame and get 
+%% the predicted angles from the model within a WCET of 50 ms. In order to determine if this was achievable, 
+%% we tested the platform by running a single model that utilized all 4 cpu cores and measured the time it 
+%% took for each frame to be processed. The performance of the DeepPicar can be seen in Figure 6. We found 
+%% that DeepPicar was able to completely process the vast majority of the provided frames within 50 ms, and 
+%% was unable to do so for very few frames.
 
-In our platform, three main real-time tasks are performed during autonomous operation. In order, these 
-operations are: (1) capturing and reading the input frame from the designated camera or video stream, 
-(2) preprocessing the acquired frame so that it is compatible with the DNN, and (3) feeding the frame 
-to, and getting the angle prediction from, the model. We aim to determine which operation(s), if any, 
-require the most time to be performed.
+%% In our platform, three main real-time tasks are performed during autonomous operation. In order, these 
+%% operations are: (1) capturing and reading the input frame from the designated camera or video stream, 
+%% (2) preprocessing the acquired frame so that it is compatible with the DNN, and (3) feeding the frame 
+%% to, and getting the angle prediction from, the model. We aim to determine which operation(s), if any, 
+%% require the most time to be performed.
 
-\begin{table}[h]
-  \centering
-  \begin{tabular} {| l | l | l | l | l |}
-    \hline
-    \textbf{operation} & \textbf{mean} & \textbf{max} & \textbf{99pct} & \textbf{stdev} \\ \hline 
-    Frame Capture & 2.28 & 4.94 & 4.54  & 0.52\\ \hline
-    Preprocessing & 3.09 & 4.60 & 3.31 & 0.10 \\ \hline
-    Angle Prediction & 37.30 & 51.03 & 45.48 & 2.75 \\  \hline
-    Total Time & 42.67 & 56.37 & 50.70 & 2.80 \\
-    \hline
-  \end{tabular}
-  \caption{Real-time performance of DeepPicar for each autonomous vehicle operation performed.}
-\end{table}
-
-In order to determine which operation(s) take the longest to execute, we measured the time it 
-took for each step to be completed. For this experiment, all four of the Pi's cpu cores were utilized, 
-and only one model was run. As is shown in Table II, the angle prediction operation consumes the 
-majority of the processing for each frame. Furthermore, the time it takes for the operation to 
-complete is volatile, and can range anywhere between 30 ms and 50 ms for any particular frame. On the 
-other hand, both the frame capture and preprocessing operations take substantially less time 
-and are relatively more consistent in their times, at 2 ms and 3 ms, respectively.
+%% In order to determine which operation(s) take the longest to execute, we measured the time it 
+%% took for each step to be completed. For this experiment, all four of the Pi's cpu cores were utilized, 
+%% and only one model was run. As is shown in Table II, the angle prediction operation consumes the 
+%% majority of the processing for each frame. Furthermore, the time it takes for the operation to 
+%% complete is volatile, and can range anywhere between 30 ms and 50 ms for any particular frame. On the 
+%% other hand, both the frame capture and preprocessing operations take substantially less time 
+%% and are relatively more consistent in their times, at 2 ms and 3 ms, respectively.
 

### 2017-10-03 paper/evaluation.tex

@@
 off or made idle so that the cpu is allowed time to cool down.
+
+\subsection{System Comparisons}
+We compare the real time performance of the Raspberry Pi 3 to two other embedded computing platforms, 
+the Intel UP Board the NVIDIA Tegra TX2, to evaluate their efficacy in vision-based autonomous 
+vehicles. We conduct the same experiments done on the DeepPicar with each platform, and analyze the 
+results.
+
+\begin{figure}[h]
+  \centering
+  \includegraphics[width=.5\textwidth]{figs/system_multicore}
+  \caption{Inferencing time across platforms based on number of cores used.}
+  \label{fig:sys_core}
+\end{figure}
+
+The results of the multicore tests can be seen in Figure~\ref{fig:sys_core}. It was found that the 
+Intel UP Board performed better in all experiments. On average, it was able to perform inferencing 
+operations, at least, twice as fast as the Raspberry Pi 3. As a result, the UP Board was able to 
+satisfy the 50 ms WCET by a very clear margin (in the worst case, it would still meet it by $\sim$20 
+ms), and, thus, demonstrates that it is more capable of autonomous driving than the Raspberry Pi 3.
+
+The Intel UP Board also performed better in the multimodel tests, as is shown in 
+Figure~\ref{fig:sys_model}. Once again, the UP was able to complete inferecing operations in half the 
+time when compared to the Pi. Furthermore, the Intel UP Board was capable of satisfying the 50 ms 
+WCET, while the Raspberry Pi 3 was unable to do so. As such, the Intel UP Board displays greater 
+potential for the simultaneous execution of multiple models during self-driving operation, while the 
+Raspberry Pi 3 would struggle to do so.
+
+\begin{figure}[h]
+  \centering
+  \includegraphics[width=.5\textwidth]{figs/system_multimodel}
+  \caption{Inferencing time across platforms based on number of models run concurrently.}
+  \label{fig:sys_model}
+\end{figure}
+
+Finally, we examined if the performance of the Intel UP Board and Tegra TX2 would be affected by the 
+addition of the synthetic benchmarks. As summarized in Figure~\ref{fig;sys_bench}, the Intel UP Board 
+did experience a change in its inferencing times, but the increase was not as drastic as is seen by 
+the Raspberry Pi 3. in the worst case, the Intel UP Board produced times that were over two times as 
+large, whereas the Raspberry Pi 3 output times that were around 10 times greater. The Intel UP Board 
+was successful in completing inferencing operations in under 50 ms when benchmarks were run on a 
+maximum of 2 cores, while the Raspberry Pi 3 failed to do so when any benchmark was introduced. As 
+such, it can be concluded that the Intel UP Board would be more capable of running computationally 
+heavy processes during autonomous driving, and that the Pi wouldn't be able to do the same.
+
+\begin{figure}[h]
+  \centering
+  \includegraphics[width=.5\textwidth]{figs/system_benchmark}
+  \caption{Inferencing time across platforms based on number of cores used.}
+  \label{fig:sys_bench}
+\end{figure}
+
+In the comparison of the real-time capabilities of three embedded computing platforms, we found that 
+the Raspberry Pi 3 performed the worst. When compared to the Intel UP Board, inferencing on the Pi 
+took twice as long across all multicore and multimodel experiments. The difference was even more 
+noticeable in the the addition of computationally heavy synthetic benchmarks that had a much more 
+dire effect on the Pi.

### 2017-10-03 paper/evaluation.tex
figure

@@
 
-\subsection{Multicore Performance}
+\subsection{Effect of the Core Count to Inference Timing}
 It may not always be the case that all four cores of the Raspberry Pi 3's Cortex A-53 CPU can be used 
@@
 vehicle platform.
+
+\begin{figure}[h]
+  \centering
+  \includegraphics[width=.5\textwidth]{figs/perf_vs_corecnt}
+  \caption{Average control loop execution time vs. \#of CPU cores used
+    for inferencing by Tensorflow.}
+  \label{fig:perf-vs-corecnt}
+\end{figure}
 

### 2017-10-03 paper/overview.tex
typo
figure

@@
   \centering
-  \begin{tabular}{|r|r|}
+  \begin{tabular}{|c|r|}
     \hline
-    Item                    & cost (\$) \\
+    Item                    & Cost (\$) \\
     \hline
@@
     250K parameters. The architecture is identical to the one used
-    in~\cite{Bojarski2016}.}
+    in NVIDIA's real self-driving car~\cite{Bojarski2016}.}
   \label{fig:architecture}
@@
 
-Nevertheless, the trained models were still able to achieve reasonable
-a degree of accuracy, successfully navigating several different tracks 
+Nevertheless, the trained models were still able to achieve a reasonable
+degree of accuracy, successfully navigating several different tracks 
 we trained. The source code, build instruction, and a collection of

### 2017-10-04 paper/evaluation.tex
eval C. revision

--- 2017-10-03 paper/evaluation.tex
@@
 
-%% In our platform, three main real-time tasks are performed during autonomous operation. In order, these 
-%% operations are: (1) capturing and reading the input frame from the designated camera or video stream, 
-%% (2) preprocessing the acquired frame so that it is compatible with the DNN, and (3) feeding the frame 
-%% to, and getting the angle prediction from, the model. We aim to determine which operation(s), if any, 
-%% require the most time to be performed.
+\subsection{Effect of the Core Count to Inference Timing}
 
-%% In order to determine which operation(s) take the longest to execute, we measured the time it 
-%% took for each step to be completed. For this experiment, all four of the Pi's cpu cores were utilized, 
-%% and only one model was run. As is shown in Table II, the angle prediction operation consumes the 
-%% majority of the processing for each frame. Furthermore, the time it takes for the operation to 
-%% complete is volatile, and can range anywhere between 30 ms and 50 ms for any particular frame. On the 
-%% other hand, both the frame capture and preprocessing operations take substantially less time 
-%% and are relatively more consistent in their times, at 2 ms and 3 ms, respectively.
+In this experiment, we investigate the scalability of performing
+inference operation of DeepPicar's nueral network with respect to the
+number of cores. As noted earlier, the Raspberry Pi 3 platform has
+four Cortex-A53 cores and the Tensorflow 
+provides a programmable mechanism to to adjust how many cores to be
+used by the library. Leveraging this feature, we repeat the
+same experiment described in the previous subsection with varying
+number of CPU cores---from one to four.
 
-%% In evaluating the real-time efficacy of the DeepPicar, the methodology is consistent across all 
-%% experiments. The performance of the platform is measured over a set of 1001 video frames that are 
-%% each individually fed to the model. The processing time for the first frame is omitted as, due to 
-%% cache warmup, it is uncharacteristically high and doesn't
-%% accurately represent the Picar's capabilities.  
-%% A WCET of 50 ms, or 20 Hz, is used as a baseline to assess the platform's ability to complete all 
-%% necessary real-time operations in a timely manner.
+\begin{figure}[h]
+  \centering
+  \includegraphics[width=.45\textwidth]{figs/perf_vs_corecnt}
+  \caption{Average control loop execution time vs. \#of CPU
+    cores. \fixme{instead of the total control loop time, use the
+      average of DNN inferencing time.}}
+  \label{fig:perf-vs-corecnt}
+\end{figure}
 
-%% We seek to determine if the DeepPicar is capable of consistently executing all necessary functions 
-%% before their given deadlines. In the case of our platform, it has to process every given frame and get 
-%% the predicted angles from the model within a WCET of 50 ms. In order to determine if this was achievable, 
-%% we tested the platform by running a single model that utilized all 4 cpu cores and measured the time it 
-%% took for each frame to be processed. The performance of the DeepPicar can be seen in Figure 6. We found 
-%% that DeepPicar was able to completely process the vast majority of the provided frames within 50 ms, and 
-%% was unable to do so for very few frames.
+%% \begin{table}[h]
+%%   \centering
+%%   \begin{tabular} {| c | l | l | l | l | l | l | l | l | l |}
+%%     \hline
+%%     \textbf{\#of cores} & \textbf{Mean} & \textbf{Max} & \textbf{99pct} & \textbf{Stdev} \\ \hline 
+%%     1 & 61.96 & 66.00 & 63.31 & 0.51\\ \hline
+%%     2 & 50.49 & 71.55 & 70.03 & 3.16 \\ \hline
+%%     3 & 48.11 & 72.22 & 58.45 & 4.18 \\ \hline
+%%     4 & 42.67 & 56.37 & 50.70 & 2.80 \\
+%%     \hline
+%%   \end{tabular}
+%%   \caption{Execution time statistics vs. \#of CPU cores.}
+%% \end{table}
 
-%% In our platform, three main real-time tasks are performed during autonomous operation. In order, these 
-%% operations are: (1) capturing and reading the input frame from the designated camera or video stream, 
-%% (2) preprocessing the acquired frame so that it is compatible with the DNN, and (3) feeding the frame 
-%% to, and getting the angle prediction from, the model. We aim to determine which operation(s), if any, 
-%% require the most time to be performed.
+Figure~\ref{fig:perf-vs-corecnt} shows the average execution times of
+the control loop as we vary the number of cores used by the
+Tensorflow. As expected as we assign more cores, the average execution
+time decreases---from 61.96 ms on a single core to 42.67 ms on four
+cores. However, the improvement is far from being ideal as we expected
+linear scaling. In particular, from 2 cores to 3 cores, the
+improvement is mere 2.38ms (or 4\%). In short, we find that the
+scalability of DeepPicar's deep neural network is not ideal on the
+platform. We do not know whether it is due to the limitations of
+Tensorflow's multicore implementation or it's the model's inherent
+characteristics.
 
-%% In order to determine which operation(s) take the longest to execute, we measured the time it 
-%% took for each step to be completed. For this experiment, all four of the Pi's cpu cores were utilized, 
-%% and only one model was run. As is shown in Table II, the angle prediction operation consumes the 
-%% majority of the processing for each frame. Furthermore, the time it takes for the operation to 
-%% complete is volatile, and can range anywhere between 30 ms and 50 ms for any particular frame. On the 
-%% other hand, both the frame capture and preprocessing operations take substantially less time 
-%% and are relatively more consistent in their times, at 2 ms and 3 ms, respectively.
+The poor scalability opens up the possibility of consolidating
+multiple different tasks or different nueral network models rather
+than allocating all cores for a single nueral network model. For
+example, it is conceivable to use four cameras and four different
+neural network models, each of which is trained seperately and
+executed on a single dedicated core. Assuming we use the same network
+architecture for all models, then one might exepct to achieve up to
+15 Hz using one core (given 1 core can deliver 62ms average
+execution time). In the next experiment, we investigate the
+feasibility of such a scenario.
 
-\subsection{Multicore Performance}
-It may not always be the case that all four cores of the Raspberry Pi 3's Cortex A-53 CPU can be used 
-solely for the purpose of operating an autonomous vehicle. Thus, we test how the number of cores 
-utilized for real-time operations affects the DeepPicar's overall ability to function as an autonomous 
-vehicle platform.
+%% It may not always be the case that all four cores of the Raspberry Pi
+%% 3's Cortex A-53 CPU can be used solely for the purpose of operating an
+%% autonomous vehicle. Thus, we test how the number of cores utilized for
+%% real-time operations affects the DeepPicar's overall ability to
+%% function as an autonomous vehicle platform.
 
-\begin{table}[h]
-  \centering
-  \begin{tabular} {| l | l | l | l | l | l | l | l | l | l |}
-    \hline
-    \textbf{num cores} & \textbf{mean} & \textbf{max} & \textbf{99pct} & \textbf{stdev} \\ \hline 
-    1 & 61.96 & 66.00 & 63.31 & 0.51\\ \hline
-    2 & 50.49 & 71.55 & 70.03 & 3.16 \\ \hline
-    3 & 48.11 & 72.22 & 58.45 & 4.18 \\ \hline
-    4 & 42.67 & 56.37 & 50.70 & 2.80 \\
-    \hline
-  \end{tabular}
-  \caption{Real-time performance of DeepPicar depending on the number of cores used.}
-\end{table}
-
-The performance of the DeepPicar is summarized in Table III. The DeepPicar performed better, on average, 
-when it utilized more cores. With 4 cores, it was able to meet the vast majority of its deadlines, doing 
-so in almost 99\% of the input frames. The platform performed the worst when using only 1 core, 
-as it was unable to meet any of its deadlines.  On average, using 3 cores instead of 2 only had a 
-performance increase of approximately 2 ms, so the addition of one core in that specific case offers 
-relatively little improvement. One important observation is that the performance was more consistent 
-when only 1 core was used. As a result, the use of multiple cores is very beneficial in terms of 
-reducing the time it takes to complete inference operations, but may result in times that are more 
-volatile.
+%% The performance of the DeepPicar is summarized in Table III. The
+%% DeepPicar performed better, on average, when it utilized more
+%% cores. With 4 cores, it was able to meet the vast majority of its
+%% deadlines, doing so in almost 99\% of the input frames. The platform
+%% performed the worst when using only 1 core, as it was unable to meet
+%% any of its deadlines.  On average, using 3 cores instead of 2 only had
+%% a performance increase of approximately 2 ms, so the addition of one
+%% core in that specific case offers  relatively little improvement. One
+%% important observation is that the performance was more consistent when
+%% only 1 core was used. As a result, the use of multiple cores is very
+%% beneficial in terms of reducing the time it takes to complete
+%% inference operations, but may result in times that are more volatile.
 
@@
 dire effect on the Pi.
+
+%% In our platform, three main real-time tasks are performed during autonomous operation. In order, these 
+%% operations are: (1) capturing and reading the input frame from the designated camera or video stream, 
+%% (2) preprocessing the acquired frame so that it is compatible with the DNN, and (3) feeding the frame 
+%% to, and getting the angle prediction from, the model. We aim to determine which operation(s), if any, 
+%% require the most time to be performed.
+
+%% In order to determine which operation(s) take the longest to execute, we measured the time it 
+%% took for each step to be completed. For this experiment, all four of the Pi's cpu cores were utilized, 
+%% and only one model was run. As is shown in Table II, the angle prediction operation consumes the 
+%% majority of the processing for each frame. Furthermore, the time it takes for the operation to 
+%% complete is volatile, and can range anywhere between 30 ms and 50 ms for any particular frame. On the 
+%% other hand, both the frame capture and preprocessing operations take substantially less time 
+%% and are relatively more consistent in their times, at 2 ms and 3 ms, respectively.
+
+%% In evaluating the real-time efficacy of the DeepPicar, the methodology is consistent across all 
+%% experiments. The performance of the platform is measured over a set of 1001 video frames that are 
+%% each individually fed to the model. The processing time for the first frame is omitted as, due to 
+%% cache warmup, it is uncharacteristically high and doesn't
+%% accurately represent the Picar's capabilities.  
+%% A WCET of 50 ms, or 20 Hz, is used as a baseline to assess the platform's ability to complete all 
+%% necessary real-time operations in a timely manner.
+
+%% We seek to determine if the DeepPicar is capable of consistently executing all necessary functions 
+%% before their given deadlines. In the case of our platform, it has to process every given frame and get 
+%% the predicted angles from the model within a WCET of 50 ms. In order to determine if this was achievable, 
+%% we tested the platform by running a single model that utilized all 4 cpu cores and measured the time it 
+%% took for each frame to be processed. The performance of the DeepPicar can be seen in Figure 6. We found 
+%% that DeepPicar was able to completely process the vast majority of the provided frames within 50 ms, and 
+%% was unable to do so for very few frames.
+
+%% In our platform, three main real-time tasks are performed during autonomous operation. In order, these 
+%% operations are: (1) capturing and reading the input frame from the designated camera or video stream, 
+%% (2) preprocessing the acquired frame so that it is compatible with the DNN, and (3) feeding the frame 
+%% to, and getting the angle prediction from, the model. We aim to determine which operation(s), if any, 
+%% require the most time to be performed.
+
+%% In order to determine which operation(s) take the longest to execute, we measured the time it 
+%% took for each step to be completed. For this experiment, all four of the Pi's cpu cores were utilized, 
+%% and only one model was run. As is shown in Table II, the angle prediction operation consumes the 
+%% majority of the processing for each frame. Furthermore, the time it takes for the operation to 
+%% complete is volatile, and can range anywhere between 30 ms and 50 ms for any particular frame. On the 
+%% other hand, both the frame capture and preprocessing operations take substantially less time 
+%% and are relatively more consistent in their times, at 2 ms and 3 ms, respectively.

### 2017-10-04 paper/evaluation.tex
figure position
perf_vs_modelcnt

@@
   \centering
-  \includegraphics[width=.5\textwidth]{Total_Processing_Time}
+  \includegraphics[width=.45\textwidth]{Total_Processing_Time}
   \caption{DeepPicar's control loop processing times over 1000 input image frames.}
@@
 
-\subsection{Multimodel Performance}
-We also test the capability of the DeepPicar to run multiple models at the same time, and whether each 
-model can successfully perform within the given deadline. Specifically, the platform is tested in the 
-cases of running 2 and 4 models simultaneously. For each case, all models are allocated an equal number 
-of cores. That is, 2 models are given 2 cores each and 4 models are given 1 core each.
+\subsection{Effect of Co-scheduling Multiple DNN Models}
+
+We also test the capability of the DeepPicar to run multiple models at
+the same time, and whether each model can successfully perform within
+the given deadline. Specifically, the platform is tested in the cases
+of running 2 and 4 models simultaneously. For each case, all models
+are allocated an equal number of cores. That is, 2 models are given 2
+cores each and 4 models are given 1 core each.
 
@@
   \centering
-  \includegraphics[width=.5\textwidth]{figs/2ModelChart}
-  \caption{ Change in inferencing time when 2 models are run concurrently. }
+  \includegraphics[width=.45\textwidth]{figs/perf_vs_modelcnt}
+  \caption{Timing impact of co-scheduling multiple DNNs. X-axis shows
+    the system configuration: \#of DNN models x \#of CPU cores\/DNN. For
+    example, '4x1' means running four DNN models each of which is
+    assigned to run on one CPU core, where as '2x2' means running two DNN
+    models, each of which is assigned to run on two CPU cores.}
+  \label{fig:perf-vs-multimodel}
 \end{figure}
+
+%% \begin{figure}[h]
+%%   \centering
+%%   \includegraphics[width=.5\textwidth]{figs/2ModelChart}
+%%   \caption{ Change in inferencing time when 2 models are run concurrently. }
+%% \end{figure}
 

### 2017-10-04 paper/evaluation.tex
revision evaluation c. multimodel:

@@
 
-We also test the capability of the DeepPicar to run multiple models at
-the same time, and whether each model can successfully perform within
-the given deadline. Specifically, the platform is tested in the cases
-of running 2 and 4 models simultaneously. For each case, all models
-are allocated an equal number of cores. That is, 2 models are given 2
-cores each and 4 models are given 1 core each.
+In this experiment, we launch multiple instances of DeepPicar's DNN
+model at the same time and measure its impact to their inference
+timings.
+
+%% We also test the capability of the DeepPicar to run multiple models at
+%% the same time, and whether each model can successfully perform within
+%% the given deadline. Specifically, the platform is tested in the cases
+%% of running 2 and 4 models simultaneously. For each case, all models
+%% are allocated an equal number of cores. That is, 2 models are given 2
+%% cores each and 4 models are given 1 core each.
 
@@
   \includegraphics[width=.45\textwidth]{figs/perf_vs_modelcnt}
-  \caption{Timing impact of co-scheduling multiple DNNs. X-axis shows
-    the system configuration: \#of DNN models x \#of CPU cores\/DNN. For
-    example, '4x1' means running four DNN models each of which is
-    assigned to run on one CPU core, where as '2x2' means running two DNN
-    models, each of which is assigned to run on two CPU cores.}
+  \caption{Timing impact of co-scheduling multiple DNNs. 1x1: one DNN
+    model using one CPU core; 4x1: four DNN models each using one CPU;
+    1x2: one DNN model using two CPU cores; 2x2: two DNN models each
+    using two CPU cores.} 
   \label{fig:perf-vs-multimodel}
@@
 %% \end{figure}
+
+Figure~\ref{fig:perf-vs-multimodel} shows the results. In the figure,
+X-axis shows the system configuration: \#of DNN models x \#of CPU
+cores\/DNN. For example, '4x1' means running four DNN models each of
+which is assigned to run on one CPU core, whereas '2x2' means running
+two DNN models, each of which is assigned to run on two CPU
+cores. The Y-axis shows the average inference timing.
+Two bars in the left show the impact of co-scheduling four DNN
+models. Compared to executing a single DNN model on one CPU core
+(1x1), when four DNN models are co-scheduled (4x1), each model
+suffers an average inference time increase of approximately 15 ms,
+$\sim$24\%. On the other hand, when two DNN models, each using two CPU
+cores, are co-scheduled, the average inference timing is increased by
+about 7 ms, or 10\%, comapred to the baseline of running one model
+using two CPU cores.
+
+%% The results for the 2 model test are outlined in Figure 3 and Table
+%% IV, and the 4 model test is summarized in  Figure 4 Table V. In the
+%% the 2 model test, both of the models showed average inference time
+%% increases of around  5-7 ms, $\sim$10\%, when compared to a baseline
+%% of 1 model running on two cores.
 
@@
 
-The results for the 2 model test are outlined in Figure 3 and Table IV, and the 4 model test is summarized in 
-Figure 4 Table V. In the the 2 model test, both of the models showed average inference time increases of around 
-5-7 ms, $\sim$10\%, when compared to a baseline of 1 model running on two cores. The difference was even 
-greater in the 4 model test, as each model displayed an average inference time increase of 
-approximately 15 ms, $\sim$30\%, when compared to a single model running on 1 core.
-
-The increase in inference times, however, was not the result of increased cache misses due to 
-additional accesses. For all models, the number of L1 misses was always $\sim$1.6\% of all 
-references. In the 2 model test, L2 cache misses remained at $\sim$11\% of all references, and each 
-model in the 4 model test missed $\sim$13\% of all L2 references.
+The increase in inference times, however, was not the result of
+increased cache misses due to additional accesses. For all models, the
+number of L1 misses was always $\sim$1.6\% of all  references. In the
+2 model test, L2 cache misses remained at $\sim$11\% of all
+references, and each model in the 4 model test missed $\sim$13\% of
+all L2 references.
 

### 2017-10-04 paper/evaluation.tex
eval D multimodel revision

@@
 BCM2837 SoC, which has a quad-core 64bit ARM Cortex-A53 cluster,
-running at up to 1.2GHz. The chip also includes Broadcom's Videocore IV
+running at up to 1.2GHz. Each core has 16K private I\&D caches, and all
+cores share a 512KB L2 cache.
+The chip also includes Broadcom's Videocore IV
 GPU, although we did not use the GPU in our evaluation due to the lack
@@
 model at the same time and measure its impact to their inference
-timings.
+timings. In particular, we are interested in how shared resource
+contention affect inference timing.
 
@@
     using two CPU cores.} 
-  \label{fig:perf-vs-multimodel}
+  \label{fig:perf-vs-modelcnt}
 \end{figure}
@@
 %% \end{figure}
+%% \begin{figure}[h]
+%%   \centering
+%%   \includegraphics[width=.5\textwidth]{figs/4ModelChart}
+%%   \caption{Change in inferencing time when 4 models are run simultaneously. }
+%% \end{figure}
 
-Figure~\ref{fig:perf-vs-multimodel} shows the results. In the figure,
+Figure~\ref{fig:perf-vs-modelcnt} shows the results. In the figure,
 X-axis shows the system configuration: \#of DNN models x \#of CPU
@@
 
-\begin{table}[h]
-\centering
-  \begin{tabular} {| l | l | l |}
-  \hline
-  \textbf{num models} & 1 & 2 \\ \hline
-  \textbf{L1 refs} &  3.04E+10 &  3.04E+10 \\ \hline
-  \textbf{L1 misses} & 4.78E+08 & 4.91E+08 \\ \hline
-  \textbf{L1 miss \%} & 1.58 & 1.61 \\ \hline
-  \textbf{L2 refs} & 3.31E+09 &  3.91E+09 \\ \hline
-  \textbf{L2 misses} & 3.68E+08 & 4.62E+08 \\ \hline
-  \textbf{L2 miss \%} & 11.12 & 10.88 \\ \hline
-  \end{tabular}
-  \caption{ Effect of 2 simulatneous models on cache accesses. }
-\end{table}
+%% \begin{table}[h]
+%% \centering
+%%   \begin{tabular} {| l | l | l |}
+%%   \hline
+%%   \textbf{num models} & 1 & 2 \\ \hline
+%%   \textbf{L1 refs} &  3.04E+10 &  3.04E+10 \\ \hline
+%%   \textbf{L1 misses} & 4.78E+08 & 4.91E+08 \\ \hline
+%%   \textbf{L1 miss \%} & 1.58 & 1.61 \\ \hline
+%%   \textbf{L2 refs} & 3.31E+09 &  3.91E+09 \\ \hline
+%%   \textbf{L2 misses} & 3.68E+08 & 4.62E+08 \\ \hline
+%%   \textbf{L2 miss \%} & 11.12 & 10.88 \\ \hline
+%%   \end{tabular}
+%%   \caption{ Effect of 2 simulatneous models on cache accesses. }
+%% \end{table}
 
-The increase in inference times, however, was not the result of
-increased cache misses due to additional accesses. For all models, the
-number of L1 misses was always $\sim$1.6\% of all  references. In the
-2 model test, L2 cache misses remained at $\sim$11\% of all
-references, and each model in the 4 model test missed $\sim$13\% of
-all L2 references.
+%% \begin{table}[h]
+%% \centering
+%%   \begin{tabular} {| l | l | l |}
+%%   \hline
+%%   \textbf{num models} & 1 & 4 \\  \hline
+%%   \textbf{L1 refs} & 2.78E+10 & 2.79E+10 \\ \hline
+%%   \textbf{L1 misses} & 4.36E+08 & 4.53E+08 \\ \hline
+%%   \textbf{L1 miss \%} & 1.57 & 1.63 \\ \hline
+%%   \textbf{L2 refs} & 2.83E+09 & 3.36E+09 \\ \hline 
+%%   \textbf{L2 misses} & 3.59E+08 & 4.43E+08 \\ \hline
+%%   \textbf{L2 miss \%} & 12.68 & 13.19 \\ 
+%% \hline
+%%   \end{tabular}
+%%   \caption{ Effect of 4 concurrent models on cache accesses. }
+%% \end{table}
 
@@
   \centering
-  \includegraphics[width=.5\textwidth]{figs/4ModelChart}
-  \caption{Change in inferencing time when 4 models are run simultaneously. }
+  \includegraphics[width=.45\textwidth]{figs/l2missrate_vs_modelcnt}
+  \caption{L2 cache miss rates of different neural network and core
+    assignments. X-axis is the same as Figure~\ref{fig:perf-vs-modelcnt}.} 
+  \label{fig:l2missrate-vs-modelcnt}
 \end{figure}
 
-\begin{table}[h]
-\centering
-  \begin{tabular} {| l | l | l |}
-  \hline
-  \textbf{num models} & 1 & 4 \\  \hline
-  \textbf{L1 refs} & 2.78E+10 & 2.79E+10 \\ \hline
-  \textbf{L1 misses} & 4.36E+08 & 4.53E+08 \\ \hline
-  \textbf{L1 miss \%} & 1.57 & 1.63 \\ \hline
-  \textbf{L2 refs} & 2.83E+09 & 3.36E+09 \\ \hline 
-  \textbf{L2 misses} & 3.59E+08 & 4.43E+08 \\ \hline
-  \textbf{L2 miss \%} & 12.68 & 13.19 \\ 
-\hline
-  \end{tabular}
-  \caption{ Effect of 4 concurrent models on cache accesses. }
-\end{table}
+These increases in inference times in the co-scheduled scenarios are
+expected and likely caused by contention in the shared hardware
+resources such as the shared L2 cache and/or the DRAM controller.
+To further analyze the source of contention, we use hardwawre
+performance counters of the processor. Specifically, we measure L2
+miss rates of the DNN models first in isolation and then after
+co-scheduling another models. If the shared L2 cache is the primary
+source of inteference, then the measured L2 miss rates would
+increase. Figure~\ref{fig:l2missrate-vs-modelcnt} shows the results.
+As can be see in the figure, L2 miss rates are largely unchanged
+regardless whether multiple models are co-scheduled or not. This
+suggests that the shared L2 cache is not the bottleneck that caused
+execution time increases. In other words, DNN models appear to be not
+sensitive to the shared L2 cache space. % L2 partitioning is probably
+                                % not going to be useful.
+Instead, we hypothesize that it is likely caused by the memory
+controller---the only other major shared hardware source---where
+memory requests from different CPU cores contend, which would result
+in increased memory access latency. While some Intel processors
+provide uncore hardware counters that can measure average memory
+access latency~\cite{ye2016maracas}, we were not able to identify
+whether such hardware counters exist in the BCM2837 processor of
+Raspberry Pi 3 due to the lack of documentation. Instead, in the next
+experiment, we use memory intensive synthetic benchmark to test the
+hypothesis.
 
-In all multimodel tests, it was found that a greater number of models run simultaneously resulted in 
-interference that led to a noticable increase in the average inferencing time that wasn't 
-attributable to a change in cache misses. Instead, the overall time increases can most likely be 
-traced to an increase in cache latency. Even if the models had a consistent number of cache hits, it 
-is highly probable that the contention for the cache increased the access time for each model, 
-consequently increasing the time it took for the models to execute their operations.
+%% The increase in inference times, however, was not the result of
+%% increased cache misses due to additional accesses. For all models, the
+%% number of L1 misses was always $\sim$1.6\% of all  references. In the
+%% 2 model test, L2 cache misses remained at $\sim$11\% of all
+%% references, and each model in the 4 model test missed $\sim$13\% of
+%% all L2 references.
 
-In terms of real-time performance under a WCET of 50 ms, the DeepPicar was unsatisfactory in all 
-multimodel tests. On average, the 2 model test saw models miss their deadline by 6-8 ms, and the 4 
-model test was worse as each model went over their deadlines by 27-29 ms. As a result, it can be 
-ascertained that, if a WCET of 50 ms is required, DeepPicar would not be able to perform as necessary.
+%% In all multimodel tests, it was found that a greater number of models
+%% run simultaneously resulted in interference that led to a noticable
+%% increase in the average inferencing time that wasn't attributable to a
+%% change in cache misses. Instead, the overall time increases can most
+%% likely be traced to an increase in cache latency. Even if the models
+%% had a consistent number of cache hits, it is highly probable that the
+%% contention for the cache increased the access time for each model,
+%% consequently increasing the time it took for the models to execute
+%% their operations.
+
+%% In terms of real-time performance under a WCET of 50
+%% ms, the DeepPicar was unsatisfactory in all multimodel tests. On
+%% average, the 2 model test saw models miss their deadline by 6-8 ms,
+%% and the 4 model test was worse as each model went over their deadlines
+%% by 27-29 ms. As a result, it can be ascertained that, if a WCET of 50
+%% ms is required, DeepPicar would not be able to perform as necessary.
 

### 2017-10-04 paper/evaluation.tex

@@
 Raspberry Pi 3 due to the lack of documentation. Instead, in the next
-experiment, we use memory intensive synthetic benchmark to test the
+experiment, we use memory intensive synthetic benchmarks to test the
 hypothesis.
@@
 
-\subsection{Benchmark Performance}
-In order to determine how cache misses affects the performance of the DeepPicar, we test its ability by 
-running synthetic benchmarks concurrently with the model. In each experiment, we run a single model on a 
-number of cores, and a synthetic benchmark on the remaining idle cores. The effects of the benchmarks can
-seen in Figure 9 and Table VI.
+\subsection{Effect of Co-scheduling Memory Performance Hogs}
+
+In order to determine how contended DRAM requests affect the DNN
+inference timing of the DeepPicar, we use a synthetic memory
+intensive benchmark from the IsolBench suite~\cite{Valsan2016}.
+
+we test its ability by  running synthetic benchmarks
+concurrently with the model. In each experiment, we run a single model
+on a  number of cores, and a synthetic benchmark on the remaining idle
+cores. The effects of the benchmarks can seen in Figure 9 and Table
+VI.
 

### 2017-10-04 paper/evaluation.tex

@@
 
-we test its ability by  running synthetic benchmarks
-concurrently with the model. In each experiment, we run a single model
-on a  number of cores, and a synthetic benchmark on the remaining idle
-cores. The effects of the benchmarks can seen in Figure 9 and Table
-VI.
+In each experiment, we run a single DNN model one core, and
+co-scheduled an increasing number of the memory intensive synthetic
+benchmark on the remaining idle cores.
+Figure~\ref{fig:}
+The effects of the benchmarks can seen in Figure 9 and Table VI.
 
@@
   \includegraphics[width=.5\textwidth]{figs/BenchmarkChart}
-  \caption{ Effect of benchmarks on inferencing time. }
+  \caption{Effect of benchmarks on inferencing time.}
+  \label{fig:}
 \end{figure}

### 2017-10-04 paper/evaluation.tex
eval. E. bandwidth revision

@@
 co-scheduled an increasing number of the memory intensive synthetic
-benchmark on the remaining idle cores.
-Figure~\ref{fig:}
-The effects of the benchmarks can seen in Figure 9 and Table VI.
+benchmarks~\footnote{We use the \emph{Bandwidth} benchmark in the
+  IsolBench suite, with the following command line parameters: \texttt{\$
+  bandwdith -a write -m 16384}}, on the remaining idle cores.
 
@@
   \centering
-  \includegraphics[width=.5\textwidth]{figs/BenchmarkChart}
-  \caption{Effect of benchmarks on inferencing time.}
+  \includegraphics[width=.45\textwidth]{figs/perf_vs_bandwidth}
+  \caption{Effect of memory performance hogs on the DNN
+    inferencing. The DNN model uses Core 0 and memory-hog co-runnes
+    use the rest of the cores.}
   \label{fig:}
@@
 
-\begin{table}[h]
-\centering
-  \begin{tabular} {| l | l | l | l | l | l |}
-  \hline
-  \textbf{num cores} & 1 & 2 & 3 \\ \hline
-  \textbf{L1 refs} & 3.16E+10 & 3.17E+10 & 3.06E+10 \\ \hline
-  \textbf{L1 misses} & 5.41E+08 & 5.35E+08 & 4.98E+08 \\ \hline
-  \textbf{L1 miss \%} & 1.71 & 1.69 & 1.62 \\ \hline
-  \textbf{L2 refs} & 3.49E+09 & 3.23E+09 & 3.21E+09 \\ \hline
-  \textbf{L2 misses} & 8.05E+08 & 7.54E+08 & 4.98E+08 \\ \hline
-  \textbf{L2 miss \%} & 23.05 & 23.37 & 15.51 \\ 
-  \hline
-  \end{tabular}
-  \caption{ Effect of benchmarks on cache accesses. }
-\end{table}
+%% \begin{table}[h]
+%% \centering
+%%   \begin{tabular} {| l | l | l | l | l | l |}
+%%   \hline
+%%   \textbf{num cores} & 1 & 2 & 3 \\ \hline
+%%   \textbf{L1 refs} & 3.16E+10 & 3.17E+10 & 3.06E+10 \\ \hline
+%%   \textbf{L1 misses} & 5.41E+08 & 5.35E+08 & 4.98E+08 \\ \hline
+%%   \textbf{L1 miss \%} & 1.71 & 1.69 & 1.62 \\ \hline
+%%   \textbf{L2 refs} & 3.49E+09 & 3.23E+09 & 3.21E+09 \\ \hline
+%%   \textbf{L2 misses} & 8.05E+08 & 7.54E+08 & 4.98E+08 \\ \hline
+%%   \textbf{L2 miss \%} & 23.05 & 23.37 & 15.51 \\ 
+%%   \hline
+%%   \end{tabular}
+%%   \caption{ Effect of benchmarks on cache accesses. }
+%% \end{table}
 
-The presence of the benchmark(s) had a noticable effect on the DeepPicar, with all tests showing 
-increases in inferencing operations. The least change happened when the model was run on 3 cores, and a 
-single benchmark was run on the final core. Even with a slight increase in the number of L2 cache 
-misses, inferencing was able to be completed in under 100 ms. The other tests, however, showed dramatic 
-changes when the model was run on 1 and 2 cores. In both instances, L2 misses rose to $\sim$23\% of all 
-references. The additional benchmark in the 1 core test was even more detrimental as the average 
-inferencing time was at least 150 ms longer than when the model used 2 cores.
+Figure~\ref{fig:} shows the normalized execution time and L2 miss-rate
+of the DNN model running on the Core 0 as a function of the number of
+co-scheduled memory intensive synthetic benchmarks. First, as we
+increase the number of co-runners, the DNN model's execution times are
+increased---by up to 9.6X---even though the DNN model is running on a
+dedicated core (Core0). On the other hand, the DNN model's L2
+cache-miss rates are not increase as much. This suggests that
+the DNN model's exeuction increase cannot be fully explained by
+increased L2 cache-misses. Instead, as we hypothesized in the previous
+experiment, the increased memory pressure from the co-scheduled memory
+intensive benchmarks is the primary cause of the DNN model's execution
+time increase. Therefore, we conclude that DeepPicar's DNN model is
+more senstive to DRAM access latency than L2 cache space.
 
-With the introduction of synthetic benchmarks, the DeepPicar failed to comply with the 50 ms WCET across 
-all tests run. Even in the best case, inferencing operations take twice as long as the deadline to 
-complete. As a result, if benchmarks were required to run during autonomous operation, the DeepPicar 
-wouldn't be capable of meeting its deadlines.
+This observation suggests that shared cache partitioning
+techniques~\cite{Gracioli2015,Kim2016} may not be effective isolation
+solutions for DeepPicar as its AI workload is more sensitive to memory
+performance. Instead, memory controller focused isolation solutions,
+either hardware or software-based ones (e.g.,~\cite{Guo2017,Yun2013}),
+may be more important. Although our observation is made on a single
+hardware platform running on a single DNN workload, we suspect that
+many AI workloads may exhibit similar characteristics.
+
+%% The presence of the benchmark(s) had a noticable effect on the
+%% DeepPicar, with all tests showing increases in inferencing
+%% operations. The least change happened when the model was run on 3
+%% cores, and a single benchmark was run on the final core. Even with a
+%% slight increase in the number of L2 cache misses, inferencing was able
+%% to be completed in under 100 ms. The other tests, however, showed
+%% dramatic changes when the model was run on 1 and 2 cores. In both
+%% instances, L2 misses rose to $\sim$23\% of all references. The
+%% additional benchmark in the 1 core test was even more detrimental as
+%% the average inferencing time was at least 150 ms longer than when the
+%% model used 2 cores.
+
+%% With the introduction of synthetic benchmarks, the DeepPicar failed to
+%% comply with the 50 ms WCET across all tests run. Even in the best
+%% case, inferencing operations take twice as long as the deadline to
+%% complete. As a result, if benchmarks were required to run during
+%% autonomous operation, the DeepPicar wouldn't be capable of meeting its
+%% deadlines.
 
 \subsection{Summary}
-We found that the DeepPicar is capable of successfully completing all necessary inferencing 
-operations within a given deadline of 50 ms. Between these operations, it was discovered that the 
-angle prediction time of the model was the dominating step in the processing time of a single frame. 
-The other operations were found to execute in very little time, with either one taking 5 ms at most.
-
-When running a single model, DeepPicar performed the best when it used all 4 cpu cores, as it was 
-able to process a frame in 43 ms on average. This feat can still be accomplished when running a 
-single model on 3 cpu cores, and is also possible when using 2 cores. The only time in which the Pi 
-can't run a single model under 50 ms is when only 1 core is used. This was also the case when 
-multiple models were run concurrently as no models were able to consistently complete inference 
+We found that the DeepPicar is capable of successfully completing all
+necessary inferencing operations within a given deadline of 50
+ms. Between these operations, it was discovered that the angle
+prediction time of the model was the dominating step in the processing
+time of a single frame. The other operations were found to execute in
+very little time, with either one taking 5 ms at most. When running a
+single model, DeepPicar performed the best when it used all 4 cpu
+cores, as it was able to process a frame in 43 ms on average. This
+feat can still be accomplished when running a single model on 3 cpu
+cores, and is also possible when using 2 cores. The only time in which
+the Pi can't run a single model under 50 ms is when only 1 core is
+used. This was also the case when multiple models were run
+concurrently as no models were able to consistently complete inference
 operations in under 50 ms. 
-
-However, the DeepPicar was shown to be capable of handling other potential WCETs. If given a larger 
-deadline, the capabilities of DeepPicar as an autonomous vehicle platform would be greatly increased. 
-For example, if tested with a deadline of 66.67 ms, or 15 fps, the platform would have passed several 
-of the experiments performed. In actuality, the DeepPicar would have only been unsatisfactory in the 
-4 model test where it would have missed the deadlines by $\sim$10 ms, and the benchmark tests, where 
+However, the DeepPicar was shown to be capable of handling other
+potential WCETs. If given a larger deadline, the capabilities of
+DeepPicar as an autonomous vehicle platform would be greatly
+increased. For example, if tested with a deadline of 66.67 ms, or 15
+fps, the platform would have passed several of the experiments
+performed. In actuality, the DeepPicar would have only been
+unsatisfactory in the 4 model test where it would have missed the
+deadlines by $\sim$10 ms, and the benchmark tests, where  
 it would still miss by $\sim$33 ms at best.
@@
 \subsection{Performance Requirements}
-In the utilization of the Raspberry Pi 3 in our platform, there are a few factors that need to be 
-considered and/or enforced in order to guarantee that the Pi is able to consistently perform at a 
-desired level. Specifically, these issues all have the potential to negatively affect the cpu clock 
-speed/frequency, which would result in decreased performance. While, in the above experiments, the cpu 
-operated at a preferred clock speed of 1.2 GHz, it is entirely possible for the cpu to operate at a 
-lower frequency if the following problems are not taken into account.
+In the utilization of the Raspberry Pi 3 in our platform, there are a
+few factors that need to be considered and/or enforced in order to
+guarantee that the Pi is able to consistently perform at a desired
+level. Specifically, these issues all have the potential to negatively
+affect the cpu clock speed/frequency, which would result in decreased
+performance. While, in the above experiments, the cpu operated at a
+preferred clock speed of 1.2 GHz, it is entirely possible for the cpu
+to operate at a lower frequency if the following problems are not
+taken into account.
 
-The most notable issue that can affect the cpu clock speed is that of the power supplied to the 
-Raspberry Pi. In essence, it is necessary that the Pi be supplied with 2 Amps, as any less could 
-hinder the Pi's ability to maintain a 1.2 GHz frequency. In experiments conducted with a power supply that 
-only provided 1 Amp, the Pi was unable to sustain a 1.2 GHz clock speed and, instead, fluctuated 
-between operating at 600 MHz and 1.2 GHz. As a result, it is necessary, or at least highly 
-recommended, that the power supply used for the Raspberry Pi 3 be capable of outputting 2 Amps, 
-otherwise optimal performance isn't guaranteed.
-
-Another factor that can affect clock speed is that of the cpu's temperature. Some model operations can 
-be computationally intensive, thus it is possible for the temperature of the cpu to become relatively 
-high. This can be especially problematic in situations where multiple models are running 
-simultaneously on the Pi. Consequently, thermal throttling may be used to decrease the clock 
-speed so that the cpu temperature stays at a safe level. Thus the Raspberry Pi may not be suited 
-for prolonged use, especially in cases where the workload is relatively larger, such as running multiple 
-models. Rather, the Pi seems to be better suited for running in set periods, after which it is turned 
-off or made idle so that the cpu is allowed time to cool down.
+The most notable issue that can affect the cpu clock speed is that of
+the power supplied to the Raspberry Pi. In essence, it is necessary
+that the Pi be supplied with 2 Amps, as any less could hinder the Pi's
+ability to maintain a 1.2 GHz frequency. In experiments conducted with
+a power supply that only provided 1 Amp, the Pi was unable to sustain
+a 1.2 GHz clock speed and, instead, fluctuated between operating at
+600 MHz and 1.2 GHz. As a result, it is necessary, or at least highly
+recommended, that the power supply used for the Raspberry Pi 3 be
+capable of outputting 2 Amps, otherwise optimal performance isn't
+guaranteed.  Another factor that can affect clock speed is that of the
+cpu's temperature. Some model operations can be computationally
+intensive, thus it is possible for the temperature of the cpu to
+become relatively high. This can be especially problematic in
+situations where multiple models are running  simultaneously on the
+Pi. Consequently, thermal throttling may be used to decrease the clock
+speed so that the cpu temperature stays at a safe level. Thus the
+Raspberry Pi may not be suited for prolonged use, especially in cases
+where the workload is relatively larger, such as running multiple
+models. Rather, the Pi seems to be better suited for running in set
+periods, after which it is turned  off or made idle so that the cpu is
+allowed time to cool down.
 

### 2017-10-04 paper/evaluation.tex
eval summary

@@
 time decreases---from 61.96 ms on a single core to 42.67 ms on four
-cores. However, the improvement is far from being ideal as we expected
+cores (a 30\% improvement). However, the improvement is far from ideal
 linear scaling. In particular, from 2 cores to 3 cores, the
@@
 \subsection{Summary}
-We found that the DeepPicar is capable of successfully completing all
-necessary inferencing operations within a given deadline of 50
-ms. Between these operations, it was discovered that the angle
-prediction time of the model was the dominating step in the processing
-time of a single frame. The other operations were found to execute in
-very little time, with either one taking 5 ms at most. When running a
-single model, DeepPicar performed the best when it used all 4 cpu
-cores, as it was able to process a frame in 43 ms on average. This
-feat can still be accomplished when running a single model on 3 cpu
-cores, and is also possible when using 2 cores. The only time in which
-the Pi can't run a single model under 50 ms is when only 1 core is
-used. This was also the case when multiple models were run
-concurrently as no models were able to consistently complete inference
-operations in under 50 ms. 
-However, the DeepPicar was shown to be capable of handling other
-potential WCETs. If given a larger deadline, the capabilities of
-DeepPicar as an autonomous vehicle platform would be greatly
-increased. For example, if tested with a deadline of 66.67 ms, or 15
-fps, the platform would have passed several of the experiments
-performed. In actuality, the DeepPicar would have only been
-unsatisfactory in the 4 model test where it would have missed the
-deadlines by $\sim$10 ms, and the benchmark tests, where  
-it would still miss by $\sim$33 ms at best.
+So far, we have systematically evaluated DeepPicar's real-time
+characteristics from the perspective of end-to-end deep learning based
+real-time control and made several observations.
+
+First, we find that DeepPicar's computing platform,
+the Raspberry Pi 3 Model B, offers adquate computing capacity to
+to perform real-time control of the RC car at 20 Hz frequency (or
+50ms per control loop). Given the complexity of the DNN we use, we are
+pleasantly suprised by this finding. The time breakdown has shown that
+the DNN inferencing operation, perforemd by the Tensorflow library, is
+dominating the execution time, which is expected.
+
+Second, we find that the scalability of Tensorflow's DNN inference 
+implementation is not ideal: We find that using all four cores are 
+only about 30\% better than using just a single core.
+
+Third, we find that consolidating multiple DNN models---on different CPU
+cores---is feasible based on the following two observations: (1)
+DNN performance on a single is not much worse than on multicore; (2)
+multiple DNN models running simultaneously do not cause severe
+interference with each other.
+
+Lastly, we however find that consolidating memory (DRAM) intensive
+applications could jeopadize DNN performance because DNN performance
+appear very sensitive to memory performance; we observe up to 9.6X
+slowdown in DNN performance by co-scheduling synthetic memory
+bandwidth intensive applications on idle cores.
+
+%% is capable of successfully completing all
+%% necessary inferencing operations within a given deadline of 50
+%% ms. Between these operations, it was discovered that the angle
+%% prediction time of the model was the dominating step in the processing
+%% time of a single frame. The other operations were found to execute in
+%% very little time, with either one taking 5 ms at most
+
+%% When running a
+%% single model, DeepPicar performed the best when it used all 4 cpu
+%% cores, as it was able to process a frame in 43 ms on average. This
+%% feat can still be accomplished when running a single model on 3 cpu
+%% cores, and is also possible when using 2 cores. The only time in which
+%% the Pi can't run a single model under 50 ms is when only 1 core is
+%% used. This was also the case when multiple models were run
+%% concurrently as no models were able to consistently complete inference
+%% operations in under 50 ms.
+
+%% However, the DeepPicar was shown to be capable of handling other
+%% potential WCETs. If given a larger deadline, the capabilities of
+%% DeepPicar as an autonomous vehicle platform would be greatly
+%% increased. For example, if tested with a deadline of 66.67 ms, or 15
+%% fps, the platform would have passed several of the experiments
+%% performed. In actuality, the DeepPicar would have only been
+%% unsatisfactory in the 4 model test where it would have missed the
+%% deadlines by $\sim$10 ms, and the benchmark tests, where  
+%% it would still miss by $\sim$33 ms at best.
 

### 2017-10-04 paper/evaluation.tex
reorg: add comparison section

@@
 including NVIDIA's Jetson TK2, later in
-Section~\ref{sec:eval-platforms}), the fact that the low-cost
+Section~\ref{sec:comparison}), the fact that the low-cost
 Raspberry Pi 3 can achieve similar real-time control performance is
@@
 \subsection{Summary}
-So far, we have systematically evaluated DeepPicar's real-time
+So far, we have evaluated DeepPicar's real-time
 characteristics from the perspective of end-to-end deep learning based
-real-time control and made several observations.
+real-time control, and made several observations.
 
@@
 the Raspberry Pi 3 Model B, offers adquate computing capacity to
-to perform real-time control of the RC car at 20 Hz frequency (or
-50ms per control loop). Given the complexity of the DNN we use, we are
-pleasantly suprised by this finding. The time breakdown has shown that
-the DNN inferencing operation, perforemd by the Tensorflow library, is
-dominating the execution time, which is expected.
+perform real-time control of the RC car at 20 Hz frequency (or
+50ms per control loop). Given the complexity of the DNN we use, we
+were pleasantly suprised by this finding. The time breakdown shows that
+the DNN inferencing operation, perforemd by the Tensorflow library,
+dominates the execution time, which is expected.
 
-Second, we find that the scalability of Tensorflow's DNN inference 
-implementation is not ideal: We find that using all four cores are 
+Second, we find that scalability of Tensorflow's DNN inference 
+implementation is limited: We find that using all four cores are 
 only about 30\% better than using just a single core.
@@
 Third, we find that consolidating multiple DNN models---on different CPU
-cores---is feasible based on the following two observations: (1)
-DNN performance on a single is not much worse than on multicore; (2)
-multiple DNN models running simultaneously do not cause severe
-interference with each other.
+cores---is feasible as we find: (1) DNN performance using a single
+core is not much worse than using multiple cores; (2) multiple DNN
+models running simultaneously do not cause severe interference with
+each other.
 
-Lastly, we however find that consolidating memory (DRAM) intensive
-applications could jeopadize DNN performance because DNN performance
-appear very sensitive to memory performance; we observe up to 9.6X
-slowdown in DNN performance by co-scheduling synthetic memory
+Lastly, we however find that consolidating memory (DRAM) performance
+intensive applications could jeopadize DNN performance because DNN
+performance appear very sensitive to memory performance; we observe up
+to 9.6X slowdown in DNN performance by co-scheduling synthetic memory 
 bandwidth intensive applications on idle cores.
@@
 %% it would still miss by $\sim$33 ms at best.
-
-\subsection{Performance Requirements}
-In the utilization of the Raspberry Pi 3 in our platform, there are a
-few factors that need to be considered and/or enforced in order to
-guarantee that the Pi is able to consistently perform at a desired
-level. Specifically, these issues all have the potential to negatively
-affect the cpu clock speed/frequency, which would result in decreased
-performance. While, in the above experiments, the cpu operated at a
-preferred clock speed of 1.2 GHz, it is entirely possible for the cpu
-to operate at a lower frequency if the following problems are not
-taken into account.
-
-The most notable issue that can affect the cpu clock speed is that of
-the power supplied to the Raspberry Pi. In essence, it is necessary
-that the Pi be supplied with 2 Amps, as any less could hinder the Pi's
-ability to maintain a 1.2 GHz frequency. In experiments conducted with
-a power supply that only provided 1 Amp, the Pi was unable to sustain
-a 1.2 GHz clock speed and, instead, fluctuated between operating at
-600 MHz and 1.2 GHz. As a result, it is necessary, or at least highly
-recommended, that the power supply used for the Raspberry Pi 3 be
-capable of outputting 2 Amps, otherwise optimal performance isn't
-guaranteed.  Another factor that can affect clock speed is that of the
-cpu's temperature. Some model operations can be computationally
-intensive, thus it is possible for the temperature of the cpu to
-become relatively high. This can be especially problematic in
-situations where multiple models are running  simultaneously on the
-Pi. Consequently, thermal throttling may be used to decrease the clock
-speed so that the cpu temperature stays at a safe level. Thus the
-Raspberry Pi may not be suited for prolonged use, especially in cases
-where the workload is relatively larger, such as running multiple
-models. Rather, the Pi seems to be better suited for running in set
-periods, after which it is turned  off or made idle so that the cpu is
-allowed time to cool down.
-
-\subsection{System Comparisons}
-We compare the real time performance of the Raspberry Pi 3 to two other embedded computing platforms, 
-the Intel UP Board the NVIDIA Tegra TX2, to evaluate their efficacy in vision-based autonomous 
-vehicles. We conduct the same experiments done on the DeepPicar with each platform, and analyze the 
-results.
-
-\begin{figure}[h]
-  \centering
-  \includegraphics[width=.5\textwidth]{figs/system_multicore}
-  \caption{Inferencing time across platforms based on number of cores used.}
-  \label{fig:sys_core}
-\end{figure}
-
-The results of the multicore tests can be seen in Figure~\ref{fig:sys_core}. It was found that the 
-Intel UP Board performed better in all experiments. On average, it was able to perform inferencing 
-operations, at least, twice as fast as the Raspberry Pi 3. As a result, the UP Board was able to 
-satisfy the 50 ms WCET by a very clear margin (in the worst case, it would still meet it by $\sim$20 
-ms), and, thus, demonstrates that it is more capable of autonomous driving than the Raspberry Pi 3.
-
-The Intel UP Board also performed better in the multimodel tests, as is shown in 
-Figure~\ref{fig:sys_model}. Once again, the UP was able to complete inferecing operations in half the 
-time when compared to the Pi. Furthermore, the Intel UP Board was capable of satisfying the 50 ms 
-WCET, while the Raspberry Pi 3 was unable to do so. As such, the Intel UP Board displays greater 
-potential for the simultaneous execution of multiple models during self-driving operation, while the 
-Raspberry Pi 3 would struggle to do so.
-
-\begin{figure}[h]
-  \centering
-  \includegraphics[width=.5\textwidth]{figs/system_multimodel}
-  \caption{Inferencing time across platforms based on number of models run concurrently.}
-  \label{fig:sys_model}
-\end{figure}
-
-Finally, we examined if the performance of the Intel UP Board and Tegra TX2 would be affected by the 
-addition of the synthetic benchmarks. As summarized in Figure~\ref{fig;sys_bench}, the Intel UP Board 
-did experience a change in its inferencing times, but the increase was not as drastic as is seen by 
-the Raspberry Pi 3. in the worst case, the Intel UP Board produced times that were over two times as 
-large, whereas the Raspberry Pi 3 output times that were around 10 times greater. The Intel UP Board 
-was successful in completing inferencing operations in under 50 ms when benchmarks were run on a 
-maximum of 2 cores, while the Raspberry Pi 3 failed to do so when any benchmark was introduced. As 
-such, it can be concluded that the Intel UP Board would be more capable of running computationally 
-heavy processes during autonomous driving, and that the Pi wouldn't be able to do the same.
-
-\begin{figure}[h]
-  \centering
-  \includegraphics[width=.5\textwidth]{figs/system_benchmark}
-  \caption{Inferencing time across platforms based on number of cores used.}
-  \label{fig:sys_bench}
-\end{figure}
-
-In the comparison of the real-time capabilities of three embedded computing platforms, we found that 
-the Raspberry Pi 3 performed the worst. When compared to the Intel UP Board, inferencing on the Pi 
-took twice as long across all multicore and multimodel experiments. The difference was even more 
-noticeable in the the addition of computationally heavy synthetic benchmarks that had a much more 
-dire effect on the Pi.
 

### 2017-10-05 paper/main.tex
use hyperref package
reorg: add comparison section
Add draft abstract

@@
 
-%\usepackage[pdftex]{hyperref}
+\usepackage[pdftex]{hyperref}
 \usepackage{rotating,tabularx}
@@
 \begin{abstract}
-Abstract goes here.
+In order to effectively research autonomous vehicles, a real-time capable platform is required. Such 
+platforms can be hard to obtain, or build, as the necessary components can be expensive. In this paper, 
+we present a low-cost alternative platform that utilizes a Raspberry Pi 3 Model B, along with a single 
+external camera sensor, to perform all necessary self-driving operations. A convolutional neural network 
+(CNN) is employed and trained with input from a human driver to navigate a custom made track. We display 
+the real-time efficacy of our platform in experiments with varying amount of computational resources 
+and/or intensive AI workloads. To gain a better understanding of our platform's potential, we compare it 
+to other embedded computing platforms to determine which are suitable for autonomous driving research.
 \end{abstract}
@@
 %-------------------------------------------------------------------------
-\bibliographystyle{plain}
+\bibliographystyle{abbrv}
 \bibliography{reference}
+
+\appendix
 \end{document}

### 2017-10-05 paper/evaluation.tex
camera latency.

@@
 
-\subsection{Summary}
+\begin{table*}[h]
+  \centering
+  \begin{tabular}{|c|c|c|c|}
+    \hline
+    Item    & Raspberry Pi 3 (B)   & Intel UP                  & NVIDIA Jetson TX2\\
+    \hline
+            & BCM2837              & X5-Z8350 (Cherry Trail)   & Tegra TX2 \\
+    CPU     & 4x Cortex-A53@1.2GHz/512KB L2  &
+              4x Atom@1.92GHz/2MB L2 &
+              4x Cortex-A57@2GHz/2MB L2 \\
+            &              &              & 2x Denver@2.0GHz/2MB L2 (not used)  \\
+    \hline
+    GPU     &  VideoCore IV (not used)    &
+               Intel HD 400 Graphics (not used) &
+               Pascal 256 CUDA cores   \\
+    \hline
+    Memory  & 1GB LPDDR2   &  2GB DDR3L     & 8GB LPDDR4              \\
+    \hline
+  \end{tabular}
+  \caption{Compared embedded computing platforms}
+  \label{tbl:patforms}
+\end{table*}
+
+\subsection{Summary of the Findings}
 So far, we have evaluated DeepPicar's real-time

### 2017-10-05 paper/overview.tex
figure position
camera latency.

@@
 
-\begin{figure}[h]
+\begin{figure}[t]
   \centering
@@
 
-\begin{table}[h]
+\begin{table}[t]
   \centering
@@
 only supports three discrete angles---left (-30$^{\circ}$), center
-(0$^{\circ}$), and right (+30$^{\circ}$)---as control
-inputs. Currently, we approximate the network generated real-valued
+(0$^{\circ}$), and right (+30$^{\circ}$)---as control inputs.
+Currently, we approximate the network generated real-valued
 angle to the closest one of the three angles, which may
@@
 In the future, we plan to use different (more expensive) RC car
-platforms that can precisely control the car's steering angle. We
-would like to stress, however, that the use of different RC car
+platforms that can precisely control the car's steering angle.
+We would like to stress, however, that the use of different RC car
 platforms has no impact in terms of computational 
@@
 
-Nevertheless, the trained models were still able to achieve a reasonable
-degree of accuracy, successfully navigating several different tracks 
-we trained. The source code, build instruction, and a collection of
-self-driving videos of the DeepPicar can be found at:
-\url{https://github.com/heechul/picar}.
+Another issue we observed in training/testing the network, which could
+affect the network performance, is camera latency. In DeepPicar's
+context, the camera latency is the time from the time the camera
+sensor observes the scene to the time the computer actually read the
+digitized image data. Unfortunately, this time can be substantially
+long depending on the camera and the computer performance. We
+experimentally measure DeepPicar's camera latency as
+follows~\footnote{We follow the method and the tool found at
+  https://www.makehardware.com/2016/03/29/finding-a-low-latency-webcam/}.
+First, we point the DeepPicar's camera toward to a computer monitor, which is
+conntected to the Raspberry Pi 3's HDMI output. We then launch a
+simple program that overlay the current time with the camera's preview
+screen on the monitor. Then, infinite number of time strings will be
+shown on the monitor screen and the time difference between the two
+successive time strings represent the camera
+latency. Figure~\ref{fig:camera-latency} shows the captured moniotor
+screen. The measured camera latency is about 300~350 ms. This is
+significantly higher than a human's perception latency, which is known
+to be as fast as 13 ms~\cite{ThomasBurger2015}. Higher camera latency
+could negatively affect to the control performance as the DNN would
+analyze stale scenes. In the future, we plan to identify and use
+low-latency cameras.
+
+\begin{figure}[t]
+  \centering
+  \includegraphics[width=.4\textwidth]{figs/camera_latency}
+  \caption{Measuring DeepPicar's camera latency.}
+  \label{fig:camera-latency}
+\end{figure}
+
+% https://www.pubnub.com/blog/2015-02-09-how-fast-is-realtime-human-perception-and-technology/
+
+Despite aforementioned issues, the trained models were still able to
+achieve a reasonable degree of accuracy, successfully navigating
+several different tracks we trained. The source code, build
+instruction, and a collection of self-driving videos of the DeepPicar
+can be found at: \url{https://github.com/heechul/picar}.
 %% \begin{center}

### 2017-10-05 paper/overview.tex

@@
 
-Despite aforementioned issues, the trained models were still able to
+Despite the aforementioned issues, the trained models were still able to
 achieve a reasonable degree of accuracy, successfully navigating

### 2017-10-05 paper/appendix.tex
..

\subsection{System-level Factors Affecting Real-Time Performance}
In the utilization of the Raspberry Pi 3 in our platform, there are a
few factors that need to be considered and/or enforced in order to
guarantee that the Pi is able to consistently perform at a desired
level. Specifically, these issues all have the potential to negatively
affect the cpu clock speed/frequency, which would result in decreased
performance. While, in the above experiments, the cpu operated at a
preferred clock speed of 1.2 GHz, it is entirely possible for the cpu
to operate at a lower frequency if the following problems are not
taken into account.

The most notable issue that can affect the cpu clock speed is that of
the power supplied to the Raspberry Pi. In essence, it is necessary
that the Pi be supplied with 2 Amps, as any less could hinder the Pi's
ability to maintain a 1.2 GHz frequency. In experiments conducted with
a power supply that only provided 1 Amp, the Pi was unable to sustain
a 1.2 GHz clock speed and, instead, fluctuated between operating at
600 MHz and 1.2 GHz. As a result, it is necessary, or at least highly
recommended, that the power supply used for the Raspberry Pi 3 be
capable of outputting 2 Amps, otherwise optimal performance isn't
guaranteed.

Another factor that can affect clock speed is that of the
cpu's temperature. Some model operations can be computationally
intensive, thus it is possible for the temperature of the cpu to
become relatively high. This can be especially problematic in
situations where multiple models are running  simultaneously on the
Pi. Consequently, thermal throttling may be used to decrease the clock
speed so that the cpu temperature stays at a safe level. Thus the
Raspberry Pi may not be suited for prolonged use, especially in cases
where the workload is relatively larger, such as running multiple
models. Rather, the Pi seems to be better suited for running in set
periods, after which it is turned  off or made idle so that the cpu is
allowed time to cool down.

### 2017-10-05 paper/comparison.tex

%
% raspberry p3 spec: https://www.raspberrypi.org/magpi/raspberry-pi-3-specs-benchmarks/

\section{Embedded Computing Platform Comparison}\label{sec:comparison}

In this section, we compare three computing platforms---the Raspberry
Pi 3, the Intel UP~\footnote{http://www.up-board.org/up/} and NVIDIA
Jetson
TX2~\footnote{http://www.nvidia.com/object/embedded-systems-dev-kits-modules.html}---from
the point of view of supporting vision-based end-to-end deep learning
based autonomous vehicles.

Table~\ref{tbl:platforms}

Our basic approach is to use the same DeepPicar's software, and repeat
the experiments in Section~\ref{sec:evaluation} on each hardware
platform and compare the results.

\begin{figure}[h]
  \centering
  \includegraphics[width=.5\textwidth]{figs/system_multicore}
  \caption{Inferencing time across platforms based on number of cores used.}
  \label{fig:sys_core}
\end{figure}

The results of the multicore tests can be seen in Figure~\ref{fig:sys_core}. It was found that the 
Intel UP Board performed better in all experiments. On average, it was able to perform inferencing 
operations, at least, twice as fast as the Raspberry Pi 3. As a result, the UP Board was able to 
satisfy the 50 ms WCET by a very clear margin (in the worst case, it would still meet it by $\sim$20 
ms), and, thus, demonstrates that it is more capable of autonomous driving than the Raspberry Pi 3.

The Intel UP Board also performed better in the multimodel tests, as is shown in 
Figure~\ref{fig:sys_model}. Once again, the UP was able to complete inferecing operations in half the 
time when compared to the Pi. Furthermore, the Intel UP Board was capable of satisfying the 50 ms 
WCET, while the Raspberry Pi 3 was unable to do so. As such, the Intel UP Board displays greater 
potential for the simultaneous execution of multiple models during self-driving operation, while the 
Raspberry Pi 3 would struggle to do so.

\begin{figure}[h]
  \centering
  \includegraphics[width=.5\textwidth]{figs/system_multimodel}
  \caption{Inferencing time across platforms based on number of models run concurrently.}
  \label{fig:sys_model}
\end{figure}

Finally, we examined if the performance of the Intel UP Board and Tegra TX2 would be affected by the 
addition of the synthetic benchmarks. As summarized in Figure~\ref{fig;sys_bench}, the Intel UP Board 
did experience a change in its inferencing times, but the increase was not as drastic as is seen by 
the Raspberry Pi 3. in the worst case, the Intel UP Board produced times that were over two times as 
large, whereas the Raspberry Pi 3 output times that were around 10 times greater. The Intel UP Board 
was successful in completing inferencing operations in under 50 ms when benchmarks were run on a 
maximum of 2 cores, while the Raspberry Pi 3 failed to do so when any benchmark was introduced. As 
such, it can be concluded that the Intel UP Board would be more capable of running computationally 
heavy processes during autonomous driving, and that the Pi wouldn't be able to do the same.

\begin{figure}[h]
  \centering
  \includegraphics[width=.5\textwidth]{figs/system_benchmark}
  \caption{Inferencing time across platforms based on number of cores used.}
  \label{fig:sys_bench}
\end{figure}

In the comparison of the real-time capabilities of three embedded computing platforms, we found that 
the Raspberry Pi 3 performed the worst. When compared to the Intel UP Board, inferencing on the Pi 
took twice as long across all multicore and multimodel experiments. The difference was even more 
noticeable in the the addition of computationally heavy synthetic benchmarks that had a much more 
dire effect on the Pi.

### 2017-10-05 paper/overview.tex
..

--- 2017-10-05 paper/overview.tex
@@
 
-Despite aforementioned issues, the trained models were still able to
+Despite these issues, the trained models were still able to
 achieve a reasonable degree of accuracy, successfully navigating

### 2017-10-05 paper/intro.tex
Add draft summary of findings to intro

@@
 % how we trained (maybe moved to a later section)
-We apply a standard imitation learning methodology to train the cat to
+We apply a standard imitation learning methodology to train the car to
 follow tracks on the ground. We collect data for
@@
 % what are our findings?
-\fixme{TBD: Summary of our findings}
+From our experiments, we discover that the DeepPicar is capable of 
+completing control loops in under 50 ms, or 20 hz, and can do so 
+almost 99\% of the time. The number of cores used by the platform has 
+an unexpectedly low impact on its performance, only decreasing the 
+control loop times by 30\% when using 4 cores instead of 1.  Our 
+platform shows the potential to run multiple neural network models 
+simultaneously, as there are no indications of direct interference 
+between them. Instead, we show that memory intensive co-runners can 
+be detrimental to the real-time performance of the DeepPicar. 
+Finally, while we find that the Raspberry Pi 3 takes longer to 
+complete control loops when compared to other embedded computing 
+platforms, namely the Intel UP Board and the NVIDIA Tegra TX2, we 
+still conclude that the Pi 3 is a viable option for autonomous 
+vehicle research.
 

### 2017-10-05 paper/appendix.tex
Add draft training section to appendix
Also made some minor changes.

@@
+\subsection{DNN Training and Testing}
+We train a neural network model across a set of 30 completed runs on 
+the track seen in Figure~\ref{fig:track} by a human pilot. Half of 
+the runs saw the car driving one way along the track, while the 
+remaining half were of the car driving in the opposite direction on 
+the track. In total, we collected 2,556 frames for training and 2,609 
+frames for validation. In each training step, we use a batch size of 
+100 frames which are used to further develop the model. We do this 
+across 2,000 training steps. When a model was trained with the 
+aformentioned data, the training loss was 0.0188 and the validation 
+loss was 0.0132.
+
+\begin{figure}[t]
+  \centering
+  \includegraphics[width=.55\textwidth]{figs/TrainingLoss}
+  \caption{DeepPicar's control loop processing times over 1000 input 
+image frames.}
+  \label{fig:modelloss}
+\end{figure}
+
+We also trained models where center frames (where the car was going 
+straight) and curve frames (where the car was turning) were equally 
+selected in each batch, and were pleasantly surprised at the change 
+in the car's performance. With the second model, the car displayed a 
+greater ability to stay in the center of the track (on the white 
+tape). This was problematic, however, as the loss values changed with 
+the training loss being 0.009 and the validation loss becoming 
+0.0869, indicating that overfitting had become a problem. We believe, 
+though, that this can be rectified through the collection of 
+additional data for training and validation.
+
 \subsection{System-level Factors Affecting Real-Time Performance}

### 2017-10-05 paper/background.tex
Add draft training section to appendix
Also made some minor changes.

@@
 
-Using deep nueral networks involves two distinct
+Using deep neural networks involves two distinct
 phases~\cite{NVIDIA2015}. The first
 phase is \emph{training} during which the weights of the network is
-incrementally updated by backpropating errors it sees from the
+incrementally updated by backpropagating errors it sees from the
 training examples. Once the network is trained---i.e., the weights of

### 2017-10-05 paper/intro.tex
Add draft training section to appendix
Also made some minor changes.

@@
 any human intervention. Deep neural networks (DNNs) have been
-sucessfully applied in various perception and control tasks in
+successfully applied in various perception and control tasks in
 recent years, and they are important workloads for autonomous vehicles
@@
 cars~\cite{Bojarski2016} and other robots. It is expected that more
-DNN based Artifical Intelligence workloads may be used in future
+DNN based Artificial Intelligence workloads may be used in future
 autonomous vehicles.
@@
 % big problem
-Executing these AI workloads on a embedded computing platform that can
+Executing these AI workloads on an embedded computing platform that can
 be used in a car poses several additional challenges. First, many AI
@@
 real-time requirements. For example, latency in a vision-based object
-detection task may directly linked to safety of the vehicle. This
+detection task may be directly linked to safety of the vehicle. This
 requires a high computing capacity as well as means to guaranteeing

### 2017-10-05 paper/appendix.tex
Update control loop graph
Also changed location of training loss graph
Add (unlikely) graph of network
The graph ended up being very large. The one currently being used will
have to be changed since it is very hard to see.

@@
 aformentioned data, the training loss was 0.0188 and the validation 
-loss was 0.0132.
+loss was 0.0132. The change of the loss value over the course of model training can be seen in Figure~\ref{fig:modelloss}.
 
-\begin{figure}[t]
+\begin{figure}[h]
   \centering
   \includegraphics[width=.55\textwidth]{figs/TrainingLoss}
-  \caption{DeepPicar's control loop processing times over 1000 input 
-image frames.}
+  \caption{Change in loss value throughout training.}
   \label{fig:modelloss}
@@
 selected in each batch, and were pleasantly surprised at the change 
-in the car's performance. With the second model, the car displayed a 
+in the car's performance. With the second method, the car displayed a 
 greater ability to stay in the center of the track (on the white 
@@
 additional data for training and validation.
+
+\begin{figure}[h]
+  \centering
+  \includegraphics[width=.55\textwidth]{figs/graph_run=.pdf}
+  \caption{ Network visualization. }
+  \label{fig:}
+\end{figure}
+\fixme{Is all of the network graph necessary or only certain parts?}
 

### 2017-10-06 paper/comparison.tex
Add TX2 discussion/results to comparison
Also made some minor changes.

@@
   \centering
-  \includegraphics[width=.5\textwidth]{figs/system_multicore}
-  \caption{Inferencing time across platforms based on number of cores used.}
+  \includegraphics[width=.5\textwidth]{figs/compare_core}
+  \caption{Inferencing time across platforms based on number of cores 
+used.}
   \label{fig:sys_core}
@@
 
-The results of the multicore tests can be seen in Figure~\ref{fig:sys_core}. It was found that the 
-Intel UP Board performed better in all experiments. On average, it was able to perform inferencing 
-operations, at least, twice as fast as the Raspberry Pi 3. As a result, the UP Board was able to 
-satisfy the 50 ms WCET by a very clear margin (in the worst case, it would still meet it by $\sim$20 
-ms), and, thus, demonstrates that it is more capable of autonomous driving than the Raspberry Pi 3.
+The results of the multicore tests can be seen in 
+Figure~\ref{fig:sys_core}. It was found that the both the Tegra TX2 
+and the Intel UP Board performed better than the Raspberry Pi 3 in 
+all experiments. On average, the UP was able to perform inferencing 
+operations, at least, twice as fast as the Raspberry Pi 3. The 
+difference was even greater with the TX2, as control loops, in the 
+worst case, were 3x faster using only the CPU and 10x quicker when 
+using the GPU. As a result, they were all able to satisfy the 50 ms 
+WCET by a very clear margin (in the worst case, the UP Board would 
+still meet it by $\sim$20 ms), and, thus, demonstrates that they are 
+more capable of autonomous driving than the Raspberry Pi 3.
 
-The Intel UP Board also performed better in the multimodel tests, as is shown in 
-Figure~\ref{fig:sys_model}. Once again, the UP was able to complete inferecing operations in half the 
-time when compared to the Pi. Furthermore, the Intel UP Board was capable of satisfying the 50 ms 
-WCET, while the Raspberry Pi 3 was unable to do so. As such, the Intel UP Board displays greater 
-potential for the simultaneous execution of multiple models during self-driving operation, while the 
-Raspberry Pi 3 would struggle to do so.
+The UP Board and TX2 also performed better in the multimodel tests, 
+as is shown in Figure~\ref{fig:sys_model}. Once again, they were able 
+to complete control loops in a fraction of the time when compared to 
+the Pi. Furthermore, both of the platforms satisfied the 50 ms WCET 
+in all experiments, while the Raspberry Pi 3 was unable to do so. As 
+such, the UP Board and TX2 display greater potential for the 
+simultaneous execution of multiple models during self-driving 
+operation, while the Raspberry Pi 3 would struggle to do so.
 
@@
   \centering
-  \includegraphics[width=.5\textwidth]{figs/system_multimodel}
-  \caption{Inferencing time across platforms based on number of models run concurrently.}
+  \includegraphics[width=.5\textwidth]{figs/compare_model}
+  \caption{Inferencing time across platforms based on number of 
+models run concurrently.}
   \label{fig:sys_model}
@@
 
-Finally, we examined if the performance of the Intel UP Board and Tegra TX2 would be affected by the 
-addition of the synthetic benchmarks. As summarized in Figure~\ref{fig;sys_bench}, the Intel UP Board 
-did experience a change in its inferencing times, but the increase was not as drastic as is seen by 
-the Raspberry Pi 3. in the worst case, the Intel UP Board produced times that were over two times as 
-large, whereas the Raspberry Pi 3 output times that were around 10 times greater. The Intel UP Board 
-was successful in completing inferencing operations in under 50 ms when benchmarks were run on a 
-maximum of 2 cores, while the Raspberry Pi 3 failed to do so when any benchmark was introduced. As 
-such, it can be concluded that the Intel UP Board would be more capable of running computationally 
-heavy processes during autonomous driving, and that the Pi wouldn't be able to do the same.
+Finally, we examined if the performance of the Intel UP Board and 
+Tegra TX2 would be affected by the addition of the synthetic 
+benchmarks. As summarized in Figure~\ref{fig;sys_bench}, the both 
+experienced a change in their inferencing times, but the increases 
+were not as drastic as is seen by the Raspberry Pi 3. in the worst 
+case, both the UP Board and TX2 (CPU only and GPU) produced times 
+that were over two times as large, whereas the Raspberry Pi 3 output 
+times that were around 10 times greater. The Intel UP Board was 
+successful in completing inferencing operations in under 50 ms when 
+benchmarks were run on a maximum of 2 cores, and the TX2 was 
+successful in all experiments irrespective of the computing resources 
+utilized. Comparetively, both performed better than the Raspberry Pi 
+3 which failed to meet its deadlines when any benchmark was 
+introduced. As such, it can be concluded that the Intel UP Board 
+would be more capable of running computationally heavy processes 
+during autonomous driving, and that the Pi wouldn't be able to do the 
+same.
 
@@
   \centering
-  \includegraphics[width=.5\textwidth]{figs/system_benchmark}
-  \caption{Inferencing time across platforms based on number of cores used.}
+  \includegraphics[width=.5\textwidth]{figs/compare_benchmark}
+  \caption{Inferencing time across platforms based on number of cores 
+used.}
   \label{fig:sys_bench}
@@
 
-In the comparison of the real-time capabilities of three embedded computing platforms, we found that 
-the Raspberry Pi 3 performed the worst. When compared to the Intel UP Board, inferencing on the Pi 
-took twice as long across all multicore and multimodel experiments. The difference was even more 
-noticeable in the the addition of computationally heavy synthetic benchmarks that had a much more 
-dire effect on the Pi.
+In the comparison of the real-time capabilities of three embedded 
+computing platforms, we found that the Raspberry Pi 3 performed the 
+worst. When compared to the Intel UP Board and the NVIDIA Tegra TX2, 
+control loops on the Pi were shown to take noticably longer to 
+complete, regardless of the number of cores utilized or the number of 
+neural network models run simultaneoulsy. The difference was even 
+more noticeable in the the addition of computationally heavy 
+synthetic benchmarks that had a much more dire effect on the Pi. In 
+the worst case, an average control loop on the Pi took an abnormally 
+long amount of time to finish. Compared to the TX2 with GPU support, 
+those times were 34.8X times greater. From this, we find that the UP 
+Board and Tegra TX2 are superior platforms for real-time CNN 
+operations. However, due to the Pi's ability to satisfy the 50 ms 
+WCET in some experiments, we conclude that the Raspberry Pi 3 is 
+feasible for autonomous research.

### 2017-10-06 paper/related.tex
Updated related works section

@@
 Research on improving the performance of neural networks \cite{Jouppi2017} also attributes to the viability of using these models in small systems.
+With less powerful onboard computing devices, it was necessary for these kinds of robotics systems to use an external computer for the algorithm such as is used in \cite{LeCun:2005}.
 
-Together with hardware and platform improvements, there have been significant improvements in the algorithms such as a neural network algorithm capable of surpassing human level control at atari games \cite{DBLP}.
+Together with hardware and platform improvements, there have been significant improvements in the algorithms such as a neural network algorithm capable of surpassing human level control at Atari games \cite{DBLP}.
 A popular technique had been to take sensor input and determine a depth map, such as the model used in \cite{Michels:2005}.
 End-to-end models such as the one used in \cite{Bojarski2016} can improve the efficiency of the model.
-Improving the flexibility of the model and it abilty to adapt to different situations and avoiding overfitting to training data continues to be an important topic in autonomous driving \cite{Pomerleau1989}.
+Improving the flexibility of the model and it ability to adapt to different situations and avoiding overfitting to training data continues to be an important topic in autonomous driving \cite{Pomerleau1989}.
 
@@
 This approach, while certainly effective, presents a higher barrier to entry for research groups.
-Smaller, more focused platforms, such as the one found in \cite{Michels:2005} present a more accesible way forward.
+Recently, there has been a movement to make research towards autonomous driving more accessible.
+\cite{Kato:2015} have provided an open platform for developing autonomous vehicles using publicly available vehicles and sensors.
+Smaller, more focused platforms, such as the one found in \cite{Michels:2005} also present an accessible way forward.

### 2017-10-06 paper/main.tex
add authors
appendix
..

@@
 \title{DeepPicar:​ ​A​ ​Low-cost​ ​Deep​ ​Neural​ ​Network-based​ ​Autonomous​ ​Car}
-\author{Author One, Author Two\\
-\{author1,author2\}@ku.edu\\
+\author{Michael Garrett Bechtel, Elise McEllhiney, Heechul Yun\\
+\{m783b224, e908m429, heechul.yun\}@ku.edu\\
 University of Kansas, USA\\ 
@@
 \begin{abstract}
-In order to effectively research autonomous vehicles, a real-time capable platform is required. Such 
-platforms can be hard to obtain, or build, as the necessary components can be expensive. In this paper, 
-we present a low-cost alternative platform that utilizes a Raspberry Pi 3 Model B, along with a single 
-external camera sensor, to perform all necessary self-driving operations. A convolutional neural network 
-(CNN) is employed and trained with input from a human driver to navigate a custom made track. We display 
-the real-time efficacy of our platform in experiments with varying amount of computational resources 
-and/or intensive AI workloads. To gain a better understanding of our platform's potential, we compare it 
-to other embedded computing platforms to determine which are suitable for autonomous driving research.
+In order to effectively research autonomous vehicles, a real-time
+capable platform is required. Such platforms can be hard to obtain, or
+build, as the necessary components can be expensive. In this paper,
+we present a low-cost alternative platform that utilizes a Raspberry
+Pi 3 Model B, along with a single external camera sensor, to perform
+all necessary self-driving operations. A convolutional neural network
+(CNN) is employed and trained with input from a human driver to
+navigate a custom made track. We display the real-time efficacy of our
+platform in experiments with varying amounts of computational resources
+and/or intensive AI workloads. To gain a better understanding of our
+platform's potential, we compare it to other embedded computing
+platforms to determine which are suitable for autonomous driving
+research.
 \end{abstract}
@@
 \input{evaluation}
+\input{comparison}
 \input{related}
@@
 \appendix
+\input{appendix}
 \end{document}

### 2017-10-06 paper/comparison.tex
comparision revision (WIP)

@@
 based autonomous vehicles. 
+Table~\ref{tbl:platforms} shows architectural features of the three
+platforms~\footnote{Note that the GPUs of the Raspberry Pi 3 and Intel
+  Up are not used in evaluation due to lack of software (TensorFlow)
+support. Also, the two Denvor cores in Tegra TX2 are not used in
+evaluation due to TensorFlow issues.}.
 
-Table~\ref{tbl:platforms}
-
-Our basic approach is to use the same DeepPicar's software, and repeat
+Our basic approach is to use the same DeepPicar software, and repeat
 the experiments in Section~\ref{sec:evaluation} on each hardware
 platform and compare the results. 
+For Tegra TX2, we have two different system configurations,
+which differ in whether TensorFlow is configured to use its GPU or
+only the CPU cores. Thus, the total four system configurations are
+compared.
 
@@
 
-The results of the multicore tests can be seen in 
-Figure~\ref{fig:sys_core}. It was found that the both the Tegra TX2 
+Figure~\ref{fig:sys_core} shows the average control loop completion
+timing as a function of the number of CPU cores on each of the four
+system configurations.
+It was found that the both the Tegra TX2 
 and the Intel UP Board performed better than the Raspberry Pi 3 in 

### 2017-10-06 paper/evaluation.tex
Update control loop graph
Also changed location of training loss graph
comparision revision (WIP)

@@
 
-The Raspberry Pi 3 Model B platform used in DeepPicar equip a Boardcom
+The Raspberry Pi 3 Model B platform used in DeepPicar equips a Boardcom
 BCM2837 SoC, which has a quad-core 64bit ARM Cortex-A53 cluster,
@@
   \centering
-  \includegraphics[width=.45\textwidth]{Total_Processing_Time}
+  \includegraphics[width=.5\textwidth]{figs/Fig7_new}
   \caption{DeepPicar's control loop processing times over 1000 input image frames.}
@@
   \caption{Average control loop execution time vs. \#of CPU
-    cores. \fixme{instead of the total control loop time, use the
-      average of DNN inferencing time.}}
+    cores.}
   \label{fig:perf-vs-corecnt}
 \end{figure}
+    %% \fixme{instead of the total control loop time, use the
+    %%   average of DNN inferencing time.}}
 
@@
   \caption{Compared embedded computing platforms}
-  \label{tbl:patforms}
+  \label{tbl:platforms}
 \end{table*}

### 2017-10-06 paper/comparison.tex
..

@@
 platforms~\footnote{Note that the GPUs of the Raspberry Pi 3 and Intel
-  Up are not used in evaluation due to lack of software (TensorFlow)
+  Up are not used in evaluation due to the lack of software (TensorFlow)
 support. Also, the two Denvor cores in Tegra TX2 are not used in
@@
 timing as a function of the number of CPU cores on each of the four
-system configurations.
-It was found that the both the Tegra TX2 
-and the Intel UP Board performed better than the Raspberry Pi 3 in 
-all experiments. On average, the UP was able to perform inferencing 
-operations, at least, twice as fast as the Raspberry Pi 3. The 
-difference was even greater with the TX2, as control loops, in the 
-worst case, were 3x faster using only the CPU and 10x quicker when 
-using the GPU. As a result, they were all able to satisfy the 50 ms 
+system configurations. First, Tegra TX2 and the Intel UP exhibit
+superio performance  compared with the Raspberry Pi 3 in 
+all experiments. On average, the Intel UP is able to perform
+inferencing operations more than twice as fast as the Raspberry Pi
+3. The  difference is even greater with the TX2 as the performance is
+about 3X faster using only the CPU and 10X faster when using the GPU.
+As a result, they were all able to satisfy the 50 ms 
 WCET by a very clear margin (in the worst case, the UP Board would 

### 2017-10-06 paper/evaluation.tex
..

@@
 
-In each experiment, we run a single DNN model one core, and
-co-scheduled an increasing number of the memory intensive synthetic
+We run a single DNN model one core, and
+co-schedule an increasing number of the memory intensive synthetic
 benchmarks~\footnote{We use the \emph{Bandwidth} benchmark in the

### 2017-10-06 paper/comparison.tex
figure 12 revision

@@
   \includegraphics[width=.5\textwidth]{figs/compare_core}
-  \caption{Inferencing time across platforms based on number of cores 
-used.}
+  \caption{Average control loop execution time.} 
   \label{fig:sys_core}
@@
 Figure~\ref{fig:sys_core} shows the average control loop completion
-timing as a function of the number of CPU cores on each of the four
-system configurations. First, Tegra TX2 and the Intel UP exhibit
-superio performance  compared with the Raspberry Pi 3 in 
-all experiments. On average, the Intel UP is able to perform
-inferencing operations more than twice as fast as the Raspberry Pi
-3. The  difference is even greater with the TX2 as the performance is
-about 3X faster using only the CPU and 10X faster when using the GPU.
-As a result, they were all able to satisfy the 50 ms 
-WCET by a very clear margin (in the worst case, the UP Board would 
-still meet it by $\sim$20 ms), and, thus, demonstrates that they are 
-more capable of autonomous driving than the Raspberry Pi 3.
+timing of the four system configurations we tested as a function of
+the number of CPU cores used.
+First, both the Intel UP and Tegra TX2 exhibit superior performance
+compared with the Raspberry Pi 3. 
+When all four CPU cores are used, the Intel UP is 2.53X faster than
+the Pi 3, while TX (CPU) and TX (GPU) are 3.6X and 7.6X times faster,
+respectively, than the Pi 3. 
+As a result, they are all able to satisfy the 50 ms 
+WCET by a clear margin
+%%  (in the worst case, the UP Board would 
+%% still meet it by $\sim$20 ms) --> where's the data showing this?
+, and, in case of TX2, 50 Hz or even 100 Hz real-time control is
+feasible with the help of its GPU.
 

### 2017-10-06 paper/comparison.tex
figure 13
figure 14 revision

@@
 Table~\ref{tbl:platforms} shows architectural features of the three
-platforms~\footnote{Note that the GPUs of the Raspberry Pi 3 and Intel
+platforms~\footnote{The GPUs of the Raspberry Pi 3 and Intel
   Up are not used in evaluation due to the lack of software (TensorFlow)
-support. Also, the two Denvor cores in Tegra TX2 are not used in
+support. Also, the two Denver cores in Tegra TX2 are not used in
 evaluation due to TensorFlow issues.}.
@@
   \centering
-  \includegraphics[width=.5\textwidth]{figs/compare_core}
+  \includegraphics[width=.45\textwidth]{figs/compare_core}
   \caption{Average control loop execution time.} 
@@
 , and, in case of TX2, 50 Hz or even 100 Hz real-time control is
-feasible with the help of its GPU.
-
-The UP Board and TX2 also performed better in the multimodel tests, 
-as is shown in Figure~\ref{fig:sys_model}. Once again, they were able 
-to complete control loops in a fraction of the time when compared to 
-the Pi. Furthermore, both of the platforms satisfied the 50 ms WCET 
-in all experiments, while the Raspberry Pi 3 was unable to do so. As 
-such, the UP Board and TX2 display greater potential for the 
-simultaneous execution of multiple models during self-driving 
-operation, while the Raspberry Pi 3 would struggle to do so.
+feasible with the help of its GPU. Another observation is that TX2
+(GPU) does not change much as most of the neural network computation
+is done at the GPU.
 
@@
   \centering
-  \includegraphics[width=.5\textwidth]{figs/compare_model}
-  \caption{Inferencing time across platforms based on number of 
-models run concurrently.}
+  \includegraphics[width=.45\textwidth]{figs/compare_model}
+  \caption{Average control loop execution time when multiple DNN
+    models are co-scheduled. }
   \label{fig:sys_model}
@@
 
-Finally, we examined if the performance of the Intel UP Board and 
-Tegra TX2 would be affected by the addition of the synthetic 
-benchmarks. As summarized in Figure~\ref{fig;sys_bench}, the both 
+The UP Board and TX2 also perform much better when multiple DNN models
+are co-scheduled. Figure~\ref{fig:sys_model} shows the results of the
+multi-model co-scheduling experiment. Once again, they can comfortably
+20 Hz real-time performance for all of the co-scheduled DNN control
+loops, and in case of TX 2 (GPU), 100 Hz real-time control is still
+feasible. Given that the GPU must be shared among the co-scheduled DNN
+models, the result suggests that TX2's GPU has sufficient capacity to
+accomodate multiple instances of the DNN model we tested.
+
+%% to complete control loops in a fraction of the time when compared to 
+%% the Pi. Furthermore, both of the platforms satisfied the 50 ms WCET 
+%% in all experiments, while the Raspberry Pi 3 was unable to do so. As 
+%% such, the UP Board and TX2 display greater potential for the 
+%% simultaneous execution of multiple models during self-driving 
+%% operation, while the Raspberry Pi 3 would struggle to do so.
+
+\begin{figure}[h]
+  \centering
+  \includegraphics[width=.45\textwidth]{figs/compare_benchmark}
+  \caption{Inferencing time across platforms based on number of cores 
+used.}
+  \label{fig:sys_bench}
+\end{figure}
+
+Finally, we compare the effect of co-scheduling memory intensive
+synthetic benchmarks on the DNN control loop timing.
+Figure~\ref{fig:sys_bench} shows the results. 
 experienced a change in their inferencing times, but the increases 
@@
 same.
-
-\begin{figure}[h]
-  \centering
-  \includegraphics[width=.5\textwidth]{figs/compare_benchmark}
-  \caption{Inferencing time across platforms based on number of cores 
-used.}
-  \label{fig:sys_bench}
-\end{figure}
 

### 2017-10-06 paper/evaluation.tex
figure 14 revision

@@
   \includegraphics[width=.45\textwidth]{figs/perf_vs_modelcnt}
-  \caption{Timing impact of co-scheduling multiple DNNs. 1x1: one DNN
-    model using one CPU core; 4x1: four DNN models each using one CPU;
-    1x2: one DNN model using two CPU cores; 2x2: two DNN models each
+  \caption{Timing impact of co-scheduling multiple DNNs. 1Nx1C: one DNN
+    model using one CPU core; 4Nx1C: four DNN models each using one CPU;
+    1Nx2C: one DNN model using two CPU cores; 2Nx2C: two DNN models each
     using two CPU cores.} 
@@
     \hline
-            & BCM2837              & X5-Z8350 (Cherry Trail)   & Tegra TX2 \\
+            & BCM2837              & X5-Z8350 (Cherry Trail)   & Tegra X2 \\
     CPU     & 4x Cortex-A53@1.2GHz/512KB L2  &

### 2017-10-06 paper/intro.tex
Adjust paper roadmap in intro
The previous roadmap covered the incorrect sections.

@@
 
-The remaining sections of the paper are as follows: Section II gives
-an overview of the platform, including the high-level system and the
-methods used for training and inference. Section III discusses the
-ways/methodologies in which training data was collected. Section IV
-outlines the network architecture. This is followed by the real-time
-DNN inferencing in Section V. Section VI reviews the online/offline
-training done with the platform, with an evaluation given in Section
-VII. Section VIII gives a discussion of related works, and the paper
-finishes with conclusions in Section IX.
+The remaining sections of the paper are as follows: Section II 
+provides a background of applications in autonomous driving. Section 
+III gives an overview of the DeepPicar platform, including the 
+high-level system and the methods used for training and inference. 
+Section IV presents our evaluation of the platform and how different 
+factors can affect performance. Section V gives a comparison between 
+the Raspberry Pi 3 and other embedded computing platforms to 
+determine their suitably for autonomous driving research. Related 
+works are discussed in Section VI, and the paper finishes with 
+conclusions in Section VII.
 

### 2017-10-06 paper/comparison.tex
comparison revision done

@@
 
-Finally, we compare the effect of co-scheduling memory intensive
-synthetic benchmarks on the DNN control loop timing.
+Finally, we compare the effect of co-scheduling memory bandwidth
+intensive synthetic benchmarks on the DNN control loop timing. 
 Figure~\ref{fig:sys_bench} shows the results.
-experienced a change in their inferencing times, but the increases 
-were not as drastic as is seen by the Raspberry Pi 3. in the worst 
-case, both the UP Board and TX2 (CPU only and GPU) produced times 
-that were over two times as large, whereas the Raspberry Pi 3 output 
-times that were around 10 times greater. The Intel UP Board was 
-successful in completing inferencing operations in under 50 ms when 
-benchmarks were run on a maximum of 2 cores, and the TX2 was 
-successful in all experiments irrespective of the computing resources 
-utilized. Comparetively, both performed better than the Raspberry Pi 
-3 which failed to meet its deadlines when any benchmark was 
-introduced. As such, it can be concluded that the Intel UP Board 
-would be more capable of running computationally heavy processes 
-during autonomous driving, and that the Pi wouldn't be able to do the 
-same.
+As discussed in 
+Section~\ref{sec:eval-memhog}, we observe dramatic execution time
+increase, up to 9.4X, in Raspberry Pi 3 as we increase the number of
+co-scheduled tasks. We also observe increased control loop execution
+timing in Intel Up and TX2, but the degree of the increase is not as
+dramatic as the Pi3. Compared to their respective solo timing (i.e.,
+the model runs on a single core in isolation), Intel UP suffers up to
+2.3X execution time increase; TX2 (CPU) and TX2 (GPU) suffer up to
+2.2X and 2.5X increase, respectively. This is somewhat suprising
+because Raspberry Pi 3's cores are in-oder architecture based while
+the cores in Intel Up and NVIDIA TX2 are out-of-order architecture
+based and that the memory intensive tasks on out-of-order cores can
+generate more memory traffic. We suspect that this is because the
+memory subsystems in the Intel UP and TX2 platforms provide higher
+bandwidth and fairness than the memory subsystem of the Pi 3.
 
-In the comparison of the real-time capabilities of three embedded 
-computing platforms, we found that the Raspberry Pi 3 performed the 
-worst. When compared to the Intel UP Board and the NVIDIA Tegra TX2, 
-control loops on the Pi were shown to take noticably longer to 
-complete, regardless of the number of cores utilized or the number of 
-neural network models run simultaneoulsy. The difference was even 
-more noticeable in the the addition of computationally heavy 
-synthetic benchmarks that had a much more dire effect on the Pi. In 
-the worst case, an average control loop on the Pi took an abnormally 
-long amount of time to finish. Compared to the TX2 with GPU support, 
-those times were 34.8X times greater. From this, we find that the UP 
-Board and Tegra TX2 are superior platforms for real-time CNN 
-operations. However, due to the Pi's ability to satisfy the 50 ms 
-WCET in some experiments, we conclude that the Raspberry Pi 3 is 
-feasible for autonomous research.
+Another intersting observation is that TX2 (GPU) also suffers
+considerable execution time increase (2.5X) despite the fact that the
+co-scheduled synthetic tasks do not utilize the GPU. In other words,
+the DNN model has dedicated access to the GPU. This is, however, a
+known characteristic of integrated CPU-GPU architectue based
+platforms that both CPU and GPU share the same memory
+subsystem~\cite{Ali2017}.
+
+%% experienced a change in their inferencing times, but the increases 
+%% were not as drastic as is seen by the Raspberry Pi 3. in the worst 
+%% case, both the UP Board and TX2 (CPU only and GPU) produced times 
+%% that were over two times as large, whereas the Raspberry Pi 3 output 
+%% times that were around 10 times greater. The Intel UP Board was 
+%% successful in completing inferencing operations in under 50 ms when 
+%% benchmarks were run on a maximum of 2 cores, and the TX2 was 
+%% successful in all experiments irrespective of the computing resources 
+%% utilized. Comparetively, both performed better than the Raspberry Pi 
+%% 3 which failed to meet its deadlines when any benchmark was 
+%% introduced. As such, it can be concluded that the Intel UP Board 
+%% would be more capable of running computationally heavy processes 
+%% during autonomous driving, and that the Pi wouldn't be able to do the 
+%% same.
+
+In summary, we find that today's embedded computing platforms, even as
+inexpensive as a Raspberry Pi 3, are powerful enough to support
+vision and end-to-end deep learning based real-time control
+applications. Furthermore, availability of CPU cores and GPU on these
+platforms allow consolidating mutiple deep neural network based AI
+workloads. However, shared resource contention among these diverse
+computing resources remain an important issue that must be understood
+and controlled especially for safety-critical applications.
+
+%% the comparison of the real-time capabilities of three embedded 
+%% computing platforms, we found that the Raspberry Pi 3 performed the 
+%% worst. When compared to the Intel UP Board and the NVIDIA Tegra TX2, 
+%% control loops on the Pi were shown to take noticably longer to 
+%% complete, regardless of the number of cores utilized or the number of 
+%% neural network models run simultaneoulsy. The difference was even 
+%% more noticeable in the the addition of computationally heavy 
+%% synthetic benchmarks that had a much more dire effect on the Pi. In 
+%% the worst case, an average control loop on the Pi took an abnormally 
+%% long amount of time to finish. Compared to the TX2 with GPU support, 
+%% those times were 34.8X times greater. From this, we find that the UP 
+%% Board and Tegra TX2 are superior platforms for real-time CNN 
+%% operations. However, due to the Pi's ability to satisfy the 50 ms 
+%% WCET in some experiments, we conclude that the Raspberry Pi 3 is 
+%% feasible for autonomous research.

### 2017-10-06 paper/conclusion.tex
integrate related work with background section
revise conclusion

@@
 \section{Conclusion}\label{sec:conclusion}
+We presented DeepPicar, a low cost autonomous car platform that is
+inexpensive to build but is based on a state-of-the art AI technology:
+End-to-end deep learning based real-time control.
+Specifically, DeepPicar uses a deep convolutional neural network to
+predict steering angles of the car directly from a camera input data
+in real-time. Importantly, DeepPicar's neural network architecture is
+identical to that of NVIDIA's real self-driving car.
 
-We present a low cost autonomous vehicle platform that uses state-of-the art AI techniques and 
-technologies. It employs a deep convolutional neural network (CNN) that is trained to effectively 
-navigate the car by predicting the steering angle given an input frame. We have tested the real time 
-efficacy of our platform and found that it is capable of performing inferencing operations within a WCET 
-of 50 ms, or 20 fps, when a single model is run on more than one core. Further studies should be 
-conducted on the interference that occurs between cores to allow for improved performance in utilizing 
-multiple models for autonomous vehicles. With minimal further research, the DeepPicar can 
-become a valudable, reliable and inexpensive tool in the continued research of autonomous vehicles.
+Despite the complexity of the neural network, DeepPicar uses a
+low-cost Raspberry Pi 3 quad-core computer as a sole computing
+resource. We systematically analyzed the real-time characteristics of
+the Pi3 platform in the context of deep-learning based real-time
+control appilcations, with a special emphasis on real-time deep neural
+network inferencing.
+We also evaluated other more powerful embedded computing
+platforms to better understand achievable real-time performance of
+DeepPicar's deep-learning based control system and the impact of
+computing hardware architectures.
+We find all tested embedded platforms, including the Pi3, are capable
+of supporting deep-learning based real-time control, from 20 Hz up to
+100 Hz, depending on the platform and its system
+configuration. However, shared resource contention remains an
+important issue that must be considered in applying deep-learning
+models on shared memory based embedded computing platforms.
+
+%% We evaluated the real time efficacy of our platform and found
+%% that it is capable of performing inferencing operations within a WCET
+%% of 50 ms, or 20 fps, when a single model is run on more than one
+%% core.
+
+As future work, we plan to apply shared resource management
+techniques~\cite{Yun2013,yun2014rtas} on the DeepPicar platform and
+evaluate their impact to the real-time performance of the system. We
+also plan to improve the prediction accuracy by feeding more data and
+upgrade the RC car hardware platform to enable more precise steering
+angle control.
+
+%% Further studies should be conducted on the interference that
+%% occurs between cores to allow for improved performance in utilizing
+%% multiple models for autonomous vehicles. With minimal further
+%% research, the DeepPicar can become a valudable, reliable and
+%% inexpensive tool in the continued research of autonomous vehicles.

### 2017-10-06 paper/related.tex
integrate related work with background section
revise conclusion

@@
 
-Autonomous vehicles are becoming an increasingly interesting research topic as computing platforms capable of safely processing sensor input into safe vehicle controls become smaller and more affordable.
-The model proposed in End-To-End Learning for Self-Driving Cars \cite{Bojarski2016} is viable on a platform as simple as a Raspberry Pi.
+%  NN for autonomous systems
+\cite{Bojarski2016}
 
-Given smaller, affordable platforms with multiple cores and even GPU based parallel processing such as the Tegra X1 \cite{NVIDIA2015} and the recently released Tegra X2 \cite{Amert2017}, allocation and management of available shared CPU and GPU resources \cite{Kim2016} becomes increasingly important.
-Research on improving the performance of neural networks \cite{Jouppi2017} also attributes to the viability of using these models in small systems.
-With less powerful onboard computing devices, it was necessary for these kinds of robotics systems to use an external computer for the algorithm such as is used in \cite{LeCun:2005}.
+% open testbed
+\cite{Kato:2015}
+~\cite{shin2017project}
+~\cite{upennf1tenth}
 
-Together with hardware and platform improvements, there have been significant improvements in the algorithms such as a neural network algorithm capable of surpassing human level control at Atari games \cite{DBLP}.
-A popular technique had been to take sensor input and determine a depth map, such as the model used in \cite{Michels:2005}.
-End-to-end models such as the one used in \cite{Bojarski2016} can improve the efficiency of the model.
-Improving the flexibility of the model and it ability to adapt to different situations and avoiding overfitting to training data continues to be an important topic in autonomous driving \cite{Pomerleau1989}.
+% real-time performance evaluation of embedded computing platforms
+~\cite{Otterness2017} vision workload
+\cite{Amert2017} TX2 scheduling algorithm
+\cite{NVIDIA2015} dnn workload real-time inference
 
-Producing large platforms such as DAVE \cite{LeCun:04} has often been the approach to developing the technology for self-driving vehicles.
-This approach, while certainly effective, presents a higher barrier to entry for research groups.
-Recently, there has been a movement to make research towards autonomous driving more accessible.
-\cite{Kato:2015} have provided an open platform for developing autonomous vehicles using publicly available vehicles and sensors.
-Smaller, more focused platforms, such as the one found in \cite{Michels:2005} also present an accessible way forward.
+% acceleration
+\cite{Jouppi2017} TPU
+
+Autonomous vehicles are becoming an increasingly interesting research
+topic as computing platforms capable of safely processing sensor input
+into safe vehicle controls become smaller and more affordable.  The
+model proposed in End-To-End Learning for Self-Driving Cars
+\cite{Bojarski2016} is viable on a platform as simple as a Raspberry
+Pi.
+
+Given smaller, affordable platforms with multiple cores and even GPU
+based parallel processing such as the Tegra X1 \cite{NVIDIA2015} and
+the recently released Tegra X2 \cite{Amert2017}, allocation and
+management of available shared CPU and GPU resources \cite{Kim2016}
+becomes increasingly important.  Research on improving the performance
+of neural networks \cite{Jouppi2017} also attributes to the viability
+of using these models in small systems.  With less powerful onboard
+computing devices, it was necessary for these kinds of robotics
+systems to use an external computer for the algorithm such as is used
+in \cite{LeCun:2005}.
+
+Together with hardware and platform improvements, there have been
+significant improvements in the algorithms such as a neural network
+algorithm capable of surpassing human level control at Atari games
+\cite{DBLP}.  A popular technique had been to take sensor input and
+determine a depth map, such as the model used in
+\cite{Michels:2005}. End-to-end models such as the one used in
+\cite{Bojarski2016} can improve the efficiency of the model.
+Improving the flexibility of the model and it ability to adapt to
+different situations and avoiding overfitting to training data
+continues to be an important topic in autonomous driving
+\cite{Pomerleau1989}.
+
+Producing large platforms such as DAVE \cite{LeCun:04} has often been
+the approach to developing the technology for self-driving vehicles.
+This approach, while certainly effective, presents a higher barrier to
+entry for research groups.  Recently, there has been a movement to
+make research towards autonomous driving more accessible.
+\cite{Kato:2015} have provided an open platform for developing
+autonomous vehicles using publicly available vehicles and sensors.
+Smaller, more focused platforms, such as the one found in
+\cite{Michels:2005} also present an accessible way forward.

### 2017-10-06 paper/background.tex
integrate related work with background section
revise conclusion
abstract

@@
-\section{Background} \label{sec:background}
+\section{Background and Related Work} \label{sec:background}
 
-In this section, we provide background on the application of deep
-learning in robotics, particularly autonomous vehicles.
+In this section, we provide background and related work on the
+application of deep learning in robotics, particularly autonomous
+vehicles.
 

### 2017-10-06 paper/main.tex
integrate related work with background section
revise conclusion
abstract

--- 2017-10-02 paper/main.tex
@@
 \begin{abstract}
-Abstract goes here.
+We present DeepPicar, a low-cost deep neural network (DNN) based
+autonomous car platform. DeepPicar is a small scale
+replication of a real self-driving car called Dave-2 by NVIDIA, which
+drove public roads using a deep convolutional network (CNN), which
+takes images from a front-facing camera as input and produces
+car steering angles as output. DeepPicar uses the exact same network
+architecture---9 layers, 27 million connections and 24K
+parameters---and can be trained to drives itself using a small web
+camera in real-time on a modest Raspberry Pi 3 quad-core computing
+platform. Using DeepPicar, we analyze Pi 3's computing capabilities to 
+support end-to-end deep learning based real-time control of autonomous
+vehicles. We also systematically compare other contemporary embedded
+computing platforms using the DeepPicar's deep learning control system
+as a workload.
+We find all tested platforms, including the Pi3, are capable of
+supporting deep-learning based real-time control, from 20 Hz up to 100
+Hz, depending on hardware platform. 
+However, shared resource contention remains an
+important issue that must be considered in applying deep-learning
+models on shared memory based embedded computing platforms.
+
+%% In order to effectively research autonomous vehicles, a real-time
+%% capable platform is required. Such platforms can be hard to obtain, or
+%% build, as the necessary components can be expensive. In this paper,
+%% we present a low-cost alternative platform that utilizes a Raspberry
+%% Pi 3 Model B, along with a single external camera sensor, to perform
+%% all necessary self-driving operations. A convolutional neural network
+%% (CNN) is employed and trained with input from a human driver to
+%% navigate a custom made track. We display the real-time efficacy of our
+%% platform in experiments with varying amounts of computational resources
+%% and/or intensive AI workloads. To gain a better understanding of our
+%% platform's potential, we compare it to other embedded computing
+%% platforms to determine which are suitable for autonomous driving
+%% research.
 \end{abstract}
@@
 \input{evaluation}
-\input{related}
+\input{comparison}
+% \input{related}
 \input{conclusion}
 %-------------------------------------------------------------------------
-\bibliographystyle{plain}
+\bibliographystyle{abbrv}
 \bibliography{reference}
+
+\appendix
+\input{appendix}
 \end{document}

### 2017-10-06 paper/main.tex
abstract again

@@
 replication of a real self-driving car called Dave-2 by NVIDIA, which
-drove public roads using a deep convolutional network (CNN), which
+drove public roads using a deep convolutional network (CNN) that
 takes images from a front-facing camera as input and produces
@@
 architecture---9 layers, 27 million connections and 24K
-parameters---and can be trained to drives itself using a small web
-camera in real-time on a modest Raspberry Pi 3 quad-core computing
-platform. Using DeepPicar, we analyze Pi 3's computing capabilities to 
+parameters---and can be trained to drives itself in real-time using a
+web camera and a modest Raspberry Pi 3 quad-core platform.
+Using DeepPicar, we analyze Pi 3's computing capabilities to 
 support end-to-end deep learning based real-time control of autonomous
 vehicles. We also systematically compare other contemporary embedded
-computing platforms using the DeepPicar's deep learning control system
-as a workload.
+computing platforms using the DeepPicar's CNN based real-time control
+software as a workload. 
 We find all tested platforms, including the Pi3, are capable of

### 2017-10-06 paper/comparison.tex
integrate related work with background section
revise conclusion
..

@@
   \includegraphics[width=.45\textwidth]{figs/compare_benchmark}
-  \caption{Inferencing time across platforms based on number of cores 
-used.}
+  \caption{Average control loop execution time in the presence of an
+    increasing number of memory intensive applications on idle CPU cores.}
   \label{fig:sys_bench}
@@
 platforms that both CPU and GPU share the same memory
-subsystem~\cite{Ali2017}.
+subsystem~\cite{Ali2017}. As a result, TX2 (GPU) fails to meet the
+10ms deadline for 100 Hz control that would have been feasible if
+there is no contention between the CPU cores and the GPU.
 

### 2017-10-06 paper/main.tex
..

@@
 architecture---9 layers, 27 million connections and 24K
-parameters---and can be trained to drives itself in real-time using a
+parameters---and can be trained to drives itself, in real-time, using a
 web camera and a modest Raspberry Pi 3 quad-core platform.

### 2017-10-06 paper/appendix.tex
abstract
appendix and figures

--- 2017-10-05 paper/appendix.tex
@@
+\subsection{DNN Training and Testing}
+We have trained and tested the deep neural network with several
+different track conditions, different combinations of input
+data, and different hyper parameters. In the following, we describe
+details on two of the training methods that performed reasonably well.
+
+In the first method, we trained the neural network model across a set 
+of 30 completed runs on the track seen in Figure~\ref{fig:track} by a
+human pilot. Half of the runs saw the car driving one way along the
+track, while the remaining half were of the car driving in the
+opposite direction on the track.
+In total, we collected 2,556 frames for training and 2,609 
+frames for validation.
+The weights of the network is initialized using Xavier
+initializer~\cite{Glorot2010}, which is known to provide better
+initial values than the random weight assignment method.
+In each training step, we use a batch
+size of 100 frames, which are randomly selected among all the
+collected training images, to optimize the network.
+We repeat this across 2,000 training steps. When a model was trained
+with the  aformentioned data, the training loss was 0.0188 and the
+validation  loss was 0.0132.
+The change of the loss value over the course of model training can be
+seen in Figure~\ref{fig:modelloss}.
+
+\begin{figure*}[t]
+  \centering
+  \includegraphics[width=.9\textwidth]{figs/TrainingLoss}
+  \caption{Change in loss value throughout training.}
+  \label{fig:modelloss}
+\end{figure*}
+
+In the second method, we use the same data and parameters above except
+that now images are labeled as 'curved' and 'straight' and we pick an
+equal number of images from each category at each training step to
+update the model. In other words, we try to remove bias in selecting
+images. We find that the car performed better in practice by applying
+this approach as the car displayed a greater ability to stay in the
+center of the track (on the white tape).
+However, we find that there is huge discrepency between the training
+loss and the validation loss as the former was 0.09 while the latter
+was 0.0869---a 10X difference---indicating that the model suffers from
+the overfitting problem.
+
+We continue to investigate ways to achieve better prediction accuracy
+in training the network as well as improving the performance of the RC
+car platform, especially on precise steering angle control.
+
+%% We believe, though, that this can be
+%% rectified through the collection of additional data for training and
+%% validation.
+
+%% \begin{figure*}[h]
+%%   \centering
+%%   \includegraphics[width=.9\textwidth]{figs/graph_run=.pdf}
+%%   \caption{ Network visualization. }
+%%   \label{fig:}
+%% \end{figure*}
+%% \fixme{Is all of the network graph necessary or only certain parts?}
+
 \subsection{System-level Factors Affecting Real-Time Performance}
-In the utilization of the Raspberry Pi 3 in our platform, there are a
-few factors that need to be considered and/or enforced in order to
-guarantee that the Pi is able to consistently perform at a desired
-level. Specifically, these issues all have the potential to negatively
-affect the cpu clock speed/frequency, which would result in decreased
-performance. While, in the above experiments, the cpu operated at a
-preferred clock speed of 1.2 GHz, it is entirely possible for the cpu
-to operate at a lower frequency if the following problems are not
-taken into account.
+In using the Raspberry Pi 3 platform, there are a
+few system-level factors, namely power supply and temperature, that
+need to be considered to achieve consistent and high real-time
+performance.
 
-The most notable issue that can affect the cpu clock speed is that of
-the power supplied to the Raspberry Pi. In essence, it is necessary
-that the Pi be supplied with 2 Amps, as any less could hinder the Pi's
-ability to maintain a 1.2 GHz frequency. In experiments conducted with
-a power supply that only provided 1 Amp, the Pi was unable to sustain
-a 1.2 GHz clock speed and, instead, fluctuated between operating at
-600 MHz and 1.2 GHz. As a result, it is necessary, or at least highly
-recommended, that the power supply used for the Raspberry Pi 3 be
-capable of outputting 2 Amps, otherwise optimal performance isn't
-guaranteed.
+%% Specifically, these issues all have the potential to negatively
+%% affect the cpu clock speed/frequency, which would result in decreased
+%% performance. 
+In all our experiments on the Raspberry Pi 3, the CPU  operated at a
+preferred clock speed of 1.2 GHz. However, without care, it is
+possible for the CPU may be operating at a lower frequency.
 
-Another factor that can affect clock speed is that of the
-cpu's temperature. Some model operations can be computationally
-intensive, thus it is possible for the temperature of the cpu to
-become relatively high. This can be especially problematic in
-situations where multiple models are running  simultaneously on the
-Pi. Consequently, thermal throttling may be used to decrease the clock
-speed so that the cpu temperature stays at a safe level. Thus the
-Raspberry Pi may not be suited for prolonged use, especially in cases
-where the workload is relatively larger, such as running multiple
-models. Rather, the Pi seems to be better suited for running in set
-periods, after which it is turned  off or made idle so that the cpu is
-allowed time to cool down.
+An important factor is CPU thermal throttling, which can affect CPU
+clock speed if the CPU temperature is too high (Pi 3's firmware is
+configured to throttle at 85C).
+DNN model operations are computationally intensive, thus it is
+possible for the temperature of the CPU to rise quickly. This can be
+especially problematic in situations where multiple DNN models are
+running simultaneously on the Pi 3.
+If the temperature reaches the threshold, the Pi 3's thermal
+throttling kicks in and decreases the clock speed down to 600MHz---a
+half of the maximum 1.2GHz---so that the CPU's temperature stays at a
+safe level.
+We found that without proper cooling solutions (heatsink or fan), 
+prolonged use of the system would result in CPU frequency decrease
+that may affect evaluation.
+
+Another factor to consider is power supply. From our experiences, the
+Pi 3 frequency throttling also kicks in when the power source can not
+provide the required minimum of 2A current.
+In experiments conducted with a power supply that only provided 1 Amp,
+the Pi was unable to sustain a 1.2 GHz clock speed and, instead,
+fluctuated between operating at 600 MHz and 1.2 GHz. As a result, it
+is necessary, or at least highly recommended, that the power supply
+used for the Raspberry Pi 3 be capable of outputting 2 Amps, otherwise
+optimal performance isn't guaranteed.
+
+Our initial experiment results suffered from these issues after which
+we always carefully monitored the current operating frequencies of the
+CPU cores during the experiments to ensure the correctness and
+repeatability of the results.
+
+%% Thus the
+%% Raspberry Pi may not be suited for prolonged use, especially in cases
+%% where the workload is relatively larger, such as running multiple
+%% models. Rather, the Pi seems to be better suited for running in set
+%% periods, after which it is turned  off or made idle so that the cpu is
+%% allowed time to cool down.

### 2017-10-07 paper/intro.tex
Small edits to intro

@@
 years as many companies are actively developing related hardware
-and software technologies toward fully autonomous driving capability without
-any human intervention. Deep neural networks (DNNs) have been
+and software technologies toward fully autonomous driving capability with
+no human intervention. Deep neural networks (DNNs) have been
 successfully applied in various perception and control tasks in
-recent years, and they are important workloads for autonomous vehicles
+recent years.  They are important workloads for autonomous vehicles
 as well. For example, Tesla Model S was known to use a specialized
@@
 detection task may be directly linked to safety of the vehicle. This
-requires a high computing capacity as well as means to guaranteeing
+requires a high computing capacity as well as the means to guaranteeing
 the timings. On the other hand, the computing hardware platform must
@@
 To understand what kind of computing hardware is needed for AI
-workloads, we need a testbed and realistic workloads. While a real car
-testbed would be most ideal, it is not only highly expensive but also
+workloads, we need a testbed and realistic workloads. While using a real car
+based testbed would be most ideal, it is not only highly expensive but also
 poses serious safety concerns that hinder development and exploration.
@@
   \item We systematically compare real-time computing capabilities of
-    multiple embedded computing platforms in the context of a
+    multiple embedded computing platforms in the context of
     vision-based autonomous driving.
@@
 The remaining sections of the paper are as follows: Section II 
-provides a background of applications in autonomous driving. Section 
-III gives an overview of the DeepPicar platform, including the 
+provides a background of applications in autonomous driving and related works.
+ Section III gives an overview of the DeepPicar platform, including the 
 high-level system and the methods used for training and inference. 
@@
 the Raspberry Pi 3 and other embedded computing platforms to 
-determine their suitably for autonomous driving research. Related 
-works are discussed in Section VI, and the paper finishes with 
-conclusions in Section VII.
+determine their suitably for autonomous driving research. The paper finishes with 
+conclusions in Section VI.
 

### 2017-10-07 paper/background.tex
Editing background section

@@
 produce steering control commands bypassing all intermediary steps and
-hand-written rules in conventional robotics control approach. The
-NVIDIA's latest effort reports that their trained CNN successfully
-self-drive their modified cars in public roads without human
-intervention~\cite{Bojarski2016}.
+hand-written rules in conventional robotics control approach.  
+NVIDIA's latest effort reports that their trained CNN
+autonomously controls their modified cars on public roads without human
+intervention successfully~\cite{Bojarski2016}.
 
@@
 as input to produce predicted output (e.g., predicted image
-classification). In general, the training phase is most compute
+classification). In general, the training phase is most computationally
 intensive and requires high throughput, which is generally not
 available on embedded platforms. The inferencing phase, on the
-other hand, is relatively less compute intensive and latency becomes
-as important, if not more, as compute throughput because many
+other hand, is relatively less computationally intensive and latency becomes
+as important, if not more, as computational throughput because many
 use-cases have strict real-time requirements (e.g., search query
@@
 %% autonomous vehicle must satisfy two often conflicting goals:
-On the one hand, the platform must provide 
+The platform must provide 
 enough computing capacity for real-time processing of computationally
 expensive AI workloads (deep neural networks);
-On the other hand, the platform must also satisfy various
-constrains such as cost, size, weight, and power consumption limits.
+The platform must also satisfy various
+constraints such as cost, size, weight, and power consumption limits.
 
 Accelerating AI workloads, especially inferencing
-operations, has gotten a lot of attentions from academia and industry
+operations, has received a lot of attention from academia and industry
 in recent years as applications of deep learning are broadening to
@@
 interference can nullify the benefits of consolidation~\cite{Kim2016}.
-Because of these reasons, finding a right computing platform is
+Because of these reasons, finding a good computing platform is a
 non-trivial task, one that requires deep understanding of the
-workloads and the hardware platform that executes the workloads.
+workloads and the hardware platform being utilized.
 

### 2017-10-07 paper/overview.tex
Edited overview

@@
 context, the camera latency is the time from the time the camera
-sensor observes the scene to the time the computer actually read the
-digitized image data. Unfortunately, this time can be substantially
-long depending on the camera and the computer performance. We
+sensor observes the scene to the time the computer actually reads the
+digitized image data. Unfortunately, this time can be significantly
+long depending on camera and the computer performance. We
 experimentally measure DeepPicar's camera latency as
@@
 conntected to the Raspberry Pi 3's HDMI output. We then launch a
-simple program that overlay the current time with the camera's preview
-screen on the monitor. Then, infinite number of time strings will be
+simple program that overlays the current time with the camera's preview
+screen on the monitor. Then, an infinite number of time strings will be
 shown on the monitor screen and the time difference between the two
@@
 screen. The measured camera latency is about 300~350 ms. This is
-significantly higher than a human's perception latency, which is known
+significantly higher than the latency of human perception, which is known
 to be as fast as 13 ms~\cite{ThomasBurger2015}. Higher camera latency

### 2017-10-07 paper/intro.tex
Changes to abstract and into

@@
 as well. For example, Tesla Model S was known to use a specialized
-chip (MobileEye EyeQ), which uses a deep neural network for vision-based
+chip (MobileEye EyeQ), which used a deep neural network for vision-based
 real-time obstacle detection and avoidance. More recently, researchers
@@
 the timings. On the other hand, the computing hardware platform must
-also satisfy cost, size, weight, and power constraints, which require
+also satisfy cost, size, weight, and power constraints, which require a
 highly efficient computing platform. These two conflicting
@@
 workloads, we need a testbed and realistic workloads. While using a real car
-based testbed would be most ideal, it is not only highly expensive but also
+based testbed would be most ideal, it is not only highly expensive, but also
 poses serious safety concerns that hinder development and exploration.
@@
 goals are (1) to analyze real-time issues in DNN based end-to-end
-control; (2) to evaluate real-time performance of contemporary embedded
+control; and (2) to evaluate real-time performance of contemporary embedded
 multicore platforms for the AI workloads.
@@
 In this paper, we present DeepPicar, a low-cost autonomous car
-platform for research and education. From hardware perspective,
+platform for research and education. From a hardware perspective,
 DeepPicar is comprised of a Raspberry Pi 3 Model B quad-core
@@
 The network architecture is identical to that of NVIDIA's DAVE-2
-self-driving car~\cite{Bojarski2016}, which uses much more powerful
+self-driving car~\cite{Bojarski2016}, which uses a much more powerful
 computer (Drive PX computer~\cite{drivepx}) than Raspberry Pi~3.
 We chose to use Pi 3 not only because of its
-affordability but also because it is representative
-of today's mainstream low-end embedded multicore platforms, found in
+affordability, but also because it is representative
+of today's mainstream low-end embedded multicore platforms found in
 smartphones and other embedded devices.
@@
 network offline using the collected data on a desktop computer, which
-equips a NVIDIA GTX 1060 GPU. Finally, the trained network is copied
-back to the Raspberry Pi, which is then used to perform interference
+is equipped with a NVIDIA GTX 1060 GPU. Finally, the trained network is copied
+back to the Raspberry Pi, which is then used to perform inference
 operations---locally on the Pi---in the car's main control loop in
 real-time. For real-time control, each inference operation must
-be completed within the desired control period. (e.g., 50ms period for
-20Hz control frequency.)
+be completed within the desired control period (e.g., 50ms period for
+20Hz control frequency).
 % how we evaluated (in terms real-time performance)
@@
 platform shows the potential to run multiple neural network models 
-simultaneously, as there are no indications of direct interference 
+simultaneously as there are no indications of direct interference 
 between them. Instead, we show that memory intensive co-runners can 
@@
   \item We present the design and implementation of a
-    low-cost autonomous vehicle testbed DeepPicar, which utilizes the
+    low-cost autonomous vehicle testbed DeepPicar, which utilizes
     state-of-the-art artificial intelligence techniques.

### 2017-10-07 paper/main.tex
Changes to abstract and into

--- 2017-10-06 paper/main.tex
@@
 architecture---9 layers, 27 million connections and 24K
-parameters---and can be trained to drives itself in real-time using a
+parameters---and can be trained to drive itself, in real-time, using a
 web camera and a modest Raspberry Pi 3 quad-core platform.
@@
 software as a workload. 
-We find all tested platforms, including the Pi3, are capable of
+We find all tested platforms, including the Pi 3, are capable of
 supporting deep-learning based real-time control, from 20 Hz up to 100
-Hz, depending on hardware platform. 
+Hz depending on hardware platform. 
 However, shared resource contention remains an

### 2017-10-07 paper/evaluation.tex
comparison revision done
Edited evaluation

@@
 GPU, although we did not use the GPU in our evaluation due to the lack
-of sofware support (Tensorflow is not compatible with the GPU).
+of sofware support (Tensorflow is not compatible with the Raspberry Pi's GPU).
 For software, we use Ubuntu MATE 16.04, Tensorflow 1.1 and python
@@
 the 99 percentail time is 50.70ms.
-This means that the DeepPicar can operates
+This means that the DeepPicar can operate
 at about 20 Hz control frequency in real-time using only the on-board
@@
 performance differences among multiple embedded computing platforms,
-including NVIDIA's Jetson TK2, later in
+including NVIDIA's Jetson TX2, later in
 Section~\ref{sec:comparison}), the fact that the low-cost
@@
 In this experiment, we investigate the scalability of performing
-inference operation of DeepPicar's nueral network with respect to the
+inference operations of DeepPicar's neural network with respect to the
 number of cores. As noted earlier, the Raspberry Pi 3 platform has
-four Cortex-A53 cores and the Tensorflow 
-provides a programmable mechanism to to adjust how many cores to be
+four Cortex-A53 cores and Tensorflow 
+provides a programmable mechanism to to adjust how many cores are to be
 used by the library. Leveraging this feature, we repeat the
 same experiment described in the previous subsection with varying
-number of CPU cores---from one to four.
+numbers of CPU cores---from one to four.
 
@@
 
-Figure~\ref{fig:perf-vs-corecnt} shows the average execution times of
-the control loop as we vary the number of cores used by the
+Figure~\ref{fig:perf-vs-corecnt} shows the average execution time of
+the control loop as we vary the number of cores used by
 Tensorflow. As expected as we assign more cores, the average execution
@@
 The poor scalability opens up the possibility of consolidating
-multiple different tasks or different nueral network models rather
-than allocating all cores for a single nueral network model. For
+multiple different tasks or different neural network models rather
+than allocating all cores for a single neural network model. For
 example, it is conceivable to use four cameras and four different
-neural network models, each of which is trained seperately and
+neural network models, each of which is trained separately and
 executed on a single dedicated core. Assuming we use the same network
-architecture for all models, then one might exepct to achieve up to
+architecture for all models, then one might expect to achieve up to
 15 Hz using one core (given 1 core can deliver 62ms average
@@
 timings. In particular, we are interested in how shared resource
-contention affect inference timing.
+contention affects inference timing.
 
@@
 
-Figure~\ref{fig:perf-vs-modelcnt} shows the results. In the figure,
+Figure~\ref{fig:perf-vs-modelcnt} shows the results. In the figure, the
 X-axis shows the system configuration: \#of DNN models x \#of CPU
@@
 cores, are co-scheduled, the average inference timing is increased by
-about 7 ms, or 10\%, comapred to the baseline of running one model
+about 7 ms, or 10\%, compared to the baseline of running one model
 using two CPU cores.
@@
 suggests that the shared L2 cache is not the bottleneck that caused
-execution time increases. In other words, DNN models appear to be not
+execution time increases. In other words, DNN models appear not to be
 sensitive to the shared L2 cache space. % L2 partitioning is probably
@@
 in increased memory access latency. While some Intel processors
-provide uncore hardware counters that can measure average memory
+provide oncore hardware counters that can measure average memory
 access latency~\cite{ye2016maracas}, we were not able to identify
@@
 
-\subsection{Effect of Co-scheduling Memory Performance Hogs}
+\subsection{Effect of Co-scheduling Memory Performance Hogs}\label{sec:eval-memhog}
 
@@
   \caption{Effect of memory performance hogs on the DNN
-    inferencing. The DNN model uses Core 0 and memory-hog co-runnes
+    inferencing. The DNN model uses Core 0 and memory-hog co-runners
     use the rest of the cores.}
@@
 dedicated core (Core0). On the other hand, the DNN model's L2
-cache-miss rates are not increase as much. This suggests that
+cache-miss rates did not increase as much. This suggests that
 the DNN model's exeuction increase cannot be fully explained by
@@
 were pleasantly suprised by this finding. The time breakdown shows that
-the DNN inferencing operation, perforemd by the Tensorflow library,
+the DNN inferencing operation, performed by the Tensorflow library,
 dominates the execution time, which is expected.
@@
 Second, we find that scalability of Tensorflow's DNN inference 
-implementation is limited: We find that using all four cores are 
+implementation is limited: We find that using all four cores is 
 only about 30\% better than using just a single core.
@@
 
-Lastly, we however find that consolidating memory (DRAM) performance
+Lastly, we find that consolidating memory (DRAM) performance
 intensive applications could jeopadize DNN performance because DNN
-performance appear very sensitive to memory performance; we observe up
+performance appears very sensitive to memory performance; we observe up
 to 9.6X slowdown in DNN performance by co-scheduling synthetic memory 

### 2017-10-07 paper/conclusion.tex

@@
 We presented DeepPicar, a low cost autonomous car platform that is
-inexpensive to build but is based on a state-of-the art AI technology:
+inexpensive to build, but is based on state-of-the-art AI technology:
 End-to-end deep learning based real-time control.
 Specifically, DeepPicar uses a deep convolutional neural network to
-predict steering angles of the car directly from a camera input data
+predict steering angles of the car directly from camera input data
 in real-time. Importantly, DeepPicar's neural network architecture is
@@
 resource. We systematically analyzed the real-time characteristics of
-the Pi3 platform in the context of deep-learning based real-time
+the Pi 3 platform in the context of deep-learning based real-time
 control appilcations, with a special emphasis on real-time deep neural
 network inferencing.
-We also evaluated other more powerful embedded computing
+We also evaluated other, more powerful, embedded computing
 platforms to better understand achievable real-time performance of
@@
 also plan to improve the prediction accuracy by feeding more data and
-upgrade the RC car hardware platform to enable more precise steering
+upgrading the RC car hardware platform to enable more precise steering
 angle control.

### 2017-10-07 paper/comparison.tex
Edited comparison

@@
 the number of CPU cores used.
-First, both the Intel UP and Tegra TX2 exhibit superior performance
+First, both the Intel UP and Tegra TX2 exhibit superior performance when
 compared with the Raspberry Pi 3. 
@@
 timing in Intel Up and TX2, but the degree of the increase is not as
-dramatic as the Pi3. Compared to their respective solo timing (i.e.,
+dramatic as the Pi3. Compared to their respective solo timings (i.e.,
 the model runs on a single core in isolation), Intel UP suffers up to
@@
 2.2X and 2.5X increase, respectively. This is somewhat suprising
-because Raspberry Pi 3's cores are in-oder architecture based while
+because Raspberry Pi 3's cores are in-order architecture based while
 the cores in Intel Up and NVIDIA TX2 are out-of-order architecture

### 2017-10-07 paper/evaluation.tex
Changes to evaluation

@@
 of sofware support (Tensorflow is not compatible with the Raspberry Pi's GPU).
-For software, we use Ubuntu MATE 16.04, Tensorflow 1.1 and python
+For software, we use Ubuntu MATE 16.04, Tensorflow 1.1 and Python
 2.7. We disabled DVFS (dynamic voltage frequency scaling) and
-configure the clock speed of each core at the maximum 1.2GHz.
+configured the clock speed of each core at the maximum 1.2GHz.
 
@@
 First, as expected, we find that the inference operation
-dominates the control loop execution time, accounting about 85\% of
+dominates the control loop execution time, accounting for about 85\% of
 the execution time.
@@
 resources. We consider these results respectable given the complexity
-of the deep neural network  and the fact that the inference operation
+of the deep neural network, and the fact that the inference operation
 performed by Tensorflow only utilizes the CPU cores of the
@@
 the control loop as we vary the number of cores used by
-Tensorflow. As expected as we assign more cores, the average execution
+Tensorflow. As expected, as we assign more cores, the average execution
 time decreases---from 61.96 ms on a single core to 42.67 ms on four
@@
 In this experiment, we launch multiple instances of DeepPicar's DNN
-model at the same time and measure its impact to their inference
+model at the same time and measure its impact on their inference
 timings. In particular, we are interested in how shared resource
@@
 X-axis shows the system configuration: \#of DNN models x \#of CPU
-cores\/DNN. For example, '4x1' means running four DNN models each of
+cores/DNN. For example, '4x1' means running four DNN models each of
 which is assigned to run on one CPU core, whereas '2x2' means running
@@
 miss rates of the DNN models first in isolation and then after
-co-scheduling another models. If the shared L2 cache is the primary
+co-scheduling other models. If the shared L2 cache is the primary
 source of inteference, then the measured L2 miss rates would
@@
 suggests that the shared L2 cache is not the bottleneck that caused
-execution time increases. In other words, DNN models appear not to be
+execution time increases. In other words, DNN models don't appear to be
 sensitive to the shared L2 cache space. % L2 partitioning is probably
@@
 increased---by up to 9.6X---even though the DNN model is running on a
-dedicated core (Core0). On the other hand, the DNN model's L2
+dedicated core (Core 0). On the other hand, the DNN model's L2
 cache-miss rates did not increase as much. This suggests that
@@
 Lastly, we find that consolidating memory (DRAM) performance
-intensive applications could jeopadize DNN performance because DNN
-performance appears very sensitive to memory performance; we observe up
+intensive applications could jeopadize DNN performance, because DNN
+performance appears to be very sensitive to memory performance; we observe up
 to 9.6X slowdown in DNN performance by co-scheduling synthetic memory 

### 2017-10-07 paper/appendix.tex
Changes to comparison and appendix

@@
 frames for validation.
-The weights of the network is initialized using Xavier
+The weights of the network are initialized using Xavier
 initializer~\cite{Glorot2010}, which is known to provide better
@@
 
-In the second method, we use the same data and parameters above except
-that now images are labeled as 'curved' and 'straight' and we pick an
+In the second method, we use the same data and parameters as  above 
+except that now images are labeled as 'curved' and 'straight' and we pick an
 equal number of images from each category at each training step to
@@
 center of the track (on the white tape).
-However, we find that there is huge discrepency between the training
-loss and the validation loss as the former was 0.09 while the latter
-was 0.0869---a 10X difference---indicating that the model suffers from
+However, we find that there is a huge discrepency between the training
+loss and the validation loss as the former was 0.09, while the latter
+was 0.00869---a 10X difference---indicating that the model suffers from
 the overfitting problem.
@@
 We continue to investigate ways to achieve better prediction accuracy
-in training the network as well as improving the performance of the RC
+in training the network, as well as improving the performance of the RC
 car platform, especially on precise steering angle control.
@@
 In experiments conducted with a power supply that only provided 1 Amp,
-the Pi was unable to sustain a 1.2 GHz clock speed and, instead,
+the Pi was unable to sustain a 1.2 GHz clock speed, and instead,
 fluctuated between operating at 600 MHz and 1.2 GHz. As a result, it
@@
 
-Our initial experiment results suffered from these issues after which
+Our initial experiment results suffered from these issues, after which
 we always carefully monitored the current operating frequencies of the

### 2017-10-07 paper/comparison.tex
Changes to comparison and appendix

@@
 multi-model co-scheduling experiment. Once again, they can comfortably
-20 Hz real-time performance for all of the co-scheduled DNN control
+satisfy 20 Hz real-time performance for all of the co-scheduled DNN control
 loops, and in case of TX 2 (GPU), 100 Hz real-time control is still
@@
 10ms deadline for 100 Hz control that would have been feasible if
-there is no contention between the CPU cores and the GPU.
+there was no contention between the CPU cores and the GPU.
 
@@
 workloads. However, shared resource contention among these diverse
-computing resources remain an important issue that must be understood
-and controlled especially for safety-critical applications.
+computing resources remains an important issue that must be understood
+and controlled, especially for safety-critical applications.
 

### 2017-10-07 paper/intro.tex
minor fix in intro
minor edits in the intro

@@
 poses serious safety concerns that hinder development and exploration.
-Therefore, there is a strong need for safer and less costly testbeds
-that still employ state-of-the-art AI technologies. There are already
-several relatively inexpensive RC-car based testbeds, such as MIT
+Therefore, there is a strong need for safer and less costly
+testbeds. There are already several relatively inexpensive RC-car
+based testbeds, such as MIT 
 RaceCar~\cite{shin2017project} and UPenn's F1$/$10 racecar~\cite{upennf1tenth}.
-However, these RC-car testbeds still cost more than \$3,000.
+However, these RC-car testbeds still cost more than \$3,000, requiring
+considerable investment.
 
@@
 Instead, we want to build a low cost testbed that still employs the
-state-of-the art AI technologies. Specifically, we focus on a deep
-convolutional neural network (CNN) based end-to-end control system,
+state-of-the art AI technologies. Specifically, we focus on a end-to-end
+deep learning based real-time control system,
 which was developed for a real self-driving car, NVIDIA
 DAVE-2~\cite{Bojarski2016}, and use the same methodology on a
-smaller, safer and less costly setup. In developing the testbed, our
+smaller and less costly setup. In developing the testbed, our
 goals are (1) to analyze real-time issues in DNN based end-to-end
 control; and (2) to evaluate real-time performance of contemporary embedded
-multicore platforms for the AI workloads.
+platforms for such workload.
 

### 2017-10-07 paper/intro.tex
intro revision

@@
 % what are our findings?
-From our experiments, we discover that the DeepPicar is capable of 
+Using the DeepPicar platform, we systematically analyzed its real-time
+capabilities in the context of deep-learning based real-time
+control, especially on real-time deep neural network inferencing.
+We also evaluated other, more powerful, embedded computing
+platforms to better understand achievable real-time performance of
+DeepPicar's deep-learning based control system and the impact of
+computing hardware architectures.
+
+We find that the DeepPicar is capable of 
 completing control loops in under 50 ms, or 20 hz, and can do so 
-almost 99\% of the time. The number of cores used by the platform has 
-an unexpectedly low impact on its performance, only decreasing the 
-control loop times by 30\% when using 4 cores instead of 1.  Our 
-platform shows the potential to run multiple neural network models 
-simultaneously as there are no indications of direct interference 
-between them. Instead, we show that memory intensive co-runners can 
-be detrimental to the real-time performance of the DeepPicar. 
-Finally, while we find that the Raspberry Pi 3 takes longer to 
-complete control loops when compared to other embedded computing 
-platforms, namely the Intel UP Board and the NVIDIA Tegra TX2, we 
-still conclude that the Pi 3 is a viable option for autonomous 
-vehicle research.
+almost 99\% of the time.
+Other tested embedded platforms, Intel Up and NVIDIA TX2, offer even
+better performance, capable of supporting deep-learning based real-time
+control up to 100 Hz control frequency on the TX2 when the GPU is used.
+However, in all platforms, shared resource contention remains an
+important issue as we observe up to 9.6X control loop execution time
+increase, mostly due to increase in the nueral network inferencing
+operation, when memory performance intensive applications are
+co-scheduled on idle CPU cores.
+
+%% that must be considered in applying deep-learning
+%% models on shared memory based embedded computing platforms.
+%% The number of cores used to process the neural network by the platform has 
+%% an unexpectedly low impact on its performance, only decreasing the 
+%% control loop times by 30\% when using 4 cores instead of 1.  Our 
+%% platform shows the potential to run multiple neural network models 
+%% simultaneously as there are no indications of direct interference 
+%% between them. Instead, we show that memory intensive co-runners can 
+%% be detrimental to the real-time performance of the DeepPicar. 
+%% Finally, while we find that the Raspberry Pi 3 takes longer to 
+%% complete control loops when compared to other embedded computing 
+%% platforms, namely the Intel UP Board and the NVIDIA Tegra TX2, we 
+%% still conclude that the Pi 3 is a viable option for autonomous 
+%% vehicle research.
 

### 2017-10-11 paper/appendix.tex
intro revision
Minor changes to paper

@@
+\begin{figure*}[t]
+  \centering
+  \includegraphics[width=.9\textwidth]{figs/TrainingLoss}
+  \caption{Change in loss value throughout training.}
+  \label{fig:modelloss}
+\end{figure*}
+
 \subsection{DNN Training and Testing}
@@
 different track conditions, different combinations of input
-data, and different hyper parameters. In the following, we describe
-details on two of the training methods that performed reasonably well.
+data, and different hyper parameters. In the following paragraphs, we 
+describe details on two of the training methods that performed 
+reasonably well.
 
@@
 
-\begin{figure*}[t]
-  \centering
-  \includegraphics[width=.9\textwidth]{figs/TrainingLoss}
-  \caption{Change in loss value throughout training.}
-  \label{fig:modelloss}
-\end{figure*}
-
 In the second method, we use the same data and parameters as  above 
@@
 in training the network, as well as improving the performance of the RC
-car platform, especially on precise steering angle control.
+car platform, especially related to precise steering angle control.
 

### 2017-10-11 paper/comparison.tex
Minor changes to paper

@@
 As a result, they are all able to satisfy the 50 ms 
-WCET by a clear margin
+WCET by a clear margin,
 %%  (in the worst case, the UP Board would 
 %% still meet it by $\sim$20 ms) --> where's the data showing this?
-, and, in case of TX2, 50 Hz or even 100 Hz real-time control is
+and, in case of TX2, 50 Hz or even 100 Hz real-time control is
 feasible with the help of its GPU. Another observation is that TX2
-(GPU) does not change much as most of the neural network computation
+(GPU) does not change much, as most of the neural network computation
 is done at the GPU.
@@
 As discussed in 
-Section~\ref{sec:eval-memhog}, we observe dramatic execution time
-increase, up to 9.4X, in Raspberry Pi 3 as we increase the number of
+Section~\ref{sec:eval-memhog}, we observed dramatic execution time
+increases, up to 9.4X, in Raspberry Pi 3 as we increased the number of
 co-scheduled tasks. We also observe increased control loop execution
@@
 the cores in Intel Up and NVIDIA TX2 are out-of-order architecture
-based and that the memory intensive tasks on out-of-order cores can
+based, and that the memory intensive tasks on out-of-order cores can
 generate more memory traffic. We suspect that this is because the
@@
 known characteristic of integrated CPU-GPU architectue based
-platforms that both CPU and GPU share the same memory
+platforms where both CPU and GPU share the same memory
 subsystem~\cite{Ali2017}. As a result, TX2 (GPU) fails to meet the

### 2017-10-11 paper/evaluation.tex
Minor changes to paper

@@
 factors in determining a proper control loop latency. While a standard
-control system may be comprized of multiple control loops with
-differning control frequencies---e.g., a inner control loop for lower-level
+control system may be comprised of multiple control loops with
+differing control frequencies---e.g., a inner control loop for lower-level
 PD control, a outer loop for motion planning, etc.---DeepPicar's
 control loop is a single layer, as shown earlier in
-Figure~\ref{fig:controlloop}, as a single deep neural network
+Figure~\ref{fig:controlloop}, since a single deep neural network
 replaces the traditional multi-layer control pipline. (Recall
@@
 the execution time.
-Second, more importantly, we also find that the measured average
+Second, and more importantly, we also find that the measured average
 execution time of a single control loop is 42.67 ms, or 23.4 Hz and
-the 99 percentail time is 50.70ms.
+the 99 percentail time is 50.70 ms.
 This means that the DeepPicar can operate
@@
 four Cortex-A53 cores and Tensorflow 
-provides a programmable mechanism to to adjust how many cores are to be
+provides a programmable mechanism to adjust how many cores are to be
 used by the library. Leveraging this feature, we repeat the
@@
 linear scaling. In particular, from 2 cores to 3 cores, the
-improvement is mere 2.38ms (or 4\%). In short, we find that the
+improvement is mere 2.38 ms (or 4\%). In short, we find that the
 scalability of DeepPicar's deep neural network is not ideal on the
 platform. We do not know whether it is due to the limitations of
-Tensorflow's multicore implementation or it's the model's inherent
+Tensorflow's multicore implementation or if it's the model's inherent
 characteristics.
@@
 architecture for all models, then one might expect to achieve up to
-15 Hz using one core (given 1 core can deliver 62ms average
+15 Hz using one core (given 1 core can deliver 62 ms average
 execution time). In the next experiment, we investigate the
@@
 cores. The Y-axis shows the average inference timing.
-Two bars in the left show the impact of co-scheduling four DNN
+The two bars on the left show the impact of co-scheduling four DNN
 models. Compared to executing a single DNN model on one CPU core
@@
 expected and likely caused by contention in the shared hardware
-resources such as the shared L2 cache and/or the DRAM controller.
-To further analyze the source of contention, we use hardwawre
+resources, such as the shared L2 cache and/or the DRAM controller.
+To further analyze the source of contention, we use hardware
 performance counters of the processor. Specifically, we measure L2
@@
 co-scheduling other models. If the shared L2 cache is the primary
-source of inteference, then the measured L2 miss rates would
+source of inteference, then the measured L2 miss rates will
 increase. Figure~\ref{fig:l2missrate-vs-modelcnt} shows the results.
 As can be see in the figure, L2 miss rates are largely unchanged
-regardless whether multiple models are co-scheduled or not. This
+regardless of whether multiple models are co-scheduled or not. This
 suggests that the shared L2 cache is not the bottleneck that caused
@@
 dedicated core (Core 0). On the other hand, the DNN model's L2
-cache-miss rates did not increase as much. This suggests that
+cache-miss rates do not increase as much. This suggests that
 the DNN model's exeuction increase cannot be fully explained by
@@
 techniques~\cite{Gracioli2015,Kim2016} may not be effective isolation
-solutions for DeepPicar as its AI workload is more sensitive to memory
+solutions for DeepPicar, as its AI workload is more sensitive to memory
 performance. Instead, memory controller focused isolation solutions,
@@
 perform real-time control of the RC car at 20 Hz frequency (or
-50ms per control loop). Given the complexity of the DNN we use, we
+50ms per control loop). Given the complexity of the DNN used, we
 were pleasantly suprised by this finding. The time breakdown shows that
@@
 Second, we find that scalability of Tensorflow's DNN inference 
-implementation is limited: We find that using all four cores is 
+implementation is limited. We find that using all four cores is 
 only about 30\% better than using just a single core.

### 2017-10-11 paper/intro.tex
Minor changes to paper

@@
 To understand what kind of computing hardware is needed for AI
-workloads, we need a testbed and realistic workloads. While using a real car
-based testbed would be most ideal, it is not only highly expensive, but also
+workloads, we need a testbed and realistic workloads. While using a real 
+car-based testbed would be most ideal, it is not only highly expensive, but also
 poses serious safety concerns that hinder development and exploration.

### 2017-10-11 paper/appendix.tex
Add picar-mini models and source data
MiniModel
train=evens in range(30,60)
val = odds in range(30,60)
use_normal_category = True
train loss = 0.009152407757937908
val loss = 0.08686625957489014

MiniModel2
train=evens in range(30,60)
val = odds in range(30,60)
use_normal_category = False
train loss = 0.018782231956720352
val loss = 0.013224977999925613

@@
 However, we find that there is a huge discrepency between the training
-loss and the validation loss as the former was 0.09, while the latter
-was 0.00869---a 10X difference---indicating that the model suffers from
+loss and the validation loss as the former was 0.009, while the latter
+was 0.0869---a 10X difference---indicating that the model suffers from
 the overfitting problem.

### 2017-10-13 test-model.py
Remove unused models/source data

@@ -26,6 +26,8 @@ config = tf.ConfigProto(intra_op_parallelism_threads=NCPU, inter_op_parallelism_
                         allow_soft_placement=True, device_count = {'CPU': 1})
 # sess = tf.Session(config=config)
 
+NFRAMES = 1000
+
 sess = tf.InteractiveSession(config=config)
 saver = tf.train.Saver()
 model_name = 'model.ckpt'
@@ -38,6 +40,7 @@ epoch_ids = [6] # DBG - heechul
 
 tot_time_list = []
 
+curFrame = 0
 for epoch_id in epoch_ids:
     print '---------- processing video for epoch {} ----------'.format(epoch_id)
     # vid_path = cm.jn(params.data_dir, 'epoch{:0>2}_front.mkv'.format(epoch_id))
@@ -52,28 +55,30 @@ for epoch_id in epoch_ids:
     print 'performing inference...'
     time_start = time.time()
     for frame_id in xrange(frame_count):
-        cam_start = time.time()
-        ret, img = cap.read()
-        assert ret
+        if curFrame < NFRAMES:
+            cam_start = time.time()
+            ret, img = cap.read()
+            assert ret
 
-        prep_start = time.time()
-        img = preprocess.preprocess(img)
+            prep_start = time.time()
+            img = preprocess.preprocess(img)
 
-        pred_start = time.time()
-        rad = model.y.eval(feed_dict={model.x: [img], model.keep_prob: 1.0})[0][0]
-        deg = rad2deg(rad)
-        pred_end   = time.time()
+            pred_start = time.time()
+            rad = model.y.eval(feed_dict={model.x: [img], model.keep_prob: 1.0})[0][0]
+            deg = rad2deg(rad)
+            pred_end   = time.time()
 
-        cam_time  = (prep_start - cam_start)*1000
-        prep_time = (pred_start - prep_start)*1000
-        pred_time = (pred_end - pred_start)*1000
-        tot_time  = (pred_end - cam_start)*1000
+            cam_time  = (prep_start - cam_start)*1000
+            prep_time = (pred_start - prep_start)*1000
+            pred_time = (pred_end - pred_start)*1000
+            tot_time  = (pred_end - cam_start)*1000
 
-        print 'pred: {:0.2f} deg. took: {:0.2f} ms | cam={:0.2f} prep={:0.2f} pred={:0.2f}'.format(deg, tot_time, cam_time, prep_time, pred_time)
[…]

### 2017-10-13 paper/background.tex
Changes to paper

@@
 (CNN); and most recently in NVIDIA's DAVE-2
-project~\cite{Bojarski2016}, using 9 layer CNN. In all these projects,
+project~\cite{Bojarski2016}, using a 9 layer CNN. In all of these projects,
 the neural network models take raw image pixels as input and directly
-produce steering control commands bypassing all intermediary steps and
-hand-written rules in conventional robotics control approach.  
+produce steering control commands, bypassing all intermediary steps and
+hand-written rules used in conventional robotics control approach.  
 NVIDIA's latest effort reports that their trained CNN
 autonomously controls their modified cars on public roads without human
-intervention successfully~\cite{Bojarski2016}.
+intervention~\cite{Bojarski2016}.
 
@@
 phases~\cite{NVIDIA2015}. The first
-phase is \emph{training} during which the weights of the network is
+phase is \emph{training} during which the weights of the network are
 incrementally updated by backpropagating errors it sees from the
@@
 the network minimize errors in the training examples---the next phase
-is \emph{inferencing} during which unseen data is fed to the network
+is \emph{inferencing}, during which unseen data is fed to the network
 as input to produce predicted output (e.g., predicted image
@@
 other hand, is relatively less computationally intensive and latency becomes
-as important, if not more, as computational throughput because many
+as important, if not moreso, as computational throughput, because many
 use-cases have strict real-time requirements (e.g., search query
@@
 Real-time embedded systems, such as an autonomous vehicle, present
-unique challenges for deep learning as the computing platforms of such
+unique challenges for deep learning, as the computing platforms of such
 systems must satisfy two often conflicting goals~\cite{Otterness2017}:
@@
 in recent years as applications of deep learning are broadening to
-areas real-time embedded systems such as autonomous vehicles. These
-efforts include development of various heterogeneous architecture
-based system-on-a-chips (SOCs) that may include multiple cores, GPU,
+areas of real-time embedded systems such as autonomous vehicles. These
+efforts include the development of various heterogeneous architecture-based 
+system-on-a-chips (SOCs) that may include multiple cores, GPU,
 DSP, FPGA and neural network optimized ASIC hardware~\cite{Jouppi2017}.
@@
 interference can nullify the benefits of consolidation~\cite{Kim2016}.
-Because of these reasons, finding a good computing platform is a
+For these reasons, finding a good computing platform is a
 non-trivial task, one that requires deep understanding of the
@@
 
-The primary objectives of this study are (1) to understand the
-necessary computing performance for applying AI technology based
+The primary objectives of this study are to understand (1) the
+necessary computing performance for applying AI technology-based
 robotics systems, and (2) what kind of computing architecture and
-runtime supports are most appropriate for such workload. Toward to
+runtime supports are most appropriate for such workload. To
 achieve these goals, we implement a low-cost autonomous car platform
-as a case-study and systematically conduct experiments, as we will 
+as a case-study and systematically conduct experiments, which we will 
 describe in the subsequent sections.

### 2017-10-13 paper/comparison.tex
Changes to paper

@@
 TX2~\footnote{http://www.nvidia.com/object/embedded-systems-dev-kits-modules.html}---from
-the point of view of supporting vision-based end-to-end deep learning
+the point of view of supporting end-to-end deep learning
 based autonomous vehicles. 
@@
 
-The UP Board and TX2 also perform much better when multiple DNN models
+The Intel UP board and TX2 also perform much better when multiple DNN models
 are co-scheduled. Figure~\ref{fig:sys_model} shows the results of the
@@
 accomodate multiple instances of the DNN model we tested.
-
-%% to complete control loops in a fraction of the time when compared to 
-%% the Pi. Furthermore, both of the platforms satisfied the 50 ms WCET 
-%% in all experiments, while the Raspberry Pi 3 was unable to do so. As 
-%% such, the UP Board and TX2 display greater potential for the 
-%% simultaneous execution of multiple models during self-driving 
-%% operation, while the Raspberry Pi 3 would struggle to do so.
 
@@
 
-Another intersting observation is that TX2 (GPU) also suffers
+Another interesting observation is that TX2 (GPU) also suffers
 considerable execution time increase (2.5X) despite the fact that the
@@
 
-%% experienced a change in their inferencing times, but the increases 
-%% were not as drastic as is seen by the Raspberry Pi 3. in the worst 
-%% case, both the UP Board and TX2 (CPU only and GPU) produced times 
-%% that were over two times as large, whereas the Raspberry Pi 3 output 
-%% times that were around 10 times greater. The Intel UP Board was 
-%% successful in completing inferencing operations in under 50 ms when 
-%% benchmarks were run on a maximum of 2 cores, and the TX2 was 
-%% successful in all experiments irrespective of the computing resources 
-%% utilized. Comparetively, both performed better than the Raspberry Pi 
-%% 3 which failed to meet its deadlines when any benchmark was 
-%% introduced. As such, it can be concluded that the Intel UP Board 
-%% would be more capable of running computationally heavy processes 
-%% during autonomous driving, and that the Pi wouldn't be able to do the 
-%% same.
-
 In summary, we find that today's embedded computing platforms, even as
@@
 and controlled, especially for safety-critical applications.
-
-%% the comparison of the real-time capabilities of three embedded 
-%% computing platforms, we found that the Raspberry Pi 3 performed the 
-%% worst. When compared to the Intel UP Board and the NVIDIA Tegra TX2, 
-%% control loops on the Pi were shown to take noticably longer to 
-%% complete, regardless of the number of cores utilized or the number of 
-%% neural network models run simultaneoulsy. The difference was even 
-%% more noticeable in the the addition of computationally heavy 
-%% synthetic benchmarks that had a much more dire effect on the Pi. In 
-%% the worst case, an average control loop on the Pi took an abnormally 
-%% long amount of time to finish. Compared to the TX2 with GPU support, 
-%% those times were 34.8X times greater. From this, we find that the UP 
-%% Board and Tegra TX2 are superior platforms for real-time CNN 
-%% operations. However, due to the Pi's ability to satisfy the 50 ms 
-%% WCET in some experiments, we conclude that the Raspberry Pi 3 is 
-%% feasible for autonomous research.

### 2017-10-13 paper/conclusion.tex
Changes to paper

@@
 
-%% We evaluated the real time efficacy of our platform and found
-%% that it is capable of performing inferencing operations within a WCET
-%% of 50 ms, or 20 fps, when a single model is run on more than one
-%% core.
-
 As future work, we plan to apply shared resource management
 techniques~\cite{Yun2013,yun2014rtas} on the DeepPicar platform and
-evaluate their impact to the real-time performance of the system. We
+evaluate their impact on the real-time performance of the system. We
 also plan to improve the prediction accuracy by feeding more data and
@@
 angle control.
-
-%% Further studies should be conducted on the interference that
-%% occurs between cores to allow for improved performance in utilizing
-%% multiple models for autonomous vehicles. With minimal further
-%% research, the DeepPicar can become a valudable, reliable and
-%% inexpensive tool in the continued research of autonomous vehicles.

### 2017-10-13 paper/evaluation.tex
Changes to paper

@@
 The Raspberry Pi 3 Model B platform used in DeepPicar equips a Boardcom
-BCM2837 SoC, which has a quad-core 64bit ARM Cortex-A53 cluster,
+BCM2837 SoC, which has a quad-core ARM Cortex-A53 cluster,
 running at up to 1.2GHz. Each core has 16K private I\&D caches, and all
@@
 GPU, although we did not use the GPU in our evaluation due to the lack
-of sofware support (Tensorflow is not compatible with the Raspberry Pi's GPU).
-For software, we use Ubuntu MATE 16.04, Tensorflow 1.1 and Python
+of sofware support (TensorFlow is not compatible with the Raspberry Pi's GPU).
+For software, we use Ubuntu MATE 16.04, TensorFlow 1.1 and Python
 2.7. We disabled DVFS (dynamic voltage frequency scaling) and
-configured the clock speed of each core at the maximum 1.2GHz.
+configured the clock speed of each core statically at the maximum 1.2GHz.
 
@@
 computation time and the type of particular physical system are
-factors in determining a proper control loop latency. While a standard
+factors in determining a proper control loop frequency. While a standard
 control system may be comprised of multiple control loops with
@@
 breakdown of each control loop. Note that all four CPU cores of the
-Raspberry Pi 3 were used by the Tensorflow library when performing the
+Raspberry Pi 3 were used by the TensorFlow library when performing the
 DNN inference operations.
@@
 of the deep neural network, and the fact that the inference operation
-performed by Tensorflow only utilizes the CPU cores of the
+performed by TensorFlow only utilizes the CPU cores of the
 Raspberry Pi 3 (its GPU is not supported by Tensorflow.)
@@
 In comparison, NVIDIA's DAVE-2 system, which has the exact same neural
-network architecture, is reportedly run at 30 Hz~\cite{Bojarski2016},
+network architecture, reportedly run at 30 Hz~\cite{Bojarski2016},
 just a bit faster than the DeepPicar. Although we believe it was not
@@
 number of cores. As noted earlier, the Raspberry Pi 3 platform has
-four Cortex-A53 cores and Tensorflow 
+four Cortex-A53 cores and TensorFlow 
 provides a programmable mechanism to adjust how many cores are to be
@@
 \end{figure}
-    %% \fixme{instead of the total control loop time, use the
-    %%   average of DNN inferencing time.}}
-
-%% \begin{table}[h]
-%%   \centering
-%%   \begin{tabular} {| c | l | l | l | l | l | l | l | l | l |}
-%%     \hline
-%%     \textbf{\#of cores} & \textbf{Mean} & \textbf{Max} & \textbf{99pct} & \textbf{Stdev} \\ \hline 
-%%     1 & 61.96 & 66.00 & 63.31 & 0.51\\ \hline
-%%     2 & 50.49 & 71.55 & 70.03 & 3.16 \\ \hline
-%%     3 & 48.11 & 72.22 & 58.45 & 4.18 \\ \hline
-%%     4 & 42.67 & 56.37 & 50.70 & 2.80 \\
-%%     \hline
-%%   \end{tabular}
-%%   \caption{Execution time statistics vs. \#of CPU cores.}
-%% \end{table}
 
@@
 the control loop as we vary the number of cores used by
-Tensorflow. As expected, as we assign more cores, the average execution
+TensorFlow. As expected, as we assign more cores, the average execution
 time decreases---from 61.96 ms on a single core to 42.67 ms on four
@@
 platform. We do not know whether it is due to the limitations of
-Tensorflow's multicore implementation or if it's the model's inherent
+TensorFlow's multicore implementation or if it's the model's inherent
 characteristics.
@@
 
-%% It may not always be the case that all four cores of the Raspberry Pi
-%% 3's Cortex A-53 CPU can be used solely for the purpose of operating an
-%% autonomous vehicle. Thus, we test how the number of cores utilized for
-%% real-time operations affects the DeepPicar's overall ability to
-%% function as an autonomous vehicle platform.
-
-%% The performance of the DeepPicar is summarized in Table III. The
-%% DeepPicar performed better, on average, when it utilized more
-%% cores. With 4 cores, it was able to meet the vast majority of its
-%% deadlines, doing so in almost 99\% of the input frames. The platform
-%% performed the worst when using only 1 core, as it was unable to meet
-%% any of its deadlines.  On average, using 3 cores instead of 2 only had
-%% a performance increase of approximately 2 ms, so the addition of one
-%% core in that specific case offers  relatively little improvement. One
-%% important observation is that the performance was more consistent when
-%% only 1 core was used. As a result, the use of multiple cores is very
-%% beneficial in terms of reducing the time it takes to complete
-%% inference operations, but may result in times that are more volatile.
-
 \subsection{Effect of Co-scheduling Multiple DNN Models}
@@
 model at the same time and measure its impact on their inference
-timings. In particular, we are interested in how shared resource
+timings. In other words, we are interested in how shared resource
 contention affects inference timing.
-
-%% We also test the capability of the DeepPicar to run multiple models at
-%% the same time, and whether each model can successfully perform within
-%% the given deadline. Specifically, the platform is tested in the cases
-%% of running 2 and 4 models simultaneously. For each case, all models
-%% are allocated an equal number of cores. That is, 2 models are given 2
-%% cores each and 4 models are given 1 core each.
 
@@
   \caption{Timing impact of co-scheduling multiple DNNs. 1Nx1C: one DNN
-    model using one CPU core; 4Nx1C: four DNN models each using one CPU;
-    1Nx2C: one DNN model using two CPU cores; 2Nx2C: two DNN models each
-    using two CPU cores.} 
+    model using one core; 4Nx1C: four DNN models each using one core;
+    1Nx2C: one DNN model using two cores; 2Nx2C: two DNN models each
+    using two cores.} 
   \label{fig:perf-vs-modelcnt}
@@
 
-%% \begin{figure}[h]
-%%   \centering
-%%   \includegraphics[width=.5\textwidth]{figs/2ModelChart}
-%%   \caption{ Change in inferencing time when 2 models are run concurrently. }
-%% \end{figure}
-%% \begin{figure}[h]
-%%   \centering
-%%   \includegraphics[width=.5\textwidth]{figs/4ModelChart}
-%%   \caption{Change in inferencing time when 4 models are run simultaneously. }
-%% \end{figure}
-
 Figure~\ref{fig:perf-vs-modelcnt} shows the results. In the figure, the
 X-axis shows the system configuration: \#of DNN models x \#of CPU
-cores/DNN. For example, '4x1' means running four DNN models each of
-which is assigned to run on one CPU core, whereas '2x2' means running
+cores/DNN. For example, '4Nx1C' means running four DNN models each of
+which is assigned to run on one CPU core, whereas '2Nx2C' means running
 two DNN models, each of which is assigned to run on two CPU
@@
 models. Compared to executing a single DNN model on one CPU core
-(1x1), when four DNN models are co-scheduled (4x1), each model
+(1Nx1C), when four DNN models are co-scheduled (4Nx1C), each model
 suffers an average inference time increase of approximately 15 ms,
 $\sim$24\%. On the other hand, when two DNN models, each using two CPU
-cores, are co-scheduled, the average inference timing is increased by
+cores, are co-scheduled (2Nx2C), the average inference timing is increased by
 about 7 ms, or 10\%, compared to the baseline of running one model
-using two CPU cores.
-
-%% The results for the 2 model test are outlined in Figure 3 and Table
-%% IV, and the 4 model test is summarized in  Figure 4 Table V. In the
-%% the 2 model test, both of the models showed average inference time
-%% increases of around  5-7 ms, $\sim$10\%, when compared to a baseline
-%% of 1 model running on two cores.
-
-%% \begin{table}[h]
-%% \centering
-%%   \begin{tabular} {| l | l | l |}
-%%   \hline
-%%   \textbf{num models} & 1 & 2 \\ \hline
-%%   \textbf{L1 refs} &  3.04E+10 &  3.04E+10 \\ \hline
-%%   \textbf{L1 misses} & 4.78E+08 & 4.91E+08 \\ \hline
-%%   \textbf{L1 miss \%} & 1.58 & 1.61 \\ \hline
-%%   \textbf{L2 refs} & 3.31E+09 &  3.91E+09 \\ \hline
-%%   \textbf{L2 misses} & 3.68E+08 & 4.62E+08 \\ \hline
-%%   \textbf{L2 miss \%} & 11.12 & 10.88 \\ \hline
-%%   \end{tabular}
-%%   \caption{ Effect of 2 simulatneous models on cache accesses. }
-%% \end{table}
-
-%% \begin{table}[h]
-%% \centering
-%%   \begin{tabular} {| l | l | l |}
-%%   \hline
-%%   \textbf{num models} & 1 & 4 \\  \hline
-%%   \textbf{L1 refs} & 2.78E+10 & 2.79E+10 \\ \hline
-%%   \textbf{L1 misses} & 4.36E+08 & 4.53E+08 \\ \hline
-%%   \textbf{L1 miss \%} & 1.57 & 1.63 \\ \hline
-%%   \textbf{L2 refs} & 2.83E+09 & 3.36E+09 \\ \hline 
-%%   \textbf{L2 misses} & 3.59E+08 & 4.43E+08 \\ \hline
-%%   \textbf{L2 miss \%} & 12.68 & 13.19 \\ 
-%% \hline
-%%   \end{tabular}
-%%   \caption{ Effect of 4 concurrent models on cache accesses. }
-%% \end{table}
+using two CPU cores (1Nx2C).
 
@@
 in increased memory access latency. While some Intel processors
-provide oncore hardware counters that can measure average memory
+provide incore hardware counters that can measure average memory
 access latency~\cite{ye2016maracas}, we were not able to identify
@@
 hypothesis.
-
-%% The increase in inference times, however, was not the result of
-%% increased cache misses due to additional accesses. For all models, the
-%% number of L1 misses was always $\sim$1.6\% of all  references. In the
-%% 2 model test, L2 cache misses remained at $\sim$11\% of all
-%% references, and each model in the 4 model test missed $\sim$13\% of
-%% all L2 references.
-
-%% In all multimodel tests, it was found that a greater number of models
-%% run simultaneously resulted in interference that led to a noticable
-%% increase in the average inferencing time that wasn't attributable to a
-%% change in cache misses. Instead, the overall time increases can most
-%% likely be traced to an increase in cache latency. Even if the models
-%% had a consistent number of cache hits, it is highly probable that the
-%% contention for the cache increased the access time for each model,
-%% consequently increasing the time it took for the models to execute
-%% their operations.
-
-%% In terms of real-time performance under a WCET of 50
-%% ms, the DeepPicar was unsatisfactory in all multimodel tests. On
-%% average, the 2 model test saw models miss their deadline by 6-8 ms,
-%% and the 4 model test was worse as each model went over their deadlines
-%% by 27-29 ms. As a result, it can be ascertained that, if a WCET of 50
-%% ms is required, DeepPicar would not be able to perform as necessary.
 
@@
 
-%% \begin{table}[h]
-%% \centering
-%%   \begin{tabular} {| l | l | l | l | l | l |}
-%%   \hline
-%%   \textbf{num cores} & 1 & 2 & 3 \\ \hline
-%%   \textbf{L1 refs} & 3.16E+10 & 3.17E+10 & 3.06E+10 \\ \hline
-%%   \textbf{L1 misses} & 5.41E+08 & 5.35E+08 & 4.98E+08 \\ \hline
-%%   \textbf{L1 miss \%} & 1.71 & 1.69 & 1.62 \\ \hline
-%%   \textbf{L2 refs} & 3.49E+09 & 3.23E+09 & 3.21E+09 \\ \hline
-%%   \textbf{L2 misses} & 8.05E+08 & 7.54E+08 & 4.98E+08 \\ \hline
-%%   \textbf{L2 miss \%} & 23.05 & 23.37 & 15.51 \\ 
-%%   \hline
-%%   \end{tabular}
-%%   \caption{ Effect of benchmarks on cache accesses. }
-%% \end{table}
-
 Figure~\ref{fig:} shows the normalized execution time and L2 miss-rate
@@
 the DNN model's exeuction increase cannot be fully explained by
-increased L2 cache-misses. Instead, as we hypothesized in the previous
+increases in L2 cache-misses. Instead, as we hypothesized in the previous
 experiment, the increased memory pressure from the co-scheduled memory
-intensive benchmarks is the primary cause of the DNN model's execution
+intensive benchmarks is likely the primary cause of the DNN model's execution
 time increase. Therefore, we conclude that DeepPicar's DNN model is
@@
 many AI workloads may exhibit similar characteristics.
-
-%% The presence of the benchmark(s) had a noticable effect on the
-%% DeepPicar, with all tests showing increases in inferencing
-%% operations. The least change happened when the model was run on 3
-%% cores, and a single benchmark was run on the final core. Even with a
-%% slight increase in the number of L2 cache misses, inferencing was able
-%% to be completed in under 100 ms. The other tests, however, showed
-%% dramatic changes when the model was run on 1 and 2 cores. In both
-%% instances, L2 misses rose to $\sim$23\% of all references. The
-%% additional benchmark in the 1 core test was even more detrimental as
-%% the average inferencing time was at least 150 ms longer than when the
-%% model used 2 cores.
-
-%% With the introduction of synthetic benchmarks, the DeepPicar failed to
-%% comply with the 50 ms WCET across all tests run. Even in the best
-%% case, inferencing operations take twice as long as the deadline to
-%% complete. As a result, if benchmarks were required to run during
-%% autonomous operation, the DeepPicar wouldn't be capable of meeting its
-%% deadlines.
 
@@
 
-Second, we find that scalability of Tensorflow's DNN inference 
+Second, we find that scalability of Tensorflow's DNN 
 implementation is limited. We find that using all four cores is 
@@
 bandwidth intensive applications on idle cores.
-
-%% is capable of successfully completing all
-%% necessary inferencing operations within a given deadline of 50
-%% ms. Between these operations, it was discovered that the angle
-%% prediction time of the model was the dominating step in the processing
-%% time of a single frame. The other operations were found to execute in
-%% very little time, with either one taking 5 ms at most
-
-%% When running a
-%% single model, DeepPicar performed the best when it used all 4 cpu
-%% cores, as it was able to process a frame in 43 ms on average. This
-%% feat can still be accomplished when running a single model on 3 cpu
-%% cores, and is also possible when using 2 cores. The only time in which
-%% the Pi can't run a single model under 50 ms is when only 1 core is
-%% used. This was also the case when multiple models were run
-%% concurrently as no models were able to consistently complete inference
-%% operations in under 50 ms.
-
-%% However, the DeepPicar was shown to be capable of handling other
-%% potential WCETs. If given a larger deadline, the capabilities of
-%% DeepPicar as an autonomous vehicle platform would be greatly
-%% increased. For example, if tested with a deadline of 66.67 ms, or 15
-%% fps, the platform would have passed several of the experiments
-%% performed. In actuality, the DeepPicar would have only been
-%% unsatisfactory in the 4 model test where it would have missed the
-%% deadlines by $\sim$10 ms, and the benchmark tests, where  
-%% it would still miss by $\sim$33 ms at best.
-
-%% In our platform, three main real-time tasks are performed during autonomous operation. In order, these 
-%% operations are: (1) capturing and reading the input frame from the designated camera or video stream, 
-%% (2) preprocessing the acquired frame so that it is compatible with the DNN, and (3) feeding the frame 
-%% to, and getting the angle prediction from, the model. We aim to determine which operation(s), if any, 
-%% require the most time to be performed.
-
-%% In order to determine which operation(s) take the longest to execute, we measured the time it 
-%% took for each step to be completed. For this experiment, all four of the Pi's cpu cores were utilized, 
-%% and only one model was run. As is shown in Table II, the angle prediction operation consumes the 
-%% majority of the processing for each frame. Furthermore, the time it takes for the operation to 
-%% complete is volatile, and can range anywhere between 30 ms and 50 ms for any particular frame. On the 
-%% other hand, both the frame capture and preprocessing operations take substantially less time 
-%% and are relatively more consistent in their times, at 2 ms and 3 ms, respectively.
-
-%% In evaluating the real-time efficacy of the DeepPicar, the methodology is consistent across all 
-%% experiments. The performance of the platform is measured over a set of 1001 video frames that are 
-%% each individually fed to the model. The processing time for the first frame is omitted as, due to 
-%% cache warmup, it is uncharacteristically high and doesn't
-%% accurately represent the Picar's capabilities.  
-%% A WCET of 50 ms, or 20 Hz, is used as a baseline to assess the platform's ability to complete all 
-%% necessary real-time operations in a timely manner.
-
-%% We seek to determine if the DeepPicar is capable of consistently executing all necessary functions 
-%% before their given deadlines. In the case of our platform, it has to process every given frame and get 
-%% the predicted angles from the model within a WCET of 50 ms. In order to determine if this was achievable, 
-%% we tested the platform by running a single model that utilized all 4 cpu cores and measured the time it 
-%% took for each frame to be processed. The performance of the DeepPicar can be seen in Figure 6. We found 
-%% that DeepPicar was able to completely process the vast majority of the provided frames within 50 ms, and 
-%% was unable to do so for very few frames.
-
-%% In our platform, three main real-time tasks are performed during autonomous operation. In order, these 
-%% operations are: (1) capturing and reading the input frame from the designated camera or video stream, 
-%% (2) preprocessing the acquired frame so that it is compatible with the DNN, and (3) feeding the frame 
-%% to, and getting the angle prediction from, the model. We aim to determine which operation(s), if any, 
-%% require the most time to be performed.
-
-%% In order to determine which operation(s) take the longest to execute, we measured the time it 
-%% took for each step to be completed. For this experiment, all four of the Pi's cpu cores were utilized, 
-%% and only one model was run. As is shown in Table II, the angle prediction operation consumes the 
-%% majority of the processing for each frame. Furthermore, the time it takes for the operation to 
-%% complete is volatile, and can range anywhere between 30 ms and 50 ms for any particular frame. On the 
-%% other hand, both the frame capture and preprocessing operations take substantially less time 
-%% and are relatively more consistent in their times, at 2 ms and 3 ms, respectively.

### 2017-10-13 paper/intro.tex
Changes to paper

@@
 % big problem
-Executing these AI workloads on an embedded computing platform that can
-be used in a car poses several additional challenges. First, many AI
-workloads in vehicles are computationally demanding and have strict
-real-time requirements. For example, latency in a vision-based object
+Executing these AI workloads on an embedded computing platform 
+poses several additional challenges. First, many AI workloads in vehicles 
+are computationally demanding and have strict real-time requirements. 
+For example, latency in a vision-based object
 detection task may be directly linked to safety of the vehicle. This
@@
 requirements  complicate the platform selection process as observed in
-~\cite{Otterness2017}, which evaluated vision workloads on
-NVIDIA's Tegra TX1 platform.
+~\cite{Otterness2017}.
 
@@
 In this paper, we present DeepPicar, a low-cost autonomous car
-platform for research and education. From a hardware perspective,
+platform for research. From a hardware perspective,
 DeepPicar is comprised of a Raspberry Pi 3 Model B quad-core
@@
 
-%% that must be considered in applying deep-learning
-%% models on shared memory based embedded computing platforms.
-%% The number of cores used to process the neural network by the platform has 
-%% an unexpectedly low impact on its performance, only decreasing the 
-%% control loop times by 30\% when using 4 cores instead of 1.  Our 
-%% platform shows the potential to run multiple neural network models 
-%% simultaneously as there are no indications of direct interference 
-%% between them. Instead, we show that memory intensive co-runners can 
-%% be detrimental to the real-time performance of the DeepPicar. 
-%% Finally, while we find that the Raspberry Pi 3 takes longer to 
-%% complete control loops when compared to other embedded computing 
-%% platforms, namely the Intel UP Board and the NVIDIA Tegra TX2, we 
-%% still conclude that the Pi 3 is a viable option for autonomous 
-%% vehicle research.
-
 % contributions
@@
 
-%% The DeepPicar is comprised of a Raspberry
-%% Pi 3 Model B, a camera and a low cost RC car.
-
-%% By creating and using the Picar platform, we seek to accomplish the
-%% following three goals. First, the significant reduction of the overall
-%% cost required
-%% would make autonomous driving research more accessible to interested
-%% individuals/parties. Furthermore, it would remove the concerns of
-%% creating replacement platforms as each part, or the entire system,
-%% could be replaced relatively inexpensively. Second, the utilization of
-%% a more cost-efficient platform would allow for a better understanding
-%% of the performance requirements necessary for efficiently operating
-%% autonomous vehicles. Finally, we strive to achieve the ability to
-%% compare the real-time performance of multiple embedded computing
-%% platforms.
-
-%% We display the efficacy of the Picar by training and testing the
-%% DeepTesla DNN \cite{} in a custom-made environment. We find that the
-%% Picar is capable of consistently accomplishing all real-time
-%% operations required for autonomous driving in a 50 millisecond frame.
-
 The remaining sections of the paper are as follows: Section II 
@@
 conclusions in Section VI.
-
-%% it is necessary that all real-time operations are successfully
-%% performed prior to their deadlines. This requires that the AV platform
-%% be capable of completing all necessary computations in a timely
-%% manner, while maintaining a high level of accuracy. Consequently, the
-%% cost of creating a platform capable of self-driving can be relatively
-%% high, especially in the current cost-sensitive climate of the
-%% automotive industry \cite{}. The overall price for an autonomous
-%% vehicle platform currently acts as a bottleneck, especially from an
-%% academic standpoint, where cost is an important factor. For many
-%% researchers and students, the charge associated with autonomous
-%% vehicle research is a significant barrier. These same parties may also
-%% be deterred from participating in autonomous vehicle research due to
-%% the potential for their developed platforms to break in the
-%% experimental process. The idea of building one relatively expensive
-%% platform may not be too daunting for some, but the notion of
-%% potentially making multiple identical platforms in case of accidents
-%% may be overwhelming.
-%% Toward this end, we explore the possibility of
-%% developing a low-cost system that still employs state-of-the-art AI
-%% technologies and is capable of executing real-time autonomous vehicle
-%% operations.

### 2017-10-13 paper/main.tex
Changes to paper

@@
 replication of a real self-driving car called Dave-2 by NVIDIA, which
-drove public roads using a deep convolutional network (CNN) that
-takes images from a front-facing camera as input and produces
+drove on public roads using a deep convolutional neural network (CNN), 
+that takes images from a front-facing camera as input and produces
 car steering angles as output. DeepPicar uses the exact same network
@@
 models on shared memory based embedded computing platforms.
-
-%% In order to effectively research autonomous vehicles, a real-time
-%% capable platform is required. Such platforms can be hard to obtain, or
-%% build, as the necessary components can be expensive. In this paper,
-%% we present a low-cost alternative platform that utilizes a Raspberry
-%% Pi 3 Model B, along with a single external camera sensor, to perform
-%% all necessary self-driving operations. A convolutional neural network
-%% (CNN) is employed and trained with input from a human driver to
-%% navigate a custom made track. We display the real-time efficacy of our
-%% platform in experiments with varying amounts of computational resources
-%% and/or intensive AI workloads. To gain a better understanding of our
-%% platform's potential, we compare it to other embedded computing
-%% platforms to determine which are suitable for autonomous driving
-%% research.
 \end{abstract}

### 2017-10-13 paper/overview.tex
Changes to paper

@@
 multicore platform, Raspberry Pi 3. Because Pi 3's computing
-performance is much lower than that of the DRIVE PX platform, used in
-DAVE-2, we are interested in if and how we can process
+performance is much lower than that of the DRIVE PX platform used in
+DAVE-2, we are interested in if, and how, we can process
 computationally expensive neural network operations in
 real-time. Specifically, inferencing (forward pass processing)
-operation must be completed within each control period
+operations must be completed within each control period
 duration---e.g., WCET of 50ms for 20Hz control frequency---locally on
@@
 (SBC), a Pololu DRV8835 motor driver, a Playstation Eye webcam, a
-battery, and a 1:24 scale RC car. Table~\ref{tbl:carbom} shows
+battery, and a 1:24 scale RC car. Table~\ref{tbl:carbom} shows the
 estimated cost of the system.
@@
 
-For the neural network architecture, we adopt a tensorflow version of
-NVIDIA's DAVE-2 convolutional neural network (CNN), published by
+For the neural network architecture, we adopt a TensorFlow version of
+NVIDIA DAVE-2's convolutional neural network (CNN), published by Dr.
 Fridman at  MIT~\footnote{https://github.com/lexfridman/deeptesla}. As
-in NVIDIA's DAVE-2, the CNN takes raw color image (200x66 RGB pixels)
-as input and produce a single steering angle value as
+in DAVE-2, the CNN takes a raw color image (200x66 RGB pixels)
+as input and produces a single steering angle value as
 output. Figure~\ref{fig:architecture} shows the network architecture, which
@@
 on a small track we created (Figure~\ref{fig:track}) to record
-timestamped video and contol commands. The stored data is then copied 
-to a desktop computer, which equips a NVIDIA GTX 1060 GPU, where we
-train the network to accerate training speed. 
+timestamped videos and contol commands. The stored data is then copied 
+to a desktop computer, which is equipped with a NVIDIA GTX 1060 GPU, 
+where we train the network to accelerate training speed. 
 For comparison, training the network on the Raspbeery Pi 3 takes
@@
 Once the network is trained on the desktop computer, the trained model
-is copied back to the Raspberry Pi computer. The network is then used
+is copied back to the Raspberry Pi. The network is then used
 by the car's main controller, which feeds a image frame from the web
 camera as input to the network. The produced steering angle output is
-then converted as the PWM value of the steering motor of the
-car. Figure~\ref{fig:controlloop} shows simplified psuedo code of the
-controller's main loop. Among the five steps, step 3---network
-inferencing---is most computationally intensive and dominates the
+then converted as the PWM values of the steering motor of the
+car. Figure~\ref{fig:controlloop} shows simplified pseudo code of the
+controller's main loop. Among the five steps, step 3, network
+inferencing, is the most computationally intensive and dominates the
 execution time.
@@
 introduce inaccuracy in control.
-In the future, we plan to use different (more expensive) RC car
+In the future, we plan to use a different (more expensive) RC car
 platforms that can precisely control the car's steering angle.
@@
 affect the network performance, is camera latency. In DeepPicar's
-context, the camera latency is the time from the time the camera
+context, the camera latency is from the time the camera
 sensor observes the scene to the time the computer actually reads the
 digitized image data. Unfortunately, this time can be significantly
-long depending on camera and the computer performance. We
+long depending on camera and the performance of the Pi. We
 experimentally measure DeepPicar's camera latency as
@@
 conntected to the Raspberry Pi 3's HDMI output. We then launch a 
-simple program that overlays the current time with the camera's preview
+program that overlays the current time with the camera's preview
 screen on the monitor. Then, an infinite number of time strings will be
-shown on the monitor screen and the time difference between the two
+shown on the monitor screen and the time difference between two
 successive time strings represent the camera
 latency. Figure~\ref{fig:camera-latency} shows the captured moniotor
-screen. The measured camera latency is about 300~350 ms. This is
+screen. The measured camera latency is about 300-350 ms. This is
 significantly higher than the latency of human perception, which is known
 to be as fast as 13 ms~\cite{ThomasBurger2015}. Higher camera latency
-could negatively affect to the control performance as the DNN would
+could negatively affect to the control performance, because the DNN would
 analyze stale scenes. In the future, we plan to identify and use
@@
 can be found at: \url{https://github.com/heechul/picar}.
-%% \begin{center}
-%% \end{center}
-
-%% We employ a small and relatively inexpensive RC car that is capable of
-%% performing basic automotive operations. However, the car we use
-%% doesn't replicate the capabilities of other autonomous vehicle
-%% platforms, as it is more simplistic in nature. Notably, the RC car we
-%% use only has three possible options in terms of turning: center, left,
-%% and right. As such, control of the car may be less precise at times,
-%% and may negatively affect performance. The car is capable of other
-%% operations that aren't used within the scope of this
-%% platform. Specifically, the car is able to drive in reverse and the
-%% driving speed can be changed. In our experiments, we have the RC car
-%% drive forward, and at a constant speed.
-
-%% Our platform also comes equipped with a camera that is used for both
-%% recording training videos and providing input frames to the model
-%% whenever the Picar is self-driving. We chose the Playstation Eye
-%% Camera due to its ability to reach and maintain higher fps levels
-%% while also remaining relatively low-cost. One concern, however, is the
-%% affect of camera latency on the self-driving performance of our
-%% platform. While humans are able to see environmental changes in
-%% real-time, the same can't be said for cameras since there is a delay
-%% between when frame capture by the camera, and when it is read by the
-%% Raspberry Pi 3. As a result, it is possible for the model to be given
-%% an input frame that is different from the real world, thus impacting
-%% the performance of the Picar.
-
-%% Compared to other existing autonomous vehicle platforms, such as the
-%% F1/10 and NVIDIA DAVE-2, the Picar platform is capable of performing
-%% the same operations while using more cost inexpensive components. As
-%% shown by Table 1, the components selected for the Picar all cost
-%% considerably less than the other  platforms when focusing on the
-%% common components (camera, power supply, etc.). Including the other
-%% components used in the F1/10 and DAVE-2, such as the additional
-%% sensors employed by both, the difference  in cost would be even
-%% greater.
-
-%% \subsection{Model Training}
-%% In order to operate the Picar as an autonomous vehicle, we utilize the
-%% DeepTesla library, which is capable of training a deep neural network
-%% (DNN) with end-to-end learning, that could then be used by the  Picar
-%% for autonomous driving. Please note that it is more efficient to
-%% execute the actual model training  process on a different system with
-%% an NVIDIA GPU, rather than on the Raspberry Pi 3 itself. This is
-%% because DeepTesla trains the DNN with the Tensorflow library, which
-%% greatly benefits from the use of  gpu-enabled operations. For
-%% comparison, training a model on the Pi took approximately 6 hours,
-%% whereas  the time it took on a computer with a NVIDIA GPU was only
-%% around 5 minutes!
-
-%% We train our model on a custom made track/lane composed of black and
-%% white duct tape, where the black  tape represents the bounds of the
-%% track, and the white tape marks the center of the lane. The model is
-%% taught to stay close to the center of the lane for as long as
-%% possible, and to turn whenever it  reaches/crosses the outer lane
-%% bounds so that it remains on the track. For data, we navigate and
-%% record  the car going both ways across the track, and use the
-%% collected videos to train the model(s) that will  be used later for
-%% angle prediction while the car is self-driving.
-
-%% \begin{table*}[h]
-%%   \centering
-%%   \begin{tabular} {| l | l | l | l | l |}
-%%     \hline
-%%     \textbf{Platform} & \textbf{Picar} & \textbf{F1/10} & \textbf{NVIDIA DAVE-2}\\ \hline  
-%%     Car & Mini RC Car (\$10) & Traxxas 1/10th car platform (\$299.97) & TBD\\ \hline
-%%     Embedded system & Raspberry Pi 3 Model B (\$35.00) & NVIDIA Jetson TK1 (\$192.99) & NVIDIA DRIVE PX \\ \hline
-%%     CPU & Cortex A-53 quad core & Cortex A-15 quad core & 8x Cortex A-57 quad core, 8x Cortex A-53 quad core \\ \hline
-%%     Camera & Playstation Eye camera (\$6.99) & ZED camera (\$449.00) & 3x cameras\\ \hline
-%%     Power Supply & Mobile battery (\$5.99) & Energizer battery pack (\$159.00) & TBD\\
-%%     \hline
-%%   \end{tabular}
-%%   \caption{Comparison of components used in autonomous vehicle platforms.}
-%% \end{table*}

### 2017-10-14 PiBits/ServoBlaster/user/test.sh
cleanup

@@ -1,7 +1,0 @@
-for duty in `seq 100`; do
-	echo "duty=${duty}%"
-	echo 0=${duty}% > /dev/servoblaster
-	echo "press key"
-	read xx
-done
-

### 2017-10-14 mjpg-streamer/mjpg-streamer-experimental/plugins/input_uvc/v4l2uvc.c

@@ -1,1018 +1,0 @@
-/*******************************************************************************
-# Linux-UVC streaming input-plugin for MJPG-streamer                           #
-#                                                                              #
-# This package work with the Logitech UVC based webcams with the mjpeg feature #
-#                                                                              #
-# Copyright (C) 2005 2006 Laurent Pinchart &&  Michel Xhaard                   #
-#                    2007 Lucas van Staden                                     #
-#                    2007 Tom Stöveken                                         #
-#                                                                              #
-# This program is free software; you can redistribute it and/or modify         #
-# it under the terms of the GNU General Public License as published by         #
-# the Free Software Foundation; either version 2 of the License, or            #
-# (at your option) any later version.                                          #
-#                                                                              #
-# This program is distributed in the hope that it will be useful,              #
-# but WITHOUT ANY WARRANTY; without even the implied warranty of               #
-# MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the                #
-# GNU General Public License for more details.                                 #
-#                                                                              #
-# You should have received a copy of the GNU General Public License            #
-# along with this program; if not, write to the Free Software                  #
-# Foundation, Inc., 59 Temple Place, Suite 330, Boston, MA  02111-1307  USA    #
-#                                                                              #
-*******************************************************************************/
-
-#include <stdlib.h>
-#include <errno.h>
-#include "v4l2uvc.h"
-#include "huffman.h"
-#include "dynctrl.h"
-
-static int debug = 0;
-
[…]

### 2017-10-14 picar-med-kbd-drv8835.py

@@ -1,160 +1,0 @@
-#!/usr/bin/python
-import os
-import time
-import atexit
-import serial
-import cv2
-import math
-import numpy as np
-import pygame
-from pololu_drv8835_rpi import motors, MAX_SPEED
-
-motors.setSpeeds(0, 0)
-cap = cv2.VideoCapture(0)
-# cap.set(3,320)
-# cap.set(4,240)
-
-# ser = serial.Serial('/dev/ttyACM0', 115200, timeout=1)
-# ser = serial.Serial('/dev/ttyUSB0', 115200, timeout=1)
-fourcc = cv2.VideoWriter_fourcc(*'XVID')
-vidfile = cv2.VideoWriter('out-video.avi', fourcc, 15.0, (640,480))
-keyfile = open('out-key.csv', 'w+')
-keyfile_btn = open('out-key-btn.csv', 'w+')
-keyfile.write("ts_micro,frame,wheel\n")
-keyfile_btn.write("ts_micro,frame,btn,speed\n")
-rec_start_time = 0
-cur_speed = MAX_SPEED   # 0...MAX_SPEED
-cur_steer = MAX_SPEED
-
-def stop():
-        motors.motor2.setSpeed(0)
-        motors.motor1.setSpeed(0)        
-
-# 
-def ffw():
-        global cur_speed
-        motors.motor2.setSpeed(cur_speed)
-def rew():
-        global cur_speed        
-        motors.motor2.setSpeed(-cur_speed)
-# W - throttle speed up         
-def throttle_up():
-        global cur_speed        
-        cur_speed = min(MAX_SPEED, cur_speed + MAX_SPEED/10)
-        print "cur_speed:", cur_speed
-# X - throttle speed down 
-def throttle_down():
-        global cur_speed        
-        cur_speed = max(0, cur_speed - MAX_SPEED/10)
-        print "cur_speed:", cur_speed        
-        
-def left():
-        global cur_steer
-        motors.motor1.setSpeed(cur_steer)
-def right():
-        global cur_steer
-        motors.motor1.setSpeed(-cur_steer)
-        
-# E - throttle speed up         
-def steer_up():
-        global cur_steer
-        cur_steer = min(MAX_SPEED, cur_steer + MAX_SPEED/10)
-        print "cur_steer:", cur_steer        
-# C - throttle speed down 
-def steer_down():
-        global cur_steer        
-        cur_steer = max(0, cur_steer - MAX_SPEED/10)
[…]

### 2017-10-14 src/teleop_twist_keyboard.py
cleanup

@@ -1,132 +1,0 @@
-#!/usr/bin/env python
-import roslib; roslib.load_manifest('teleop_twist_keyboard')
-import rospy
-
-from geometry_msgs.msg import Twist
-
-import sys, select, termios, tty
-
-msg = """
-Reading from the keyboard  and Publishing to Twist!
----------------------------
-Moving around:
-   u    i    o
-   j    k    l
-   m    ,    .
-
-For Holonomic mode (strafing), hold down the shift key:
----------------------------
-   U    I    O
-   J    K    L
-   M    <    >
-
-t : up (+z)
-b : down (-z)
-
-anything else : stop
-
-q/z : increase/decrease max speeds by 10%
-w/x : increase/decrease only linear speed by 10%
-e/c : increase/decrease only angular speed by 10%
-
-CTRL-C to quit
-"""
-
-moveBindings = {
-		'i':(1,0,0,0),
-		'o':(1,0,0,-1),
-		'j':(0,0,0,1),
-		'l':(0,0,0,-1),
-		'u':(1,0,0,1),
-		',':(-1,0,0,0),
-		'.':(-1,0,0,1),
-		'm':(-1,0,0,-1),
-		'O':(1,-1,0,0),
-		'I':(1,0,0,0),
-		'J':(0,1,0,0),
-		'L':(0,-1,0,0),
-		'U':(1,1,0,0),
-		'<':(-1,0,0,0),
-		'>':(-1,-1,0,0),
-		'M':(-1,1,0,0),
-		't':(0,0,1,0),
-		'b':(0,0,-1,0),
-	       }
-
-speedBindings={
-		'q':(1.1,1.1),
-		'z':(.9,.9),
-		'w':(1.1,1),
-		'x':(.9,1),
-		'e':(1,1.1),
-		'c':(1,.9),
-	      }
-
-def getKey():
-	tty.setraw(sys.stdin.fileno())
-	select.select([sys.stdin], [], [], 0)
-	key = sys.stdin.read(1)
-	termios.tcsetattr(sys.stdin, termios.TCSADRAIN, settings)
-	return key
-
-speed = .5
-turn = 1
-
-def vels(speed,turn):
-	return "currently:\tspeed %s\tturn %s " % (speed,turn)
-
-if __name__=="__main__":
-    	settings = termios.tcgetattr(sys.stdin)
-	
-	pub = rospy.Publisher('cmd_vel', Twist, queue_size = 1)
[…]

### 2017-10-14 src/tx2_driver.py

@@ -1,120 +1,0 @@
-#!/usr/bin/env python
-import rospy
-import math
-import serial
-
-import os
-import time
-import atexit
-
-from geometry_msgs.msg import Twist 
-from sensor_msgs.msg import Imu
-from tf.transformations import euler_from_quaternion
-
-rad2degrees = 180.0/math.pi
-degree2rad = math.pi / 180.0
-
-precision = 2 #round to this number of digits
-
-yaw_offset = 0 #used to align animation upon key press
-yaw_measured = 0
-
-ser = serial.Serial('/dev/ttyACM0', 115200)
-
-def turnOff():
-	os.system("killall -HUP mencoder")
-	
-atexit.register(turnOff)
-
-def right(rad):
-    ser.write(b'r\n')
-    print "right"
-
-def left(rad):
-    ser.write(b'l\n')
-    print "left"
-
-def center():
-    ser.write(b'c\n')
-    print "center"
-
-def ffw(vel):
-    ser.write(b'f\n')        
-    print "forward"
-
-def rew(vel):
-    ser.write(b'b\n')        
-    print "backward"
-
-def stop():
-    ser.write(b's\n')
-    print "stop"
-    
-def callback(msg):
-    rospy.loginfo("Received a /cmd_vel message!")
-    rospy.loginfo("Linear Components: [%f, %f, %f]" %
-                  (msg.linear.x, msg.linear.y, msg.linear.z))
-    rospy.loginfo("Angular Components: [%f, %f, %f]" %
-                  (msg.angular.x, msg.angular.y, msg.angular.z))
-    # Do velocity processing here:
-    # Use the kinematics of your robot to map linear and angular velocities
-    # into motor commands
-
-    # speed
-    if msg.linear.x < 0.01 and msg.linear.x > -0.01:
-        stop()
-    elif msg.linear.x > 0.01:
-        ffw(msg.linear.x)
-    else:
-        rew(msg.linear.x)
-    
-    # turn 
-    if msg.angular.z < -0.01:
-        right(msg.angular.z)
-    elif msg.angular.z > 0.01:
-        left(msg.angular.z)
-    else:
-        center()
-
-def processIMU_message(imuMsg):
-    global yaw_offset, yaw_measured
-
-    roll=0
-    pitch=0
-    yaw=0
-
-    quaternion = (
-      imuMsg.orientation.x,
-      imuMsg.orientation.y,
-      imuMsg.orientation.z,
[…]

### 2017-10-16 paper/appendix.tex
Changes to paper

@@
 frames for validation.
-The weights of the network are initialized using Xavier
+The weights of the network are initialized using a Xavier
 initializer~\cite{Glorot2010}, which is known to provide better
@@
 was 0.0869---a 10X difference---indicating that the model suffers from
-the overfitting problem.
+an overfitting problem.
 
@@
 car platform, especially related to precise steering angle control.
-
-%% We believe, though, that this can be
-%% rectified through the collection of additional data for training and
-%% validation.
-
-%% \begin{figure*}[h]
-%%   \centering
-%%   \includegraphics[width=.9\textwidth]{figs/graph_run=.pdf}
-%%   \caption{ Network visualization. }
-%%   \label{fig:}
-%% \end{figure*}
-%% \fixme{Is all of the network graph necessary or only certain parts?}
 
@@
 
-%% Specifically, these issues all have the potential to negatively
-%% affect the cpu clock speed/frequency, which would result in decreased
-%% performance. 
 In all our experiments on the Raspberry Pi 3, the CPU  operated at a
 preferred clock speed of 1.2 GHz. However, without care, it is
-possible for the CPU may be operating at a lower frequency.
+possible for the CPU to operate at a lower frequency.
 
@@
 If the temperature reaches the threshold, the Pi 3's thermal
-throttling kicks in and decreases the clock speed down to 600MHz---a
+throttling kicks in and decreases the clock speed down to 600MHz---
 half of the maximum 1.2GHz---so that the CPU's temperature stays at a
@@
 repeatability of the results.
-
-%% Thus the
-%% Raspberry Pi may not be suited for prolonged use, especially in cases
-%% where the workload is relatively larger, such as running multiple
-%% models. Rather, the Pi seems to be better suited for running in set
-%% periods, after which it is turned  off or made idle so that the cpu is
-%% allowed time to cool down.

### 2017-10-16 paper/background.tex
Changes to paper

@@
 produce steering control commands, bypassing all intermediary steps and
-hand-written rules used in conventional robotics control approach.  
+hand-written rules used in the conventional robotics control approach.  
 NVIDIA's latest effort reports that their trained CNN
@@
 as input to produce predicted output (e.g., predicted image
-classification). In general, the training phase is most computationally
+classification). In general, the training phase is more computationally
 intensive and requires high throughput, which is generally not

### 2017-10-16 paper/evaluation.tex
Changes to paper

@@
 
-For real-time control of a car (or any robots), the control loop
+For real-time control of a car (or any robot), the control loop
 frequency must be sufficiently high so that the car can quickly
@@
 control performance improves when the frequency is higher, though
-computation time and the type of particular physical system are
+computation time and the type of the particular physical system are
 factors in determining a proper control loop frequency. While a standard
 control system may be comprised of multiple control loops with
-differing control frequencies---e.g., a inner control loop for lower-level
-PD control, a outer loop for motion planning, etc.---DeepPicar's
+differing control frequencies---e.g., an inner control loop for lower-level
+PD control, an outer loop for motion planning, etc.---DeepPicar's
 control loop is a single layer, as shown earlier in
 Figure~\ref{fig:controlloop}, since a single deep neural network
-replaces the traditional multi-layer control pipline. (Recall
+replaces the traditional multi-layer control pipline. (Refer to
 Figure~\ref{fig:end-to-end-control} on the differences between the
 standard robotics control vs. end-to-end deep learning approach).
-This means that the DNN interference operation must be completed
+This means that the DNN inference operation must be completed
 within the inner-most control loop update frequency. To understand
@@
 processing times of the DeepPicar over 1000 image frames (one per each
-control loop). We omit the first frame processing time for cache
+control loop). We omit the first frame's processing time for cache
 warmup. Table~\ref{tbl:control-loop-breakdown} shows the time
@@
 execution time of a single control loop is 42.67 ms, or 23.4 Hz and
-the 99 percentail time is 50.70 ms.
+the 99 percentile time is 50.70 ms.
 This means that the DeepPicar can operate
-at about 20 Hz control frequency in real-time using only the on-board
-Raspberry Pi 3 computing platform, without needing any remote computing
-resources. We consider these results respectable given the complexity
+at about a 20 Hz control frequency in real-time using only the on-board
+Raspberry Pi 3 computing platform, as no remote computing resources 
+were necessary. We consider these results respectable given the complexity
 of the deep neural network, and the fact that the inference operation
 performed by TensorFlow only utilizes the CPU cores of the
-Raspberry Pi 3 (its GPU is not supported by Tensorflow.)
+Raspberry Pi 3 (its GPU is not supported by Tensorflow).
 
 In comparison, NVIDIA's DAVE-2 system, which has the exact same neural
-network architecture, reportedly run at 30 Hz~\cite{Bojarski2016},
+network architecture, reportedly runs at 30 Hz~\cite{Bojarski2016},
 just a bit faster than the DeepPicar. Although we believe it was not

### 2017-10-16 paper/intro.tex
Changes to paper

@@
 testbeds. There are already several relatively inexpensive RC-car
-based testbeds, such as MIT 
+based testbeds, such as MIT's 
 RaceCar~\cite{shin2017project} and UPenn's F1$/$10 racecar~\cite{upennf1tenth}.
@@
 self-driving car~\cite{Bojarski2016}, which uses a much more powerful
-computer (Drive PX computer~\cite{drivepx}) than Raspberry Pi~3.
-We chose to use Pi 3 not only because of its
+computer (Drive PX computer~\cite{drivepx}) than a Raspberry Pi~3.
+We chose to use a Pi 3 not only because of its
 affordability, but also because it is representative
@@
 real-time. For real-time control, each inference operation must
-be completed within the desired control period (e.g., 50ms period for
-20Hz control frequency).
+be completed within the desired control period (e.g., 50 ms period for
+20 Hz control frequency).
 % how we evaluated (in terms real-time performance)
@@
 % what are our findings?
-Using the DeepPicar platform, we systematically analyzed its real-time
+Using the DeepPicar platform, we systematically analyze its real-time
 capabilities in the context of deep-learning based real-time
@@
   \item We present the design and implementation of a
-    low-cost autonomous vehicle testbed DeepPicar, which utilizes
+    low-cost autonomous vehicle testbed, DeepPicar, which utilizes
     state-of-the-art artificial intelligence techniques.

### 2017-10-16 paper/main.tex
Changes to paper

@@
 car steering angles as output. DeepPicar uses the exact same network
-architecture---9 layers, 27 million connections and 24K
+architecture---9 layers, 27 million connections and 250K
 parameters---and can be trained to drive itself, in real-time, using a
 web camera and a modest Raspberry Pi 3 quad-core platform.
-Using DeepPicar, we analyze Pi 3's computing capabilities to 
+Using DeepPicar, we analyze the Pi 3's computing capabilities to 
 support end-to-end deep learning based real-time control of autonomous

### 2017-10-16 paper/overview.tex
Changes to paper

@@
 NVIDIA's DAVE-2 system on a smaller scale with using a low cost
-multicore platform, Raspberry Pi 3. Because Pi 3's computing
+multicore platform, Raspberry Pi 3. Because the Pi 3's computing
 performance is much lower than that of the DRIVE PX platform used in
@@
 operations must be completed within each control period
-duration---e.g., WCET of 50ms for 20Hz control frequency---locally on
+duration---e.g., a WCET of 50ms for 20Hz control frequency---locally on
 the Pi 3 platform, although training of the network (back-propagation
@@
 by the car's main controller, which feeds a image frame from the web
-camera as input to the network. The produced steering angle output is
-then converted as the PWM values of the steering motor of the
-car. Figure~\ref{fig:controlloop} shows simplified pseudo code of the
-controller's main loop. Among the five steps, step 3, network
-inferencing, is the most computationally intensive and dominates the
+camera as input to the network. In each control period, the produced 
+steering angle output is then converted as the PWM values of the 
+steering motor of the car. Figure~\ref{fig:controlloop} shows simplified 
+pseudo code of the controller's main loop. Among the five steps, the 3rd step, 
+network inferencing, is the most computationally intensive and dominates the
 execution time.
@@
 In the future, we plan to use a different (more expensive) RC car
-platforms that can precisely control the car's steering angle.
+platform that can precisely control the car's steering angle.
 We would like to stress, however, that the use of different RC car
-platforms has no impact in terms of computational 
+platforms has no impact on the computational 
 aspects of the system, and that our main focus of this study is
@@
 digitized image data. Unfortunately, this time can be significantly
-long depending on camera and the performance of the Pi. We
+long depending on the camera and the performance of the Pi. We
 experimentally measure DeepPicar's camera latency as
@@
   https://www.makehardware.com/2016/03/29/finding-a-low-latency-webcam/}.
-First, we point the DeepPicar's camera toward to a computer monitor, which is
+First, we point the DeepPicar's camera towards a computer monitor, which is
 conntected to the Raspberry Pi 3's HDMI output. We then launch a 

### 2017-10-16 paper/conclusion.tex
Changes to paper
Minor changes to paper

@@
 Despite the complexity of the neural network, DeepPicar uses a
-low-cost Raspberry Pi 3 quad-core computer as a sole computing
+low-cost Raspberry Pi 3 quad-core computer as its sole computing
 resource. We systematically analyzed the real-time characteristics of
@@
 computing hardware architectures.
-We find all tested embedded platforms, including the Pi3, are capable
+We find all tested embedded platforms, including the Pi 3, are capable
 of supporting deep-learning based real-time control, from 20 Hz up to

### 2017-11-26 paper/background.tex
typo

@@
 inputs~\cite{Levine2016}. Figure~\ref{fig:end-to-end-control}
-shows the differences between two approaches.
+shows the differences between the two approaches.
 
@@
 system-on-a-chips (SOCs) that may include multiple cores, GPU,
-DSP, FPGA and neural network optimized ASIC hardware~\cite{Jouppi2017}.
+DSP, FPGA, and neural network optimized ASIC hardware~\cite{Jouppi2017}.
 Consolidating multiple tasks on SoCs with lots of shared hardware
@@
 research area, which is orthogonal to improving raw
-performance. Consolidation is necessary for efficiency but unmanaged 
+performance. Consolidation is necessary for efficiency, but unmanaged 
 interference can nullify the benefits of consolidation~\cite{Kim2016}.
 For these reasons, finding a good computing platform is a
-non-trivial task, one that requires deep understanding of the
+non-trivial task, one that requires a deep understanding of the
 workloads and the hardware platform being utilized.

### 2017-11-26 paper/intro.tex
typo

@@
 computer (Drive PX computer~\cite{drivepx}) than a Raspberry Pi~3.
-We chose to use a Pi 3 not only because of its
-affordability, but also because it is representative
+We chose to use a Pi 3 not only because it is affordable, but also because it is representative
 of today's mainstream low-end embedded multicore platforms found in

### 2017-11-29 benchmark_timings_x2.sh
updates TX2 script for perf

@@ -13,14 +13,14 @@ echo "The output files can be found in: logs/test-model/$identification"
 
 echo 2-parallel-single-core
 ./bandwidth -a write -m 16384 -t 10000 -c 3 &
-taskset -c 0 python test-model.py 1 > logs/test-model-and-benchmark/$identification/1-sc-bench_1-sc-model_cpu0.txt
+perf stat -o logs/test-model-and-benchmark/$identification/1-sc-bench_1-sc-model_cpu0_perf.txt -e cache-misses,cache-references,LLC-loads,LLC-load-misses,LLC-stores,LLC-store-misses taskset -c 0 python test-model.py 1 > logs/test-model-and-benchmark/$identification/1-sc-bench_1-sc-model_cpu0.txt
 jobs -p | xargs kill
 echo end 2-parallel-single-core
 
 echo 3-parallel-single-core
 ./bandwidth -a write -m 16384 -t 10000 -c 3 &
 ./bandwidth -a write -m 16384 -t 10000 -c 4 &
-taskset -c 0 python test-model.py 1 > logs/test-model-and-benchmark/$identification/2-sc-bench_1-sc-model_cpu0.txt
+perf stat -o logs/test-model-and-benchmark/$identification/2-sc-bench_1-sc-model_cpu0_perf.txt -e cache-misses,cache-references,LLC-loads,LLC-load-misses,LLC-stores,LLC-store-misses taskset -c 0 python test-model.py 1 > logs/test-model-and-benchmark/$identification/2-sc-bench_1-sc-model_cpu0.txt
 jobs -p | xargs kill
 echo end 3-parallel-single-core
 
@@ -28,25 +28,25 @@ echo 4-parallel-single-core
 ./bandwidth -a write -m 16384 -t 10000 -c 3 &
 ./bandwidth -a write -m 16384 -t 10000 -c 4 &
 ./bandwidth -a write -m 16384 -t 10000 -c 5 &
-taskset -c 0 python test-model.py 1 > logs/test-model-and-benchmark/$identification/3-sc-bench_1-sc-model_cpu0.txt
[…]

### 2017-11-29 benchmark_timings_x2_cpu.sh

@@ -13,14 +13,14 @@ echo "The output files can be found in: logs/test-model/$identification"
 
 echo 2-parallel-single-core
 ./bandwidth -a write -m 16384 -t 10000 -c 3 &
-CUDA_VISIBLE_DEVICES='' taskset -c 0 python test-model.py 1 > logs/test-model-and-benchmark/$identification/1-sc-bench_1-sc-model_cpu0.txt
+CUDA_VISIBLE_DEVICES='' perf stat -o logs/test-model-and-benchmark/$identification/1-sc-bench_1-sc-model_cpu0_perf.txt -e cache-misses,cache-references,LLC-loads,LLC-load-misses,LLC-stores,LLC-store-misses taskset -c 0 python test-model.py 1 > logs/test-model-and-benchmark/$identification/1-sc-bench_1-sc-model_cpu0.txt
 jobs -p | xargs kill
 echo end 2-parallel-single-core
 
 echo 3-parallel-single-core
 ./bandwidth -a write -m 16384 -t 10000 -c 3 &
 ./bandwidth -a write -m 16384 -t 10000 -c 4 &
-CUDA_VISIBLE_DEVICES='' taskset -c 0 python test-model.py 1 > logs/test-model-and-benchmark/$identification/2-sc-bench_1-sc-model_cpu0.txt
+CUDA_VISIBLE_DEVICES='' perf stat -o logs/test-model-and-benchmark/$identification/2-sc-bench_1-sc-model_cpu0_perf.txt -e cache-misses,cache-references,LLC-loads,LLC-load-misses,LLC-stores,LLC-store-misses taskset -c 0 python test-model.py 1 > logs/test-model-and-benchmark/$identification/2-sc-bench_1-sc-model_cpu0.txt
 jobs -p | xargs kill
 echo end 3-parallel-single-core
 
@@ -28,25 +28,25 @@ echo 4-parallel-single-core
 ./bandwidth -a write -m 16384 -t 10000 -c 3 &
 ./bandwidth -a write -m 16384 -t 10000 -c 4 &
 ./bandwidth -a write -m 16384 -t 10000 -c 5 &
-CUDA_VISIBLE_DEVICES='' taskset -c 0 python test-model.py 1 > logs/test-model-and-benchmark/$identification/3-sc-bench_1-sc-model_cpu0.txt
[…]

### 2017-11-29 paper/background.tex

--- 2017-10-13 paper/background.tex
@@
 autonomously controls their modified cars on public roads without human
-intervention~\cite{Bojarski2016}.
+intervention successfully~\cite{Bojarski2016}.
 
@@
 phases~\cite{NVIDIA2015}. The first
-phase is \emph{training} during which the weights of the network are
+phase is \emph{training} during which the weights of the network is
 incrementally updated by backpropagating errors it sees from the
@@
 
-The primary objectives of this study are to understand (1) the
+The primary objectives of this study are (1) to understand the
 necessary computing performance for applying AI technology-based

### 2017-11-29 paper/conclusion.tex

--- 2017-10-07 paper/conclusion.tex
@@
 techniques~\cite{Yun2013,yun2014rtas} on the DeepPicar platform and
-evaluate their impact to the real-time performance of the system. We
+evaluate their impact on the real-time performance of the system. We
 also plan to improve the prediction accuracy by feeding more data and

### 2017-11-29 paper/overview.tex
Minor changes to paper

--- 2017-10-07 paper/overview.tex
@@
 Fridman at  MIT~\footnote{https://github.com/lexfridman/deeptesla}. As
-in NVIDIA's DAVE-2, the CNN takes raw color image (200x66 RGB pixels)
-as input and produce a single steering angle value as
+in NVIDIA's DAVE-2, the CNN takes raw color images (200x66 RGB pixels)
+as input and produces a single steering angle value as
 output. Figure~\ref{fig:architecture} shows the network architecture, which
@@
 affect the network performance, is camera latency. In DeepPicar's
-context, the camera latency is the time from the time the camera
+context, the camera latency is the measurement from the time the camera
 sensor observes the scene to the time the computer actually reads the
@@
 screen on the monitor. Then, an infinite number of time strings will be
-shown on the monitor screen and the time difference between the two
+shown on the monitor screen and the time difference between two
 successive time strings represent the camera
 latency. Figure~\ref{fig:camera-latency} shows the captured moniotor
-screen. The measured camera latency is about 300~350 ms. This is
+screen. The measured camera latency is about 300-350 ms. This is
 significantly higher than the latency of human perception, which is known
 to be as fast as 13 ms~\cite{ThomasBurger2015}. Higher camera latency
-could negatively affect to the control performance as the DNN would
+could negatively affect to the control performance, as the DNN would
 analyze stale scenes. In the future, we plan to identify and use

### 2017-11-29 test-model_timings_x2.sh
perf tx2 test-models

@@ -12,49 +12,49 @@ echo "This test will take approximately 45 minutes to complete"
 echo "The output files can be found in: logs/test-model/$identification"
 
 echo single-core
-taskset -c 0 python test-model.py 1 > logs/test-model/$identification/single-core.txt &
+perf stat -o logs/test-model/$identification/single-core_perf.txt -e cache-misses,cache-references,LLC-loads,LLC-load-misses,LLC-stores,LLC-store-misses taskset -c 0 python test-model.py 1 > logs/test-model/$identification/single-core.txt &
 wait
 echo end single-core
 
 echo 2-parallel-single-core
-taskset -c 0 python test-model.py 1 > logs/test-model/$identification/2-single-core_cpu0.txt &
-taskset -c 3 python test-model.py 1 > logs/test-model/$identification/2-single-core_cpu3.txt &
+perf stat -o logs/test-model/$identification/2-single-core_cpu0_perf.txt -e cache-misses,cache-references,LLC-loads,LLC-load-misses,LLC-stores,LLC-store-misses taskset -c 0 python test-model.py 1 > logs/test-model/$identification/2-single-core_cpu0.txt &
+perf stat -o logs/test-model/$identification/2-single-core_cpu3_perf.txt -e cache-misses,cache-references,LLC-loads,LLC-load-misses,LLC-stores,LLC-store-misses taskset -c 3 python test-model.py 1 > logs/test-model/$identification/2-single-core_cpu3.txt &
 wait
 echo end 2-parallel-single-core
 
 echo 3-parallel-single-core
-taskset -c 0 python test-model.py 1 > logs/test-model/$identification/3-single-core_cpu0.txt &
-taskset -c 3 python test-model.py 1 > logs/test-model/$identification/3-single-core_cpu3.txt &
-taskset -c 4 python test-model.py 1 > logs/test-model/$identification/3-single-core_cpu4.txt &
+perf stat -o logs/test-model/$identification/3-single-core_cpu0_perf.txt -e cache-misses,cache-references,LLC-loads,LLC-load-misses,LLC-stores,LLC-store-misses taskset -c 0 python test-model.py 1 > logs/test-model/$identification/3-single-core_cpu0.txt &
[…]

### 2017-11-29 test-model_timings_x2_cpu.sh

@@ -12,49 +12,49 @@ echo "This test will take approximately 45 minutes to complete"
 echo "The output files can be found in: logs/test-model/$identification"
 
 echo single-core
-CUDA_VISIBLE_DEVICES='' taskset -c 0 python test-model.py 1 > logs/test-model/$identification/single-core.txt &
+CUDA_VISIBLE_DEVICES='' perf stat -o logs/test-model/$identification/single-core_perf.txt -e cache-misses,cache-references,LLC-loads,LLC-load-misses,LLC-stores,LLC-store-misses taskset -c 0 python test-model.py 1 > logs/test-model/$identification/single-core.txt &
 wait
 echo end single-core
 
 echo 2-parallel-single-core
-CUDA_VISIBLE_DEVICES='' taskset -c 0 python test-model.py 1 > logs/test-model/$identification/2-single-core_cpu0.txt &
-CUDA_VISIBLE_DEVICES='' taskset -c 3 python test-model.py 1 > logs/test-model/$identification/2-single-core_cpu3.txt &
+CUDA_VISIBLE_DEVICES='' perf stat -o logs/test-model/$identification/2-single-core_cpu0_perf.txt -e cache-misses,cache-references,LLC-loads,LLC-load-misses,LLC-stores,LLC-store-misses taskset -c 0 python test-model.py 1 > logs/test-model/$identification/2-single-core_cpu0.txt &
+CUDA_VISIBLE_DEVICES='' perf stat -o logs/test-model/$identification/2-single-core_cpu3_perf.txt -e cache-misses,cache-references,LLC-loads,LLC-load-misses,LLC-stores,LLC-store-misses taskset -c 3 python test-model.py 1 > logs/test-model/$identification/2-single-core_cpu3.txt &
 wait
 echo end 2-parallel-single-core
 
 echo 3-parallel-single-core
-CUDA_VISIBLE_DEVICES='' taskset -c 0 python test-model.py 1 > logs/test-model/$identification/3-single-core_cpu0.txt &
-CUDA_VISIBLE_DEVICES='' taskset -c 3 python test-model.py 1 > logs/test-model/$identification/3-single-core_cpu3.txt &
-CUDA_VISIBLE_DEVICES='' taskset -c 4 python test-model.py 1 > logs/test-model/$identification/3-single-core_cpu4.txt &
[…]

### 2018-01-01 picar-mini-kbd-drv8835.py
Threaded camera to reduce camera latency
PS3 Eye camera's minimum fps is 30Hz, while the DeepPicar's main
loop's control frequency is 20Hz.

Previously, cap.read() was performed once per each control loop,
which results in reading a stale buffered image in the camera.
The time difference---camera latency---was about 300-350ms.

By decoupling the camera reading as a separate thread, now the
control loop can fetch the most recent image frame, ignoring any
buffered stale image frames. With this update, the camera latency
is reduced to 0-50ms (i.e., up to 1 control loop latency).

@@ -8,6 +8,8 @@ import math
 import numpy as np
 import pygame
 import sys
+from threading import Thread
+
 # import rospy
 
 from pololu_drv8835_rpi import motors, MAX_SPEED
@@ -16,25 +18,6 @@ def deg2rad(deg):
     return deg * math.pi / 180.0
 def rad2deg(rad):
     return 180.0 * rad / math.pi
-
-motors.setSpeeds(0, 0)
-cap = cv2.VideoCapture(0)
-cap.set(3,640)
-cap.set(4,480)
-
-# ser = serial.Serial('/dev/ttyACM0', 115200, timeout=1)
-# ser = serial.Serial('/dev/ttyUSB0', 115200, timeout=1)
-fourcc = cv2.VideoWriter_fourcc(*'XVID')
-vidfile = cv2.VideoWriter('out-video.avi', fourcc, 15.0, (640,480))
-keyfile = open('out-key.csv', 'w+')
-keyfile_btn = open('out-key-btn.csv', 'w+')
-keyfile.write("ts_micro,frame,wheel\n")
-keyfile_btn.write("ts_micro,frame,btn,speed\n")
-rec_start_time = 0
-SET_SPEED = MAX_SPEED * 5 / 10 
-cur_speed = SET_SPEED
-print "MAX speed:", MAX_SPEED
-print "cur speed:", cur_speed
 
 def stop():
     global cur_speed
@@ -64,39 +47,88 @@ def turnOff():
     stop()
     center()
 
-atexit.register(turnOff)
-
-view_video = False
-frame_id = 0
-null_frame = np.zeros((160,120,3), np.uint8)
-cv2.imshow('frame', null_frame)
-
-angle = 0.0
-btn   = 107
-period = 0.05 # sec (=50ms)
-
-if len(sys.argv) == 2:
-    SET_SPEED = int(sys.argv[1])
-    print "Set new speed: ", SET_SPEED
-
-# rospy.init_node('my_node_name')
-# r = rospy.Rate(20) # 10hz
-
 def g_tick():
     t = time.time()
     count = 0
     while True:
         count += 1
         yield max(t + count*period - time.time(),0)
+
+class Camera:
+    def __init__(self, res=(320, 240), fps=30):
+        print "Initilize camera."
+        self.cap = cv2.VideoCapture(0)
+        self.cap.set(3, res[0]) # width
+        self.cap.set(4, res[1]) # height
+        self.cap.set(5, fps)
+        self.frame = None
+        self.enabled = True
[…]

### 2018-01-01 view-video.py

@@ -23,11 +23,14 @@ frame_count = cm.frame_count(vid_path)
 cap = cv2.VideoCapture(vid_path)
 imgs = []
 rows = cm.fetch_csv_data(csv_path)
+print ("%d %d" % (frame_count, len(rows)))
 assert frame_count == len(rows)
+ts_init = rows[0]['ts_micro']
 for i in range(frame_count):
     ret, img = cap.read()
-    text = 'frame: {}, wheel: {}'.format(i, rows[i]['wheel'])
-    cv2.putText(img, text, (50,50), font, 1, (255,255,255))
+    text = '{:.3f} {:3d} {:.3f}'.format(float((rows[i]['ts_micro']-ts_init))/1000.0,
+                                  rows[i]['frame'], rows[i]['wheel'])
+    cv2.putText(img, text, (5,50), font, 1.0, (255,255,255))
     imgs.append(img)
 
 frame_id = 0

### 2018-01-01 paper/comparison.tex

@@
 platform and compare the results. 
-For Tegra TX2, we have two different system configurations,
+For the Jetson TX2, we have two different system configurations,
 which differ in whether TensorFlow is configured to use its GPU or
@@
 the number of CPU cores used.
-First, both the Intel UP and Tegra TX2 exhibit superior performance when
+First, both the Intel UP and Jetson TX2 exhibit superior performance when
 compared with the Raspberry Pi 3. 
 When all four CPU cores are used, the Intel UP is 2.53X faster than
-the Pi 3, while TX (CPU) and TX (GPU) are 3.6X and 7.6X times faster,
+the Pi 3, while the TX2 (CPU) and TX2 (GPU) are 3.6X and 7.6X times faster,
 respectively, than the Pi 3. 
@@
 
-The Intel UP board and TX2 also perform much better when multiple DNN models
+The Intel UP board and Jetson TX2 also perform much better when multiple DNN models
 are co-scheduled. Figure~\ref{fig:sys_model} shows the results of the
@@
 satisfy 20 Hz real-time performance for all of the co-scheduled DNN control
-loops, and in case of TX 2 (GPU), 100 Hz real-time control is still
+loops, and in the case of the TX2 (GPU), 100 Hz real-time control is still
 feasible. Given that the GPU must be shared among the co-scheduled DNN
-models, the result suggests that TX2's GPU has sufficient capacity to
+models, the results suggest that the TX2's GPU has sufficient capacity to
 accomodate multiple instances of the DNN model we tested.
@@
 co-scheduled tasks. We also observe increased control loop execution
-timing in Intel Up and TX2, but the degree of the increase is not as
-dramatic as the Pi3. Compared to their respective solo timings (i.e.,
-the model runs on a single core in isolation), Intel UP suffers up to
+timing in the Intel Up and Jetson TX2, but the degree of the increase is 
+not as dramatic as the Pi 3. Compared to their respective solo timings 
+(i.e., the model runs on a single core in isolation), Intel UP suffers up to
 2.3X execution time increase; TX2 (CPU) and TX2 (GPU) suffer up to
-2.2X and 2.5X increase, respectively. This is somewhat suprising
-because Raspberry Pi 3's cores are in-order architecture based while
-the cores in Intel Up and NVIDIA TX2 are out-of-order architecture
+2.2X and 2.5X increases, respectively. This is somewhat suprising
+because the Raspberry Pi 3's cores are in-order architecture based while
+the cores in the Intel Up and NVIDIA TX2 are out-of-order architecture
 based, and that the memory intensive tasks on out-of-order cores can
@@
 
-Another interesting observation is that TX2 (GPU) also suffers
+Another interesting observation is that the TX2 (GPU) also suffers
 considerable execution time increase (2.5X) despite the fact that the
@@
 platforms where both CPU and GPU share the same memory
-subsystem~\cite{Ali2017}. As a result, TX2 (GPU) fails to meet the
+subsystem~\cite{Ali2017}. As a result, the TX2 (GPU) fails to meet the
 10ms deadline for 100 Hz control that would have been feasible if

### 2018-01-01 paper/evaluation.tex

@@
 In comparison, NVIDIA's DAVE-2 system, which has the exact same neural
-network architecture, reportedly runs at 30 Hz~\cite{Bojarski2016},
-just a bit faster than the DeepPicar. Although we believe it was not
+network architecture, reportedly runs at 30 Hz~\cite{Bojarski2016}, which 
+is just a bit faster than the DeepPicar. Although we believe it was not
 limited by their computing platform (we will experimentally compare
@@
 time decreases---from 61.96 ms on a single core to 42.67 ms on four
-cores (a 30\% improvement). However, the improvement is far from ideal
+cores (a 30\% improvement). However, the improvement is far from an ideal
 linear scaling. In particular, from 2 cores to 3 cores, the

### 2018-01-02 camera-opencv-latency.py
camera latency

@@ -16,7 +16,7 @@ import cv2
 ###############################################
 width=320
 height=240
-fps=15
+fps=20
 camera_channel=0
 fname="camera_latency"
 ###############################################
@@ -26,7 +26,7 @@ cv2.namedWindow("preview")
 vc = cv2.VideoCapture(camera_channel)
 vc.set(3, width)
 vc.set(4, height)
-# vc.set(cv2.CAP_PROP_FPS,fps) 
+vc.set(5, fps) 
 
 ################################################
 time.sleep(2)
@@ -52,9 +52,10 @@ try:
         if key == ord('q'): # exit on ESC
             break
         elif key == ord('s'):
-            cv2.imwrite(fname+str(ii)+".jpg", frame_o)
+            cv2.imwrite(fname+str(ii)+".png", frame_o)
             ii+=1
         rval, frame_o = vc.read()
+        
           
 finally: 
     vc.release()

### 2018-01-03 data_shuffled.py
new training function
instead of separating train/val data sets, it now takes
all data and use params.train_pct percent of them as
'train' set and the rest as 'val' set.

@@ -33,7 +33,50 @@ for p in purposes:
     for c in categories:
         imgs_cat[p][c] = []
         wheels_cat[p][c] = []
-    
+
+def load_imgs_v2():
+    global imgs
+    global wheels
+
+    for epoch_id in epochs:
+        print 'processing and loading epoch {} into memorys. train:{}, val:{}'.format(
+            epoch_id, len(imgs['train']), len(imgs['val']))
+        
+        # vid_path = cm.jn(data_dir, 'epoch{:0>2}_front.mkv'.format(epoch_id))
+        vid_path = cm.jn(data_dir, 'out-video-{}.avi'.format(epoch_id))
+
+        if not os.path.isfile(vid_path):
+            continue
+
+        frame_count = cm.frame_count(vid_path)
+        cap = cv2.VideoCapture(vid_path)
+
+
+        # csv_path = cm.jn(data_dir, 'epoch{:0>2}_steering.csv'.format(epoch_id))
+        csv_path = cm.jn(data_dir, 'out-key-{}.csv'.format(epoch_id))
+        assert os.path.isfile(csv_path)
+
+        rows = cm.fetch_csv_data(csv_path)
+        print len(rows), frame_count
+        assert frame_count == len(rows)
+
+        for row in rows:
+            ret, img = cap.read()
+            if not ret:
+                break
+
+            img = preprocess.preprocess(img)
+            angle = float(row['wheel']))
+            
+            if random.random() < params.train_pct:
+                imgs['train'].append(img)
+                wheels['train'].append(angle)
+            else:
+                imgs['val'].append(img)
+                wheels['val'].append(angle)
+
+        cap.release()
+        
 # load all preprocessed training images into memory
 def load_imgs():
     global imgs

### 2018-01-03 params.py

@@ -1,3 +1,4 @@
+
 #!/usr/bin/env python 
 from __future__ import division
 
@@ -29,8 +30,9 @@ assert os.path.isdir(data_dir)
 if not os.path.isdir(out_dir):
     os.makedirs(out_dir)
 
-epochs = OrderedDict()
+# epochs = OrderedDict()
+# epochs['train'] = [7,8,9,10]
+# epochs['val'] = [11,12]
 
-epochs['train'] = [7,8,9,10]
-epochs['val'] = [11,12]
-
+epochs = [7,8,9,10,11,12]
+train_pct = 0.8

### 2018-01-03 train.py

@@ -36,6 +36,13 @@ if write_summary:
     merged_summary_op = tf.summary.merge_all()
 
 saver = tf.train.Saver()
+
+# model_name = 'model.ckpt'
+# model_path = cm.jn(params.save_dir, model_name)
+# print "Load model file into the TF session"
+# saver.restore(sess, model_path)
+# print "Done..."
+
 time_start = time.time()
 
 # op to write logs to Tensorboard
@@ -43,7 +50,7 @@ if write_summary:
     summary_writer = tf.summary.FileWriter(params.save_dir, graph=tf.get_default_graph())
 
 if params.shuffle_training:
-    data.load_imgs()
+    data.load_imgs_v2()
 
 # center, curve 50:50%
 if params.use_category_normal:

### 2018-01-03 data_shuffled.py
new train function bug fix.

@@ -38,7 +38,7 @@ def load_imgs_v2():
     global imgs
     global wheels
 
-    for epoch_id in epochs:
+    for epoch_id in epochs['all']:
         print 'processing and loading epoch {} into memorys. train:{}, val:{}'.format(
             epoch_id, len(imgs['train']), len(imgs['val']))
         
@@ -50,7 +50,6 @@ def load_imgs_v2():
 
         frame_count = cm.frame_count(vid_path)
         cap = cv2.VideoCapture(vid_path)
-
 
         # csv_path = cm.jn(data_dir, 'epoch{:0>2}_steering.csv'.format(epoch_id))
         csv_path = cm.jn(data_dir, 'out-key-{}.csv'.format(epoch_id))
@@ -66,16 +65,18 @@ def load_imgs_v2():
                 break
 
             img = preprocess.preprocess(img)
-            angle = float(row['wheel']))
+            angle = float(row['wheel'])
             
             if random.random() < params.train_pct:
                 imgs['train'].append(img)
-                wheels['train'].append(angle)
+                wheels['train'].append([angle])
             else:
                 imgs['val'].append(img)
-                wheels['val'].append(angle)
+                wheels['val'].append([angle])
 
         cap.release()
+
+    print 'Total data: train:{}, val:{}'.format(len(imgs['train']), len(imgs['val']))
         
 # load all preprocessed training images into memory
 def load_imgs():
@@ -117,6 +118,7 @@ def load_imgs():
                 imgs[p].append(img)
 
             wheels[p].extend(yy)
+            # print "DBG:", wheels[p]
             assert len(imgs[p]) == len(wheels[p])
 
             cap.release()

### 2018-01-03 params.py

@@ -7,7 +7,7 @@ from collections import OrderedDict
 
 batch_size = 255
 
-VER="-donkey-T78910_V1112"
+VER="-donkey-T789101112"
 save_dir = os.path.abspath('../picar/models' + VER)
 
 training_steps = 2000
@@ -30,9 +30,9 @@ assert os.path.isdir(data_dir)
 if not os.path.isdir(out_dir):
     os.makedirs(out_dir)
 
-# epochs = OrderedDict()
+epochs = OrderedDict()
 # epochs['train'] = [7,8,9,10]
 # epochs['val'] = [11,12]
 
-epochs = [7,8,9,10,11,12]
+epochs['all'] = [7,8,9,10,11,12]
 train_pct = 0.8

### 2018-01-03 camera-pi.py
Delete camera-pi.py

@@ -1,8 +1,0 @@
-from picamera import PiCamera
-from time import sleep
-
-camera = PiCamera()
-
-camera.start_preview()
-sleep(10)
-camera.stop_preview()

### 2018-01-03 setup.sh
Delete setup.sh

@@ -1,7 +1,0 @@
-roscore &
-sleep 5
-roslaunch rtimulib_ros rtimulib_ros.launch &  # publish 'imu' channel
-rosrun teleop_twist_joy teleop_node &         # publish 'cmd_vel' output from joystick input
-./servod &              # enable pwm output for the speed motor
-./src/motor_driver.py & # subscribe 'cmd_vel' channel
-# rosrun teleop_twist_keyboard teleop_twist_keyboard.py &

### 2018-01-03 picar-mini-dnn-drv8835_btn.py
Delete picar-mini-dnn-drv8835_btn.py

@@ -1,195 +1,0 @@
-#!/usr/bin/python
-from __future__ import division
-import tensorflow as tf
-import model_btn as model
-import params_btn as params
-import local_common as cm
-import preprocess
-
-import random
-import os
-import time
-import atexit
-import serial
-import cv2
-import math
-import numpy as np
-import pygame
-from pololu_drv8835_rpi import motors, MAX_SPEED
-
-def deg2rad(deg):
-    return deg * math.pi / 180.0
-def rad2deg(rad):
-    return 180.0 * rad / math.pi
-
-sess = tf.InteractiveSession()
-saver = tf.train.Saver()
-model_name = 'model.ckpt'
-model_path = cm.jn(params.save_dir, model_name)
-saver.restore(sess, model_path)
-
-motors.setSpeeds(0, 0)
-cap = cv2.VideoCapture(0)
-cap.set(3,640)
-cap.set(4,480)
-
-# ser = serial.Serial('/dev/ttyACM0', 115200, timeout=1)
-# ser = serial.Serial('/dev/ttyUSB0', 115200, timeout=1)
-fourcc = cv2.VideoWriter_fourcc(*'XVID')
-vidfile = cv2.VideoWriter('out-video.avi', fourcc, 15.0, (640,480))
-keyfile = open('out-key.csv', 'w+')
-keyfile_btn = open('out-key-btn.csv', 'w+')
-keyfile.write("ts_micro,frame,wheel\n")
-keyfile_btn.write("ts_micro,frame,btn,speed\n")
-rec_start_time = 0
-SET_SPEED = MAX_SPEED # /2 + 2*MAX_SPEED/10
-cur_speed = SET_SPEED
-print "MAX speed:", MAX_SPEED
-
-def stop():
-    global cur_speed
-    cur_speed = 0
-    motors.motor2.setSpeed(int(cur_speed))
-        
-def ffw():
-    global cur_speed
-    # cur_speed = min(MAX_SPEED, cur_speed + MAX_SPEED/10)
-    cur_speed = SET_SPEED
-    motors.motor2.setSpeed(int(cur_speed))
-
-def rew():
-    global cur_speed        
-    # cur_speed = max(-MAX_SPEED, cur_speed - MAX_SPEED/10)
-    cur_speed = SET_SPEED
-    motors.motor2.setSpeed(int(cur_speed))
-
-def center():
-    motors.motor1.setSpeed(0)
-def left():
[…]

### 2018-01-03 picar-mini-kbd-arduino.py
Delete picar-mini-kbd-arduino.py

@@ -1,129 +1,0 @@
-#!/usr/bin/python
-import os
-import time
-import atexit
-import serial
-import cv2
-import math
-import numpy as np
-import pygame
-
-cap = cv2.VideoCapture(0)
-cap.set(3,320)
-cap.set(4,240)
-ser = serial.Serial('/dev/ttyACM0', 115200, timeout=1)
-fourcc = cv2.VideoWriter_fourcc(*'XVID')
-vidfile = cv2.VideoWriter('out-video.avi', fourcc, 20.0, (320,240))
-keyfile = open('out-key.csv', 'w+')
-keyfile_btn = open('out-key-btn.csv', 'w+')
-keyfile.write("ts_micro,frame,wheel\n")
-keyfile_btn.write("ts_micro,frame,btn,speed\n")
-rec_start_time = 0
-
-def stop():
-        ser.write("s\n")
-def ffw():
-        ser.write("f\n")
-def rew():
-        ser.write("b\n")
-
-def center():
-        ser.write("c\n")
-def left():
-        ser.write("l\n")
-def right():
-        ser.write("r\n")
-
-def degree2rad(deg):
-        return deg * math.pi / 180.0
-
-def turnOff():
-        stop()
-        center()
-
-atexit.register(turnOff)
-
-view_video = False
-frame_id = 0
-null_frame = np.zeros((160,120,3), np.uint8)
-cv2.imshow('frame', null_frame)
-
-angle = 0.0
-btn = ''
-
-while (True):
-        # read a frame
-        ret, frame = cap.read()
-        ts = int(time.time() * 1000000)
-        
-        if ret == False:
-		print "Camera is not ready"
-                break
-
-        if view_video == True:
-                cv2.imshow('frame', frame)
-
-        ch = cv2.waitKey(50) & 0xFF
-
-        if ch == ord('j'):
-                left()
-                print "left"
-                angle = degree2rad(-30)
-                btn   = ch
-        elif ch == ord('k'):
-                center()
-                print "center"
-                angle = degree2rad(0)
-                btn   = ch                
-        elif ch == ord('l'):
-                right()
-                print "right"
-                angle = degree2rad(30)
-                btn   = ch                
-        elif ch == ord('a'):
-                ffw()
-                print "accel"
-        elif ch == ord('s'):
-                stop()
-                print "stop"                
-                btn   = ch
-        elif ch == ord('z'):
-                rew()
-                print "reverse"
-        elif ch == ord('q'):
-                break
[…]

### 2018-01-03 params.py
Update params.py
cleanup

@@ -5,21 +5,17 @@ import os
 from collections import OrderedDict
 
 batch_size = 100
-# save_dir = os.path.abspath('models')
-save_dir = os.path.abspath('../picar/models-v22-lr')
 training_steps = 2000
 img_height = 66
 img_width = 200
 img_channels = 3
 write_summary = True
-use_category_normal = True # if ture, center/curve images are equally selected.
+use_category_normal = False # if ture, center/curve images are equally selected.
 
 # change this to the directory that contains the source videos
-data_dir = os.path.abspath('../picar/epochs')
-out_dir = os.path.abspath('../picar/output-v22-lr')
-
-# data_dir = os.path.abspath('epochs')
-# out_dir = os.path.abspath('.output')
+save_dir = os.path.abspath('models')
+data_dir = os.path.abspath('epochs')
+out_dir = os.path.abspath('output')
 
 shuffle_training = True
 
@@ -28,51 +24,5 @@ if not os.path.isdir(out_dir):
     os.makedirs(out_dir)
 
 epochs = OrderedDict()
-# 8/23 picar-v2.0, 50ms fix
-
-epochs['train'] = [401, 402, 403, 404, 409, 410, 411, 412]
-# epochs['train'] += [406, 407, 408] # dagger
-epochs['val'] = [405, 413]
-
-#epochs['train'] += [414, 415, 416, 417, 419, 420, 421, 422]
-#epochs['val'] += [418, 423]
-
-# 8/21 picar-v2.0
-# left = 302 - 306
-# right = 307 - 311
-# epochs['train'] = [302, 303, 304, 305, 307, 308, 309, 310]
-# epochs['val'] = [306, 311] 
-
-# center following.
-# left  = 120-135 (18)
-# right = 117-119, 136 - 144, 100-101 (14)
-# epochs['train'] = [100, 101, 117, 118, 119, 120, 121, 122, 123, 124, 125, 126, 127, 128, 129, 130, 131, 132, 133, 134, 135, 136, 137, 138, 139, 140, 141, 142, 143, 144]
-# epochs['val'] =   [145, 146]
-
-
-# # right only
-# epochs['train'] = [100, 101, 110, 111, 117, 118]
-# epochs['val'] =   [112, 119]
-
-# right & left (120-124)
-# epochs['train'] = [100, 101, 110, 111, 117, 118, 120, 121, 122, 123]
[…]

### 2018-01-04 paper/main.tex
Change paper to ECRTS format

--- 2017-10-16 paper/main.tex
@@
-%
-%  $Author: ienne $
-%  $Date: 1995/09/15 15:20:59 $
-%  $Revision: 1.4 $
-%
+\documentclass[a4paper,UKenglish]{lipics-v2016}
+%This is a template for producing LIPIcs articles. 
+%See lipics-manual.pdf for further information.
+%for A4 paper format use option "a4paper", for US-letter use option "letterpaper"
+%for british hyphenation rules use option "UKenglish", for american hyphenation rules use option "USenglish"
+% for section-numbered lemmas etc., use "numberwithinsect"
 
-% \documentclass[10pt,journal,cspaper,compsoc]{IEEEtran}   %%%tc version
-\documentclass[10pt, conference]{IEEEtran}
-%\documentclass[conference,compsoc]{IEEEtran}
-%\documentclass[10pt, conference]{IEEEtran}
-%\documentclass[times, 10pt,onecolumn]{article}
-\usepackage{amsmath, amssymb, enumerate}
+\usepackage{microtype}%if unwanted, comment out or use option "draft"
 
-%%%%%%%%%%%%%%%% page control%%%%%%%%%%%%%%%%%
-%\usepackage[margin=0.75in]{geometry}
+%\graphicspath{{./graphics/}}%helpful if your graphic files are in another directory
 
-%\linespread{0.991}  %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% this is really useful
-%\usepackage{cite}
-\usepackage{fancybox}
-\usepackage{amsfonts}
-%\usepackage{algorithm}
-%\usepackage[noend]{algorithmic}
-\usepackage[usenames]{color}
-%\usepackage{colortbl}
-%\usepackage[ figure, boxed, vlined]{algorithm2e}
-%\usepackage[linesnumbered,vlined]{algorithm2e}
-%\usepackage[lined,boxed]{algorithm2e}
-\usepackage{listings}
+\bibliographystyle{plainurl}% the recommended bibstyle
 
-\usepackage[linesnumbered,vlined]{algorithm2e}
-\usepackage{graphicx}
-\usepackage{times}
-\usepackage{psfrag}
-\usepackage{subfigure}
-\usepackage{caption}
-%\usepackage{subcaption}
-\usepackage{multirow}
-%\usepackage{setspace}
-%\usepackage{listings}
-\usepackage{epsfig}
-%\usepackage{epstopdf}
-%\usepackage[font=small,labelfont=bf]{caption}
-\usepackage{url}
+% Author macros::begin %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\title{DeepPicar: A Low-cost Deep Neural Network-based Autonomous Car}
+\titlerunning{DeepPicar: A Low-cost Autonomous Car} %optional, in case that the title is too long; the running title should fit into the top page column
 
-\usepackage{color}
-\def\fixme#1{\typeout{FIXED in page \thepage : {#1}}
-%\bgroup \color{red}{} \egroup}
-\bgroup \color{red}{[FIXME: {#1}]} \egroup}
+%% Please provide for each author the \author and \affil macro, even when authors have the same affiliation, i.e. for each author there needs to be the  \author and \affil macros
+\author[1]{Michael G. Bechtel}
+\author[2]{Elise McEllhiney}
+\author[3]{Heechul Yun}
+\affil[1]{University of Kansas, Lawrence, United States\\
+  \texttt{m783b224@ku.edu}}
+\affil[2]{University of Kansas, Lawrence, United States\\
+  \texttt{e908m429@ku.edu}}
+\affil[3]{University of Kansas, Lawrence, United States\\
+  \texttt{heechul.yun@ku.edu}}
+\authorrunning{M.\,G. Bechtel, E. McEllhiney, H. Yun} %mandatory. First: Use abbreviated first/middle names. Second (only in severe cases): Use first author plus 'et. al.'
 
-\usepackage[pdftex]{hyperref}
-\usepackage{rotating,tabularx}
+\Copyright{Michael G. Bechtel, Elise McEllhiney, Heechul Yun}%mandatory, please use full first names. LIPIcs license is "CC-BY";  http://creativecommons.org/licenses/by/3.0/
 
-\interfootnotelinepenalty=10000
+\subjclass{Dummy classification -- please refer to \url{http://www.acm.org/about/class/ccs98-html}}% mandatory: Please choose ACM 1998 classifications from http://www.acm.org/about/class/ccs98-html .
+E.g., cite as "F.1.1 Models of Computation".
+\keywords{Dummy keyword -- please provide 1--5 keywords}% mandatory: Please provide 1-5 keywords
+% Author macros::end %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
-%% Define a new 'leo' style for the package that will use a smaller font.
-\makeatletter
-\def\url@leostyle{%
-  \@ifundefined{selectfont}{\def\UrlFont{\sf}}{\def\UrlFont{\small\ttfamily}}}
-\makeatother
-
-%\documentstyle[times,art10,twocolumn,latex8]{article}
-
-%-------------------------------------------------------------------------
-% take the % away on next line to produce the final camera-ready version
-\pagestyle{plain}
-%\thispagestyle{empty}
-%\pagestyle{empty}
-
-\newtheorem{theorem}{Theorem}
-\newtheorem{lemma}[theorem]{Lemma}
-
-%% remaining budget share, used in task stall section.
-\newcommand{\bottomrule}{\hline}
-\newcommand{\toprule}{\hline}
-\newcommand{\midrule}{\hline}
-%-------------------------------------------------------------------------
+%Editor-only macros:: begin (do not touch as author)%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\EventEditors{Sebastian Altmeyer}
+\EventNoEds{1}
+\EventLongTitle{30th Euromicro Conference on Real-Time Systems (ECRTS 2018)}
+\EventShortTitle{ECRTS 2018}
+\EventAcronym{ECRTS}
+\EventYear{2018}
+\EventDate{July 3--6, 2018}
+\EventLocation{Barcelona, Spain}
+\EventLogo{}
+\SeriesVolume{XX}
+\ArticleNo{YY}
+% Editor-only macros::end %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \begin{document}
-
-\title{DeepPicar:​ ​A​ ​Low-cost​ ​Deep​ ​Neural​ ​Network-based​ ​Autonomous​ ​Car}
-\author{Michael Garrett Bechtel, Elise McEllhiney, Heechul Yun\\
-\{m783b224, e908m429, heechul.yun\}@ku.edu\\
-University of Kansas, USA\\ 
-}
 
@@
 %-------------------------------------------------------------------------
-\bibliographystyle{abbrv}
+%\bibliographystyle{abbrv}
 \bibliography{reference}

### 2018-01-05 sync-video.py
Skip frames to synchronize with the steering data.
Example:

$ python sync-video.py 40 6

skip 6 frames from the epoch_id 40 and save the converted data.

source (unsynchronized data):
  epochs/out-video-6.avi
  epochs/out-key-6.csv

destination (re-synchronized data):
  epochs-conv/out-video-6.avi
  epochs-conv/out-key-6.csv

@@ -1,0 +1,58 @@
+import cv2
+import local_common as cm
+import os
+import sys
+
+font = cv2.FONT_HERSHEY_SIMPLEX
+
+if len(sys.argv) == 3:
+    epoch_id = int(sys.argv[1])
+    skip_count = int(sys.argv[2])
+    vid_path = 'epochs/out-video-{}.avi'.format(epoch_id)
+    csv_path = 'epochs/out-key-{}.csv'.format(epoch_id)
+    conv_vid_path = 'epochs-conv/out-video-{}.avi'.format(epoch_id)
+    conv_csv_path = 'epochs-conv/out-key-{}.csv'.format(epoch_id)
+else:
+    sys.exit(1)
+
+print ("epoch_id: {}, skip_count: {}".format(epoch_id, skip_count))
+
+assert os.path.isfile(vid_path)
+assert os.path.isfile(csv_path)
+
+frame_count = cm.frame_count(vid_path)
+cap = cv2.VideoCapture(vid_path)
+cam_width = int(cap.get(3))
+cam_height = int(cap.get(4))
+cam_fps = int(cap.get(5))
+
+print ("w: {}, h:{}, fps: {}".format(cam_width, cam_height, cam_fps))
+
+rows = cm.fetch_csv_data(csv_path)
+assert frame_count == len(rows)
+
+# data files
+fourcc = cv2.cv.CV_FOURCC(*'XVID')
+vidfile = cv2.VideoWriter(conv_vid_path, fourcc,
+                          cam_fps, (cam_width, cam_height))
+keyfile = open(conv_csv_path, 'w+')
+keyfile.write("ts_micro,frame,wheel\n")
+
+# skip video frames
+for i in range(skip_count):
+    ret, img = cap.read()
+
+# record adjusted data
+for i in range(frame_count-skip_count):    
+    # write input (angle)
+    text = "{},{},{}\n".format(rows[i]['ts_micro'],
+                               rows[i]['frame'], rows[i]['wheel'])
+    keyfile.write(text)
+    ret, img = cap.read()
+    vidfile.write(img)
+
+print("Close files.")
+cap.release()
+keyfile.close()
+vidfile.release()
+    

### 2018-01-05 run.py
epoch based parallelized visualization.

@@ -10,12 +10,15 @@ import params
 import sys
 import os
 import preprocess
-import visualize_parallel
-# import visualize
+# import visualize_parallel
+import visualize
 import time
 import math
 
 import local_common as cm
+
+from joblib import Parallel, delayed 
+import multiprocessing
 
 def deg2rad(deg):
         return deg * math.pi / 180.0
@@ -29,9 +32,8 @@ model_path = cm.jn(params.save_dir, model_name)
 saver.restore(sess, model_path)
 
 epoch_ids = sorted(list(set(itertools.chain(*params.epochs.values()))))
-# epoch_ids = [6]
 
-for epoch_id in epoch_ids:
+def process_epoch(epoch_id):
     print '---------- processing video for epoch {} ----------'.format(epoch_id)
     # vid_path = cm.jn(params.data_dir, 'epoch{:0>2}_front.mkv'.format(epoch_id))
 
@@ -72,13 +74,22 @@ for epoch_id in epoch_ids:
     cap.release()
 
     fps = frame_count / (time.time() - time_start)
-    
-    print 'completed inference, total frames: {}, average fps: {} Hz'.format(frame_count, round(fps, 1))
+    print ('completed inference, total frames: {}, average fps: {} Hz'.format(frame_count, round(fps, 1)))
+    # print "Machine Steering:", machine_steering
+    return machine_steering
 
-    print "Machine Steering:", machine_steering
-
-    print 'performing visualization...'
-    visualize_parallel.visualize(epoch_id, machine_steering, params.out_dir,
+def visualize_epoch(job):
+    epoch_id = job[0]
+    machine_steering = job[1]
+    visualize.visualize(epoch_id, machine_steering, params.out_dir,
                         verbose=True, frame_count_limit=None)
-    
-    
+job_list = []
+for epoch_id in epoch_ids:
+    steering = process_epoch(epoch_id)
+    job_list.append([epoch_id, steering])
+
+print "visualize epochs in parallel"    
+num_cores = 4 # multiprocessing.cpu_count()
+Parallel (n_jobs = num_cores) (delayed (visualize_epoch) (job) for job in job_list)
+
+     

### 2018-01-05 visualize.py

@@ -1,0 +1,238 @@
+#!/usr/bin/env python
+from __future__ import division
+
+import subprocess as sp
+import os
+import io
+import sys
+import re
+from copy import deepcopy
+import psycopg2
+import psycopg2.extras
+import subprocess
+from operator import itemgetter
+from collections import OrderedDict, Counter
+import cv2
+from fractions import Fraction
+import csv
+import errno
+import time
+import shutil
+import numpy as np
+import datetime
+import matplotlib.pyplot as plt
+import PIL
+from pprint import pprint
+import multiprocessing
+import string
+import copy
+import json
+import random
+import params
+import math
+
+import local_common as cm
+
+def rad2deg(rad):
+    return 180.0 * rad / math.pi
+
+def get_human_steering(epoch_id):
+    epoch_dir = params.data_dir
+    assert os.path.isdir(epoch_dir)
+    steering_path = cm.jn(epoch_dir, 'out-key-{}.csv'.format(epoch_id))
+    # steering_path = cm.jn(epoch_dir, 'epoch{:0>2}_steering.csv'.format(epoch_id))
+    assert os.path.isfile(steering_path)
+    
+    rows = cm.fetch_csv_data(steering_path)
+    human_steering = [rad2deg(row['wheel']) for row in rows]
+
+    return human_steering
+
+def visualize(epoch_id, machine_steering, out_dir, perform_smoothing=False,
+              verbose=False, verbose_progress_step = 100, frame_count_limit = None):
+    epoch_dir = params.data_dir
+    human_steering = get_human_steering(epoch_id)
+    assert len(human_steering) == len(machine_steering)
+
+    # testing: artificially magnify steering to test steering wheel visualization
+    # human_steering = list(np.array(human_steering) * 10)
+    # machine_steering = list(np.array(machine_steering) * 10)
+
+    # testing: artificially alter machine steering to test that the disagreement coloring is working
+    # delta = 0
+    # for i in xrange(len(machine_steering)):
+    #     delta += random.uniform(-1, 1)
+    #     machine_steering[i] += delta
+    
+    if perform_smoothing:
+        machine_steering = list(cm.smooth(np.array(machine_steering)))
[…]

### 2018-01-05 visualize_parallel.py

@@ -1,267 +1,0 @@
-#!/usr/bin/env python
-from __future__ import division
-
-import subprocess as sp
-import os
-import io
-import sys
-import re
-from copy import deepcopy
-import psycopg2
-import psycopg2.extras
-import subprocess
-from operator import itemgetter
-from collections import OrderedDict, Counter
-import cv2
-from fractions import Fraction
-import csv
-import errno
-import time
-import shutil
-import numpy as np
-import datetime
-import matplotlib.pyplot as plt
-import PIL
-from pprint import pprint
-import multiprocessing
-import string
-import copy
-import json
-import random
-import params
-import math
-from joblib import Parallel, delayed 
-import multiprocessing
-import local_common as cm
-
-machine_steering = []
-human_steering = []
-cam_images = []
-
-wimg = cm.imread(os.path.abspath("images/wheel-tesla-image-150.png"), cv2.IMREAD_UNCHANGED)
-timg_green_agree = cm.imread(os.path.abspath("images/text-green-agree.png"), cv2.IMREAD_UNCHANGED)
-timg_ground_truth = cm.imread(os.path.abspath("images/text-ground-truth.png"), cv2.IMREAD_UNCHANGED)
-timg_learned_control = cm.imread(os.path.abspath("images/text-learned-control.png"), cv2.IMREAD_UNCHANGED)
-timg_red_disagree = cm.imread(os.path.abspath("images/text-red-disagree.png"), cv2.IMREAD_UNCHANGED)
-timg_tesla_control_autopilot = cm.imread(os.path.abspath("images/text-tesla-control-autopilot.png"), cv2.IMREAD_UNCHANGED)
-timg_tesla_control_human = cm.imread(os.path.abspath("images/text-tesla-control-human.png"), cv2.IMREAD_UNCHANGED)
-# timg_ = cm.imread(os.path.abspath("images/text-.png"), cv2.IMREAD_UNCHANGED)
-
-def rad2deg(rad):
-    return 180.0 * rad / math.pi
-    
-def get_human_steering(epoch_id):
-    epoch_dir = params.data_dir
-    assert os.path.isdir(epoch_dir)
-    steering_path = cm.jn(epoch_dir, 'out-key-{}.csv'.format(epoch_id))
-    # steering_path = cm.jn(epoch_dir, 'epoch{:0>2}_steering.csv'.format(epoch_id))
-    assert os.path.isfile(steering_path)
-    
[…]

### 2018-01-05 params.py
Set category_normal as default.
It appears to work much better now based on the visulazation results

@@ -5,74 +5,25 @@ import os
 from collections import OrderedDict
 
 batch_size = 100
-# save_dir = os.path.abspath('models')
-save_dir = os.path.abspath('../picar/models-v22-lr')
 training_steps = 2000
 img_height = 66
 img_width = 200
 img_channels = 3
+
 write_summary = True
 use_category_normal = True # if ture, center/curve images are equally selected.
+shuffle_training = True
 
 # change this to the directory that contains the source videos
-data_dir = os.path.abspath('../picar/epochs')
-out_dir = os.path.abspath('../picar/output-v22-lr')
-
-# data_dir = os.path.abspath('epochs')
-# out_dir = os.path.abspath('.output')
-
-shuffle_training = True
+save_dir = os.path.abspath('models-5conv_3fc')
+data_dir = os.path.abspath('epochs-conv')
+out_dir = os.path.abspath('output-5conv_3fc')
 
 assert os.path.isdir(data_dir)
 if not os.path.isdir(out_dir):
     os.makedirs(out_dir)
 
 epochs = OrderedDict()
-# 8/23 picar-v2.0, 50ms fix
-
-epochs['train'] = [401, 402, 403, 404, 409, 410, 411, 412]
-# epochs['train'] += [406, 407, 408] # dagger
-epochs['val'] = [405, 413]
-
-#epochs['train'] += [414, 415, 416, 417, 419, 420, 421, 422]
-#epochs['val'] += [418, 423]
-
-# 8/21 picar-v2.0
-# left = 302 - 306
-# right = 307 - 311
-# epochs['train'] = [302, 303, 304, 305, 307, 308, 309, 310]
-# epochs['val'] = [306, 311] 
-
-# center following.
-# left  = 120-135 (18)
-# right = 117-119, 136 - 144, 100-101 (14)
-# epochs['train'] = [100, 101, 117, 118, 119, 120, 121, 122, 123, 124, 125, 126, 127, 128, 129, 130, 131, 132, 133, 134, 135, 136, 137, 138, 139, 140, 141, 142, 143, 144]
-# epochs['val'] =   [145, 146]
-
-
-# # right only
-# epochs['train'] = [100, 101, 110, 111, 117, 118]
-# epochs['val'] =   [112, 119]
-
-# right & left (120-124)
[…]

### 2018-01-05 sync-video.py
opencv3 support

@@ -32,7 +32,8 @@ rows = cm.fetch_csv_data(csv_path)
 assert frame_count == len(rows)
 
 # data files
-fourcc = cv2.cv.CV_FOURCC(*'XVID')
+# fourcc = cv2.cv.CV_FOURCC(*'XVID')
+fourcc = cv2.VideoWriter_fourcc(*'XVID')
 vidfile = cv2.VideoWriter(conv_vid_path, fourcc,
                           cam_fps, (cam_width, cam_height))
 keyfile = open(conv_csv_path, 'w+')

### 2018-01-05 model.py
make it a symbolic link

@@ -1,92 +1,1 @@
-#!/usr/bin/env python 
-from __future__ import division
-
-import tensorflow as tf
-import params
-
-def weight_variable(name, shape):
-    return tf.get_variable(name, shape=shape, initializer=tf.contrib.layers.xavier_initializer())
-    # initial = tf.truncated_normal(shape, stddev=0.1)
-    # return tf.Variable(initial)
-
-def bias_variable(shape):
-    initial = tf.constant(0.1, shape=shape)
-    return tf.Variable(initial)
-
-def conv2d(x, W, stride):
-    return tf.nn.conv2d(x, W, strides=[1, stride, stride, 1], padding='VALID')
-
-x = tf.placeholder(tf.float32, shape=[None, params.img_height, params.img_width, params.img_channels])
-y_ = tf.placeholder(tf.float32, shape=[None, 1])
-
-x_image = x
-
-# first convolutional layer
-W_conv1 = weight_variable("wc1", [5, 5, 3, 24])
-b_conv1 = bias_variable([24])
-
-h_conv1 = tf.nn.relu(conv2d(x_image, W_conv1, 2) + b_conv1)
-
-# second convolutional layer
-W_conv2 = weight_variable("wc2", [5, 5, 24, 36])
-b_conv2 = bias_variable([36])
-
-h_conv2 = tf.nn.relu(conv2d(h_conv1, W_conv2, 2) + b_conv2)
-
-# third convolutional layer
-W_conv3 = weight_variable("wc3", [5, 5, 36, 48])
-b_conv3 = bias_variable([48])
-
-h_conv3 = tf.nn.relu(conv2d(h_conv2, W_conv3, 2) + b_conv3)
-
-# fourth convolutional layer
-W_conv4 = weight_variable("wc4", [3, 3, 48, 64])
-b_conv4 = bias_variable([64])
-
-h_conv4 = tf.nn.relu(conv2d(h_conv3, W_conv4, 1) + b_conv4)
-
-# fifth convolutional layer
-W_conv5 = weight_variable("wc5", [3, 3, 64, 64])
-b_conv5 = bias_variable([64])
-
-h_conv5 = tf.nn.relu(conv2d(h_conv4, W_conv5, 1) + b_conv5)
-
-# fully connected layer 1
[…]

### 2018-01-06 params.py
params for 5conv-3fc-home model

@@ -15,15 +15,15 @@ use_category_normal = True # if ture, center/curve images are equally selected.
 shuffle_training = True
 
 # change this to the directory that contains the source videos
-save_dir = os.path.abspath('models-5conv_3fc')
 data_dir = os.path.abspath('epochs-conv')
-out_dir = os.path.abspath('output-5conv_3fc')
+save_dir = os.path.abspath('models-5conv_3fc-home')
+out_dir = os.path.abspath('output-5conv_3fc-home')
 
 assert os.path.isdir(data_dir)
 if not os.path.isdir(out_dir):
     os.makedirs(out_dir)
 
 epochs = OrderedDict()
-epochs['train'] = range(30, 50)
-epochs['val'] = range(50, 60)
+epochs['train'] = [100, 101, 103, 104] # range(30, 50)
+epochs['val'] = [102, 105] # range(50, 60)
 

### 2018-01-16 modelOld.py
Update model.py and add compatible models
The old model still exists in modelOld.py.

@@ -50,15 +50,22 @@ W_conv5 = weight_variable("wc5", [3, 3, 64, 64])
 b_conv5 = bias_variable([64])
 
 h_conv5 = tf.nn.relu(conv2d(h_conv4, W_conv5, 1) + b_conv5)
+
+# fully connected layer 1
+W_fc1 = weight_variable("fc1", [1152, 1164])
+b_fc1 = bias_variable([1164])
+
 h_conv5_flat = tf.reshape(h_conv5, [-1, 1152])
+h_fc1 = tf.nn.relu(tf.matmul(h_conv5_flat, W_fc1) + b_fc1)
 
 keep_prob = tf.placeholder(tf.float32)
+h_fc1_drop = tf.nn.dropout(h_fc1, keep_prob)
 
 # fully connected layer 2
-W_fc2 = weight_variable("fc2", [1152, 100])
+W_fc2 = weight_variable("fc2", [1164, 100])
 b_fc2 = bias_variable([100])
 
-h_fc2 = tf.nn.relu(tf.matmul(h_conv5_flat, W_fc2) + b_fc2)
+h_fc2 = tf.nn.relu(tf.matmul(h_fc1_drop, W_fc2) + b_fc2)
 
 h_fc2_drop = tf.nn.dropout(h_fc2, keep_prob)
 

### 2018-01-16 test-model3.py
Update params.py and add new test-model files
Added save directory variables for the additional models that are
accessed by the new folders.

@@ -31,7 +31,7 @@ NFRAMES = 1000
 sess = tf.InteractiveSession(config=config)
 saver = tf.train.Saver()
 model_name = 'model.ckpt'
-model_path = cm.jn(params.save_dir2, model_name)
+model_path = cm.jn(params.save_dir3, model_name)
 saver.restore(sess, model_path)
 
 epoch_ids = sorted(list(set(itertools.chain(*params.epochs.values()))))

### 2018-01-16 test-model4.py

@@ -31,7 +31,7 @@ NFRAMES = 1000
 sess = tf.InteractiveSession(config=config)
 saver = tf.train.Saver()
 model_name = 'model.ckpt'
-model_path = cm.jn(params.save_dir2, model_name)
+model_path = cm.jn(params.save_dir4, model_name)
 saver.restore(sess, model_path)
 
 epoch_ids = sorted(list(set(itertools.chain(*params.epochs.values()))))

### 2018-01-16 benchmark_timings_Pi_RT.sh
Add scripts that use RT priority
The new scripts are identical to their original counterparts except that
they add RT priority (sudo chrt -f 99) to any command where a test-model
file is called.

Also updated test-model_timings_Pi.sh to use the different models, and
replaced the bandwidth executable, which was 64-bit, with a 32-bit
counterpart so that it can run on the platforms.

@@ -1,0 +1,57 @@
+timestamp=$(date +%y-%m-%d_%H-%M-%S)
+
+echo "Which board is this test being run on?  Input a string and press [ENTER]"
+read board
+
+identification=$board
+identification+="_"
+identification+=$timestamp
+
+mkdir -p logs/test-model-and-benchmark/$identification
+echo "This test will take approximately 30 minutes to complete"
+echo "The output files can be found in: logs/test-model/$identification"
+
+echo 2-parallel-single-core
+./bandwidth -a write -m 16384 -t 10000 -c 1 &
+sudo chrt -f 99 perf stat -o logs/test-model-and-benchmark/$identification/1-sc-bench_1-sc-model_cpu0_perf.txt -e cache-misses,cache-references,LLC-loads,LLC-load-misses,LLC-stores,LLC-store-misses taskset -c 0 python test-model.py 1 > logs/test-model-and-benchmark/$identification/1-sc-bench_1-sc-model_cpu0.txt
+jobs -p | xargs kill
+echo end 2-parallel-single-core
+sleep 300
+
+echo 3-parallel-single-core
+./bandwidth -a write -m 16384 -t 10000 -c 1 &
+./bandwidth -a write -m 16384 -t 10000 -c 2 &
+sudo chrt -f 99 perf stat -o logs/test-model-and-benchmark/$identification/2-sc-bench_1-sc-model_cpu0_perf.txt -e cache-misses,cache-references,LLC-loads,LLC-load-misses,LLC-stores,LLC-store-misses taskset -c 0 python test-model.py 1 > logs/test-model-and-benchmark/$identification/2-sc-bench_1-sc-model_cpu0.txt
+jobs -p | xargs kill
+echo end 3-parallel-single-core
+sleep 300
+
+echo 4-parallel-single-core
+./bandwidth -a write -m 16384 -t 10000 -c 1 &
+./bandwidth -a write -m 16384 -t 10000 -c 2 &
+./bandwidth -a write -m 16384 -t 10000 -c 3 &
[…]

### 2018-01-16 test-model_timings_Pi.sh

@@ -19,24 +19,24 @@ sleep 300
 
 echo 2-parallel-single-core
 perf stat -o logs/test-model/$identification/2-single-core_cpu0_perf.txt -e cache-misses,cache-references,LLC-loads,LLC-load-misses,LLC-stores,LLC-store-misses taskset -c 0 python test-model.py 1 > logs/test-model/$identification/2-single-core_cpu0.txt &
-perf stat -o logs/test-model/$identification/2-single-core_cpu1_perf.txt -e cache-misses,cache-references,LLC-loads,LLC-load-misses,LLC-stores,LLC-store-misses taskset -c 1 python test-model.py 1 > logs/test-model/$identification/2-single-core_cpu1.txt &
+perf stat -o logs/test-model/$identification/2-single-core_cpu1_perf.txt -e cache-misses,cache-references,LLC-loads,LLC-load-misses,LLC-stores,LLC-store-misses taskset -c 1 python test-model2.py 1 > logs/test-model/$identification/2-single-core_cpu1.txt &
 wait
 echo end 2-parallel-single-core
 sleep 300
 
 echo 3-parallel-single-core
 perf stat -o logs/test-model/$identification/3-single-core_cpu0_perf.txt -e cache-misses,cache-references,LLC-loads,LLC-load-misses,LLC-stores,LLC-store-misses taskset -c 0 python test-model.py 1 > logs/test-model/$identification/3-single-core_cpu0.txt &
-perf stat -o logs/test-model/$identification/3-single-core_cpu1_perf.txt -e cache-misses,cache-references,LLC-loads,LLC-load-misses,LLC-stores,LLC-store-misses taskset -c 1 python test-model.py 1 > logs/test-model/$identification/3-single-core_cpu1.txt &
-perf stat -o logs/test-model/$identification/3-single-core_cpu2_perf.txt -e cache-misses,cache-references,LLC-loads,LLC-load-misses,LLC-stores,LLC-store-misses taskset -c 2 python test-model.py 1 > logs/test-model/$identification/3-single-core_cpu2.txt &
+perf stat -o logs/test-model/$identification/3-single-core_cpu1_perf.txt -e cache-misses,cache-references,LLC-loads,LLC-load-misses,LLC-stores,LLC-store-misses taskset -c 1 python test-model2.py 1 > logs/test-model/$identification/3-single-core_cpu1.txt &
[…]

### 2018-01-16 test-model_timings_Pi_RT.sh

@@ -1,0 +1,68 @@
+timestamp=$(date +%y-%m-%d_%H-%M-%S)
+
+echo "Which board is this test being run on?  Input a string and press [ENTER]"
+read board
+
+identification=$board
+identification+="_"
+identification+=$timestamp
+
+mkdir -p logs/test-model/$identification
+echo "This test will take approximately 45 minutes to complete"
+echo "The output files can be found in: logs/test-model/$identification"
+
+echo single-core
+sudo chrt -f 99 perf stat -o logs/test-model/$identification/single-core_perf.txt -e cache-misses,cache-references,LLC-loads,LLC-load-misses,LLC-stores,LLC-store-misses taskset -c 0 python test-model.py 1 > logs/test-model/$identification/single-core.txt &
+wait
+echo end single-core
+sleep 300
+
+echo 2-parallel-single-core
+sudo chrt -f 99 perf stat -o logs/test-model/$identification/2-single-core_cpu0_perf.txt -e cache-misses,cache-references,LLC-loads,LLC-load-misses,LLC-stores,LLC-store-misses taskset -c 0 python test-model.py 1 > logs/test-model/$identification/2-single-core_cpu0.txt &
+sudo chrt -f 99 perf stat -o logs/test-model/$identification/2-single-core_cpu1_perf.txt -e cache-misses,cache-references,LLC-loads,LLC-load-misses,LLC-stores,LLC-store-misses taskset -c 1 python test-model2.py 1 > logs/test-model/$identification/2-single-core_cpu1.txt &
+wait
+echo end 2-parallel-single-core
+sleep 300
+
+echo 3-parallel-single-core
+sudo chrt -f 99 perf stat -o logs/test-model/$identification/3-single-core_cpu0_perf.txt -e cache-misses,cache-references,LLC-loads,LLC-load-misses,LLC-stores,LLC-store-misses taskset -c 0 python test-model.py 1 > logs/test-model/$identification/3-single-core_cpu0.txt &
+sudo chrt -f 99 perf stat -o logs/test-model/$identification/3-single-core_cpu1_perf.txt -e cache-misses,cache-references,LLC-loads,LLC-load-misses,LLC-stores,LLC-store-misses taskset -c 1 python test-model2.py 1 > logs/test-model/$identification/3-single-core_cpu1.txt &
[…]

### 2018-01-16 run.py
use half cores for rendering

@@ -35,16 +35,16 @@ epoch_ids = sorted(list(set(itertools.chain(*params.epochs.values()))))
 
 def process_epoch(epoch_id):
     print '---------- processing video for epoch {} ----------'.format(epoch_id)
-    # vid_path = cm.jn(params.data_dir, 'epoch{:0>2}_front.mkv'.format(epoch_id))
-
+    vid_path = cm.jn(params.data_dir, 'out-video-{}.avi'.format(epoch_id))
+    frame_count = cm.frame_count(vid_path)        
+    
     vid_scaled_path = cm.jn(params.data_dir, 'out-video-{}-scaled.avi'.format(epoch_id))
     if not os.path.exists(vid_scaled_path):
-        vid_path = cm.jn(params.data_dir, 'out-video-{}.avi'.format(epoch_id))
         assert os.path.isfile(vid_path)
         os.system("ffmpeg -i " + vid_path + " -vf scale=1280:720 " + vid_scaled_path)
         print("ffmpeg -i " + vid_path + " -vf scale=1280:720 " + vid_scaled_path)
-    frame_count = cm.frame_count(vid_path)
     vid_path = vid_scaled_path
+    
     cap = cv2.VideoCapture(vid_path)
 
     machine_steering = []
@@ -89,7 +89,7 @@ for epoch_id in epoch_ids:
     job_list.append([epoch_id, steering])
 
 print "visualize epochs in parallel"    
-num_cores = 4 # multiprocessing.cpu_count()
+num_cores = int(multiprocessing.cpu_count()/2)
 Parallel (n_jobs = num_cores) (delayed (visualize_epoch) (job) for job in job_list)
 
      

### 2018-01-16 train.py
check stored model file

@@ -26,9 +26,12 @@ train_step = tf.train.AdamOptimizer(1e-4).minimize(loss)
 saver = tf.train.Saver()
 
 model_load_path = cm.jn(params.save_dir, params.model_load_file)
-if os.path.exists(model_load_path):
+print model_load_path
+if os.path.exists(model_load_path + ".index"):
+    print ("Loading initial weights from %s" % model_load_path)
     saver.restore(sess, model_load_path)
 else:
+    print ("Initialize weights.")
     sess.run(tf.global_variables_initializer())
 
 # create a summary to monitor cost tensor

### 2018-01-16 data_shuffled.py
bug fix?

@@ -47,12 +47,13 @@ def load_imgs():
             # vid_path = cm.jn(data_dir, 'epoch{:0>2}_front.mkv'.format(epoch_id))
             vid_path = cm.jn(data_dir, 'out-video-{}.avi'.format(epoch_id))
 
+            print "DBG:", vid_path
             assert os.path.isfile(vid_path)
 
-            print "DBG:", vid_path
             frame_count = cm.frame_count(vid_path)
-            cap = cv2.VideoCapture(vid_path)
             print "DBG:", frame_count
+
+            cap = cv2.VideoCapture(vid_path)
 
             # csv_path = cm.jn(data_dir, 'epoch{:0>2}_steering.csv'.format(epoch_id))
             csv_path = cm.jn(data_dir, 'out-key-{}.csv'.format(epoch_id))

### 2018-01-16 params.py
..

@@ -18,8 +18,8 @@ use_picar_mini = True
 # change this to the directory that contains the source videos
 # "-5conv_3fc_noreuse" "-5conv_3fc" "-3conv_1pool" "-5conv_3fc_noreuse_nocn"
 model="model-5conv_3fc"
-model_load_file="model-5conv_3fc_noreuse.ckpt"
-model_save_file="model-5conv_3fc_noreuse.ckpt"
+model_save_file="model-5conv_3fc-home_night.ckpt"
+model_load_file=model_save_file 
 save_dir = os.path.abspath('models')
 data_dir = os.path.abspath('epochs-conv')
 out_dir = os.path.abspath('output')
@@ -30,8 +30,8 @@ if not os.path.isdir(out_dir):
     os.makedirs(out_dir)
 
 epochs = OrderedDict()
-epochs['train'] = range(30, 50) + range(60, 68) 
-epochs['val'] = range(50, 60) + range(68, 70)
+epochs['train'] = range(106, 110) # range(30, 50) + range(60, 68) 
+epochs['val']   = range(110, 112) # range(50, 60) + range(68, 70)
 
 # data description
 # range(30, 60)        # ittc building. michael. resynchronized

### 2018-01-17 palloc-setup.sh
Add palloc figures and scripts
Also added palloc-setup.sh, which is used to enable PALLOC.

@@ -1,0 +1,5 @@
+echo 0x00003000 > /sys/kernel/debug/palloc/palloc_mask
+cgcreate -g palloc:part1
+echo 0 > /sys/fs/cgroup/palloc/part1/palloc.bins
+echo 1 > /sys/kernel/debug/palloc/use_palloc
+echo 2 > /sys/kernel/debug/palloc/debug_level

### 2018-01-24 paper/conclusion.tex
Add initial cache partition section
Also made minor changes in other sections of the paper, and added the
script used to max performance on the TX2.

--- 2017-10-16 paper/conclusion.tex
@@
 
-As future work, we plan to apply shared resource management
+As future work, we plan to apply shared DRAM resource management
 techniques~\cite{Yun2013,yun2014rtas} on the DeepPicar platform and

### 2018-01-24 paper/evaluation.tex
Change figure sizes
Many of the figures were relatively small and hard to see, so the sizes
were increased.
Add initial cache partition section
Also made minor changes in other sections of the paper, and added the
script used to max performance on the TX2.

@@
   \centering
-  \includegraphics[width=.5\textwidth]{figs/Fig7_new}
+  \includegraphics[width=.75\textwidth]{figs/Fig7_new}
   \caption{DeepPicar's control loop processing times over 1000 input image frames.}
@@
   \centering
-  \includegraphics[width=.45\textwidth]{figs/perf_vs_corecnt}
+  \includegraphics[width=.7\textwidth]{figs/perf_vs_corecnt}
   \caption{Average control loop execution time vs. \#of CPU
@@
   \centering
-  \includegraphics[width=.45\textwidth]{figs/perf_vs_modelcnt}
+  \includegraphics[width=.7\textwidth]{figs/perf_vs_modelcnt}
   \caption{Timing impact of co-scheduling multiple DNNs. 1Nx1C: one DNN
@@
   \centering
-  \includegraphics[width=.45\textwidth]{figs/l2missrate_vs_modelcnt}
+  \includegraphics[width=.7\textwidth]{figs/l2missrate_vs_modelcnt}
   \caption{L2 cache miss rates of different neural network and core
@@
   \centering
-  \includegraphics[width=.45\textwidth]{figs/perf_vs_bandwidth}
+  \includegraphics[width=.7\textwidth]{figs/perf_vs_bandwidth}
   \caption{Effect of memory performance hogs on the DNN
@@
 
-\begin{table*}[h]
+\subsection{Effect of Cache Partitioning}
+
+In order to determine if the DNN is mostly affected by shared DRAM 
+contention, we partition the L2 cache of the Pi3 and reperform the 
+same experiments to see if there are any noticable changes. For the 
+partition, we employ PALLOC~\cite{yun2014rtas}, a color-based page 
+allocator that works on the kernel level. For the bit mask, we select 
+bits 12, 13 and 14 as those bits are used to specifically access the 
+shared L2 cache, as can be seen by \ref{fig:cache-mapping}. This 
+results in 2\textsuperscript{3} = 8 colors, which we then assign to the Pi3's 
+physical cores such that each core has two unique colors (colors 0 
+and 1 are assigned to core 0, colors 2 and 3 are assigned to core 1, 
+etc.).
+
+\begin{figure}[h]
+  \centering
+  \includegraphics[width=\textwidth]{figs/cache-mapping}
+  \caption{A memory mapping of the cache sets for the Broadcom 
+BCM2837 processor used in the Raspberry Pi 3 Model B.}
+  \label{fig:cache-mapping}
+\end{figure}
+
+If performance improves as a result of partitioning the shared cache 
+then we know that the DNN, to some extent, relies on shared cache 
+performance in addition to shared DRAM performance. On the other 
+hand, if there is no improvement in performance, then it can be 
+observed that the shared memory is more critical to the performance 
+of the DNN.
+
+\begin{figure}[h]
+  \centering
+  \includegraphics[width=.7\textwidth]{figs/palloc_multicore}
+  \caption{Average control loop execution time vs. \#of CPU
+    cores with cache partitioning enabled.}
+  \label{fig:palloc_multicore}
+\end{figure}
+
+\begin{figure}[h]
+  \centering
+  \includegraphics[width=.7\textwidth]{figs/palloc_multicore}
+  \caption{Timing impact of co-scheduling multiple DNNs when cache 
+partitioning is enabled.}
+  \label{fig:palloc_multimodel}
+\end{figure}
+
+The results can be seen in Figures \ref{fig:palloc_multicore} - 
+\ref{fig:palloc_bandwidth_l2missrate}. As 
+expected, based off of previous results, we find that the performance 
+of the DNN remains consistent when cache partitioning is employed. 
+When fewer cores are used, DNN performance is slightly affected as 
+the entire L2 cache isn't available to the network. In the case of 
+running multiple models simultaneously, no performance improvements 
+are seen, which further shows that the shared cache has no affect on 
+the DNN's real-time performance.
+
+\begin{figure}[h]
+  \centering
+  \includegraphics[width=.7\textwidth]{figs/palloc_bandwidth_exectime}
+  \caption{Timing impact of co-scheduling memory intensive co-runners 
+when cache partitioning is enabled.}
+  \label{fig:palloc_bandwidth_exectime}
+\end{figure}
+
+In the case of running the DNN alongside synthetic memory-intensive 
+benchmarks, however, we found that partitioning the L2 cache led to a 
+decrease in performance as the number of co-runners increased. 
+However, the L2 miss rates of the model decrease as more benchmarks 
+are introduced. As such, this means that the model encounters greater 
+interence from the memory contention of the co-runners.
+
+\begin{figure}[h]
+  \centering
+  \includegraphics[width=.7\textwidth]{figs/palloc_bandwidth_l2missrate}
+  \caption{L2 miss rate impact of co-scheduling memory intensive co-
+runners when cache partitioning is enabled.}
+  \label{fig:palloc_bandwidth_l2missrate}
+\end{figure}
+
+By partitioning the shared L2 cache, we find that no noticable 
+improvements are gained and that performance remains consistent, or 
+decreases, depending on the circumstances. As a result, we find that 
+cache partitioning is not an effective isolation mechanism, and that 
+the performance of the shared DRAM is of greater importance to the 
+real-time efficiacy of the DeepPicar.
+
+\begin{table*}[!h]
   \centering
@@
     Memory  & 1GB LPDDR2   &  2GB DDR3L     & 8GB LPDDR4              \\
+    \hline
+	Peak Memory Bandwidth & 8.5 GB/s & 12.8 GB/s & 59.7 GB/s \\
+	\hline
+	Measured Bandwidth & 2127.94 MB/s & 3951.94 MB/s & 4447.90 MB/s \\
 	\hline
@@
 
-Lastly, we find that consolidating memory (DRAM) performance
+Fourth, we find that consolidating memory (DRAM) performance
 intensive applications could jeopadize DNN performance, because DNN
-performance appears to be very sensitive to memory performance; we observe up
-to 9.6X slowdown in DNN performance by co-scheduling synthetic memory 
-bandwidth intensive applications on idle cores.
+performance appears to be very sensitive to memory performance; we 
+observe up to 11.6X slowdown in DNN performance by co-scheduling 
+synthetic memory bandwidth intensive applications on idle cores.
+
+Lastly, we find that partitioning the shared L2 cache provides 
+practically no benefits to the real-time performance of the DNN. To the 
+contrary, performance is negatively affected when fewer cores are used, 
+due to less cache being available, and memory intensive co-runners are 
+introduced, as they generate more memory contention. Furthermore, we 
+find that performance remains consistent when multiple models are run 
+simultaneously. As such, we conclude that the shared L2 cache is not 
+important to the DNN's real-time performance.

### 2018-01-24 paper/intro.tex
Add initial cache partition section
Also made minor changes in other sections of the paper, and added the
script used to max performance on the TX2.

@@
 We find that the DeepPicar is capable of 
-completing control loops in under 50 ms, or 20 hz, and can do so 
-almost 99\% of the time.
+completing control loops in under 33.$\overline{\mbox{33}}$ 
+ms, or 30 hz, and can do so 100\% of the time.
 Other tested embedded platforms, Intel Up and NVIDIA TX2, offer even
@@
 However, in all platforms, shared resource contention remains an
-important issue as we observe up to 9.6X control loop execution time
+important issue as we observe up to 11.6X control loop execution time
 increase, mostly due to increase in the nueral network inferencing

### 2018-01-25 paper/overview.tex
Change figure sizes
Many of the figures were relatively small and hard to see, so the sizes
were increased.
Change section on camera latency

--- 2017-10-16 paper/overview.tex
@@
   \centering
-  \includegraphics[width=.4\textwidth]{figs/DeepPicar_platform}
+  \includegraphics[width=.55\textwidth]{figs/DeepPicar_platform}
   \caption{DeepPicar platform.}
@@
   \centering
-  \includegraphics[width=.4\textwidth]{figs/architecture}
+  \includegraphics[width=.6\textwidth]{figs/architecture}
   \caption{DeepPicar's neural network architecture: 9 layers (5
@@
   \centering
-  \includegraphics[width=.4\textwidth]{figs/track_new2}
+  \includegraphics[width=.65\textwidth]{figs/track_new2}
   \caption{One of the custom tracks used for training/testing.}
@@
 sensor observes the scene to the time the computer actually reads the
-digitized image data. Unfortunately, this time can be significantly
-long depending on the camera and the performance of the Pi. We
-experimentally measure DeepPicar's camera latency as
-follows~\footnote{We follow the method and the tool found at
-  https://www.makehardware.com/2016/03/29/finding-a-low-latency-webcam/}.
-First, we point the DeepPicar's camera towards a computer monitor, which is
-conntected to the Raspberry Pi 3's HDMI output. We then launch a 
-program that overlays the current time with the camera's preview
-screen on the monitor. Then, an infinite number of time strings will be
-shown on the monitor screen and the time difference between two
-successive time strings represent the camera
-latency. Figure~\ref{fig:camera-latency} shows the captured moniotor
-screen. The measured camera latency is about 300-350 ms. This is
-significantly higher than the latency of human perception, which is known
-to be as fast as 13 ms~\cite{ThomasBurger2015}. Higher camera latency
-could negatively affect to the control performance, because the DNN would
-analyze stale scenes. In the future, we plan to identify and use
-low-latency cameras.
-
-\begin{figure}[t]
-  \centering
-  \includegraphics[width=.4\textwidth]{figs/camera_latency}
-  \caption{Measuring DeepPicar's camera latency.}
-  \label{fig:camera-latency}
-\end{figure}
+digitized image data. This time can be noticable depending on the camera 
+used and the performance of the Pi. We experimentally measure the camera
+latency and found it to be around 50-100 ms. This is higher than the 
+latency of human perception, which is known to be as fast as 13 
+ms~\cite{ThomasBurger2015}. Higher camera latency could negatively 
+affect control performance, because the DNN would analyze stale scenes. 
+In the future, we plan to identify and use low-latency cameras.
 

### 2018-01-25 paper/evaluation.tex
Update cache mapping and relevant section

@@
 allocator that works on the kernel level. For the bit mask, we select 
-bits 12, 13 and 14 as those bits are used to specifically access the 
-shared L2 cache, as can be seen by \ref{fig:cache-mapping}. This 
-results in 2\textsuperscript{3} = 8 colors, which we then assign to the Pi3's 
+bits 12, 13 and 14 as they can be used to access the L2 cache,
+as can be seen by \ref{fig:cache-mapping}. This results in 
+2\textsuperscript{3} = 8 colors, which we then assign to the Pi3's 
 physical cores such that each core has two unique colors (colors 0 
 and 1 are assigned to core 0, colors 2 and 3 are assigned to core 1, 
-etc.).
+etc.). In the case of bit 12, since we use 2 colors for each partition,
+the L1D cache is not partitioned in our experiments.
 
@@
 running multiple models simultaneously, no performance improvements 
-are seen, which further shows that the shared cache has no affect on 
+are seen, which further shows that the shared cache has little affect on 
 the DNN's real-time performance.
@@
 improvements are gained and that performance remains consistent, or 
-decreases, depending on the circumstances. As a result, we find that 
+decreases, depending on the circumstances. As a result, we conclude that 
 cache partitioning is not an effective isolation mechanism, and that 
@@
 simultaneously. As such, we conclude that the shared L2 cache is not 
-important to the DNN's real-time performance.
+essential to the DNN's real-time performance.

### 2018-01-25 input_kbd.py
zero waiting on select

@@ -37,7 +37,7 @@ def stop(state):
     
 def read_single_keypress():
     state = init()    
-    r, w, e = select.select([sys.stdin], [], [], 0.001)
+    r, w, e = select.select([sys.stdin], [], [], 0.000)
     input = ' '
     for s in r:
         if s == sys.stdin:
@@ -52,7 +52,7 @@ if __name__ == "__main__":
         key = read_single_keypress()
         if key == "q":
             break
-        elif key != None:
+        elif key != ' ':
             print ("%s pressed\r" % key)
         
 

### 2018-01-25 params.py
camera-webcam option description

@@ -6,6 +6,7 @@ from collections import OrderedDict
 
 ##########################################################
 # camera module selection
+#   "camera-webcam"
 ##########################################################
 camera="camera-webcam"
 

### 2018-01-25 actuator-drv8835.py
drv8835 test

@@ -1,15 +1,14 @@
 from pololu_drv8835_rpi import motors, MAX_SPEED
 
 # init
-def init():
+def init(default_speed=50):
     motors.setSpeeds(0, 0)
+    set_speed(default_speed)
 
 # throttle
 cur_speed = MAX_SPEED
 
 def set_speed(speed):
-
-
     global cur_speed
     speed = int(MAX_SPEED * speed / 100)
     cur_speed = min(MAX_SPEED, speed)

### 2018-01-25 input_kbd.py

@@ -36,16 +36,18 @@ def stop(state):
     fcntl.fcntl(fd, fcntl.F_SETFL, state[0])
     
 def read_single_keypress():
+    state = init()    
     r, w, e = select.select([sys.stdin], [], [], 0.001)
+    input = ' '
     for s in r:
         if s == sys.stdin:
             input = sys.stdin.read(1)
-            return input
-    return None
+            break
+    stop(state)
+    return input
 
 if __name__ == "__main__":
-    state = init()
-    
+
     while True:
         key = read_single_keypress()
         if key == "q":
@@ -53,5 +55,5 @@ if __name__ == "__main__":
         elif key != None:
             print ("%s pressed\r" % key)
         
-    stop(state)
+
     

### 2018-01-25 params.py

@@ -13,7 +13,7 @@ camera="camera-webcam"
 # actuator selection
 #   "actuator-drv8835", "actuator-adafruit_hat"
 ##########################################################
-actuator="actuator-adafruit_hat"
+actuator="actuator-drv8835"
 
 ##########################################################
 # model selection

### 2018-01-25 picar-mini-kbd-common.py

@@ -52,7 +52,6 @@ def g_tick():
 def turn_off():    
     actuator.stop()
     camera.stop()
-    input_kbd.stop()
     
     keyfile.close()
     keyfile_btn.close()
@@ -97,9 +96,6 @@ if use_dnn == True:
     model_load_path = cm.jn(params.save_dir, params.model_load_file)
     saver.restore(sess, model_load_path)
     print ("Done..")
-
-# initialize kbd input 
-input_kbd.init()
 
 # initlaize deeppicar modules
 actuator.init(cfg_throttle)

### 2018-01-26 paper/main.tex
Update template and abstract

@@
 \usepackage{microtype}%if unwanted, comment out or use option "draft"
+\usepackage{adjustbox}
 
@@
 
-\subjclass{Dummy classification -- please refer to \url{http://www.acm.org/about/class/ccs98-html}}% mandatory: Please choose ACM 1998 classifications from http://www.acm.org/about/class/ccs98-html .
-E.g., cite as "F.1.1 Models of Computation".
-\keywords{Dummy keyword -- please provide 1--5 keywords}% mandatory: Please provide 1-5 keywords
+\subjclass{C.4 PERFORMANCE OF SYSTEMS}% mandatory: Please choose ACM 1998 classifications from http://www.acm.org/about/class/ccs98-html . E.g., cite as "F.1.1 Models of Computation". 
+\keywords{Self-driving car, CNN, Raspberry Pi 3}% mandatory: Please provide 1-5 keywords
 % Author macros::end %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 car steering angles as output. DeepPicar uses the exact same network
-architecture---9 layers, 27 million connections and 250K
+architecture---8 layers, 27 million connections and 250K
 parameters---and can be trained to drive itself, in real-time, using a

### 2018-01-26 paper/intro.tex
Update introduction

@@
 value as output at each control period in \emph{real-time}.
-The network has 9 layers, about 27 million connections
+The network has 8 layers, about 27 million connections
 and 250 thousand parameters (weights).
@@
 real-time. For real-time control, each inference operation must
-be completed within the desired control period (e.g., 50 ms period for
-20 Hz control frequency).
+be completed within the desired control period (e.g., 33.$\overline{\mbox{33}}$ ms
+ period for 30 Hz control frequency).
 % how we evaluated (in terms real-time performance)
@@
 ms, or 30 hz, and can do so 100\% of the time.
-Other tested embedded platforms, Intel Up and NVIDIA TX2, offer even
-better performance, capable of supporting deep-learning based real-time
-control up to 100 Hz control frequency on the TX2 when the GPU is used.
-However, in all platforms, shared resource contention remains an
-important issue as we observe up to 11.6X control loop execution time
-increase, mostly due to increase in the nueral network inferencing
+Other tested embedded platforms, Intel UP and NVIDIA TX2, offer even
+better performance, and are capable of supporting deep-learning based 
+real-time control up to 100 Hz control frequency on the TX2 when the GPU 
+is used. However, in all platforms, shared resource contention remains 
+an important issue as we observe up to 11.6X control loop execution time
+increase, mostly due to increase in the neural network inferencing
 operation, when memory performance intensive applications are

### 2018-01-26 paper/overview.tex
Update platform overview

@@
 NVIDIA's DAVE-2 system on a smaller scale with using a low cost
-multicore platform, Raspberry Pi 3. Because the Pi 3's computing
+multicore platform, the Raspberry Pi 3. Because the Pi 3's computing
 performance is much lower than that of the DRIVE PX platform used in
@@
 operations must be completed within each control period
-duration---e.g., a WCET of 50ms for 20Hz control frequency---locally on
-the Pi 3 platform, although training of the network (back-propagation
-for weight updates) can be done offline and remotely using a desktop
-computer.
+duration---e.g., a WCET of 33.$\overline{\mbox{33}}$ ms for 30Hz control 
+frequency---locally on the Pi 3 platform, although training of the 
+network (back-propagation for weight updates) can be done offline and 
+remotely using a desktop computer.
 
@@
 output. Figure~\ref{fig:architecture} shows the network architecture, which
-is comprised of 9 layers, 250K parameters, and about 27 million
+is comprised of 8 layers, 250K parameters, and about 27 million
 connections.
@@
   \centering
-  \includegraphics[width=.6\textwidth]{figs/architecture}
-  \caption{DeepPicar's neural network architecture: 9 layers (5
-    convolutional, 4 fully-connected layers), 27 million connections,
+  \includegraphics[width=\textwidth]{figs/architecture}
+  \caption{DeepPicar's neural network architecture: 8 layers (5
+    convolutional, 3 fully-connected layers), 27 million connections,
     250K parameters. The architecture is identical to the one used
@@
   \label{fig:architecture}
+\end{figure}
+
+\begin{figure}[t]
+  \centering
+  \includegraphics[width=.65\textwidth]{figs/track_new2}
+  \caption{One of the custom tracks used for training/testing.}
+  \label{fig:track}
+\end{figure}
+
+\begin{figure}[t]
+   \lstset{language=python,
+           basicstyle=\ttfamily\small,
+           keywordstyle=\color{blue}\ttfamily,
+           stringstyle=\color{red}\ttfamily,
+           commentstyle=\color{green}\ttfamily
+          }  
+  \lstinputlisting[language=python]{control.py}
+  \caption{Control loop}
+  \label{fig:controlloop}
 \end{figure}
@@
 
-\begin{figure}[t]
-  \centering
-  \includegraphics[width=.65\textwidth]{figs/track_new2}
-  \caption{One of the custom tracks used for training/testing.}
-  \label{fig:track}
-\end{figure}
-
 % inferencing on pi3
@@
 execution time.
-
-\begin{figure}[t]
-   \lstset{language=python,
-           basicstyle=\ttfamily\small,
-           keywordstyle=\color{blue}\ttfamily,
-           stringstyle=\color{red}\ttfamily,
-           commentstyle=\color{green}\ttfamily
-          }  
-  \lstinputlisting[language=python]{control.py}
-  \caption{Control loop}
-  \label{fig:controlloop}
-\end{figure}
 
@@
 digitized image data. This time can be noticable depending on the camera 
-used and the performance of the Pi. We experimentally measure the camera
+used and the performance of the Pi. We experimentally measured the camera
 latency and found it to be around 50-100 ms. This is higher than the 

### 2018-01-26 paper/evaluation.tex
Update evaluation section

@@
 (4) the effect of co-scheduling memory bandwidth intensive co-runners.
+(5) the effect of partitioning the shared L2 cache.
 
@@
 BCM2837 SoC, which has a quad-core ARM Cortex-A53 cluster,
-running at up to 1.2GHz. Each core has 16K private I\&D caches, and all
-cores share a 512KB L2 cache.
+running at up to 1.2GHz. Each core has a 16K private Instruction cache
+and a 32K private Data cache, and all cores share a 512KB L2 cache.
 The chip also includes Broadcom's Videocore IV
@@
     \textbf{Operation} & \textbf{Mean} & \textbf{Max} &   \textbf{99pct.} & \textbf{Stdev.} \\ \hline
-    Image capture        & 2.28  &  4.94 &  4.54  & 0.52 \\ \hline
-    Image pre-processing & 3.09  &  4.60 &  3.31  & 0.10 \\ \hline
-    DNN inferencing      & 37.30 & 51.03 & 45.48  & 2.75 \\ \hline
-    Total Time           & 42.67 & 56.37 & 50.70  & 2.80 \\ \hline
+    Image capture        & 1.61  &  1.81 &  1.75  & 0.05 \\ \hline
+    Image pre-processing & 2.77  &  2.90 &  2.87  & 0.04 \\ \hline
+    DNN inferencing      & 18.49 & 19.30 & 18.99  & 0.20 \\ \hline
+    Total Time           & 22.87 & 23.74 & 23.38  & 0.20 \\ \hline
   \end{tabular}
@@
 First, as expected, we find that the inference operation
-dominates the control loop execution time, accounting for about 85\% of
+dominates the control loop execution time, accounting for about 81\% of
 the execution time.
 Second, and more importantly, we also find that the measured average
-execution time of a single control loop is 42.67 ms, or 23.4 Hz and
-the 99 percentile time is 50.70 ms.
+execution time of a single control loop is 22.87 ms, or 43.7 Hz and
+the 99 percentile time is 23.38 ms.
 This means that the DeepPicar can operate
-at about a 20 Hz control frequency in real-time using only the on-board
+at a 30 Hz control frequency in real-time using only the on-board
 Raspberry Pi 3 computing platform, as no remote computing resources 
-were necessary. We consider these results respectable given the complexity
+were necessary. We consider these results surprising given the complexity
 of the deep neural network, and the fact that the inference operation
@@
 In comparison, NVIDIA's DAVE-2 system, which has the exact same neural
-network architecture, reportedly runs at 30 Hz~\cite{Bojarski2016}, which 
-is just a bit faster than the DeepPicar. Although we believe it was not
+network architecture, reportedly runs also at 30 Hz~\cite{Bojarski2016}. 
+Although we believe it was not
 limited by their computing platform (we will experimentally compare
@@
 \subsection{Effect of the Core Count to Inference Timing}
+
+\begin{figure}[h]
+  \centering
+  \includegraphics[width=.7\textwidth]{figs/perf_vs_corecnt}
+  \caption{Average control loop execution time vs. \#of CPU
+    cores.}
+  \label{fig:perf-vs-corecnt}
+\end{figure}
 
@@
 
-\begin{figure}[h]
-  \centering
-  \includegraphics[width=.7\textwidth]{figs/perf_vs_corecnt}
-  \caption{Average control loop execution time vs. \#of CPU
-    cores.}
-  \label{fig:perf-vs-corecnt}
-\end{figure}
-
 Figure~\ref{fig:perf-vs-corecnt} shows the average execution time of
@@
 TensorFlow. As expected, as we assign more cores, the average execution
-time decreases---from 61.96 ms on a single core to 42.67 ms on four
-cores (a 30\% improvement). However, the improvement is far from an ideal
-linear scaling. In particular, from 2 cores to 3 cores, the
-improvement is mere 2.38 ms (or 4\%). In short, we find that the
+time decreases---from 46.30 ms on a single core to 22.86 ms on four
+cores (over a 100\% improvement). However, the improvement is far from an ideal
+linear scaling. In particular, from 3 cores to 4 cores, the
+improvement is mere 2.80 ms, or 12%. In short, we find that the
 scalability of DeepPicar's deep neural network is not ideal on the
@@
 architecture for all models, then one might expect to achieve up to
-15 Hz using one core (given 1 core can deliver 62 ms average
+20 Hz using one core (given 1 core can deliver 46 ms average
 execution time). In the next experiment, we investigate the
@@
 (1Nx1C), when four DNN models are co-scheduled (4Nx1C), each model
-suffers an average inference time increase of approximately 15 ms,
+suffers an average inference time increase of approximately 11 ms,
 $\sim$24\%. On the other hand, when two DNN models, each using two CPU
 cores, are co-scheduled (2Nx2C), the average inference timing is increased by
-about 7 ms, or 10\%, compared to the baseline of running one model
+about 4 ms, or 13\%, compared to the baseline of running one model
 using two CPU cores (1Nx2C).
@@
 increase. Figure~\ref{fig:l2missrate-vs-modelcnt} shows the results.
-As can be see in the figure, L2 miss rates are largely unchanged
-regardless of whether multiple models are co-scheduled or not. This
-suggests that the shared L2 cache is not the bottleneck that caused
-execution time increases. In other words, DNN models don't appear to be
-sensitive to the shared L2 cache space. % L2 partitioning is probably
-                                % not going to be useful.
+As can be see in the figure, L2 miss rates increase as more models are 
+co-scheduled, but the increase is less than expected.
+This suggests that the shared L2 cache is not the main bottleneck that 
+causes execution time increases. In other words, DNN models don't appear 
+to be significantly affected by the shared L2 cache space. 
+% L2 partitioning is probably not going to be useful.
 Instead, we hypothesize that it is likely caused by the memory
@@
 
-We run a single DNN model one core, and
+We run a single DNN model on one core, and
 co-schedule an increasing number of the memory intensive synthetic
@@
 increase the number of co-runners, the DNN model's execution times are
-increased---by up to 9.6X---even though the DNN model is running on a
+increased---by up to 11.6X---even though the DNN model is running on a
 dedicated core (Core 0). On the other hand, the DNN model's L2
 cache-miss rates do not increase as much. This suggests that
-the DNN model's exeuction increase cannot be fully explained by
+the DNN model's execution increase cannot be fully explained by
 increases in L2 cache-misses. Instead, as we hypothesized in the previous
@@
 
-In order to determine if the DNN is mostly affected by shared DRAM 
-contention, we partition the L2 cache of the Pi3 and reperform the 
-same experiments to see if there are any noticable changes. For the 
-partition, we employ PALLOC~\cite{yun2014rtas}, a color-based page 
-allocator that works on the kernel level. For the bit mask, we select 
-bits 12, 13 and 14 as they can be used to access the L2 cache,
-as can be seen by \ref{fig:cache-mapping}. This results in 
-2\textsuperscript{3} = 8 colors, which we then assign to the Pi3's 
-physical cores such that each core has two unique colors (colors 0 
-and 1 are assigned to core 0, colors 2 and 3 are assigned to core 1, 
-etc.). In the case of bit 12, since we use 2 colors for each partition,
-the L1D cache is not partitioned in our experiments.
-
 \begin{figure}[h]
@@
 \end{figure}
-
-If performance improves as a result of partitioning the shared cache 
-then we know that the DNN, to some extent, relies on shared cache 
-performance in addition to shared DRAM performance. On the other 
-hand, if there is no improvement in performance, then it can be 
-observed that the shared memory is more critical to the performance 
-of the DNN.
 
@@
   \centering
-  \includegraphics[width=.7\textwidth]{figs/palloc_multicore}
+  \includegraphics[width=.7\textwidth]{figs/palloc_multimodel}
   \caption{Timing impact of co-scheduling multiple DNNs when cache 
@@
 \end{figure}
+
+\begin{figure}[h]
+  \centering
+  \includegraphics[width=.7\textwidth]{figs/palloc_bandwidth_exectime}
+  \caption{Timing impact of co-scheduling memory intensive co-runners 
+when cache partitioning is enabled.}
+  \label{fig:palloc_bandwidth_exectime}
+\end{figure}
+
+\begin{figure}[h]
+  \centering
+  \includegraphics[width=.7\textwidth]{figs/palloc_bandwidth_l2missrate}
+  \caption{L2 miss rate impact of co-scheduling memory intensive co-
+runners when cache partitioning is enabled.}
+  \label{fig:palloc_bandwidth_l2missrate}
+\end{figure}
+
+In order to determine if the DNN is mostly affected by the DRAM 
+controller, we partition the L2 cache of the Pi3 and reperform the 
+same experiments to see if there are any noticable changes. For the 
+partition, we employ PALLOC~\cite{yun2014rtas}, a color-based page 
+allocator that works on the kernel level. For the bit mask, we select 
+bits 12, 13 and 14 as they can be used to access the L2 cache,
+as can be seen by Figure \ref{fig:cache-mapping}. This results in 
+2\textsuperscript{3} = 8 colors, which we then assign to the Pi3's 
+physical cores such that each core has two unique colors (colors 0 
+and 1 are assigned to core 0, colors 2 and 3 are assigned to core 1, 
+etc.). In the case of bit 12, since we use 2 colors for each partition,
+the L1 Data cache is not partitioned in our experiments.
+
+If performance improves as a result of partitioning the shared cache 
+then we know that the DNN, to some extent, relies on shared cache 
+performance in addition to shared DRAM performance. On the other 
+hand, if there is no improvement in performance, then it can be 
+observed that the shared memory is more critical to the performance 
+of the DNN.
 
@@
 
-\begin{figure}[h]
-  \centering
-  \includegraphics[width=.7\textwidth]{figs/palloc_bandwidth_exectime}
-  \caption{Timing impact of co-scheduling memory intensive co-runners 
-when cache partitioning is enabled.}
-  \label{fig:palloc_bandwidth_exectime}
-\end{figure}
-
 In the case of running the DNN alongside synthetic memory-intensive 
@@
 
-\begin{figure}[h]
-  \centering
-  \includegraphics[width=.7\textwidth]{figs/palloc_bandwidth_l2missrate}
-  \caption{L2 miss rate impact of co-scheduling memory intensive co-
-runners when cache partitioning is enabled.}
-  \label{fig:palloc_bandwidth_l2missrate}
-\end{figure}
-
 By partitioning the shared L2 cache, we find that no noticable 
@@
 cache partitioning is not an effective isolation mechanism, and that 
-the performance of the shared DRAM is of greater importance to the 
-real-time efficiacy of the DeepPicar.
-
-\begin{table*}[!h]
-  \centering
-  \begin{tabular}{|c|c|c|c|}
-    \hline
-    Item    & Raspberry Pi 3 (B)   & Intel UP                  & NVIDIA Jetson TX2\\
-    \hline
-            & BCM2837              & X5-Z8350 (Cherry Trail)   & Tegra X2 \\
-    CPU     & 4x Cortex-A53@1.2GHz/512KB L2  &
-              4x Atom@1.92GHz/2MB L2 &
-              4x Cortex-A57@2GHz/2MB L2 \\
-            &              &              & 2x Denver@2.0GHz/2MB L2 (not used)  \\
-    \hline
-    GPU     &  VideoCore IV (not used)    &
-               Intel HD 400 Graphics (not used) &
-               Pascal 256 CUDA cores   \\
-    \hline
-    Memory  & 1GB LPDDR2   &  2GB DDR3L     & 8GB LPDDR4              \\
-    \hline
-	Peak Memory Bandwidth & 8.5 GB/s & 12.8 GB/s & 59.7 GB/s \\
-	\hline
-	Measured Bandwidth & 2127.94 MB/s & 3951.94 MB/s & 4447.90 MB/s \\
-	\hline
-  \end{tabular}
-  \caption{Compared embedded computing platforms}
-  \label{tbl:platforms}
-\end{table*}
+the performance of the shared DRAM controller is of greater importance 
+to the real-time efficiacy of the DeepPicar.
 
@@
 First, we find that DeepPicar's computing platform,
-the Raspberry Pi 3 Model B, offers adquate computing capacity to
-perform real-time control of the RC car at 20 Hz frequency (or
-50ms per control loop). Given the complexity of the DNN used, we
-were pleasantly suprised by this finding. The time breakdown shows that
-the DNN inferencing operation, performed by the Tensorflow library,
-dominates the execution time, which is expected.
+the Raspberry Pi 3 Model B, offers computing capacity to
+perform real-time control of the RC car at 30 Hz frequency (or
+33.$\overline{\mbox{33}}$ ms  per control loop). Given the complexity of 
+the DNN used, we were pleasantly surprised by this finding. 
+The time breakdown shows that the DNN inferencing operation, 
+performed by the Tensorflow library, dominates the execution time, 
+which is expected.
 
 Second, we find that scalability of Tensorflow's DNN 
-implementation is limited. We find that using all four cores is 
-only about 30\% better than using just a single core.
+implementation is limited. We find that using all four cores can double 
+performance of the DNN, but compared to using three cores, performance
+improvement is relatively minimal.
 

### 2018-01-26 paper/comparison.tex
Change figure sizes
Many of the figures were relatively small and hard to see, so the sizes
were increased.
Update platform comparison section

@@
 \section{Embedded Computing Platform Comparison}\label{sec:comparison}
+
+\begin{table*}[h]
+  \centering
+  \begin{adjustbox}{width=1\textwidth}
+  \begin{tabular}{|c|c|c|c|}
+    \hline
+    Item    & Raspberry Pi 3 (B)   & Intel UP                  & NVIDIA Jetson TX2\\
+    \hline
+            & BCM2837              & X5-Z8350 (Cherry Trail)   & Tegra X2 \\
+    CPU     & 4x Cortex-A53@1.2GHz/512KB L2  &
+              4x Atom@1.92GHz/2MB L2 &
+              4x Cortex-A57@2GHz/2MB L2 \\
+            &              &              & 2x Denver@2.0GHz/2MB L2 (not used)  \\
+    \hline
+    GPU     &  VideoCore IV (not used)    &
+               Intel HD 400 Graphics (not used) &
+               Pascal 256 CUDA cores   \\
+    \hline
+    Memory  & 1GB LPDDR2   &  2GB DDR3L     & 8GB LPDDR4              \\
+    \hline
+	Peak Memory Bandwidth & 8.5 GB/s & 12.8 GB/s & 59.7 GB/s \\
+	\hline
+	Measured Bandwidth & 2127.94 MB/s & 3951.94 MB/s & 4447.90 MB/s \\
+	\hline
+  \end{tabular}
+  \end{adjustbox}
+  \caption{Compared embedded computing platforms}
+  \label{tbl:platforms}
+\end{table*}
 
@@
 the experiments in Section~\ref{sec:evaluation} on each hardware
-platform and compare the results. 
+platform and compare the results. We do not, however, repeat the cache
+partitioned experiments done on the Pi 3, as we did not believe our 
+findings warranted further experimentation on additional platforms. 
 For the Jetson TX2, we have two different system configurations,
@@
   \centering
-  \includegraphics[width=.45\textwidth]{figs/compare_core}
+  \includegraphics[width=.7\textwidth]{figs/compare_core}
   \caption{Average control loop execution time.} 
@@
 compared with the Raspberry Pi 3. 
-When all four CPU cores are used, the Intel UP is 2.53X faster than
-the Pi 3, while the TX2 (CPU) and TX2 (GPU) are 3.6X and 7.6X times faster,
+When all four CPU cores are used, the Intel UP is 1.33X faster than
+the Pi 3, while the TX2 (CPU) and TX2 (GPU) are 2.79X and 4.16X times faster,
 respectively, than the Pi 3. 
-As a result, they are all able to satisfy the 50 ms 
+As a result, they are all able to satisfy the 33.$\overline{\mbox{33}}$ ms 
 WCET by a clear margin,
@@
   \centering
-  \includegraphics[width=.45\textwidth]{figs/compare_model}
+  \includegraphics[width=.7\textwidth]{figs/compare_model}
   \caption{Average control loop execution time when multiple DNN
@@
   \centering
-  \includegraphics[width=.45\textwidth]{figs/compare_benchmark}
+  \includegraphics[width=.7\textwidth]{figs/compare_benchmark}
   \caption{Average control loop execution time in the presence of an
@@
   \label{fig:sys_bench}
+\end{figure}
+
+\begin{figure}[h]
+  \centering
+  \includegraphics[width=.7\textwidth]{figs/compare_l2missrate}
+  \caption{L2 miss rates in the presence of an increasing number of 
+			memory intensive applications on idle CPU cores.}
+  \label{fig:sys_l2miss}
 \end{figure}
@@
 Section~\ref{sec:eval-memhog}, we observed dramatic execution time
-increases, up to 9.4X, in Raspberry Pi 3 as we increased the number of
+increases, up to 11.6X, in Raspberry Pi 3 as we increased the number of
 co-scheduled tasks. We also observe increased control loop execution
@@
 2.3X execution time increase; TX2 (CPU) and TX2 (GPU) suffer up to
-2.2X and 2.5X increases, respectively. This is somewhat suprising
+2.2X and 2.1X increases, respectively. This is somewhat suprising
 because the Raspberry Pi 3's cores are in-order architecture based while
@@
 
+However, the L2 miss rates of the platforms follow the same expected 
+pattern, as can be seen in Figure~\ref{fig:sys_l2miss}. Namely, the 
+percentage of L2 miss rates increases are more 
+memory intensive co-runners are introduced. We found that the Pi 3 had 
+slightly higher L2 miss rates than the TX2, both CPU and GPU, but lower 
+miss rates than the UP, which had the highest l2 miss rates. Also, we
+found that the TX2 had similar miss rates when it used either the CPU
+or the GPU, which was expected as the GPU on the TX2 shares a majority 
+of the available DRAM.
+
 Another interesting observation is that the TX2 (GPU) also suffers
-considerable execution time increase (2.5X) despite the fact that the
+considerable execution time increase (2.1X) despite the fact that the
 co-scheduled synthetic tasks do not utilize the GPU. In other words,

### 2018-01-26 paper/appendix.tex
Update conclusion and appendix

@@
-\begin{figure*}[t]
-  \centering
-  \includegraphics[width=.9\textwidth]{figs/TrainingLoss}
-  \caption{Change in loss value throughout training.}
-  \label{fig:modelloss}
-\end{figure*}
-
 \subsection{DNN Training and Testing}
@@
 
+\begin{figure}[h]
+  \centering
+  \includegraphics[width=1.1\textwidth, height=5cm]{figs/TrainingLoss}
+  \caption{Change in loss value throughout training.}
+  \label{fig:modelloss}
+\end{figure}
+
 In the first method, we trained the neural network model across a set 
@@
 opposite direction on the track.
-In total, we collected 2,556 frames for training and 2,609 
+In total, we collected 2,481 frames for training and 2,534 
 frames for validation.
@@
 We repeat this across 2,000 training steps. When a model was trained
-with the  aformentioned data, the training loss was 0.0188 and the
-validation  loss was 0.0132.
+with the  aformentioned data, the training loss was 0.0151 and the
+validation  loss was 0.0251.
 The change of the loss value over the course of model training can be
@@
 However, we find that there is a huge discrepency between the training
-loss and the validation loss as the former was 0.009, while the latter
-was 0.0869---a 10X difference---indicating that the model suffers from
-an overfitting problem.
+loss and the validation loss as the former was 0.0126, while the latter
+was 0.0725 --- almost a 6X difference---indicating that the model 
+suffers from an overfitting problem.
 

### 2018-01-26 paper/conclusion.tex
Update conclusion and appendix

@@
 the Pi 3 platform in the context of deep-learning based real-time
-control appilcations, with a special emphasis on real-time deep neural
+control applications, with a special emphasis on real-time deep neural
 network inferencing.

### 2018-01-26 paper/comparison.tex
Minor changes to comparison section

@@
 multi-model co-scheduling experiment. Once again, they can comfortably
-satisfy 20 Hz real-time performance for all of the co-scheduled DNN control
+satisfy 30 Hz real-time performance for all of the co-scheduled DNN control
 loops, and in the case of the TX2 (GPU), 100 Hz real-time control is still
-feasible. Given that the GPU must be shared among the co-scheduled DNN
-models, the results suggest that the TX2's GPU has sufficient capacity to
-accomodate multiple instances of the DNN model we tested.
+feasible. However, we found that the TX2's GPU may not have the capacity
+necessary to support multiple models running simultaneously. 
+It was unable to co-schedule 4 DNN models due to insufficient memory, 
+hence why there is no data entry for the 4Nx1C test. In our experiements,
+we found that the TX2's GPU could run at most 3 models concurrently, as 
+any more would result in one of the models being terminated.
 

### 2018-01-26 paper/comparison.tex
Update multimodel data for TX2-GPU
Added results for 4Nx1C test. Added relevant data, removed old data, and
updated Figure 17.

--- 2018-01-26 paper/comparison.tex
@@
 multi-model co-scheduling experiment. Once again, they can comfortably
-satisfy 20 Hz real-time performance for all of the co-scheduled DNN control
+satisfy 30 Hz real-time performance for all of the co-scheduled DNN control
 loops, and in the case of the TX2 (GPU), 100 Hz real-time control is still

### 2018-01-28 paper/background.tex
Change figure sizes
Many of the figures were relatively small and hard to see, so the sizes
were increased.
Add Minje's part

--- 2017-11-26 paper/background.tex
@@
-\section{Background and Related Work} \label{sec:background}
+\section{Background} \label{sec:background}
 
-In this section, we provide background and related work on the
-application of deep learning in robotics, particularly autonomous
-vehicles.
+In this section, we provide background on the application of deep
+learning in robotics, particularly autonomous vehicles.
+
+\subsection{Deep Convolutional Neural Network}
+\fixme{CNN background}
 
@@
   \centering
-  \includegraphics[width=.5\textwidth]{figs/endtoend_redrawn}
+  \includegraphics[width=.75\textwidth]{figs/endtoend_redrawn}
   \caption{Standard robotics control vs. DNN based end-to-end
@@
 \subsection{Embedded Computing Platforms for Real-Time Inferencing}
+\fixme{we can consider moving this section to a dedicated related work
+  section at the end.}
 Real-time embedded systems, such as an autonomous vehicle, present
@@
 
-The primary objectives of this study are to understand (1) the
+The \emph{primary objectives of this study} are to understand (1) the
 necessary computing performance for applying AI technology-based

### 2018-01-28 paper/related.tex
Added DonkeyCar to related works as footnote
Add Minje's part

--- 2017-10-03 paper/related.tex
@@
-\section{Related Work}\label{sec:related}
+\section{Related Work}
+% discusstion
+% - model is still quite small. cam we use bigger CNNs?
+\fixme{What we want to say:
+  1) We showed DNN is feasible on today's embedded computing
+  platforms.
+  2) But, the CNN model is rather smaller-kind with respect to today's
+  state-of-the-art CNNs, which are much bigger.
+  3) show table, borrowed from google's object detection api
+  paper~\cite{huang2017speed}. 
+  4) Applying these much more complex DNNs would improve network accuracy
+  but they are obviously challenging for embedded computing
+  platforms. We already discussed computing platform improvement that
+  improve performance of executing a given DNN workload in
+  section 2, so here we discuss model compression that reduces
+  the necessary computing load itself.}
 
-Autonomous vehicles are becoming an increasingly interesting research topic as computing platforms capable of safely processing sensor input into safe vehicle controls become smaller and more affordable.
-The model proposed in End-To-End Learning for Self-Driving Cars \cite{Bojarski2016} is viable on a platform as simple as a Raspberry Pi.
+% related work on efficient DNN representation and processing.
+While repeated feedforward and backpropagation operations during
+the training time account for most of the computational cost in
+deep learning, there is a growing need for an improved efficiency
+during the inferencing time as well, especially because of the
+potential of utilizing Deep Neural Networks (DNN) for the real-time
+pattern recognition tasks in embedded systems, as our paper
+exemplifies.
 
-Given smaller, affordable platforms with multiple cores and even GPU based parallel processing such as the Tegra X1 \cite{NVIDIA2015} and the recently released Tegra X2 \cite{Amert2017}, allocation and management of available shared CPU and GPU resources \cite{Kim2016} becomes increasingly important.
-Research on improving the performance of neural networks \cite{Jouppi2017} also attributes to the viability of using these models in small systems.
+When a DNN is deployed in those implementations with limited
+resources, such as memory and power, the floating-point operations
+involved in the large matrix multiplications are a burdensome task.
+Many recent advances in network compression showed promising results
+in reducing such computational cost during the feedforward
+process. The fundamental assumption in those techniques is that the
+DNNs are redundant in their structure and representation. For example,
+network pruning can thin out the network and provides a more condensed
+topology~\cite{han2015deep}.
+Another common compression method is to reduce the
+quantization level of the network parameters, so that arithmetic
+defined with the floating-point operations are replaced with low-bit
+fixed-point counterparts. To this end, single bit quantization of the
+network parameters or ternary quantization have been recently proposed
+~\cite{hwang2014fixed,soudry2014expectation,kim2016bitwise,rastegari2016xnor,hubara2016binarized,beauchamp2006embedded,govindu2004analysis}.
+In those networks, the inner product between the
+originally real-valued parameter vectors is defined with XNOR followed
+by bit counting, so that the network can greatly minimize the
+computational cost in the hardware implementations. This drastic
+quantization can produce some additional performance loss, but those
+new binarized or ternarized systems provide a simple quantization
+noise injection mechanism during training so that the additional error
+is minimized to an acceptable level.
 
-Together with hardware and platform improvements, there have been significant improvements in the algorithms such as a neural network algorithm capable of surpassing human level control at atari games \cite{DBLP}.
-A popular technique had been to take sensor input and determine a depth map, such as the model used in \cite{Michels:2005}.
-End-to-end models such as the one used in \cite{Bojarski2016} can improve the efficiency of the model.
-Improving the flexibility of the model and it abilty to adapt to different situations and avoiding overfitting to training data continues to be an important topic in autonomous driving \cite{Pomerleau1989}.
+The XNOR operation and bit counting have been known to be efficient in
+hardware implementations. In~\cite{rastegari2016xnor}, it was shown
+that the binarized convolution can substitute the expensive
+convolutional feedforward operations in a regular Convolutional Neural
+Network (CNN), %% by using only about 1.5\% of the memory space, while
+providing 20 to 60 times faster feedforward. Binary weights were also
+able to provide 7 times faster feedforward than a floating-point
+network for the hand written digit recognition task as well as 23
+times faster matrix multiplication tasks on
+GPU~\cite{hubara2016binarized}.
+Moreover, FPGA implementations showed that the XNOR operation is 200
+times cheaper than floating-point multiplications with 
+single precision~\cite{beauchamp2006embedded,govindu2004analysis}.
+XNOR-POP is another hardware implementation that reduced
+the energy consumption of a CNN by 98.7\%~\cite{jiang2017xnor}.
 
-Producing large platforms such as DAVE \cite{LeCun:04} has often been the approach to developing the technology for self-driving vehicles.
-This approach, while certainly effective, presents a higher barrier to entry for research groups.
-Smaller, more focused platforms, such as the one found in \cite{Michels:2005} present a more accesible way forward.
+% related work on testbeds
+\fixme{discuss MIT racecar, UPENN car, donkey car and other testbeds?
+  The points of discussion may include: 1) they are expensive. 2)
+  donkey car is similar to us, but still a bit more expnsive, and more
+  importantly there is no existing studies which systematically evaluate
+  real-time performance of using DNN on embedded computing
+  platforms. Our work provides quantitive real-time performance
+  evaluation results, using a known DNN used in a real self-driving
+  car. Our work is fully reproducible at a very low cost, lowring the
+  barrier to entry in researching self-driving cars, helping the
+  community blah blah blah.}

### 2018-01-29 paper/evaluation.tex
Minor change to evaluation sections
Update PALLOC multicore experiment
Updated Figure 12 with the new data

@@
 timings. In other words, we are interested in how shared resource
-contention affects inference timing.
+contention affects inference timing. These instances are not identical
+copies of the same model, but are instead copies of different models. 
+This is done to ensure that no L2 cache memory is shared between the 
+models as they are running.
 
@@
 
-The results can be seen in Figures \ref{fig:palloc_multicore} - 
-\ref{fig:palloc_bandwidth_l2missrate}. As 
-expected, based off of previous results, we find that the performance 
-of the DNN remains consistent when cache partitioning is employed. 
-When fewer cores are used, DNN performance is slightly affected as 
-the entire L2 cache isn't available to the network. In the case of 
-running multiple models simultaneously, no performance improvements 
-are seen, which further shows that the shared cache has little affect on 
-the DNN's real-time performance.
+We run the DNN model with different amounts of L2 cache being
+available based on the number of colors being used. For example, if two 
+colors are used, then 25\% of the L2 cache is available, whereas if all 
+8 colors are used, 100\% of the L2 cache is available. Particularly, we 
+test the performance of the model when it utilizes a single core and
+all four of the available cores. The results can be seen in 
+\ref{fig:palloc_multicore}. As expected, the amount of L2 cache 
+available to the model does not result in any significant changes in 
+performance. As such, we find that a single model is not sensitive to 
+the shared L2 cache.
 
-In the case of running the DNN alongside synthetic memory-intensive 
-benchmarks, however, we found that partitioning the L2 cache led to a 
-decrease in performance as the number of co-runners increased. 
-However, the L2 miss rates of the model decrease as more benchmarks 
-are introduced. As such, this means that the model encounters greater 
-interence from the memory contention of the co-runners.
+We also co-schedule multiple models to see how cache partitioning affects
+interference between them. Each model uses an equal number of colors, and
+are thus given equal amounts of L2 cache space, and there is no overlap
+in the colors used, so no L2 cache is shared between models. As can be 
+seen in \ref{fig:palloc_multimodel}, the performance of the models 
+remains the same as when no cache partitioning was employed. Based of 
+these results, we find that contention for the L2 cache is not the main
+source of interference between multiple DNN models.
+
+Finally, we run a single model alongside memory intensive co-runners 
+again, but ensure that each task is given equal amounts of L2 cache space.
+For this, we assign two colors to each task, regardless of the number of
+tasks being scheduled. As a result, the entire cache is not utilized in
+experiments with less than four tasks. The results can be seen in 
+\ref{fig:palloc_bandwidth_exectime}. With zero or one co-runners, the DNN
+performance doesnt't change relative to when the cache wasn't partitioned.
+However, we find that with two or more co-runners, DNN performance becomes
+worse. In the case of two co-runners, average control loop time increases
+by 66.31 ms, or 26\%, and with three co-runners, the average time
+is 265.64 ms, of 49\%, greater.
+
+%TODO: Talk about the prefetcher, or something, for why the benchmark performance is worse.
 
@@
 improvements are gained and that performance remains consistent, or 
-decreases, depending on the circumstances. As a result, we conclude that 
-cache partitioning is not an effective isolation mechanism, and that 
+decreases, depending on the circumstances. In all experiements, the DNN
+shows no sensitivity to the shared L2 cache. As a result, we conclude 
+that cache partitioning is not an effective isolation mechanism, and that 
 the performance of the shared DRAM controller is of greater importance 

### 2018-01-29 paper/main.tex
Minor change to paper
Update keywords
Add Minje's part

@@
 \usepackage{adjustbox}
+\usepackage{color}
+\def\fixme#1{\typeout{FIXED in page \thepage : {#1}}
+%  \bgroup \color{red}{} \egroup}
+\bgroup \color{red}{[FIXME: {#1}]} \egroup}
 
@@
 \title{DeepPicar: A Low-cost Deep Neural Network-based Autonomous Car}
-\titlerunning{DeepPicar: A Low-cost Autonomous Car} %optional, in case that the title is too long; the running title should fit into the top page column
+\titlerunning{DeepPicar: A Low-cost Autonomous Car}
+%optional, in case that the title is too long; the running title should fit into the top page column
 
@@
 \author[2]{Elise McEllhiney}
-\author[3]{Heechul Yun}
+\author[3]{Minje Kim}
+\author[4]{Heechul Yun}
 \affil[1]{University of Kansas, Lawrence, United States\\
-  \texttt{m783b224@ku.edu}}
+  \texttt{mbechtel@ku.edu}}
 \affil[2]{University of Kansas, Lawrence, United States\\
   \texttt{e908m429@ku.edu}}
-\affil[3]{University of Kansas, Lawrence, United States\\
+\affil[3]{Indiana University, Bloomington, United States\\
+  \texttt{e908m429@ku.edu}}
+\affil[4]{University of Kansas, Lawrence, United States\\
   \texttt{heechul.yun@ku.edu}}
@@
 \subjclass{C.4 PERFORMANCE OF SYSTEMS}% mandatory: Please choose ACM 1998 classifications from http://www.acm.org/about/class/ccs98-html . E.g., cite as "F.1.1 Models of Computation". 
-\keywords{Self-driving car, CNN, Raspberry Pi 3}% mandatory: Please provide 1-5 keywords
+\keywords{Real-time, Autonomous car, Convolutional neural network, Case study}% mandatory: Please provide 1-5 keywords
 % Author macros::end %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \input{comparison}
-% \input{related}
+\input{related}
 \input{conclusion}

### 2018-01-29 paper/evaluation.tex
Minor change to evaluation section

@@
 by 66.31 ms, or 26\%, and with three co-runners, the average time
-is 265.64 ms, of 49\%, greater.
-
-%TODO: Talk about the prefetcher, or something, for why the benchmark performance is worse.
+is 265.64 ms, of 49\%, greater. This difference is most likely due to the
+L2's prefetcher, which has the potential to generate additional memory
+contention and, thus, affect the performance of the DNN.
+At the same time, though, the L2 miss rates of the model don't correlate 
+to the timing increases seen, as is shown in 
+\ref{fig:palloc_bandwidth_l2missrate}. With partitioning enabled, L2 miss
+rates are higher when no co-runners are present, which is expected, but 
+are the same or lower when co-runners are introduced. If the model was 
+sensitive to the L2 cache, then performance would have improved as 
+proportionaly fewer L2 misses occurred. Since this is not the case, it 
+can be seen that the DNN is more sensitive to the performance of the DRAM
+controller, and not the shared L2 cache.
 

### 2018-01-29 paper/comparison.tex
Add platform costs to table 3
Minor change to comparison section

@@
 	Measured Bandwidth & 2127.94 MB/s & 3951.94 MB/s & 4447.90 MB/s \\
+	\hline
+	Cost (\$) & 35 & 100 & 600 \\
 	\hline
@@
 memory subsystems in the Intel UP and TX2 platforms provide higher
-bandwidth and fairness than the memory subsystem of the Pi 3.
-
-However, the L2 miss rates of the platforms follow the same expected 
+bandwidth and fairness than the memory subsystem of the Pi 3. However, 
+the L2 miss rates of the platforms follow the same expected 
 pattern, as can be seen in Figure~\ref{fig:sys_l2miss}. Namely, the 
@@
 memory intensive co-runners are introduced. We found that the Pi 3 had 
-slightly higher L2 miss rates than the TX2, both CPU and GPU, but lower 
-miss rates than the UP, which had the highest l2 miss rates. Also, we
-found that the TX2 had similar miss rates when it used either the CPU
-or the GPU, which was expected as the GPU on the TX2 shares a majority 
-of the available DRAM.
+slightly higher L2 miss rates than the UP and TX2.
 

### 2018-01-29 paper/evaluation.tex
Change Pi3 specs in evaluation setup
Changed the size of the L1I cache.
Update evaluation section
Added the new figures and discussion regarding their findings.

--- 2018-01-29 paper/evaluation.tex
@@
 BCM2837 SoC, which has a quad-core ARM Cortex-A53 cluster,
-running at up to 1.2GHz. Each core has a 16K private Instruction cache
+running at up to 1.2GHz. Each core has a 32K private Instruction cache
 and a 32K private Data cache, and all cores share a 512KB L2 cache.
@@
   \includegraphics[width=.7\textwidth]{figs/palloc_multicore}
-  \caption{Average control loop execution time vs. \#of CPU
-    cores with cache partitioning enabled.}
+  \caption{Average control loop execution time vs. \# % of L2 cache
+  space available.}
   \label{fig:palloc_multicore}
+\end{figure}
+
+\begin{figure}[h]
+  \centering
+  \includegraphics[width=.7\textwidth]{figs/palloc_multicore_l2missrate}
+  \caption{L2 miss rate impact of limiting the amount of L2 cache space
+  available to the DNN.}
+  \label{fig:palloc_multicore_l2missrate}
 \end{figure}
@@
   \label{fig:palloc_bandwidth_exectime}
+\end{figure}
+
+\begin{figure}[h]
+  \centering
+  \includegraphics[width=.7\textwidth]{figs/palloc_bandwidth_bus}
+  \caption{Bus access impact of co-scheduling memory intensive co-runners 
+when cache partitioning is enabled.}
+  \label{fig:palloc_bandwidth_bus}
 \end{figure}
@@
 available to the model does not result in any significant changes in 
-performance. As such, we find that a single model is not sensitive to 
-the shared L2 cache.
+performance. This can also be seen by the L2 miss rates of the model 
+seen in \ref{fig:palloc_multicore_l2missrate}. As more cache space 
+becomes avaiable, the percentage of L2 misses decreases, which doesn't 
+affect the real-time performance of the DNN. As such, we find that a 
+single model is not sensitive to the shared L2 cache.
 
@@
 is 265.64 ms, of 49\%, greater.
-
-%TODO: Talk about the prefetcher, or something, for why the benchmark performance is worse.
+At the same time, though, the L2 miss rates of the model don't correlate 
+to the timing increases seen, as is shown in 
+\ref{fig:palloc_bandwidth_l2missrate}. Instead, we find that the number
+of bus accesses increases when cache partitions was used, and they mirror
+the timing increases of the model. These additional bus accesses are
+most likely caused by the L2 prefetcher, as it can generate additional bus
+accesses that aren't counted as L2 misses. As a result, we find that cache
+partitioning is detrimental to the performance of the DNN so long as the 
+L2 prefetcher is enabled (we have not yet found a way to disable it on
+the Pi 3). Furthermore, no cache sensitivity was displayed by the model,
+thus showing that the performance of the L2 cache is not vital for DNN 
+performance when co-runners are introduced.
 

### 2018-01-30 paper/intro.tex
intro revision

@@
 
-We find that the DeepPicar is capable of 
-completing control loops in under 33.$\overline{\mbox{33}}$ 
-ms, or 30 hz, and can do so 100\% of the time.
+% Findings:
+%
+% - we find DNN processing is an ideal real-time workload from the
+%   perspective of timing predictability as the amount of computation
+%   needed is fixed at design time and doesn't change over input.
+%
+% - real-time processing of non-trivial DNN inferencing is feasible
+%   on today's embedded multicore platform, even as inexpensive as
+%   raspberry pi3.
+%
+% - all tested platforms have enough computing capacity to consolidate 
+%   multiple DeepPicar's CNN workloads or other compute intensive tasks
+%   simultaneously.
+%
+% - however, when consolidating multiple workloads shared resource
+%   contention is an important issue. we carefully analyzed
+%   sensitivity of DeepPicar's workload with respect to shared
+%   hardware resources.
+%
+% - in particular, we find DNN processing is highly sensitive to
+%   memory performance. performance severely affected when memory
+%   bandwidth is contented.
+%
+% - however, we find that dnn processing is not sensitive to cache
+%   space. we find applying cache partitioning does not help improve
+%   isolation of DNN workload when memory bandwidth is contended.
+From the systematic study, we made a number of interesting
+observations from the perspective of real-time systems.
+% predictability
+First, we find that DNN inferencing is, in a sense, an ideal real-time
+workload as the amount of computation needed to compete a single
+inference is fixed at the DNN architecture design time and does not
+change at runtime over different input (e.g., different image frame).
+This means that it is feasible to achieve highly predictable and
+tightly bounded worst-case execution time estimation.
+Second,
+
+% performance on low-cost embedded multicore
+Second, we find that real-time processing of DeepPicar's DNN-based
+control loop is feasible on today's embedded computing platforms, even 
+as inexpensive as Raspberry Pi 3. We find that the control loop completes 
+under 33.$\overline{\mbox{33}}$ ms, or 30 Hz, using just two cores of
+the Raspberry Pi 3's quad-core CPU (w/o using its GPU), and can
+do so 100\% of the time; we can achieve 20 Hz control frequency using
+just one core (out of four) of the Raspberry Pi 3's quad-core CPU.
 Other tested embedded platforms, Intel UP and NVIDIA TX2, offer even
 better performance, and are capable of supporting deep-learning based 
-real-time control up to 100 Hz control frequency on the TX2 when the GPU 
-is used. However, in all platforms, shared resource contention remains 
-an important issue as we observe up to 11.6X control loop execution time
-increase, mostly due to increase in the neural network inferencing
-operation, when memory performance intensive applications are
-co-scheduled on idle CPU cores.
+real-time control over 100 Hz control frequency on the TX2 when its
+embedded GPU is used.
+We find that all the tested embedded computing platforms have 
+enough computing capacity to consolidate multiple instances of
+DeepPicar's DNN workload.
+
+% all platforms have enough capacity to consolidate
+Third, contention in shared hardware resources remains an important
+issue when consolidating multiple workloads. 
+In particular, we find that DNN inferencing is highly sensitive to
+memory performance inference as we observe up to 11.6X control loop
+execution time increase, when memory performance intensive
+applications are co-scheduled on idle CPU cores.
+We find, however, that DNN inferencing is not sensitive to cache
+space; we show that applying cache partitioning has negligible impact
+in providing timing isolation on consolidated scenarios. This finding
+suggests that in order to guarantee real-time performance of DNN-based
+control applications, solutions to guarantee its memory performance
+are much more important.
 
@@
 \begin{itemize}
-  \item We present the design and implementation of a
-    low-cost autonomous vehicle testbed, DeepPicar, which utilizes
-    state-of-the-art artificial intelligence techniques.
-  \item We provide an analysis and case-study of real-time issues in the
-    DeepPicar.
-  \item We systematically compare real-time computing capabilities of
-    multiple embedded computing platforms in the context of
-    vision-based autonomous driving.
+  \item We provide the design and implementation of a low-cost
+    autonomous car testbed, DeepPicar, which is based on a
+    state-of-the-art end-to-end deep learning, as
+    open-source~\footnote{The source code, build 
+instruction, and a collection of self-driving videos of the DeepPicar
+can be found at: \url{https://github.com/heechul/picar}}, with the
+    goal of lowering the economic and safety-related barriers to study
+    autonomous cars.
+
+  \item We systematically analyze the characteristics of DeepPicar's
+    DNN inferencing workload on a number of representative
+    contemporary embedded computing platforms from the perspective of
+    real-time systems. In particular, to the best of our knowledge,
+    our work is first to investigate the impacts of shared resource
+    contention on DNN inferencing workloads in the context of
+    real-time systems.
+
 \end{itemize}

### 2018-01-30 paper/related.tex
intro revision

@@
   state-of-the-art CNNs, which are much bigger.
-  3) show table, borrowed from google's object detection api
+  3) disscuss google's object detection api
   paper~\cite{huang2017speed}. 
-  4) Applying these much more complex DNNs would improve network accuracy
-  but they are obviously challenging for embedded computing
-  platforms. We already discussed computing platform improvement that
-  improve performance of executing a given DNN workload in
-  section 2, so here we discuss model compression that reduces
-  the necessary computing load itself.}
+  4) Applying these much more complex DNNs (such as google's detection
+  models) may be challenging for embedded computing
+  platforms.
+  5) Two general approaches: (1) make it faster (computing workload
+  remains the same) (2) reduce the workload itself (model compression
+  or reduced precision etc.)}
 
@@
 % related work on testbeds
-\fixme{discuss MIT racecar, UPENN car, donkey car and other testbeds?
+\fixme{maybe we can also discuss MIT racecar, UPENN car, donkey car and other testbeds?
   The points of discussion may include: 1) they are expensive. 2)

### 2018-01-30 paper/intro.tex
Minor changes to intro

@@
 For example, latency in a vision-based object
-detection task may be directly linked to safety of the vehicle. This
+detection task may be directly linked to the safety of the vehicle. This
 requires a high computing capacity as well as the means to guaranteeing
@@
 and 250 thousand parameters (weights).
-The network architecture is identical to that of NVIDIA's DAVE-2
+The network architecture is almost identical to that of NVIDIA's DAVE-2
 self-driving car~\cite{Bojarski2016}, which uses a much more powerful
@@
 tightly bounded worst-case execution time estimation.
-Second,
 

### 2018-01-30 paper/overview.tex
intro revision
Minor changes to overview

@@
 NVIDIA DAVE-2's convolutional neural network (CNN), published by Dr.
-Fridman at  MIT~\footnote{https://github.com/lexfridman/deeptesla}. As
+Fridman at  MIT~\footnote{https://github.com/lexfridman/deeptesla}. The 
+main difference is that we do not utilize a normalization layer, and 
+instead initialize the weights using a Xavier initialization. As
 in DAVE-2, the CNN takes a raw color image (200x66 RGB pixels)
@@
     convolutional, 3 fully-connected layers), 27 million connections,
-    250K parameters. The architecture is identical to the one used
-    in NVIDIA's real self-driving car~\cite{Bojarski2016}.}
+    250K parameters. The architecture is almost identical to the one 
+	used in NVIDIA's real self-driving car~\cite{Bojarski2016}.}
   \label{fig:architecture}
@@
 achieve a reasonable degree of accuracy, successfully navigating
-several different tracks we trained. The source code, build
-instruction, and a collection of self-driving videos of the DeepPicar
-can be found at: \url{https://github.com/heechul/picar}.
+several different tracks we trained.

### 2018-01-30 paper/related.tex
Add initial paragraphs in related works
Added initial sections on the Google Object Detection API and Donkey Car
to the related works section.

@@
   or reduced precision etc.)}
+
+While we were able to show that a smaller embedded platform like the 
+Raspberry Pi 3 can successfully run a DNN, it should be noted that the
+network architecture, and hence model, used was relatively small. Other 
+available DNN's can be much larger compared to the architecture of the
+DAVE-2. Such DNN's can be seen in Google's Object Detection API~\cite{huang2017speed}.
+The smallest model available in the API is the MobileNet model, which
+has 3,191,072 parameters and is approximately 12.7 larger than the
+model used by the DeepPicar. The other models all have parameter sizes 
+of at least 10 million. Due to the considerable differences in the 
+model sizes, it is likely that running them on platforms like the Pi 3,
+and still get acceptable real-time performance would be difficult, 
+if not infeasible.
 
@@
   community blah blah blah.}
+
+Another autonomous vehicle platform is the Donkey Car\footnote{https://donkeycar.com}.
+Like the DeepPicar, the Donkey Car is a low cost platform that can be 
+utilized for researching self-driving cars. However, comparatively, the
+DeepPicar is approximately \$200 cheaper than the Donkey Car. 
+Furthermore, the Donkey Car was created with an emphasis on model accuracy
+and the ability of the car to navigate the desired environment. We are 
+more focused on the real-time performance of the DeepPicar, and whether
+the Raspberry Pi 3 embedded computing platform can properly support autonomous 
+vehicle operations. To the best of our knowledge, no other studies have
+been done to evaluate the capabilities of the Pi 3 for running DNNs and
+performing autonomous vehicle operations.

### 2018-01-31 paper/intro.tex
Minor changes to abstract and intro

@@
 inference is fixed at the DNN architecture design time and does not
-change at runtime over different input (e.g., different image frame).
+change at runtime over different inputs (e.g., different image frame).
 This means that it is feasible to achieve highly predictable and
@@
 control loop is feasible on today's embedded computing platforms, even 
-as inexpensive as Raspberry Pi 3. We find that the control loop completes 
-under 33.$\overline{\mbox{33}}$ ms, or 30 Hz, using just two cores of
-the Raspberry Pi 3's quad-core CPU (w/o using its GPU), and can
+as inexpensive as the Raspberry Pi 3. We find that the control loop 
+completes in under 33.$\overline{\mbox{33}}$ ms, or 30 Hz, using just two 
+cores of the Raspberry Pi 3's quad-core CPU (w/o using its GPU), and can
 do so 100\% of the time; we can achieve 20 Hz control frequency using
-just one core (out of four) of the Raspberry Pi 3's quad-core CPU.
+just one core of the Raspberry Pi 3's quad-core CPU.
 Other tested embedded platforms, Intel UP and NVIDIA TX2, offer even
@@
 In particular, we find that DNN inferencing is highly sensitive to
-memory performance inference as we observe up to 11.6X control loop
+memory performance inference as we observe up to an 11.6X control loop
 execution time increase, when memory performance intensive

### 2018-01-31 paper/main.tex
Minor changes to abstract and intro

@@
 \affil[3]{Indiana University, Bloomington, United States\\
-  \texttt{e908m429@ku.edu}}
+  \texttt{minje@indiana.edu}}
 \affil[4]{University of Kansas, Lawrence, United States\\
@@
 that takes images from a front-facing camera as input and produces
-car steering angles as output. DeepPicar uses the exact same network
-architecture---8 layers, 27 million connections and 250K
+car steering angles as output. DeepPicar uses almost the exact same 
+network architecture---8 layers, 27 million connections and 250K
 parameters---and can be trained to drive itself, in real-time, using a

### 2018-01-31 paper/background.tex
Changes to paper

@@
 The use of neural networks for end-to-end control of autonomous
-car was first demonstrated in late 1980s~\cite{Pomerleau1989},
+car was first demonstrated in the late 1980s~\cite{Pomerleau1989},
 using a small 3-layer fully connected neural network; and subsequently
@@
 efforts include the development of various heterogeneous architecture-based 
-system-on-a-chips (SOCs) that may include multiple cores, GPU,
+system-on-a-chips (SoCs) that may include multiple cores, GPU,
 DSP, FPGA, and neural network optimized ASIC hardware~\cite{Jouppi2017}.

### 2018-01-31 paper/comparison.tex
Changes to paper

@@
   Up are not used in evaluation due to the lack of software (TensorFlow)
-support. Also, the two Denver cores in Tegra TX2 are not used in
+support. Also, the two Denver cores in the Tegra TX2 are not used in
 evaluation due to TensorFlow issues.}.
@@
 %% still meet it by $\sim$20 ms) --> where's the data showing this?
-and, in case of TX2, 50 Hz or even 100 Hz real-time control is
-feasible with the help of its GPU. Another observation is that TX2
-(GPU) does not change much, as most of the neural network computation
-is done at the GPU.
+and, in the case of the TX2, 50 Hz or even 100 Hz real-time control is
+feasible with the help of its GPU. Another observation is that the TX2
+(GPU) performance does not change much, as most of the neural network 
+computation is done at the GPU.
 
@@
 models, the results suggest that the TX2's GPU has sufficient capacity to
-accomodate multiple instances of the DNN model we tested.
+accomodate multiple instances of the DNN models we tested.
 
@@
 bandwidth and fairness than the memory subsystem of the Pi 3. However, 
-the L2 miss rates of the platforms follow the same expected 
+the L2 miss rates of the Pi 3 and TX2 platforms follow the same expected 
 pattern, as can be seen in Figure~\ref{fig:sys_l2miss}. Namely, the 
 percentage of L2 miss rates increases are more 
-memory intensive co-runners are introduced. We found that the Pi 3 had 
-slightly higher L2 miss rates than the UP and TX2.
+memory intensive co-runners are introduced. This was not the case for 
+the UP board, where the model running on two cores generated the highest
+percentage of L2 cache misses. This is most likely due to the CPU 
+architecture of the UP's CPU which is Intel based, whereas both the Pi 3
+and TX2 have ARM based CPUs.
 

### 2018-01-31 paper/evaluation.tex
Changes to paper

@@
 (3) the effect of co-scheduling multiple deep neural network models,
+(4) the effect of co-scheduling memory bandwidth intensive co-runners,
 and
-(4) the effect of co-scheduling memory bandwidth intensive co-runners.
 (5) the effect of partitioning the shared L2 cache.
@@
 linear scaling. In particular, from 3 cores to 4 cores, the
-improvement is mere 2.80 ms, or 12%. In short, we find that the
+improvement is mere 2.80 ms, or 12\%. In short, we find that the
 scalability of DeepPicar's deep neural network is not ideal on the
@@
 tasks being scheduled. As a result, the entire cache is not utilized in
-experiments with less than four tasks. The results can be seen in 
+experiments with less than four tasks. The results can be seen in Figure
 \ref{fig:palloc_bandwidth_exectime}. With zero or one co-runners, the DNN
@@
 \ref{fig:palloc_bandwidth_l2missrate}. Instead, we find that the number
-of bus accesses increases when cache partitions was used, and they mirror
-the timing increases of the model. These additional bus accesses are
+of bus accesses increases when cache partitions are used, and they mirror
+the timing increases of the model. This can be seen in Figure 
+\ref{fig:palloc_bandwidth_bus}. These additional bus accesses are
 most likely caused by the L2 prefetcher, as it can generate additional bus
@@
 practically no benefits to the real-time performance of the DNN. To the 
-contrary, performance is negatively affected when fewer cores are used, 
-due to less cache being available, and memory intensive co-runners are 
-introduced, as they generate more memory contention. Furthermore, we 
+contrary, performance is constant regardless of the amount of cache 
+space used. We also find that performance becomes worse when memory
+intensive co-runners are introduced as they generate more memory 
+contention when partitioning is enabled. Furthermore, we 
 find that performance remains consistent when multiple models are run 

### 2018-02-01 paper/background.tex
remove fixmes

@@
 
-\subsection{Deep Convolutional Neural Network}
-\fixme{CNN background}
+%% \subsection{Deep Convolutional Neural Network}
+%% \fixme{CNN background}
 
@@
 \subsection{Embedded Computing Platforms for Real-Time Inferencing}
-\fixme{we can consider moving this section to a dedicated related work
-  section at the end.}
+%% \fixme{we can consider moving this section to a dedicated related work
+%%   section at the end.}
 Real-time embedded systems, such as an autonomous vehicle, present

### 2018-02-01 paper/related.tex
remove fixmes

--- 2018-01-30 paper/related.tex
@@
 % - model is still quite small. cam we use bigger CNNs?
-\fixme{What we want to say:
-  1) We showed DNN is feasible on today's embedded computing
-  platforms.
-  2) But, the CNN model is rather smaller-kind with respect to today's
-  state-of-the-art CNNs, which are much bigger.
-  3) disscuss google's object detection api
-  paper~\cite{huang2017speed}. 
-  4) Applying these much more complex DNNs (such as google's detection
-  models) may be challenging for embedded computing
-  platforms.
-  5) Two general approaches: (1) make it faster (computing workload
-  remains the same) (2) reduce the workload itself (model compression
-  or reduced precision etc.)}
+%% \fixme{What we want to say:
+%%   1) We showed DNN is feasible on today's embedded computing
+%%   platforms.
+%%   2) But, the CNN model is rather smaller-kind with respect to today's
+%%   state-of-the-art CNNs, which are much bigger.
+%%   3) disscuss google's object detection api
+%%   paper~\cite{huang2017speed}. 
+%%   4) Applying these much more complex DNNs (such as google's detection
+%%   models) may be challenging for embedded computing
+%%   platforms.
+%%   5) Two general approaches: (1) make it faster (computing workload
+%%   remains the same) (2) reduce the workload itself (model compression
+%%   or reduced precision etc.)}
 
-% related work on efficient DNN representation and processing.
-While repeated feedforward and backpropagation operations during
-the training time account for most of the computational cost in
-deep learning, there is a growing need for an improved efficiency
-during the inferencing time as well, especially because of the
-potential of utilizing Deep Neural Networks (DNN) for the real-time
-pattern recognition tasks in embedded systems, as our paper
-exemplifies.
+%% {\bf DNN optimization for embedded systems.}
 
+While we showed that a low-cost embedded platform such as the 
+Raspberry Pi 3 can successfully implement CNN-based autonomous driving
+in real-time, it should be noted that the CNN we used here is relatively
+small and recent state-of-the-art CNNs tend to be much larger.
+For example, the CNN based object detector models evaluated in
+Google's recent study~\cite{huang2017speed} have between 3M to 54M
+parameters. Using such large CNN models will be challenging on
+resource constrainted embedded computing platforms, especially for
+real-time applications such as self-driving cars.
+
+%% available DNN's can be much larger compared to the architecture of the
+%% DAVE-2. Such DNN's can be seen in Google's Object Detection API
+%% The smallest model available in the API is the MobileNet model, which
+%% has 3,191,072 parameters and is approximately 12.7 larger than the
+%% model used by the DeepPicar. The other models all have parameter sizes 
+%% of at least 10 million. Due to the considerable differences in the 
+%% model sizes, it is likely that running them on platforms like the Pi 3,
+%% and still get acceptable real-time performance would be difficult, 
+%% if not infeasible.
+
+%% % related work on efficient DNN representation and processing.
+%% While repeated feedforward and backpropagation operations during
+%% the training time account for most of the computational cost in
+%% deep learning, there is a growing need for an improved efficiency
+%% during the inferencing time as well, especially because of the
+%% potential of utilizing Deep Neural Networks (DNN) for the real-time
+%% pattern recognition tasks in embedded systems, as our paper
+%% exemplifies.
+
+While continuing performance improvmenets in embedded computing
+platforms will certainly make processing of these complex CNNs faster,
+another actively investigate approrach is to reduce the required
+computational complexity itself.
 When a DNN is deployed in those implementations with limited
@@
 
-% related work on testbeds
-\fixme{maybe we can also discuss MIT racecar, UPENN car, donkey car and other testbeds?
-  The points of discussion may include: 1) they are expensive. 2)
-  donkey car is similar to us, but still a bit more expnsive, and more
-  importantly there is no existing studies which systematically evaluate
-  real-time performance of using DNN on embedded computing
-  platforms. Our work provides quantitive real-time performance
-  evaluation results, using a known DNN used in a real self-driving
-  car. Our work is fully reproducible at a very low cost, lowring the
-  barrier to entry in researching self-driving cars, helping the
-  community blah blah blah.}
+We believe these reseach can make complex DNNs more accessible for
+real-time embedded systems.
+
+%% % related work on testbeds
+%% \fixme{maybe we can also discuss MIT racecar, UPENN car, donkey car and other testbeds?
+%%   The points of discussion may include: 1) they are expensive. 2)
+%%   donkey car is similar to us, but still a bit more expnsive, and more
+%%   importantly there is no existing studies which systematically evaluate
+%%   real-time performance of using DNN on embedded computing
+%%   platforms. Our work provides quantitive real-time performance
+%%   evaluation results, using a known DNN used in a real self-driving
+%%   car. Our work is fully reproducible at a very low cost, lowring the
+%%   barrier to entry in researching self-driving cars, helping the
+%%   community blah blah blah.}
+
+%% Another autonomous vehicle platform is the Donkey Car\footnote{https://donkeycar.com}.
+%% Like the DeepPicar, the Donkey Car is a low cost platform that can be 
+%% utilized for researching self-driving cars. However, comparatively, the
+%% DeepPicar is approximately \$200 cheaper than the Donkey Car. 
+%% Furthermore, the Donkey Car was created with an emphasis on model accuracy
+%% and the ability of the car to navigate the desired environment. We are 
+%% more focused on the real-time performance of the DeepPicar, and whether
+%% the Raspberry Pi 3 embedded computing platform can properly support autonomous 
+%% vehicle operations. To the best of our knowledge, no other studies have
+%% been done to evaluate the capabilities of the Pi 3 for running DNNs and
+%% performing autonomous vehicle operations.

### 2018-02-01 paper/intro.tex
video links

@@
     open-source~\footnote{The source code, build 
-instruction, and a collection of self-driving videos of the DeepPicar
+instruction the DeepPicar
 can be found at: \url{https://github.com/heechul/picar}}, with the

### 2018-02-01 paper/comparison.tex
revision

@@
 Table~\ref{tbl:platforms} shows architectural features of the three
-platforms~\footnote{The GPUs of the Raspberry Pi 3 and Intel
-  Up are not used in evaluation due to the lack of software (TensorFlow)
-support. Also, the two Denver cores in the Tegra TX2 are not used in
-evaluation due to TensorFlow issues.}.
+platforms~\footnote{The GPU of Intel UP and the two Denver cores in the
+  Tegra TX2 are not used in evaluation due to TensorFlow issues.}.
 

### 2018-02-01 paper/evaluation.tex
revision

@@
 the execution time.
-Second, and more importantly, we also find that the measured average
+
+Second, we find that the measured average
 execution time of a single control loop is 22.87 ms, or 43.7 Hz and
@@
 This means that the DeepPicar can operate
-at a 30 Hz control frequency in real-time using only the on-board
+at a up to 40 Hz control frequency in real-time using only the on-board
 Raspberry Pi 3 computing platform, as no remote computing resources 
@@
 Raspberry Pi 3 (its GPU is not supported by Tensorflow).
-
 In comparison, NVIDIA's DAVE-2 system, which has the exact same neural
-network architecture, reportedly runs also at 30 Hz~\cite{Bojarski2016}. 
+network architecture, reportedly runs at 30 Hz~\cite{Bojarski2016}. 
 Although we believe it was not
@@
 including NVIDIA's Jetson TX2, later in
-Section~\ref{sec:comparison}), the fact that the low-cost
-Raspberry Pi 3 can achieve similar real-time control performance is
+Section~\ref{sec:comparison}), the fact that a low-cost
+Raspberry Pi 3 can achieve comparable real-time control performance is
 surprising.
+
+Lastly, we find that the control loop execution timing is highly
+predictable and shows very little variance over different input image
+frames. This is because the amount of computation needed to perform
+a DNN inferencing operation is fixed at the DNN architecture design
+time and does not change at runtime over different inputs (i.e.,
+different image frames). This predictable timing behavior is a highly
+desirable property for real-time systems, making DNN inferencing an
+attractive real-time workload.
 
@@
 scalability of DeepPicar's deep neural network is not ideal on the
-platform. We do not know whether it is due to the limitations of
-TensorFlow's multicore implementation or if it's the model's inherent
-characteristics.
+platform.
 
-The poor scalability opens up the possibility of consolidating
-multiple different tasks or different neural network models rather
-than allocating all cores for a single neural network model. For
-example, it is conceivable to use four cameras and four different
-neural network models, each of which is trained separately and
-executed on a single dedicated core. Assuming we use the same network
+As noted in~\cite{NVIDIA2015}, DNN inferencing is inherently more
+difficult to parallelize than training because the easiest
+parallelization option, batching (i.e., processing multiple images in
+parallel), is not available or is limited. Specifically, in DeepPicar,
+only one image frame, obtained from the camera, can be processed at a
+time. Thus, more fine-grained algorithmic parallelization is needed to
+improve inference performance~\cite{NVIDIA2015}, which generally does
+not scale very well.
+
+On the other hand, the poor scalability opens up the possibility of
+consolidating multiple different tasks or different neural network
+models rather than allocating all cores for a single neural network
+model.
+For example, it is conceivable to use four cameras and four different
+neural network models, each of which is trained separately for
+different purpose and executed on a single dedicated core.
+Assuming we use the same network
 architecture for all models, then one might expect to achieve up to
 20 Hz using one core (given 1 core can deliver 46 ms average
-execution time). In the next experiment, we investigate the
+execution time).
+In the next experiment, we investigate the
 feasibility of such a scenario.
@@
 
-\begin{figure}[h]
-  \centering
-  \includegraphics[width=.7\textwidth]{figs/l2missrate_vs_modelcnt}
-  \caption{L2 cache miss rates of different neural network and core
-    assignments. X-axis is the same as Figure~\ref{fig:perf-vs-modelcnt}.} 
-  \label{fig:l2missrate-vs-modelcnt}
-\end{figure}
+These increases in inference times in the co-scheduled scenarios are
+expected and are caused by contention in the shared hardware
+resources, such as the shared L2 cache and/or the DRAM controller.
 
-These increases in inference times in the co-scheduled scenarios are
-expected and likely caused by contention in the shared hardware
-resources, such as the shared L2 cache and/or the DRAM controller.
-To further analyze the source of contention, we use hardware
-performance counters of the processor. Specifically, we measure L2
-miss rates of the DNN models first in isolation and then after
-co-scheduling other models. If the shared L2 cache is the primary
-source of inteference, then the measured L2 miss rates will
-increase. Figure~\ref{fig:l2missrate-vs-modelcnt} shows the results.
-As can be see in the figure, L2 miss rates increase as more models are 
-co-scheduled, but the increase is less than expected.
-This suggests that the shared L2 cache is not the main bottleneck that 
-causes execution time increases. In other words, DNN models don't appear 
-to be significantly affected by the shared L2 cache space. 
-% L2 partitioning is probably not going to be useful.
-Instead, we hypothesize that it is likely caused by the memory
-controller---the only other major shared hardware source---where
-memory requests from different CPU cores contend, which would result
-in increased memory access latency. While some Intel processors
-provide incore hardware counters that can measure average memory
-access latency~\cite{ye2016maracas}, we were not able to identify
-whether such hardware counters exist in the BCM2837 processor of
-Raspberry Pi 3 due to the lack of documentation. Instead, in the next
-experiment, we use memory intensive synthetic benchmarks to test the
-hypothesis.
+%% \begin{figure}[h]
+%%   \centering
+%%   \includegraphics[width=.7\textwidth]{figs/l2missrate_vs_modelcnt}
+%%   \caption{L2 cache miss rates of different neural network and core
+%%     assignments. X-axis is the same as Figure~\ref{fig:perf-vs-modelcnt}.} 
+%%   \label{fig:l2missrate-vs-modelcnt}
+%% \end{figure}
+
+%% To further analyze the source of contention, we use hardware
+%% performance counters of the processor. Specifically, we measure L2
+%% miss rates of the DNN models first in isolation and then after
+%% co-scheduling other models. If the shared L2 cache is the primary
+%% source of inteference, then the measured L2 miss rates will
+%% increase. Figure~\ref{fig:l2missrate-vs-modelcnt} shows the results.
+%% As can be see in the figure, L2 miss rates increase as more models are 
+%% co-scheduled, but the increase is less than expected.
+%% This suggests that the shared L2 cache is not the main bottleneck that 
+%% causes execution time increases. In other words, DNN models don't appear 
+%% to be significantly affected by the shared L2 cache space. 
+%% % L2 partitioning is probably not going to be useful.
+%% Instead, we hypothesize that it is likely caused by the memory
+%% controller---the only other major shared hardware source---where
+%% memory requests from different CPU cores contend, which would result
+%% in increased memory access latency. While some Intel processors
+%% provide incore hardware counters that can measure average memory
+%% access latency~\cite{ye2016maracas}, we were not able to identify
+%% whether such hardware counters exist in the BCM2837 processor of
+%% Raspberry Pi 3 due to the lack of documentation. Instead, in the next
+%% experiment, we use memory intensive synthetic benchmarks to test the
+%% hypothesis.
 
@@
 
-In order to determine how contended DRAM requests affect the DNN
-inference timing of the DeepPicar, we use a synthetic memory
+In this experiment, we investigate the impact of contended DRAM
+requests to the DNN inference timing of DeepPicar.
+For the experiment, we use a synthetic memory
 intensive benchmark from the IsolBench suite~\cite{Valsan2016}.
-
 We run a single DNN model on one core, and
@@
 dedicated core (Core 0). On the other hand, the DNN model's L2
-cache-miss rates do not increase as much. This suggests that
-the DNN model's execution increase cannot be fully explained by
-increases in L2 cache-misses. Instead, as we hypothesized in the previous
-experiment, the increased memory pressure from the co-scheduled memory
-intensive benchmarks is likely the primary cause of the DNN model's execution
-time increase. Therefore, we conclude that DeepPicar's DNN model is
-more senstive to DRAM access latency than L2 cache space.
+cache-miss rates do not increase as much.
+In other words, the DNN model's execution increase is not fully
+explained by increased contention in the L2 cache space.
+Instead, the high memory bandwidth pressure from the co-scheduled
+memory intensive benchmarks is likely the primary cause.
+
+%% Instead, as we hypothesized in the previous
+%% experiment, the increased memory pressure from the co-scheduled memory
+%% intensive benchmarks is likely the primary cause of the DNN model's execution
+%% time increase. Therefore, we conclude that DeepPicar's DNN model is
+%% more senstive to DRAM access latency than L2 cache space.
 
@@
 techniques~\cite{Gracioli2015,Kim2016} may not be effective isolation
-solutions for DeepPicar, as its AI workload is more sensitive to memory
-performance. Instead, memory controller focused isolation solutions,
-either hardware or software-based ones (e.g.,~\cite{Guo2017,Yun2013}),
-may be more important. Although our observation is made on a single
-hardware platform running on a single DNN workload, we suspect that
-many AI workloads may exhibit similar characteristics.
+solutions for DeepPicar's DNN processing workload, as cache
+partitioning does not provide memory performance guarantees.
 
-\subsection{Effect of Cache Partitioning}
+%% Instead, memory controller focused isolation solutions,
+%% either hardware or software-based ones (e.g.,~\cite{Guo2017,Yun2013}),
+%% may be more important. Although our observation is made on a single
+%% hardware platform running on a single DNN workload, we suspect that
+%% many AI workloads may exhibit similar characteristics.
 
-\begin{figure}[h]
-  \centering
-  \includegraphics[width=\textwidth]{figs/cache-mapping}
-  \caption{A memory mapping of the cache sets for the Broadcom 
-BCM2837 processor used in the Raspberry Pi 3 Model B.}
-  \label{fig:cache-mapping}
-\end{figure}
-
-\begin{figure}[h]
-  \centering
-  \includegraphics[width=.7\textwidth]{figs/palloc_multicore}
-  \caption{Average control loop execution time vs. \# % of L2 cache
-  space available.}
-  \label{fig:palloc_multicore}
-\end{figure}
-
-\begin{figure}[h]
-  \centering
-  \includegraphics[width=.7\textwidth]{figs/palloc_multicore_l2missrate}
-  \caption{L2 miss rate impact of limiting the amount of L2 cache space
-  available to the DNN.}
-  \label{fig:palloc_multicore_l2missrate}
-\end{figure}
-
-\begin{figure}[h]
-  \centering
-  \includegraphics[width=.7\textwidth]{figs/palloc_multimodel}
-  \caption{Timing impact of co-scheduling multiple DNNs when cache 
-partitioning is enabled.}
-  \label{fig:palloc_multimodel}
-\end{figure}
-
-\begin{figure}[h]
-  \centering
-  \includegraphics[width=.7\textwidth]{figs/palloc_bandwidth_exectime}
-  \caption{Timing impact of co-scheduling memory intensive co-runners 
-when cache partitioning is enabled.}
-  \label{fig:palloc_bandwidth_exectime}
-\end{figure}
-
-\begin{figure}[h]
-  \centering
-  \includegraphics[width=.7\textwidth]{figs/palloc_bandwidth_bus}
-  \caption{Bus access impact of co-scheduling memory intensive co-runners 
-when cache partitioning is enabled.}
-  \label{fig:palloc_bandwidth_bus}
-\end{figure}
-
-\begin{figure}[h]
-  \centering
-  \includegraphics[width=.7\textwidth]{figs/palloc_bandwidth_l2missrate}
-  \caption{L2 miss rate impact of co-scheduling memory intensive co-
-runners when cache partitioning is enabled.}
-  \label{fig:palloc_bandwidth_l2missrate}
-\end{figure}
-
-In order to determine if the DNN is mostly affected by the DRAM 
-controller, we partition the L2 cache of the Pi3 and reperform the 
-same experiments to see if there are any noticable changes. For the 
-partition, we employ PALLOC~\cite{yun2014rtas}, a color-based page 
-allocator that works on the kernel level. For the bit mask, we select 
-bits 12, 13 and 14 as they can be used to access the L2 cache,
-as can be seen by Figure \ref{fig:cache-mapping}. This results in 
-2\textsuperscript{3} = 8 colors, which we then assign to the Pi3's 
-physical cores such that each core has two unique colors (colors 0 
-and 1 are assigned to core 0, colors 2 and 3 are assigned to core 1, 
-etc.). In the case of bit 12, since we use 2 colors for each partition,
-the L1 Data cache is not partitioned in our experiments.
-
-If performance improves as a result of partitioning the shared cache 
-then we know that the DNN, to some extent, relies on shared cache 
-performance in addition to shared DRAM performance. On the other 
-hand, if there is no improvement in performance, then it can be 
-observed that the shared memory is more critical to the performance 
-of the DNN.
-
-We run the DNN model with different amounts of L2 cache being
-available based on the number of colors being used. For example, if two 
-colors are used, then 25\% of the L2 cache is available, whereas if all 
-8 colors are used, 100\% of the L2 cache is available. Particularly, we 
-test the performance of the model when it utilizes a single core and
-all four of the available cores. The results can be seen in 
-\ref{fig:palloc_multicore}. As expected, the amount of L2 cache 
-available to the model does not result in any significant changes in 
-performance. This can also be seen by the L2 miss rates of the model 
-seen in \ref{fig:palloc_multicore_l2missrate}. As more cache space 
-becomes avaiable, the percentage of L2 misses decreases, which doesn't 
-affect the real-time performance of the DNN. As such, we find that a 
-single model is not sensitive to the shared L2 cache.
-
-We also co-schedule multiple models to see how cache partitioning affects
-interference between them. Each model uses an equal number of colors, and
-are thus given equal amounts of L2 cache space, and there is no overlap
-in the colors used, so no L2 cache is shared between models. As can be 
-seen in \ref{fig:palloc_multimodel}, the performance of the models 
-remains the same as when no cache partitioning was employed. Based of 
-these results, we find that contention for the L2 cache is not the main
-source of interference between multiple DNN models.
-
-Finally, we run a single model alongside memory intensive co-runners 
-again, but ensure that each task is given equal amounts of L2 cache space.
-For this, we assign two colors to each task, regardless of the number of
-tasks being scheduled. As a result, the entire cache is not utilized in
-experiments with less than four tasks. The results can be seen in Figure
-\ref{fig:palloc_bandwidth_exectime}. With zero or one co-runners, the DNN
-performance doesnt't change relative to when the cache wasn't partitioned.
-However, we find that with two or more co-runners, DNN performance becomes
-worse. In the case of two co-runners, average control loop time increases
-by 66.31 ms, or 26\%, and with three co-runners, the average time
-is 265.64 ms, of 49\%, greater.
-At the same time, though, the L2 miss rates of the model don't correlate 
-to the timing increases seen, as is shown in 
-\ref{fig:palloc_bandwidth_l2missrate}. Instead, we find that the number
-of bus accesses increases when cache partitions are used, and they mirror
-the timing increases of the model. This can be seen in Figure 
-\ref{fig:palloc_bandwidth_bus}. These additional bus accesses are
-most likely caused by the L2 prefetcher, as it can generate additional bus
-accesses that aren't counted as L2 misses. As a result, we find that cache
-partitioning is detrimental to the performance of the DNN so long as the 
-L2 prefetcher is enabled (we have not yet found a way to disable it on
-the Pi 3). Furthermore, no cache sensitivity was displayed by the model,
-thus showing that the performance of the L2 cache is not vital for DNN 
-performance when co-runners are introduced.
-
-By partitioning the shared L2 cache, we find that no noticable 
-improvements are gained and that performance remains consistent, or 
-decreases, depending on the circumstances. In all experiements, the DNN
-shows no sensitivity to the shared L2 cache. As a result, we conclude 
-that cache partitioning is not an effective isolation mechanism, and that 
-the performance of the shared DRAM controller is of greater importance 
-to the real-time efficiacy of the DeepPicar.
+% \input{evaluation-cachepart}
 
@@
 
-Lastly, we find that partitioning the shared L2 cache provides 
-practically no benefits to the real-time performance of the DNN. To the 
-contrary, performance is constant regardless of the amount of cache 
-space used. We also find that performance becomes worse when memory
-intensive co-runners are introduced as they generate more memory 
-contention when partitioning is enabled. Furthermore, we 
-find that performance remains consistent when multiple models are run 
-simultaneously. As such, we conclude that the shared L2 cache is not 
-essential to the DNN's real-time performance.
+%% Lastly, we find that partitioning the shared L2 cache provides 
+%% practically no benefits to the real-time performance of the DNN. To the 
+%% contrary, performance is constant regardless of the amount of cache 
+%% space used. We also find that performance becomes worse when memory
+%% intensive co-runners are introduced as they generate more memory 
+%% contention when partitioning is enabled. Furthermore, we 
+%% find that performance remains consistent when multiple models are run 
+%% simultaneously. As such, we conclude that the shared L2 cache is not 
+%% essential to the DNN's real-time performance.

### 2018-02-01 paper/overview.tex
Changes to paper
video links
revision

@@
 
-\begin{figure}[t]
-  \centering
-  \includegraphics[width=.55\textwidth]{figs/DeepPicar_platform}
-  \caption{DeepPicar platform.}
-  \label{fig:overview}
-\end{figure}
-
 In this section, we provide an overview of our DeepPicar platform. In
-developing DeepPicar, one of our primary goals is to replicate the
-NVIDIA's DAVE-2 system on a smaller scale with using a low cost
-multicore platform, the Raspberry Pi 3. Because the Pi 3's computing
-performance is much lower than that of the DRIVE PX platform used in
-DAVE-2, we are interested in if, and how, we can process
+developing DeepPicar, one of our primary goals is to faithfully
+replicate NVIDIA's DAVE-2 system on a smaller scale with using a low cost
+multicore platform, the Raspberry Pi 3. Because the Raspberry Pi 3's
+computing performance is much lower than that of the DRIVE PX platform
+used in DAVE-2, we are interested in if, and how, we can process
 computationally expensive neural network operations in
@@
 
+\begin{figure}[h]
+  \centering
+  \includegraphics[width=.55\textwidth]{figs/DeepPicar_platform}
+  \caption{DeepPicar platform.}
+  \label{fig:overview}
+\end{figure}
+
 Figure~\ref{fig:overview} shows the DeepPicar, which is comprised of a
@@
 
-\begin{table}[t]
+\begin{table}[h]
   \centering
@@
 
-For the neural network architecture, we adopt a TensorFlow version of
-NVIDIA DAVE-2's convolutional neural network (CNN), published by Dr.
-Fridman at  MIT~\footnote{https://github.com/lexfridman/deeptesla}. The 
-main difference is that we do not utilize a normalization layer, and 
-instead initialize the weights using a Xavier initialization. As
-in DAVE-2, the CNN takes a raw color image (200x66 RGB pixels)
-as input and produces a single steering angle value as
-output. Figure~\ref{fig:architecture} shows the network architecture, which
-is comprised of 8 layers, 250K parameters, and about 27 million
-connections.
-
-\begin{figure}[t]
+\begin{figure}[h]
   \centering
-  \includegraphics[width=\textwidth]{figs/architecture}
+  \includegraphics[width=0.7\textwidth]{figs/architecture}
   \caption{DeepPicar's neural network architecture: 8 layers (5
@@
 
-\begin{figure}[t]
+For the neural network architecture, we implement
+NVIDIA DAVE-2's convolutional neural network (CNN) using TensorFlow,
+based on ~\cite{deeptesla}. Note, however, that the CNN model
+in~\cite{deeptesla} was considerably bigger than NVIDIA's CNN as it contained an
+additional fully-connected layer of approximately 1.3M additional
+parameters. We eliminated the additional layer to faithfully recreate
+the NVIDIA's original CNN model.
+%% The main difference is that we do not utilize a normalization layer, and 
+%% instead initialize the weights using a Xavier initialization. 
+As in DAVE-2, the CNN takes a raw color image (200x66 RGB pixels)
+as input and produces a single steering angle value as
+output. Figure~\ref{fig:architecture} shows the network architecture
+used in this paper, which
+is comprised of 8 layers, 250K parameters, and about 27 million
+connections. It is identical to NVIDIA DAVE-2's CNN except that we do
+not apply a normalization layer as it does not include trainable
+parameters and its computational demand is minimal.
+
+\begin{figure}[h]
   \centering
@@
   \label{fig:track}
-\end{figure}
-
-\begin{figure}[t]
-   \lstset{language=python,
-           basicstyle=\ttfamily\small,
-           keywordstyle=\color{blue}\ttfamily,
-           stringstyle=\color{red}\ttfamily,
-           commentstyle=\color{green}\ttfamily
-          }  
-  \lstinputlisting[language=python]{control.py}
-  \caption{Control loop}
-  \label{fig:controlloop}
 \end{figure}
@@
 desktop computer using the GTX 1060 GPU.
+
+\begin{figure}[h]
+   \lstset{language=python,
+           basicstyle=\ttfamily\small,
+           keywordstyle=\color{blue}\ttfamily,
+           stringstyle=\color{red}\ttfamily,
+           commentstyle=\color{green}\ttfamily
+          }  
+  \lstinputlisting[language=python]{control.py}
+  \caption{Control loop}
+  \label{fig:controlloop}
+\end{figure}
 
@@
 achieve a reasonable degree of accuracy, successfully navigating
-several different tracks we trained.
+several different tracks we trained. In our testing, we found that the
+DeepPicar could remain on a track for over 10
+minutes~\footnote{A collection of self-driving videos can be found at: \url{https://photos.app.goo.gl/ce93sU7jPk4ywO8u2}}.

### 2018-02-01 paper/evaluation-cachepart.tex
cache partition section as a separate file.
for now it is not included.

\subsection{Effect of Cache Partitioning}

\begin{figure}[h]
  \centering
  \includegraphics[width=\textwidth]{figs/cache-mapping}
  \caption{A memory mapping of the cache sets for the Broadcom 
BCM2837 processor used in the Raspberry Pi 3 Model B.}
  \label{fig:cache-mapping}
\end{figure}

\begin{figure}[h]
  \centering
  \includegraphics[width=.7\textwidth]{figs/palloc_multicore}
  \caption{Average control loop execution time vs.  % of L2 cache
    L2 cache partition size.}
  \label{fig:palloc_multicore}
\end{figure}

%% \begin{figure}[h]
%%   \centering
%%   \includegraphics[width=.7\textwidth]{figs/palloc_multicore_l2missrate}
%%   \caption{L2 miss rate impact of limiting the amount of L2 cache space
%%   available to the DNN.}
%%   \label{fig:palloc_multicore_l2missrate}
%% \end{figure}

%% \begin{figure}[h]
%%   \centering
%%   \includegraphics[width=.7\textwidth]{figs/palloc_multimodel}
%%   \caption{Timing impact of co-scheduling multiple DNNs when cache 
%% partitioning is enabled.}
%%   \label{fig:palloc_multimodel}
%% \end{figure}

\begin{figure}[h]
  \centering
  \includegraphics[width=.7\textwidth]{figs/palloc_bandwidth_exectime}
  \caption{Timing impact of co-scheduling memory intensive co-runners 
when cache partitioning is enabled.}
  \label{fig:palloc_bandwidth_exectime}
\end{figure}

%% \begin{figure}[h]
%%   \centering
%%   \includegraphics[width=.7\textwidth]{figs/palloc_bandwidth_bus}
%%   \caption{Bus access impact of co-scheduling memory intensive co-runners 
%% when cache partitioning is enabled.}
%%   \label{fig:palloc_bandwidth_bus}
%% \end{figure}

%% \begin{figure}[h]
%%   \centering
%%   \includegraphics[width=.7\textwidth]{figs/palloc_bandwidth_l2missrate}
%%   \caption{L2 miss rate impact of co-scheduling memory intensive co-
%% runners when cache partitioning is enabled.}
%%   \label{fig:palloc_bandwidth_l2missrate}
%% \end{figure}

In order to validate the effectiveness of cache partitioning, we patch
the kernel with PALLOC~\cite{yun2014rtas}, which is a kernel-level
color-aware page allocator, supporting cache partitioning. 
Figure~\ref{fig:cache-mapping} shows the physical address mapping of
Raspberry Pi3's BCM2837 processor. We use the bit 12, 13, 14 for
coloring, which results in 8 page colors.

%% determine if the DNN is mostly affected by the DRAM 
%% controller, we partition the L2 cache of the Pi3 and reperform the 
%% same experiments to see if there are any noticable changes. For the 
%% partition, we employ PALLOC~\cite{yun2014rtas}, a color-based page 
%% allocator that works on the kernel level.

%% For the bit mask, we select 
%% bits 12, 13 and 14 as they can be used to access the L2 cache,
%% as can be seen by Figure \ref{fig:cache-mapping}.
%% This results in 
%% 2\textsuperscript{3} = 8 colors, which we then assign to the Pi3's 
%% physical cores such that each core has two unique colors (colors 0 
%% and 1 are assigned to core 0, colors 2 and 3 are assigned to core 1, 
%% etc.). In the case of bit 12, since we use 2 colors for each partition,
%% the L1 Data cache is not partitioned in our experiments.

In the first experiment, we investigate the cache space sensitivity of
the DeepPicar's DNN-based control loop. Using PALLOC, we create 4
different cgroups which are configured to use 2, 4,
6, 8 colors (25\%, 50\%, 75\%, 100\% of the L2 cache
space, respectively). We then execute the control loop on each cache
partition (cgroup) and measure the average processing time.

If performance improves as a result of partitioning the shared cache 
then we know that the DNN, to some extent, relies on shared cache 
performance in addition to shared DRAM performance. On the other 
hand, if there is no improvement in performance, then it can be 
observed that the shared memory is more critical to the performance 
of the DNN.

We run the DNN model with different amounts of L2 cache being
available based on the number of colors being used. For example, if two 
colors are used, then 25\% of the L2 cache is available, whereas if all 
8 colors are used, 100\% of the L2 cache is available. Particularly, we 
test the performance of the model when it utilizes a single core and
all four of the available cores. The results can be seen in 
\ref{fig:palloc_multicore}. As expected, the amount of L2 cache 
available to the model does not result in any significant changes in 
performance. This can also be seen by the L2 miss rates of the model 
seen in \ref{fig:palloc_multicore_l2missrate}. As more cache space 
becomes avaiable, the percentage of L2 misses decreases, which doesn't 
affect the real-time performance of the DNN. As such, we find that a 
single model is not sensitive to the shared L2 cache.

We also co-schedule multiple models to see how cache partitioning affects
interference between them. Each model uses an equal number of colors, and
are thus given equal amounts of L2 cache space, and there is no overlap
in the colors used, so no L2 cache is shared between models. As can be 
seen in \ref{fig:palloc_multimodel}, the performance of the models 
remains the same as when no cache partitioning was employed. Based of 
these results, we find that contention for the L2 cache is not the main
source of interference between multiple DNN models.

Finally, we run a single model alongside memory intensive co-runners 
again, but ensure that each task is given equal amounts of L2 cache space.
For this, we assign two colors to each task, regardless of the number of
tasks being scheduled. As a result, the entire cache is not utilized in
experiments with less than four tasks. The results can be seen in Figure
\ref{fig:palloc_bandwidth_exectime}. With zero or one co-runners, the DNN
performance doesnt't change relative to when the cache wasn't partitioned.
However, we find that with two or more co-runners, DNN performance becomes
worse. In the case of two co-runners, average control loop time increases
by 66.31 ms, or 26\%, and with three co-runners, the average time
is 265.64 ms, of 49\%, greater.
At the same time, though, the L2 miss rates of the model don't correlate 
to the timing increases seen, as is shown in 
\ref{fig:palloc_bandwidth_l2missrate}. Instead, we find that the number
of bus accesses increases when cache partitions are used, and they mirror
the timing increases of the model. This can be seen in Figure 
\ref{fig:palloc_bandwidth_bus}. These additional bus accesses are
most likely caused by the L2 prefetcher, as it can generate additional bus
accesses that aren't counted as L2 misses. As a result, we find that cache
partitioning is detrimental to the performance of the DNN so long as the 
L2 prefetcher is enabled (we have not yet found a way to disable it on
the Pi 3). Furthermore, no cache sensitivity was displayed by the model,
thus showing that the performance of the L2 cache is not vital for DNN 
performance when co-runners are introduced.

By partitioning the shared L2 cache, we find that no noticable 
improvements are gained and that performance remains consistent, or 
decreases, depending on the circumstances. In all experiements, the DNN
shows no sensitivity to the shared L2 cache. As a result, we conclude 
that cache partitioning is not an effective isolation mechanism, and that 
the performance of the shared DRAM controller is of greater importance 
to the real-time efficiacy of the DeepPicar.

### 2018-02-01 paper/evaluation.tex
cache partition section as a separate file.
for now it is not included.

@@
 (3) the effect of co-scheduling multiple deep neural network models,
-(4) the effect of co-scheduling memory bandwidth intensive co-runners,
 and 
-(5) the effect of partitioning the shared L2 cache.
+(4) the effect of co-scheduling memory bandwidth intensive co-runners.
+%% and
+%% (5) the effect of partitioning the shared L2 cache.
 

### 2018-02-01 benchmark_timings_Pi.sh
Update scripts for Pi3
Also removed the RT scripts for both the test-model and benchmark
experiements.

@@ -13,7 +13,7 @@ echo "The output files can be found in: logs/test-model/$identification"
 
 echo 2-parallel-single-core
 ./bandwidth -a write -m 16384 -t 10000 -c 1 &
-perf stat -o logs/test-model-and-benchmark/$identification/1-sc-bench_1-sc-model_cpu0_perf.txt -e cache-misses,cache-references,LLC-loads,LLC-load-misses,LLC-stores,LLC-store-misses taskset -c 0 python test-model.py 1 > logs/test-model-and-benchmark/$identification/1-sc-bench_1-sc-model_cpu0.txt
+sudo chrt -f 99 perf stat -o logs/test-model-and-benchmark/$identification/1-sc-bench_1-sc-model_cpu0_perf.txt -e cache-misses,cache-references,LLC-loads,LLC-load-misses,r19,rc2 taskset -c 0 python test-model.py 1 > logs/test-model-and-benchmark/$identification/1-sc-bench_1-sc-model_cpu0.txt
 jobs -p | xargs kill
 echo end 2-parallel-single-core
 sleep 300
@@ -21,7 +21,7 @@ sleep 300
 echo 3-parallel-single-core
 ./bandwidth -a write -m 16384 -t 10000 -c 1 &
 ./bandwidth -a write -m 16384 -t 10000 -c 2 &
-perf stat -o logs/test-model-and-benchmark/$identification/2-sc-bench_1-sc-model_cpu0_perf.txt -e cache-misses,cache-references,LLC-loads,LLC-load-misses,LLC-stores,LLC-store-misses taskset -c 0 python test-model.py 1 > logs/test-model-and-benchmark/$identification/2-sc-bench_1-sc-model_cpu0.txt
+sudo chrt -f 99 perf stat -o logs/test-model-and-benchmark/$identification/2-sc-bench_1-sc-model_cpu0_perf.txt -e cache-misses,cache-references,LLC-loads,LLC-load-misses,r19,rc2 taskset -c 0 python test-model.py 1 > logs/test-model-and-benchmark/$identification/2-sc-bench_1-sc-model_cpu0.txt
 jobs -p | xargs kill
 echo end 3-parallel-single-core
 sleep 300
@@ -30,13 +30,13 @@ echo 4-parallel-single-core
 ./bandwidth -a write -m 16384 -t 10000 -c 1 &
 ./bandwidth -a write -m 16384 -t 10000 -c 2 &
[…]

### 2018-02-01 test-model_timings_Pi_PALLOC.sh

@@ -1,0 +1,90 @@
+timestamp=$(date +%y-%m-%d_%H-%M-%S)
+
+echo "Which board is this test being run on?  Input a string and press [ENTER]"
+read board
+
+identification=$board
+identification+="_"
+identification+=$timestamp
+
+mkdir -p logs/test-model/$identification
+echo "This test will take approximately 45 minutes to complete"
+echo "The output files can be found in: logs/test-model/$identification"
+
+echo single-core
+echo $$ > /sys/fs/cgroup/palloc/part1/tasks
+chrt -f 99 perf stat -o logs/test-model/$identification/single-core_perf.txt -e cache-misses,cache-references,LLC-loads,LLC-load-misses,LLC-stores,LLC-store-misses taskset -c 0 python test-model.py 1 > logs/test-model/$identification/single-core.txt &
+wait
+echo end single-core
+sleep 300
+
+echo 2-parallel-single-core
+echo $$ > /sys/fs/cgroup/palloc/part1/tasks
+chrt -f 99 perf stat -o logs/test-model/$identification/2-single-core_cpu0_perf.txt -e cache-misses,cache-references,LLC-loads,LLC-load-misses,LLC-stores,LLC-store-misses taskset -c 0 python test-model.py 1 > logs/test-model/$identification/2-single-core_cpu0.txt &
+
+echo $$ > /sys/fs/cgroup/palloc/part2/tasks
+chrt -f 99 perf stat -o logs/test-model/$identification/2-single-core_cpu1_perf.txt -e cache-misses,cache-references,LLC-loads,LLC-load-misses,LLC-stores,LLC-store-misses taskset -c 1 python test-model2.py 1 > logs/test-model/$identification/2-single-core_cpu1.txt &
+wait
+echo end 2-parallel-single-core
+sleep 300
+
+echo 3-parallel-single-core
+echo $$ > /sys/fs/cgroup/palloc/part1/tasks
+chrt -f 99 perf stat -o logs/test-model/$identification/3-single-core_cpu0_perf.txt -e cache-misses,cache-references,LLC-loads,LLC-load-misses,LLC-stores,LLC-store-misses taskset -c 0 python test-model.py 1 > logs/test-model/$identification/3-single-core_cpu0.txt &
+
+echo $$ > /sys/fs/cgroup/palloc/part2/tasks
[…]

### 2018-02-01 test-model_timings_Pi_RT.sh

@@ -1,68 +1,0 @@
-timestamp=$(date +%y-%m-%d_%H-%M-%S)
-
-echo "Which board is this test being run on?  Input a string and press [ENTER]"
-read board
-
-identification=$board
-identification+="_"
-identification+=$timestamp
-
-mkdir -p logs/test-model/$identification
-echo "This test will take approximately 45 minutes to complete"
-echo "The output files can be found in: logs/test-model/$identification"
-
-echo single-core
-sudo chrt -f 99 perf stat -o logs/test-model/$identification/single-core_perf.txt -e cache-misses,cache-references,LLC-loads,LLC-load-misses,LLC-stores,LLC-store-misses taskset -c 0 python test-model.py 1 > logs/test-model/$identification/single-core.txt &
-wait
-echo end single-core
-sleep 300
-
-echo 2-parallel-single-core
-sudo chrt -f 99 perf stat -o logs/test-model/$identification/2-single-core_cpu0_perf.txt -e cache-misses,cache-references,LLC-loads,LLC-load-misses,LLC-stores,LLC-store-misses taskset -c 0 python test-model.py 1 > logs/test-model/$identification/2-single-core_cpu0.txt &
-sudo chrt -f 99 perf stat -o logs/test-model/$identification/2-single-core_cpu1_perf.txt -e cache-misses,cache-references,LLC-loads,LLC-load-misses,LLC-stores,LLC-store-misses taskset -c 1 python test-model2.py 1 > logs/test-model/$identification/2-single-core_cpu1.txt &
-wait
-echo end 2-parallel-single-core
-sleep 300
-
-echo 3-parallel-single-core
-sudo chrt -f 99 perf stat -o logs/test-model/$identification/3-single-core_cpu0_perf.txt -e cache-misses,cache-references,LLC-loads,LLC-load-misses,LLC-stores,LLC-store-misses taskset -c 0 python test-model.py 1 > logs/test-model/$identification/3-single-core_cpu0.txt &
-sudo chrt -f 99 perf stat -o logs/test-model/$identification/3-single-core_cpu1_perf.txt -e cache-misses,cache-references,LLC-loads,LLC-load-misses,LLC-stores,LLC-store-misses taskset -c 1 python test-model2.py 1 > logs/test-model/$identification/3-single-core_cpu1.txt &
[…]

### 2018-02-01 test-model_timings_Pi_parts.sh

@@ -1,0 +1,68 @@
+timestamp=$(date +%y-%m-%d_%H-%M-%S)
+
+echo "Which board is this test being run on?  Input a string and press [ENTER]"
+read board
+
+identification=$board
+identification+="_"
+identification+=$timestamp
+
+mkdir -p logs/test-model/$identification
+echo "This test will take approximately 45 minutes to complete"
+echo "The output files can be found in: logs/test-model/$identification"
+
+echo quad-core-4parts
+echo $$ > /sys/fs/cgroup/palloc/part8/tasks
+chrt -f 99 perf stat -o logs/test-model/$identification/quad-core_4part_perf.txt -e cache-misses,cache-references,LLC-loads,LLC-load-misses,r19,rc2 taskset -c 0,1,2,3 python test-model.py 4 > logs/test-model/$identification/quad-core_4part.txt &
+wait
+echo end quad-core
+sleep 300
+
+echo single-core-4parts
+echo $$ > /sys/fs/cgroup/palloc/part8/tasks
+chrt -f 99 perf stat -o logs/test-model/$identification/single-core_4part_perf.txt -e cache-misses,cache-references,LLC-loads,LLC-load-misses,r19,rc2 taskset -c 0 python test-model.py 1 > logs/test-model/$identification/single-core_4part.txt &
+wait
+echo end single-core
+sleep 300
+
+echo quad-core-3parts
+echo $$ > /sys/fs/cgroup/palloc/part7/tasks
+chrt -f 99 perf stat -o logs/test-model/$identification/quad-core_3part_perf.txt -e cache-misses,cache-references,LLC-loads,LLC-load-misses,r19,rc2 taskset -c 0,1,2,3 python test-model.py 4 > logs/test-model/$identification/quad-core_3part.txt &
+wait
+echo end quad-core
+sleep 300
+
+echo single-core-3parts
+echo $$ > /sys/fs/cgroup/palloc/part7/tasks
+chrt -f 99 perf stat -o logs/test-model/$identification/single-core_3part_perf.txt -e cache-misses,cache-references,LLC-loads,LLC-load-misses,r19,rc2 taskset -c 0 python test-model.py 1 > logs/test-model/$identification/single-core_3part.txt &
+wait
+echo end single-core
+sleep 300
+
+echo quad-core-2parts
+echo $$ > /sys/fs/cgroup/palloc/part5/tasks
[…]

### 2018-02-01 test-model_timings_x2.sh
Updates scripts for the UP and TX2

@@ -12,49 +12,50 @@ echo "This test will take approximately 45 minutes to complete"
 echo "The output files can be found in: logs/test-model/$identification"
 
 echo single-core
-taskset -c 0 python test-model.py 1 > logs/test-model/$identification/single-core.txt &
+sudo chrt -f 99 perf stat -o logs/test-model/$identification/single-core_perf.txt -e r16,r17 taskset -c 0 python test-model.py 1 > logs/test-model/$identification/single-core.txt &
 wait
 echo end single-core
 
 echo 2-parallel-single-core
-taskset -c 0 python test-model.py 1 > logs/test-model/$identification/2-single-core_cpu0.txt &
-taskset -c 3 python test-model.py 1 > logs/test-model/$identification/2-single-core_cpu3.txt &
+sudo chrt -f 99 perf stat -o logs/test-model/$identification/2-single-core_cpu0_perf.txt -e r16,r17 taskset -c 0 python test-model.py 1 > logs/test-model/$identification/2-single-core_cpu0.txt &
+sudo chrt -f 99 perf stat -o logs/test-model/$identification/2-single-core_cpu3_perf.txt -e r16,r17 taskset -c 3 python test-model.py 1 > logs/test-model/$identification/2-single-core_cpu3.txt &
 wait
 echo end 2-parallel-single-core
 
 echo 3-parallel-single-core
-taskset -c 0 python test-model.py 1 > logs/test-model/$identification/3-single-core_cpu0.txt &
-taskset -c 3 python test-model.py 1 > logs/test-model/$identification/3-single-core_cpu3.txt &
-taskset -c 4 python test-model.py 1 > logs/test-model/$identification/3-single-core_cpu4.txt &
+sudo chrt -f 99 perf stat -o logs/test-model/$identification/3-single-core_cpu0_perf.txt -e r16,r17 taskset -c 0 python test-model.py 1 > logs/test-model/$identification/3-single-core_cpu0.txt &
+sudo chrt -f 99 perf stat -o logs/test-model/$identification/3-single-core_cpu3_perf.txt -e r16,r17 taskset -c 3 python test-model.py 1 > logs/test-model/$identification/3-single-core_cpu3.txt &
[…]

### 2018-02-01 test-model_timings_x2_cpu.sh

@@ -12,49 +12,49 @@ echo "This test will take approximately 45 minutes to complete"
 echo "The output files can be found in: logs/test-model/$identification"
 
 echo single-core
-CUDA_VISIBLE_DEVICES='' taskset -c 0 python test-model.py 1 > logs/test-model/$identification/single-core.txt &
+sudo CUDA_VISIBLE_DEVICES='' chrt -f 99 ./perf stat -o logs/test-model/$identification/single-core_perf.txt -e r16,r17 taskset -c 0 python test-model.py 1 > logs/test-model/$identification/single-core.txt &
 wait
 echo end single-core
 
 echo 2-parallel-single-core
-CUDA_VISIBLE_DEVICES='' taskset -c 0 python test-model.py 1 > logs/test-model/$identification/2-single-core_cpu0.txt &
-CUDA_VISIBLE_DEVICES='' taskset -c 3 python test-model.py 1 > logs/test-model/$identification/2-single-core_cpu3.txt &
+sudo CUDA_VISIBLE_DEVICES='' chrt -f 99 ./perf stat -o logs/test-model/$identification/2-single-core_cpu0_perf.txt -e r16,r17 taskset -c 0 python test-model.py 1 > logs/test-model/$identification/2-single-core_cpu0.txt &
+sudo CUDA_VISIBLE_DEVICES='' chrt -f 99 ./perf stat -o logs/test-model/$identification/2-single-core_cpu3_perf.txt -e r16,r17 taskset -c 3 python test-model2.py 1 > logs/test-model/$identification/2-single-core_cpu3.txt &
 wait
 echo end 2-parallel-single-core
 
 echo 3-parallel-single-core
-CUDA_VISIBLE_DEVICES='' taskset -c 0 python test-model.py 1 > logs/test-model/$identification/3-single-core_cpu0.txt &
-CUDA_VISIBLE_DEVICES='' taskset -c 3 python test-model.py 1 > logs/test-model/$identification/3-single-core_cpu3.txt &
-CUDA_VISIBLE_DEVICES='' taskset -c 4 python test-model.py 1 > logs/test-model/$identification/3-single-core_cpu4.txt &
[…]

### 2018-02-01 paper/overview.tex
..

@@
 \section{DeepPicar Overview}
-
-In this section, we provide an overview of our DeepPicar platform. In
-developing DeepPicar, one of our primary goals is to faithfully
-replicate NVIDIA's DAVE-2 system on a smaller scale with using a low cost
-multicore platform, the Raspberry Pi 3. Because the Raspberry Pi 3's
-computing performance is much lower than that of the DRIVE PX platform
-used in DAVE-2, we are interested in if, and how, we can process
-computationally expensive neural network operations in
-real-time. Specifically, inferencing (forward pass processing)
-operations must be completed within each control period
-duration---e.g., a WCET of 33.$\overline{\mbox{33}}$ ms for 30Hz control 
-frequency---locally on the Pi 3 platform, although training of the 
-network (back-propagation for weight updates) can be done offline and 
-remotely using a desktop computer.
 
@@
 \end{figure}
-
-Figure~\ref{fig:overview} shows the DeepPicar, which is comprised of a
-set of inexpensive components: a Raspberry Pi 3 Single Board Computer
-(SBC), a Pololu DRV8835 motor driver, a Playstation Eye webcam, a
-battery, and a 1:24 scale RC car. Table~\ref{tbl:carbom} shows the
-estimated cost of the system.
 
@@
 
+In this section, we provide an overview of our DeepPicar platform.
+In developing DeepPicar, one of our primary goals is to faithfully
+replicate NVIDIA's DAVE-2 system on a smaller scale with using a low cost
+multicore platform, the Raspberry Pi 3. Because the Raspberry Pi 3's
+computing performance is much lower than that of the DRIVE PX platform
+used in DAVE-2, we are interested in if, and how, we can process
+computationally expensive neural network operations in
+real-time. Specifically, inferencing (forward pass processing)
+operations must be completed within each control period
+duration---e.g., a WCET of 33.$\overline{\mbox{33}}$ ms for 30Hz control 
+frequency---locally on the Pi 3 platform, although training of the 
+network (back-propagation for weight updates) can be done offline and 
+remotely using a desktop computer.
+Figure~\ref{fig:overview} shows the DeepPicar, which is comprised of a
+set of inexpensive components: a Raspberry Pi 3 Single Board Computer
+(SBC), a Pololu DRV8835 motor driver, a Playstation Eye webcam, a
+battery, and a 1:24 scale RC car. Table~\ref{tbl:carbom} shows the
+estimated cost of the system.
+
 \begin{figure}[h]
   \centering
-  \includegraphics[width=0.7\textwidth]{figs/architecture}
+  \includegraphics[width=0.65\textwidth]{figs/architecture}
   \caption{DeepPicar's neural network architecture: 8 layers (5
     convolutional, 3 fully-connected layers), 27 million connections,
-    250K parameters. The architecture is almost identical to the one 
+    250K parameters. The CNN architecture is identical to the one 
 	used in NVIDIA's real self-driving car~\cite{Bojarski2016}.}
@@
 For the neural network architecture, we implement
-NVIDIA DAVE-2's convolutional neural network (CNN) using TensorFlow,
-based on ~\cite{deeptesla}. Note, however, that the CNN model
-in~\cite{deeptesla} was considerably bigger than NVIDIA's CNN as it contained an
-additional fully-connected layer of approximately 1.3M additional
-parameters. We eliminated the additional layer to faithfully recreate
-the NVIDIA's original CNN model.
+NVIDIA DAVE-2's convolutional neural network (CNN) using a TensorFlow
+based CNN model in ~\cite{deeptesla}. Note, however, that the CNN model
+in~\cite{deeptesla} was considerably bigger than NVIDIA's CNN model as
+it contained an additional fully-connected layer of approximately 1.3M
+additional parameters. We eliminated the additional layer to
+faithfully recreate the NVIDIA's original CNN model.
 %% The main difference is that we do not utilize a normalization layer, and 
@@
 output. Figure~\ref{fig:architecture} shows the network architecture
-used in this paper, which
-is comprised of 8 layers, 250K parameters, and about 27 million
-connections. It is identical to NVIDIA DAVE-2's CNN except that we do
-not apply a normalization layer as it does not include trainable
-parameters and its computational demand is minimal.
+used in this paper, which is comprised of 8 layers, 250K parameters,
+and about 27 million connections as in NVIDIA DAVE-2's. Note, however,
+that we did not apply the normalization mentioned
+in~\cite{Bojarski2016}, as it does not include trainable parameters
+and its computational demand with respect to the overall CNN
+processing is minimal.
 
@@
   \centering
-  \includegraphics[width=.65\textwidth]{figs/track_new2}
+  \includegraphics[width=.5\textwidth]{figs/track_new2}
   \caption{One of the custom tracks used for training/testing.}
@@
 To collect the training data, a human pilot manually drives the RC car
-on a small track we created (Figure~\ref{fig:track}) to record
+on tracks we created (Figure~\ref{fig:track}) to record
 timestamped videos and contol commands. The stored data is then copied 
-to a desktop computer, which is equipped with a NVIDIA GTX 1060 GPU, 
+to a desktop computer, which is equipped with a NVIDIA GTX 1050ti GPU, 
 where we train the network to accelerate training speed. 
 For comparison, training the network on the Raspbeery Pi 3 takes
-approximately 4 hours, whereas it takes only about 5 minutes on the
-desktop computer using the GTX 1060 GPU.
+approximately 4 hours, whereas it takes only about 4 minutes on the
+desktop computer using the GTX 1050ti GPU.
 
@@
 
-Another issue we observed in training/testing the network, which could
-affect the network performance, is camera latency. In DeepPicar's
+Other issues we observed in training/testing the network, which
+can affect the prediction accuracy of the CNN, are camera and actuator
+(motor) control latencies. In DeepPicar's
 context, the camera latency is from the time the camera
@@
 digitized image data. This time can be noticable depending on the camera 
-used and the performance of the Pi. We experimentally measured the camera
-latency and found it to be around 50-100 ms. This is higher than the 
-latency of human perception, which is known to be as fast as 13 
-ms~\cite{ThomasBurger2015}. Higher camera latency could negatively 
-affect control performance, because the DNN would analyze stale scenes. 
-In the future, we plan to identify and use low-latency cameras.
-
+used and the performance of the Pi. Higher camera latency could
+negatively affect control performance, because the DNN would analyze
+stale scenes. Actuator (motor) control latency also can be a
+factor, because moving the sterring motor to a desired position takes
+time. 
+In our platform, the combined latency is measured to be around
+50 ms. In other words, in the recorded video and control data, a
+control action (left, center, right) taken by the CNN appears to be
+applied 50 ms later. Considering that camera's framerate is 30Hz
+(33.3ms/frame), this is about two frames after the control action.
+%% We experimentally measured the camera
+%% latency and found it to be around 50-100 ms.
+If this value is too high, control performance may suffer. 
+For reference, the latency of human perception is known to be as fast
+as 13 ms~\cite{ThomasBurger2015}.
 % https://www.pubnub.com/blog/2015-02-09-how-fast-is-realtime-human-perception-and-technology/
@@
 several different tracks we trained. In our testing, we found that the
-DeepPicar could remain on a track for over 10
-minutes~\footnote{A collection of self-driving videos can be found at: \url{https://photos.app.goo.gl/ce93sU7jPk4ywO8u2}}.
+DeepPicar could remain on a fairly challenging track for over 10
+minutes at a moderate speed (50\% throttle), at which point we stop
+the experiment, and more than one minute at a higher speed (75\%
+throttle)~\footnote{A collection of self-driving videos can be found
+  at: \url{https://photos.app.goo.gl/ce93sU7jPk4ywO8u2}}. Running at
+higher speed is inherently more challenging because the CNN controller
+has less time to recover from mistakes (bad predictions).

### 2018-02-01 palloc-setup.sh
Remove debug level from palloc setup
Caused additional overhead that reduced performance

@@ -24,4 +24,4 @@ echo 0-7 > /sys/fs/cgroup/palloc/part8/palloc.bins #Multicore 4c
 
 #Enable PALLOC and set debug level
 echo 1 > /sys/kernel/debug/palloc/use_palloc
-echo 2 > /sys/kernel/debug/palloc/debug_level
+#echo 2 > /sys/kernel/debug/palloc/debug_level

### 2018-02-01 paper/related.tex
..

@@
 
-While we showed that a low-cost embedded platform such as the 
-Raspberry Pi 3 can successfully implement CNN-based autonomous driving
-in real-time, it should be noted that the CNN we used here is relatively
-small and recent state-of-the-art CNNs tend to be much larger.
+In this paper, we demonstrated that a low-cost embedded platform such
+as the  Raspberry Pi 3 can successfully run a CNN-based autonomous
+driving control system in real-time. However, it should be noted that
+the CNN we used here is relatively small compared to recent
+state-of-the-art CNNs, which are increasingly larger and deeper.
 For example, the CNN based object detector models evaluated in
@@
 platforms will certainly make processing of these complex CNNs faster,
-another actively investigate approrach is to reduce the required
+another actively investigated approrach is to reduce the required
 computational complexity itself.
@@
 
-We believe these reseach can make complex DNNs more accessible for
-real-time embedded systems.
+These reseach efforts are expected to make complex DNNs more
+accessible for real-time embedded systems.
 

### 2018-02-01 paper/overview.tex
Add side-by-side pictures of tracks
revision

@@
   \centering
-  \includegraphics[width=.5\textwidth]{figs/track_new2}
-  \caption{One of the custom tracks used for training/testing.}
+%% <<<<<<< HEAD
+%%   \includegraphics[width=.5\textwidth]{figs/track_new2}
+%%   \caption{One of the custom tracks used for training/testing.}
+%%   \label{fig:track}
+%% =======
+  \begin{subfigure}{0.4\textwidth}
+    \includegraphics[width=\textwidth]{figs/track_new2}
+    \caption{Track 1.}
     \label{fig:track}
+  \end{subfigure}
+  \hfill
+  \begin{subfigure}{0.52\textwidth}
+    \includegraphics[width=\textwidth]{figs/track-2}
+    \caption{Track 2.}
+    \label{fig:track2}
+  \end{subfigure}
+  \caption{Custom tracks used for training/testing}
+  \label{fig:tracks}
+%% >>>>>>> f3814396d8701264f2439b42fbe8bbb608c41452
 \end{figure}
@@
 If this value is too high, control performance may suffer.
+Our initial prototype suffered this problem as the observed camera and
+control latency was as high as 300 ms, which negatively affected
+control performance.
 For reference, the latency of human perception is known to be as fast
@@
 higher speed is inherently more challenging because the CNN controller
-has less time to recover from mistakes (bad predictions).
+has less time to recover from mistakes (bad predictions).  Also, we
+find that the prediction accuracy is significantly affected by the
+quality of training data as well as various environmental factors such
+as the lighting condition. We plan to investigate more systematic ways
+to improve the CNN's prediction accuracies.

### 2018-02-01 paper/overview.tex
..

@@
 platform that can precisely control the car's steering angle.
-We would like to stress, however, that the use of different RC car
-platforms has no impact on the computational 
-aspects of the system, and that our main focus of this study is
-not in improving network accuracy but in closely replicating the
-DAVE-2's network architecture and its real-time characteristics.
 
@@
 to improve the CNN's prediction accuracies.
+
+We would like to stress, however, that the the issues related to the
+CNN's accuracies have no impact on the \emph{computational 
+aspects of the system}, and that our main focus of this study is
+not in improving network accuracy but in closely replicating the
+DAVE-2's network architecture and studying its real-time
+characteristics.

### 2018-02-01 paper/related.tex
related and future

@@
 
-These reseach efforts are expected to make complex DNNs more
-accessible for real-time embedded systems.
+These research efforts are expected to make complex DNNs more
+accessible for real-time embedded systems. We plan to investigate the
+feasibility of these approaches in the context of DeepPicar so that we
+can use even more resource constrained micro-controller class
+computing platforms instead of Raspberry Pi3.
 

### 2018-02-01 paper/intro.tex
intro revision

@@
 % predictability
-First, we find that DNN inferencing is, in a sense, an ideal real-time
-workload as the amount of computation needed to compete a single
-inference is fixed at the DNN architecture design time and does not
-change at runtime over different inputs (e.g., different image frame).
-This means that it is feasible to achieve highly predictable and
-tightly bounded worst-case execution time estimation.
+First, we find that DNN inferencing is highly predictable---from the
+timing perspective---as the amount of computation needed to complete a
+single inference is fixed at the DNN architecture design time and does
+not change at runtime over different inputs (e.g., different image
+frames). This predictable timing behavior is obviously a desirable
+property for real-time systems.
+%% , which gives another reason for the
+%% real-time community to be interested in the use of DNNs.
 
 % performance on low-cost embedded multicore
-Second, we find that real-time processing of DeepPicar's DNN-based
+Second, we find that real-time processing of DeepPicar's CNN-based
 control loop is feasible on today's embedded computing platforms, even 
@@
 cores of the Raspberry Pi 3's quad-core CPU (w/o using its GPU), and can
-do so 100\% of the time; we can achieve 20 Hz control frequency using
+do so 100\% of the time; or we can achieve 20 Hz control frequency using
 just one core of the Raspberry Pi 3's quad-core CPU.
@@
 In particular, we find that DNN inferencing is highly sensitive to
-memory performance inference as we observe up to an 11.6X control loop
-execution time increase, when memory performance intensive
+memory performance interference as we observe up to an 11.6X control
+loop execution time increase, when memory performance intensive
 applications are co-scheduled on idle CPU cores.
-We find, however, that DNN inferencing is not sensitive to cache
-space; we show that applying cache partitioning has negligible impact
-in providing timing isolation on consolidated scenarios. This finding
+On the other hand, we find that DNN inferencing is not sensitive to
+cache space.
+%% ; we show that applying cache partitioning has negligible impact
+%% in providing timing isolation on consolidated scenarios.
+This finding
 suggests that in order to guarantee real-time performance of DNN-based
-control applications, solutions to guarantee its memory performance
-are much more important.
+control applications, solutions to guarantee their memory performance
+are more important.
 
@@
 \begin{itemize}
-  \item We provide the design and implementation of a low-cost
-    autonomous car testbed, DeepPicar, which is based on a
-    state-of-the-art end-to-end deep learning, as
+  \item We present DeepPicar, a low-cost autonomous car testbed, which
+    is based on a state-of-the-art end-to-end deep learning, as 
     open-source~\footnote{The source code, build 
@@
 
-  \item We systematically analyze the characteristics of DeepPicar's
+  \item We systematically analyze real-time characteristics of DeepPicar's
     DNN inferencing workload on a number of representative
-    contemporary embedded computing platforms from the perspective of
-    real-time systems. In particular, to the best of our knowledge,
-    our work is first to investigate the impacts of shared resource
-    contention on DNN inferencing workloads in the context of
-    real-time systems.
+    contemporary embedded computing platforms. To the
+    best of our knowledge, no prior work investigated the
+    impacts of shared resource contention on DNN inferencing workloads
+    in the context of real-time systems.
 
@@
 
-The remaining sections of the paper are as follows: Section II 
+The remaining sections of the paper are as follows.
+Section~\ref{sec:background}
 provides a background of applications in autonomous driving and related works.
- Section III gives an overview of the DeepPicar platform, including the 
-high-level system and the methods used for training and inference. 
-Section IV presents our evaluation of the platform and how different 
-factors can affect performance. Section V gives a comparison between 
+Section~\ref{sec:overview} gives an overview of the DeepPicar
+platform, including the high-level system and the methods used for
+training and inference. Section~\ref{sec:evaluation} presents our
+evaluation of the platform and how different factors can affect
+performance. Section~\ref{sec:comparison} gives a comparison between  
 the Raspberry Pi 3 and other embedded computing platforms to 
-determine their suitably for autonomous driving research. The paper finishes with 
-conclusions in Section VI.
+determine their suitably for autonomous driving research.
+We discuss more detailed training methods and system-level issues
+in~\ref{sec:discussion}. We review related work in
+Section~\ref{sec:related} and conclude in
+Section~\ref{sec:conclusion}.

### 2018-02-02 paper/background.tex
Minor changes to paper

@@
 simplifies the standard control pipeline by applying deep neural
-networks to directly produce control outputs from senor
+networks to directly produce control outputs from sensor
 inputs~\cite{Levine2016}. Figure~\ref{fig:end-to-end-control}
@@
 %% autonomous vehicle must satisfy two often conflicting goals:
-The platform must provide 
+(1) The platform must provide 
 enough computing capacity for real-time processing of computationally
-expensive AI workloads (deep neural networks);
-The platform must also satisfy various
+expensive AI workloads (deep neural networks); and
+(2) The platform must also satisfy various
 constraints such as cost, size, weight, and power consumption limits.

### 2018-02-02 paper/comparison.tex
Minor changes to paper

@@
 In this section, we compare three computing platforms---the Raspberry
-Pi 3, the Intel UP~\footnote{http://www.up-board.org/up/} and NVIDIA
+Pi 3, the Intel UP~\footnote{http://www.up-board.org/up/} and the NVIDIA
 Jetson
@@
 the experiments in Section~\ref{sec:evaluation} on each hardware
-platform and compare the results. We do not, however, repeat the cache
-partitioned experiments done on the Pi 3, as we did not believe our 
-findings warranted further experimentation on additional platforms. 
+platform and compare the results. 
+%We do not, however, repeat the cache
+%partitioned experiments done on the Pi 3, as we did not believe our 
+%findings warranted further experimentation on additional platforms. 
 For the Jetson TX2, we have two different system configurations,
 which differ in whether TensorFlow is configured to use its GPU or
-only the CPU cores. Thus, the total four system configurations are
+only the CPU cores. Thus, a total of four system configurations are
 compared.
@@
 Section~\ref{sec:eval-memhog}, we observed dramatic execution time
-increases, up to 11.6X, in Raspberry Pi 3 as we increased the number of
+increases, up to 11.5X, in Raspberry Pi 3 as we increased the number of
 co-scheduled tasks. We also observe increased control loop execution

### 2018-02-02 paper/evaluation.tex
Minor changes to paper

@@
 increase the number of co-runners, the DNN model's execution times are
-increased---by up to 11.6X---even though the DNN model is running on a
+increased---by up to 11.5X---even though the DNN model is running on a
 dedicated core (Core 0). On the other hand, the DNN model's L2
@@
 
-% \input{evaluation-cachepart}
+%% \input{evaluation-cachepart}
 
@@
 performance appears to be very sensitive to memory performance; we 
-observe up to 11.6X slowdown in DNN performance by co-scheduling 
+observe up to 11.5X slowdown in DNN performance by co-scheduling 
 synthetic memory bandwidth intensive applications on idle cores.

### 2018-02-02 paper/intro.tex
Minor changes to paper

@@
 is equipped with a NVIDIA GTX 1060 GPU. Finally, the trained network is copied
-back to the Raspberry Pi, which is then used to perform inference
-operations---locally on the Pi---in the car's main control loop in
+back to the Raspberry Pi 3, which is then used to perform inference
+operations---locally on the Pi 3---in the car's main control loop in
 real-time. For real-time control, each inference operation must
@@
 just one core of the Raspberry Pi 3's quad-core CPU.
-Other tested embedded platforms, Intel UP and NVIDIA TX2, offer even
+Other tested embedded platforms, the Intel UP and NVIDIA TX2, offer even
 better performance, and are capable of supporting deep-learning based 
@@
 In particular, we find that DNN inferencing is highly sensitive to
-memory performance interference as we observe up to an 11.6X control
+memory performance interference as we observe up to an 11.5X control
 loop execution time increase, when memory performance intensive
@@
     is based on a state-of-the-art end-to-end deep learning, as 
-    open-source~\footnote{The source code, build 
-instruction the DeepPicar
+    open-source~\footnote{The source code, and build 
+instructions for the DeepPicar
 can be found at: \url{https://github.com/heechul/picar}}, with the

### 2018-02-02 paper/overview.tex
intro revision
Minor changes to paper

@@
-\section{DeepPicar Overview}
+\section{DeepPicar Overview}\label{sec:overview}
 
@@
 In developing DeepPicar, one of our primary goals is to faithfully
-replicate NVIDIA's DAVE-2 system on a smaller scale with using a low cost
+replicate NVIDIA's DAVE-2 system on a smaller scale using a low cost
 multicore platform, the Raspberry Pi 3. Because the Raspberry Pi 3's
@@
 To collect the training data, a human pilot manually drives the RC car
-on tracks we created (Figure~\ref{fig:track}) to record
+on tracks we created (Figure~\ref{fig:tracks}) to record
 timestamped videos and contol commands. The stored data is then copied 
@@
 Once the network is trained on the desktop computer, the trained model
-is copied back to the Raspberry Pi. The network is then used
+is copied back to the Raspberry Pi 3. The network is then used
 by the car's main controller, which feeds a image frame from the web
@@
 stale scenes. Actuator (motor) control latency also can be a
-factor, because moving the sterring motor to a desired position takes
+factor, because moving the steering motor to a desired position takes
 time. 
@@
 applied 50 ms later. Considering that camera's framerate is 30Hz
-(33.3ms/frame), this is about two frames after the control action.
+(33.$\overline{\mbox{33}}$ ms/frame), this is about two frames after the 
+control action.
 %% We experimentally measured the camera
@@
 achieve a reasonable degree of accuracy, successfully navigating
-several different tracks we trained. In our testing, we found that the
-DeepPicar could remain on a fairly challenging track for over 10
-minutes at a moderate speed (50\% throttle), at which point we stop
+several different tracks we used for training. In our testing, we found 
+that the DeepPicar could remain on a fairly challenging track for over 
+10 minutes at a moderate speed (50\% throttle), at which point we stopped
 the experiment, and more than one minute at a higher speed (75\%

### 2018-02-02 paper/comparison.tex
comparison section revision.

@@
 intensive synthetic benchmarks on the DNN control loop timing.
-Figure~\ref{fig:sys_bench} shows the results.
-As discussed in 
+
+Figure~\ref{fig:sys_bench} shows the results. As discussed in 
 Section~\ref{sec:eval-memhog}, we observed dramatic execution time
@@
 2.3X execution time increase; TX2 (CPU) and TX2 (GPU) suffer up to
-2.2X and 2.1X increases, respectively. This is somewhat suprising
+2.2X and 2.1X increases, respectively. This is somewhat surprising
 because the Raspberry Pi 3's cores are in-order architecture based while
@@
 based, and that the memory intensive tasks on out-of-order cores can
-generate more memory traffic. We suspect that this is because the
+generate more memory traffic. We believe that this is because the
 memory subsystems in the Intel UP and TX2 platforms provide higher
-bandwidth and fairness than the memory subsystem of the Pi 3. However, 
-the L2 miss rates of the Pi 3 and TX2 platforms follow the same expected 
-pattern, as can be seen in Figure~\ref{fig:sys_l2miss}. Namely, the 
-percentage of L2 miss rates increases are more 
-memory intensive co-runners are introduced. This was not the case for 
-the UP board, where the model running on two cores generated the highest
-percentage of L2 cache misses. This is most likely due to the CPU 
-architecture of the UP's CPU which is Intel based, whereas both the Pi 3
-and TX2 have ARM based CPUs.
+performance than the memory subsystem of the Pi 3 as suggested from
+the measured memory bandwidth results in Table~\ref{tbl:platforms}
+('Measured Bandwidth').
 
@@
 the DNN model has dedicated access to the GPU. This is, however, a
-known characteristic of integrated CPU-GPU architectue based
+known characteristic of integrated CPU-GPU architecture based
 platforms where both CPU and GPU share the same memory
@@
 there was no contention between the CPU cores and the GPU.
+
+Figure~\ref{fig:sys_l2miss} shows the L2 miss rates of DNN control
+loop. The key takeaway of this result is that the L2 miss rate is a
+poor indicator to the execution time of the control loop. As shown
+already, Raspberry Pi 3's increased L2 misses do not fully explain
+increase in processing times. In case of Intel UP, as the number of
+interfering Bandwidth benchmark instances increases, the L2 miss rates
+even drop slightly. In case of TX2 (CPU), from solo to one co-runner,
+L2 miss rate increase almost twice, but then the execution time
+increase is only about less than 10\%. However, when there are
+co-runners, even though the L2 miss rates are largely unchanged, the
+DNN processing time increases significantly. TX2 (GPU) also shows
+similar behavior. All of these results essentially point to the fact
+that DNN inferencing workload is largely memory performance sensitive
+but not L2 cache space sensitive.
+
+%% However, the L2 miss rates of the Pi 3 and TX2 platforms follow the same expected 
+%% pattern, as can be seen in Figure~\ref{fig:sys_l2miss}. Namely, the 
+%% percentage of L2 miss rates increases are more 
+%% memory intensive co-runners are introduced. This was not the case for 
+%% the UP board, where the model running on two cores generated the highest
+%% percentage of L2 cache misses. This is most likely due to the CPU 
+%% architecture of the UP's CPU which is Intel based, whereas both the Pi 3
+%% and TX2 have ARM based CPUs.
 

### 2018-02-02 paper/intro.tex
comparison section revision.

@@
 % performance on low-cost embedded multicore
-Second, we find that real-time processing of DeepPicar's CNN-based
-control loop is feasible on today's embedded computing platforms, even 
+Second, we find that real-time processing of DeepPicar's CNN is
+feasible on today's embedded computing platforms, even  
 as inexpensive as the Raspberry Pi 3. We find that the control loop 
@@
 do so 100\% of the time; or we can achieve 20 Hz control frequency using
-just one core of the Raspberry Pi 3's quad-core CPU.
+just one core.
 Other tested embedded platforms, the Intel UP and NVIDIA TX2, offer even

### 2018-02-02 paper/overview.tex
comparison section revision.

@@
 timestamped videos and contol commands. The stored data is then copied 
-to a desktop computer, which is equipped with a NVIDIA GTX 1050ti GPU, 
+to a desktop computer, which is equipped with a NVIDIA GTX 1050 Ti GPU, 
 where we train the network to accelerate training speed. 
@@
 approximately 4 hours, whereas it takes only about 4 minutes on the
-desktop computer using the GTX 1050ti GPU.
+desktop computer using the GTX 1050 Ti GPU.
 
@@
 
-Note that although the steering angle output of the network $angle$ is
+Note that although the steering angle output of the network ($angle$) is
 a continuous real value, the RC car platform we used unfortunately

### 2018-02-02 paper/main.tex
revision
Minor changes to paper
8 -> 9 layers.

@@
   \texttt{heechul.yun@ku.edu}}
-\authorrunning{M.\,G. Bechtel, E. McEllhiney, H. Yun} %mandatory. First: Use abbreviated first/middle names. Second (only in severe cases): Use first author plus 'et. al.'
+\authorrunning{M.\,G. Bechtel, E. McEllhiney, M. Kim, H. Yun} %mandatory. First: Use abbreviated first/middle names. Second (only in severe cases): Use first author plus 'et. al.'
 
-\Copyright{Michael G. Bechtel, Elise McEllhiney, Heechul Yun}%mandatory, please use full first names. LIPIcs license is "CC-BY";  http://creativecommons.org/licenses/by/3.0/
+\Copyright{Michael G. Bechtel, Elise McEllhiney, Minje Kim, Heechul Yun}%mandatory, please use full first names. LIPIcs license is "CC-BY";  http://creativecommons.org/licenses/by/3.0/
 
@@
 car steering angles as output. DeepPicar uses almost the exact same 
-network architecture---8 layers, 27 million connections and 250K
+network architecture---9 layers, 27 million connections and 250K
 parameters---and can be trained to drive itself, in real-time, using a
@@
 \input{comparison}
+\input{discussion}
 \input{related}
@@
 \bibliography{reference}
-
-\appendix
-\input{appendix}
+%% \appendix
 \end{document}

### 2018-02-02 paper/overview.tex
8 -> 9 layers.

@@
   \centering
-  \includegraphics[width=.55\textwidth]{figs/DeepPicar_platform}
+  \includegraphics[width=.65\textwidth]{figs/DeepPicar_platform}
   \caption{DeepPicar platform.}
@@
   \includegraphics[width=0.65\textwidth]{figs/architecture}
-  \caption{DeepPicar's neural network architecture: 8 layers (5
-    convolutional, 3 fully-connected layers), 27 million connections,
+  \caption{DeepPicar's neural network architecture: 9 layers (5
+    convolutional, 4 fully-connected layers), 27 million connections,
     250K parameters. The CNN architecture is identical to the one 
@@
 output. Figure~\ref{fig:architecture} shows the network architecture
-used in this paper, which is comprised of 8 layers, 250K parameters,
+used in this paper, which is comprised of 9 layers, 250K parameters,
 and about 27 million connections as in NVIDIA DAVE-2's. Note, however,
@@
 %% =======
-  \begin{subfigure}{0.4\textwidth}
+  \begin{subfigure}{0.6\textwidth}
     \includegraphics[width=\textwidth]{figs/track_new2}
@@
   \hfill
-  \begin{subfigure}{0.52\textwidth}
+  \begin{subfigure}{0.7\textwidth}
     \includegraphics[width=\textwidth]{figs/track-2}

### 2018-02-02 paper/comparison.tex
Minor changes to comparison

@@
 poor indicator to the execution time of the control loop. As shown
-already, Raspberry Pi 3's increased L2 misses do not fully explain
-increase in processing times. In case of Intel UP, as the number of
+already, the Raspberry Pi 3's increased L2 misses do not fully explain
+increases in processing times. In case of the Intel UP, as the number of
 interfering Bandwidth benchmark instances increases, the L2 miss rates
-even drop slightly. In case of TX2 (CPU), from solo to one co-runner,
+even drop slightly. In case of the TX2 (CPU), from solo to one co-runner,
 L2 miss rate increase almost twice, but then the execution time
-increase is only about less than 10\%. However, when there are
+increase is a little less than 10\%. However, when there are
 co-runners, even though the L2 miss rates are largely unchanged, the
-DNN processing time increases significantly. TX2 (GPU) also shows
+DNN processing time increases significantly. The TX2 (GPU) also shows
 similar behavior. All of these results essentially point to the fact

### 2018-02-02 paper/evaluation.tex
Add bus accesses to Figure 9

--- 2018-02-01 paper/evaluation.tex
@@
 In other words, the DNN model's execution increase is not fully
-explained by increased contention in the L2 cache space.
-Instead, the high memory bandwidth pressure from the co-scheduled
+explained by increased contention in the L2 cache space. Instead, the 
+total number of bus accesses made while the model is running increases
+dramatically, which explains why the DNN's performance decreased.
+As such, the high memory bandwidth pressure from the co-scheduled
 memory intensive benchmarks is likely the primary cause.

### 2018-02-02 paper/evaluation.tex
..

--- 2018-02-01 paper/evaluation.tex
@@
 
-Figure~\ref{fig:} shows the normalized execution time and L2 miss-rate
+Figure~\ref{fig:} shows the normalized execution time (first group) and
+L2 miss-rate (last group)
 of the DNN model running on the Core 0 as a function of the number of
-co-scheduled memory intensive synthetic benchmarks. First, as we
+co-scheduled memory intensive synthetic benchmarks. We also shows the
+normalized total system bus access count (middle group), which is
+representative of memory bandwidth usage of all the cores.
+First, as we
 increase the number of co-runners, the DNN model's execution times are
-increased---by up to 11.6X---even though the DNN model is running on a
-dedicated core (Core 0). On the other hand, the DNN model's L2
-cache-miss rates do not increase as much.
+increased exponentially---by up to 11.6X---even though the DNN model
+is running on a dedicated core (Core 0). On the other hand, the DNN
+model's L2 cache-miss rates do not increase as much.
 In other words, the DNN model's execution increase is not fully
 explained by increased contention in the L2 cache space.
-Instead, the high memory bandwidth pressure from the co-scheduled
-memory intensive benchmarks is likely the primary cause.
+Instead, the exponential increase in the total number of bus accesses,
+due to the memory-hog co-runners, appears to be more closely
+correlated with the DNN model's execution time increase. This is
+the high memory bandwidth pressure from the co-scheduled
+memory intensive benchmarks would have increased the DNN model's
+memory access latency.
 

### 2018-02-02 paper/overview.tex
..
19 page exact. typo

@@
 %% =======
-  \begin{subfigure}{0.6\textwidth}
+  \begin{subfigure}{0.55\textwidth}
     \includegraphics[width=\textwidth]{figs/track_new2}
@@
 
-We would like to stress, however, that the the issues related to the
+We would like to stress, however, that the issues related to the
 CNN's accuracies have no impact on the \emph{computational 

### 2018-02-04 picar-mini-dnn-drv8835-v2.py
use v2 dnn controller as default.

@@ -1,270 +1,0 @@
-#!/usr/bin/python
-import os
-import time
-import atexit
-import serial
-import cv2
-import math
-import numpy as np
-import sys
-from threading import Thread
-
-print ("Load TF")
-import tensorflow as tf
-import params
-model = __import__(params.model)
-import local_common as cm
-import preprocess
-
-print ("Load Model")
-sess = tf.InteractiveSession()    
-saver = tf.train.Saver()
-model_load_path = cm.jn(params.save_dir, params.model_load_file)
-saver.restore(sess, model_load_path)
-print ("Done..")
-    
-from pololu_drv8835_rpi import motors, MAX_SPEED
-
-def deg2rad(deg):
-    return deg * math.pi / 180.0
-def rad2deg(rad):
-    return 180.0 * rad / math.pi
-
-def stop():
-    global cur_speed
-    cur_speed = 0
-    motors.motor2.setSpeed(int(cur_speed))
-        
-def ffw():
-    global cur_speed
-    # cur_speed = min(MAX_SPEED, cur_speed + MAX_SPEED/10)
-    cur_speed = SET_SPEED    
-    motors.motor2.setSpeed(int(cur_speed))
-
-def rew():
-    global cur_speed        
-    # cur_speed = max(-MAX_SPEED, cur_speed - MAX_SPEED/10)
-    cur_speed = SET_SPEED
-    motors.motor2.setSpeed(-int(cur_speed))
-
-def center():
-    motors.motor1.setSpeed(0)
-def left():
-    motors.motor1.setSpeed(MAX_SPEED)
-def right():
-    motors.motor1.setSpeed(-MAX_SPEED)
-
-
-def get_control(degree):
-    if degree < 15 and degree > -15:
-        return 0
-    elif degree >= 15:
-        return 1
-    elif degree <= -15:
-        return -1
-
-def get_ctl_str(control):
-    if control == -1:
-        return "left"
-    elif control == 0:
-        return "center"
-    elif control == 1:
-        return "right"
-    else:
-        raise ValueError
-    
-def turnOff(stat):
-    stop()
-    center()
-    print "Control loop statistics"
-    print "======================="
-    print "min/avg/99pct/max (sec): {:.3f} {:.3f} {:.3f} {:.3f}".format(
-        np.min(stat), np.mean(stat), np.percentile(stat, 99), np.max(stat))
[…]

### 2018-02-04 picar-mini-dnn-drv8835.py

@@ -1,12 +1,4 @@
 #!/usr/bin/python
-from __future__ import division
-import tensorflow as tf
-import params
-model = __import__(params.model)
-import local_common as cm
-import preprocess
-
-import random
 import os
 import time
 import atexit
@@ -14,30 +6,134 @@ import serial
 import cv2
 import math
 import numpy as np
-from pololu_drv8835_rpi import motors, MAX_SPEED
 import sys
+from threading import Thread
+
+print ("Load TF")
+import tensorflow as tf
+import params
+model = __import__(params.model)
+import local_common as cm
+import preprocess
+
+print ("Load Model")
+sess = tf.InteractiveSession()    
+saver = tf.train.Saver()
+model_load_path = cm.jn(params.save_dir, params.model_load_file)
+saver.restore(sess, model_load_path)
+print ("Done..")
+    
+from pololu_drv8835_rpi import motors, MAX_SPEED
 
 def deg2rad(deg):
     return deg * math.pi / 180.0
 def rad2deg(rad):
     return 180.0 * rad / math.pi
 
-sess = tf.InteractiveSession()
-saver = tf.train.Saver()
-model_load_path = cm.jn(params.save_dir, params.model_load_file)
-saver.restore(sess, model_load_path)
+def stop():
+    global cur_speed
+    cur_speed = 0
+    motors.motor2.setSpeed(int(cur_speed))
+        
+def ffw():
+    global cur_speed
+    # cur_speed = min(MAX_SPEED, cur_speed + MAX_SPEED/10)
+    cur_speed = SET_SPEED    
+    motors.motor2.setSpeed(int(cur_speed))
+
+def rew():
+    global cur_speed        
+    # cur_speed = max(-MAX_SPEED, cur_speed - MAX_SPEED/10)
+    cur_speed = SET_SPEED
+    motors.motor2.setSpeed(-int(cur_speed))
+
+def center():
+    motors.motor1.setSpeed(0)
+def left():
+    motors.motor1.setSpeed(MAX_SPEED)
+def right():
+    motors.motor1.setSpeed(-MAX_SPEED)
+
+
+def get_control(degree):
+    if degree < 15 and degree > -15:
+        return 0
+    elif degree >= 15:
+        return 1
+    elif degree <= -15:
+        return -1
+
+def get_ctl_str(control):
+    if control == -1:
+        return "left"
+    elif control == 0:
+        return "center"
[…]

### 2018-02-04 data_shuffled_btn.py
remove btn versions

@@ -1,88 +1,0 @@
-#!/usr/bin/env python
-from __future__ import division
-
-import random
-import os
-import sys
-from collections import OrderedDict
-import cv2
-import params
-import preprocess
-
-import local_common as cm
-
-################ parameters ###############
-data_dir = params.data_dir
-epochs = params.epochs
-img_height = params.img_height
-img_width = params.img_width
-img_channels = params.img_channels
-
-purposes = ['train', 'val']
-imgs = OrderedDict()
-wheels = OrderedDict()
-for purpose in purposes:
-    imgs[purpose] = []
-    wheels[purpose] = []
-
-# load all preprocessed training images into memory
-def load_imgs():
-    global imgs
-    global wheels
-
-    for p in purposes:
-        for epoch_id in epochs[p]:
-            print 'processing and loading "{}" epoch {} into memory, current num of imgs is {}...'.format(
-                p, epoch_id, len(imgs[p]))
-
-            # vid_path = cm.jn(data_dir, 'epoch{:0>2}_front.mkv'.format(epoch_id))
-            vid_path = cm.jn(data_dir, 'out-video-{}.avi'.format(epoch_id))
-
-            assert os.path.isfile(vid_path)
-
-            frame_count = cm.frame_count(vid_path)
-            cap = cv2.VideoCapture(vid_path)
-
-            # csv_path = cm.jn(data_dir, 'epoch{:0>2}_steering.csv'.format(epoch_id))
-            csv_path = cm.jn(data_dir, 'out-key-btn-{}.csv'.format(epoch_id))
-            assert os.path.isfile(csv_path)
-
-            print "DBG:", csv_path
-            rows = cm.fetch_csv_data(csv_path)
-            print len(rows), frame_count
-            assert frame_count == len(rows)
-            yy = [[int(row['btn'])] for row in rows]
-
-            while True:
-                ret, img = cap.read()
-                if not ret:
-                    break
-
-                img = preprocess.preprocess(img)
-                imgs[p].append(img)
-
-            wheels[p].extend(yy)
-            assert len(imgs[p]) == len(wheels[p])
-
-            cap.release()
-
-def load_batch(purpose):
-    p = purpose
-    assert len(imgs[p]) == len(wheels[p])
-    n = len(imgs[p])
-    assert n > 0
-
-    ii = random.sample(xrange(0, n), params.batch_size)
-    assert len(ii) == params.batch_size
-
[…]

### 2018-02-04 model_btn.py

@@ -1,98 +1,0 @@
-#!/usr/bin/env python 
-from __future__ import division
-
-import tensorflow as tf
-import params_btn as params
-
-def weight_variable(shape):
-    initial = tf.truncated_normal(shape, stddev=0.1)
-    return tf.Variable(initial)
-
-def bias_variable(shape):
-    initial = tf.constant(0.1, shape=shape)
-    return tf.Variable(initial)
-
-def conv2d(x, W, stride):
-    return tf.nn.conv2d(x, W, strides=[1, stride, stride, 1], padding='VALID')
-
-x = tf.placeholder(tf.float32, shape=[None, params.img_height, params.img_width, params.img_channels])
-y_ = tf.placeholder(tf.int32, shape=[None, 1]) # 106 - left, 107 - center, 108 - right
-yy_ = y_ - 106
-y_oh_yy_ = tf.one_hot(yy_, 3)
-y_oh_ = tf.reshape(y_oh_yy_, [-1, 3])
-
-x_image = x
-
-# first convolutional layer
-W_conv1 = weight_variable([5, 5, 3, 24])
-b_conv1 = bias_variable([24])
-
-h_conv1 = tf.nn.relu(conv2d(x_image, W_conv1, 2) + b_conv1)
-
-# second convolutional layer
-W_conv2 = weight_variable([5, 5, 24, 36])
-b_conv2 = bias_variable([36])
-
-h_conv2 = tf.nn.relu(conv2d(h_conv1, W_conv2, 2) + b_conv2)
-
-# third convolutional layer
-W_conv3 = weight_variable([5, 5, 36, 48])
-b_conv3 = bias_variable([48])
-
-h_conv3 = tf.nn.relu(conv2d(h_conv2, W_conv3, 2) + b_conv3)
-
-# fourth convolutional layer
-W_conv4 = weight_variable([3, 3, 48, 64])
-b_conv4 = bias_variable([64])
-
-h_conv4 = tf.nn.relu(conv2d(h_conv3, W_conv4, 1) + b_conv4)
-
-# fifth convolutional layer
-W_conv5 = weight_variable([3, 3, 64, 64])
-b_conv5 = bias_variable([64])
-
-h_conv5 = tf.nn.relu(conv2d(h_conv4, W_conv5, 1) + b_conv5)
-
[…]

### 2018-02-04 picar-mini-dnn-drv8835_btn.py

@@ -1,194 +1,0 @@
-#!/usr/bin/python
-from __future__ import division
-import tensorflow as tf
-import params
-model = __import__(params.model)
-import local_common as cm
-import preprocess
-
-import random
-import os
-import time
-import atexit
-import serial
-import cv2
-import math
-import numpy as np
-import pygame
-from pololu_drv8835_rpi import motors, MAX_SPEED
-
-def deg2rad(deg):
-    return deg * math.pi / 180.0
-def rad2deg(rad):
-    return 180.0 * rad / math.pi
-
-sess = tf.InteractiveSession()
-saver = tf.train.Saver()
-model_load_path = cm.jn(params.save_dir, params.model_load_file)
-saver.restore(sess, model_load_path)
-
-motors.setSpeeds(0, 0)
-cap = cv2.VideoCapture(0)
-cap.set(3,640)
-cap.set(4,480)
-
-# ser = serial.Serial('/dev/ttyACM0', 115200, timeout=1)
-# ser = serial.Serial('/dev/ttyUSB0', 115200, timeout=1)
-fourcc = cv2.VideoWriter_fourcc(*'XVID')
-vidfile = cv2.VideoWriter('out-video.avi', fourcc, 15.0, (640,480))
-keyfile = open('out-key.csv', 'w+')
-keyfile_btn = open('out-key-btn.csv', 'w+')
-keyfile.write("ts_micro,frame,wheel\n")
-keyfile_btn.write("ts_micro,frame,btn,speed\n")
-rec_start_time = 0
-SET_SPEED = MAX_SPEED # /2 + 2*MAX_SPEED/10
-cur_speed = SET_SPEED
-print "MAX speed:", MAX_SPEED
-
-def stop():
-    global cur_speed
-    cur_speed = 0
-    motors.motor2.setSpeed(int(cur_speed))
-        
-def ffw():
-    global cur_speed
-    # cur_speed = min(MAX_SPEED, cur_speed + MAX_SPEED/10)
-    cur_speed = SET_SPEED
-    motors.motor2.setSpeed(int(cur_speed))
-
-def rew():
-    global cur_speed        
-    # cur_speed = max(-MAX_SPEED, cur_speed - MAX_SPEED/10)
-    cur_speed = SET_SPEED
-    motors.motor2.setSpeed(int(cur_speed))
-
-def center():
-    motors.motor1.setSpeed(0)
-def left():
[…]

### 2018-02-04 test-model_btn.py

@@ -1,74 +1,0 @@
-#!/usr/bin/env python 
-from __future__ import division
-
-import tensorflow as tf
-import model_btn as model
-import cv2
-import subprocess as sp
-import itertools
-import params_btn as params
-import sys
-import os
-import preprocess
-# import visualize
-import time
-import math
-
-import local_common as cm
-
-def deg2rad(deg):
-        return deg * math.pi / 180.0
-def rad2deg(rad):
-        return 180.0 * rad / math.pi
-
-sess = tf.InteractiveSession()
-saver = tf.train.Saver()
-model_load_path = cm.jn(params.save_dir, params.model_load_file)
-saver.restore(sess, model_load_path)
-
-epoch_ids = sorted(list(set(itertools.chain(*params.epochs.values()))))
-
-# epoch_ids = [6] # DBG - heechul
-
-for epoch_id in epoch_ids:
-    print '---------- processing video for epoch {} ----------'.format(epoch_id)
-    # vid_path = cm.jn(params.data_dir, 'epoch{:0>2}_front.mkv'.format(epoch_id))
-
-    vid_path = cm.jn(params.data_dir, 'out-video-{}.avi'.format(epoch_id))
-    assert os.path.isfile(vid_path)
-    frame_count = cm.frame_count(vid_path)
-    cap = cv2.VideoCapture(vid_path)
-
-    machine_steering = []
-
-    print 'performing inference...'
-    time_start = time.time()
-    for frame_id in xrange(frame_count):
-        ret, img = cap.read()
-        assert ret
-
-        prep_start = time.time()
-        img = preprocess.preprocess(img)
-
-        pred_start = time.time()
-        btn = model.y.eval(feed_dict={model.x: [img], model.keep_prob: 1.0})
-        pred_end   = time.time()
-
-        prep_time = pred_start - prep_start
-        pred_time = pred_end - pred_start
-
-        print 'pred: {} took {} ms'.format(btn, pred_time * 1000)
-
-        machine_steering.append(btn)
-
-    cap.release()
-
-    fps = frame_count / (time.time() - time_start)
-    
-    print 'completed inference, total frames: {}, average fps: {} Hz'.format(frame_count, round(fps, 1))
-
-    print "Machine Steering:", machine_steering
-
-    # print 'performing visualization...'
-    # visualize.visualize(epoch_id, machine_steering, params.out_dir,
-    #                     verbose=True, frame_count_limit=None)

### 2018-02-04 train_btn.py

@@ -1,96 +1,0 @@
-#!/usr/bin/env python 
-from __future__ import division
-
-import os
-import tensorflow as tf
-import params
-model = __import__(params.model)
-import time
-
-if params.shuffle_training:
-    import data_shuffled_btn as data
-else:
-    import data_ordered as data
-
-import local_common as cm
-
-write_summary = params.write_summary
-
-sess = tf.InteractiveSession()
-loss = tf.reduce_mean(tf.nn.softmax_cross_entropy_with_logits(logits=model.y_lg, labels=model.y_oh_))
-# loss = tf.reduce_mean(tf.square(tf.subtract(model.y_, model.y)))
-# loss = tf.reduce_mean(tf.square(tf.sub(model.y_, model.y)))
-#         + tf.add_n([tf.nn.l2_loss(v) for v in train_vars]) * L2NormConst
-train_step = tf.train.AdamOptimizer(1e-4).minimize(loss)
-
-sess.run(tf.global_variables_initializer())
-
-
-# create a summary to monitor cost tensor
-if write_summary:
-    tf.summary.scalar("loss", loss)
-
-# merge all summaries into a single op
-if write_summary:
-    merged_summary_op = tf.summary.merge_all()
-
-saver = tf.train.Saver()
-time_start = time.time()
-
-# op to write logs to Tensorboard
-if write_summary:
-    summary_writer = tf.summary.FileWriter(params.save_dir, graph=tf.get_default_graph())
-
-if params.shuffle_training:
-    data.load_imgs()
-
-    # norm_input = sess.run(model.yy_, feed_dict={model.x: txx, model.y_: tyy, model.keep_prob: 1.0})    
-    # enco_input = sess.run(model.y_oh_yy_, feed_dict={model.x: txx, model.y_: tyy, model.keep_prob: 1.0})
-    # encoded_input = sess.run(model.y_oh_, feed_dict={model.x: txx, model.y_: tyy, model.keep_prob: 1.0})    
-    # for n, e, p,y in zip(norm_input, enco_input, encoded_input, tyy):
-    #     print("N={}, E={}, E2={}, O={}".format(n, e, p, y))
-    
-for i in xrange(params.training_steps):
-    txx, tyy = data.load_batch('train')
-              
-    train_step.run(feed_dict={model.x: txx, model.y_: tyy, model.keep_prob: 0.8})
-
-    # write logs at every iteration
-    if write_summary:
[…]

### 2018-02-04 actuator-adafruit_hat.py
modular architecture to support multiple actuators.

@@ -1,0 +1,55 @@
+from Adafruit_MotorHAT import Adafruit_MotorHAT, Adafruit_DCMotor
+
+# init
+mh = Adafruit_MotorHAT(addr=0x60)
+
+def init():
+    global steering, throttle
+    steering = mh.getMotor(1)
+    throttle = mh.getMotor(4)
+    
+# throttle
+cur_speed = 0
+MAX_SPEED = 255
+
+def set_speed(speed):
+    global cur_speed
+    speed = int(MAX_SPEED * speed / 100)
+    cur_speed = min(MAX_SPEED, speed)
+    print ("speed: %d" % cur_speed)
+
+def get_speed():
+    global cur_speed    
+    return int(cur_speed * 100 / MAX_SPEED)
+
+def stop():
+    throttle.setSpeed(0)
+    steering.setSpeed(0)
+    throttle.run(Adafruit_MotorHAT.RELEASE)
+    steering.run(Adafruit_MotorHAT.RELEASE)
+    
+def ffw():
+    global cur_speed        
+    throttle.setSpeed(cur_speed)    
+    throttle.run(Adafruit_MotorHAT.FORWARD)    
+
+def rew():
+    global cur_speed        
+    throttle.setSpeed(cur_speed)    
+    throttle.run(Adafruit_MotorHAT.BACKWARD)
+
+# steering
+def center():
+    steering.setSpeed(0)
+def left():
+    steering.setSpeed(MAX_SPEED)
+    steering.run(Adafruit_MotorHAT.BACKWARD)        
+def right():
+    steering.setSpeed(MAX_SPEED)
+    steering.run(Adafruit_MotorHAT.FORWARD)
+
+
+# exit    
+def turn_off():
+    stop()
+    center()

### 2018-02-04 actuator-drv8835.py

@@ -1,0 +1,40 @@
+from pololu_drv8835_rpi import motors, MAX_SPEED
+
+# init
+def init():
+    motors.setSpeeds(0, 0)
+
+# throttle
+cur_speed = MAX_SPEED
+
+def set_speed(speed):
+
+
+    global cur_speed
+    speed = int(MAX_SPEED * speed / 100)
+    cur_speed = min(MAX_SPEED, speed)
+
+def get_speed():
+    return int(cur_speed * 100 / MAX_SPEED)
+
+def stop():
+    motors.motor2.setSpeed(0)
+        
+def ffw():
+    motors.motor2.setSpeed(cur_speed)
+
+def rew():
+    motors.motor2.setSpeed(-cur_speed)
+
+# steering
+def center():
+    motors.motor1.setSpeed(0)
+def left():
+    motors.motor1.setSpeed(MAX_SPEED)
+def right():
+    motors.motor1.setSpeed(-MAX_SPEED)
+
+# exit    
+def turn_off():
+    stop()
+    center()

### 2018-02-04 params.py

@@ -15,10 +15,14 @@ shuffle_training = True
 use_category_normal = True # if ture, center/curve images are equally selected.
 use_picar_mini = True
 
+# actuator selection
+#   "actuator-drv8835", "actuator-adafruit_hat"
+actuator="actuator-adafruit_hat"
+
 # change this to the directory that contains the source videos
 # "-5conv_3fc_noreuse" "-5conv_3fc" "-3conv_1pool" "-5conv_3fc_noreuse_nocn"
+#   model_load_file="model-5conv_3fc_noreuse.ckpt"  # <-- solved ittc building maze.
 model="model-5conv_3fc"
-# model_load_file="model-5conv_3fc_noreuse.ckpt"  # <-- solved ittc building maze. 
 model_load_file="model-5conv_3fc-home_night.ckpt"
 model_save_file=model_load_file
 save_dir = os.path.abspath('models')

### 2018-02-04 picar-mini-kbd-common.py

@@ -1,0 +1,171 @@
+#!/usr/bin/python
+import os
+import time
+import atexit
+import serial
+import cv2
+import math
+import numpy as np
+import sys
+from threading import Thread
+import params
+actuator = __import__(params.actuator)
+
+def deg2rad(deg):
+    return deg * math.pi / 180.0
+def rad2deg(rad):
+    return 180.0 * rad / math.pi
+
+def g_tick():
+    t = time.time()
+    count = 0
+    while True:
+        count += 1
+        yield max(t + count*period - time.time(),0)
+
+class Camera:
+    def __init__(self, res=(320, 240), fps=30):
+        print "Initilize camera."
+        self.cap = cv2.VideoCapture(0)
+        self.cap.set(3, res[0]) # width
+        self.cap.set(4, res[1]) # height
+        self.cap.set(5, fps)
+        self.frame = None
+        self.enabled = True
+        
+    def update(self):
+        while self.enabled:
+            ret, self.frame = self.cap.read() # blocking read.
+
+    def read_frame(self):
+        return self.frame
+
+    def shutdown(self):
+        print ("Close the camera.")
+        self.enabled = False;        
+        self.cap.release()
+
+cfg_width = 320
+cfg_height = 240
+cfg_fps = 30
+
+cam = Camera((cfg_width, cfg_height), cfg_fps)
+cam_thr = Thread(target=cam.update, args=())
+cam_thr.start()
+time.sleep(2)
+
+view_video = False
+frame_id = 0
+
+angle = 0.0
+btn   = 107
+period = 0.05 # sec (=50ms)
+
+# ser = serial.Serial('/dev/ttyACM0', 115200, timeout=1)
+# ser = serial.Serial('/dev/ttyUSB0', 115200, timeout=1)
+# fourcc = cv2.VideoWriter_fourcc(*'XVID')
+fourcc = cv2.cv.CV_FOURCC(*'XVID')
+vidfile = cv2.VideoWriter('out-video.avi', fourcc, 
+	cfg_fps, (cfg_width,cfg_height))
+
+keyfile = open('out-key.csv', 'w+')
+keyfile_btn = open('out-key-btn.csv', 'w+')
+keyfile.write("ts_micro,frame,wheel\n")
+keyfile_btn.write("ts_micro,frame,btn,speed\n")
+rec_start_time = 0
+
+default_speed = 50
+actuator.init()
+actuator.set_speed(default_speed)
[…]

### 2018-02-05 actuator-adafruit_hat.py
modular architecture for camera and other cleanup

@@ -3,15 +3,18 @@ from Adafruit_MotorHAT import Adafruit_MotorHAT, Adafruit_DCMotor
 # init
 mh = Adafruit_MotorHAT(addr=0x60)
 
-def init():
-    global steering, throttle
-    steering = mh.getMotor(1)
-    throttle = mh.getMotor(4)
-    
 # throttle
 cur_speed = 0
 MAX_SPEED = 255
 
+# Actuator API:
+#   init, stop, set/get_speed, ffw, rew, left, right, center
+def init(default_speed=50):
+    global steering, throttle
+    steering = mh.getMotor(1)
+    throttle = mh.getMotor(4)
+    set_speed(default_speed)
+    
 def set_speed(speed):
     global cur_speed
     speed = int(MAX_SPEED * speed / 100)
@@ -19,22 +22,13 @@ def set_speed(speed):
     print ("speed: %d" % cur_speed)
 
 def get_speed():
-    global cur_speed    
     return int(cur_speed * 100 / MAX_SPEED)
 
-def stop():
-    throttle.setSpeed(0)
-    steering.setSpeed(0)
-    throttle.run(Adafruit_MotorHAT.RELEASE)
-    steering.run(Adafruit_MotorHAT.RELEASE)
-    
 def ffw():
-    global cur_speed        
     throttle.setSpeed(cur_speed)    
     throttle.run(Adafruit_MotorHAT.FORWARD)    
 
 def rew():
-    global cur_speed        
     throttle.setSpeed(cur_speed)    
     throttle.run(Adafruit_MotorHAT.BACKWARD)
 
@@ -48,8 +42,10 @@ def right():
     steering.setSpeed(MAX_SPEED)
     steering.run(Adafruit_MotorHAT.FORWARD)
 
-
-# exit    
-def turn_off():
-    stop()
-    center()
+# stop    
+def stop():
+    throttle.setSpeed(0)
+    steering.setSpeed(0)
+    throttle.run(Adafruit_MotorHAT.RELEASE)
+    steering.run(Adafruit_MotorHAT.RELEASE)
+    

### 2018-02-05 camera-webcam.py

@@ -1,0 +1,42 @@
+import cv2
+from threading import Thread,Lock
+import time
+
+enabled = False
+cap = None
+frame = None
+
+# public API
+# init(), read_frame(), stop()
+
+def init(res=(320, 240), fps=30):
+    print "Initilize camera."
+    global cap, enabled,frame, cam_thr
+
+    cap = cv2.VideoCapture(0)
+
+    cap.set(3, res[0]) # width
+    cap.set(4, res[1]) # height
+    cap.set(5, fps)
+
+    # start the camera thread
+    enabled = True
+    cam_thr = Thread(target=__update, args=())
+    cam_thr.start()
+    time.sleep(0.1)
+    print ("camera init completed.")
+
+def __update():
+    global frame
+    while enabled:
+        ret, frame = cap.read() # blocking read.
+    print ("Camera thread finished...")
+    cap.release()        
+
+def read_frame():
+    return frame
+
+def stop():
+    global frame, enabled
+    print ("Close the camera.")
+    enabled = False

### 2018-02-05 params.py

@@ -4,27 +4,47 @@ from __future__ import division
 import os
 from collections import OrderedDict
 
+##########################################################
+# camera module selection
+##########################################################
+camera="camera-webcam"
+
+##########################################################
+# actuator selection
+#   "actuator-drv8835", "actuator-adafruit_hat"
+##########################################################
+actuator="actuator-adafruit_hat"
+
+##########################################################
+# model selection
+#   "model-5conv_3fc"   <-- nvidia dave-2 model
+#   "model-5conv_4fc"   <-- deeptesla model
+#
+# model checkpoint file selection
+#   "model-5conv_3fc_noreuse.ckpt"  <-- ittc building maze
+#   "model-5conv_3fc-home_night.ckpt" <-- kitchen@night
+##########################################################
+model="model-5conv_3fc"
+model_load_file="model-5conv_3fc-home_night.ckpt"
+model_save_file=model_load_file
+
+##########################################################
+# Training options
+##########################################################
 batch_size = 100
 training_steps = 2000
 img_height = 66
 img_width = 200
 img_channels = 3
-
 write_summary = True
-shuffle_training = True
-use_category_normal = True # if ture, center/curve images are equally selected.
-use_picar_mini = True
+shuffle_training = True 
+use_category_normal = True # if ture, center/curve images
+                           # are equally selected.
+use_picar_mini = True # visualization fix for picar mini
 
-# actuator selection
-#   "actuator-drv8835", "actuator-adafruit_hat"
-actuator="actuator-adafruit_hat"
-
-# change this to the directory that contains the source videos
-# "-5conv_3fc_noreuse" "-5conv_3fc" "-3conv_1pool" "-5conv_3fc_noreuse_nocn"
-#   model_load_file="model-5conv_3fc_noreuse.ckpt"  # <-- solved ittc building maze.
-model="model-5conv_3fc"
-model_load_file="model-5conv_3fc-home_night.ckpt"
-model_save_file=model_load_file
+##########################################################
+# Directories
+##########################################################
 save_dir = os.path.abspath('models')
 data_dir = os.path.abspath('epochs-conv')
 out_dir = os.path.abspath('output')

### 2018-02-05 picar-mini-kbd-common.py

@@ -7,10 +7,32 @@ import cv2
 import math
 import numpy as np
 import sys
-from threading import Thread
 import params
+
+##########################################################
+# import deeppicar's sensor/actuator modules
+##########################################################
+camera   = __import__(params.camera)
 actuator = __import__(params.actuator)
 
+##########################################################
+# global variable initialization
+##########################################################
+view_video = False
+frame_id = 0
+
+angle = 0.0
+btn   = 107
+period = 0.05 # sec (=50ms)
+
+cfg_width = 320
+cfg_height = 240
+cfg_fps = 30
+cfg_throttle = 50
+
+##########################################################
+# local functions
+##########################################################
 def deg2rad(deg):
     return deg * math.pi / 180.0
 def rad2deg(rad):
@@ -23,46 +45,22 @@ def g_tick():
         count += 1
         yield max(t + count*period - time.time(),0)
 
-class Camera:
-    def __init__(self, res=(320, 240), fps=30):
-        print "Initilize camera."
-        self.cap = cv2.VideoCapture(0)
-        self.cap.set(3, res[0]) # width
-        self.cap.set(4, res[1]) # height
-        self.cap.set(5, fps)
-        self.frame = None
-        self.enabled = True
-        
-    def update(self):
-        while self.enabled:
-            ret, self.frame = self.cap.read() # blocking read.
+def turn_off():    
+    actuator.stop()
+    camera.stop()
+    
+    keyfile.close()
+    keyfile_btn.close()
+    vidfile.release()
 
-    def read_frame(self):
-        return self.frame
-
-    def shutdown(self):
-        print ("Close the camera.")
-        self.enabled = False;        
-        self.cap.release()
-
-cfg_width = 320
-cfg_height = 240
-cfg_fps = 30
-
-cam = Camera((cfg_width, cfg_height), cfg_fps)
-cam_thr = Thread(target=cam.update, args=())
-cam_thr.start()
-time.sleep(2)
-
-view_video = False
-frame_id = 0
-
-angle = 0.0
-btn   = 107
-period = 0.05 # sec (=50ms)
+##########################################################
+# program begins
[…]

### 2018-02-05 test-camera-opencv.py

@@ -1,12 +1,19 @@
 import numpy as np
 import cv2
+import time
 
 cap = cv2.VideoCapture(0)
 
+cap.set(3, 320) # width
+cap.set(4, 240) # height
+cap.set(5, 30)
+
 while(True):
+    ts = time.time()    
     # Capture frame-by-frame
     ret, frame = cap.read()
-
+    dur = int((time.time() - ts) * 1000);
+    
     # Our operations on the frame come here
     # gray = cv2.cvtColor(frame, cv2.COLOR_BGR2GRAY)
 
@@ -14,6 +21,7 @@ while(True):
     cv2.imshow('frame',frame)
     if cv2.waitKey(1) & 0xFF == ord('q'):
         break
+    print ("%.3f %d(ms)" % (ts, dur))    
 
 # When everything done, release the capture
 cap.release()

### 2018-02-06 picar-mini-kbd-common.py
support dnn and kbd control in a single file
use argparser to support

-d (or --dnn) turn on dnn
-t XX (or --throttle XX) set the throttle value

@@ -8,6 +8,7 @@ import math
 import numpy as np
 import sys
 import params
+import argparse
 
 ##########################################################
 # import deeppicar's sensor/actuator modules
@@ -18,17 +19,19 @@ actuator = __import__(params.actuator)
 ##########################################################
 # global variable initialization
 ##########################################################
+use_dnn = False
 view_video = False
+
 frame_id = 0
 
 angle = 0.0
-btn   = 107
+btn   = ord('k') # center
 period = 0.05 # sec (=50ms)
 
 cfg_width = 320
 cfg_height = 240
 cfg_fps = 30
-cfg_throttle = 50
+cfg_throttle = 50 # 50% power.
 
 ##########################################################
 # local functions
@@ -56,21 +59,42 @@ def turn_off():
 ##########################################################
 # program begins
 ##########################################################
-if len(sys.argv) == 2:
-    cfg_throttle = int(sys.argv[1])
-    print ("Set new speed: %d" % cfg_throttle)
+parser = argparse.ArgumentParser(description='DeepPicar main')
+parser.add_argument("-d", "--dnn", help="Enable DNN", action="store_true")
+parser.add_argument("-t", "--throttle", help="throttle percent. [0-100]%", type=int)
+args = parser.parse_args()
 
-# ser = serial.Serial('/dev/ttyACM0', 115200, timeout=1)
-# fourcc = cv2.VideoWriter_fourcc(*'XVID')
-fourcc = cv2.cv.CV_FOURCC(*'XVID')
-vidfile = cv2.VideoWriter('out-video.avi', fourcc, 
-	cfg_fps, (cfg_width,cfg_height))
+if args.dnn:
+    print ("DNN is on")
+    use_dnn = True
+if args.throttle:
+    cfg_throttle = args.throttle
+    print ("throttle = %d pct" % (args.throttle))
 
+# create files for data recording
 keyfile = open('out-key.csv', 'w+')
 keyfile_btn = open('out-key-btn.csv', 'w+')
 keyfile.write("ts_micro,frame,wheel\n")
 keyfile_btn.write("ts_micro,frame,btn,speed\n")
 rec_start_time = 0
+# fourcc = cv2.VideoWriter_fourcc(*'XVID')
+fourcc = cv2.cv.CV_FOURCC(*'XVID')
+vidfile = cv2.VideoWriter('out-video.avi', fourcc, 
+	cfg_fps, (cfg_width,cfg_height))
+
+# initilize dnn model
[…]

### 2018-02-06 picar-mini-kbd-common.py
print stats only during recording.

@@ -202,7 +202,7 @@ while True:
             print "recorded 1000 frames"
             break
 
-    print ("%.3f %d %.3f %d %d(ms)" %
+        print ("%.3f %d %.3f %d %d(ms)" %
            (ts, frame_id, angle, btn, int((time.time() - ts)*1000)))
 
 

### 2018-02-06 picar-mini-kbd-common.py
minor cleanup

@@ -28,9 +28,8 @@ angle = 0.0
 btn   = ord('k') # center
 period = 0.05 # sec (=50ms)
 
-cfg_width = 320
-cfg_height = 240
-cfg_fps = 30
+cfg_cam_res = (320, 240)
+cfg_cam_fps = 30
 cfg_throttle = 50 # 50% power.
 
 ##########################################################
@@ -80,7 +79,7 @@ rec_start_time = 0
 # fourcc = cv2.VideoWriter_fourcc(*'XVID')
 fourcc = cv2.cv.CV_FOURCC(*'XVID')
 vidfile = cv2.VideoWriter('out-video.avi', fourcc, 
-	cfg_fps, (cfg_width,cfg_height))
+                          cfg_cam_fps, cfg_cam_res)
 
 # initilize dnn model
 if use_dnn == True:
@@ -98,10 +97,10 @@ if use_dnn == True:
 
 # initlaize deeppicar modules
 actuator.init(cfg_throttle)
-camera.init(res=(cfg_width,cfg_height), fps=cfg_fps)
+camera.init(res=cfg_cam_res, fps=cfg_cam_fps)
 atexit.register(turn_off)
 
-null_frame = np.zeros((cfg_width,cfg_height,3), np.uint8)
+null_frame = np.zeros((cfg_cam_res[0],cfg_cam_res[1],3), np.uint8)
 cv2.imshow('frame', null_frame)
 
 g = g_tick()

### 2018-02-06 input_kbd.py
console support.

@@ -1,0 +1,57 @@
+import os
+import time
+import atexit
+import termios, fcntl, sys, os
+import select
+
+def init():
+    fd = sys.stdin.fileno()
+    # save old state
+    flags_save = fcntl.fcntl(fd, fcntl.F_GETFL)
+    attrs_save = termios.tcgetattr(fd)
+    # make raw - the way to do this comes from the termios(3) man page.
+    attrs = list(attrs_save) # copy the stored version to update
+    # iflag
+    attrs[0] &= ~(termios.IGNBRK | termios.BRKINT | termios.PARMRK 
+                  | termios.ISTRIP | termios.INLCR | termios. IGNCR 
+                  | termios.ICRNL | termios.IXON )
+    # oflag
+    attrs[1] &= ~termios.OPOST
+    # cflag
+    attrs[2] &= ~(termios.CSIZE | termios. PARENB)
+    attrs[2] |= termios.CS8
+    # lflag
+    attrs[3] &= ~(termios.ECHONL | termios.ECHO | termios.ICANON
+                  | termios.ISIG | termios.IEXTEN)
+    termios.tcsetattr(fd, termios.TCSANOW, attrs)
+    # turn off non-blocking
+    fcntl.fcntl(fd, fcntl.F_SETFL, flags_save & ~os.O_NONBLOCK)
+    # read a single keystroke
+    return (flags_save, attrs_save)
+
+def stop(state):
+    fd = sys.stdin.fileno()
+    # restore old state
+    termios.tcsetattr(fd, termios.TCSAFLUSH, state[1])
+    fcntl.fcntl(fd, fcntl.F_SETFL, state[0])
+    
+def read_single_keypress():
+    r, w, e = select.select([sys.stdin], [], [], 0.001)
+    for s in r:
+        if s == sys.stdin:
+            input = sys.stdin.read(1)
+            return input
+    return None
+
+if __name__ == "__main__":
+    state = init()
+    
+    while True:
+        key = read_single_keypress()
+        if key == "q":
+            break
+        elif key != None:
+            print ("%s pressed\r" % key)
+        
+    stop(state)
+    

### 2018-02-06 picar-mini-kbd-common.py

@@ -10,6 +10,8 @@ import sys
 import params
 import argparse
 
+import input_kbd
+
 ##########################################################
 # import deeppicar's sensor/actuator modules
 ##########################################################
@@ -50,6 +52,7 @@ def g_tick():
 def turn_off():    
     actuator.stop()
     camera.stop()
+    input_kbd.stop()
     
     keyfile.close()
     keyfile_btn.close()
@@ -95,13 +98,16 @@ if use_dnn == True:
     saver.restore(sess, model_load_path)
     print ("Done..")
 
+# initialize kbd input 
+input_kbd.init()
+
 # initlaize deeppicar modules
 actuator.init(cfg_throttle)
 camera.init(res=cfg_cam_res, fps=cfg_cam_fps)
 atexit.register(turn_off)
 
-null_frame = np.zeros((cfg_cam_res[0],cfg_cam_res[1],3), np.uint8)
-cv2.imshow('frame', null_frame)
+# null_frame = np.zeros((cfg_cam_res[0],cfg_cam_res[1],3), np.uint8)
+# cv2.imshow('frame', null_frame)
 
 g = g_tick()
 
@@ -117,24 +123,25 @@ while True:
         
     if view_video == True:
         cv2.imshow('frame', frame)
-
-    ch = cv2.waitKey(1) & 0xFF
-
+        ch = cv2.waitKey(1) & 0xFF
+    else:
+        ch = ord(input_kbd.read_single_keypress())
+    
     if ch == ord('j'):
         actuator.left()
         print ("left")
         angle = deg2rad(-30)
-        btn   = ch
+        btn   = ord('j')
     elif ch == ord('k'):
         actuator.center()
         print ("center")
         angle = deg2rad(0)
-        btn   = ch                
+        btn   = ord('k')
     elif ch == ord('l'):
         actuator.right()
         print ("right")
         angle = deg2rad(30)
-        btn   = ch               
+        btn   = ord('l')         
     elif ch == ord('a'):
         actuator.ffw()
         print ("accel")

### 2018-02-06 picar-mini-kbd-common.py
dnn mode bug fix.

@@ -83,12 +83,18 @@ fourcc = cv2.cv.CV_FOURCC(*'XVID')
 vidfile = cv2.VideoWriter('out-video.avi', fourcc, 
                           cfg_cam_fps, cfg_cam_res)
 
+# initlaize deeppicar modules
+actuator.init(cfg_throttle)
+camera.init(res=cfg_cam_res, fps=cfg_cam_fps)
+atexit.register(turn_off)
+
 # initilize dnn model
 if use_dnn == True:
+    print ("Load TF")
     import tensorflow as tf
+    model = __import__(params.model)
     import local_common as cm
     import preprocess
-    model = __import__(params.model)
 
     print ("Load Model")
     sess = tf.InteractiveSession()    
@@ -96,11 +102,6 @@ if use_dnn == True:
     model_load_path = cm.jn(params.save_dir, params.model_load_file)
     saver.restore(sess, model_load_path)
     print ("Done..")
-
-# initlaize deeppicar modules
-actuator.init(cfg_throttle)
-camera.init(res=cfg_cam_res, fps=cfg_cam_fps)
-atexit.register(turn_off)
 
 # null_frame = np.zeros((cfg_cam_res[0],cfg_cam_res[1],3), np.uint8)
 # cv2.imshow('frame', null_frame)

### 2018-02-06 params.py
epochs as default data store directory

@@ -47,7 +47,7 @@ use_picar_mini = True # visualization fix for picar mini
 # Directories
 ##########################################################
 save_dir = os.path.abspath('models')
-data_dir = os.path.abspath('epochs-conv')
+data_dir = os.path.abspath('epochs')
 out_dir = os.path.abspath('output')
 
 if not os.path.isdir(data_dir):

### 2018-02-07 paper/comparison.tex

@@
 Section~\ref{sec:eval-memhog}, we observed dramatic execution time
-increases, up to 11.5X, in Raspberry Pi 3 as we increased the number of
+increases, up to 11.6X, in Raspberry Pi 3 as we increased the number of
 co-scheduled tasks. We also observe increased control loop execution

### 2018-02-07 paper/intro.tex
8 -> 9 layers.

@@
 value as output at each control period in \emph{real-time}.
-The network has 8 layers, about 27 million connections
+The network has 9 layers, about 27 million connections
 and 250 thousand parameters (weights).
@@
 In particular, we find that DNN inferencing is highly sensitive to
-memory performance interference as we observe up to an 11.5X control
+memory performance interference as we observe up to an 11.6X control
 loop execution time increase, when memory performance intensive

### 2018-04-04 paper/background.tex
..
Initial commit

@@
 control, which together form a processing pipeline~\cite{Bojarski2016}.
-Recently, researchers are exploring another approach that dramatically
+Recently, researchers have begun exploring another approach that dramatically
 simplifies the standard control pipeline by applying deep neural
@@
   \centering
-  \includegraphics[width=.75\textwidth]{figs/endtoend_redrawn}
+  \includegraphics[width=.5\textwidth]{figs/endtoend_redrawn}
   \caption{Standard robotics control vs. DNN based end-to-end
@@
 The use of neural networks for end-to-end control of autonomous
-car was first demonstrated in the late 1980s~\cite{Pomerleau1989},
+cars was first demonstrated in the late 1980s~\cite{Pomerleau1989},
 using a small 3-layer fully connected neural network; and subsequently
@@
 as important, if not moreso, as computational throughput, because many
-use-cases have strict real-time requirements (e.g., search query
+use cases have strict real-time requirements (e.g., search query
 latency).
@@
 operations, has received a lot of attention from academia and industry
-in recent years as applications of deep learning are broadening to
+in recent years as applications of deep learning are broadening to include
 areas of real-time embedded systems such as autonomous vehicles. These
@@
 DSP, FPGA, and neural network optimized ASIC hardware~\cite{Jouppi2017}.
-Consolidating multiple tasks on SoCs with lots of shared hardware
+Consolidating multiple tasks on SoCs with a lot of shared hardware
 resources while guaranteeing real-time performance is also an active
@@
 robotics systems, and (2) what kind of computing architecture and
-runtime supports are most appropriate for such workload. To
+runtime supports are most appropriate for such a workload. To
 achieve these goals, we implement a low-cost autonomous car platform

### 2018-04-04 paper/comparison.tex
Initial commit

@@
   \centering
-  \includegraphics[width=.7\textwidth]{figs/compare_core}
+  \includegraphics[width=.45\textwidth]{figs/compare_core}
   \caption{Average control loop execution time.} 
@@
 (GPU) performance does not change much, as most of the neural network 
-computation is done at the GPU.
+computation is done by the GPU.
 
@@
   \centering
-  \includegraphics[width=.7\textwidth]{figs/compare_model}
+  \includegraphics[width=.45\textwidth]{figs/compare_model}
   \caption{Average control loop execution time when multiple DNN
@@
   \centering
-  \includegraphics[width=.7\textwidth]{figs/compare_benchmark}
+  \includegraphics[width=.45\textwidth]{figs/compare_benchmark}
   \caption{Average control loop execution time in the presence of an
@@
   \centering
-  \includegraphics[width=.7\textwidth]{figs/compare_l2missrate}
+  \includegraphics[width=.45\textwidth]{figs/compare_l2missrate}
   \caption{L2 miss rates in the presence of an increasing number of 
@@
 intensive synthetic benchmarks on the DNN control loop timing. 
-
 Figure~\ref{fig:sys_bench} shows the results. As discussed in 
@@
 interfering Bandwidth benchmark instances increases, the L2 miss rates
-even drop slightly. In case of the TX2 (CPU), from solo to one co-runner,
-L2 miss rate increase almost twice, but then the execution time
+drop slightly. In the case of the TX2 (CPU), from solo to one co-runner,
+L2 miss rates almost double, but then the execution time
 increase is a little less than 10\%. However, when there are
@@
 
-In summary, we find that today's embedded computing platforms, even as
+In summary, we find that todays embedded computing platforms, even as
 inexpensive as a Raspberry Pi 3, are powerful enough to support

### 2018-04-04 paper/evaluation-cachepart.tex
Initial commit

@@
 
-\begin{figure}[h]
+\begin{figure}
   \centering
-  \includegraphics[width=\textwidth]{figs/cache-mapping}
+  \includegraphics[width=.5\textwidth]{figs/cache-mapping}
   \caption{A memory mapping of the cache sets for the Broadcom 
@@
   \label{fig:cache-mapping}
-\end{figure}
-
-\begin{figure}[h]
-  \centering
-  \includegraphics[width=.7\textwidth]{figs/palloc_multicore}
-  \caption{Average control loop execution time vs.  % of L2 cache
-    L2 cache partition size.}
-  \label{fig:palloc_multicore}
 \end{figure}
@@
 %%   \centering
-%%   \includegraphics[width=.7\textwidth]{figs/palloc_multimodel}
-%%   \caption{Timing impact of co-scheduling multiple DNNs when cache 
-%% partitioning is enabled.}
-%%   \label{fig:palloc_multimodel}
+%%   \includegraphics[width=.7\textwidth]{figs/palloc_bandwidth_modelbus}
+%%   \caption{Model bus access impact of co-scheduling memory intensive co-runners 
+%% when cache partitioning is enabled.}
+%%   \label{fig:palloc_bandwidth_modelbus}
 %% \end{figure}
-
-\begin{figure}[h]
-  \centering
-  \includegraphics[width=.7\textwidth]{figs/palloc_bandwidth_exectime}
-  \caption{Timing impact of co-scheduling memory intensive co-runners 
-when cache partitioning is enabled.}
-  \label{fig:palloc_bandwidth_exectime}
-\end{figure}
 
@@
 %%   \includegraphics[width=.7\textwidth]{figs/palloc_bandwidth_bus}
-%%   \caption{Bus access impact of co-scheduling memory intensive co-runners 
+%%   \caption{Total bus access impact of co-scheduling memory intensive co-runners 
 %% when cache partitioning is enabled.}
@@
 
+\begin{figure}[h]
+  \centering
+  \includegraphics[width=.45\textwidth]{figs/palloc_multicore}
+  \caption{Average control loop execution time vs.  % of L2 cache
+    L2 cache partition size.}
+  \label{fig:palloc_multicore}
+\end{figure}
+
 In the first experiment, we investigate the cache space sensitivity of
@@
 different cgroups which are configured to use 2, 4,
-6, 8 colors (25\%, 50\%, 75\%, 100\% of the L2 cache
+6, and 8 colors (25\%, 50\%, 75\%, and 100\% of the L2 cache
 space, respectively). We then execute the control loop on each cache
@@
 of the DNN.
-
-We run the DNN model with different amounts of L2 cache being
-available based on the number of colors being used. For example, if two 
-colors are used, then 25\% of the L2 cache is available, whereas if all 
-8 colors are used, 100\% of the L2 cache is available. Particularly, we 
-test the performance of the model when it utilizes a single core and
-all four of the available cores. The results can be seen in 
+%% We run the DNN model with different amounts of L2 cache being
+%% available based on the number of colors being used. For example, if two 
+%% colors are used, then 25\% of the L2 cache is available, whereas if all 
+%% 8 colors are used, 100\% of the L2 cache is available. Particularly, we 
+%% test the performance of the model when it utilizes a single core and
+%% all four of the available cores. 
+The results can be seen in 
 \ref{fig:palloc_multicore}. As expected, the amount of L2 cache 
 available to the model does not result in any significant changes in 
-performance. This can also be seen by the L2 miss rates of the model 
-seen in \ref{fig:palloc_multicore_l2missrate}. As more cache space 
-becomes avaiable, the percentage of L2 misses decreases, which doesn't 
-affect the real-time performance of the DNN. As such, we find that a 
-single model is not sensitive to the shared L2 cache.
+performance. 
+%% This can also be seen by the L2 miss rates of the model 
+%% seen in \ref{fig:palloc_multicore_l2missrate}. As more cache space 
+%% becomes avaiable, the percentage of L2 misses decreases, which doesn't 
+%% affect the real-time performance of the DNN. As such, we find that a 
+%% single model is not sensitive to the shared L2 cache.
+
+\begin{figure}[h]
+  \centering
+  \includegraphics[width=.45\textwidth]{figs/palloc_multimodel}
+  \caption{Timing impact of co-scheduling multiple DNNs when cache 
+partitioning is enabled.}
+  \label{fig:palloc_multimodel}
+\end{figure}
 
@@
 in the colors used, so no L2 cache is shared between models. As can be 
-seen in \ref{fig:palloc_multimodel}, the performance of the models 
+seen in Figure \ref{fig:palloc_multimodel}, the performance of the models 
 remains the same as when no cache partitioning was employed. Based of 
@@
 source of interference between multiple DNN models.
+
+\begin{figure}[h]
+  \centering
+  \includegraphics[width=.45\textwidth]{figs/palloc_bandwidth_exectime}
+  \caption{Timing impact of co-scheduling memory intensive co-runners 
+when cache partitioning is enabled.}
+  \label{fig:palloc_bandwidth_exectime}
+\end{figure}
 
@@
 experiments with less than four tasks. The results can be seen in Figure
-\ref{fig:palloc_bandwidth_exectime}. With zero or one co-runners, the DNN
+\ref{fig:palloc_bandwidth_exectime}. When the model runs by itself, performance
+is slightly worse due to less cache space being available (this was also seen in 
+Figure \ref{fig:palloc_multicore}). With one or two co-runners, the DNN
 performance doesnt't change relative to when the cache wasn't partitioned.
-However, we find that with two or more co-runners, DNN performance becomes
-worse. In the case of two co-runners, average control loop time increases
-by 66.31 ms, or 26\%, and with three co-runners, the average time
-is 265.64 ms, of 49\%, greater.
-At the same time, though, the L2 miss rates of the model don't correlate 
-to the timing increases seen, as is shown in 
-\ref{fig:palloc_bandwidth_l2missrate}. Instead, we find that the number
-of bus accesses increases when cache partitions are used, and they mirror
-the timing increases of the model. This can be seen in Figure 
-\ref{fig:palloc_bandwidth_bus}. These additional bus accesses are
-most likely caused by the L2 prefetcher, as it can generate additional bus
-accesses that aren't counted as L2 misses. As a result, we find that cache
-partitioning is detrimental to the performance of the DNN so long as the 
-L2 prefetcher is enabled (we have not yet found a way to disable it on
-the Pi 3). Furthermore, no cache sensitivity was displayed by the model,
-thus showing that the performance of the L2 cache is not vital for DNN 
-performance when co-runners are introduced.
+When three co-runners are present, DNN performance becomes slightly
+better, with a improvement of \textasciitilde6.5\%. However, as the 
+performance improvement is minimal, we do not believe that cache 
+partitioning is effective for protecting the model from interference from
+other memory intensive processes.
+%% At the same time, though, the L2 miss rates of the model don't correlate 
+%% to the timing increases seen, as is shown in 
+%% \ref{fig:palloc_bandwidth_l2missrate}. Instead, we find that the number
+%% of bus accesses increases when cache partitions are used, and they mirror
+%% the timing increases of the model. This can be seen in Figure 
+%% \ref{fig:palloc_bandwidth_bus}. These additional bus accesses are
+%% most likely caused by the L2 prefetcher, as it can generate additional bus
+%% accesses that aren't counted as L2 misses. As a result, we find that cache
+%% partitioning is detrimental to the performance of the DNN so long as the 
+%% 0L2 prefetcher is enabled (we have not yet found a way to disable it on
+%% the Pi 3). Furthermore, no cache sensitivity was displayed by the model,
+%% thus showing that the performance of the L2 cache is not vital for DNN 
+%% performance when co-runners are introduced.
 
@@
 improvements are gained and that performance remains consistent, or 
-decreases, depending on the circumstances. In all experiements, the DNN
-shows no sensitivity to the shared L2 cache. As a result, we conclude 
-that cache partitioning is not an effective isolation mechanism, and that 
-the performance of the shared DRAM controller is of greater importance 
-to the real-time efficiacy of the DeepPicar.
+slightly decreases, depending on the circumstances. In all experiements,
+the DNN shows no sensitivity to the shared L2 cache. As a result, we 
+conclude that cache partitioning is not an effective isolation mechanism, 
+and that the performance of the shared DRAM controller is of greater 
+importance to the real-time efficiacy of the DeepPicar.

### 2018-04-04 paper/evaluation-drampart.tex
Initial commit

\subsection{Effect of Memory Bandwidth Throttling}

To determine the how sensitive the model is to the shared DRAM 
memory, we use MemGuard, a memory bandwidth reservation system that 
can limit the amount of bandwidth each core receives in a period. For 
all tests run using MemGuard, we use a period of one second.

\begin{figure}[h]
  \centering
  \includegraphics[width=.45\textwidth]{figs/memguard_multicore}
  \caption{Average control loop execution time vs.
    memory bandwidth reservation size. }
  \label{fig:memguard_multicore}
\end{figure}

We first run a single model on one core, Core 0, while varying the 
core's bandwidth reservation size, from 400 MB/s down to 100 MB/s. 
The results can be seen in Figure \ref{fig:memguard_multicore}. When less 
memory bandwidth is available to the core the model's performance 
noticably decreases, and it performs the best when no memory bandwidth 
throttling is implemented. Compared to a reservation size of 100 
MB/s, disabling memory bandwidth throttling results in 
\textasciitilde3x performance improvement. If the model was memory 
insensitive, then the amount of bandwidth avaiable to it would have 
little effect on its performance. However, since that is not the 
case, we conclude that, to a notable extent, the neural network model 
is dependent on shared DRAM memory performance.

\begin{figure}[h]
  \centering
  \includegraphics[width=.45\textwidth]{figs/memguard_multimodel}
  \caption{Timing impact of co-scheduling multiple DNNs when memory
bandwidth throttling is enabled. }
  \label{fig:memguard_multimodel}
\end{figure}

We also test the effects of memory bandwidth throttling in the case 
of multiple models running concurrently on the Pi 3 by rerunning the 
4Nx1C experiment. We use the same memory bandwidth reservation sizes 
from the previous experiment. The results can be seen in Figure
\ref{fig:memguard_multimodel}. Once again, the performance of the 
models are affected by the amount of memory bandwidth that is 
available to the Pi 3's cores during each period, meaning that they 
are all memory dependent.

\begin{figure}[h]
  \centering
  \includegraphics[width=.45\textwidth]{figs/memguard_bandwidth}
  \caption{Timing impact of co-scheduling memory intensive read
co-runners when memory bandwidth throttling is enabled. }
  \label{fig:memguard_bandwidth}
\end{figure}

Finally, we test the performance of the model in the presence of 
memory intensive read co-runners in order to determine the effects of 
limiting the memory bandwidth of the co-runners. For each of the 
experiments, we give the model's core, Core 0, a constant 1000 MB/s 
memory bandwidth reservation, and then vary the amount of bandwidth 
the co-runners' cores receive following the same scheme used in the 
previous experiments. The results can be see in Figure
\ref{fig:memguard_bandwidth}. When less memory bandwidth is given to 
the read co-runners, the performance of the model does improve. In 
the case of one co-runner, the performance of the model remains 
constant even as the reservation sizes are decreased, however, it 
still improves compared to when throttling wasn't enabled. With two 
and three co-runners, performance continually improves as less 
bandwidth is given to the benchmarks. This is especially true with 
three co-runners as the improvement is linear, which indicates that 
there is a direct correlation between the co-runner bandwidth size 
and model performance. In all cases, model performance improved, to 
some extent, through the use of memory bandwidth throttling on the 
co-runners. Since this would not be the case if the model was memory 
insensitive, we find that the model is memory dependent in the 
presence of co-runners.

Based on all of the memory bandwidth throttling experiments 
performed, it is evident that the model is dependent on the shared 
DRAM memory, to a noticable extent. When a single or multiple models 
were run, their performance(s) decreased as less memory bandwidth was 
made available to them. Furthermore, when the bandwidths of memory 
intensive read co-runners were limited, the performance of the model 
increased. As such, we conclude that the model is dependent on the 
performance of the shared DRAM memory.

### 2018-04-04 paper/evaluation.tex
Minor changes
Initial commit

@@
 
-The Raspberry Pi 3 Model B platform used in DeepPicar equips a Boardcom
+The Raspberry Pi 3 Model B platform used in DeepPicar equips a Broadcom
 BCM2837 SoC, which has a quad-core ARM Cortex-A53 cluster,
@@
 control performance improves when the frequency is higher, though
-computation time and the type of the particular physical system are
+computation time and the type of particular physical system are
 factors in determining a proper control loop frequency. While a standard
@@
   \centering
-  \includegraphics[width=.75\textwidth]{figs/Fig7_new}
+  \includegraphics[width=.5\textwidth]{figs/Fig7_new}
   \caption{DeepPicar's control loop processing times over 1000 input image frames.}
@@
 This means that the DeepPicar can operate
-at a up to 40 Hz control frequency in real-time using only the on-board
+at up to 40 Hz control frequency in real-time using only the on-board
 Raspberry Pi 3 computing platform, as no remote computing resources 
@@
   \centering
-  \includegraphics[width=.7\textwidth]{figs/perf_vs_corecnt}
+  \includegraphics[width=.45\textwidth]{figs/perf_vs_corecnt}
   \caption{Average control loop execution time vs. \#of CPU
@@
   \centering
-  \includegraphics[width=.7\textwidth]{figs/perf_vs_modelcnt}
+  \includegraphics[width=.45\textwidth]{figs/perf_vs_modelcnt}
   \caption{Timing impact of co-scheduling multiple DNNs. 1Nx1C: one DNN
@@
   IsolBench suite, with the following command line parameters: \texttt{\$
-  bandwdith -a write -m 16384}}, on the remaining idle cores.
+  bandwdith -a write -m 16384 -t 1000}}, on the remaining idle cores.
 
@@
   \centering
-  \includegraphics[width=.7\textwidth]{figs/perf_vs_bandwidth}
-  \caption{Effect of memory performance hogs on the DNN
-    inferencing. The DNN model uses Core 0 and memory-hog co-runners
-    use the rest of the cores.}
-  \label{fig:}
+  \includegraphics[width=.45\textwidth]{figs/perf_vs_bandwidth}
+  \caption{Average processing time vs. the number of memory
+    memory intensive co-runners introduced.}
+  \label{fig:perf_vs_bandwidth}
 \end{figure}
 
-Figure~\ref{fig:} shows the normalized execution time (first group) and
-L2 miss-rate (last group)
-of the DNN model running on the Core 0 as a function of the number of
-co-scheduled memory intensive synthetic benchmarks. We also shows the
-normalized total system bus access count (middle group), which is
-representative of memory bandwidth usage of all the cores.
+\begin{figure}[h]
+  \centering
+  \includegraphics[width=.45\textwidth]{figs/perf_vs_bandwidth_mem}
+  \caption{Effect of memory performance hogs on the shared resources. 
+    The DNN model uses Core 0 and memory-hog co-runners
+    use the rest of the cores.}
+  \label{fig:perf_vs_bandwidth_mem}
+\end{figure}
+
+Figure~\ref{fig:perf_vs_bandwidth} shows the execution time of the DNN 
+model and Figure~\ref{fig:perf_vs_bandwidth_mem} shows both the normalized
+total memory bus accesses (first group) and L2 miss rates (second group).
+In both figures, the DNN model is running on Core 0 as a function of the 
+number of co-scheduled memory intensive synthetic benchmarks.
+%% We also show 
+%% the normalized total system bus access count (middle group), which is
+%% representative of memory bandwidth usage of all the cores.
 First, as we
@@
 due to the memory-hog co-runners, appears to be more closely
-correlated with the DNN model's execution time increase. This is
+correlated with the DNN model's execution time increase. This means
 the high memory bandwidth pressure from the co-scheduled
-memory intensive benchmarks would have increased the DNN model's
+memory intensive benchmarks increased the DNN model's
 memory access latency.
@@
 
-%% \input{evaluation-cachepart}
+\input{evaluation-cachepart}
+\input{evaluation-drampart}
 
@@
 
-%% Lastly, we find that partitioning the shared L2 cache provides 
-%% practically no benefits to the real-time performance of the DNN. To the 
-%% contrary, performance is constant regardless of the amount of cache 
-%% space used. We also find that performance becomes worse when memory
-%% intensive co-runners are introduced as they generate more memory 
-%% contention when partitioning is enabled. Furthermore, we 
-%% find that performance remains consistent when multiple models are run 
-%% simultaneously. As such, we conclude that the shared L2 cache is not 
-%% essential to the DNN's real-time performance.
+Fifth, we find that partitioning the shared L2 cache provides 
+practically no benefits to the real-time performance of the DNN. To the 
+contrary, performance is constant regardless of the amount of cache 
+space used. Furthermore, we find that performance remains consistent 
+when multiple models are run simultaneously, and when memory intensive
+read co-runners are introduced. As such, we conclude that 
+the shared L2 cache is not essential to the DNN's real-time performance.
+
+Lastly, we find that memory bandwidth throttling results in significant
+changes to model performance. When less bandwidth is given to models,
+whether they're running alone or concurrently, we find that their performance
+decreases noticably. When memory intensive co-runners are
+present, we find that limiting the available bandwidth of the co-runners
+actually results in improved model performance. As such, we conclude
+that the shared DRAM memory is essential to the DNN's real-time 
+performance.

### 2018-04-04 paper/intro.tex
Initial commit

@@
 Instead, we want to build a low cost testbed that still employs the
-state-of-the art AI technologies. Specifically, we focus on a end-to-end
+state-of-the art AI technologies. Specifically, we focus on an end-to-end
 deep learning based real-time control system,
 which was developed for a real self-driving car, NVIDIA
-DAVE-2~\cite{Bojarski2016}, and use the same methodology on a
+DAVE-2~\cite{Bojarski2016}, and uses the same methodology on a
 smaller and less costly setup. In developing the testbed, our
@@
 control; and (2) to evaluate real-time performance of contemporary embedded
-platforms for such workload.
+platforms for such a workload.
 
@@
 The network architecture is almost identical to that of NVIDIA's DAVE-2
-self-driving car~\cite{Bojarski2016}, which uses a much more powerful
+self-driving car~\cite{Bojarski2016}, which uses a significantly more powerful
 computer (Drive PX computer~\cite{drivepx}) than a Raspberry Pi~3.
@@
 %   isolation of DNN workload when memory bandwidth is contended.
-From the systematic study, we made a number of interesting
+From the systematic study, we have made a number of interesting
 observations from the perspective of real-time systems.
@@
 Second, we find that real-time processing of DeepPicar's CNN is
-feasible on today's embedded computing platforms, even  
+feasible on today's embedded computing platforms, even one  
 as inexpensive as the Raspberry Pi 3. We find that the control loop 
@@
 In particular, we find that DNN inferencing is highly sensitive to
-memory performance interference as we observe up to an 11.6X control
-loop execution time increase, when memory performance intensive
+memory performance interference, as we observe up to an 11.6X control
+loop execution time increase when memory performance intensive
 applications are co-scheduled on idle CPU cores.
@@
 can be found at: \url{https://github.com/heechul/picar}}, with the
-    goal of lowering the economic and safety-related barriers to study
-    autonomous cars.
+    goal of lowering the economic and safety-related barriers to the study
+    of autonomous cars.
 
@@
     DNN inferencing workload on a number of representative
-    contemporary embedded computing platforms. To the
+    contemporary embedded computing platforms. While larger embedded
+    platforms have been evaluated, as in  \cite{Otterness2017}, to the
     best of our knowledge, no prior work investigated the
     impacts of shared resource contention on DNN inferencing workloads
-    in the context of real-time systems.
+    in the context of smaller embedded platforms like the Raspberry Pi 3.
 
@@
 evaluation of the platform and how different factors can affect
-performance. Section~\ref{sec:comparison} gives a comparison between  
+performance. Section~\ref{sec:comparison} offers a comparison between  
 the Raspberry Pi 3 and other embedded computing platforms to 
 determine their suitably for autonomous driving research.
-We discuss more detailed training methods and system-level issues
-in~\ref{sec:discussion}. We review related work in
+%% We discuss more detailed training methods and system-level issues
+%% in~\ref{sec:discussion}. 
+We review related work in
 Section~\ref{sec:related} and conclude in

### 2018-04-04 paper/overview.tex
Initial commit

@@
   \centering
-  \includegraphics[width=.65\textwidth]{figs/DeepPicar_platform}
+  \includegraphics[width=.4\textwidth]{figs/DeepPicar_platform}
   \caption{DeepPicar platform.}
@@
   \centering
-  \includegraphics[width=0.65\textwidth]{figs/architecture}
+  \includegraphics[width=0.4\textwidth]{figs/architecture}
   \caption{DeepPicar's neural network architecture: 9 layers (5
@@
 based CNN model in ~\cite{deeptesla}. Note, however, that the CNN model
-in~\cite{deeptesla} was considerably bigger than NVIDIA's CNN model as
+in~\cite{deeptesla} was considerably larger than NVIDIA's CNN model as
 it contained an additional fully-connected layer of approximately 1.3M
@@
   \centering
-%% <<<<<<< HEAD
-%%   \includegraphics[width=.5\textwidth]{figs/track_new2}
-%%   \caption{One of the custom tracks used for training/testing.}
-%%   \label{fig:track}
-%% =======
-  \begin{subfigure}{0.55\textwidth}
+  \begin{subfigure}{0.4\textwidth}
     \includegraphics[width=\textwidth]{figs/track_new2}
@@
   \hfill
-  \begin{subfigure}{0.7\textwidth}
+  \begin{subfigure}{0.4\textwidth}
     \includegraphics[width=\textwidth]{figs/track-2}
@@
   \label{fig:tracks}
-%% >>>>>>> f3814396d8701264f2439b42fbe8bbb608c41452
 \end{figure}
@@
 timestamped videos and contol commands. The stored data is then copied 
-to a desktop computer, which is equipped with a NVIDIA GTX 1050 Ti GPU, 
+to a desktop computer, which is equipped with a NVIDIA GTX 1060 GPU, 
 where we train the network to accelerate training speed. 
@@
 is copied back to the Raspberry Pi 3. The network is then used
-by the car's main controller, which feeds a image frame from the web
+by the car's main controller, which feeds an image frame from the web
 camera as input to the network. In each control period, the produced 
-steering angle output is then converted as the PWM values of the 
+steering angle output is then converted into the PWM values of the 
 steering motor of the car. Figure~\ref{fig:controlloop} shows simplified 
@@
 (motor) control latencies. In DeepPicar's
-context, the camera latency is from the time the camera
+context, the camera latency  measured from the time the camera
 sensor observes the scene to the time the computer actually reads the
@@
 quality of training data as well as various environmental factors such
-as the lighting condition. We plan to investigate more systematic ways
+as lighting conditions. We plan to investigate more systematic ways
 to improve the CNN's prediction accuracies.

### 2018-04-04 paper/related.tex
Initial commit

@@
-\section{Related Work}
+\section{Related Work}\label{sec:related}
 % discusstion
@@
 
-While continuing performance improvmenets in embedded computing
-platforms will certainly make processing of these complex CNNs faster,
+While continuing performance improvements in embedded computing
+platforms will certainly make processing these complex CNNs faster,
 another actively investigated approrach is to reduce the required
@@
 involved in the large matrix multiplications are a burdensome task.
-Many recent advances in network compression showed promising results
+Many recent advances in network compression have shown promising results
 in reducing such computational cost during the feedforward

### 2018-04-04 picar-mini-kbd-drv8835.py

@@ -177,7 +177,7 @@ while True:
         frame_id += 1
         
         # write input (angle)
-        str = "{},{},{}\n".format(int(ts*1000), frame_id, angle)
+        str = "{},{},{},{}\n".format(int(ts*1000), frame_id, angle, cur_speed)
         keyfile.write(str)
         
         # write input (button: left, center, stop, speed)

### 2018-04-15 paper/comparison.tex
Updates to paper

@@
   \centering
-  \includegraphics[width=.45\textwidth]{figs/compare_benchmark}
+  \includegraphics[width=.45\textwidth]{figs/compare_benchmark_read}
   \caption{Average control loop execution time in the presence of an
-    increasing number of memory intensive applications on idle CPU cores.}
-  \label{fig:sys_bench}
+    increasing number of memory intensive read applications on idle CPU cores.}
+  \label{fig:sys_bench_read}
 \end{figure}
@@
   \centering
-  \includegraphics[width=.45\textwidth]{figs/compare_l2missrate}
-  \caption{L2 miss rates in the presence of an increasing number of 
-			memory intensive applications on idle CPU cores.}
-  \label{fig:sys_l2miss}
+  \includegraphics[width=.45\textwidth]{figs/compare_benchmark}
+  \caption{Average control loop execution time in the presence of an
+    increasing number of memory intensive write applications on idle CPU cores.}
+  \label{fig:sys_bench}
 \end{figure}
@@
 intensive synthetic benchmarks on the DNN control loop timing. 
-Figure~\ref{fig:sys_bench} shows the results. As discussed in 
+Figure~\ref{fig:sys_bench_read} and ~\ref{fig:sys_bench} shows 
+the results for read and write co-runners, respectively. For read co-runners,
+the performance of all platforms decreased as additional BwRead processes
+were introduced, but none of them displayed any significant or unexpected
+execution time increases. The Pi 3 suffered an increase of up to 1.6X; the 
+Intel UP increased up to 1.3X; the TX2 (CPU) and TX2(GPU) saw increases 
+up to1.6X and 2.2X, respectively. For write co-runners, as discussed in 
 Section~\ref{sec:eval-memhog}, we observed dramatic execution time
-increases, up to 11.6X, in Raspberry Pi 3 as we increased the number of
+increases, up to 11.6X, in the Raspberry Pi 3 as we increased the number of
 co-scheduled tasks. We also observe increased control loop execution
@@
 (i.e., the model runs on a single core in isolation), Intel UP suffers up to
-2.3X execution time increase; TX2 (CPU) and TX2 (GPU) suffer up to
-2.2X and 2.1X increases, respectively. This is somewhat surprising
+2.3X execution time increase; TX2 (CPU) and TX2 (GPU) both suffer up to
+2.3X increases. This is somewhat surprising
 because the Raspberry Pi 3's cores are in-order architecture based while
@@
 Another interesting observation is that the TX2 (GPU) also suffers
-considerable execution time increase (2.1X) despite the fact that the
+considerable execution time increase (2.3X) despite the fact that the
 co-scheduled synthetic tasks do not utilize the GPU. In other words,
@@
 
-Figure~\ref{fig:sys_l2miss} shows the L2 miss rates of DNN control
-loop. The key takeaway of this result is that the L2 miss rate is a
+\begin{figure}[h]
+  \centering
+  \includegraphics[width=.45\textwidth]{figs/compare_readl2}
+  \caption{L2 miss rates in the presence of an increasing number of 
+			memory intensive read applications on idle CPU cores.}
+  \label{fig:sys_readl2miss}
+\end{figure}
+
+\begin{figure}[h]
+  \centering
+  \includegraphics[width=.45\textwidth]{figs/compare_l2missrate}
+  \caption{L2 miss rates in the presence of an increasing number of 
+			memory intensive write applications on idle CPU cores.}
+  \label{fig:sys_l2miss}
+\end{figure}
+
+Figures~\ref{fig:sys_readl2miss} and ~\ref{fig:sys_l2miss} show
+the L2 miss rates of DNN control loops when read and write co-runners
+are introduced, respectively. The key takeaway of this result is that the 
+L2 miss rate is a
 poor indicator to the execution time of the control loop. As shown
@@
 co-runners, even though the L2 miss rates are largely unchanged, the
-DNN processing time increases significantly. The TX2 (GPU) also shows
+CNN processing time increases significantly. The TX2 (GPU) also shows
 similar behavior. All of these results essentially point to the fact
@@
 but not L2 cache space sensitive.
+
+\begin{figure}[h]
+  \centering
+  \includegraphics[width=.45\textwidth]{figs/compare_readbus}
+  \caption{Bus accesses in the presence of an increasing number of 
+			memory intensive read applications on idle CPU cores.}
+  \label{fig:sys_readbus}
+\end{figure}
+
+\begin{figure}[h]
+  \centering
+  \includegraphics[width=.45\textwidth]{figs/compare_writebus}
+  \caption{Bus accesses in the presence of an increasing number of 
+			memory intensive write applications on idle CPU cores.}
+  \label{fig:sys_writebus}
+\end{figure}
+
+\color{red}FIXME: I believe the bus access results for the Intel UP are wrong. 
+I couldn't get the IDI off-core response event to work, so I used the bus-cycles 
+perf event, which I think is incorrect.
+
+\color{black}Figures~\ref{fig:sys_readbus} and ~\ref{fig:sys_writebus} show the
+normalized bus accesses for each of the platforms as read and write
+co-runners are introduced, respectively. Unlike the L2 miss rates of the
+model, the total number of bus accesses more closely align with the 
+execution time increases that are seen in each of the platforms. For the Pi 3, both 
+the execution time and bus accesses increase at exponential rates. In the TX2(GPU), 
+the bus accesses increase in a similar linear fashion to the execution time. The main outlier
+is the TX2(CPU) which has bus accesses increase at a faster rate similar to that of the Pi 3.
+However, the bus accesses seem to increase linearly, which is also seen in the
+execution time. As a result, we find that the total number of bus accesses have a higher 
+correlation to the model execution time than L2 miss rates.
 

### 2018-04-15 paper/conclusion.tex
Updates to paper

@@
 in real-time. Importantly, DeepPicar's neural network architecture is
-identical to that of NVIDIA's real self-driving car.
+almost identical to that of NVIDIA's real self-driving car.
 
@@
 
-As future work, we plan to apply shared DRAM resource management
-techniques~\cite{Yun2013,yun2014rtas} on the DeepPicar platform and
-evaluate their impact on the real-time performance of the system. We
-also plan to improve the prediction accuracy by feeding more data and
+As future work, 
+%% we plan to apply shared DRAM resource management
+%% techniques~\cite{Yun2013,yun2014rtas} on the DeepPicar platform and
+%%evaluate their impact on the real-time performance of the system. 
+we plan to improve the prediction accuracy by feeding more data and
 upgrading the RC car hardware platform to enable more precise steering

### 2018-04-15 paper/evaluation-cachepart.tex
Updates to paper

@@
 Figure~\ref{fig:cache-mapping} shows the physical address mapping of
-Raspberry Pi3's BCM2837 processor. We use the bit 12, 13, 14 for
-coloring, which results in 8 page colors.
+Raspberry Pi3's BCM2837 processor. We use the bits 13 and 14 for
+coloring, which results in 4 page colors.
 
@@
   \includegraphics[width=.45\textwidth]{figs/palloc_multicore}
-  \caption{Average control loop execution time vs.  % of L2 cache
-    L2 cache partition size.}
+  \caption{Cache space sensitivity of the control loop exectuion time.}
   \label{fig:palloc_multicore}
@@
 the DeepPicar's DNN-based control loop. Using PALLOC, we create 4
-different cgroups which are configured to use 2, 4,
-6, and 8 colors (25\%, 50\%, 75\%, and 100\% of the L2 cache
+different cgroups which are configured to use 1, 2,
+3, and 4 colors (25\%, 50\%, 75\%, and 100\% of the L2 cache
 space, respectively). We then execute the control loop on each cache
@@
 If performance improves as a result of partitioning the shared cache 
-then we know that the DNN, to some extent, relies on shared cache 
-performance in addition to shared DRAM performance. On the other 
+then we know that the DNN, to some extent, relies on the shared cache 
+in addition to the shared DRAM. On the other 
 hand, if there is no improvement in performance, then it can be 
-observed that the shared memory is more critical to the performance 
-of the DNN.
+observed that the shared memory is more critical for the DNN.
 %% We run the DNN model with different amounts of L2 cache being
@@
 
-\begin{figure}[h]
-  \centering
-  \includegraphics[width=.45\textwidth]{figs/palloc_multimodel}
-  \caption{Timing impact of co-scheduling multiple DNNs when cache 
-partitioning is enabled.}
-  \label{fig:palloc_multimodel}
-\end{figure}
+%% \begin{figure}[h]
+%%  \centering
+%%   \includegraphics[width=.45\textwidth]{figs/palloc_multimodel}
+%%   \caption{Timing impact of co-scheduling multiple DNNs when cache 
+%% partitioning is enabled.}
+%%   \label{fig:palloc_multimodel}
+%% \end{figure}
 
-We also co-schedule multiple models to see how cache partitioning affects
-interference between them. Each model uses an equal number of colors, and
-are thus given equal amounts of L2 cache space, and there is no overlap
-in the colors used, so no L2 cache is shared between models. As can be 
-seen in Figure \ref{fig:palloc_multimodel}, the performance of the models 
-remains the same as when no cache partitioning was employed. Based of 
-these results, we find that contention for the L2 cache is not the main
-source of interference between multiple DNN models.
+%% We also co-schedule multiple models to see how cache partitioning affects
+%% interference between them. Each model uses an equal number of colors, and
+%% are thus given equal amounts of L2 cache space, and there is no overlap
+%% in the colors used, so no L2 cache is shared between models. As can be 
+%% seen in Figure \ref{fig:palloc_multimodel}, the performance of the models 
+%% remains the same as when no cache partitioning was employed. Based of 
+%% these results, we find that contention for the L2 cache is not the main
+%% source of interference between multiple DNN models.
 
@@
 
-Finally, we run a single model alongside memory intensive co-runners 
+We also run a single model alongside three memory intensive read co-runners 
 again, but ensure that each task is given equal amounts of L2 cache space.
-For this, we assign two colors to each task, regardless of the number of
-tasks being scheduled. As a result, the entire cache is not utilized in
-experiments with less than four tasks. The results can be seen in Figure
-\ref{fig:palloc_bandwidth_exectime}. When the model runs by itself, performance
-is slightly worse due to less cache space being available (this was also seen in 
-Figure \ref{fig:palloc_multicore}). With one or two co-runners, the DNN
-performance doesnt't change relative to when the cache wasn't partitioned.
-When three co-runners are present, DNN performance becomes slightly
-better, with a improvement of \textasciitilde6.5\%. However, as the 
-performance improvement is minimal, we do not believe that cache 
-partitioning is effective for protecting the model from interference from
-other memory intensive processes.
+For this, we assign one color to each task. The results can be seen in Figure
+\ref{fig:palloc_bandwidth_exectime}. When cache partitioning is enabled, 
+DNN performance becomes slightly better, with a improvement of 
+\textasciitilde6.5\%. However, as the performance improvement is minimal, 
+we do not believe that cache partitioning is effective for protecting the model 
+from interference from other memory intensive processes.
 %% At the same time, though, the L2 miss rates of the model don't correlate 
@@
 By partitioning the shared L2 cache, we find that no noticable 
-improvements are gained and that performance remains consistent, or 
-slightly decreases, depending on the circumstances. In all experiements,
-the DNN shows no sensitivity to the shared L2 cache. As a result, we 
-conclude that cache partitioning is not an effective isolation mechanism, 
-and that the performance of the shared DRAM controller is of greater 
-importance to the real-time efficiacy of the DeepPicar.
+improvements are gained and that performance remains consistent. 
+In all experiements, the model shows no sensitivity to the shared L2 
+cache. As a result, we conclude that cache partitioning is not an 
+effective isolation mechanism, and that the performance of the shared 
+DRAM controller is of greater importance to the real-time efficiacy of 
+the DeepPicar.

### 2018-04-15 paper/evaluation-drampart.tex
Updates to paper

@@
 To determine the how sensitive the model is to the shared DRAM 
-memory, we use MemGuard, a memory bandwidth reservation system that 
+memory, we use MemGuard\cite{yun2014rtas}, a memory bandwidth reservation system that 
 can limit the amount of bandwidth each core receives in a period. For 
@@
   \includegraphics[width=.45\textwidth]{figs/memguard_multicore}
-  \caption{Average control loop execution time vs.
-    memory bandwidth reservation size. }
+  \caption{ Memory bandwidth sensitivity of the control loop 
+execution time. }
   \label{fig:memguard_multicore}
@@
 We first run a single model on one core, Core 0, while varying the 
-core's bandwidth reservation size, from 400 MB/s down to 100 MB/s. 
+core's bandwidth reservation size, from 500 MB/s down to 100 MB/s. 
 The results can be seen in Figure \ref{fig:memguard_multicore}. When less 
 memory bandwidth is available to the core the model's performance 
-noticably decreases, and it performs the best when no memory bandwidth 
-throttling is implemented. Compared to a reservation size of 100 
+noticably decreases.
+%% , and it performs the best when no memory bandwidth 
+%% throttling is implemented. 
+Compared to a reservation size of 100 
 MB/s, disabling memory bandwidth throttling results in 
 \textasciitilde3x performance improvement. If the model was memory 
-insensitive, then the amount of bandwidth avaiable to it would have 
+insensitive, then the amount of bandwidth available to it would have 
 little effect on its performance. However, since that is not the 
@@
 
-\begin{figure}[h]
-  \centering
-  \includegraphics[width=.45\textwidth]{figs/memguard_multimodel}
-  \caption{Timing impact of co-scheduling multiple DNNs when memory
-bandwidth throttling is enabled. }
-  \label{fig:memguard_multimodel}
-\end{figure}
+%% \begin{figure}[h]
+%%   \centering
+%%   \includegraphics[width=.45\textwidth]{figs/memguard_multimodel}
+%%   \caption{Timing impact of co-scheduling multiple DNNs when memory
+%% bandwidth throttling is enabled. }
+%%   \label{fig:memguard_multimodel}
+%% \end{figure}
 
-We also test the effects of memory bandwidth throttling in the case 
-of multiple models running concurrently on the Pi 3 by rerunning the 
-4Nx1C experiment. We use the same memory bandwidth reservation sizes 
-from the previous experiment. The results can be seen in Figure
-\ref{fig:memguard_multimodel}. Once again, the performance of the 
-models are affected by the amount of memory bandwidth that is 
-available to the Pi 3's cores during each period, meaning that they 
-are all memory dependent.
+%% We also test the effects of memory bandwidth throttling in the case 
+%% of multiple models running concurrently on the Pi 3 by rerunning the 
+%% 4Nx1C experiment. We use the same memory bandwidth reservation sizes 
+%% from the previous experiment. The results can be seen in Figure
+%% \ref{fig:memguard_multimodel}. Once again, the performance of the 
+%% models are affected by the amount of memory bandwidth that is 
+%% available to the Pi 3's cores during each period, meaning that they 
+%% are all memory dependent.
 
@@
 
-Finally, we test the performance of the model in the presence of 
-memory intensive read co-runners in order to determine the effects of 
+\begin{figure}[h]
+  \centering
+  \includegraphics[width=.45\textwidth]{figs/memguard_bwwrite}
+  \caption{Timing impact of co-scheduling memory intensive write
+co-runners when memory bandwidth throttling is enabled. }
+  \label{fig:memguard_bwwrite}
+\end{figure}
+
+We also test the performance of the model in the presence of three
+memory intensive read and write benchmarks to determine the effects of 
 limiting the memory bandwidth of the co-runners. For each of the 
 experiments, we give the model's core, Core 0, a constant 1000 MB/s 
-memory bandwidth reservation, and then vary the amount of bandwidth 
-the co-runners' cores receive following the same scheme used in the 
-previous experiments. The results can be see in Figure
-\ref{fig:memguard_bandwidth}. When less memory bandwidth is given to 
-the read co-runners, the performance of the model does improve. In 
-the case of one co-runner, the performance of the model remains 
-constant even as the reservation sizes are decreased, however, it 
-still improves compared to when throttling wasn't enabled. With two 
-and three co-runners, performance continually improves as less 
-bandwidth is given to the benchmarks. This is especially true with 
-three co-runners as the improvement is linear, which indicates that 
-there is a direct correlation between the co-runner bandwidth size 
-and model performance. In all cases, model performance improved, to 
-some extent, through the use of memory bandwidth throttling on the 
-co-runners. Since this would not be the case if the model was memory 
-insensitive, we find that the model is memory dependent in the 
-presence of co-runners.
+memory bandwidth reservation size, and then vary the amount of bandwidth 
+the co-runners' cores receive. For read co-runners, we give them an initial
+reservation size of 500 MB/s, and decrease that to a minimum of 100 MB/s.
+In the case of write co-runners, MemGuard accounts for L2 refills but not
+write-backs, so each benchmark would effectively have twice the bandwidth
+of their reservation size. Due to this, we give write co-runners an initial 
+reservation size of 250 MB/s and decrease it to a minimum of 50 MB/s.
+The results can be see in Figures \ref{fig:memguard_bandwidth} and
+\ref{fig:memguard_bwwrite}. When less memory bandwidth is given to 
+the both the  read and write co-runners, the performance of the model 
+improves. 
+%% In the case of one co-runner, the performance of the model remains 
+%% constant even as the reservation sizes are decreased, however, it 
+%% still improves compared to when throttling wasn't enabled. 
+In both cases, when the co-runners were given a minimal reservation size,
+the performance of the model was much closer to its solo execution time.
+This is especially noteworthy when write co-runners were present, as the 
+model didn't suffer the same 11.6X slowdown that was seen without
+bandwidth throttling enabled. 
+%% When memory intensive Bandwidth benchmarks were present, model performance
+%% improved, to a notable extent, through the use of memory bandwidth 
+%% throttling on the co-runners. 
+Since this would not be the case if the 
+model was memory insensitive, we find that the model is memory 
+dependent in the presence of co-runners.
 
 Based on all of the memory bandwidth throttling experiments 
-performed, it is evident that the model is dependent on the shared 
-DRAM memory, to a noticable extent. When a single or multiple models 
-were run, their performance(s) decreased as less memory bandwidth was 
-made available to them. Furthermore, when the bandwidths of memory 
-intensive read co-runners were limited, the performance of the model 
-increased. As such, we conclude that the model is dependent on the 
-performance of the shared DRAM memory.
+performed, it is evident that the shared DRAM memory is important for
+the DNN model. When a single model was run, its 
+performance decreased as less memory bandwidth was 
+made available to it. Furthermore, when the bandwidths of memory 
+intensive read and write co-runners were limited, the performance of 
+the model improved. As such, we conclude that the model is dependent on the 
+shared DRAM memory.

### 2018-04-15 paper/evaluation.tex
Updates to paper

@@
 
-\begin{figure}[t]
+\begin{figure}[h]
   \centering
@@
 
+\begin{figure}[h]
+  \centering
+  \includegraphics[width=.45\textwidth]{figs/a53_vs_a7}
+  \caption{DNN model performance on different in-order
+ARM Cortex cores.}
+  \label{fig:a53_vs_a7}
+\end{figure}
+
+We do find the performance of the model in the presence of 
+write co-runners surprising, as it has been previously shown in experiments
+on another in-order ARM core, the Cortex A7, that write interference was not as 
+significant\cite{Valsan2016}. We also tested the model, in the presence of write 
+co-runners, on the A7 cores of the Odroid XU4 platform to see if  the same model 
+behavior would also be seen there. The results can be seen in Figure ~\ref{fig:a53_vs_a7}. 
+As expected, when the model is run on A7 cores, the intereference of the write
+co-runners is not as impactful on model performance as it can still operate at 
+under 10 Hz.
+
 %% Instead, memory controller focused isolation solutions,
@@
 implementation is limited. We find that using all four cores can double 
-performance of the DNN, but compared to using three cores, performance
+performance of the CNN, but compared to using three cores, performance
 improvement is relatively minimal.
@@
 space used. Furthermore, we find that performance remains consistent 
-when multiple models are run simultaneously, and when memory intensive
-read co-runners are introduced. As such, we conclude that 
-the shared L2 cache is not essential to the DNN's real-time performance.
+when memory intensive read co-runners are introduced, even though
+the model has its own isolated cache space. As such, we 
+conclude that the shared L2 cache is not essential to the DNN's real-time 
+performance.
 
 Lastly, we find that memory bandwidth throttling results in significant
-changes to model performance. When less bandwidth is given to models,
-whether they're running alone or concurrently, we find that their performance
-decreases noticably. When memory intensive co-runners are
+changes to model performance. When less bandwidth is given to the model,
+we find that it performs worse. When memory intensive co-runners are
 present, we find that limiting the available bandwidth of the co-runners
-actually results in improved model performance. As such, we conclude
+results in significantly improved model execution times. As such, we conclude
 that the shared DRAM memory is essential to the DNN's real-time 

### 2018-04-15 paper/main.tex
Updates to paper

@@
-\documentclass[a4paper,UKenglish]{lipics-v2016}
-%This is a template for producing LIPIcs articles. 
-%See lipics-manual.pdf for further information.
-%for A4 paper format use option "a4paper", for US-letter use option "letterpaper"
-%for british hyphenation rules use option "UKenglish", for american hyphenation rules use option "USenglish"
-% for section-numbered lemmas etc., use "numberwithinsect"
-
-\usepackage{microtype}%if unwanted, comment out or use option "draft"
+\documentclass[conference]{IEEEtran}
+\IEEEoverridecommandlockouts
+% The preceding line is only needed to identify funding in the first footnote. If that is unneeded, please comment it out.
+\usepackage{cite}
+\usepackage{amsmath,amssymb,amsfonts}
+\usepackage{algorithmic}
+\usepackage{graphicx}
+\usepackage{textcomp}
 \usepackage{adjustbox}
+\usepackage{subcaption}
+\usepackage{listings}
 \usepackage{color}
-\def\fixme#1{\typeout{FIXED in page \thepage : {#1}}
-%  \bgroup \color{red}{} \egroup}
-\bgroup \color{red}{[FIXME: {#1}]} \egroup}
-
-%\graphicspath{{./graphics/}}%helpful if your graphic files are in another directory
-
-\bibliographystyle{plainurl}% the recommended bibstyle
-
-% Author macros::begin %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\title{DeepPicar: A Low-cost Deep Neural Network-based Autonomous Car}
-\titlerunning{DeepPicar: A Low-cost Autonomous Car}
-%optional, in case that the title is too long; the running title should fit into the top page column
-
-%% Please provide for each author the \author and \affil macro, even when authors have the same affiliation, i.e. for each author there needs to be the  \author and \affil macros
-\author[1]{Michael G. Bechtel}
-\author[2]{Elise McEllhiney}
-\author[3]{Minje Kim}
-\author[4]{Heechul Yun}
-\affil[1]{University of Kansas, Lawrence, United States\\
-  \texttt{mbechtel@ku.edu}}
-\affil[2]{University of Kansas, Lawrence, United States\\
-  \texttt{e908m429@ku.edu}}
-\affil[3]{Indiana University, Bloomington, United States\\
-  \texttt{minje@indiana.edu}}
-\affil[4]{University of Kansas, Lawrence, United States\\
-  \texttt{heechul.yun@ku.edu}}
-\authorrunning{M.\,G. Bechtel, E. McEllhiney, M. Kim, H. Yun} %mandatory. First: Use abbreviated first/middle names. Second (only in severe cases): Use first author plus 'et. al.'
-
-\Copyright{Michael G. Bechtel, Elise McEllhiney, Minje Kim, Heechul Yun}%mandatory, please use full first names. LIPIcs license is "CC-BY";  http://creativecommons.org/licenses/by/3.0/
-
-\subjclass{C.4 PERFORMANCE OF SYSTEMS}% mandatory: Please choose ACM 1998 classifications from http://www.acm.org/about/class/ccs98-html . E.g., cite as "F.1.1 Models of Computation". 
-\keywords{Real-time, Autonomous car, Convolutional neural network, Case study}% mandatory: Please provide 1-5 keywords
-% Author macros::end %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%Editor-only macros:: begin (do not touch as author)%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\EventEditors{Sebastian Altmeyer}
-\EventNoEds{1}
-\EventLongTitle{30th Euromicro Conference on Real-Time Systems (ECRTS 2018)}
-\EventShortTitle{ECRTS 2018}
-\EventAcronym{ECRTS}
-\EventYear{2018}
-\EventDate{July 3--6, 2018}
-\EventLocation{Barcelona, Spain}
-\EventLogo{}
-\SeriesVolume{XX}
-\ArticleNo{YY}
-% Editor-only macros::end %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\usepackage{hyperref}
+\usepackage{fixme}
+%\usepackage{authblk}
+\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
+    T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}
 \begin{document}
 
+\title{DeepPicar: A Low-cost Deep Neural Network-based Autonomous Car}
+
+\author{\IEEEauthorblockN{Michael G. Bechtel}
+\IEEEauthorblockA{\textit{Electrical Engineering and Computer Science} \\
+\textit{University of Kansas}\\
+Lawrence, USA \\
+mbechtel@ku.edu}
+\and
+\IEEEauthorblockN{Elise McEllhiney}
+\IEEEauthorblockA{\textit{Electrical Engineering and Computer Science} \\
+\textit{University of Kansas}\\
+Lawrence, USA \\
+e908m428@ku.edu}
+\and
+\IEEEauthorblockN{Minje Kim}
+\IEEEauthorblockA{\textit{Electrical Engineering and Computer Science} \\
+\textit{Indiana University Bloomington}\\
+Bloomington, USA \\
+minje@indiana.edu}
+\and
+\IEEEauthorblockN{Heechul Yun}
+\IEEEauthorblockA{\textit{Electrical Engineering and Computer Science} \\
+\textit{University of Kansas}\\
+Lawrence, USA \\
+heechul.yun@ku.edu}
+}
+
 \maketitle
-\thispagestyle{empty}
+
 \begin{abstract}
@@
 
+\begin{IEEEkeywords}
+Real-time, Autonomous car, Convolutional neural network, Case study
+\end{IEEEkeywords}
+
 %-------------------------------------------------------------------------
@@
 \input{evaluation}
+%\input{evaluation-cachepart}
 \input{comparison}
-\input{discussion}
+%\input{discussion}
 \input{related}
@@
 %-------------------------------------------------------------------------
-%\bibliographystyle{abbrv}
+\bibliographystyle{abbrv}
 \bibliography{reference}
-%% \appendix
+
+\appendix
+\input{appendix}
 \end{document}

### 2018-04-15 paper/overview.tex
Updates to paper

@@
   \begin{subfigure}{0.4\textwidth}
-    \includegraphics[width=\textwidth]{figs/track-2}
+    \includegraphics[width=\textwidth]{figs/track-3}
     \caption{Track 2.}
@@
 timestamped videos and contol commands. The stored data is then copied 
-to a desktop computer, which is equipped with a NVIDIA GTX 1060 GPU, 
+to a desktop computer, which is equipped with a NVIDIA Titan Xp GPU, 
 where we train the network to accelerate training speed. 
@@
 approximately 4 hours, whereas it takes only about 4 minutes on the
-desktop computer using the GTX 1050 Ti GPU.
+desktop computer using the Titan Xp GPU.
 

### 2018-04-16 paper/background.tex
intro - overview revision

@@
 as important, if not moreso, as computational throughput, because many
-use cases have strict real-time requirements (e.g., search query
-latency).
+use cases have strict real-time requirements.
+
+%% (e.g., search query latency)
 
@@
 unique challenges for deep learning, as the computing platforms of such
-systems must satisfy two often conflicting goals~\cite{Otterness2017}:
+systems must satisfy two often conflicting goals:
 %% Recent successes in AI, including NVIDIA's DAVE-2 showing, are due
@@
 (2) The platform must also satisfy various
-constraints such as cost, size, weight, and power consumption limits.
+constraints such as cost, size, weight, and power consumption limits~\cite{Otterness2017}.
 
@@
 necessary computing performance for applying AI technology-based
-robotics systems, and (2) what kind of computing architecture and
-runtime supports are most appropriate for such a workload. To
-achieve these goals, we implement a low-cost autonomous car platform
+robotics systems, and (2) the characteristics of the computing
+platform to support such a workload.
+To achieve these goals, we implement a low-cost autonomous car platform
 as a case-study and systematically conduct experiments, which we will 

### 2018-04-16 paper/evaluation.tex
intro - overview revision

@@
 and a 32K private Data cache, and all cores share a 512KB L2 cache.
-The chip also includes Broadcom's Videocore IV
-GPU, although we did not use the GPU in our evaluation due to the lack
-of sofware support (TensorFlow is not compatible with the Raspberry Pi's GPU).
-For software, we use Ubuntu MATE 16.04, TensorFlow 1.1 and Python
-2.7. We disabled DVFS (dynamic voltage frequency scaling) and
+The chip also includes Broadcom's Videocore IV GPU, although we did
+not use the GPU in our evaluation, due to the lack of sofware support
+\footnote{TensorFlow currently only support NVIDIA's GPUs.}.
+For software, we use Ubuntu MATE 16.04, TensorFlow 1.1.
+We disabled DVFS (dynamic voltage frequency scaling) and
 configured the clock speed of each core statically at the maximum 1.2GHz.
@@
 of the deep neural network, and the fact that the inference operation
-performed by TensorFlow only utilizes the CPU cores of the
-Raspberry Pi 3 (its GPU is not supported by Tensorflow).
+performed by TensorFlow only utilizes the CPU cores of the Raspberry Pi 3.
 In comparison, NVIDIA's DAVE-2 system, which has the exact same neural

### 2018-04-16 paper/intro.tex
Updates to paper
intro - overview revision

@@
 To understand what kind of computing hardware is needed for AI
-workloads, we need a testbed and realistic workloads. While using a real 
-car-based testbed would be most ideal, it is not only highly expensive, but also
-poses serious safety concerns that hinder development and exploration.
-Therefore, there is a strong need for safer and less costly
-testbeds. There are already several relatively inexpensive RC-car
-based testbeds, such as MIT's 
-RaceCar~\cite{shin2017project} and UPenn's F1$/$10 racecar~\cite{upennf1tenth}.
-However, these RC-car testbeds still cost more than \$3,000, requiring
-considerable investment.
+workloads, we need a testbed and realistic workloads. While using a
+real car-based testbed would be most ideal, it is not only highly
+expensive, but also poses serious safety concerns that hinder
+development and exploration. Therefore, there is a need for safer and
+less costly testbeds.
 
-% our goals
-Instead, we want to build a low cost testbed that still employs the
-state-of-the art AI technologies. Specifically, we focus on an end-to-end
-deep learning based real-time control system,
-which was developed for a real self-driving car, NVIDIA
-DAVE-2~\cite{Bojarski2016}, and uses the same methodology on a
-smaller and less costly setup. In developing the testbed, our
-goals are (1) to analyze real-time issues in DNN based end-to-end
-control; and (2) to evaluate real-time performance of contemporary embedded
-platforms for such a workload.
+%% There are already several relatively inexpensive RC-car
+%% based testbeds, such as MIT's 
+%% RaceCar~\cite{shin2017project} and UPenn's F1$/$10 racecar~\cite{upennf1tenth}.
+%% However, these RC-car testbeds still cost more than \$3,000, requiring
+%% considerable investment.
+
+%% % our goals
+%% Instead, we want to build a low cost testbed that still employs the
+%% state-of-the art AI technologies. Specifically, we focus on an end-to-end
+%% deep learning based real-time control system,
+%% which was developed for a real self-driving car, NVIDIA
+%% DAVE-2~\cite{Bojarski2016}, and uses the same methodology on a
+%% smaller and less costly setup. In developing the testbed, our
+%% goals are (1) to analyze real-time issues in DNN based end-to-end
+%% control; and (2) to evaluate real-time performance of contemporary embedded
+%% platforms for such a workload.
 
@@
 In this paper, we present DeepPicar, a low-cost autonomous car
-platform for research. From a hardware perspective,
+testbed for research. From a hardware perspective,
 DeepPicar is comprised of a Raspberry Pi 3 Model B quad-core
@@
 components (less than \$100 in total).
-The DeepPicar, however, employs state-of-the-art AI
-technologies, including a vision-based end-to-end control system that
-utilizes a deep convolutional neural network (CNN).
-The network receives an image frame from a single forward
+The DeepPicar, however, employs a state-of-the-art AI
+technology, namely deep learning based end-to-end real-time control,
+which utilizes a deep convolutional neural network (CNN).
+The CNN receives an image frame from a single forward
 looking camera as input and generates a predicted steering angle
 value as output at each control period in \emph{real-time}.
-The network has 9 layers, about 27 million connections
+The CNN has 9 layers, about 27 million connections
 and 250 thousand parameters (weights).
-The network architecture is almost identical to that of NVIDIA's DAVE-2
-self-driving car~\cite{Bojarski2016}, which uses a significantly more powerful
-computer (Drive PX computer~\cite{drivepx}) than a Raspberry Pi~3.
-We chose to use a Pi 3 not only because it is affordable, but also because it is representative
-of today's mainstream low-end embedded multicore platforms found in
-smartphones and other embedded devices.
+DeepPicar's CNN architecture is identical to that of NVIDIA's
+real-sized self-driving car, called DAVE-2~\cite{Bojarski2016}, which
+drove on public roads without human driver's intervention while only
+using the CNN, which runs on a significantly more powerful computing
+platform (Drive PX~\cite{drivepx}).
+
+%% , which uses a significantly more
+%% powerful computer (Drive PX computer~\cite{drivepx}) than a Raspberry
+%% Pi~3.
+%% We chose to use a Pi 3 not only because it is affordable, but 
+%% also because it is representative of today's mainstream low-end
+%% embedded multicore platforms found in smartphones and other embedded
+%% devices.
 
@@
 % how we trained (maybe moved to a later section)
-We apply a standard imitation learning methodology to train the car to
-follow tracks on the ground. We collect data for
-training and validation by manually
-controlling the RC car and recording the vision (from the webcam
-mounted on the RC-car) and the human control inputs. We then train the
-network offline using the collected data on a desktop computer, which
-is equipped with a NVIDIA GTX 1060 GPU. Finally, the trained network is copied
-back to the Raspberry Pi 3, which is then used to perform inference
-operations---locally on the Pi 3---in the car's main control loop in
-real-time. For real-time control, each inference operation must
-be completed within the desired control period (e.g., 33.$\overline{\mbox{33}}$ ms
- period for 30 Hz control frequency).
+%% We apply a standard imitation learning methodology to train the car to
+%% follow tracks on the ground. We collect data for
+%% training and validation by manually
+%% controlling the RC car and recording the vision (from the webcam
+%% mounted on the RC-car) and the human control inputs. We then train the
+%% network offline using the collected data on a desktop computer, which
+%% is equipped with a NVIDIA GTX 1050 Ti GPU. Finally, the trained network is copied
+%% back to the Raspberry Pi 3, which is then used to perform inference
+%% operations---locally on the Pi 3---in the car's main control loop in
+%% real-time. For real-time control, each inference operation must
+%% be completed within the desired control period (e.g., 33.$\overline{\mbox{33}}$ ms
+%%  period for 30 Hz control frequency).
 % how we evaluated (in terms real-time performance)
@@
 % what are our findings?
-Using the DeepPicar platform, we systematically analyze its real-time
+Using DeepPicar, we systematically analyze its real-time
 capabilities in the context of deep-learning based real-time
-control, especially on real-time deep neural network inferencing.
+control, especially on real-time deep neural network \emph{inferencing}.
 We also evaluated other, more powerful, embedded computing
@@
 From the systematic study, we have made a number of interesting
-observations from the perspective of real-time systems.
+observations.
 % predictability
-First, we find that DNN inferencing is highly predictable---from the
-timing perspective---as the amount of computation needed to complete a
-single inference is fixed at the DNN architecture design time and does
-not change at runtime over different inputs (e.g., different image
+First, we find that CNN inferencing is highly time predictable as the
+amount of computation needed to complete a 
+single inference is fixed at the network architecture design time and
+does not change at runtime over different inputs (e.g., different image
 frames). This predictable timing behavior is obviously a desirable
@@
 % performance on low-cost embedded multicore
-Second, we find that real-time processing of DeepPicar's CNN is
+Second, we find that real-time processing of CNN inferencing is
 feasible on today's embedded computing platforms, even one   
-as inexpensive as the Raspberry Pi 3. We find that the control loop 
-completes in under 33.$\overline{\mbox{33}}$ ms, or 30 Hz, using just two 
-cores of the Raspberry Pi 3's quad-core CPU (w/o using its GPU), and can
-do so 100\% of the time; or we can achieve 20 Hz control frequency using
-just one core.
-Other tested embedded platforms, the Intel UP and NVIDIA TX2, offer even
-better performance, and are capable of supporting deep-learning based 
-real-time control over 100 Hz control frequency on the TX2 when its
-embedded GPU is used.
+as inexpensive as the Raspberry Pi 3. For example, just one Cortex-A53
+core of the Raspberry Pi3 is able to complete the CNN inferencing task
+in less than 50ms, sufficient to support 20 Hz control loop. 
+%% We find that the control loop 
+%% completes in under 33.$\overline{\mbox{33}}$ ms, or 30 Hz, using just two 
+%% cores of the Raspberry Pi 3's quad-core CPU (w/o using its GPU), and can
+%% do so 100\% of the time; or we can achieve 20 Hz control frequency using
+%% just one core.
+%% Other tested embedded platforms, the Intel UP and NVIDIA TX2, offer even
+%% better performance, and are capable of supporting deep-learning based 
+%% real-time control over 100 Hz control frequency on the TX2 when its
+%% embedded GPU is used.
 We find that all the tested embedded computing platforms have 
@@
 issue when consolidating multiple workloads. 
-In particular, we find that DNN inferencing is highly sensitive to
-memory performance interference, as we observe up to an 11.6X control
-loop execution time increase when memory performance intensive
-applications are co-scheduled on idle CPU cores.
-On the other hand, we find that DNN inferencing is not sensitive to
-cache space.
+We observe up to 11.6X control
+loop execution time increase due to shared resource contention when
+memory intensive  applications are co-scheduled on idle CPU cores.
+We find that DeepPicar's inferencing task is highly
+sensitive to memory bandwidth, but not cache space.
+
 %% ; we show that applying cache partitioning has negligible impact
 %% in providing timing isolation on consolidated scenarios.
-This finding
-suggests that in order to guarantee real-time performance of DNN-based
-control applications, solutions to guarantee their memory performance
-are more important.
+%% This finding
+%% suggests that in order to guarantee real-time performance of DNN-based
+%% control applications, solutions to guarantee their memory performance
+%% are more important.
 
@@
 \begin{itemize}
-  \item We present DeepPicar, a low-cost autonomous car testbed, which
-    is based on a state-of-the-art end-to-end deep learning, as 
-    open-source~\footnote{The source code, and build 
-instructions for the DeepPicar
-can be found at: \url{https://github.com/heechul/picar}}, with the
-    goal of lowering the economic and safety-related barriers to the study
-    of autonomous cars.
+  \item We present DeepPicar, a low-cost autonomous car testbed using
+    a state-of-the-art CNN for end-to-end real-time control, to lower
+    economic and safety-related barriers to the study autonomous
+    cars.~\footnote{The source code, dataset, and build instructions
+      of DeepPicar can be found at:\fixme{Michael github url}}
 
-  \item We systematically analyze real-time characteristics of DeepPicar's
-    DNN inferencing workload on a number of representative
-    contemporary embedded computing platforms. While larger embedded
-    platforms have been evaluated, as in  \cite{Otterness2017}, to the
-    best of our knowledge, no prior work investigated the
-    impacts of shared resource contention on DNN inferencing workloads
-    in the context of smaller embedded platforms like the Raspberry Pi 3.
+  \item We provide extensive systematic experimental results using
+    DeepPicar's CNN inferencing workload on a number of representative
+    contemporary embedded computing platforms. The results provide
+    insights for real-time processing of the CNN workload and the
+    effect of consolidation.
 
+    %% While larger embedded
+    %% platforms have been evaluated, as in  \cite{Otterness2017}, to the
+    %% best of our knowledge, no prior work investigated the
+    %% impacts of shared resource contention on DNN inferencing workloads
+    %% in the context of smaller embedded platforms like the Raspberry Pi 3.
 \end{itemize}

### 2018-04-16 paper/overview.tex
intro - overview revision

@@
 used in this paper, which is comprised of 9 layers, 250K parameters,
-and about 27 million connections as in NVIDIA DAVE-2's. Note, however,
-that we did not apply the normalization mentioned
-in~\cite{Bojarski2016}, as it does not include trainable parameters
-and its computational demand with respect to the overall CNN
-processing is minimal.
+and about 27 million connections as in NVIDIA DAVE-2's.
+%% Note, however,
+%% that we did not apply the normalization mentioned
+%% in~\cite{Bojarski2016}, as it does not include trainable parameters
+%% and its computational demand with respect to the overall CNN
+%% processing is minimal.
 
@@
 where we train the network to accelerate training speed.
-For comparison, training the network on the Raspbeery Pi 3 takes
-approximately 4 hours, whereas it takes only about 4 minutes on the
-desktop computer using the Titan Xp GPU.
+
+%% For comparison, training the network on the Raspbeery Pi 3 takes
+%% approximately 4 hours, whereas it takes only about 4 minutes on the
+%% desktop computer using the Titan Xp GPU.
 
@@
 is copied back to the Raspberry Pi 3. The network is then used
-by the car's main controller, which feeds an image frame from the web
-camera as input to the network. In each control period, the produced 
-steering angle output is then converted into the PWM values of the 
-steering motor of the car. Figure~\ref{fig:controlloop} shows simplified 
+by the car's main controlller, which feeds an image frame from the web
+camera as input to the network at each control period. The network
+produced steering angle output is then converted into the PWM values
+of the car's steering motor. Figure~\ref{fig:controlloop} shows simplified 
 pseudo code of the controller's main loop. Among the five steps, the 3rd step, 
@@
 introduce inaccuracy in control.
-In the future, we plan to use a different (more expensive) RC car
-platform that can precisely control the car's steering angle.
+%% In the future, we plan to use a different (more expensive) RC car
+%% platform that can precisely control the car's steering angle.
 
@@
 (motor) control latencies. In DeepPicar's
-context, the camera latency  measured from the time the camera
+context, the camera latency is from the time the camera
 sensor observes the scene to the time the computer actually reads the
@@
 applied 50 ms later. Considering that camera's framerate is 30Hz
-(33.$\overline{\mbox{33}}$ ms/frame), this is about two frames after the 
-control action.
+(33.3 ms/frame), this is about two frames after the control action.
 %% We experimentally measured the camera
@@
 throttle)~\footnote{A collection of self-driving videos can be found
-  at: \url{https://photos.app.goo.gl/ce93sU7jPk4ywO8u2}}. Running at
+  at: \fixme{Michael: link your video}
+  %% \url{https://photos.app.goo.gl/ce93sU7jPk4ywO8u2}
+}. Running at
 higher speed is inherently more challenging because the CNN controller
@@
 
-We would like to stress, however, that the issues related to the
-CNN's accuracies have no impact on the \emph{computational 
-aspects of the system}, and that our main focus of this study is
-not in improving network accuracy but in closely replicating the
-DAVE-2's network architecture and studying its real-time
-characteristics.
+We would like to stress, however, that
+%% the issues related to the
+%% CNN's accuracies have no impact on the \emph{computational 
+%% aspects of the system}, and that 
+our main focus of this study is not in improving the network accuracy
+but in closely replicating the DAVE-2's network architecture and
+studying its real-time characteristics, which will be presented in the
+subsequent section.

### 2018-04-17 paper/main.tex
intro - overview revision
author style fix

@@
 \usepackage{hyperref}
-\usepackage{fixme}
+% \usepackage{fixme}
 %\usepackage{authblk}
+\def\fixme#1{\typeout{FIXED in page \thepage : {#1}}
+%  \bgroup \color{red}{} \egroup}
+\bgroup \color{red}{[FIXME: {#1}]} \egroup}
+
 \def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
@@
 
-\author{\IEEEauthorblockN{Michael G. Bechtel}
-\IEEEauthorblockA{\textit{Electrical Engineering and Computer Science} \\
-\textit{University of Kansas}\\
-Lawrence, USA \\
-mbechtel@ku.edu}
-\and
-\IEEEauthorblockN{Elise McEllhiney}
-\IEEEauthorblockA{\textit{Electrical Engineering and Computer Science} \\
-\textit{University of Kansas}\\
-Lawrence, USA \\
-e908m428@ku.edu}
-\and
-\IEEEauthorblockN{Minje Kim}
-\IEEEauthorblockA{\textit{Electrical Engineering and Computer Science} \\
-\textit{Indiana University Bloomington}\\
-Bloomington, USA \\
-minje@indiana.edu}
-\and
-\IEEEauthorblockN{Heechul Yun}
-\IEEEauthorblockA{\textit{Electrical Engineering and Computer Science} \\
-\textit{University of Kansas}\\
-Lawrence, USA \\
-heechul.yun@ku.edu}
+\author{Michael G. Bechtel$^\dagger$, Elise McEllhiney$^\dagger$, Minje Kim$^\star$, Heechul Yun$^\dagger$\\
+  $^\dagger$ University of Kansas, USA. \{mbechtel, e908m428, heechul.yun\}@ku.edu \\
+  $^\star$ Indiana University, USA. minje@indiana.edu \\
 }
+
+%% \IEEEauthorblockA{\textit{Electrical Engineering and Computer Science} \\
+%% \textit{University of Kansas}\\
+%% Lawrence, USA \\
+%% mbechtel@ku.edu}
+%% \and
+%% \IEEEauthorblockN{Elise McEllhiney}
+%% \IEEEauthorblockA{\textit{Electrical Engineering and Computer Science} \\
+%% \textit{University of Kansas}\\
+%% Lawrence, USA \\
+%% e908m428@ku.edu}
+%% \and
+%% \IEEEauthorblockN{Minje Kim}
+%% \IEEEauthorblockA{\textit{Electrical Engineering and Computer Science} \\
+%% \textit{Indiana University}\\
+%% Bloomington, USA \\
+%% minje@indiana.edu}
+%% \and
+%% \IEEEauthorblockN{Heechul Yun}
+%% \IEEEauthorblockA{\textit{Electrical Engineering and Computer Science} \\
+%% \textit{University of Kansas}\\
+%% Lawrence, USA \\
+%% heechul.yun@ku.edu}
+%% }
 

### 2018-04-17 paper/evaluation.tex
revision up to Section 4.E

@@
 contention affects inference timing. These instances are not identical
-copies of the same model, but are instead copies of different models. 
+copies of the same model, but are instead copies of different models
+(but of the same network architecture). 
 This is done to ensure that no L2 cache memory is shared between the 
@@
 These increases in inference times in the co-scheduled scenarios are
-expected and are caused by contention in the shared hardware
-resources, such as the shared L2 cache and/or the DRAM controller.
+expected due to contention in the shared hardware
+resources, such as the shared cache and DRAM.
 
@@
 
-\subsection{Effect of Co-scheduling Memory Performance Hogs}\label{sec:eval-memhog}
+\subsection{Effect of Co-scheduling Synthetic Benchmarks}\label{sec:eval-memhog}
 
-In this experiment, we investigate the impact of contended DRAM
-requests to the DNN inference timing of DeepPicar.
-For the experiment, we use a synthetic memory
-intensive benchmark from the IsolBench suite~\cite{Valsan2016}.
-We run a single DNN model on one core, and
-co-schedule an increasing number of the memory intensive synthetic
-benchmarks~\footnote{We use the \emph{Bandwidth} benchmark in the
-  IsolBench suite, with the following command line parameters: \texttt{\$
-  bandwdith -a write -m 16384 -t 1000}}, on the remaining idle cores.
+In this experiment, we further investigate the impact of shared
+resource contention to the DNN inference timing of DeepPicar using  
+synthetic memory benchmarks. Specifically, we use the \emph{Bandwidth}
+benchmark from the IsolBench suite~\cite{Valsan2016}, which
+sequentially reads or writes a big array; we henceforth refer BwRead 
+as Bandwidth with read accesses and BwWrite as the one with write
+accesses. The experiment setup is as follows. We run a single DNN
+model on one core, and co-schedule an increasing number of the
+Bandwidth benchmark instances on the other cores. We repeat the
+experiement first with BwRead~\footnote{\texttt{\$ bandwdith -a read
+    -m 16384 -t 1000}} and next with BwWrite~\footnote{\texttt{\$
+    bandwdith -a write -m 16384 -t 1000}}.
 
@@
 
-\begin{figure}[h]
-  \centering
-  \includegraphics[width=.45\textwidth]{figs/perf_vs_bandwidth_mem}
-  \caption{Effect of memory performance hogs on the shared resources. 
-    The DNN model uses Core 0 and memory-hog co-runners
-    use the rest of the cores.}
-  \label{fig:perf_vs_bandwidth_mem}
-\end{figure}
+%% \begin{figure}[h]
+%%   \centering
+%%   \includegraphics[width=.45\textwidth]{figs/perf_vs_bandwidth_mem}
+%%   \caption{Effect of memory performance hogs on the shared resources. 
+%%     The DNN model uses Core 0 and memory-hog co-runners
+%%     use the rest of the cores.}
+%%   \label{fig:perf_vs_bandwidth_mem}
+%% \end{figure}
 
-Figure~\ref{fig:perf_vs_bandwidth} shows the execution time of the DNN 
-model and Figure~\ref{fig:perf_vs_bandwidth_mem} shows both the normalized
-total memory bus accesses (first group) and L2 miss rates (second group).
-In both figures, the DNN model is running on Core 0 as a function of the 
-number of co-scheduled memory intensive synthetic benchmarks.
-%% We also show 
-%% the normalized total system bus access count (middle group), which is
-%% representative of memory bandwidth usage of all the cores.
-First, as we
-increase the number of co-runners, the DNN model's execution times are
-increased exponentially---by up to 11.6X---even though the DNN model
-is running on a dedicated core (Core 0). On the other hand, the DNN
-model's L2 cache-miss rates do not increase as much.
-In other words, the DNN model's execution increase is not fully
-explained by increased contention in the L2 cache space.
-Instead, the exponential increase in the total number of bus accesses,
-due to the memory-hog co-runners, appears to be more closely
-correlated with the DNN model's execution time increase. This means
-the high memory bandwidth pressure from the co-scheduled
-memory intensive benchmarks increased the DNN model's
-memory access latency.
+Figure~\ref{fig:perf_vs_bandwidth} shows the execution times of the DNN 
+model against varying number of BwRead and BwWrite co-runners. First,
+note that BwWrite cause significantly high execution time incrase---up
+to 11.6X---on the DNN inferencing timing while BwRead causes
+relatively much smaller time increase. While execution time increase
+was expected, the degree of which is quite surprising---\emph{if the DNN
+controller was driving an actual car, it would result in crash of the
+car!}
+
+\fixme{story: is cache space contention the reason? traditional
+  solution is cache partitioning. so, next we evaluate if cache
+  partitioning can solve the problem. spolier alert: it does not.}
+
+%% Figure~\ref{fig:perf_vs_bandwidth_mem} shows both the normalized
+%% total memory bus accesses (first group) and L2 miss rates (second group).
+%% In both figures, the DNN model is running on Core 0 as a function of the 
+%% number of co-scheduled memory intensive synthetic benchmarks.
+%% %% We also show 
+%% %% the normalized total system bus access count (middle group), which is
+%% %% representative of memory bandwidth usage of all the cores.
+%% First, as we
+%% increase the number of co-runners, the DNN model's execution times are
+%% increased exponentially---by up to 11.6X---even though the DNN model
+%% is running on a dedicated core (Core 0). On the other hand, the DNN
+%% model's L2 cache-miss rates do not increase as much.
+%% In other words, the DNN model's execution increase is not fully
+%% explained by increased contention in the L2 cache space.
+%% Instead, the exponential increase in the total number of bus accesses,
+%% due to the memory-hog co-runners, appears to be more closely
+%% correlated with the DNN model's execution time increase. This means
+%% the high memory bandwidth pressure from the co-scheduled
+%% memory intensive benchmarks increased the DNN model's
+%% memory access latency.
 
@@
 
-This observation suggests that shared cache partitioning
-techniques~\cite{Gracioli2015,Kim2016} may not be effective isolation
-solutions for DeepPicar's DNN processing workload, as cache
-partitioning does not provide memory performance guarantees.
+%% This observation suggests that shared cache partitioning
+%% techniques~\cite{Gracioli2015,Kim2016} may not be effective isolation
+%% solutions for DeepPicar's DNN processing workload, as cache
+%% partitioning does not provide memory performance guarantees.
 
@@
   \caption{DNN model performance on different in-order
-ARM Cortex cores.}
+ARM Cortex cores. \fixme{this may be better in the plaform comparison section}.}
   \label{fig:a53_vs_a7}
@@
 
-We do find the performance of the model in the presence of 
-write co-runners surprising, as it has been previously shown in experiments
-on another in-order ARM core, the Cortex A7, that write interference was not as 
-significant\cite{Valsan2016}. We also tested the model, in the presence of write 
-co-runners, on the A7 cores of the Odroid XU4 platform to see if  the same model 
-behavior would also be seen there. The results can be seen in Figure ~\ref{fig:a53_vs_a7}. 
-As expected, when the model is run on A7 cores, the intereference of the write
-co-runners is not as impactful on model performance as it can still operate at 
-under 10 Hz.
+%% We do find the performance of the model in the presence of 
+%% write co-runners surprising, as it has been previously shown in experiments
+%% on another in-order ARM core, the Cortex A7, that write interference was not as 
+%% significant\cite{Valsan2016}. We also tested the model, in the presence of write 
+%% co-runners, on the A7 cores of the Odroid XU4 platform to see if  the same model 
+%% behavior would also be seen there. The results can be seen in Figure ~\ref{fig:a53_vs_a7}. 
+%% As expected, when the model is run on A7 cores, the intereference of the write
+%% co-runners is not as impactful on model performance as it can still operate at 
+%% under 10 Hz.
 

### 2018-04-17 paper/evaluation.tex
revision 4.E (2nd)

@@
 cores/DNN. For example, '4Nx1C' means running four DNN models each of
-which is assigned to run on one CPU core, whereas '2Nx2C' means running
+which is assigned to run on one CPU core, whereas `2Nx2C' means running
 two DNN models, each of which is assigned to run on two CPU
@@
 These increases in inference times in the co-scheduled scenarios are
-expected due to contention in the shared hardware
-resources, such as the shared cache and DRAM.
+expected because it is well-known that co-scheduled tasks on a
+multicore platform interfere with each other due to contention in the
+shared hardware resources, such as the shared
+cache~\cite{Gracioli2015} and DRAM~\cite{Yun2013}.
 
@@
 
-\subsection{Effect of Co-scheduling Synthetic Benchmarks}\label{sec:eval-memhog}
+\subsection{Effect of Co-scheduling Synthetic Memory Intensive Tasks}\label{sec:eval-memhog}
 
-In this experiment, we further investigate the impact of shared
-resource contention to the DNN inference timing of DeepPicar using  
-synthetic memory benchmarks. Specifically, we use the \emph{Bandwidth}
+In this experiment, we investigate \emph{worst-case} impact of shared 
+resource contention to DeepPicar's DNN inference timing using  
+a synthetic memory benchmark. Specifically, we use the \emph{Bandwidth}
 benchmark from the IsolBench suite~\cite{Valsan2016}, which
@@
 Bandwidth benchmark instances on the other cores. We repeat the
-experiement first with BwRead~\footnote{\texttt{\$ bandwdith -a read
-    -m 16384 -t 1000}} and next with BwWrite~\footnote{\texttt{\$
-    bandwdith -a write -m 16384 -t 1000}}.
+experiement first with BwRead and next with BwWrite.
+
+%% ~\footnote{\texttt{\$ bandwdith -a read  -m 16384 -t 1000}} 
+%% ~\footnote{\texttt{\$ bandwdith -a write -m 16384 -t 1000}}.
 
@@
 Figure~\ref{fig:perf_vs_bandwidth} shows the execution times of the DNN 
-model against varying number of BwRead and BwWrite co-runners. First,
-note that BwWrite cause significantly high execution time incrase---up
+model against varying number of BwRead and BwWrite co-runners. Note
+that BwWrite cause significantly high execution time incrase---up 
 to 11.6X---on the DNN inferencing timing while BwRead causes
@@
 was expected, the degree of which is quite surprising---\emph{if the DNN
-controller was driving an actual car, it would result in crash of the
+controller was driving an actual car, it would result in a crash of the
 car!}
 
-\fixme{story: is cache space contention the reason? traditional
-  solution is cache partitioning. so, next we evaluate if cache
-  partitioning can solve the problem. spolier alert: it does not.}
+Given the importance of predictable timing on real-time control, such
+as our DNN based control task, we wanted to understand (1) the main
+source of the timing and (2) to evaluate existing isolation methods to
+avoid this kind of timing interference. Specifically, shared cache
+space and DRAM bandwidth are two most well-known sources of contention
+in multicore. Thus, in the following, we investigate test if and how
+they influence the observed timing interference and how effective are
+the existing mitigation methods. To our surprise, we find that neither
+the cache space nor DRAM bandwidth is the root cause of the
+DeepPicar's observed worst-case execution time increase.
+
+%% \fixme{story: is cache space contention the reason? traditional
+%%   solution is cache partitioning. so, next we evaluate if cache
+%%   partitioning can solve the problem. spolier alert: it does not.}
 
@@
 %% partitioning does not provide memory performance guarantees.
-
-\begin{figure}[h]
-  \centering
-  \includegraphics[width=.45\textwidth]{figs/a53_vs_a7}
-  \caption{DNN model performance on different in-order
-ARM Cortex cores. \fixme{this may be better in the plaform comparison section}.}
-  \label{fig:a53_vs_a7}
-\end{figure}
 

### 2018-04-17 paper/evaluation-cachepart.tex
Section 4.F revision

@@
 \subsection{Effect of Cache Partitioning}
-
-\begin{figure}
-  \centering
-  \includegraphics[width=.5\textwidth]{figs/cache-mapping}
-  \caption{A memory mapping of the cache sets for the Broadcom 
-BCM2837 processor used in the Raspberry Pi 3 Model B.}
-  \label{fig:cache-mapping}
-\end{figure}
 
@@
 
-In order to validate the effectiveness of cache partitioning, we patch
-the kernel with PALLOC~\cite{yun2014rtas}, which is a kernel-level
-color-aware page allocator, supporting cache partitioning. 
-Figure~\ref{fig:cache-mapping} shows the physical address mapping of
-Raspberry Pi3's BCM2837 processor. We use the bits 13 and 14 for
-coloring, which results in 4 page colors.
+Cache partitioning is a well-known technique to improve isolation in
+real-time systems by giving a dedicated cache space to individaul task
+or core. In this experiment, we use PALLOC~\cite{yun2014rtas}, a
+page-coloring based cache partitioning solution for Linux, to
+investigate the effect of cache partitioning in protecting DeepPicar's
+DNN based controller.
+
+\begin{figure}
+  \centering
+  \includegraphics[width=.5\textwidth]{figs/cache-mapping}
+  \caption{Physical address mapping of L1/L2 caches of Broadcom
+    BCM2837 processor in Raspberry Pi 3.}
+  \label{fig:cache-mapping}
+\end{figure}
+
+Page coloring is a OS technique that controls the physical addresses
+of memory pages. By allocating pages over non-overlapping sets of the
+cache, the OS can effectively partition the cache.
+Figure~\ref{fig:cache-mapping} shows the physical address
+mapping of Raspberry Pi 3's BCM2837 processor, which has 32K private
+L1 I\&D (4way) caches and a share 512KB L2 (16 way) cache. In order not
+to partition the private L1 caches, we use the bit 13 and 14 for
+coloring, which results in 4 usuable page colors.
 
@@
   \includegraphics[width=.45\textwidth]{figs/palloc_multicore}
-  \caption{Cache space sensitivity of the control loop exectuion time.}
+  \caption{Cache space sensitivity of the DNN controller. \fixme{I
+      think 'no part.' bar is not necessary here.}}
   \label{fig:palloc_multicore}
@@
 In the first experiment, we investigate the cache space sensitivity of
-the DeepPicar's DNN-based control loop. Using PALLOC, we create 4
-different cgroups which are configured to use 1, 2,
-3, and 4 colors (25\%, 50\%, 75\%, and 100\% of the L2 cache
-space, respectively). We then execute the control loop on each cache
-partition (cgroup) and measure the average processing time.
+DeepPicar's DNN-based control loop. Using PALLOC, we create 4
+different cgroups which are configured to use 4, 3,
+2, and 1 colors (100\%, 75\%, 50\% and 25\% of the L2 cache
+space, respectively). We then execute the DNN control loop (inference)
+on each cache partition (cgroup) and measure the average processing
+time.
 
-If performance improves as a result of partitioning the shared cache 
-then we know that the DNN, to some extent, relies on the shared cache 
-in addition to the shared DRAM. On the other 
-hand, if there is no improvement in performance, then it can be 
-observed that the shared memory is more critical for the DNN.
+Figure~\ref{fig:palloc_multicore} shows the result. As can be seen,
+the DNN inference timing is hardly changed at all regardless of the
+size of the allocated L2 cache space. In other words, we find that
+the DNN workload is largely insensitive to L2 cache space.
+
+The next experiment further validates this finding. In this
+experiment, we repeat the experiment in
+Section~\ref{sec:eval-memhog}---i.e., co-scheduling the DNN model and
+three Bandwidth (BwRead or BwWrite) instances---but this time we
+ensure that each task is given equal amounts of L2 cache space by
+assigning one color to each task's cache partition.
+
+\begin{figure}[h]
+  \centering
+  \includegraphics[width=.45\textwidth]{figs/palloc_bandwidth_exectime}
+  \caption{Average processing time vs. the number of memory
+memory intensive co-runners. Unlike
+Figure~\ref{fig:perf_vs_bandwidth}, the shared L2 cache is equally
+partitioned and each task is given a dedicated cache
+partition. \fixme{how about changing the figure style to be the same
+  as Figure 9.}}
+  \label{fig:palloc_bandwidth_exectime}
+\end{figure}
+
+Figure \ref{fig:palloc_bandwidth_exectime} shows the
+result. For BwRead, when cache partitioning is enabled, DNN
+performance becomes slightly better, with a improvement of  
+\textasciitilde6.5\%. For BwWrite, cache partitioning actually makes
+the DNN performance slightly worse that that of w/o partitioning. In
+either case, the main observation is that cache partitioning is not
+effective in providing timing isolation to our DNN workload.
+
+%% By partitioning the shared L2 cache, we find that no noticable 
+%% improvements are gained and that performance remains consistent. 
+%% In all experiements, the model shows no sensitivity to the shared L2 
+%% cache. As a result, we conclude that cache partitioning is not an 
+%% effective isolation mechanism, and that the performance of the shared 
+%% DRAM controller is of greater importance to the real-time efficiacy of 
+%% the DeepPicar.
+
+%% If performance improves as a result of partitioning the shared cache 
+%% then we know that the DNN, to some extent, relies on the shared cache 
+%% in addition to the shared DRAM. On the other 
+%% hand, if there is no improvement in performance, then it can be 
+%% observed that the shared memory is more critical for the DNN.
+
 %% We run the DNN model with different amounts of L2 cache being
@@
 %% all four of the available cores.
-The results can be seen in 
-\ref{fig:palloc_multicore}. As expected, the amount of L2 cache 
-available to the model does not result in any significant changes in 
-performance. 
+
+%% As expected, the amount of L2 cache 
+%% available to the model does not result in any significant changes in 
+%% performance. 
 %% This can also be seen by the L2 miss rates of the model 
@@
 
-\begin{figure}[h]
-  \centering
-  \includegraphics[width=.45\textwidth]{figs/palloc_bandwidth_exectime}
-  \caption{Timing impact of co-scheduling memory intensive co-runners 
-when cache partitioning is enabled.}
-  \label{fig:palloc_bandwidth_exectime}
-\end{figure}
-
-We also run a single model alongside three memory intensive read co-runners 
-again, but ensure that each task is given equal amounts of L2 cache space.
-For this, we assign one color to each task. The results can be seen in Figure
-\ref{fig:palloc_bandwidth_exectime}. When cache partitioning is enabled, 
-DNN performance becomes slightly better, with a improvement of 
-\textasciitilde6.5\%. However, as the performance improvement is minimal, 
-we do not believe that cache partitioning is effective for protecting the model 
-from interference from other memory intensive processes.
 %% At the same time, though, the L2 miss rates of the model don't correlate 
@@
 %% performance when co-runners are introduced.
-
-By partitioning the shared L2 cache, we find that no noticable 
-improvements are gained and that performance remains consistent. 
-In all experiements, the model shows no sensitivity to the shared L2 
-cache. As a result, we conclude that cache partitioning is not an 
-effective isolation mechanism, and that the performance of the shared 
-DRAM controller is of greater importance to the real-time efficiacy of 
-the DeepPicar.

### 2018-04-17 paper/evaluation.tex
Section 4.F revision

@@
 
-\begin{figure}[h]
-  \centering
-  \includegraphics[width=.45\textwidth]{figs/perf_vs_corecnt}
-  \caption{Average control loop execution time vs. \#of CPU
-    cores.}
-  \label{fig:perf-vs-corecnt}
-\end{figure}
-
 In this experiment, we investigate the scalability of performing
@@
 numbers of CPU cores---from one to four.
+
+\begin{figure}[h]
+  \centering
+  \includegraphics[width=.45\textwidth]{figs/perf_vs_corecnt}
+  \caption{Average control loop execution time vs. \#of CPU
+    cores.}
+  \label{fig:perf-vs-corecnt}
+\end{figure}
 
@@
 X-axis shows the system configuration: \#of DNN models x \#of CPU
-cores/DNN. For example, '4Nx1C' means running four DNN models each of
+cores/DNN. For example, `4Nx1C' means running four DNN models each of
 which is assigned to run on one CPU core, whereas `2Nx2C' means running
@@
 suffers an average inference time increase of approximately 11 ms,
-$\sim$24\%. On the other hand, when two DNN models, each using two CPU
+or 24\%. On the other hand, when two DNN models, each using two CPU
 cores, are co-scheduled (2Nx2C), the average inference timing is increased by
@@
 
-\subsection{Effect of Co-scheduling Synthetic Memory Intensive Tasks}\label{sec:eval-memhog}
+\subsection{Effect of Co-scheduling Synthetic Memory Intensive
+  Tasks}\label{sec:eval-memhog}
 

### 2018-04-17 paper/evaluation-drampart.tex
Combine Figures 14 and 15

@@
   \includegraphics[width=.45\textwidth]{figs/memguard_bandwidth}
-  \caption{Timing impact of co-scheduling memory intensive read
+  \caption{Timing impact of co-scheduling memory intensive
 co-runners when memory bandwidth throttling is enabled. }
@@
 
-\begin{figure}[h]
-  \centering
-  \includegraphics[width=.45\textwidth]{figs/memguard_bwwrite}
-  \caption{Timing impact of co-scheduling memory intensive write
-co-runners when memory bandwidth throttling is enabled. }
-  \label{fig:memguard_bwwrite}
-\end{figure}
+%% \begin{figure}[h]
+%%   \centering
+%%   \includegraphics[width=.45\textwidth]{figs/memguard_bwwrite}
+%%   \caption{Timing impact of co-scheduling memory intensive write
+%% co-runners when memory bandwidth throttling is enabled. }
+%%   \label{fig:memguard_bwwrite}
+%% \end{figure}
 
@@
 of their reservation size. Due to this, we give write co-runners an initial 
-reservation size of 250 MB/s and decrease it to a minimum of 50 MB/s.
-The results can be see in Figures \ref{fig:memguard_bandwidth} and
-\ref{fig:memguard_bwwrite}. When less memory bandwidth is given to 
-the both the  read and write co-runners, the performance of the model 
-improves. 
+reservation size of 250 MB/s and decrease it to a minimum of 50 MB/s, but
+their effective bandwidths are the same as the read co-runners.
+The results can be see in Figures \ref{fig:memguard_bandwidth}. 
+When less memory bandwidth is given to both the read and write 
+co-runners, the performance of the model improves. 
 %% In the case of one co-runner, the performance of the model remains 

### 2018-04-18 paper/evaluation-cachepart.tex
revision up to 4.H

@@
 Cache partitioning is a well-known technique to improve isolation in
-real-time systems by giving a dedicated cache space to individaul task
-or core. In this experiment, we use PALLOC~\cite{yun2014rtas}, a
+real-time systems by giving a dedicated cache space to each individaul
+task or core. In this experiment, we use PALLOC~\cite{yun2014rtas}, a
 page-coloring based cache partitioning solution for Linux, to
@@
 \textasciitilde6.5\%. For BwWrite, cache partitioning actually makes
-the DNN performance slightly worse that that of w/o partitioning. In
-either case, the main observation is that cache partitioning is not
-effective in providing timing isolation to our DNN workload.
+the DNN performance slightly worse that that of w/o partitioning.
+
+In summary, we find that the DNN inferencing workload is not sensitive
+to cache space and that cache partitioning is not effective in
+providing timing isolation to our DNN workload.
 

### 2018-04-18 paper/evaluation-drampart.tex
revision up to 4.H

@@
 
-To determine the how sensitive the model is to the shared DRAM 
-memory, we use MemGuard\cite{yun2014rtas}, a memory bandwidth reservation system that 
-can limit the amount of bandwidth each core receives in a period. For 
-all tests run using MemGuard, we use a period of one second.
+In this subsection, we examine the DNN workload's memory bandwidth
+sensitivity and the effect of memory bandwidth throttling in providing
+isolation. For the experiments, we use MemGuard\cite{Yun2013}, which 
+can limit the amount of memory bandwidth each core receives. MemGuard
+operates periodically, at an 1ms interval, and uses hardware
+performance counters to throttle cores if they exceed their given
+bandwidth budgets within each regulation period (i.e., 1ms) by
+scheduling high-priority idle kernel threads.
+
+In the first experiment, we measure the performance of the DNN model
+running on a single core, Core 0, first w/o using MemGuard and then w/
+using MemGuard while varying the core's bandwidth throttling parameter
+from 500 MB/s down to 100 MB/s.
 
@@
 
-We first run a single model on one core, Core 0, while varying the 
-core's bandwidth reservation size, from 500 MB/s down to 100 MB/s. 
-The results can be seen in Figure \ref{fig:memguard_multicore}. When less 
-memory bandwidth is available to the core the model's performance 
-noticably decreases.
-%% , and it performs the best when no memory bandwidth 
-%% throttling is implemented. 
-Compared to a reservation size of 100 
-MB/s, disabling memory bandwidth throttling results in 
-\textasciitilde3x performance improvement. If the model was memory 
-insensitive, then the amount of bandwidth available to it would have 
-little effect on its performance. However, since that is not the 
-case, we conclude that, to a notable extent, the neural network model 
-is dependent on shared DRAM memory performance.
+Figure \ref{fig:memguard_multicore} shows the result. When the DNN
+model executing core is throttled at 400 MB/s or more, performance of
+the model is largely the same as the non-throttled case. However, as
+we decreate the assigned memory badnwdith below 300 MB/s, we start to
+observe noticably decreases in the model's performance. In other
+words, the DNN model is sensitive to memory bandwidth and that it
+requires 400 MB/s or more bandwidth should be given to ensure ideal
+performance.
+
+In the next experiment, we again repeat the experiment in
+Section~\ref{sec:eval-memhog}---i.e., co-scheduling the DNN model and
+three Bandwidth (BwRead or BwWrite) instances---but this time we
+throttle the cores that execute the co-runners using MemGuard by 
+varying their bandwidth budgets to see ther impact in the DNN model's
+performance.
+
+\begin{figure}[h]
+  \centering
+  \includegraphics[width=.45\textwidth]{figs/memguard_bandwidth}
+  \caption{Timing impact of co-scheduling memory intensive
+co-runners when memory bandwidth throttling is enabled. }
+  \label{fig:memguard_bandwidth}
+\end{figure}
+
+Figures \ref{fig:memguard_bandwidth} shows the result.
+Note that, in case of BwWrite co-runners, we have assigned half the
+bandwidth budget because MemGuard currently only account L2 refills
+but not write-backs, which effectively allow twice the allocated
+memory bandwidth budget. Regardless, the result clearly shows that
+limiting co-runners's memory bandwidth is effective in protecting the
+DNN model's performance.
+
+In summary, we find that the DNN inferencing workload is sensitive to
+memory bandwidth and that memory bandwidth throttling is effective in
+improving performance isolation of the DNN workload.
 
@@
 
-\begin{figure}[h]
-  \centering
-  \includegraphics[width=.45\textwidth]{figs/memguard_bandwidth}
-  \caption{Timing impact of co-scheduling memory intensive
-co-runners when memory bandwidth throttling is enabled. }
-  \label{fig:memguard_bandwidth}
-\end{figure}
-
 %% \begin{figure}[h]
@@
 
-We also test the performance of the model in the presence of three
-memory intensive read and write benchmarks to determine the effects of 
-limiting the memory bandwidth of the co-runners. For each of the 
-experiments, we give the model's core, Core 0, a constant 1000 MB/s 
-memory bandwidth reservation size, and then vary the amount of bandwidth 
-the co-runners' cores receive. For read co-runners, we give them an initial
-reservation size of 500 MB/s, and decrease that to a minimum of 100 MB/s.
-In the case of write co-runners, MemGuard accounts for L2 refills but not
-write-backs, so each benchmark would effectively have twice the bandwidth
-of their reservation size. Due to this, we give write co-runners an initial 
-reservation size of 250 MB/s and decrease it to a minimum of 50 MB/s, but
-their effective bandwidths are the same as the read co-runners.
-The results can be see in Figures \ref{fig:memguard_bandwidth}. 
-When less memory bandwidth is given to both the read and write 
-co-runners, the performance of the model improves. 
 %% In the case of one co-runner, the performance of the model remains 
@@
 %% still improves compared to when throttling wasn't enabled. 
-In both cases, when the co-runners were given a minimal reservation size,
-the performance of the model was much closer to its solo execution time.
-This is especially noteworthy when write co-runners were present, as the 
-model didn't suffer the same 11.6X slowdown that was seen without
-bandwidth throttling enabled. 
-%% When memory intensive Bandwidth benchmarks were present, model performance
-%% improved, to a notable extent, through the use of memory bandwidth 
-%% throttling on the co-runners. 
-Since this would not be the case if the 
-model was memory insensitive, we find that the model is memory 
-dependent in the presence of co-runners.
+%% In both cases, when the co-runners were given a minimal reservation size,
+%% the performance of the model was much closer to its solo execution time.
+%% This is especially noteworthy when write co-runners were present, as the 
+%% model didn't suffer the same 11.6X slowdown that was seen without
+%% bandwidth throttling enabled. 
+%% %% When memory intensive Bandwidth benchmarks were present, model performance
+%% %% improved, to a notable extent, through the use of memory bandwidth 
+%% %% throttling on the co-runners. 
+%% Since this would not be the case if the 
+%% model was memory insensitive, we find that the model is memory 
+%% dependent in the presence of co-runners.
 
-Based on all of the memory bandwidth throttling experiments 
-performed, it is evident that the shared DRAM memory is important for
-the DNN model. When a single model was run, its 
-performance decreased as less memory bandwidth was 
-made available to it. Furthermore, when the bandwidths of memory 
-intensive read and write co-runners were limited, the performance of 
-the model improved. As such, we conclude that the model is dependent on the 
-shared DRAM memory.
+%% Based on all of the memory bandwidth throttling experiments 
+%% performed, it is evident that the shared DRAM memory is important for
+%% the DNN model. When a single model was run, its 
+%% performance decreased as less memory bandwidth was 
+%% made available to it. Furthermore, when the bandwidths of memory 
+%% intensive read and write co-runners were limited, the performance of 
+%% the model improved. As such, we conclude that the model is dependent on the 
+%% shared DRAM memory.

### 2018-04-18 paper/evaluation.tex
revision up to 4.H

@@
 not use the GPU in our evaluation, due to the lack of sofware support
-\footnote{TensorFlow currently only support NVIDIA's GPUs.}.
+\footnote{TensorFlow currently only supports NVIDIA's GPUs.}.
 For software, we use Ubuntu MATE 16.04, TensorFlow 1.1.
@@
 space and DRAM bandwidth are two most well-known sources of contention
-in multicore. Thus, in the following, we investigate test if and how
+in multicore. Thus, in the following, we investigate if and how
 they influence the observed timing interference and how effective are
@@
 
-First, we find that DeepPicar's computing platform,
-the Raspberry Pi 3 Model B, offers computing capacity to
-perform real-time control of the RC car at 30 Hz frequency (or
-33.$\overline{\mbox{33}}$ ms  per control loop). Given the complexity of 
-the DNN used, we were pleasantly surprised by this finding. 
-The time breakdown shows that the DNN inferencing operation, 
-performed by the Tensorflow library, dominates the execution time, 
-which is expected.
+First, we find that DeepPicar's computing platform, the Raspberry Pi
+3, offers sufficient computing capacity to perform deap-learning based
+real-time control at 30 Hz frequency (or 33.3 ms per control
+loop). Given the complexity of the DNN used, we were pleasantly
+surprised by this finding. 
+%% The time breakdown shows that the DNN inferencing operation, 
+%% performed by the Tensorflow library, dominates the execution time, 
+%% which is expected.
 
@@
 core is not much worse than using multiple cores; (2) multiple DNN
-models running simultaneously do not cause severe interference with
-each other.
+models running simultaneously cause moderate interference.
 
 Fourth, we find that consolidating memory (DRAM) performance
-intensive applications could jeopadize DNN performance, because DNN
-performance appears to be very sensitive to memory performance; we 
-observe up to 11.6X slowdown in DNN performance by co-scheduling 
-synthetic memory bandwidth intensive applications on idle cores.
+intensive applications (especially write-heavy ones) can jeopadize
+the DNN inferencing task's performance even when all tasks have
+dedicated cores due to contention in shared hardware resources; we
+observe up to 11.6X slowdown.
 
-Fifth, we find that partitioning the shared L2 cache provides 
-practically no benefits to the real-time performance of the DNN. To the 
-contrary, performance is constant regardless of the amount of cache 
-space used. Furthermore, we find that performance remains consistent 
-when memory intensive read co-runners are introduced, even though
-the model has its own isolated cache space. As such, we 
-conclude that the shared L2 cache is not essential to the DNN's real-time 
-performance.
+Lastly, we find that the DNN inferencing workload is not sensitive
+to cache space and that cache partitioning is not effective in
+providing timing isolation to our DNN workload. On the other hand, we
+find that the DNN workload is sensitive to memory bandwidth and that
+memory bandwidth throttling is effective in improving performance
+isolation.
 
-Lastly, we find that memory bandwidth throttling results in significant
-changes to model performance. When less bandwidth is given to the model,
-we find that it performs worse. When memory intensive co-runners are
-present, we find that limiting the available bandwidth of the co-runners
-results in significantly improved model execution times. As such, we conclude
-that the shared DRAM memory is essential to the DNN's real-time 
-performance.
+%% To the 
+%% contrary, performance is constant regardless of the amount of cache 
+%% space used. Furthermore, we find that performance remains consistent 
+%% when memory intensive read co-runners are introduced, even though
+%% the model has its own isolated cache space. As such, we 
+%% conclude that the shared L2 cache is not essential to the DNN's real-time 
+%% performance.
+
+%% Lastly, we find that memory bandwidth throttling results in significant
+%% changes to model performance. When less bandwidth is given to the model,
+%% we find that it performs worse. When memory intensive co-runners are
+%% present, we find that limiting the available bandwidth of the co-runners
+%% results in significantly improved model execution times. As such, we conclude
+%% that the shared DRAM memory is essential to the DNN's real-time 
+%% performance.

### 2018-04-18 paper/intro.tex
Add GitHub link

@@
     cars.~\footnote{The source code, dataset, and build instructions
-      of DeepPicar can be found at:\fixme{Michael github url}}
+      of DeepPicar can be found at: \url{https://github.com/mbechtel2/DeepPicar-v2}}
 

### 2018-04-18 paper/overview.tex
Add link to videos

@@
 throttle)~\footnote{A collection of self-driving videos can be found
-  at: \fixme{Michael: link your video}
+  at: \url{https://photos.app.goo.gl/q40QFieD5iI9yXU42}
   %% \url{https://photos.app.goo.gl/ce93sU7jPk4ywO8u2}

### 2018-04-18 paper/comparison.tex
revision 4.E (2nd)
comparsion and conclusion revision

--- 2018-04-04 paper/comparison.tex
@@
 \section{Embedded Computing Platform Comparison}\label{sec:comparison}
+
+%% \begin{figure}[h]
+%%   \centering
+%%   \includegraphics[width=.45\textwidth]{figs/a53_vs_a7}
+%%   \caption{DNN model performance on different in-order
+%% ARM Cortex cores. \fixme{this may be better in the plaform comparison section}.}
+%%   \label{fig:a53_vs_a7}
+%% \end{figure}
 
@@
 respectively, than the Pi 3. 
-As a result, they are all able to satisfy the 33.$\overline{\mbox{33}}$ ms 
+As a result, they are all able to satisfy the 33.3 ms 
 WCET by a clear margin,
@@
 
-The Intel UP board and Jetson TX2 also perform much better when multiple DNN models
-are co-scheduled. Figure~\ref{fig:sys_model} shows the results of the
+Figure~\ref{fig:sys_model} shows the results of the
 multi-model co-scheduling experiment. Once again, they can comfortably
@@
   \centering
-  \includegraphics[width=.45\textwidth]{figs/compare_benchmark}
-  \caption{Average control loop execution time in the presence of an
-    increasing number of memory intensive applications on idle CPU cores.}
+  \begin{subfigure}{0.4\textwidth}
+    \includegraphics[width=\textwidth]{figs/compare_benchmark_read}
+    \caption{BwRead}
+    \label{fig:sys_bench_read}
+  \end{subfigure}
+  \hfill
+  \begin{subfigure}{0.4\textwidth}
+    \includegraphics[width=\textwidth]{figs/compare_benchmark}
+    \caption{BwWrite}
+    \label{fig:sys_bench_write}    
+  \end{subfigure}
+  \caption{Average processing time vs. the number of memory
+    memory intensive co-runners.}
   \label{fig:sys_bench}    
@@
 
-\begin{figure}[h]
-  \centering
-  \includegraphics[width=.45\textwidth]{figs/compare_l2missrate}
-  \caption{L2 miss rates in the presence of an increasing number of 
-			memory intensive applications on idle CPU cores.}
-  \label{fig:sys_l2miss}
-\end{figure}
+Figure~\ref{fig:sys_bench} shows the results of synthetic memory
+intensive task co-scheduling experiment. 
+For read co-runners (BwRead), the performance of all platforms
+gradually decreased as additional BwRead instances are introduced: up
+to 1.6X for Pi 3, up to 1.3X for Intel UP, and up to 1.6X and 2.2X for the TX2
+(CPU) and TX2(GPU), respectively.
+For write co-runners (BwWrite), however, we observe generally more
+noticieable execution time increases. In particular, as we discussed
+earlier in Section~\ref{sec:evaluation}, the Pi 3 suffers up to 11.6X
+execution time increase. We also observe increased control loop
+execution timing in the Intel Up and Jetson TX2, although the degree
+of the increase is not as dramatic as the Pi 3.
 
-Finally, we compare the effect of co-scheduling memory bandwidth
-intensive synthetic benchmarks on the DNN control loop timing. 
-Figure~\ref{fig:sys_bench} shows the results. As discussed in 
-Section~\ref{sec:eval-memhog}, we observed dramatic execution time
-increases, up to 11.6X, in Raspberry Pi 3 as we increased the number of
-co-scheduled tasks. We also observe increased control loop execution
-timing in the Intel Up and Jetson TX2, but the degree of the increase is 
-not as dramatic as the Pi 3. Compared to their respective solo timings 
-(i.e., the model runs on a single core in isolation), Intel UP suffers up to
-2.3X execution time increase; TX2 (CPU) and TX2 (GPU) suffer up to
-2.2X and 2.1X increases, respectively. This is somewhat surprising
-because the Raspberry Pi 3's cores are in-order architecture based while
-the cores in the Intel Up and NVIDIA TX2 are out-of-order architecture
-based, and that the memory intensive tasks on out-of-order cores can
-generate more memory traffic. We believe that this is because the
-memory subsystems in the Intel UP and TX2 platforms provide higher
-performance than the memory subsystem of the Pi 3 as suggested from
-the measured memory bandwidth results in Table~\ref{tbl:platforms}
-('Measured Bandwidth').
+%% Compared to their respective solo timings  
+%% (i.e., the model runs on a single core in isolation), Intel UP suffers up to
+%% 2.3X execution time increase; TX2 (CPU) and TX2 (GPU) both suffer up to
+%% 2.3X increases. This is somewhat surprising
+%% because the Raspberry Pi 3's cores are in-order architecture based while
+%% the cores in the Intel Up and NVIDIA TX2 are out-of-order architecture
+%% based, and that the memory intensive tasks on out-of-order cores can
+%% generate more memory traffic.
+%% We believe that this is because the
+%% memory subsystems in the Intel UP and TX2 platforms provide higher
+%% performance than the memory subsystem of the Pi 3 as suggested from
+%% the measured memory bandwidth results in Table~\ref{tbl:platforms}
+%% ('Measured Bandwidth').
 
 Another interesting observation is that the TX2 (GPU) also suffers
-considerable execution time increase (2.1X) despite the fact that the
+considerable execution time increase (2.3X) despite the fact that the
 co-scheduled synthetic tasks do not utilize the GPU. In other words,
@@
 
-Figure~\ref{fig:sys_l2miss} shows the L2 miss rates of DNN control
-loop. The key takeaway of this result is that the L2 miss rate is a
-poor indicator to the execution time of the control loop. As shown
-already, the Raspberry Pi 3's increased L2 misses do not fully explain
-increases in processing times. In case of the Intel UP, as the number of
-interfering Bandwidth benchmark instances increases, the L2 miss rates
-drop slightly. In the case of the TX2 (CPU), from solo to one co-runner,
-L2 miss rates almost double, but then the execution time
-increase is a little less than 10\%. However, when there are
-co-runners, even though the L2 miss rates are largely unchanged, the
-DNN processing time increases significantly. The TX2 (GPU) also shows
-similar behavior. All of these results essentially point to the fact
-that DNN inferencing workload is largely memory performance sensitive
-but not L2 cache space sensitive.
+%% \begin{figure}[h]
+%%   \centering
+%%   \includegraphics[width=.45\textwidth]{figs/compare_readl2}
+%%   \caption{L2 miss rates in the presence of an increasing number of 
+%% 			memory intensive read applications on idle CPU cores.}
+%%   \label{fig:sys_readl2miss}
+%% \end{figure}
 
-%% However, the L2 miss rates of the Pi 3 and TX2 platforms follow the same expected 
-%% pattern, as can be seen in Figure~\ref{fig:sys_l2miss}. Namely, the 
-%% percentage of L2 miss rates increases are more 
-%% memory intensive co-runners are introduced. This was not the case for 
-%% the UP board, where the model running on two cores generated the highest
-%% percentage of L2 cache misses. This is most likely due to the CPU 
-%% architecture of the UP's CPU which is Intel based, whereas both the Pi 3
-%% and TX2 have ARM based CPUs.
+%% \begin{figure}[h]
+%%   \centering
+%%   \includegraphics[width=.45\textwidth]{figs/compare_l2missrate}
+%%   \caption{L2 miss rates in the presence of an increasing number of 
+%% 			memory intensive write applications on idle CPU cores.}
+%%   \label{fig:sys_l2miss}
+%% \end{figure}
+
+%% Figures~\ref{fig:sys_readl2miss} and ~\ref{fig:sys_l2miss} show
+%% the L2 miss rates of DNN control loops when read and write co-runners
+%% are introduced, respectively. The key takeaway of this result is that the 
+%% L2 miss rate is a
+%% poor indicator to the execution time of the control loop. As shown
+%% already, the Raspberry Pi 3's increased L2 misses do not fully explain
+%% increases in processing times. In case of the Intel UP, as the number of
+%% interfering Bandwidth benchmark instances increases, the L2 miss rates
+%% drop slightly. In the case of the TX2 (CPU), from solo to one co-runner,
+%% L2 miss rates almost double, but then the execution time
+%% increase is a little less than 10\%. However, when there are
+%% co-runners, even though the L2 miss rates are largely unchanged, the
+%% CNN processing time increases significantly. The TX2 (GPU) also shows
+%% similar behavior. All of these results essentially point to the fact
+%% that DNN inferencing workload is largely memory performance sensitive
+%% but not L2 cache space sensitive.
+
+%% \begin{figure}[h]
+%%   \centering
+%%   \includegraphics[width=.45\textwidth]{figs/compare_readbus}
+%%   \caption{Bus accesses in the presence of an increasing number of 
+%% 			memory intensive read applications on idle CPU cores.}
+%%   \label{fig:sys_readbus}
+%% \end{figure}
+
+%% \begin{figure}[h]
+%%   \centering
+%%   \includegraphics[width=.45\textwidth]{figs/compare_writebus}
+%%   \caption{Bus accesses in the presence of an increasing number of 
+%% 			memory intensive write applications on idle CPU cores.}
+%%   \label{fig:sys_writebus}
+%% \end{figure}
+
+%% \color{red}FIXME: I believe the bus access results for the Intel UP are wrong. 
+%% I couldn't get the IDI off-core response event to work, so I used the bus-cycles 
+%% perf event, which I think is incorrect.
+
+%% \color{black}Figures~\ref{fig:sys_readbus} and ~\ref{fig:sys_writebus} show the
+%% normalized bus accesses for each of the platforms as read and write
+%% co-runners are introduced, respectively. Unlike the L2 miss rates of the
+%% model, the total number of bus accesses more closely align with the 
+%% execution time increases that are seen in each of the platforms. For the Pi 3, both 
+%% the execution time and bus accesses increase at exponential rates. In the TX2(GPU), 
+%% the bus accesses increase in a similar linear fashion to the execution time. The main outlier
+%% is the TX2(CPU) which has bus accesses increase at a faster rate similar to that of the Pi 3.
+%% However, the bus accesses seem to increase linearly, which is also seen in the
+%% execution time. As a result, we find that the total number of bus accesses have a higher 
+%% correlation to the model execution time than L2 miss rates.
+
+%% %% However, the L2 miss rates of the Pi 3 and TX2 platforms follow the same expected 
+%% %% pattern, as can be seen in Figure~\ref{fig:sys_l2miss}. Namely, the 
+%% %% percentage of L2 miss rates increases are more 
+%% %% memory intensive co-runners are introduced. This was not the case for 
+%% %% the UP board, where the model running on two cores generated the highest
+%% %% percentage of L2 cache misses. This is most likely due to the CPU 
+%% %% architecture of the UP's CPU which is Intel based, whereas both the Pi 3
+%% %% and TX2 have ARM based CPUs.
 

### 2018-04-18 paper/conclusion.tex
comparsion and conclusion revision

@@
 
-As future work, 
+%% As future work, we plan to improve the prediction accuracy of the
+%% system by feeding more data in training.
+As future work, we plan to investigate ways to reduce computational
+and memory overhead of DNN inferencing and to evaluate the
+effectiveness of FPGA and other specialized accelerators.
+
 %% we plan to apply shared DRAM resource management
@@
 %%evaluate their impact on the real-time performance of the system.
-we plan to improve the prediction accuracy by feeding more data and
-upgrading the RC car hardware platform to enable more precise steering
-angle control.
+
+%% and
+%% upgrading the RC car hardware platform to enable more precise steering
+%% angle control

### 2018-04-18 paper/evaluation-cachepart.tex
remove fixme. replace track 1 picture.

@@
   \includegraphics[width=.45\textwidth]{figs/palloc_multicore}
-  \caption{Cache space sensitivity of the DNN controller. \fixme{I
-      think 'no part.' bar is not necessary here.}}
+  \caption{Cache space sensitivity of the DNN controller.}
   \label{fig:palloc_multicore}
@@
   \caption{Average processing time vs. the number of memory
-memory intensive co-runners. Unlike
-Figure~\ref{fig:perf_vs_bandwidth}, the shared L2 cache is equally
-partitioned and each task is given a dedicated cache
-partition. \fixme{how about changing the figure style to be the same
-  as Figure 9.}}
+memory intensive co-runners; Each core (task) is given a equal-sized
+dedicated cache partition.}
   \label{fig:palloc_bandwidth_exectime}

### 2018-04-19 paper/related.tex
related work revision

@@
 
-In this paper, we demonstrated that a low-cost embedded platform such
-as the  Raspberry Pi 3 can successfully run a CNN-based autonomous
-driving control system in real-time. However, it should be noted that
-the CNN we used here is relatively small compared to recent
-state-of-the-art CNNs, which are increasingly larger and deeper.
-For example, the CNN based object detector models evaluated in
+There are several relatively inexpensive RC-car based autonomous car
+testbeds. MIT's RaceCar~\cite{shin2017project} and UPenn's
+F1$/$10~\cite{upennf1tenth} are both based on a Trexx 1/10 scale RC
+car and a powerful NVIDIA Jetson computing platform equipped with many
+sophisticated sensor packages such as a lidar. However, they both cost
+more than \$3,000, requiring considerable
+investment. DonkeyCar~\cite{donkeycar} is similar to our DeepPicar as
+it also uses a Raspberry Pi 3 and a CNN for end-to-end control,
+although it costs slightly more because it uses a more expensive 1/16
+scale RC car. The main contribution of our paper is not the hardware
+design but in the detailed analysis of computational characteristics
+for the CNN-based end-to-end real-time control.
+
+In this paper, we have analyzed real-time performance of a
+state-of-the art CNN, used in NVIDIA's DAVE-2 self-driving car, on a
+low-cost raspberry pi 3 quad-core platform. It should be noted,
+however, that DAVE-2's CNN we used here is relatively small compared
+to recent state-of-the-art CNNs, which are increasingly larger and
+deeper. For example, the CNN based object detector models evaluated in
 Google's recent study~\cite{huang2017speed} have between 3M to 54M
@@
 
-%% Another autonomous vehicle platform is the Donkey Car\footnote{https://donkeycar.com}.
-%% Like the DeepPicar, the Donkey Car is a low cost platform that can be 
-%% utilized for researching self-driving cars. However, comparatively, the
-%% DeepPicar is approximately \$200 cheaper than the Donkey Car. 
-%% Furthermore, the Donkey Car was created with an emphasis on model accuracy
 %% and the ability of the car to navigate the desired environment. We are 

### 2018-04-19 paper/related.tex
..

@@
 
-In this paper, we have analyzed real-time performance of a
-state-of-the art CNN, used in NVIDIA's DAVE-2 self-driving car, on a
+In this paper, we have analyzed real-time performance of a CNN, which
+was used in NVIDIA's DAVE-2 self-driving car~\cite{Bojarski2016}, on a 
 low-cost raspberry pi 3 quad-core platform. It should be noted,
-however, that DAVE-2's CNN we used here is relatively small compared
+however, that DAVE-2's CNN is relatively small compared
 to recent state-of-the-art CNNs, which are increasingly larger and
@@
 Google's recent study~\cite{huang2017speed} have between 3M to 54M
-parameters. Using such large CNN models will be challenging on
+parameters, much larger than DAVE-2's CNN.
+Using such large CNN models will be challenging on
 resource constrainted embedded computing platforms, especially for

### 2018-04-19 paper/background.tex
revision

@@
 interference can nullify the benefits of consolidation~\cite{Kim2016}.
-For these reasons, finding a good computing platform is a
-non-trivial task, one that requires a deep understanding of the
-workloads and the hardware platform being utilized.
+%% For these reasons, finding a good computing platform is a
+%% non-trivial task, one that requires a deep understanding of the
+%% workloads and the hardware platform being utilized.
 
-The \emph{primary objectives of this study} are to understand (1) the
-necessary computing performance for applying AI technology-based
-robotics systems, and (2) the characteristics of the computing
-platform to support such a workload.
+The \emph{primary objectives of this study} are (1) to understand the
+necessary computing performance to realize deep neural network based
+robotics systems, (2) to understand the characteristics of the
+computing platform to support such workloads, and (3) to evaluate the
+significance of contention in shared hardware resources and existing
+mitigation techniques to address the contention problem.
+
 To achieve these goals, we implement a low-cost autonomous car platform

### 2018-04-19 paper/evaluation-cachepart.tex
revision

@@
 Cache partitioning is a well-known technique to improve isolation in
-real-time systems by giving a dedicated cache space to each individaul
+multicore by giving a dedicated cache space to each individaul 
 task or core. In this experiment, we use PALLOC~\cite{yun2014rtas}, a
-page-coloring based cache partitioning solution for Linux, to
-investigate the effect of cache partitioning in protecting DeepPicar's
-DNN based controller.
+page-coloring based kernel-level memory allocator for Linux.
+Page coloring is a OS technique that controls the physical addresses
+of memory pages. By allocating pages over non-overlapping sets of the
+cache, the OS can effectively partition the cache.
+Using PALLOC, we investigate the effect of cache partitioning in
+protecting DeepPicar's DNN based controller.
 
-\begin{figure}
+\begin{figure} [h]
   \centering
@@
 
-Page coloring is a OS technique that controls the physical addresses
-of memory pages. By allocating pages over non-overlapping sets of the
-cache, the OS can effectively partition the cache.
 Figure~\ref{fig:cache-mapping} shows the physical address
@@
 
+In the first experiment, we investigate the cache space sensitivity of
+DeepPicar's DNN-based control loop. Using PALLOC, we create 4
+different cgroups which are configured to use 4, 3,
+2, and 1 colors (100\%, 75\%, 50\% and 25\% of the L2 cache
+space, respectively). We then execute the DNN control loop (inference)
+on one core using a different cgroup cache partition, one at a time,
+and measure the average processing time.
+
 \begin{figure}[h]
@@
 \end{figure}
-
-In the first experiment, we investigate the cache space sensitivity of
-DeepPicar's DNN-based control loop. Using PALLOC, we create 4
-different cgroups which are configured to use 4, 3,
-2, and 1 colors (100\%, 75\%, 50\% and 25\% of the L2 cache
-space, respectively). We then execute the DNN control loop (inference)
-on each cache partition (cgroup) and measure the average processing
-time.
 

### 2018-04-19 paper/evaluation.tex
..
revision

@@
 \subsection{Setup}
-
 The Raspberry Pi 3 Model B platform used in DeepPicar equips a Broadcom
 BCM2837 SoC, which has a quad-core ARM Cortex-A53 cluster,
-running at up to 1.2GHz. Each core has a 32K private Instruction cache
-and a 32K private Data cache, and all cores share a 512KB L2 cache.
+running at up to 1.2GHz. Each core has private 32K I/D caches
+and all cores share a 512KB L2 cache.
 The chip also includes Broadcom's Videocore IV GPU, although we did
@@
 configured the clock speed of each core statically at the maximum 1.2GHz.
+We use the SCHED\_FIFO real-time scheduler to schedule the DNN control
+task while using the CFS when executing memory intensive co-runners.
 
 \subsection{Inference Timing for Real-Time Control}
-
 For real-time control of a car (or any robot), the control loop
@@
 This means that the DNN inference operation must be completed
-within the inner-most control loop update frequency. To understand
-achievable control-loop update frequencies, we experimentally measured
-the execution times of DeepPicar's DNN inference operations.
+within the inner-most control loop update frequency.
+
+To understand achievable control-loop update frequencies, we
+experimentally measured the execution times of DeepPicar's DNN
+inference operations.
 
@@
   \centering
-  \includegraphics[width=.5\textwidth]{figs/Fig7_new}
+  \includegraphics[width=.45\textwidth]{figs/Fig7_new}
   \caption{DeepPicar's control loop processing times over 1000 input image frames.}
@@
 used by the library. Leveraging this feature, we repeat the
-same experiment described in the previous subsection with varying
+experiment in the previous subsection but with using varying
 numbers of CPU cores---from one to four.
@@
 improvement is mere 2.80 ms, or 12\%. In short, we find that the
-scalability of DeepPicar's deep neural network is not ideal on the
-platform.
+scalability of DeepPicar's deep neural network is not ideal.
 
@@
 improve inference performance~\cite{NVIDIA2015}, which generally does
-not scale very well.
+not scale well.
 
-On the other hand, the poor scalability opens up the possibility of
+On the other hand, the limited scalability opens up the possibility of
 consolidating multiple different tasks or different neural network
@@
 These increases in inference times in the co-scheduled scenarios are
-expected because it is well-known that co-scheduled tasks on a
+expected because co-scheduled tasks on a
 multicore platform interfere with each other due to contention in the
 shared hardware resources, such as the shared
-cache~\cite{Gracioli2015} and DRAM~\cite{Yun2013}.
+cache and DRAM~\cite{Gracioli2015,Yun2013}.
 
@@
 
-Figure~\ref{fig:perf_vs_bandwidth} shows the execution times of the DNN 
-model against varying number of BwRead and BwWrite co-runners. Note
-that BwWrite cause significantly high execution time incrase---up 
-to 11.6X---on the DNN inferencing timing while BwRead causes
-relatively much smaller time increase. While execution time increase
-was expected, the degree of which is quite surprising---\emph{if the DNN
-controller was driving an actual car, it would result in a crash of the
-car!}
+Figure~\ref{fig:perf_vs_bandwidth} shows the result. Note
+that BwWrite co-runners cause significantly high execution time
+incrases---up to 11.6X---on DNN inferencing while BwRead co-runners
+cause relatively much smaller time increases. While execution time
+increases are expected, the degree of which in the worst-case is quite
+surprising---\emph{if the DNN controller was driving an actual car, it
+  would result in a crash of the car!}
 
@@
 they influence the observed timing interference and how effective are
-the existing mitigation methods. To our surprise, we find that neither
-the cache space nor DRAM bandwidth is the root cause of the
-DeepPicar's observed worst-case execution time increase.
+the existing mitigation methods.
+
+%% To our surprise, we find that neither
+%% the cache space nor DRAM bandwidth is the root cause of the
+%% DeepPicar's observed worst-case execution time increase.
 

### 2018-04-19 paper/intro.tex
revision

--- 2018-04-16 paper/intro.tex
@@
 computing hardware architectures.
+From the systematic study, we want to answer the following questions:
+(1) how well does the DNN based real-time inferencing task perform on
+contemporary embedded multicore computing platforms? (2) how
+susceptible is the DNN inferencing task to contention in shared
+hardware resources (e.g., cache and DRAM) when multiple tasks/models
+are consolidated? (3) are the existing state-of-the-art shared resource
+isolation techniques effective in protecting real-time performance of
+the DNN inferencing task?
 
@@
 %   isolation of DNN workload when memory bandwidth is contended.
-From the systematic study, we have made a number of interesting
-observations.
-% predictability
-First, we find that CNN inferencing is highly time predictable as the
-amount of computation needed to complete a 
-single inference is fixed at the network architecture design time and
-does not change at runtime over different inputs (e.g., different image
-frames). This predictable timing behavior is obviously a desirable
-property for real-time systems.
-%% , which gives another reason for the
-%% real-time community to be interested in the use of DNNs.
+Our observations are as follows.
+First, we find that real-time processing of the DNN inferencing is
+feasible on contemporary embedded computing platforms, even one   
+as inexpensive as the Raspberry Pi 3. Second, while consolidating
+multiple DNN models and task are feasible but the impact of shared
+resource contention can be catastrophic in the worst-case as we
+observe up to 11.6X slowdown even when the DNN model was given a
+dedicated core. Third, using the state-of-the-art OS-level cache
+partitioning and memory bandwidth control
+mechanisms~\cite{yun2014rtas,Yun2013}, we find that cache space has a
+minimal performance impact to the DNN workload while memory
+performance has a significant impact.
 
+%% % predictability
+%% First, we find that CNN inferencing is highly time predictable as the
+%% amount of computation needed to complete a 
+%% single inference is fixed at the network architecture design time and
+%% does not change at runtime over different inputs (e.g., different image
+%% frames). This predictable timing behavior is obviously a desirable
+%% property for real-time systems.
+%% %% , which gives another reason for the
+%% %% real-time community to be interested in the use of DNNs.
 % performance on low-cost embedded multicore
-Second, we find that real-time processing of CNN inferencing is
-feasible on today's embedded computing platforms, even one   
-as inexpensive as the Raspberry Pi 3. For example, just one Cortex-A53
-core of the Raspberry Pi3 is able to complete the CNN inferencing task
-in less than 50ms, sufficient to support 20 Hz control loop. 
+%% For example, just one Cortex-A53
+%% core of the Raspberry Pi3 is able to complete the CNN inferencing task
+%% in less than 50ms, sufficient to support 20 Hz control loop. 
 %% We find that the control loop 
@@
 %% embedded GPU is used.
-We find that all the tested embedded computing platforms have 
-enough computing capacity to consolidate multiple instances of
-DeepPicar's DNN workload.
+%% We find that all the tested embedded computing platforms have 
+%% enough computing capacity to consolidate multiple instances of
+%% DeepPicar's DNN workload.
 
 % all platforms have enough capacity to consolidate
-Third, contention in shared hardware resources remains an important
-issue when consolidating multiple workloads. 
-We observe up to 11.6X control
-loop execution time increase due to shared resource contention when
-memory intensive  applications are co-scheduled on idle CPU cores.
-We find that DeepPicar's inferencing task is highly
-sensitive to memory bandwidth, but not cache space.
+%% Third, contention in shared hardware resources remains an important
+%% issue when consolidating multiple workloads. 
+%% We observe up to 11.6X control
+%% loop execution time increase due to shared resource contention when
+%% memory intensive  applications are co-scheduled on idle CPU cores.
+%% We find that DeepPicar's inferencing task is highly
+%% sensitive to memory bandwidth, but not cache space.
 
@@
 % contributions
-The {\bf contributions} of this paper are as follows:
+This paper makes the following {\bf contributions}:
 \begin{itemize}
-  \item We present DeepPicar, a low-cost autonomous car testbed using
-    a state-of-the-art CNN for end-to-end real-time control, to lower
-    economic and safety-related barriers to the study autonomous
-    cars.~\footnote{The source code, dataset, and build instructions
-      of DeepPicar can be found at:\fixme{Michael github url}}
-
-  \item We provide extensive systematic experimental results using
-    DeepPicar's CNN inferencing workload on a number of representative
-    contemporary embedded computing platforms. The results provide
-    insights for real-time processing of the CNN workload and the
-    effect of consolidation.
-
-    %% While larger embedded
-    %% platforms have been evaluated, as in  \cite{Otterness2017}, to the
-    %% best of our knowledge, no prior work investigated the
-    %% impacts of shared resource contention on DNN inferencing workloads
-    %% in the context of smaller embedded platforms like the Raspberry Pi 3.
+  \item We present DeepPicar, a low-cost autonomous car testbed, which
+    employs a state-of-the-art CNN based end-to-end
+    real-time control.~\footnote{The source code, datasets, and build
+      instructions of DeepPicar can be found at:
+      \url{https://github.com/mbechtel2/DeepPicar-v2}}.
+  \item We provide extensive empirical performance evaluation results
+    of the CNN inferencing workload on a number of contemporary
+    embedded computing platforms.
+  \item We apply the state-of-the-art shared resource isolation
+    techniques and evaluate their effectiveness in protecting the
+    inferencing workload in consolidated settings.
 \end{itemize}

### 2018-04-19 paper/overview.tex
remove fixme. replace track 1 picture.
revision

--- 2018-04-16 paper/overview.tex
@@
 \section{DeepPicar Overview}\label{sec:overview}
+
+In this section, we provide an overview of our DeepPicar platform.
+In developing DeepPicar, one of our primary goals is to faithfully
+replicate NVIDIA's DAVE-2 system on a smaller scale using a low cost
+multicore platform, the Raspberry Pi 3. Because Raspberry Pi 3's
+computing performance is much lower than that of the DRIVE
+PX~\cite{drivepx} platform used in DAVE-2, we are interested in if,
+and how, we can process 
+computationally expensive neural network operations in
+real-time. Specifically, inferencing (forward pass processing)
+operations must be completed within each control period
+duration---e.g., a WCET of 33.3 ms for 30Hz control 
+frequency---locally on the Pi 3 platform, although training of the 
+network (back-propagation for weight updates) can be done offline and 
+remotely using a desktop computer.
 
@@
 
-In this section, we provide an overview of our DeepPicar platform.
-In developing DeepPicar, one of our primary goals is to faithfully
-replicate NVIDIA's DAVE-2 system on a smaller scale using a low cost
-multicore platform, the Raspberry Pi 3. Because the Raspberry Pi 3's
-computing performance is much lower than that of the DRIVE PX platform
-used in DAVE-2, we are interested in if, and how, we can process
-computationally expensive neural network operations in
-real-time. Specifically, inferencing (forward pass processing)
-operations must be completed within each control period
-duration---e.g., a WCET of 33.$\overline{\mbox{33}}$ ms for 30Hz control 
-frequency---locally on the Pi 3 platform, although training of the 
-network (back-propagation for weight updates) can be done offline and 
-remotely using a desktop computer.
 Figure~\ref{fig:overview} shows the DeepPicar, which is comprised of a
@@
 estimated cost of the system.
+
+For the neural network architecture, we implement NVIDIA DAVE-2's
+convolutional neural network (CNN) using an open-source CNN model in
+~\cite{deeptesla}. Note, however, that the CNN model
+in~\cite{deeptesla} is considerably larger than NVIDIA's CNN
+model as it contains an additional fully-connected layer of
+approximately 1.3M additional parameters. We remove the additional
+layer to faithfully recreate NVIDIA's original CNN model.
+%% The main difference is that we do not utilize a normalization layer, and 
+%% instead initialize the weights using a Xavier initialization. 
+As in DAVE-2, the CNN takes a raw color image (200x66 RGB pixels)
+as input and produces a single steering angle value as output.
+Figure~\ref{fig:architecture} shows the network architecture
+used in this paper, which is comprised of 9 layers, 250K parameters,
+and about 27 million connections as in NVIDIA DAVE-2's architecture.
 
@@
 
-For the neural network architecture, we implement
-NVIDIA DAVE-2's convolutional neural network (CNN) using a TensorFlow
-based CNN model in ~\cite{deeptesla}. Note, however, that the CNN model
-in~\cite{deeptesla} was considerably larger than NVIDIA's CNN model as
-it contained an additional fully-connected layer of approximately 1.3M
-additional parameters. We eliminated the additional layer to
-faithfully recreate the NVIDIA's original CNN model.
-%% The main difference is that we do not utilize a normalization layer, and 
-%% instead initialize the weights using a Xavier initialization. 
-As in DAVE-2, the CNN takes a raw color image (200x66 RGB pixels)
-as input and produces a single steering angle value as
-output. Figure~\ref{fig:architecture} shows the network architecture
-used in this paper, which is comprised of 9 layers, 250K parameters,
-and about 27 million connections as in NVIDIA DAVE-2's.
 %% Note, however,
@@
   \begin{subfigure}{0.4\textwidth}
-    \includegraphics[width=\textwidth]{figs/track_new2}
+    \includegraphics[width=\textwidth]{figs/track-2}
     \caption{Track 1.}
@@
 is copied back to the Raspberry Pi 3. The network is then used
-by the car's main controlller, which feeds an image frame from the web
-camera as input to the network at each control period. The network
-produced steering angle output is then converted into the PWM values
+by the car's main controlller, which feeds image frames from the web
+camera as input to the network in real-time. At each control period,
+the network produced steering angle output is converted into the PWM values
 of the car's steering motor. Figure~\ref{fig:controlloop} shows simplified 
@@
 Note that although the steering angle output of the network ($angle$) is
-a continuous real value, the RC car platform we used unfortunately
-only supports three discrete angles---left (-30$^{\circ}$), center
+a continuous real value, the RC car platform we used only supports
+three discrete angles---left (-30$^{\circ}$), center 
 (0$^{\circ}$), and right (+30$^{\circ}$)---as control inputs.
-Currently, we approximate the network generated real-valued
-angle to the closest one of the three angles, which may
-introduce inaccuracy in control.
+We approximate the network generated real-valued angle to the closest
+one of the three angles. Although this may introduce inaccuracy in
+control, the observed control performance of the system is respectable,
+likely due to the inherently stocastic nature of DNN.
+
 %% In the future, we plan to use a different (more expensive) RC car
@@
 
-Other issues we observed in training/testing the network, which
-can affect the prediction accuracy of the CNN, are camera and actuator
-(motor) control latencies. In DeepPicar's
-context, the camera latency is from the time the camera
-sensor observes the scene to the time the computer actually reads the
-digitized image data. This time can be noticable depending on the camera 
-used and the performance of the Pi. Higher camera latency could
+Another factors that can potentially affect the prediction accuracy of
+the CNN, are camera and actuator (motor) control latencies. The camera
+latency is from the time the camera sensor observes the scene to the
+time the computer actually reads the digitized image data. This time
+can be noticable depending on the camera used and the data processing
+time of the computing platform. Higher camera latency could
 negatively affect control performance, because the DNN would analyze
-stale scenes. Actuator (motor) control latency also can be a
-factor, because moving the steering motor to a desired position takes
-time. 
-In our platform, the combined latency is measured to be around
-50 ms. In other words, in the recorded video and control data, a
-control action (left, center, right) taken by the CNN appears to be
-applied 50 ms later. Considering that camera's framerate is 30Hz
-(33.3 ms/frame), this is about two frames after the control action.
+stale scenes. The actuator (motor) control latency from the time
+the control output is sent to the steering motor to the time the motor
+actually moved at a desired position, which also can takes
+considerable time. In our platform, the combined latency is measured
+to be around 50 ms, which is reasonble.
+%% In other words, %% in the recorded video and control data, 
+%% a control action of the CNN appears to be applied 50 ms later.
+%% Considering that camera's framerate is 30Hz
+%% (33.3 ms/frame), this is about two frames after the control action.
 %% We experimentally measured the camera
@@
 If this value is too high, control performance may suffer.
-Our initial prototype suffered this problem as the observed camera and
-control latency was as high as 300 ms, which negatively affected
-control performance.
+Our initial prototype suffered this problem as the observed latency
+was as high as 300 ms, which negatively affected control performance.
 For reference, the latency of human perception is known to be as fast
@@
 
-Despite these issues, the trained models were still able to
-achieve a reasonable degree of accuracy, successfully navigating
-several different tracks we used for training. In our testing, we found 
-that the DeepPicar could remain on a fairly challenging track for over 
-10 minutes at a moderate speed (50\% throttle), at which point we stopped
-the experiment, and more than one minute at a higher speed (75\%
-throttle)~\footnote{A collection of self-driving videos can be found
-  at: \fixme{Michael: link your video}
+Our trained DNN models showed good prediction accuracy, successfully
+navigating several different tracks we used for training.
+For instance, the DeepPicar could remain on Track 1
+(Figure \label{fig:track}) over 10 minutes at a moderate speed (50\%
+throttle), at which point we stopped the experiment, and more than one
+minute at a higher speed (75\% throttle)~\footnote{Self-driving videos: \url{https://photos.app.goo.gl/q40QFieD5iI9yXU42}
   %% \url{https://photos.app.goo.gl/ce93sU7jPk4ywO8u2}

### 2018-04-19 paper/appendix.tex
revision
Updates to paper
revision 2

--- 2017-10-16 paper/appendix.tex
@@
-\begin{figure*}[t]
+\begin{figure*}[h]
   \centering
@@
 In the first method, we trained the neural network model across a set 
-of 30 completed runs on the track seen in Figure~\ref{fig:track} by a
+of 30 completed runs on the track seen in Figure~\ref{fig:track2} by a
 human pilot. Half of the runs saw the car driving one way along the

### 2018-04-19 paper/comparison.tex
revision 2

@@
 timing of the four system configurations we tested as a function of
-the number of CPU cores used.
-First, both the Intel UP and Jetson TX2 exhibit superior performance when
-compared with the Raspberry Pi 3. 
-When all four CPU cores are used, the Intel UP is 1.33X faster than
-the Pi 3, while the TX2 (CPU) and TX2 (GPU) are 2.79X and 4.16X times faster,
-respectively, than the Pi 3. 
-As a result, they are all able to satisfy the 33.3 ms 
-WCET by a clear margin,
-%%  (in the worst case, the UP Board would 
-%% still meet it by $\sim$20 ms) --> where's the data showing this?
-and, in the case of the TX2, 50 Hz or even 100 Hz real-time control is
-feasible with the help of its GPU. Another observation is that the TX2
-(GPU) performance does not change much, as most of the neural network 
-computation is done by the GPU.
+the number of CPU cores used. (cf. Figure~\ref{fig:perf-vs-corecnt})
+Both the Intel UP and Jetson TX2 exhibit better performance than
+Raspberry Pi 3. When all four CPU cores are used, Intel UP is
+1.33X faster than Pi 3, while TX2 (CPU) and TX2 (GPU) are 2.79X and
+4.16X times faster, respectively, than Pi 3.
+Thus, they all satisfy 33.3 ms WCET by a clear margin, and, in the
+case of the TX2, 50 Hz or even 100 Hz real-time control is feasible
+with the help of its GPU.
+Another observation is that the DNN task's performance on TX2 (GPU)
+does not change much as we increase the number of cores. This
+isbecause most of the neural network computation is done by the GPU.
 
@@
   \includegraphics[width=.45\textwidth]{figs/compare_model}
-  \caption{Average control loop execution time when multiple DNN
-    models are co-scheduled. }
+  \caption{Timing impact of co-scheduling multiple DNNs on different
+    embedded multicore platforms. %% 1Nx1C: one DNN
+    %% model using one core; 4Nx1C: four DNN models each using one core;
+    %% 1Nx2C: one DNN model using two cores; 2Nx2C: two DNN models each
+    %% using two cores.
+  }
   \label{fig:sys_model}
@@
 Figure~\ref{fig:sys_model} shows the results of the
-multi-model co-scheduling experiment. Once again, they can comfortably
+multi-model co-scheduling experiment
+(cf. Figure~\ref{fig:perf-vs-modelcnt}). Once again, they can comfortably
 satisfy 30 Hz real-time performance for all of the co-scheduled DNN control
-loops, and in the case of the TX2 (GPU), 100 Hz real-time control is still
-feasible. Given that the GPU must be shared among the co-scheduled DNN
+loops, and in the case of the TX2 (GPU), even 100 Hz real-time control
+is feasible in all co-scheduling setups.
+Given that the GPU must be shared among the co-scheduled DNN
 models, the results suggest that the TX2's GPU has sufficient capacity to
@@
   \centering
-  \begin{subfigure}{0.4\textwidth}
+  \begin{subfigure}{0.45\textwidth}
     \includegraphics[width=\textwidth]{figs/compare_benchmark_read}
@@
   \hfill
-  \begin{subfigure}{0.4\textwidth}
+  \begin{subfigure}{0.45\textwidth}
     \includegraphics[width=\textwidth]{figs/compare_benchmark}
@@
 Figure~\ref{fig:sys_bench} shows the results of synthetic memory
-intensive task co-scheduling experiment. 
+intensive task co-scheduling experiment
+(cf. Figure~\ref{fig:perf_vs_bandwidth}).
 For read co-runners (BwRead), the performance of all platforms
 gradually decreased as additional BwRead instances are introduced: up
-to 1.6X for Pi 3, up to 1.3X for Intel UP, and up to 1.6X and 2.2X for the TX2
-(CPU) and TX2(GPU), respectively.
+to 1.6X for Pi 3, up to 1.3X for Intel UP, and up to 1.6X and 2.2X for
+the TX2 (CPU) and TX2(GPU), respectively.
 For write co-runners (BwWrite), however, we observe generally more
-noticieable execution time increases. In particular, as we discussed
-earlier in Section~\ref{sec:evaluation}, the Pi 3 suffers up to 11.6X
-execution time increase. We also observe increased control loop
-execution timing in the Intel Up and Jetson TX2, although the degree
-of the increase is not as dramatic as the Pi 3.
+noticieable execution time increases. As we discussed
+earlier in Section~\ref{sec:evaluation}, Pi 3 suffers up to 11.6X
+execution time increase, while Intel UP and Jetson TX2 suffer less
+dramatic but still significant execution time increases.
 
@@
 considerable execution time increase (2.3X) despite the fact that the
-co-scheduled synthetic tasks do not utilize the GPU. In other words,
-the DNN model has dedicated access to the GPU. This is, however, a
+co-scheduled synthetic tasks do not utilize the GPU (i.e., gthe DNN
+model has dedicated access to the GPU.) This is, however, a 
 known characteristic of integrated CPU-GPU architecture based
-platforms where both CPU and GPU share the same memory
-subsystem~\cite{Ali2017}. As a result, the TX2 (GPU) fails to meet the
-10ms deadline for 100 Hz control that would have been feasible if
-there was no contention between the CPU cores and the GPU.
+platforms in which both CPU and GPU share the same memory
+subsystem~\cite{Ali2017} and therefore can suffer bandwidth contention
+as we observe in this case.
 
@@
 inexpensive as a Raspberry Pi 3, are powerful enough to support
-vision and end-to-end deep learning based real-time control
-applications. Furthermore, availability of CPU cores and GPU on these
-platforms allow consolidating mutiple deep neural network based AI
+DNN based real-time control applications. Furthermore, availability of
+CPU cores and GPU on these platforms allow consolidating mutiple DNN
 workloads. However, shared resource contention among these diverse

### 2018-04-19 paper/conclusion.tex
revision 2

@@
 in real-time. Importantly, DeepPicar's neural network architecture is
-almost identical to that of NVIDIA's real self-driving car.
+identical to that of NVIDIA's real self-driving car DAVE-2.
 
 Despite the complexity of the neural network, DeepPicar uses a
-low-cost Raspberry Pi 3 quad-core computer as its sole computing
-resource. We systematically analyzed the real-time characteristics of
-the Pi 3 platform in the context of deep-learning based real-time
-control applications, with a special emphasis on real-time deep neural
-network inferencing.
+low-cost embedded quad-core computer, the Raspberry Pi 3,  as its sole
+computing resource.
+We systematically analyzed the platform's real-time capability in
+supporting the DNN-based real-time control task.
 We also evaluated other, more powerful, embedded computing
 platforms to better understand achievable real-time performance of
-DeepPicar's deep-learning based control system and the impact of
+DeepPicar's DNN based control system and the impact of
 computing hardware architectures.
@@
 100 Hz, depending on the platform and its system
-configuration. However, shared resource contention remains an
-important issue that must be considered in applying deep-learning
-models on shared memory based embedded computing platforms.
+configuration. Futhermore, all platforms were capable of consolidating
+multiple DNN models and/or tasks.
+
+However, when consolidating multiple models/tasks, shared resource
+contention remains an important issue that must be considered to
+ensure desired real-time performance on these shared memory based
+embedded computing platforms. 
+Toward this end, we evaluated the impact of shared resource contention
+to the DNN workload in diverse consolidated workload setups, and
+evaluated the effectivness of state-of-the-art shared resource
+isolation mechanisms in protecting real-time performance of the DNN
+workload.
 

### 2018-04-19 paper/evaluation-cachepart.tex
revision 2

@@
 Figure \ref{fig:palloc_bandwidth_exectime} shows the
-result. For BwRead, when cache partitioning is enabled, DNN
-performance becomes slightly better, with a improvement of  
-\textasciitilde6.5\%. For BwWrite, cache partitioning actually makes
-the DNN performance slightly worse that that of w/o partitioning.
+result. Compared to Figure~\ref{fig:perf_vs_bandwidth} where no cache
+partioning is applied, assigning a dedicated L2 cache parititon to
+each core does not provide significant isolation benefits: For BwRead
+co-runners, cache partitioning slightly improves isolation; but for
+BwWrite co-runners, cache partitioning causes worse worst-case
+slowdown.
 

### 2018-04-19 paper/evaluation-drampart.tex
revision 2

@@
 bandwidth budgets within each regulation period (i.e., 1ms) by
-scheduling high-priority idle kernel threads.
+scheduling high-priority idle kernel threads, until the next period begins.
 
 In the first experiment, we measure the performance of the DNN model
-running on a single core, Core 0, first w/o using MemGuard and then w/
+on a single core, first w/o using MemGuard and then w/
 using MemGuard while varying the core's bandwidth throttling parameter
@@
 
-Figure \ref{fig:memguard_multicore} shows the result. When the DNN
-model executing core is throttled at 400 MB/s or more, performance of
-the model is largely the same as the non-throttled case. However, as
-we decreate the assigned memory badnwdith below 300 MB/s, we start to
-observe noticably decreases in the model's performance. In other
+Figure \ref{fig:memguard_multicore} shows the result. When core
+executing the DNN model is throttled at 400 MB/s or more, performance
+of the model is largely the same as the non-throttled case. However, as
+we decrease the assigned memory bandwidth below 300 MB/s, we start to
+observe noticeable decreases in the model's performance. In other
 words, the DNN model is sensitive to memory bandwidth and that it
@@
 In the next experiment, we again repeat the experiment in
-Section~\ref{sec:eval-memhog}---i.e., co-scheduling the DNN model and
-three Bandwidth (BwRead or BwWrite) instances---but this time we
-throttle the cores that execute the co-runners using MemGuard by 
-varying their bandwidth budgets to see ther impact in the DNN model's
+Section~\ref{sec:eval-memhog}---i.e., co-scheduling memory intensive
+synthetic tasks---but this time we
+throttle the cores of the co-runners using MemGuard and vary their
+memory bandwidth budgets to see their impact on the DNN model's 
 performance.

### 2018-04-19 paper/evaluation.tex
revision 2

@@
 
-\subsection{Summary of the Findings}
-So far, we have evaluated DeepPicar's real-time
-characteristics from the perspective of end-to-end deep learning based
-real-time control, and made several observations.
+%% \subsection{Summary of the Findings}
+%% So far, we have evaluated DeepPicar's real-time
+%% characteristics from the perspective of end-to-end deep learning based
+%% real-time control, and made several observations.
 
-First, we find that DeepPicar's computing platform, the Raspberry Pi
-3, offers sufficient computing capacity to perform deap-learning based
-real-time control at 30 Hz frequency (or 33.3 ms per control
-loop). Given the complexity of the DNN used, we were pleasantly
-surprised by this finding. 
-%% The time breakdown shows that the DNN inferencing operation, 
-%% performed by the Tensorflow library, dominates the execution time, 
-%% which is expected.
+%% First, we find that DeepPicar's computing platform, the Raspberry Pi
+%% 3, offers sufficient computing capacity to perform deap-learning based
+%% real-time control at 30 Hz frequency (or 33.3 ms per control
+%% loop). Given the complexity of the DNN used, we were pleasantly
+%% surprised by this finding. 
+%% %% The time breakdown shows that the DNN inferencing operation, 
+%% %% performed by the Tensorflow library, dominates the execution time, 
+%% %% which is expected.
 
-Second, we find that scalability of Tensorflow's DNN 
-implementation is limited. We find that using all four cores can double 
-performance of the CNN, but compared to using three cores, performance
-improvement is relatively minimal.
+%% Second, we find that scalability of Tensorflow's DNN 
+%% implementation is limited. We find that using all four cores can double 
+%% performance of the CNN, but compared to using three cores, performance
+%% improvement is relatively minimal.
 
-Third, we find that consolidating multiple DNN models---on different CPU
-cores---is feasible as we find: (1) DNN performance using a single
-core is not much worse than using multiple cores; (2) multiple DNN
-models running simultaneously cause moderate interference.
+%% Third, we find that consolidating multiple DNN models---on different CPU
+%% cores---is feasible as we find: (1) DNN performance using a single
+%% core is not much worse than using multiple cores; (2) multiple DNN
+%% models running simultaneously cause moderate interference.
 
-Fourth, we find that consolidating memory (DRAM) performance
-intensive applications (especially write-heavy ones) can jeopadize
-the DNN inferencing task's performance even when all tasks have
-dedicated cores due to contention in shared hardware resources; we
-observe up to 11.6X slowdown.
+%% Fourth, we find that consolidating memory (DRAM) performance
+%% intensive applications (especially write-heavy ones) can jeopadize
+%% the DNN inferencing task's performance even when all tasks have
+%% dedicated cores due to contention in shared hardware resources; we
+%% observe up to 11.6X slowdown.
 
-Lastly, we find that the DNN inferencing workload is not sensitive
-to cache space and that cache partitioning is not effective in
-providing timing isolation to our DNN workload. On the other hand, we
-find that the DNN workload is sensitive to memory bandwidth and that
-memory bandwidth throttling is effective in improving performance
-isolation.
+%% Lastly, we find that the DNN inferencing workload is not sensitive
+%% to cache space and that cache partitioning is not effective in
+%% providing timing isolation to our DNN workload. On the other hand, we
+%% find that the DNN workload is sensitive to memory bandwidth and that
+%% memory bandwidth throttling is effective in improving performance
+%% isolation.
 

### 2018-04-19 paper/related.tex
revision 2

@@
 F1$/$10~\cite{upennf1tenth} are both based on a Trexx 1/10 scale RC
-car and a powerful NVIDIA Jetson computing platform equipped with many
-sophisticated sensor packages such as a lidar. However, they both cost
-more than \$3,000, requiring considerable
+car and a powerful NVIDIA Jetson multicore computing platform, which
+is equipped with many sophisticated sensor packages such as a lidar.
+However, they both cost more than \$3,000, requiring considerable
 investment. DonkeyCar~\cite{donkeycar} is similar to our DeepPicar as
-it also uses a Raspberry Pi 3 and a CNN for end-to-end control,
-although it costs slightly more because it uses a more expensive 1/16
-scale RC car. The main contribution of our paper is not the hardware
-design but in the detailed analysis of computational characteristics
-for the CNN-based end-to-end real-time control.
+it also uses a Raspberry Pi 3 and a similar CNN for end-to-end
+control, although it costs more (about \$200).
+%% because it uses a more
+%% expensive 1/16 scale RC car than our 1/24 scale one.
+The main contribution of our paper is in the detailed analysis of
+computational aspects of executing a CNN-based real-time control
+workload on diverse embedded computing platforms.
 
-In this paper, we have analyzed real-time performance of a CNN, which
-was used in NVIDIA's DAVE-2 self-driving car~\cite{Bojarski2016}, on a 
-low-cost raspberry pi 3 quad-core platform. It should be noted,
+In this paper, we have analyzed real-time performance of a real-world
+CNN, which was used in NVIDIA's DAVE-2 self-driving
+car~\cite{Bojarski2016}, on a low-cost raspberry pi 3 quad-core 
+platform and other embedded multicore platforms. It should be noted, 
 however, that DAVE-2's CNN is relatively small compared
@@
 computational complexity itself.
-When a DNN is deployed in those implementations with limited
-resources, such as memory and power, the floating-point operations
-involved in the large matrix multiplications are a burdensome task.
+%% When a DNN is deployed in those implementations with limited
+%% resources, such as memory and power, the floating-point operations
+%% involved in the large matrix multiplications are a burdensome task.
 Many recent advances in network compression have shown promising results
@@
 topology~\cite{han2015deep}.
+
 Another common compression method is to reduce the
@@
 These research efforts are expected to make complex DNNs more
-accessible for real-time embedded systems. We plan to investigate the
-feasibility of these approaches in the context of DeepPicar so that we
-can use even more resource constrained micro-controller class
-computing platforms instead of Raspberry Pi3.
+accessible for a wider range of real-time embedded systems. We plan to
+investigate the feasibility of these approaches in the context of
+DeepPicar so that we can use even more resource constrained
+micro-controller class computing platforms.
 

### 2018-04-19 paper/comparison.tex
minor fixes

@@
               4x Atom@1.92GHz/2MB L2 &
-              4x Cortex-A57@2GHz/2MB L2 \\
+              4x Cortex-A57@2.0GHz/2MB L2 \\
             &              &              & 2x Denver@2.0GHz/2MB L2 (not used)  \\
@@
 	Peak Memory Bandwidth & 8.5 GB/s & 12.8 GB/s & 59.7 GB/s \\
+	%% \hline
+	%% Measured Bandwidth & 2127.94 MB/s & 3951.94 MB/s & 4447.90 MB/s \\
 	\hline
-	Measured Bandwidth & 2127.94 MB/s & 3951.94 MB/s & 4447.90 MB/s \\
-	\hline
-	Cost (\$) & 35 & 100 & 600 \\
+	Cost  & \$35 & \$100 & \$600 \\
 	\hline
@@
 In this section, we compare three computing platforms---the Raspberry
-Pi 3, the Intel UP~\footnote{http://www.up-board.org/up/} and the NVIDIA
-Jetson
-TX2~\footnote{http://www.nvidia.com/object/embedded-systems-dev-kits-modules.html}---from
-the point of view of supporting end-to-end deep learning
-based autonomous vehicles. 
+Pi 3, the Intel UP~\cite{intelup} and the NVIDIA Jetson
+TX2~\cite{nvidiajetson}---from the point of view of supporting
+end-to-end deep learning based autonomous vehicles. 
 Table~\ref{tbl:platforms} shows architectural features of the three

### 2018-04-19 paper/related.tex
minor fixes

@@
 testbeds. MIT's RaceCar~\cite{shin2017project} and UPenn's
-F1$/$10~\cite{upennf1tenth} are both based on a Trexx 1/10 scale RC
-car and a powerful NVIDIA Jetson multicore computing platform, which
+F1$/$10~\cite{upennf1tenth} are both based on a Traxxas 1/10 scale RC
+car and a NVIDIA Jetson multicore computing platform, which
 is equipped with many sophisticated sensor packages such as a lidar.

### 2018-04-20 paper/comparison.tex
Paper revisions

@@
 Both the Intel UP and Jetson TX2 exhibit better performance than
-Raspberry Pi 3. When all four CPU cores are used, Intel UP is
-1.33X faster than Pi 3, while TX2 (CPU) and TX2 (GPU) are 2.79X and
-4.16X times faster, respectively, than Pi 3.
+Raspberry Pi 3. When all four CPU cores are used, the Intel UP is
+1.33X faster than Pi 3, while the TX2 (CPU) and TX2 (GPU) are 2.79X and
+4.16X times faster that the Pi 3, respectively.
 Thus, they all satisfy 33.3 ms WCET by a clear margin, and, in the
@@
 with the help of its GPU.
-Another observation is that the DNN task's performance on TX2 (GPU)
+Another observation is that the CNN task's performance on TX2 (GPU)
 does not change much as we increase the number of cores. This
-isbecause most of the neural network computation is done by the GPU.
+is because most of the neural network computation is done by the GPU.
 
@@
   \includegraphics[width=.45\textwidth]{figs/compare_model}
-  \caption{Timing impact of co-scheduling multiple DNNs on different
+  \caption{Timing impact of co-scheduling multiple CNNs on different
     embedded multicore platforms. %% 1Nx1C: one DNN
@@
 (cf. Figure~\ref{fig:perf-vs-modelcnt}). Once again, they can comfortably
-satisfy 30 Hz real-time performance for all of the co-scheduled DNN control
+satisfy 30 Hz real-time performance for all of the co-scheduled CNN control
 loops, and in the case of the TX2 (GPU), even 100 Hz real-time control
 is feasible in all co-scheduling setups.
-Given that the GPU must be shared among the co-scheduled DNN
+Given that the GPU must be shared among the co-scheduled CNN
 models, the results suggest that the TX2's GPU has sufficient capacity to
-accomodate multiple instances of the DNN models we tested.
+accomodate multiple instances of the CNN models we tested.
 
@@
   \caption{Average processing time vs. the number of memory
-    memory intensive co-runners.}
+    intensive co-runners.}
   \label{fig:sys_bench}    
@@
 
-Figure~\ref{fig:sys_bench} shows the results of synthetic memory
-intensive task co-scheduling experiment
+Figure~\ref{fig:sys_bench} shows the results of the synthetic memory
+intensive task co-scheduling experiments
 (cf. Figure~\ref{fig:perf_vs_bandwidth}).
 For read co-runners (BwRead), the performance of all platforms
-gradually decreased as additional BwRead instances are introduced: up
-to 1.6X for Pi 3, up to 1.3X for Intel UP, and up to 1.6X and 2.2X for
+gradually decreased as additional BwRead instances were introduced: up
+to 1.6X for the Pi 3, up to 1.3X for the Intel UP, and up to 1.6X and 2.2X for
 the TX2 (CPU) and TX2(GPU), respectively.
@@
 noticieable execution time increases. As we discussed
-earlier in Section~\ref{sec:evaluation}, Pi 3 suffers up to 11.6X
-execution time increase, while Intel UP and Jetson TX2 suffer less
+earlier in Section~\ref{sec:evaluation}, the Pi 3 suffers up to 11.6X
+execution time increase, while the Intel UP and Jetson TX2 suffer less
 dramatic but still significant execution time increases.
@@
 considerable execution time increase (2.3X) despite the fact that the
-co-scheduled synthetic tasks do not utilize the GPU (i.e., gthe DNN
+co-scheduled synthetic tasks do not utilize the GPU (i.e., the CNN
 model has dedicated access to the GPU.) This is, however, a 
 known characteristic of integrated CPU-GPU architecture based
-platforms in which both CPU and GPU share the same memory
+platforms in which both the CPU and GPU share the same memory
 subsystem~\cite{Ali2017} and therefore can suffer bandwidth contention
@@
 inexpensive as a Raspberry Pi 3, are powerful enough to support
-DNN based real-time control applications. Furthermore, availability of
-CPU cores and GPU on these platforms allow consolidating mutiple DNN
+CNN based real-time control applications. Furthermore, availability of
+CPU cores and a GPU on these platforms allows consolidating multiple CNN
 workloads. However, shared resource contention among these diverse

### 2018-04-20 paper/conclusion.tex
Paper revisions

@@
 We systematically analyzed the platform's real-time capability in
-supporting the DNN-based real-time control task.
+supporting the CNN-based real-time control task.
 We also evaluated other, more powerful, embedded computing
 platforms to better understand achievable real-time performance of
-DeepPicar's DNN based control system and the impact of
+DeepPicar's CNN based control system and the impact of
 computing hardware architectures.
@@
 configuration. Futhermore, all platforms were capable of consolidating
-multiple DNN models and/or tasks.
+multiple CNN models and/or tasks.
 
@@
 Toward this end, we evaluated the impact of shared resource contention
-to the DNN workload in diverse consolidated workload setups, and
+to the CNN workload in diverse consolidated workload setups, and
 evaluated the effectivness of state-of-the-art shared resource
-isolation mechanisms in protecting real-time performance of the DNN
+isolation mechanisms in protecting real-time performance of the CNN
 workload.
@@
 As future work, we plan to investigate ways to reduce computational
-and memory overhead of DNN inferencing and to evaluate the
+and memory overhead of CNN inferencing and to evaluate the
 effectiveness of FPGA and other specialized accelerators.

### 2018-04-20 paper/evaluation-cachepart.tex
Paper revisions

--- 2018-04-19 paper/evaluation-cachepart.tex
@@
 Cache partitioning is a well-known technique to improve isolation in
-multicore by giving a dedicated cache space to each individaul 
+multicore systems by giving a dedicated cache space to each individaul 
 task or core. In this experiment, we use PALLOC~\cite{yun2014rtas}, a
@@
 Page coloring is a OS technique that controls the physical addresses
-of memory pages. By allocating pages over non-overlapping sets of the
-cache, the OS can effectively partition the cache.
-Using PALLOC, we investigate the effect of cache partitioning in
-protecting DeepPicar's DNN based controller.
+of memory pages. By allocating pages over non-overlapping cache sets,
+the OS can effectively partition the cache.
+Using PALLOC, we investigate the effect of cache partitioning on
+protecting DeepPicar's CNN based controller.
 
@@
 Figure~\ref{fig:cache-mapping} shows the physical address
-mapping of Raspberry Pi 3's BCM2837 processor, which has 32K private
-L1 I\&D (4way) caches and a share 512KB L2 (16 way) cache. In order not
-to partition the private L1 caches, we use the bit 13 and 14 for
+mapping of the Raspberry Pi 3's BCM2837 processor, which has 32K private
+L1 I\&D (4way) caches and a shared 512KB L2 (16 way) cache. In order
+to avoid partitioning the private L1 caches, we use bits 13 and 14 for
 coloring, which results in 4 usuable page colors.
@@
 In the first experiment, we investigate the cache space sensitivity of
-DeepPicar's DNN-based control loop. Using PALLOC, we create 4
+the DeepPicar's CNN-based control loop. Using PALLOC, we create 4
 different cgroups which are configured to use 4, 3,
 2, and 1 colors (100\%, 75\%, 50\% and 25\% of the L2 cache
-space, respectively). We then execute the DNN control loop (inference)
+space, respectively). We then execute the CNN control loop (inference)
 on one core using a different cgroup cache partition, one at a time,
@@
   \includegraphics[width=.45\textwidth]{figs/palloc_multicore}
-  \caption{Cache space sensitivity of the DNN controller.}
+  \caption{Cache space sensitivity of the CNN controller.}
   \label{fig:palloc_multicore}
@@
 
-Figure~\ref{fig:palloc_multicore} shows the result. As can be seen,
-the DNN inference timing is hardly changed at all regardless of the
+Figure~\ref{fig:palloc_multicore} shows the results. As can be seen,
+the CNN inference timing hardly changes at all regardless of the
 size of the allocated L2 cache space. In other words, we find that
-the DNN workload is largely insensitive to L2 cache space.
+the CNN workload is largely insensitive to L2 cache space.
 
@@
 experiment, we repeat the experiment in
-Section~\ref{sec:eval-memhog}---i.e., co-scheduling the DNN model and
+Section~\ref{sec:eval-memhog}---i.e., co-scheduling the CNN model and
 three Bandwidth (BwRead or BwWrite) instances---but this time we
@@
   \caption{Average processing time vs. the number of memory
-memory intensive co-runners; Each core (task) is given a equal-sized
+intensive co-runners; Each core (task) is given a equal-sized
 dedicated cache partition.}
@@
 Figure \ref{fig:palloc_bandwidth_exectime} shows the
-result. For BwRead, when cache partitioning is enabled, DNN
-performance becomes slightly better, with a improvement of  
-\textasciitilde6.5\%. For BwWrite, cache partitioning actually makes
-the DNN performance slightly worse that that of w/o partitioning.
+results. Compared to Figure~\ref{fig:perf_vs_bandwidth} where no cache
+partioning is applied, assigning a dedicated L2 cache parititon to
+each core does not provide significant isolation benefits. For BwRead
+co-runners, cache partitioning slightly improves isolation, but for
+BwWrite co-runners, cache partitioning causes worse worst-case
+slowdown.
 
-In summary, we find that the DNN inferencing workload is not sensitive
+In summary, we find that the CNN inferencing workload is not sensitive
 to cache space and that cache partitioning is not effective in
-providing timing isolation to our DNN workload.
+providing timing isolation for our CNN workload.
 

### 2018-04-20 paper/evaluation-drampart.tex
Paper revisions

@@
 
-In this subsection, we examine the DNN workload's memory bandwidth
+In this subsection, we examine the CNN workload's memory bandwidth
 sensitivity and the effect of memory bandwidth throttling in providing
@@
 can limit the amount of memory bandwidth each core receives. MemGuard
-operates periodically, at an 1ms interval, and uses hardware
+operates periodically, at a 1 ms interval, and uses hardware
 performance counters to throttle cores if they exceed their given
-bandwidth budgets within each regulation period (i.e., 1ms) by
+bandwidth budgets within each regulation period (i.e., 1 ms) by
 scheduling high-priority idle kernel threads, until the next period begins.
 
-In the first experiment, we measure the performance of the DNN model
+In the first experiment, we measure the performance of the CNN model
 on a single core, first w/o using MemGuard and then w/
@@
 
-Figure \ref{fig:memguard_multicore} shows the result. When core
-executing the DNN model is throttled at 400 MB/s or more, performance
+Figure \ref{fig:memguard_multicore} shows the results. When the core
+executing the CNN model is throttled at 400 MB/s or more, the performance
 of the model is largely the same as the non-throttled case. However, as
@@
 observe noticeable decreases in the model's performance. In other
-words, the DNN model is sensitive to memory bandwidth and that it
-requires 400 MB/s or more bandwidth should be given to ensure ideal
-performance.
+words, the CNN model is sensitive to memory bandwidth and it
+requires 400 MB/s or more bandwidth to ensure ideal performance.
 
-In the next experiment, we again repeat the experiment in
+In the next experiment, we repeat the experiment in
 Section~\ref{sec:eval-memhog}---i.e., co-scheduling memory intensive
@@
 throttle the cores of the co-runners using MemGuard and vary their
-memory bandwidth budgets to see their impact on the DNN model's 
+memory bandwidth budgets to see their impact on the CNN model's 
 performance.
@@
 
-Figures \ref{fig:memguard_bandwidth} shows the result.
+Figures \ref{fig:memguard_bandwidth} shows the results.
 Note that, in case of BwWrite co-runners, we have assigned half the
-bandwidth budget because MemGuard currently only account L2 refills
-but not write-backs, which effectively allow twice the allocated
-memory bandwidth budget. Regardless, the result clearly shows that
-limiting co-runners's memory bandwidth is effective in protecting the
-DNN model's performance.
+bandwidth budget because MemGuard currently only accounts for L2 refills
+but not write-backs, which effectively allows twice the allocated
+memory bandwidth budget. Regardless, the results clearly show that
+limiting the co-runners's memory bandwidth is effective in protecting the
+CNN model's performance.
 
-In summary, we find that the DNN inferencing workload is sensitive to
+In summary, we find that the CNN inferencing workload is sensitive to
 memory bandwidth and that memory bandwidth throttling is effective in
-improving performance isolation of the DNN workload.
+improving the performance isolation of the CNN workload.
 

### 2018-04-20 paper/evaluation.tex
Paper revisions

@@
 \footnote{TensorFlow currently only supports NVIDIA's GPUs.}.
-For software, we use Ubuntu MATE 16.04, TensorFlow 1.1.
+For software, we use Ubuntu MATE 16.04 and TensorFlow 1.1.
 We disabled DVFS (dynamic voltage frequency scaling) and
 configured the clock speed of each core statically at the maximum 1.2GHz.
-We use the SCHED\_FIFO real-time scheduler to schedule the DNN control
+We use the SCHED\_FIFO real-time scheduler to schedule the CNN control
 task while using the CFS when executing memory intensive co-runners.
@@
 standard robotics control vs. end-to-end deep learning approach).
-This means that the DNN inference operation must be completed
+This means that the CNN inference operation must be completed
 within the inner-most control loop update frequency.
@@
 To understand achievable control-loop update frequencies, we
-experimentally measured the execution times of DeepPicar's DNN
+experimentally measured the execution times of DeepPicar's CNN
 inference operations.
@@
     Image pre-processing & 2.77  &  2.90 &  2.87  & 0.04 \\ \hline
-    DNN inferencing      & 18.49 & 19.30 & 18.99  & 0.20 \\ \hline
+    CNN inferencing      & 18.49 & 19.30 & 18.99  & 0.20 \\ \hline
     Total Time           & 22.87 & 23.74 & 23.38  & 0.20 \\ \hline
@@
 Raspberry Pi 3 were used by the TensorFlow library when performing the
-DNN inference operations.
+CNN inference operations.
 
@@
 frames. This is because the amount of computation needed to perform
-a DNN inferencing operation is fixed at the DNN architecture design
+a CNN inferencing operation is fixed at the CNN architecture design
 time and does not change at runtime over different inputs (i.e.,
 different image frames). This predictable timing behavior is a highly
-desirable property for real-time systems, making DNN inferencing an
+desirable property for real-time systems, making CNN inferencing an
 attractive real-time workload.
@@
 
-As noted in~\cite{NVIDIA2015}, DNN inferencing is inherently more
+As noted in~\cite{NVIDIA2015}, CNN inferencing is inherently more
 difficult to parallelize than training because the easiest
@@
 
-\subsection{Effect of Co-scheduling Multiple DNN Models}
+\subsection{Effect of Co-scheduling Multiple CNN Models}
 
-In this experiment, we launch multiple instances of DeepPicar's DNN
+In this experiment, we launch multiple instances of DeepPicar's CNN
 model at the same time and measure its impact on their inference
@@
   \includegraphics[width=.45\textwidth]{figs/perf_vs_modelcnt}
-  \caption{Timing impact of co-scheduling multiple DNNs. 1Nx1C: one DNN
-    model using one core; 4Nx1C: four DNN models each using one core;
-    1Nx2C: one DNN model using two cores; 2Nx2C: two DNN models each
+  \caption{Timing impact of co-scheduling multiple CNNs. 1Nx1C: one CNN
+    model using one core; 4Nx1C: four CNN models each using one core;
+    1Nx2C: one CNN model using two cores; 2Nx2C: two CNN models each
     using two cores.} 
@@
 Figure~\ref{fig:perf-vs-modelcnt} shows the results. In the figure, the
-X-axis shows the system configuration: \#of DNN models x \#of CPU
-cores/DNN. For example, `4Nx1C' means running four DNN models each of
+X-axis shows the system configuration: \#of CNN models x \#of CPU
+cores/CNN. For example, `4Nx1C' means running four CNN models each of
 which is assigned to run on one CPU core, whereas `2Nx2C' means running
-two DNN models, each of which is assigned to run on two CPU
+two CNN models, each of which is assigned to run on two CPU
 cores. The Y-axis shows the average inference timing.
-The two bars on the left show the impact of co-scheduling four DNN
-models. Compared to executing a single DNN model on one CPU core
-(1Nx1C), when four DNN models are co-scheduled (4Nx1C), each model
+The two bars on the left show the impact of co-scheduling four CNN
+models. Compared to executing a single CNN model on one CPU core
+(1Nx1C), when four CNN models are co-scheduled (4Nx1C), each model
 suffers an average inference time increase of approximately 11 ms,
-or 24\%. On the other hand, when two DNN models, each using two CPU
+or 24\%. On the other hand, when two CNN models, each using two CPU
 cores, are co-scheduled (2Nx2C), the average inference timing is increased by
@@
 
-In this experiment, we investigate \emph{worst-case} impact of shared 
-resource contention to DeepPicar's DNN inference timing using  
+In this experiment, we investigate the \emph{worst-case} impact of shared 
+resource contention on DeepPicar's CNN inference timing using  
 a synthetic memory benchmark. Specifically, we use the \emph{Bandwidth}
 benchmark from the IsolBench suite~\cite{Valsan2016}, which
-sequentially reads or writes a big array; we henceforth refer BwRead 
+sequentially reads or writes a big array; we henceforth refer to BwRead 
 as Bandwidth with read accesses and BwWrite as the one with write
-accesses. The experiment setup is as follows. We run a single DNN
+accesses. The experiment setup is as follows: We run a single CNN
 model on one core, and co-schedule an increasing number of the
@@
   \caption{Average processing time vs. the number of memory
-    memory intensive co-runners introduced.}
+    intensive co-runners introduced.}
   \label{fig:perf_vs_bandwidth}
@@
 
-Figure~\ref{fig:perf_vs_bandwidth} shows the result. Note
-that BwWrite co-runners cause significantly high execution time
-incrases---up to 11.6X---on DNN inferencing while BwRead co-runners
+Figure~\ref{fig:perf_vs_bandwidth} shows the results. Note
+that BwWrite co-runners cause significantly higher execution time
+increases---up to 11.6X---on CNN inferencing while BwRead co-runners
 cause relatively much smaller time increases. While execution time
-increases are expected, the degree of which in the worst-case is quite
-surprising---\emph{if the DNN controller was driving an actual car, it
-  would result in a crash of the car!}
+increases are expected, the degree that is seen in the worst-case 
+is quite surprising---\emph{if the CNN controller was driving an actual 
+car, it would result in a car crash!}
 
-Given the importance of predictable timing on real-time control, such
-as our DNN based control task, we wanted to understand (1) the main
-source of the timing and (2) to evaluate existing isolation methods to
+Given the importance of predictable timing for real-time control, such
+as our CNN based control task, we wanted to: (1) understand the main
+source of the timing and (2) evaluate existing isolation methods to
 avoid this kind of timing interference. Specifically, shared cache
-space and DRAM bandwidth are two most well-known sources of contention
-in multicore. Thus, in the following, we investigate if and how
-they influence the observed timing interference and how effective are
-the existing mitigation methods.
+space and DRAM bandwidth are the two most well-known sources of contention
+in multicore systems. Thus, in the following sections, we investigate whether
+and to what extent they influence the observed timing interference and 
+the effectiveness of existing mitigation methods.
 

### 2018-04-20 paper/overview.tex
Paper revisions

@@
 control, the observed control performance of the system is respectable,
-likely due to the inherently stocastic nature of DNN.
+likely due to the inherently stocastic nature of CNNs.
 
@@
 time of the computing platform. Higher camera latency could
-negatively affect control performance, because the DNN would analyze
+negatively affect control performance, because the CNN would analyze
 stale scenes. The actuator (motor) control latency from the time
@@
 
-Our trained DNN models showed good prediction accuracy, successfully
+Our trained CNN models showed good prediction accuracy, successfully
 navigating several different tracks we used for training.

### 2018-04-20 paper/related.tex
Paper revisions

@@
 car and a NVIDIA Jetson multicore computing platform, which
-is equipped with many sophisticated sensor packages such as a lidar.
-However, they both cost more than \$3,000, requiring considerable
+is equipped with many sophisticated sensor packages, such as a lidar.
+However, they both cost more than \$3,000, requiring a considerable
 investment. DonkeyCar~\cite{donkeycar} is similar to our DeepPicar as
@@
 Google's recent study~\cite{huang2017speed} have between 3M to 54M
-parameters, much larger than DAVE-2's CNN.
+parameters, and are much larger than DAVE-2's CNN.
 Using such large CNN models will be challenging on
@@
 platforms will certainly make processing these complex CNNs faster,
-another actively investigated approrach is to reduce the required
+another actively investigated approach is to reduce the required
 computational complexity itself.
@@
 Many recent advances in network compression have shown promising results
-in reducing such computational cost during the feedforward
+in reducing such computational costs during the feedforward
 process. The fundamental assumption in those techniques is that the
-DNNs are redundant in their structure and representation. For example,
+CNNs are redundant in their structure and representation. For example,
 network pruning can thin out the network and provides a more condensed
@@
 quantization level of the network parameters, so that arithmetic
-defined with the floating-point operations are replaced with low-bit
+defined with floating-point operations are replaced with low-bit
 fixed-point counterparts. To this end, single bit quantization of the
@@
 hardware implementations. In~\cite{rastegari2016xnor}, it was shown
-that the binarized convolution can substitute the expensive
+that the binarized convolution could substitute the expensive
 convolutional feedforward operations in a regular Convolutional Neural
@@
 network for the hand written digit recognition task as well as 23
-times faster matrix multiplication tasks on
+times faster matrix multiplication tasks on a
 GPU~\cite{hubara2016binarized}.
@@
 
-These research efforts are expected to make complex DNNs more
+These research efforts are expected to make complex CNNs more
 accessible for a wider range of real-time embedded systems. We plan to

### 2018-04-20 paper/conclusion.tex
fixes and acknowledgement

@@
 We find all tested embedded platforms, including the Pi 3, are capable
-of supporting deep-learning based real-time control, from 20 Hz up to
-100 Hz, depending on the platform and its system
-configuration. Futhermore, all platforms were capable of consolidating
+of supporting the CNN based real-time control, from 20 Hz up to
+100 Hz, depending on the platform.
+Futhermore, all platforms were capable of consolidating
 multiple CNN models and/or tasks.
 
-However, when consolidating multiple models/tasks, shared resource
+However, we also find that shared resource
 contention remains an important issue that must be considered to
@@
 evaluated the effectivness of state-of-the-art shared resource
-isolation mechanisms in protecting real-time performance of the CNN
-workload.
+isolation mechanisms in protecting performance of the CNN
+based real-time control workload.
 

### 2018-04-20 paper/evaluation-cachepart.tex
fixes and acknowledgement

@@
 Cache partitioning is a well-known technique to improve isolation in
-multicore systems by giving a dedicated cache space to each individaul 
+a multicore system by giving a dedicated cache space to each individaul 
 task or core. In this experiment, we use PALLOC~\cite{yun2014rtas}, a

### 2018-04-20 paper/evaluation.tex
fixes and acknowledgement

@@
 and 
-(4) the effect of co-scheduling memory bandwidth intensive co-runners.
-%% and
-%% (5) the effect of partitioning the shared L2 cache.
+(4) the effect of co-scheduling memory bandwidth intensive co-runners,
+and
+(5) the effect of shared L2 cache partitioning and memory bandwidth
+throttling for guaranteed real-time performance.
 

### 2018-04-20 paper/main.tex
revision up to Section 4.E
fixes and acknowledgement

@@
 \author{Michael G. Bechtel$^\dagger$, Elise McEllhiney$^\dagger$, Minje Kim$^\star$, Heechul Yun$^\dagger$\\
-  $^\dagger$ University of Kansas, USA. \{mbechtel, e908m428, heechul.yun\}@ku.edu \\
+  $^\dagger$ University of Kansas, USA. \{mbechtel, elisemmc, heechul.yun\}@ku.edu \\
   $^\star$ Indiana University, USA. minje@indiana.edu \\
@@
 \begin{abstract}
-We present DeepPicar, a low-cost deep neural network (DNN) based
+We present DeepPicar, a low-cost deep neural network based
 autonomous car platform. DeepPicar is a small scale
-replication of a real self-driving car called Dave-2 by NVIDIA, which
+replication of a real self-driving car called DAVE-2 by NVIDIA. DAVE-2
 drove on public roads using a deep convolutional neural network (CNN), 
-that takes images from a front-facing camera as input and produces
-car steering angles as output. DeepPicar uses almost the exact same 
+which takes images from a front-facing camera as input and produces
+car steering angles as output. DeepPicar uses the same 
 network architecture---9 layers, 27 million connections and 250K
-parameters---and can be trained to drive itself, in real-time, using a
-web camera and a modest Raspberry Pi 3 quad-core platform.
+parameters---and can be trained to drive itself using a
+web camera and a modest Raspberry Pi 3 quad-core platform in real-time.
 Using DeepPicar, we analyze the Pi 3's computing capabilities to 
 support end-to-end deep learning based real-time control of autonomous
-vehicles. We also systematically compare other contemporary embedded
+vehicles. 
+We also systematically compare other contemporary embedded
 computing platforms using the DeepPicar's CNN based real-time control
-software as a workload. 
+workload.
+
 We find all tested platforms, including the Pi 3, are capable of
-supporting deep-learning based real-time control, from 20 Hz up to 100
+supporting the CNN based real-time control, from 20 Hz up to 100
 Hz depending on hardware platform. 
@@
 important issue that must be considered in applying deep-learning
-models on shared memory based embedded computing platforms.
+models on shared memory based embedded computing platforms; we observe
+up to 11.6X control loop execution time increase due to such shared
+resource contention. We evaluate state-of-the-art cache
+partitioning and memory bandwidth throttling techniques on the Pi 3
+in protecting the CNN workload. We find cache partition is
+ineffective, while memory bandwidth throttling is an effective
+solution.
 \end{abstract}
@@
 \input{conclusion}
+\input{acknowledgement}
 %-------------------------------------------------------------------------

### 2018-04-20 paper/overview.tex
fixes and acknowledgement

@@
   \centering
-  \includegraphics[width=.4\textwidth]{figs/DeepPicar_platform}
+  \includegraphics[width=.38\textwidth]{figs/DeepPicar_platform}
   \caption{DeepPicar platform.}
@@
 
-Note that although the steering angle output of the network ($angle$) is
-a continuous real value, the RC car platform we used only supports
+Note that although the steering angle output of the CNN ($angle$) is
+a continuous real value, the RC car we used only supports
 three discrete angles---left (-30$^{\circ}$), center 
@@
 For instance, the DeepPicar could remain on Track 1
-(Figure \label{fig:track}) over 10 minutes at a moderate speed (50\%
+(Figure~\ref{fig:track}) over 10 minutes at a moderate speed (50\%
 throttle), at which point we stopped the experiment, and more than one

### 2018-04-20 paper/related.tex
fixes and acknowledgement

@@
 Google's recent study~\cite{huang2017speed} have between 3M to 54M
-parameters, and are much larger than DAVE-2's CNN.
+parameters, which are much larger than DAVE-2's CNN.
 Using such large CNN models will be challenging on

### 2018-04-20 paper/intro.tex
intro revision

@@
 cars~\cite{Bojarski2016} and other robots. It is expected that more
-DNN based Artificial Intelligence workloads may be used in future
+DNN based artificial intelligence (AI) workloads may be used in future
 autonomous vehicles.
@@
 The DeepPicar, however, employs a state-of-the-art AI
-technology, namely deep learning based end-to-end real-time control,
+technology, namely end-to-end deep learning based real-time control,
 which utilizes a deep convolutional neural network (CNN).
@@
 drove on public roads without human driver's intervention while only
-using the CNN, which runs on a significantly more powerful computing
-platform (Drive PX~\cite{drivepx}).
+using the CNN.
+%% Note that, DAVE-2 run on a significantly more powerful computing
+%% platform (Drive PX~\cite{drivepx}).
 
@@
 Using DeepPicar, we systematically analyze its real-time
-capabilities in the context of deep-learning based real-time
-control, especially on real-time deep neural network \emph{inferencing}.
+capabilities in the context of end-to-end deep-learning based real-time
+control, especially on real-time \emph{inferencing} of the CNN.
 We also evaluated other, more powerful, embedded computing
@@
 From the systematic study, we want to answer the following questions:
-(1) how well does the DNN based real-time inferencing task perform on
-contemporary embedded multicore computing platforms? (2) how
-susceptible is the DNN inferencing task to contention in shared
+(1) How well does the CNN based real-time inferencing task perform on
+contemporary embedded multicore computing platforms? (2) How
+susceptible is the CNN inferencing task to contention in shared
 hardware resources (e.g., cache and DRAM) when multiple tasks/models
-are consolidated? (3) are the existing state-of-the-art shared resource
+are consolidated? (3) Are the existing state-of-the-art shared resource
 isolation techniques effective in protecting real-time performance of
-the DNN inferencing task?
+the CNN inferencing task?
 
@@
 Our observations are as follows.
-First, we find that real-time processing of the DNN inferencing is
+First, we find that real-time processing of the CNN inferencing is
 feasible on contemporary embedded computing platforms, even one   
 as inexpensive as the Raspberry Pi 3. Second, while consolidating
-multiple DNN models and task are feasible but the impact of shared
+multiple CNN models and task are feasible but the impact of shared
 resource contention can be catastrophic in the worst-case as we
-observe up to 11.6X slowdown even when the DNN model was given a
+observe up to 11.6X slowdown even when the CNN model was given a
 dedicated core. Third, using the state-of-the-art OS-level cache
@@
 mechanisms~\cite{yun2014rtas,Yun2013}, we find that cache space has a
-minimal performance impact to the DNN workload while memory
+minimal performance impact to the CNN workload while memory
 performance has a significant impact.
@@
     employs a state-of-the-art CNN based end-to-end
-    real-time control.~\footnote{The source code, datasets, and build
+    real-time control~\footnote{The source code, datasets, and build
       instructions of DeepPicar can be found at:
@@
 
-The remaining sections of the paper are as follows.
-Section~\ref{sec:background}
-provides a background of applications in autonomous driving and related works.
+The remainder of the paper is organized as follows.
+Section~\ref{sec:background} provides a background on the application
+of nueral networks in autonomous driving.
 Section~\ref{sec:overview} gives an overview of the DeepPicar
-platform, including the high-level system and the methods used for
-training and inference. Section~\ref{sec:evaluation} presents our
-evaluation of the platform and how different factors can affect
-performance. Section~\ref{sec:comparison} offers a comparison between  
-the Raspberry Pi 3 and other embedded computing platforms to 
-determine their suitably for autonomous driving research.
+testbed. Section~\ref{sec:evaluation} presents extensive real-time
+performance evaluation results we collected on the testbed.
+Section~\ref{sec:comparison} offers a comparison between  
+the Raspberry Pi 3 and other embedded computing platforms.
 %% We discuss more detailed training methods and system-level issues

### 2018-04-20 paper/evaluation-drampart.tex
..

@@
 sensitivity and the effect of memory bandwidth throttling in providing
-isolation. For the experiments, we use MemGuard\cite{Yun2013}, which 
-can limit the amount of memory bandwidth each core receives. MemGuard
+isolation. For the experiments, we use MemGuard\cite{Yun2013}, a Linux
+kernel module that can limit the amount of memory bandwidth each core
+receives. MemGuard
 operates periodically, at a 1 ms interval, and uses hardware
-performance counters to throttle cores if they exceed their given
-bandwidth budgets within each regulation period (i.e., 1 ms) by
-scheduling high-priority idle kernel threads, until the next period begins.
+performance counters to throttle cores, if they exceed their given
+bandwidth budgets within each regulation period (i.e., 1 ms), by
+scheduling high-priority idle kernel threads until the next period begins.
 
@@
 Figures \ref{fig:memguard_bandwidth} shows the results.
-Note that, in case of BwWrite co-runners, we have assigned half the
-bandwidth budget because MemGuard currently only accounts for L2 refills
-but not write-backs, which effectively allows twice the allocated
-memory bandwidth budget. Regardless, the results clearly show that
-limiting the co-runners's memory bandwidth is effective in protecting the
-CNN model's performance.
+As clearly can be seen in the figure, limiting the co-runners's memory
+bandwidth is effective in protecting the CNN model's performance for
+BwRead and BwWrite co-runners. The benefits are especially more
+pronounced in case of BwWrite co-runners as we throttle them more the
+CNN's performance quickly improves.
+
+%% Note that, in case of BwWrite co-runners, we assign half the
+%% bandwidth budget because MemGuard only accounts for L2 refills
+%% but not write-backs, which effectively allows twice the allocated
+%% memory bandwidth budget.
 

### 2018-04-20 paper/evaluation-drampart.tex
..

@@
   \includegraphics[width=.45\textwidth]{figs/memguard_multicore}
-  \caption{ Memory bandwidth sensitivity of the control loop 
-execution time. }
+  \caption{Memory bandwidth sensitivity of the CNN control loop.}
   \label{fig:memguard_multicore}
@@
   \includegraphics[width=.45\textwidth]{figs/memguard_bandwidth}
-  \caption{Timing impact of co-scheduling memory intensive
-co-runners when memory bandwidth throttling is enabled. }
+  \caption{Effect of throttling three memory intensive co-runners.} 
   \label{fig:memguard_bandwidth}

### 2018-04-20 paper/main.tex
abstract revision

@@
 replication of a real self-driving car called DAVE-2 by NVIDIA. DAVE-2
-drove on public roads using a deep convolutional neural network (CNN), 
+uses a deep convolutional neural network (CNN), 
 which takes images from a front-facing camera as input and produces
@@
 network architecture---9 layers, 27 million connections and 250K
-parameters---and can be trained to drive itself using a
-web camera and a modest Raspberry Pi 3 quad-core platform in real-time.
+parameters---and can drive itself in real-time using a web camera and
+a Raspberry Pi 3 quad-core platform.
 Using DeepPicar, we analyze the Pi 3's computing capabilities to 
@@
 Hz depending on hardware platform. 
-However, shared resource contention remains an
-important issue that must be considered in applying deep-learning
-models on shared memory based embedded computing platforms; we observe
-up to 11.6X control loop execution time increase due to such shared
-resource contention. We evaluate state-of-the-art cache
-partitioning and memory bandwidth throttling techniques on the Pi 3
-in protecting the CNN workload. We find cache partition is
+However, we find that shared resource contention remains an
+important issue that must be considered in applying CNN models on
+shared memory based embedded computing platforms; we observe 
+up to 11.6X execution time increase in the CNN based control loop due
+to shared resource contention. To protect the CNN workload, we also
+evaluate state-of-the-art cache partitioning and memory bandwidth
+throttling techniques on the Pi 3. We find cache partition is
 ineffective, while memory bandwidth throttling is an effective

### 2018-04-20 paper/intro.tex
intro revision

@@
 as well. For example, Tesla Model S was known to use a specialized
-chip (MobileEye EyeQ), which used a deep neural network for vision-based
-real-time obstacle detection and avoidance. More recently, researchers
-are investigating DNN based end-to-end control of
-cars~\cite{Bojarski2016} and other robots. It is expected that more
-DNN based artificial intelligence (AI) workloads may be used in future
-autonomous vehicles.
+chip (MobileEye EyeQ), which used a vision-based real-time obstacle
+detection system based on a DNN. More recently, researchers 
+are investigating DNN based end-to-end real-time control for
+robotics applications~\cite{Bojarski2016,Levine2016}. It is expected
+that more DNN based artificial intelligence (AI) workloads may be
+used in future autonomous vehicles.
 
@@
 Executing these AI workloads on an embedded computing platform 
-poses several additional challenges. First, many AI workloads in vehicles 
-are computationally demanding and have strict real-time requirements. 
-For example, latency in a vision-based object
+poses several additional challenges. First, many AI workloads,
+especially those in vehicles, are computationally demanding and have
+strict real-time requirements.
+For example, computing latency in a vision-based object
 detection task may be directly linked to the safety of the vehicle. This
@@
 DeepPicar is comprised of a Raspberry Pi 3 Model B quad-core
-computer, a web camera and a RC car, all of which are affordable
+computer, a web camera and a small RC car, all of which are affordable
 components (less than \$100 in total).
@@
 control, especially on real-time \emph{inferencing} of the CNN.
-We also evaluated other, more powerful, embedded computing
+We also evaluate other, more powerful, embedded computing
 platforms to better understand achievable real-time performance of
-DeepPicar's deep-learning based control system and the impact of
+DeepPicar's CNN based control system and the performance impact of
 computing hardware architectures.
@@
 %   isolation of DNN workload when memory bandwidth is contended.
-Our observations are as follows.
+Our main observations are as follows.
 First, we find that real-time processing of the CNN inferencing is
@@
 as inexpensive as the Raspberry Pi 3. Second, while consolidating
-multiple CNN models and task are feasible but the impact of shared
-resource contention can be catastrophic in the worst-case as we
-observe up to 11.6X slowdown even when the CNN model was given a
-dedicated core. Third, using the state-of-the-art OS-level cache
-partitioning and memory bandwidth control
-mechanisms~\cite{yun2014rtas,Yun2013}, we find that cache space has a
-minimal performance impact to the CNN workload while memory
-performance has a significant impact.
+additional CNN models and tasks on a single multicore platform is
+feasible but the impact of shared resource contention can be
+catastrophic---we observe up to 11.6X slowdown even when the CNN model
+was given a dedicated core.
+Third, our evaluation of existing cache
+partitioning and memory bandwidth throttling
+techniques~\cite{yun2014rtas,Yun2013} shows that cache partitioning is
+not effective in protecting the CNN inferencing task while memory
+bandwidth throttling is quite effective.
 

### 2018-04-20 paper/evaluation.tex
Minor changes

@@
 timings. In other words, we are interested in how shared resource
-contention affects inference timing. These instances are not identical
-copies of the same model, but are instead copies of different models
-(but of the same network architecture). 
-This is done to ensure that no L2 cache memory is shared between the 
-models as they are running.
+contention affects inference timing. For this, we create four different
+neural network models, that have the same network architecture, and
+run them simultaneously.
+%% These instances are not identical
+%% copies of the same model, but are instead copies of different models
+%% (but of the same network architecture). 
+%% This is done to ensure that no L2 cache memory is shared between the 
+%% models as they are running.
 

### 2018-04-20 paper/related.tex
related work revision

@@
 that the binarized convolution could substitute the expensive
-convolutional feedforward operations in a regular Convolutional Neural
-Network (CNN), %% by using only about 1.5\% of the memory space, while
+convolutional feedforward operations in a regular CNN, by using only
+about 1.5\% of the memory space, while 
 providing 20 to 60 times faster feedforward. Binary weights were also
@@
 
-These research efforts are expected to make complex CNNs more
-accessible for a wider range of real-time embedded systems. We plan to
+These research efforts are expected to make complex CNNs accessible
+for a wider range of real-time embedded systems. We plan to 
 investigate the feasibility of these approaches in the context of
 DeepPicar so that we can use even more resource constrained
-micro-controller class computing platforms.
+micro-controller class computing platforms in place of the current
+Raspberry Pi 3.
 

### 2018-04-20 paper/appendix.tex
appendix revision

@@
-\begin{figure*}[h]
+\begin{figure*}[t]
   \centering
@@
 collected training images, to optimize the network.
-We repeat this across 2,000 training steps. When a model was trained
-with the  aformentioned data, the training loss was 0.0188 and the
-validation  loss was 0.0132.
-The change of the loss value over the course of model training can be
-seen in Figure~\ref{fig:modelloss}.
+We repeat this across 2,000 training steps.
+%% When a model was trained
+%% with the  aformentioned data, the training loss was 0.0188 and the
+%% validation  loss was 0.0132.
+Figure~\ref{fig:modelloss} shows the change of the loss value over the
+course of model training.
 
 In the second method, we use the same data and parameters as  above 
-except that now images are labeled as 'curved' and 'straight' and we pick an
+except that now images are labeled as `curved' and `straight' and we pick an
 equal number of images from each category at each training step to
@@
 center of the track (on the white tape).
-However, we find that there is a huge discrepency between the training
-loss and the validation loss as the former was 0.009, while the latter
-was 0.0869---a 10X difference---indicating that the model suffers from
-an overfitting problem.
+However, we find that there is a discrepency between the training
+loss and the validation loss, indicating that the model may suffer
+from an overfitting problem, despite its better real-world
+performance.
 
-We continue to investigate ways to achieve better prediction accuracy
-in training the network, as well as improving the performance of the RC
-car platform, especially related to precise steering angle control.
+%% as the former was 0.009, while the latter
+%% was 0.0869
+
+We plan to investigate ways to achieve better prediction accuracy
+in training the network. We also plan to evaluate different RC
+car platforms that have different steering mechanisms to see their
+impact to the control performance.
 
@@
 few system-level factors, namely power supply and temperature, that
-need to be considered to achieve consistent and high real-time
-performance.
+need to be considered to achieve consistent performance.
 
-In all our experiments on the Raspberry Pi 3, the CPU  operated at a
-preferred clock speed of 1.2 GHz. However, without care, it is
-possible for the CPU to operate at a lower frequency.
-
+In all our experiments on the Raspberry Pi 3, the CPU  is configured
+at the maximum clock speed of 1.2 GHz. However, without care, it is
+possible for the CPU to operate at a lower frequency involuntarily. 
 An important factor is CPU thermal throttling, which can affect CPU
 clock speed if the CPU temperature is too high (Pi 3's firmware is
-configured to throttle at 85C).
-DNN model operations are computationally intensive, thus it is
-possible for the temperature of the CPU to rise quickly. This can be
-especially problematic in situations where multiple DNN models are
-running simultaneously on the Pi 3.
-If the temperature reaches the threshold, the Pi 3's thermal
+configured to throttle at 85 deg. C).
+DNN inferencing is computationally intensive, thus the temperature of
+the CPU could rise quickly. This can be especially problematic in
+situations where multiple DNN models run simultaneously on the
+Pi 3. If the temperature reaches the threshold, the Pi 3's thermal
 throttling kicks in and decreases the clock speed down to 600MHz---

### 2018-04-20 paper/appendix.tex
removed the loss figure.

@@
-\begin{figure*}[t]
-  \centering
-  \includegraphics[width=.9\textwidth]{figs/TrainingLoss}
-  \caption{Change in loss value throughout training.}
-  \label{fig:modelloss}
-\end{figure*}
+%% \begin{figure*}[t]
+%%   \centering
+%%   \includegraphics[width=.9\textwidth]{figs/TrainingLoss}
+%%   \caption{Change in loss value throughout training.}
+%%   \label{fig:modelloss}
+%% \end{figure*}
 
@@
 frames for validation.
-The weights of the network are initialized using a Xavier
-initializer~\cite{Glorot2010}, which is known to provide better
-initial values than the random weight assignment method.
+The weights of the network are initialized using the Xavier
+initializer~\cite{Glorot2010}, which is known to perform better
+than a random weight assignment.
 In each training step, we use a batch
@@
 %% validation  loss was 0.0132.
-Figure~\ref{fig:modelloss} shows the change of the loss value over the
-course of model training.
+%% Figure~\ref{fig:modelloss} shows the change of the loss value over the
+%% course of model training.
 
@@
 
-We plan to investigate ways to achieve better prediction accuracy
-in training the network. We also plan to evaluate different RC
-car platforms that have different steering mechanisms to see their
-impact to the control performance.
+%% We plan to investigate ways to achieve better prediction accuracy
+%% in training the network. We also plan to evaluate different RC
+%% car platforms that have different steering mechanisms to see their
+%% impact to the control performance.
 

### 2018-04-20 paper/main.tex
Changes to abstract

@@
 We also systematically compare other contemporary embedded
-computing platforms using the DeepPicar's CNN based real-time control
+computing platforms using the DeepPicar's CNN-based real-time control
 workload.
 
-We find all tested platforms, including the Pi 3, are capable of
-supporting the CNN based real-time control, from 20 Hz up to 100
+We find that all tested platforms, including the Pi 3, are capable of
+supporting the CNN-based real-time control, from 20 Hz up to 100
 Hz depending on hardware platform. 
@@
 evaluate state-of-the-art cache partitioning and memory bandwidth
-throttling techniques on the Pi 3. We find cache partition is
+throttling techniques on the Pi 3. We find that cache partitioning is
 ineffective, while memory bandwidth throttling is an effective

### 2018-04-20 paper/background.tex
Changes to sections I through III

@@
 phases~\cite{NVIDIA2015}. The first
-phase is \emph{training} during which the weights of the network are
+phase is \emph{training}, during which the weights of the network are
 incrementally updated by backpropagating errors it sees from the
@@
 efforts include the development of various heterogeneous architecture-based 
-system-on-a-chips (SoCs) that may include multiple cores, GPU,
+system-on-a-chip (SoCs) that may include multiple cores, GPU,
 DSP, FPGA, and neural network optimized ASIC hardware~\cite{Jouppi2017}.

### 2018-04-20 paper/intro.tex
Changes to sections I through III

@@
 highly efficient computing platform. These two conflicting
-requirements complicate the platform selection process as observed in
-~\cite{Otterness2017}.
+requirements complicate the platform selection process as observed 
+in~\cite{Otterness2017}.
 
@@
 additional CNN models and tasks on a single multicore platform is
-feasible but the impact of shared resource contention can be
+feasible, the impact of shared resource contention can be
 catastrophic---we observe up to 11.6X slowdown even when the CNN model
@@
 Section~\ref{sec:background} provides a background on the application
-of nueral networks in autonomous driving.
+of neural networks in autonomous driving.
 Section~\ref{sec:overview} gives an overview of the DeepPicar

### 2018-04-20 paper/overview.tex
intro revision
Minor changes
Changes to sections I through III

@@
-\section{DeepPicar Overview}\label{sec:overview}
+\section{DeepPicar}\label{sec:overview}
 
@@
   \centering
-  \includegraphics[width=.38\textwidth]{figs/DeepPicar_platform}
+  \includegraphics[width=.4\textwidth]{figs/DeepPicar_platform}
   \caption{DeepPicar platform.}
@@
 pseudo code of the controller's main loop. Among the five steps, the 3rd step, 
-network inferencing, is the most computationally intensive and dominates the
-execution time.
+network inferencing, is the most computationally intensive and is expected to 
+dominate the execution time.
 
@@
 
-Another factors that can potentially affect the prediction accuracy of
+Other factors that can potentially affect the prediction accuracy of
 the CNN, are camera and actuator (motor) control latencies. The camera
-latency is from the time the camera sensor observes the scene to the
+latency is defined as the time the camera sensor observes the scene to the
 time the computer actually reads the digitized image data. This time
@@
 For instance, the DeepPicar could remain on Track 1
-(Figure~\ref{fig:track}) over 10 minutes at a moderate speed (50\%
+(Figure~\ref{fig:track}) for over 10 minutes at a moderate speed (50\%
 throttle), at which point we stopped the experiment, and more than one

### 2018-04-20 paper/evaluation.tex
Minor changes to evaluation

@@
 and all cores share a 512KB L2 cache.
-The chip also includes Broadcom's Videocore IV GPU, although we did
+The chip also includes Broadcom's Videocore IV GPU, although we do
 not use the GPU in our evaluation, due to the lack of sofware support
@@
 For software, we use Ubuntu MATE 16.04 and TensorFlow 1.1.
-We disabled DVFS (dynamic voltage frequency scaling) and
-configured the clock speed of each core statically at the maximum 1.2GHz.
+We disable DVFS (dynamic voltage frequency scaling) and
+configure the clock speed of each core statically at the maximum 1.2GHz.
 We use the SCHED\_FIFO real-time scheduler to schedule the CNN control

### 2018-04-21 paper/background.tex
Paper changes

@@
 necessary computing performance to realize deep neural network based
-robotics systems, (2) to understand the characteristics of the
+robotic systems, (2) to understand the characteristics of the
 computing platform to support such workloads, and (3) to evaluate the

### 2018-04-21 paper/comparison.tex
Paper changes

@@
 end-to-end deep learning based autonomous vehicles. 
-Table~\ref{tbl:platforms} shows architectural features of the three
+Table~\ref{tbl:platforms} shows the architectural features of the three
 platforms~\footnote{The GPU of Intel UP and the two Denver cores in the

### 2018-04-21 paper/evaluation-cachepart.tex
Minor changes to evaluation
Paper changes

@@
 page-coloring based kernel-level memory allocator for Linux.
-Page coloring is a OS technique that controls the physical addresses
+Page coloring is an OS technique that controls the physical addresses
 of memory pages. By allocating pages over non-overlapping cache sets,
@@
 to avoid partitioning the private L1 caches, we use bits 13 and 14 for
-coloring, which results in 4 usuable page colors.
+coloring, which results in 4 usable page colors.
 
@@
   \caption{Average processing time vs. the number of memory
-intensive co-runners; Each core (task) is given a equal-sized
+intensive co-runners; Each core (task) is given an equal-sized
 dedicated cache partition.}

### 2018-04-21 paper/evaluation.tex
Paper changes

@@
     CNN inferencing      & 18.49 & 19.30 & 18.99  & 0.20 \\ \hline
-    Total Time           & 22.87 & 23.74 & 23.38  & 0.20 \\ \hline
+    Total Time           & 22.86 & 23.74 & 23.38  & 0.20 \\ \hline
   \end{tabular}
@@
 Second, we find that the measured average
-execution time of a single control loop is 22.87 ms, or 43.7 Hz and
+execution time of a single control loop is 22.86 ms, or 43.7 Hz and
 the 99 percentile time is 23.38 ms.
@@
 linear scaling. In particular, from 3 cores to 4 cores, the
-improvement is mere 2.80 ms, or 12\%. In short, we find that the
+improvement is a mere 2.80 ms, or 12\%. In short, we find that the
 scalability of DeepPicar's deep neural network is not ideal.
@@
 For example, it is conceivable to use four cameras and four different
-neural network models, each of which is trained separately for
+neural network models, each of which is trained separately for a
 different purpose and executed on a single dedicated core.
@@
 architecture for all models, then one might expect to achieve up to
-20 Hz using one core (given 1 core can deliver 46 ms average
+20 Hz using one core (given that 1 core can deliver 46 ms average
 execution time).
@@
 cause relatively much smaller time increases. While execution time
-increases are expected, the degree that is seen in the worst-case 
+increases are expected, the degree to which that is seen in the worst-case 
 is quite surprising---\emph{if the CNN controller was driving an actual 

### 2018-04-21 paper/overview.tex
Paper changes

@@
 operations must be completed within each control period
-duration---e.g., a WCET of 33.3 ms for 30Hz control 
+duration---e.g., a WCET of 33.3 ms for 30 Hz control 
 frequency---locally on the Pi 3 platform, although training of the 
@@
 control, the observed control performance of the system is respectable,
-likely due to the inherently stocastic nature of CNNs.
+likely due to the inherently stochastic nature of CNNs.
 
@@
 negatively affect control performance, because the CNN would analyze
-stale scenes. The actuator (motor) control latency from the time
+stale scenes. The actuator (motor) control latency is defined as the time
 the control output is sent to the steering motor to the time the motor
-actually moved at a desired position, which also can takes
+actually moves to a desired position, which also can take
 considerable time. In our platform, the combined latency is measured
-to be around 50 ms, which is reasonble.
+to be around 50 ms, which is reasonable.
 %% In other words, %% in the recorded video and control data, 
@@
 If this value is too high, control performance may suffer.
-Our initial prototype suffered this problem as the observed latency
+Our initial prototype suffered from this problem as the observed latency
 was as high as 300 ms, which negatively affected control performance.

### 2018-04-26 paper/evaluation-drampart.tex
Minor changes to evaluation

@@
 Figures \ref{fig:memguard_bandwidth} shows the results.
-As clearly can be seen in the figure, limiting the co-runners's memory
+As can clearly be seen in the figure, limiting the co-runners's memory
 bandwidth is effective in protecting the CNN model's performance for
 BwRead and BwWrite co-runners. The benefits are especially more
-pronounced in case of BwWrite co-runners as we throttle them more the
+pronounced in case of BwWrite co-runners as, when we throttle them more, the
 CNN's performance quickly improves.

### 2018-04-26 paper/main.tex

@@
 supporting the CNN-based real-time control, from 20 Hz up to 100
-Hz depending on hardware platform. 
+Hz, depending on hardware platform. 
 However, we find that shared resource contention remains an

### 2018-06-15 paper/evaluation-drampart.tex
Minor changes

@@
 operates periodically, at a 1 ms interval, and uses hardware
-performance counters to throttle cores, if they exceed their given
+performance counters to throttle cores if they exceed their given
 bandwidth budgets within each regulation period (i.e., 1 ms), by

### 2018-06-15 paper/evaluation.tex
Minor changes

@@
 cause relatively much smaller time increases. While execution time
-increases are expected, the degree to which that is seen in the worst-case 
+increases are expected, the degree to which it is seen in the worst-case 
 is quite surprising---\emph{if the CNN controller was driving an actual

### 2018-06-15 paper/related.tex
Minor changes

@@
 CNN, which was used in NVIDIA's DAVE-2 self-driving
-car~\cite{Bojarski2016}, on a low-cost raspberry pi 3 quad-core 
+car~\cite{Bojarski2016}, on a low-cost Raspberry Pi 3 quad-core 
 platform and other embedded multicore platforms. It should be noted,
